### 2014-06-10 foreground_delay_spectrum_modeling.tex
First version of the draft.

%%\documentclass[12pt,preprint]{aastex}

%% manuscript produces a one-column, double-spaced document:

%% \documentclass[10pt,manuscript]{aastex}

%% preprint2 produces a double-column, single-spaced document:
\documentclass[preprint2,iop,numberedappendix]{emulateapj}
%% \documentclass[preprint2,iop]{aastex}

%% \documentclass[preprint2,longabstract]{aastex}

%% \usepackage{ccaption}
%% \captionstyle{\raggedright}
\usepackage[caption=false]{subfig}
\usepackage{amsmath}
\usepackage{footnote}
\bibpunct{(}{)}{;}{a}{}{,} 
\captionsetup{belowskip=12pt,aboveskip=4pt}
\setlength{\textfloatsep}{10pt plus 1.0pt minus 2.0pt}
\newcommand{\dif}{\mathrm{d}}
%% \renewcommand*{\thefootnote}{\fnsymbol{footnote}}

\def\nar{{New~A~Rev.}}          % New Astronomy Review
\def\pasa{{PASA}}               % Publications of the Astron. Soc. of Australia

%% \bibliographystyle{mn2e}
%% \bibliographystyle{apj}

\shorttitle{Foreground signatures in EoR power spectra}
\shortauthors{Thyagarajan et~al.}

\def\ASU{\altaffilmark{1}}
\def\ASUtxt{\altaffiltext{1}{Arizona State University, Tempe, AZ 85287, USA; e-mail: t\_nithyanandan@asu.edu}}

% \def\RRI{\altaffilmark{1}}
% %% \def\RRItxt{\altaffiltext{1}{Raman Research Institute, Bangalore, India}}
% \def\RRItxt{\altaffiltext{1}{Raman Research Institute, Bangalore, India; e-mail: nithya.rri@gmail.com}}

% \def\NRAONM{\altaffilmark{2}}
% \def\NRAONMtxt{\altaffiltext{2}{National Radio Astronomy Observatory, Socorro, NM, USA}}

% \def\CAASTRO{\altaffilmark{3}}
% \def\CAASTROtxt{\altaffiltext{3}{ARC Centre of Excellence for All-sky Astrophysics (CAASTRO), Sydney, Australia}}

% \def\Curtin{\altaffilmark{4}}
% \def\Curtintxt{\altaffiltext{4}{Curtin University, Perth, Australia}}

% %% \def\Swinburne{\altaffilmark{4}}
% %% \def\Swinburnetxt{\altaffiltext{4}{Swinburne University of Technology, Melbourne, Australia}}

% \def\CfA{\altaffilmark{5}}
% \def\CfAtxt{\altaffiltext{5}{Harvard-Smithsonian Center for Astrophysics, Cambridge, USA}}

% \def\ANU{\altaffilmark{7}}
% \def\ANUtxt{\altaffiltext{7}{The Australian National University, Canberra, Australia}}

% \def\CSIRO{\altaffilmark{8}}
% \def\CSIROtxt{\altaffiltext{8}{CSIRO Astronomy and Space Science, NSW, Australia}}

% \def\Haystack{\altaffilmark{9}}
% \def\Haystacktxt{\altaffiltext{9}{MIT Haystack Observatory, Westford, USA}}

% \def\USydney{\altaffilmark{10}}
% \def\USydneytxt{\altaffiltext{10}{Sydney Institute for Astronomy, School of Physics, The University of Sydney, Sydney, Australia}}

% \def\MIT{\altaffilmark{11}}
% \def\MITtxt{\altaffiltext{11}{MIT Kavli Institute for Astrophysics and Space Research, Cambridge, USA}}

% \def\UW{\altaffilmark{12}}
% \def\UWtxt{\altaffiltext{12}{University of Washington, Seattle, USA}}

% \def\Victoria{\altaffilmark{13}}
% \def\Victoriatxt{\altaffiltext{13}{Victoria University of Wellington, Wellington, New Zealand}}

% \def\UWisc{\altaffilmark{14}}
% \def\UWisctxt{\altaffiltext{14}{University of Wisconsin--Milwaukee, Milwaukee, USA}}

% \def\UMelbourne{\altaffilmark{15}}
% \def\UMelbournetxt{\altaffiltext{15}{The University of Melbourne, Melbourne, Australia}}

% \def\Tata{\altaffilmark{16}}
% \def\Tatatxt{\altaffiltext{16}{Tata Institute of Fundamental Research, Pune, India}}

% \def\NRAOGBT{\altaffilmark{17}}
% \def\NRAOGBTtxt{\altaffiltext{17}{National Radio Astronomy Observatory, Charlottesville, USA}}

% \def\UTasmania{\altaffilmark{18}}
% \def\UTasmaniatxt{\altaffiltext{18}{University of Tasmania, Hobart, Australia}}

% \def\UWA{\altaffilmark{19}}
% \def\UWAtxt{\altaffiltext{19}{University of Western Australia, Perth, Australia}}

% %% \def\ICRAR{\altaffilmark{9}}
% %% \def\ICRARtxt{\altaffiltext{9}{International Centre for Radio Astronomy Research, Curtin University, Perth, Australia }}

% %\def\VUW{\altaffilmark{13}}
% %\def\VUWtxt{\altaffiltext{13}{Victoria University of Wellington, New Zealand}}

%% \definenote[thanks][conversion=set 2]

\begin{document}

%% \title{Detecting Epoch of Reionization in redshifted 21~cm: contamination in EoR window}
%% \title{Limits on the detection of the Epoch of Reionization from MWA observations of the redshifted 21~cm line}
\title{Signatures of Foreground Sky in Power Spectra of Redshifted Neutral Hydrogen from the Epoch of Reionization}

%% Use \author, \affil, and the \and command to format
%% author and affiliation information.
%% Note that \email has replaced the old \authoremail command
%% from AASTeX v4.0. You can use \email to mark an email address
%% anywhere in the paper, not just in the front matter.
%% As in the title, use \\ to force line breaks.

%% \author{NITHYANANDAN~THYAGARAJAN\altaffilmark{1}, 
%%   N.~UDAYA~SHANKAR\altaffilmark{1}, 
%%   \and RAVI~SUBRAHMANYAN\altaffilmark{1}}

%% Author list
\author{
%% Lead authors
Nithyanandan~Thyagarajan\ASU,
Daniel~C.~Jacobs\ASU,
Judd~D.~Bowman\ASU
}

%Institutional footnotes (typeset, then rearrange here to be in order)
\ASUtxt
% \RRItxt
% \NRAONMtxt
% \CAASTROtxt
% \Curtintxt
% %% \Swinburnetxt
% \CfAtxt
% \ANUtxt
% \CSIROtxt
% \Haystacktxt
% \USydneytxt
% \MITtxt
% \UWtxt
% \Victoriatxt
% \UWisctxt
% \Tatatxt
% \NRAOGBTtxt
% \UMelbournetxt
% \UTasmaniatxt
% %% \UWAtxt
% %% \ICRARtxt
% %\VUWtxt

%% \email{t\_nithyanandan@rri.res.in}

%% \clearpage

\begin{abstract}

We characterize signatures of foreground radio emission and instrument configuration on the observed power spectra of redshifted H{\sc i} 21~cm line emission from the epoch of reionization (EoR) using a baseline (antenna pair) based ``delay spectrum'' analysis technique. We use the 128-tile Murchison Widefield Array (MWA) instrument configuration for our studies. Our analysis is centered around 185~MHz with 30.72~MHz bandwidth at 40~kHz frequency resolution. We simulate the delay spectrum response of interferometers with antenna pair spacings up to $\sim 315$~m, for the first time, to an all-sky foreground model consisting of diffuse (on scales $\gtrsim\,$0\fdg 85) and compact Galactic and extragalactic emission. By matching simulated parameters to those of a subset of first season of observing with the MWA, we identify and verify key foreground and instrumental signatures seen in data. A wedge-shaped region of foreground contamination in the delay spectra is confirmed. There is a spillover from this contamination into relatively foreground--free regions ({\it EoR window}) in the delay spectrum caused due to the spectral properties of foreground emission and the instrument. Due to projection effects, diffuse emission produces edge-heavy features. Features in the inner regions of the wedge are predominantly due to compact foreground objects. This results in a ``pitchfork'' shaped delay spectrum. We provide a practical diagnostic tool for scheduling EoR observations that helps in reducing foreground contamination.

\end{abstract}

\keywords{large-scale structure of Universe --- methods: statistical --- radio continuum: galaxies --- radio lines: general --- reionization --- techniques: interferometric}

\section{Introduction}\label{intro}

After the CMB epoch, the period in the history of the universe often referred to as the {\it dark ages} was followed by the epoch of reionization (EoR). This was a period of non-linear growth of density perturbations and astrophysical evolution. Studying the epoch of reionization holds the key to understanding this evolution. Recently, oberving the redshifted 21~cm spin transition of neutral hydrogen has emerged as one of the most promising experiments to fill the gaps in our understanding of the universe's history.

Sensitive instruments such as the Square Kilometer Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}. Numerous precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum.

One of the key challenges in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which lack such symmetry. Foreground removal techniques rely on the spectral smoothness of foreground emission \citep{mor06,bow09,liu11,par12,dil13}.

Precise understanding of these foreground signatures is necessary in order to remove foreground systematics from the measured power spectrum. Considerable effort has been made in mapping the residual foreground signatures in observed power spectrum \citep{thy13,pob13,mor12,tro12,dat10,bow09} after foreground subtraction.

In this paper, in one of the most detailed characterizations to date, using an interferometer (antenna pair) based delay spectrum technique, we demonstrate the signatures of unsubtracted diffuse and compact foreground models in delay spectrum, a quantity closely related to the sought power spectrum. We explore the diversity in these signatures as a function of observation and instrument parameters and confirm our findings with data obtained from the MWA.

We have organized the rest of the paper as follows. \S\ref{sec:delay-spectrum} defines the delay spectrum and its relation to the power spectrum. \S\ref{sec:instrument} and \S\ref{sec:obsparms} describe the MWA instrumental parameters and observation parameters used in this study respectively. In \S\ref{sec:data-analysis}, we describe our data analysis. In \S\ref{sec:modeling}, we describe the building blocks of our delay spectrum modeling including the foreground models used. Here, we also establish that the results of modeling and data analysis match with each other. In \S\ref{sec:delay-spectrum-analysis}, we present a detailed analysis of the signatures of foregrounds on the modeled delay spectrum and its dependence on instrumental and observation parameters. In \S\ref{sec:fg-grading}, we use our understanding to grade the delay spectrum as a function of local sidereal time (LST) in order to obtain a useful diagnostic for planning future observations. In \S\ref{sec:summary}, we present our summary and conclusions.

\section{Delay Spectrum}\label{sec:delay-spectrum}

Interferometer array data known as {\it visibilities}, $V_f(\overline{\mathbf{u}},f)$, are correlations between electric fields measured at different frequencies ($f$) by different antenna pairs with separation vectors $\overline{\mathbf{x}}$, where $\overline{\mathbf{u}}\equiv \overline{\mathbf{x}}f/c$ and $c$ is the speed of light. $\overline{\mathbf{u}}$ is related to transverse spatial frequency modes ($\overline{\mathbf{k}}_\perp$) of brightness distribution of the sky as $\overline{\mathbf{k}}_\perp \equiv 2\pi\overline{\mathbf{u}}/D(z)$, where $D(z)$ is the transverse comoving distance at a redshift $z$. If $I(\overline{\mathbf{l}},f)$ is the emission at different frequencies on the sky as a function of direction unit vector ($\overline{\mathbf{l}}$) specified as direction cosines, $V_f(\overline{\mathbf{u}},f)$ represents the Fourier decomposition of $I(\overline{\mathbf{l}},f)$ attenuated by the antenna's angular power pattern $A(\overline{\mathbf{l}},f)$, into transverse spatial frequency components. Taking into account the instrumental bandpass weights $W_f(f)$, this relation can be written as:
\begin{align}\label{eqn:obsvis}
  V_f(\overline{\mathbf{u}},f) &= \int A(\overline{\mathbf{l}},f)\,I(\overline{\mathbf{l}},f)\,W_f(f)\,e^{-i2\pi \overline{\mathbf{u}}\cdot\overline{\mathbf{l}}}\,\dif^2 \overline{\mathbf{l}}.
\end{align}

We define the {\it delay spectrum} $V_\eta(\overline{\mathbf{u}},\eta)$ to be the inverse Fourier transform of $V_f(\overline{\mathbf{u}},f)$ along the frequency coordinate:
\begin{align}\label{eqn:delay-transform}
  V_\eta(\overline{\mathbf{u}},\eta) &= \int V_f(\overline{\mathbf{u}},f)\,W'_f(f)\,e^{i2\pi\eta f}\,\dif f,
\end{align}
where, $W'_f(f)$ is a frequency window weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}. $\eta$ represents the signal delay between the antenna pairs in an interferometer given by:
\begin{equation}\label{eqn:delay}
  \eta = \frac{\overline{\mathbf{x}}\cdot\overline{\mathbf{l}}}{c} = \frac{\overline{\mathbf{u}}\cdot\overline{\mathbf{l}}}{f}.
\end{equation}
We use the terms delay and lag interchangeably in the rest of this paper.

Since we are studying a redshifted H{\sc i} spectral line, $f$ is a measure of cosmological distance along the line of sight. Consequently, $\eta$ is a measure of line of sight spatial frequency, $k_\parallel$. Thus, $|V_\eta(\overline{\mathbf{u}},\eta)|^2$ is directly related to the spatial power spectrum of redshifted H{\sc i} distribution, $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$. However, foreground objects emit in radio frequencies and contaminate the signals from the redshifted H{\sc i}. Due to their spatial and spectral properties, $V_\eta(\overline{\mathbf{u}},\eta)$ also contains contamination from foreground radio sources \citep{thy13,tro12,mor12,bow09}. These contaminations have to be characterized precisely in order to reduce their impact on redshifted H{\sc i} power spectrum detection sensitivity. We, heareafter, examine in detail the delay spectrum of emission from unsubtracted foreground objects and compare with preliminary observations with the MWA.

\section{Instrument Parameters}\label{sec:instrument}

We use the 128-tile layout of the MWA \citep{bea12}. We choose a center frequency of 185~MHz with a bandwidth of 30.72~MHz divided into 768 frequency channels each of width 40~kHz. The choice of center frequency translates to $z\approx 6.68$ for the 21~cm spin flip transition of H{\sc i}. The bandwidth was chosen to be close to the maximum instantaneous value of the MWA in order to maximize the resolution in delay coordinates and obtain detailed characteristics of delay spectra from foregrounds. We restrict our analysis to the innermost 2000 pairs covering $|\overline{\mathbf{x}}| \leq x_\textrm{max} \approx 315$~m, or $|\overline{\mathbf{u}}| \leq u_\textrm{max} \approx 157.5$ at a frequency of 185~MHz. These interferometers are responsive to transverse spatial structures on scales $\gtrsim 1/u_\textrm{max} \approx 22$\arcmin.

The bandpass weights in the frequency window of the MWA are obtained through a 8-tap Polyphase Filter Bank (PFB) using a Kaiser window with the $\beta$ parameter set to 5. The bandpass weights consist of coarse frequency channels of width 1.28~MHz each of which contains 32 finer frequency channels of width 40~kHz. Our analysis, therefore, contains 24 coarse channels. As the data are subsequently bandpass calibrated, the bandpass weights are normalized to unity. But in the edges of the coarse channels where the bandpass correction factors are $\gtrsim 3$, this correction makes the edge channels signficantly noisy compared to the rest of the spectrum. Thus, we remove 2 fine channels from each side of the coarse channel edges from our data. This constitutes $W_f(f)$ in equation~\ref{eqn:obsvis}. The bandpass properties used in this paper are illustrated in figure~\ref{fig:bandpass}.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/bandpass_properties}
\caption{Bandpass properties used in modeling and analysis of delay spectrum of MWA visibilities. The filled circles joined by dotted lines show the 8-tap PFB bandpass shapes obtained using a Kaiser window with parameter $\beta=5$. The {\it plus} symbols joined by dashed lines are the bandpass correction factors. The dot-dashed line represents the threshold for bandpass correction factor above which frequency channels are flagged. The solid line shows the flagged and gain--corrected bandpass shape, $W_f(f)$. These properties hold for the entire bandwidth of 30.72~MHz centered around 185~MHz. \label{fig:bandpass}}
\end{figure}

Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=440$~K \citep{bow06} for all frequency channels for all interferometers throughout the course of observation. We take into account the increase in thermal uncertainty in frequency channels that results due to aforementioned bandpass correction.

The power pattern of a MWA tile, $A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}, over the entire hemisphere is modeled based on a 4$\times$4 phased array of half-wave dipoles. Hereafter, we use the terms tile and antenna interchangeably to refer to the MWA tile.

\section{Observation Parameters}\label{sec:obsparms}

One of the primary targets for EoR observations with the MWA is a patch of sky centered around RA = 00h 00m 00.00s, Dec = -30\arcdeg$\,$00\arcmin$\,$00\farcs 0. This is one of the regions of low foreground emission within 27\arcdeg$\,$ from the zenith. The MWA tracks a patch of sky through beams formed and steered electronically by changing delay settings between a phased array of dipoles in a MWA tile. These beamformer delay settings can be changed only in discrete steps. Accordingly, observations were made in a ``drift and shift'' (D-S) mode. Figure~\ref{fig:pointings} shows the direction of pointing of MWA tiles during the course of the observation. The sky is allowed to drift for a certain period of time (usually $\sim 30$~mins) before the discrete shifts in the beamformer delay settings position the beam to be centered again on the patch of sky. This process is repeated throughout the course of the observation $\approx 4.86$~hours.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/dns_pointings}
\caption{MWA tile pointing directions during the course of the observation. The x-axis refers to the Local Sidereal Time (LST) in hours. The axis on the left refers to longitudes, namely, Right Ascension (RA) and Hour Angle (HA) in degrees. Negative values of RA, HA, and LST are to be interpreted as having been wrapped around by 360\arcdeg or 24 hours. The axis on the right refers to the declination of the pointing direction. The RA, HA, and declination are plotted with solid, dashed and dotted line styles respectively. A ``drift and shift'' scheme is used. The sky is allowed to drift for $\sim 30$~mins before the beamformer delay settings change in a discrete step to center the beam around RA = 00h 00m 00.00s, Dec = -30\arcdeg$\,$00\arcmin$\,$00\farcs 0. \label{fig:pointings}}
\end{figure}

\section{Data Analysis}\label{sec:data-analysis}

{\bf Describe the data analysis here.}

Throughout this paper, $\overline{\mathbf{u}}$ is assumed to be on a coordinate system aligned with the local east and local north directions at the MWA site. For the conventions of Fourier transform and its inverse we use in this paper, in accordance with equation~\ref{eqn:delay}, signals from the sky towards east and north are recorded with positive delays on eastward and northward baselines respectively. For instance, at the beginning of the observation, the Galactic center is on the far west just about to set. Signals from this direction are recorded with a negative delay on interferometers oriented towards the east. Similarly, signals from directions eastward of the local meridian arrive with positive delays on eastward interferometers. In subsequent discussions, unless mentioned otherwise, all the snapshot visibilities are assumed to be phased to zenith.

For purposes of geometrical intuition, we restrict the orientation ($\theta_b$ measured anti-clockwise from East) of all baselines to lie in the range $-67\fdg 5 \leq \theta_b < 112\fdg 5$. Baselines in the other half-plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.

\subsection{Deconvolution along Delay Axis}\label{sec:CLEAN}

Subsequently, we obtain the delay spectrum of these visibilities using equation~\ref{eqn:delay-transform} while choosing $W'_f(f)$ to be a {\it Blackman--Nuttall} window function for its superior sidelobe suppression as described in \citet{thy13}. Due to periodic gaps in frequency bandpass occurring in intervals of 1.28~MHz (see figure~\ref{fig:bandpass}), the delay spectrum is expected to contain harmonics of actual foreground emission repeated at intervals of 0.78~$\mu$s. Hence, we have employed deconvolution by {\it CLEAN} algorithm \citep{tay99} along the delay axis \citep{par09,par12} to rid the delay spectra of such artifacts. The convolving kernel for the algorithm is given by the inverse Fourier transform of the instrumental bandpass shape.

\subsection{Delay Spectrum from Data}\label{sec:data-delay-spectrum}

Put up plots here and discuss them.

Despite the deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the data perfectly. This is predominant at short baselines. This is because the maximum delay envelope (boundary of the wedge) consists of mixed emission from large portions of the sky into a narrow range of delays and the deconvolution algorithm does not have sufficient support in delay space to fit the delay spectrum. However, we note that the purpose of deconvolution is to rid the delay spectra of instrumental artifacts to the best extent possible in order to see the underlying signatures of foreground emission. It is not our intention in this paper to use the deconvolution as a foreground removal tool. Hence, our discussion and conclusions still hold despite imperfections in deconvolution.

\section{Delay Spectrum Modeling}\label{sec:modeling}

Below we describe the foreground model we have used in modeling the measured delay spectra.

\subsection{Foreground Model}\label{sec:foreground}

An instrument such as the MWA has a very wide field of view ($\sim 30$\arcdeg at 185~MHz) for imaging purposes. In the context of EoR H{\sc i} power spectrum it has been shown that unsubtracted forground sources anywhere in the horizon directly contaminate the spatial frequency modes in the power spectrum. In addition, this contamination also spills over into the relatively cleaner regions, called {\it EoR window}, due to spectral properties of the instrument and the foregrounds \citep{thy13,pob13,ved12,par12}. Thus, it is important to consider an all-sky model for foreground objects in evaluating the features seen in the power spectrum.

\citet{bea13} estimated the sensitivity of the MWA to EoR H{\sc i} power spectrum detection taking into account theremal noise effects in the {\it Eor window}. \citet{thy13} estimated the sensitivity taking into account the spillover from foreground contamination from unsubtracted extragalactic point sources in the {\t EoR window} besides thermal noise. In this paper, we include the diffuse Galactic emission for a comprehensive understanding of foreground signatures in the power spectrum. We use an all-sky foreground emission model that consists of diffuse and bright compact components.

\subsubsection{Diffuse Foreground Model}\label{sec:DSM}

For the diffuse component, we use an all-sky radio foreground model from \citet{deo08} to estimate the emission at 185~MHz. Since this map is predominantly based on the 408~MHz all-sky map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz diffuse emission map to the same resolution. However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals. Using the same source model, we also obtain the diffuse emission maps at 140~MHz and 160~MHz to estimate the spectral index map of the diffuse emission model.

\subsubsection{Compact Foreground Model}\label{sec:CSM}

We use a combination of NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz due to their similar flux sensitivity and angular resolution, and complimentary survey footprints covering the entire sky. The SUMSS catalog covers the sky with declination $\delta < -30$\arcdeg with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec. The NVSS covers the sky with $\delta > -40$\arcdeg with a similar angular resolution and a limiting flux density of $\approx 2.5$~mJy for discrete sources.

From the SUMSS catalog, we select compact sources whose deconvolved major axes are equal to 0. From the NVSS catalog, we excluded objects that overlap with those in the SUMSS survey footprint. Compact sources from NVSS were selected if the convolved major axes were not greater than $\approx 47$\arcsec, which is almost equal to the angular resolution of the survey. Using a mean spectral index of $\langle\alpha\rangle=-0.83$ (flux density, $S(f)\propto f^\alpha$) obtained by \citet{mau03} for both NVSS and SUMSS catalog objcts, we calculate the corresponding flux densities at 185~MHz, $S_{185}$. From this subset, we choose compact objects with $S_{185}\geq 10$~Jy. We verified that our selection criteria ensure a similar areal density of sources in the two surveys.

Based on these criteria, we selected 133 sources from the SUMSS catalog and 336 sources from the NVSS catalog. The selection of such bright compact objects is not affected by minor differences in sensitivity of the two surveys. Together with the diffuse foreground model, we obtain an all-sky foreground model consisting of both compact and diffuse emission.

\subsection{Comparison with Data}\label{sec:data-vs-model}

With the aforementioned all-sky foreground model, and instrumental and observational parameters, we simulate visibilities using equation~\ref{eqn:obsvis}. We consider two snapshots, one at the beginning and one during the middle of the observation, to explain the signatures of foregrounds seen in the modeled and observed data. The first snapshot is pointed at RA = 4\fdg 387, Dec = -29\fdg 86 at 22.08 hours LST while the second is pointed at RA = +1\fdg 320, Dec = -26\fdg 70 at 0.09 hours LST. In both cases the visibilities are phased to zenith corresponding to RA = -28\fdg 8, Dec = -26\fdg 701, and RA = 1\fdg 35, Dec = -26\fdg 701 respectively. Signals from the phase center are received simulatneously with no delay between the interferometers.

Figure~\ref{fig:obsids} shows the amplitude of delay spectrum from two snapshots, one each at the beginning, and middle of the observation. Notice the close resemblance in amplitude and structure with those obtained from data.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_combined_baseline_CLEAN_visibilities_contiguous_orientations_custom_gaussian_FG_model_asm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_2_snapshots.eps}
\caption{Delay spectrum for the all-sky foreground model at two pointing positions. {\it Top:} At the beginning of the observation. {\it Bottom:} At the middle of the observation. \label{fig:obsids}}
\end{figure}

We examined the delay spectra obtained from data and through modeling in more detail by comparing the two on a few selected baselines. They are found to match  well with each other.

\section{Delay Spectrum Analysis}\label{sec:delay-spectrum-analysis}

Having established that results from modeling match those from data, we proceed to examine in further detail the signatures seen in the modeled delay spectra.

A number of factors are responsible for the characteristics noted in the delay spectra obtained from data and through modeling. We address these factors below:
\begin{itemize}

\item {\it Sky Model}: Our model of the sky consists of bright compact sources and diffuse emission. Both of these kinds of emission are anisotropic and so is the resulting composite model. In fact, the pointing for MWA observations is chosen from regions of low foreground emission. In our study, we have divided the transiting sky into four different ``bow-tie'' shaped regions with equal areas and zenith as the origin. This results in the ability to attribute the features in the delay spectrum to eight different sectors of the sky.

\item {\it LST}: The LST is also a crucial factor in determining the scheduling of the observation. The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours (see figure~\ref{fig:pointings}). For comparisons, we choose two snapshot observations at 22.09~hours and 0.08~hours LST.

\item {\it Baseline Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our interferometer baselines by their orientation ($\theta_\textrm{b}$) measured counter-clockwise from East. We use the following bins: $-67\fdg 5\le \theta_\textrm{b} < -22\fdg 5$, $-22\fdg 5\le \theta_\textrm{b} < 22\fdg 5$, $22\fdg 5\le \theta_\textrm{b} < 67\fdg 5$, and $67\fdg 5\le \theta_\textrm{b} < 112\fdg 5$. The bin centers are oriented towards South-East, East, North-East, and North respectively.

\item {\it Tile Pointing and Power Pattern}: Since the MWA tiles are phased electronically, the power pattern of the tile changes in any observing mode that tracks the source. Our observations consist of a combination of allowing the sky to drift and tracking the sky. In our study, we take into account the effect of changing power pattern of the antennas on the delay spectrum. For convenience, we refer to the first and second pointing used at the chosen LST as off-zenith and on-zenith pointings respectively.

\item {\it Instrumental Bandpass}: The instrumental bandpass characteristics have a direct effect on the delay transform owing to a Fourier relationship between the two. Due to excision of noisy edge channels repeated in each coarse channel of our bandpass, we see repeated patterns of the {\it foreground wedge} in the delay spectra. As already mentioned earlier, we have used the deconvolution algorithm ({\it CLEAN}) to rid the delay spectra of band shape effects. Since this paper focuses on studying the effects of foreground on the delay spectra, we do not pursue the characterization of artifacts from the band shape and imperfections in {\it CLEAN} deconvolution.

\end{itemize}

The signatures we henceforth identify in the delay spectrum are characterized as arising out of the aforementioned factors. Due to numerous combinations of such factors, we use unique codes for these combinations as described in table~\ref{tab:codes}. The first column refers to the place holder of the specific parameter in the code sequence from left to right. The parameter codes are specified in this order. The second column gives the parameter used in describing features in delay spectrum. The third column denotes the code describing the parameter. Snapshot index uniquely specifies the snapshot and carries a numeric code (1 and 2 for off-zenith and on-zenith snapshots respectively) leading the code sequence. Following this, the type of emission is specified as compact, diffuse, Galactic center, or Galactic plane using codes C, D, GC, or GP respectively. Although in reality, a feature could be due to a combination of emission types, the code we use refers to the predominant type of emission that can be attributed to the feature. The next code specifies the direction of this predominant source of emission. The code values for the direction of emission are the usual abbreviations of the cardinal and ordinal directions besides Z (zenith). Following this is the code for the baseline orientation bin in the half-plane specified as S, SE, E, and NE along directions south, south-east, east, and north-east respectively. Code `A' is used to refer to all baselines. Finally, the parameter describing the region of antenna power pattern coincident with the predominant emission responsible for the feature is specified as P (primary lobe) and S (secondary lobe). The secondary lobe code carries a numeral suffix to further specify uniquely the sidelobe in the direction of emission. For example, a feature code 1-D-W-E-P indicates it is from the first snapshot (1) arising due to diffuse emission (D) westward (W) of zenith observed by predominantly eastward (E) baselines centered on the primary beam (P) of the antenna power pattern. Simiarly, feature code 2-GC-W-NE-S3 implies the feature is seen in the second snapshot (2) due to the Galactic Center (GC) that is westward (W) of zenith by north-eastward (NE) baselines and the region coincides with the third sidelobe (S3) in the direction of emission.

\begin{table*}[htp]
\caption[]{Codes used to describe signatures in delay spectrum.\label{tab:codes}}
\centering
\begin{tabular}{ccc}
\tableline\tableline\tableline
Order & Parameter Description & Code \\
\tableline\tableline
1 & Snapshot index & 1, 2, \ldots \\
2 & Type of emission & C (compact), D (diffuse), GC (Galactic center), GP (Galactic plane) \\
3 & Direction of emission & E (east), N (north), W (west), S (south), Z (zenith)\\
& & NE (north-east), NW (north-west), SW (south-west), SE (south-east) \\
4 & Baseline orientation & E (east), S (south), SE (south-east), NE (north-east), A (all) \\
5 & Region in antenna power pattern & P (primary beam), Sn ($n$-th sidelobe, $n=1,2,\ldots$)\\
& (in the direction of emission) & \\
\tableline
\end{tabular}
\end{table*}

We wish to note that numerous features overlap at varying levels of significance as a result of various combiantions of parameters. We assign the features to their predominant causes. Secondly, we have used noiseless cases to clearly illustrate the observed foreground signatures. With the addition of noise in the visibilities which is subject to observing time, some of the weaker features may not be as prominently visible.

\subsection{Diffuse Foregrounds}\label{sec:diffuse}

Figure~\ref{fig:dsm} shows the noiseless delay spectra due to diffuse emission. Figure~\ref{fig:dsm_delay_spectrum_snap0} and \ref{fig:dsm_delay_spectrum_snap1} correspond to off-zenith and on-zenith snapshots respectively. In general, shorter baselines are expected to be receptive to large scales such as from diffuse emission as well as from small scales from compact objects. This effect is most clearly visible in the on-zenith snapshot noted by feature code 2-D-Z-A-P which stands for the signature observed due to diffuse emission from zenith direction on short baselines of all orientations restricted primarily to the primary beam. This is a notable feature on baselines of length $\lesssim 60$~m. For a frequency of 185~MHz, this length scale corresponds to angular scales of $\gtrsim 1\fdg 5$ on the sky. This feature is noted only for diffuse emission from inside the primary beam.

The most prominent feature, however, is 1-GC-W-E-S3 in figure~\ref{fig:dsm_delay_spectrum_snap0}. This is noted in the off-zenith snapshot caused by the bright Galactic center which is on the west at an altitude of $\sim 33$\arcdeg with $\sim 100$~minutes remaining before it sets. This emission is observed as it is coincident with the third sidelobe of the power pattern in the westward direction with a neagtive delay of value almost close to the horizon limit on eastward baselines.

\begin{figure*}[htp]
\centering
\subfloat[][Off-zenith delay spectra from diffuse foreground model]{\label{fig:dsm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_dsm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_0.eps}}
\subfloat[][On-zenith delay spectra from diffuse foreground model]{\label{fig:dsm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_dsm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_1.eps}}\\
\caption{Delay spectrum visibility amplitudes (in units of Jy~Hz) from diffuse foreground emission obtained on baselines of length $\lesssim 314.7$~m. The left and right panels correspond to off-zenith and on-zenith snapshots respectively. These are noiseless delay spectra obtained after {\it CLEAN} deconvolution along delay axis. A logarithmic color scale is used. Most prominent features are due to the bright Galactic center shining through the third sidelobe (1-GC-W-E-S3), diffuse emission of short baselines at zero delays (2-D-Z-A-P), and diffuse emission on long baselines close to the horizon limit leaking in through the second sidelobes (1-GP-NE-NE-S2, 2-GP-N-N-S2, and 2-GP-S-N-S2). Features from diffuse emission are predominantly edge-heavy resembling a two-pronged fork. See text for a detailed description of these features.}
\label{fig:dsm}
\end{figure*}

\subsubsection{Diffuse Emission On Long Baselines }\label{sec:fork}

Interestingly, in contrast to feature code 2-D-Z-A-P, all the other features in figures~\ref{fig:dsm_delay_spectrum_snap0} and \ref{fig:dsm_delay_spectrum_snap1} extend from the shortest ($\sim 7.7$~m) to longest baselines ($\sim 314.7$~m). We argue this is because even longer baselines are shortened when projected against directions along the baseline vector towards the horizon while considering an all-sky foreground model. Such a shortening due to projection effects allows diffuse emission from close to the horizon to be observed at maximum delays that correspond to the horizon limit even on long baselines. Feature 1-GP-NE-NE-S2 in figure~\ref{fig:dsm_delay_spectrum_snap0}, and features 2-GP-N-N-S2 and 2-GP-S-N-S2 in figure~\ref{fig:dsm_delay_spectrum_snap0} demonstrate this effect of shortening of baselines. 1-GP-NE-NE-S2 corresponds to emission observed from the Galactic plane from north-eastward directions by north-eastward baselines from the second sidelobe in their power pattern. Features 2-GP-N-N-S2 and 2-GP-S-N-S2 are also Galactic plane emission from southward and northward directions observed by second sidelobes of northward baselines with opposite delays. Owing to their apparent structure in delay spectra, we hereafter refer broadly to this edge-heavy structure as the ``two-pronged fork'' feature.

Thus, the most notable features from diffuse emission are those from primary beam area on short baselines, bright Galactic center from the third sidelobe, and the two-pronged fork.

\subsection{Compact Foregrounds}\label{sec:compact}

Figure~\ref{fig:csm} shows the noiseless delay spectra obtained from compact foreground objects. The off-zenith and on-zenith snapshot delay spectra are shown in figures~\ref{fig:csm_delay_spectrum_snap0} and \ref{fig:csm_delay_spectrum_snap1} respectively.

In the off-zenith snapshot, the most prominent feature 1-C-E-E-P arises due to eastward baselines observing the eastward sky at positive delays centered on the primary beam. The next prominent feature 1-C-E-N-P at zero delays arises due to the compact emission from eastward sky centered on the primary beam observed by northward baselines. Features 1-C-NE-NE-S2 and 1-C-W-E-S3 at positive and negative horizon limits are caused by north-eastward sky emission observed by the second sidelobe of north-eastward baselines and westward sky centered on the third sidelobe of eastward baselines respectively.

The most prominent feature in the on-zenith snapshot is 2-C-Z-A-P appearing at zero delays caused by compact sources at zenith observed by all baselines along their primary beams. First and second sidelobes along north and south observed by northward baselines contribute to symmetric positive and negative delays as demonstrated by features 2-C-S-N-S1, 2-C-N-N-S1, 2-C-S-N-S2, and 2-C-N-N-S2.

\begin{figure*}[htp]
\centering
\subfloat[][Off-zenith delay spectra from compact foreground objects]{\label{fig:csm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_csm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_0.eps}}
\subfloat[][On-zenith delay spectra from compact foreground objects]{\label{fig:csm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_csm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_1.eps}}\\
\caption{Same as figure~\ref{fig:dsm} but for compact foreground objects. Color scale is the same as in figure~\ref{fig:dsm}. The different features annotated are a result of emission observed by compact foreground objects through primary beam and sidelobes. Broadly, signatures in delay spectra of compact foreground objects are center-heavy in appearance. See text for a detailed description of the different features.}
\label{fig:csm}
\end{figure*}

While compact foreground objects give rise to features throughout the foreground wedge, the primary distinction between delay spectra of compact and diffuse foregrounds is that the latter are predominantly center-heavy when compared to edge-heavy features of the former.

% Figure~\ref{fig:obs-begin} shows the delay spectra obtained with the compact foreground model described in \S\ref{sec:CSM} at the beginning of the observation when the MWA tiles are phased to point at RA = 4\fdg 387, Dec = -29\fdg 86. The contours of the antenna power pattern are shown in the central panel. The flux densities of compact foreground objects multiplied by the antenna power pattern at the respective locations are shown on a logarithmic color scale right below the central panel. For reference, maps of delays between the received signals are shown for 100~m baselines with $\theta_b=0\arcdeg, 45\arcdeg, 90\arcdeg, \textrm{and}\,135\arcdeg$ to relate the strength of emission with the corresponding delay in the delay spectra.

% \begin{figure*}[htb]
% \centering
% % \subfloat[][Delay spectra at the beginning of observation]{\label{fig:SNR1D_6hrs_167_single_e0}\includegraphics[width=\linewidth]{}}
% \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_CLEAN_visibilities_custom_gaussian_FG_model_csm_all_sky_nside_128_30.7_MHz_bnw2.0_snapshot_0.eps}
% \caption{Delay spectra, compact foreground sources, and delay maps for a snapshot at the beginning of the observation pointed at RA = 4\fdg 387, Dec = -29\fdg 86 at LST = 22.08 hours. The phase center is at RA = -28\fdg 8, Dec = -26\fdg 701. The baselines have been binned by their orientation ($\theta_\textrm{b}$) anti-clockwise from east. The center-right panel shows delay spectra from baselines oriented between -22\fdg 5 $\le \theta_\textrm{b}<$+22\fdg 5. The upper-right, upper-center, and upper-left panels have baseline orientations lying between +22\fdg 5 $\le \theta_\textrm{b}<$+67\fdg 5, +67\fdg 5 $\le \theta_\textrm{b}<$+112\fdg 5, and +112\fdg 5 $\le \theta_\textrm{b}<$+157\fdg 5, respectively. The color scale in these delay spectra are given by the color bar on top. The panels diametrically opposite to aforementioned panels represent the delay map on the sky for a 100~m baseline oriented along east, north-east, north, and north-west respectively. The color scale for these delay maps is specified at the bottom. Delays are positive towards east and north sky positions for eastward and northward baseline orientations respectively. The central panel shows the compact foreground model multiplied by the antenna power pattern. The color scale is specified below the panel. Also shown are contours of the antenna power pattern. The contour levels shown are: 0.0078125, 0.03125, 0.125, and 0.5.}
% \label{fig:obs-begin}
% \end{figure*}

% \begin{figure*}[htb]
% \centering
% % \subfloat[][Delay spectra at the beginning of observation]{\label{fig:SNR1D_6hrs_167_single_e0}\includegraphics[width=\linewidth]{}}
% \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_CLEAN_visibilities_custom_gaussian_FG_model_csm_all_sky_nside_128_30.7_MHz_bnw2.0_snapshot_1.eps}
% \caption{Same as figure~\ref{fig:obs-begin} but for a snapshot during the middle of the observation pointed at RA = +1\fdg 320, Dec = -26\fdg 70 at LST = 0.09 hours. The phase center is at RA = 1\fdg 35, Dec = -26\fdg 701. }
% \label{fig:obs-middle}
% \end{figure*}

\subsection{All-Sky Composite Foreground Model}\label{sec:composite}

Delay spectra from the all-sky foreground model in our study display a composite feature set drawn from the features of compact and diffuse foreground models. Figures~\ref{fig:asm_delay_spectrum_snap0} and \ref{fig:asm_delay_spectrum_snap1} show the delay spectra for the composite all-sky foreground model. While features from both its constituents occupy and overlap in different regions of the foreground wedge, their relatives magnitudes determine the resultant signatures seen in the delay spectra. For instance, the fork-shaped features of diffuse emission are brighter by a factor $\gtrsim 10$ relative to the compact foreground features at the edges of delay spectra in the respective snapshots. Hence, features 1-C-W-E-S3 and 1-C-NE-NE-S2 in figure~\ref{fig:csm_delay_spectrum_snap0} have been masked by features 1-GC-W-E-S3 and 1-GP-NE-NE-S2 respecively from diffuse emission in figure~\ref{fig:dsm_delay_spectrum_snap0}. Similarly, the edge features 2-C-S-N-S2 and 2-C-N-N-S2 from compact objects in figure~\ref{fig:csm_delay_spectrum_snap1} have been masked by diffuse emission features 2-GP-S-N-S2 and 2-GP-N-N-S2 respectively in figure~\ref{fig:dsm_delay_spectrum_snap1}.

On the other hand, in central and inner regions of the wedge, clearly the center-heavy compact foreground model features dominate by $\gtrsim 2$ orders of magnitude relative to features arising out of diffuse emission. This is quite evident since almost all the diffuse emission features are confined to the edges of the wedge. Further, 2-D-Z-A-P in figure~\ref{fig:dsm_delay_spectrum_snap1} is completely masked by 2-C-Z-A-P in figure~\ref{fig:csm_delay_spectrum_snap1}. Thus the inner regions of the wedge consist only of compact foreground features 1-C-E-N-P and 1-C-E-E-P in figure~\ref{fig:asm_delay_spectrum_snap0} and 2-C-Z-A-P, 2-C-S-N-S1 and 2-C-N-N-S1 in figure~\ref{fig:asm_delay_spectrum_snap1}.

Figures~\ref{fig:baseline_composite_asm_delay_spectrum_snap0} and \ref{fig:baseline_composite_asm_delay_spectrum_snap1} show the delay spectra from the diffuse and compact foreground models as well as from the composite model for the off-zenith and on-zenith snapshots respectively for a selected baseline labeled as ``155-154''. This baseline is of length $\simeq 201$~m and orientation $\theta_b\simeq 13\fdg 7$. These are effectively vertical slices of figures~\ref{fig:asm_delay_spectrum_snap0} and \ref{fig:asm_delay_spectrum_snap1} through the baseline location. The horizon delay limits are shown by vertical gray lines. Diffuse emission is negligible in the off-zenith snapshot except at the negative horizon delay limit. This is identified to be feature 1-GC-W-E-S3 noted in figure~\ref{fig:asm_delay_spectrum_snap0} due to the bright Galactic center on the westward sky observed by eastward baseline through its third sidelobe. All the other features are predominantly due to compact foreground objects. The peak at delay of $\sim 0.3\,\mu$~s is identified as feature 1-C-E-E-P in figure~\ref{fig:asm_delay_spectrum_snap0}. The peak at zero delay is from feature 1-C-Z-A-S1 while the other secondary peaks arise compact sources coincident with sidelobes of the power pattern. In the on-zenith snapshot, the feature 1-C-E-E-P has transformed to feature 2-C-Z-A-P due to change in pointing position. The diffuse emission is confined to the edges labeled by features 2-D-W-E-S2 and 2-D-E-E-S2. All intermediate peaks are due to compact foreground objects observed through the different sidelobes of the antenna power pattern.

\begin{figure*}[htp]
\centering
\subfloat[][Off-zenith delay spectra from composite foreground model]{\label{fig:asm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_asm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_0.eps}}
\subfloat[][Off-zenith delay spectra from composite foreground model]{\label{fig:asm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_asm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_1.eps}}\\
\subfloat[][Off-zenith composite delay spectrum on eastward baseline]{\label{fig:baseline_composite_asm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/baseline_155-154_composite_noiseless_delay_spectrum_snapshot_0.edited.eps}}
\subfloat[][On-zenith composite delay spectrum on eastward baseline]{\label{fig:baseline_composite_asm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/baseline_155-154_composite_noiseless_delay_spectrum_snapshot_1.edited.eps}}\\
\caption{{\it Top}: Same as figure~\ref{fig:dsm} but for a composite all-sky foreground model consisting of both diffuse emission and compact objects. The color scale used is the same. The different features annotated arise from those in figures~\ref{fig:dsm} and \ref{fig:csm}. Inner regions of the wedge are dominated by center-heavy features such as 1-C-E-N-P, 1-C-E-E-P, 2-C-Z-A-P, 2-C-N-N-S1 and 2-C-S-N-S1 characteristic of compact objects. Outer regions are dominated by edge-heavy fork-shaped features such as 1-GC-W-E-S3, 1-GP-NE-NE-S2, 2-GP-S-N-S2 and 2-GP-N-N-S2 characteristic of diffuse emission. The resultant structure in delay spectrum takes the form of a {\it pitchfork}. {\it Bottom}: Slices of delay spectrum visibility amplitudes from the top panel at the location of a selected eastward baseline. The baseline vector is $\sim 201$~m in length oriented $\sim 13\fdg 7$ measured anti-clockwise from east. The foreground components are: diffuse sky model (DSM, dashed lines), sky made of compact sources (CSM, dotted lines) and all-sky composite model (ASM, solid lines). The gray vertical lines signify the wedge boundaries set by the horizon delay limit. Compact sources dominate the delay spectra in the inner part of the wedge as marked by features 1-C-E-E-P and 2-C-Z-A-P. Edge-heavy fork-shaped features such as 1-GC-W-E-S3, 2-D-W-E-S2 and 2-D-E-E-S2 from diffuse emission are more significant at the edges of the wedge. These give rise to a ``pitchfork'' appearance. Other secondary peaks inside the wedge are due to compact objects observed through sidelobes of the antenna power pattern. The secondary bumps outside the wedge in all the model components is a result of imperfect deconvolution along delay axis. Refer to text for detailed description of these features.}
\label{fig:asm}
\end{figure*}

Thus the composite all-sky foreground model results in a delay spectrum structure that captures both the edge-heavy two-pronged fork structure from diffuse foreground emission and the center-heavy features from compact foreground objects. Hereafter, we refer to this structure as a ``three-pronged pitchfork''.

\section{Foreground Grading Diagnostic For Observations}\label{sec:fg-grading}

% \section{Discussion}\label{sec:discussion}

\section{Summary}\label{sec:summary}

Our primary motivation in this work is to analyze the signatures of foreground components in the measured delay spectrum. Such an analysis will be helpful in extending our knowledge on the challenge posed by foregrounds in studying the EoR and in devising techniques to mitigate EoR signal contamination from foregrounds. This is one of the first studies that uses an all-sky foreground model to study the resultant signatures on EoR delay spectrum.

Using parameters that match the instrument and EoR observations using the MWA, and an all-sky foreground model that consists of diffuse emission on scales $\gtrsim 0\fdg 85$ and bright compact sources from the NVSS and SUMSS catalogs, we model delay spectra obtained with the MWA on each of its baselines of length $\lesssim 315$~m. We confirm that the modeled delay spectra match the data obtained with the MWA. A wedge shaped structure is clearly noted as predicted by previous studies.

We use our simulations in a noiseless scenario to understand many of the foreground signatures seen in the delay spectra. We find that the delay spectra depend on observing parameters such as antenna pointing and LST, instrument parameters such as antenna power pattern and bandpass shape, and foreground parameters such as the nature of emission, spectral index, etc. A {\it CLEAN} deconvolution in delay-space was applied to remove the contaminating harmonic effects of frequency flagging.

Notable features caused by diffuse emission include the bright Galactic center at the edge of the horizon to be observed through the far sidelobe of the antenna power pattern. As expected, diffuse emission in the primary beam of the antenna power pattern is present on short baselines of length $\lesssim 60$~m. Quite surprisingly, diffuse emission also leaves its footprint on long baselines on either side of the delay axis at the horizon delay limits. This clearly indicates that long baselines appear to be shortened in the direction of the horizon along the baseline vector and these short projected baselines are responsive to large scale diffuse emission. Thus, one of the most important findings of this study is the response of even long baselines to diffuse structures. The characteristic signature of diffuse emission appears to be a roughly symmetric edge-heavy feature taking the form of a {\it two-pronged fork}.

On the other hand, compact foreground objects predominantly map onto inner regions of the wedge. Features arising from compact sources coincident with primary beam and sidelobes of antenna power pattern are clearly noted. In general, compact objects produce delay spectrum signatures that are center-heavy, in stark contrast to those from diffuse foregrounds.

A composite all-sky foreground model consisting of diffuse and compact foregrounds gives rise to a combination of edge-heavy fork shaped features from diffuse foregrounds and center-heavy features from compact foreground objects in the delay spectrum. The compact foreground features are brighter by $\gtrsim 3$ orders of magnitude brighter than the diffuse foreground features in the central regions of delay spectrum. On the other hand, the edge-heavy diffuse foreground features in the delay spectrum are $\gtrsim 1$ order of magnitude brighter than that due to compact foregrounds at the edges of delay spectrum. This gives the appearance of a {\it pitchfork} shape to the resultant delay spectrum.

\acknowledgments

This scientific work makes use of the Murchison Radio-astronomy Observatory, operated by CSIRO. We acknowledge the Wajarri Yamatji people as the traditional owners of the Observatory site. Support for the MWA comes from the U.S. National Science Foundation (grants AST-0457585, PHY-0835713, CAREER-0847753, and AST-0908884), the Australian Research Council (LIEF grants LE0775621 and LE0882938), the U.S. Air Force Office of Scientific Research (grant FA9550-0510247), and the Centre for All-sky Astrophysics (an Australian Research Council Centre of Excellence funded by grant CE110001020). Support is also provided by the Smithsonian Astrophysical Observatory, the MIT School of Science, the Raman Research Institute, the Australian National University, and the Victoria University of Wellington (via grant MED-E1799 from the New Zealand Ministry of Economic Development and an IBM Shared University Research Grant). The Australian Federal government provides additional support via the Commonwealth Scientific and Industrial Research Organisation (CSIRO), National Collaborative Research Infrastructure Strategy, Education Investment Fund, and the Australia India Strategic Research Fund, and Astronomy Australia Limited, under contract to Curtin University. We acknowledge the iVEC Petabyte Data Store, the Initiative in Innovative Computing and the CUDA Center for Excellence sponsored by NVIDIA at Harvard University, and the International Centre for Radio Astronomy Research (ICRAR), a Joint Venture of Curtin University and The University of Western Australia, funded by the Western Australian State government.

\appendix

\par\bigskip
\bibliographystyle{apj}
\bibliography{eor}

\end{document}

### 2014-08-29 foreground_delay_spectrum_modeling.tex
From v0.1: =========
(*) Edited first half of the foreground_power_spectra_modeling.tex file upto and
including section 6.
(*) Removed unnecessary .eps files from the version and use only those needed to
compile.

@@
 We use the 128-tile Murchison Widefield Array (MWA) instrument configuration for our studies.
-Our analysis is centered around 185~MHz with 30.72~MHz bandwidth at 40~kHz frequency resolution.
-We simulate the delay spectrum response of interferometers with antenna pair spacings up to $\sim 315$~m, for the first time, to an all-sky foreground model consisting of diffuse (on scales $\gtrsim\,$0\fdg 85) and compact Galactic and extragalactic emission.
-By matching simulated parameters to those of a subset of first season of observing with the MWA, we identify and verify key foreground and instrumental signatures seen in data.
-A wedge-shaped region of foreground contamination in the delay spectra is confirmed.
-There is a spillover from this contamination into relatively foreground--free regions ({\it EoR window}) in the delay spectrum caused due to the spectral properties of foreground emission and the instrument.
-Due to projection effects, diffuse emission produces edge-heavy features.
+% Our analysis is centered around 185~MHz with 30.72~MHz bandwidth at 80~kHz frequency resolution.
+We simulate the delay spectrum response of interferometers, for the first time, to an all-sky foreground model consisting of diffuse (on scales $\gtrsim\,$0\fdg 85) and compact Galactic and extragalactic emission.
+By tying model parameters to those of a subset of first season of observing with the MWA, the ratio of our predictions to observations, subject to expected errors, are consistent with unity at $\gtrsim$~90\% confidence level.
+A wedge-shaped region of foreground contamination and a spillover from this contamination into relatively foreground--free regions ({\it EoR window}) in the delay spectrum due to spectral properties of foreground emission and the instrument is confirmed.
+Due to shortening of antenna spacings caused by projection effects, diffuse emission produces edge-heavy features in the wedge even on large antenna spacings, hitherto not predicted.
 Features in the inner regions of the wedge are predominantly due to compact foreground objects.
 This results in a ``pitchfork'' shaped delay spectrum.
-We provide a practical diagnostic tool for scheduling EoR observations that helps in reducing foreground contamination.
+We also provide a practical diagnostic tool for data analysis that helps in identifying interferometers with high foreground contamination.
 
@@
 
-After the CMB epoch, the period in the history of the universe often referred to as the {\it dark ages} was followed by the epoch of reionization (EoR).
+After the CMB epoch, the period in the history of the universe often referred to as the {\it dark ages}, followed the epoch of reionization (EoR).
 This was a period of non-linear growth of density perturbations and astrophysical evolution.
 Studying the epoch of reionization holds the key to understanding this evolution.
-Recently, oberving the redshifted 21~cm spin transition of neutral hydrogen has emerged as one of the most promising experiments to fill the gaps in our understanding of the universe's history.
+Recently, observing the redshifted 21~cm spin transition of neutral hydrogen has emerged as a very promising experiment to fill the gaps in our understanding of the universe's history.
 
@@
 In this paper, in one of the most detailed characterizations to date, using an interferometer (antenna pair) based delay spectrum technique, we demonstrate the signatures of unsubtracted diffuse and compact foreground models in delay spectrum, a quantity closely related to the sought power spectrum.
+The simplicity of this technique and usage of basic interferometric visibility data fourier transformed along frequency (delay spectrum) on individual interferometers offer very effective diagnostics for scheduling observations and data analysis, and an insightful view onto foreground signatures in the EoR H{\sc i} power spectrum.
 We explore the diversity in these signatures as a function of observation and instrument parameters and confirm our findings with data obtained from the MWA.
+This simple technique allows us to predict foreground signatures, hitherto not predicted.
 
-We have organized the rest of the paper as follows.
+We have organized this paper as follows.
 \S\ref{sec:delay-spectrum} defines the delay spectrum and its relation to the power spectrum.
 \S\ref{sec:instrument} and \S\ref{sec:obsparms} describe the MWA instrumental parameters and observation parameters used in this study respectively.
-In \S\ref{sec:data-analysis}, we describe our data analysis.
-In \S\ref{sec:modeling}, we describe the building blocks of our delay spectrum modeling including the foreground models used.
-Here, we also establish that the results of modeling and data analysis match with each other.
+Analysis of data obtained with the MWA is described in \S\ref{sec:data-analysis}.
+In \S\ref{sec:modeling}, we describe the building blocks of our delay spectrum modeling including the models used for the antenna power pattern and foreground.
+Here, we also establish that the results of modeling and data analysis agree with each other.
 In \S\ref{sec:delay-spectrum-analysis}, we present a detailed analysis of the signatures of foregrounds on the modeled delay spectrum and its dependence on instrumental and observation parameters.
-In \S\ref{sec:fg-grading}, we use our understanding to grade the delay spectrum as a function of local sidereal time (LST) in order to obtain a useful diagnostic for planning future observations.
+In \S\ref{sec:fg-grading}, we use our understanding to grade interferometers based on the amount of foreground contamination estimated by the delay spectrum technique.
 In \S\ref{sec:summary}, we present our summary and conclusions.
@@
 
-Interferometer array data known as {\it visibilities}, $V_f(\overline{\mathbf{u}},f)$, are correlations between electric fields measured at different frequencies ($f$) by different antenna pairs with separation vectors $\overline{\mathbf{x}}$, where $\overline{\mathbf{u}}\equiv \overline{\mathbf{x}}f/c$ and $c$ is the speed of light.
+Interferometer array data known as {\it visibilities}, $V_f(\overline{\mathbf{u}},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\overline{\mathbf{x}}$ and then Fourier transformed along time axis to obtain a spectrum along frequency ($f$) axis.
+Here, $\overline{\mathbf{u}}\equiv \overline{\mathbf{x}}f/c \equiv \overline{\mathbf{x}}/\lambda$, $c$ is the speed of light and $\lambda$ is the wavelength.
 $\overline{\mathbf{u}}$ is related to transverse spatial frequency modes ($\overline{\mathbf{k}}_\perp$) of brightness distribution of the sky as $\overline{\mathbf{k}}_\perp \equiv 2\pi\overline{\mathbf{u}}/D(z)$, where $D(z)$ is the transverse comoving distance at a redshift $z$.
-If $I(\overline{\mathbf{l}},f)$ is the emission at different frequencies on the sky as a function of direction unit vector ($\overline{\mathbf{l}}$) specified as direction cosines, $V_f(\overline{\mathbf{u}},f)$ represents the Fourier decomposition of $I(\overline{\mathbf{l}},f)$ attenuated by the antenna's angular power pattern $A(\overline{\mathbf{l}},f)$, into transverse spatial frequency components.
+If $I(\overline{\mathbf{l}},f)$ is the emission at different frequencies on the sky as a function of direction unit vector ($\overline{\mathbf{l}}$) specified as direction cosines, $V_f(\overline{\mathbf{u}},f)$ represents the Fourier decomposition of $I(\overline{\mathbf{l}},f)$ attenuated by the antenna's angular power pattern $A(\overline{\mathbf{l}},f)$, into transverse spatial frequency modes.
 Taking into account the instrumental bandpass weights $W_f(f)$, this relation can be written as:
@@
 \end{equation}
-We use the terms delay and lag interchangeably in the rest of this paper.
+The terms delay and lag are used interchangeably in this paper.
 
 Since we are studying a redshifted H{\sc i} spectral line, $f$ is a measure of cosmological distance along the line of sight.
-Consequently, $\eta$ is a measure of line of sight spatial frequency, $k_\parallel$.
+Consequently, $\eta$ is a measure of line of sight spatial frequency mode, $k_\parallel$.
 Thus, $|V_\eta(\overline{\mathbf{u}},\eta)|^2$ is directly related to the spatial power spectrum of redshifted H{\sc i} distribution, $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$.
@@
 These contaminations have to be characterized precisely in order to reduce their impact on redshifted H{\sc i} power spectrum detection sensitivity.
-We, heareafter, examine in detail the delay spectrum of emission from unsubtracted foreground objects and compare with preliminary observations with the MWA.
+
+In summary, $V_\eta(\overline{\mathbf{u}},\eta)$, is closely related to visibilities, $V_f(\overline{\mathbf{u}},f)$, the basic data blocks measured by an interferometer, as seen from equation~\ref{eqn:delay-transform}.
+It captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument.
+At the same time, it is very closely related to the sought power $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ in $k$--modes.
+For our analysis, we use an equivalent quantity, $V_\eta(\overline{\mathbf{x}},\eta) \equiv V_\eta(\lambda\overline{\mathbf{u}},\eta)$.
+For purposes of illustration, we often use a simplified quantity $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$.
+
+We now describe in detail an important ingredient, the MWA instrument.
 
@@
 
-We use the 128-tile layout of the MWA \citep{bea12}.
-We choose a center frequency of 185~MHz with a bandwidth of 30.72~MHz divided into 768 frequency channels each of width 40~kHz.
+We use the 128-tile layout of the MWA \citep{bea12} and choose a center frequency of 185~MHz with a bandwidth of 30.72~MHz divided into 384 frequency channels each of width 80~kHz.
 The choice of center frequency translates to $z\approx 6.68$ for the 21~cm spin flip transition of H{\sc i}.
 The bandwidth was chosen to be close to the maximum instantaneous value of the MWA in order to maximize the resolution in delay coordinates and obtain detailed characteristics of delay spectra from foregrounds.
-We restrict our analysis to the innermost 2000 pairs covering $|\overline{\mathbf{x}}| \leq x_\textrm{max} \approx 315$~m, or $|\overline{\mathbf{u}}| \leq u_\textrm{max} \approx 157.5$ at a frequency of 185~MHz.
-These interferometers are responsive to transverse spatial structures on scales $\gtrsim 1/u_\textrm{max} \approx 22$\arcmin.
+We restrict our analysis to the innermost antenna spacings with $|\overline{\mathbf{x}}| \leq x_\textrm{max} = 200$~m, or $|\overline{\mathbf{u}}| \leq u_\textrm{max} = 123$ at a frequency of 185~MHz.
+These interferometers are responsive to transverse spatial structures on scales $\gtrsim 1/u_\textrm{max} \approx 28$\arcmin.
+Hereafter, we use the terms tile and antenna interchangeably to refer to the MWA tile.
 
 The bandpass weights in the frequency window of the MWA are obtained through a 8-tap Polyphase Filter Bank (PFB) using a Kaiser window with the $\beta$ parameter set to 5.
-The bandpass weights consist of coarse frequency channels of width 1.28~MHz each of which contains 32 finer frequency channels of width 40~kHz.
+The bandpass weights consist of coarse frequency channels of width 1.28~MHz each of which contains 16 finer frequency channels of width 80~kHz.
 Our analysis, therefore, contains 24 coarse channels.
 As the data are subsequently bandpass calibrated, the bandpass weights are normalized to unity.
-But in the edges of the coarse channels where the bandpass correction factors are $\gtrsim 3$, this correction makes the edge channels signficantly noisy compared to the rest of the spectrum.
-Thus, we remove 2 fine channels from each side of the coarse channel edges from our data.
+But the edges of the coarse channels are noisy and aliased from adjacent coarse channels.
+Thus, we remove one fine channel from each side of the coarse channel edges from our data.
 This constitutes $W_f(f)$ in equation~\ref{eqn:obsvis}.
@@
 The {\it plus} symbols joined by dashed lines are the bandpass correction factors.
-The dot-dashed line represents the threshold for bandpass correction factor above which frequency channels are flagged.
-The solid line shows the flagged and gain--corrected bandpass shape, $W_f(f)$.
-These properties hold for the entire bandwidth of 30.72~MHz centered around 185~MHz.
+The solid line shows the flagged (gaps) and gain--corrected bandpass shape, $W_f(f)$.
+These shapes are repeated every 1.28~MHz for the entire bandwidth of 30.72~MHz centered around 185~MHz.
 \label{fig:bandpass}}
@@
 
-Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=440$~K \citep{bow06} for all frequency channels for all interferometers throughout the course of observation.
+Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=90$~K per polarization for all frequency channels for all interferometers throughout the course of observation.
 We take into account the increase in thermal uncertainty in frequency channels that results due to aforementioned bandpass correction.
-
-The power pattern of a MWA tile, $A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}, over the entire hemisphere is modeled based on a 4$\times$4 phased array of half-wave dipoles.
-Hereafter, we use the terms tile and antenna interchangeably to refer to the MWA tile.
 
@@
 
-One of the primary targets for EoR observations with the MWA is a patch of sky centered around RA = 00h 00m 00.00s, Dec = -30\arcdeg$\,$00\arcmin$\,$00\farcs 0.
+One of the primary targets for EoR observing with the MWA is a patch of sky centered around RA = 00h 00m 00.00s, Dec = -30\arcdeg$\,$00\arcmin$\,$00\farcs 0.
 This is one of the regions of low foreground emission within 27\arcdeg$\,$ from the zenith.
-The MWA tracks a patch of sky through beams formed and steered electronically by changing delay settings between a phased array of dipoles in a MWA tile.
+The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of an array of dipoles in a MWA tile.
 These beamformer delay settings can be changed only in discrete steps.
@@
 Figure~\ref{fig:pointings} shows the direction of pointing of MWA tiles during the course of the observation.
-The sky is allowed to drift for a certain period of time (usually $\sim 30$~mins) before the discrete shifts in the beamformer delay settings position the beam to be centered again on the patch of sky.
+The sky is allowed to drift for a certain period of time (usually $\sim 30$~mins) before the discrete shifts in the beamformer delay settings position the beam to be centered again on the same patch of sky.
 This process is repeated throughout the course of the observation $\approx 4.86$~hours.
+The LST of two snapshots we chose for our study are shown as vertical lines at -1.92~hours (wrapped by 360\arcdeg from 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings respectively.
+These pointings (phased array power patterns) are centered at RA = 4\fdg 387, Dec = -29\fdg 86 and RA = -28\fdg 8, Dec = -26\fdg 701, whereas the visibilities themselves are phased to zenith corresponding to RA = -28\fdg 8, Dec = -26\fdg 701, and RA = 1\fdg 35, Dec = -26\fdg 701 respectively.
 
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/dns_pointings}
-\caption{MWA tile pointing directions during the course of the observation.
+\includegraphics[width=\linewidth]{figures/v1_0/custom_pointings.eps}
+\caption{MWA tile beam pointing directions during the course of the observation.
 The x-axis refers to the Local Sidereal Time (LST) in hours.
@@
 The sky is allowed to drift for $\sim 30$~mins before the beamformer delay settings change in a discrete step to center the beam around RA = 00h 00m 00.00s, Dec = -30\arcdeg$\,$00\arcmin$\,$00\farcs 0.
+Dot--dashed vertical lines show the LST of off-zenith and zenith pointings used in our study.
 \label{fig:pointings}}
@@
 
+Throughout this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east and north (along local meridian) directions at the MWA site.
+For the conventions of Fourier transform and its inverse we use in this paper, in accordance with equation~\ref{eqn:delay}, signals from the sky towards east and north are recorded with positive delays on eastward and northward oriented antenna spacings respectively.
+For instance, at the beginning of the observation, the Galactic center is in the westward sky just about to set.
+Signals from this direction are recorded with a negative delay on interferometers whose antenna spacings are oriented towards the east.
+Similarly, signals from directions eastward of the local meridian arrive with positive delays on eastward oriented antenna spacings.
+
+For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all antenna spacings to lie in the range $-67\fdg 5 \leq \theta_b < 112\fdg 5$.
+Interferometers with antenna spacings oriented in the other half-plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
+
 \section{Data Analysis}\label{sec:data-analysis}
 
-{\bf Describe the data analysis here.}
+\subsection{Imaging and deconvolution}\label{sec:FHD}
 
-Throughout this paper, $\overline{\mathbf{u}}$ is assumed to be on a coordinate system aligned with the local east and local north directions at the MWA site.
-For the conventions of Fourier transform and its inverse we use in this paper, in accordance with equation~\ref{eqn:delay}, signals from the sky towards east and north are recorded with positive delays on eastward and northward baselines respectively.
-For instance, at the beginning of the observation, the Galactic center is on the far west just about to set.
-Signals from this direction are recorded with a negative delay on interferometers oriented towards the east.
-Similarly, signals from directions eastward of the local meridian arrive with positive delays on eastward interferometers.
-In subsequent discussions, unless mentioned otherwise, all the snapshot visibilities are assumed to be phased to zenith.
-
-For purposes of geometrical intuition, we restrict the orientation ($\theta_b$ measured anti-clockwise from East) of all baselines to lie in the range $-67\fdg 5 \leq \theta_b < 112\fdg 5$.
-Baselines in the other half-plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
+{\bf Describe the FHD data analysis up to the point of getting calibrated visibilities here.}
 
@@
 
-Subsequently, we obtain the delay spectrum of these visibilities using equation~\ref{eqn:delay-transform} while choosing $W'_f(f)$ to be a {\it Blackman--Nuttall} window function for its superior sidelobe suppression as described in \citet{thy13}.
+We obtain the delay spectrum of these calibrated visibilities using equation~\ref{eqn:delay-transform} while choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function.
 Due to periodic gaps in frequency bandpass occurring in intervals of 1.28~MHz (see figure~\ref{fig:bandpass}), the delay spectrum is expected to contain harmonics of actual foreground emission repeated at intervals of 0.78~$\mu$s.
@@
 
-\subsection{Delay Spectrum from Data}\label{sec:data-delay-spectrum}
+\subsection{Delay Spectrum after Delay Deconvolution}\label{sec:data-delay-spectrum}
 
-Put up plots here and discuss them.
+We show results of delay spectrum from MWA data after deconvolution along delay axis in figure~\ref{fig:fhd_data}.
+Boundaries of the {\it foreground wedge} are also shown.
+The {\it off--zenith} pointing has emission notably higher than in the {\it zenith} pointing inside the wedge boundaries.
+This is shown later to be due to response of the interferometer array to the bright Galactic center and Galactic plane in the westward sky.
+Consequently, the contamination into the {\it EoR window} is also higher.
+Emission in {\it zenith} pointing is more centrally concentrated due to phasing of power pattern to the zenith, whereas the {\it off--zenith} pointing has a power pattern phased eastward of zenith and is also responsive to emission from angles far away from zenith.
 
-Despite the deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the data perfectly.
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
+\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with MWA data for {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+White lines mark the boundaries of {\it foreground wedge} determined by the horizon limit and antenna spacing.
+The features in the wedge appear to be brighter and the spillover contamination appears to be higher in the {\it off--zenith} pointing relative to the {\it zenith} pointing due to the presence of brighter foregrounds such as the Galactic center and the Galactic plane.
+The logarithmic color scale (shown at the right) is common to both panels.\label{fig:fhd_data}}
+\end{figure}
+
+Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the data perfectly.
 This is predominant at short baselines.
@@
 
-Below we describe the foreground model we have used in modeling the measured delay spectra.
+We describe power pattern and foreground models we have used in modeling the measured delay spectra.
+
+\subsection{Power pattern}\label{sec:power_pattern}
+
+The power pattern of a MWA tile, $A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}, over the entire hemisphere is analytically modeled based on a 4$\times$4 phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane.
+We also assume that the delay corrections during the phased addition of dipole voltages in the beamformer suffer from random fluctuations of rms 0.05~ns.
+The models of power pattern used in data analysis take into account effects of mutual coupling of dipoles in the tile besides finite ground plane effects.
+Our models were found to be in good agreement with those used in the data analysis.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom.eps}
+\caption{Models of MWA tile power pattern in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates for the off--zenith (top) and zenith pointings (bottom) at 185~MHz.
+An MWA tile is modeled as a 4$\times$4 array of isotropic radiators placed 0.3~m above the ground plane.
+Random fluctutations of rms 0.05~ns have been added to delay corrections during the phased addition of voltages from these isotropic radiators.
+The logarithmic color scale (shown at the right) is common to both panels.
+\label{fig:power_pattern}}
+\end{figure}
 
@@
 
-An instrument such as the MWA has a very wide field of view ($\sim 30$\arcdeg at 185~MHz) for imaging purposes.
-In the context of EoR H{\sc i} power spectrum it has been shown that unsubtracted forground sources anywhere in the horizon directly contaminate the spatial frequency modes in the power spectrum.
+An instrument such as the MWA has a very wide field of view ($\gtrsim 20$\arcdeg at 185~MHz) for imaging purposes.
+In the context of EoR H{\sc i} power spectrum, it has been known that unsubtracted forground sources anywhere in the visible hemisphere up to the horizon directly contaminate the spatial frequency modes in the power spectrum.
 In addition, this contamination also spills over into the relatively cleaner regions, called {\it EoR window}, due to spectral properties of the instrument and the foregrounds \citep{thy13,pob13,ved12,par12}.
-Thus, it is important to consider an all-sky model for foreground objects in evaluating the features seen in the power spectrum.
+Thus, it is important to consider an all-sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting to the primary field of view.
 
@@
 \citet{thy13} estimated the sensitivity taking into account the spillover from foreground contamination from unsubtracted extragalactic point sources in the {\t EoR window} besides thermal noise.
-In this paper, we include the diffuse Galactic emission for a comprehensive understanding of foreground signatures in the power spectrum.
+In this paper, we include the diffuse Galactic emission for enhancing our understanding of foreground signatures in the power spectrum.
 We use an all-sky foreground emission model that consists of diffuse and bright compact components.
@@
 
-For the diffuse component, we use an all-sky radio foreground model from \citet{deo08} to estimate the emission at 185~MHz.
+For the diffuse component, we use an all-sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz.
 Since this map is predominantly based on the 408~MHz all-sky map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz diffuse emission map to the same resolution.
 However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals.
-Using the same source model, we also obtain the diffuse emission maps at 140~MHz and 160~MHz to estimate the spectral index map of the diffuse emission model.
+Using the same source model, we also obtain the diffuse emission maps at 170~MHz and 200~MHz to estimate the spectral index map of the diffuse emission model.
+
+A low resolution version of the diffuse foreground model used is show in figure~\ref{fig:DSM}.
+Contours of the MWA tile power pattern shown in figure~\ref{fig:power_pattern} are also overlaid.
+Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the visible hemisphere in the westward sky and the MWA tile's power receptivity is significantly high ($\gtrsim 0.125$) in that direction.
+In the {\it zenith} pointing, the Galactic plane has almost set and the contour level of power pattern in that direction is at least 16 times lesser.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\linewidth]{figures/v1_0/dsm.eps}
+\caption{Sky brightness temperature (in K) of the diffuse foreground model at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+The color scale on the right is logarithmic and is common to both panels.
+Contours of power pattern shown in figure~\ref{fig:power_pattern} are overlaid.
+The contour levels shown are 0.001953125, 0.0078125, 0.03125, 0.125, and 0.5.
+The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off--zenith} pointing and the MWA tile power receptivity is significant ($\gtrsim 0.125$) in that direction.
+In contrast, emission from the Galactic plane in {\it zenith} pointing is significantly lesser.
+\label{fig:DSM}}
+\end{figure}
 
@@
 Compact sources from NVSS were selected if the convolved major axes were not greater than $\approx 47$\arcsec, which is almost equal to the angular resolution of the survey.
-Using a mean spectral index of $\langle\alpha\rangle=-0.83$ (flux density, $S(f)\propto f^\alpha$) obtained by \citet{mau03} for both NVSS and SUMSS catalog objcts, we calculate the corresponding flux densities at 185~MHz, $S_{185}$.
+Using a mean spectral index of $\langle\alpha_\textrm{sp}\rangle=-0.83$ (flux density, $S(f)\propto f^{\alpha_\textrm{sp}}$) obtained by \citet{mau03} for both NVSS and SUMSS catalog objcts, we calculate the corresponding flux densities at 185~MHz, $S_{185}$.
 From this subset, we choose compact objects with $S_{185}\geq 10$~Jy.
+The selection of such bright compact objects is not affected by minor differences in sensitivity of the two surveys.
 We verified that our selection criteria ensure a similar areal density of sources in the two surveys.
@@
 Based on these criteria, we selected 133 sources from the SUMSS catalog and 336 sources from the NVSS catalog.
-The selection of such bright compact objects is not affected by minor differences in sensitivity of the two surveys.
 Together with the diffuse foreground model, we obtain an all-sky foreground model consisting of both compact and diffuse emission.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\linewidth]{figures/v1_0/csm.eps}
+\caption{Flux densities (in Jy) of bright compact sources at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+The color scale on the right is logarithmic and is common to both panels and corresponds to the color--coded filled circles.
+Contours of power pattern shown in figure~\ref{fig:power_pattern} are overlaid.
+The contour levels shown are 0.001953125, 0.0078125, 0.03125, 0.125, and 0.5.
+The locations of bright compact sources appear to be random.\label{fig:CSM}}
+\end{figure}
 
@@
 With the aforementioned all-sky foreground model, and instrumental and observational parameters, we simulate visibilities using equation~\ref{eqn:obsvis}.
-We consider two snapshots, one at the beginning and one during the middle of the observation, to explain the signatures of foregrounds seen in the modeled and observed data.
-The first snapshot is pointed at RA = 4\fdg 387, Dec = -29\fdg 86 at 22.08 hours LST while the second is pointed at RA = +1\fdg 320, Dec = -26\fdg 70 at 0.09 hours LST.
-In both cases the visibilities are phased to zenith corresponding to RA = -28\fdg 8, Dec = -26\fdg 701, and RA = 1\fdg 35, Dec = -26\fdg 701 respectively.
-Signals from the phase center are received simulatneously with no delay between the interferometers.
-
-Figure~\ref{fig:obsids} shows the amplitude of delay spectrum from two snapshots, one each at the beginning, and middle of the observation.
-Notice the close resemblance in amplitude and structure with those obtained from data.
+Figure~\ref{fig:sim_data} shows the amplitude of delay spectrum from {\it off--zenith} and {\it zenith} pointings.
+Notice the qualitative agreement of amplitude and structure with those obtained from data shown in figure~\ref{fig:fhd_data}.
+The Galactic center and the Galactic plane visible in the {\it off--zenith} pointing make it appear brighter in the {\it foreground wedge}.
 
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/multi_combined_baseline_CLEAN_visibilities_contiguous_orientations_custom_gaussian_FG_model_asm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_2_snapshots.eps}
-\caption{Delay spectrum for the all-sky foreground model at two pointing positions. {\it Top:} At the beginning of the observation. {\it Bottom:} At the middle of the observation. \label{fig:obsids}}
+\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
+\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+White lines mark the boundaries of {\it foreground wedge} determined by the horizon limit and antenna spacing.
+The features resemble those obtained with MWA data shown in figure~\ref{fig:fhd_data}.
+The Galactic center and Galactic plane prominently visible in the {\it off--zenith} pointing makes the emission in the {\it foreground wedge} brighter relative to that in the {\it zenith} pointing.
+The axes and color scale are identical to those in figure~\ref{fig:fhd_data}.
+\label{fig:sim_data}}
 \end{figure}
 
-We examined the delay spectra obtained from data and through modeling in more detail by comparing the two on a few selected baselines. They are found to match  well with each other.
+In order to make a quantitiative comparison of delay spectra obtained with the MWA data and our simulations, we consider the following uncertainties.
+Our foreground models are derived from other higher frequency catalogs and sky maps.
+The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observing frequency.
+Using simple error propagation, the fractional error in the delay spectrum caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the observing frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index.
+From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for compact sources from NVSS and SUMSS catalogs.
+We assume the same holds true for our diffuse sky model as well, which is predominantly derived from 408~MHz map of \citet{has82}.
+Besides, delay spectra from simulations and data each have uncertainties arising from thermal noise with rms, $\Delta V^\textrm{N}_\eta(|\overline{\mathbf{x}}|,\eta) \sim 1.4\times 10^7$~Jy~Hz, where the superscript N stands for thermal noise.
+We estimate the ratio of delay spectra from data and simulations as, $\rho = |V^\textrm{D}_\eta(|\overline{\mathbf{x}}|,\eta)|/|V^\textrm{S}_\eta(|\overline{\mathbf{x}}|,\eta)|$, where superscripts D and S denote data and simulation respectively.
+Using the aforementioned uncertainties added in quadrature, we estimate the approximate expected uncertainty in $\log_{10}\rho$ and denote it by $\Delta\,\log_{10}\rho$.
+
+Figure~\ref{fig:data-sim-ratio-snr} shows the significance of $\log_{10}\rho$ relative to the expected error $\Delta\,\log_{10}\rho$ for all $|\overline{\mathbf{x}}|$ and $\eta$ inside the {\it foreground wedge}.
+The probability densities and cumulative distributions are shown in black and gray respectively.
+Also shown are dotted vertical lines which denote equality between $\Delta\,\log_{10}\rho$ and $\log_{10}\rho$.
+The fraction of the {\it foreground wedge} that lies in between these vertical lines is $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively.
+In other words, subject to expected uncertainties, the ratio of delay spectra obtained with MWA data and simulations are consistent with unity at $\sim$~90\% and $\sim$~95\% confidence levels in the {\it off--zenith} and {\it zenith} pointings respectively.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
+\caption{Significance of logarithm of ratio of observed to simulated visibilities relative to expected uncertainty restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+Probability densities and cumulative distributions are shown in black and gray respectively.
+Region inside the dotted vertical lines denotes when the ratio, $\rho$, is consistent with unity subject to expected uncertainty.
+Also indicated in percentages are confidence levels estimated conservatively for such a consistency.
+They are $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively.
+\label{fig:data-sim-ratio-snr}}
+\end{figure}
+
+These uncertainties are presented only to confirm the qualitative agreement already seen between figures~\ref{fig:fhd_data} and \ref{fig:sim_data} and are not intended to serve as a comprehensive estimate of all uncertainties involved.
+However, since we have not accounted for numerous other uncertainties, especially in the data, such as uncertainties in the antenna power pattern, calibration, radio frequency interference (RFI), and wide field imaging effects, our estimates are conservative.
+Accounting for such uncertainties will only improve our confidence level in this comparison.
+Our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground window}.
 
@@
 
-\item {\it LST}: The LST is also a crucial factor in determining the scheduling of the observation.
-The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours (see figure~\ref{fig:pointings}).
-For comparisons, we choose two snapshot observations at 22.09~hours and 0.08~hours LST.
-
 \item {\it Baseline Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our interferometer baselines by their orientation ($\theta_\textrm{b}$) measured counter-clockwise from East.
@@
 In our study, we take into account the effect of changing power pattern of the antennas on the delay spectrum.
-For convenience, we refer to the first and second pointing used at the chosen LST as off-zenith and on-zenith pointings respectively.
+The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours (see figure~\ref{fig:pointings}).
+For comparisons, we choose two snapshot observations at LST of 22.08~hours (off--zenith pointing) and 0.09~hours (zenith pointing).
 

### 2014-09-03 foreground_delay_spectrum_modeling.tex
new abstract, updated intro

@@
 
-We characterize signatures of foreground radio emission and instrument configuration on the observed power spectra of redshifted H{\sc i} 21~cm line emission from the epoch of reionization (EoR) using a baseline (antenna pair) based ``delay spectrum'' analysis technique.
+ Detection and measurement of 21cm HI emission from the Epoch of Reionization, at redshifts above 6, is limited primarily by foreground emission.
+With much of the information coming from the spectral line that maps to time and line of site distance, foregrounds affecting the spectral domain are deserving of close attention, particularly those that are due to the instrument because they are in theory removable.
+Here we investigate in detail the chromatic appearance of flat spectrum foregrounds imprinted by the instrumental transfer function.
+Using the delay spectrum technique to measure power spectrum contamination on a per-baseline basis we demonstrate that the foreground signatures which have the largest impact on the HI signal arise from power entering far from the center of the field of view.
+Comparing data from recent Murchison Widefield Array observations with simulations broken down by sky position and type of emission, we identify smooth emission near the horizon as the largest contributing factor, followed by bright sources in the same locations.
+The observation that these contaminants only effect a fraction of the baselines leads us to propose that by down-weighting certain baselines at certain times, a large fraction of foreground contamination could be mitigated.
+%We characterize signatures of foreground radio emission and instrument configuration on the observed power spectra of redshifted H{\sc i} 21~cm line emission from the epoch of reionization (EoR) using a baseline (antenna pair) based ``delay spectrum'' analysis technique.
 We use the 128-tile Murchison Widefield Array (MWA) instrument configuration for our studies.
 % Our analysis is centered around 185~MHz with 30.72~MHz bandwidth at 80~kHz frequency resolution.
-We simulate the delay spectrum response of interferometers, for the first time, to an all-sky foreground model consisting of diffuse (on scales $\gtrsim\,$0\fdg 85) and compact Galactic and extragalactic emission.
+%We simulate the delay spectrum response of interferometers, for the first time, to an all-sky foreground model consisting of diffuse (on scales $\gtrsim\,$0\fdg 85) and compact Galactic and extragalactic emission.
 By tying model parameters to those of a subset of first season of observing with the MWA, the ratio of our predictions to observations, subject to expected errors, are consistent with unity at $\gtrsim$~90\% confidence level.
@@
 
-One of the key challenges in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
-\citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which lack such symmetry.
-Foreground removal techniques rely on the spectral smoothness of foreground emission \citep{mor06,bow09,liu11,par12,dil13}.
+Perhaps the key challenge in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
+\citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness\citep{mor06,bow09,liu11,par12,dil13}.
+Considerable effort has been made towards understanding the $k$-space behavior foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,dat10,bow09}.
+A general model of how smooth spectrum foregrounds are scattered to higher $k$-modes by the wide-field response of the instrument, into the so-called `wedge',  has emerged that provides a high level explanation for the observed foreground power spectrum.
+The conservative foreground strategy that has developed alongside this work is to throw out $k$-modes which could be contaminated (Avoidance).
+The more aggressive option is to subtract a sky model and regain access to modes lost by Avoidance.
+In both cases one large unknown is which parts of the sky are most critical to either Avoid or Subtract.
+Here we will focus primarily on the Avoidance strategy, asking which foreground components are most likely to ``leak'' from foreground modes to Reionization modes.
 
-Precise understanding of these foreground signatures is necessary in order to remove foreground systematics from the measured power spectrum.
-Considerable effort has been made in mapping the residual foreground signatures in observed power spectrum \citep{thy13,pob13,mor12,tro12,dat10,bow09} after foreground subtraction.
+Foregrounds work their way into the power spectrum by either having intrinsic spectral variability, high instrumental chromaticity, or through approximations in the estimating the power spectrum.
+Here we are focusing on the instrumental effects.
+To de-couple from possible analysis effects, such as those pointed out by \citet{Haz13}, we compute power spectra using the per-baseline approach of \citet{pob13,arp14}.
+This approximates the power spectrum as the fourier transform of the spectra output by the correlator.
 
-In this paper, in one of the most detailed characterizations to date, using an interferometer (antenna pair) based delay spectrum technique, we demonstrate the signatures of unsubtracted diffuse and compact foreground models in delay spectrum, a quantity closely related to the sought power spectrum.
-The simplicity of this technique and usage of basic interferometric visibility data fourier transformed along frequency (delay spectrum) on individual interferometers offer very effective diagnostics for scheduling observations and data analysis, and an insightful view onto foreground signatures in the EoR H{\sc i} power spectrum.
-We explore the diversity in these signatures as a function of observation and instrument parameters and confirm our findings with data obtained from the MWA.
-This simple technique allows us to predict foreground signatures, hitherto not predicted.
+ Matching real observations from the Murchison Widefield Array Epoch of Reionization observing program with instrumental simulations of the entire known sky we are examine the signature of the foreground power spectrum on each baseline.
+We discover that the foregrounds most likely to obfuscate the Epoch of Reionization power are not caused by power emitted from the same sky location as the target HI signal, and that smooth galactic emission plays a larger role then previously thought.
+We quantify this contamination by breaking the simulations down by sky region and type of emission.
+Using these we arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to downweight badly contaminated baselines.
 
-We have organized this paper as follows.
-\S\ref{sec:delay-spectrum} defines the delay spectrum and its relation to the power spectrum.
-\S\ref{sec:instrument} and \S\ref{sec:obsparms} describe the MWA instrumental parameters and observation parameters used in this study respectively.
-Analysis of data obtained with the MWA is described in \S\ref{sec:data-analysis}.
-In \S\ref{sec:modeling}, we describe the building blocks of our delay spectrum modeling including the models used for the antenna power pattern and foreground.
-Here, we also establish that the results of modeling and data analysis agree with each other.
-In \S\ref{sec:delay-spectrum-analysis}, we present a detailed analysis of the signatures of foregrounds on the modeled delay spectrum and its dependence on instrumental and observation parameters.
-In \S\ref{sec:fg-grading}, we use our understanding to grade interferometers based on the amount of foreground contamination estimated by the delay spectrum technique.
-In \S\ref{sec:summary}, we present our summary and conclusions.
+In \S\ref{sec:delay-spectrum} we introduce the delay spectrum technique, then in  \S\ref{sec:instrument}, \S\ref{sec:obsparms}, and \S\ref{sec:data-analysis} we describe the MWA setup, summarize the observing parameters and present the resulting data.
+Simulations of these observing parameters are described in \S\ref{sec:modeling} and analyzed by sky location, and type of emission in \S\ref{sec:delay-spectrum-analysis}.
+\S\ref{sec:discussion} offers a discussion along with a basic comparison to other current and future instruments.
+\S\ref{sec:conclusion} offers conclusions and an initial description of a more precise foreground avoidance technique.
 

### 2014-09-03 foreground_delay_spectrum_modeling.tex
From v0.2: =========
(*) Edited second half of the foreground_power_spectra_modeling.tex file beyond section 6.
(*) Removed unnecessary .eps files from the version and use only those needed to
compile.
(*) Simplified justification of agreement of data with simulations.

--- 2014-08-29 foreground_delay_spectrum_modeling.tex
@@
 
-We characterize signatures of foreground radio emission and instrument configuration on the observed power spectra of redshifted H{\sc i} 21~cm line emission from the epoch of reionization (EoR) using a baseline (antenna pair) based ``delay spectrum'' analysis technique.
+We characterize signatures of foreground radio emission and instrument configuration on the observed power spectra of redshifted H{\sc i} 21~cm line emission from the epoch of reionization (EoR) using an antenna pair based ``delay spectrum'' analysis technique.
 We use the 128-tile Murchison Widefield Array (MWA) instrument configuration for our studies.
 % Our analysis is centered around 185~MHz with 30.72~MHz bandwidth at 80~kHz frequency resolution.
-We simulate the delay spectrum response of interferometers, for the first time, to an all-sky foreground model consisting of diffuse (on scales $\gtrsim\,$0\fdg 85) and compact Galactic and extragalactic emission.
-By tying model parameters to those of a subset of first season of observing with the MWA, the ratio of our predictions to observations, subject to expected errors, are consistent with unity at $\gtrsim$~90\% confidence level.
-A wedge-shaped region of foreground contamination and a spillover from this contamination into relatively foreground--free regions ({\it EoR window}) in the delay spectrum due to spectral properties of foreground emission and the instrument is confirmed.
-Due to shortening of antenna spacings caused by projection effects, diffuse emission produces edge-heavy features in the wedge even on large antenna spacings, hitherto not predicted.
+We simulate the delay spectrum response of interferometers, for the first time, to an all-sky foreground model consisting of diffuse and compact Galactic and extragalactic emission.
+By tying model parameters to those of a subset of first season of observing with the MWA, our simulations agree with the observed data to within expected uncertainty.
+A wedge--shaped region of foreground contamination and a spillover from this contamination into relatively foreground--free regions ({\it EoR window}) in the delay spectrum due to spectral properties of foreground emission and the instrument is confirmed.
+Due to shortening of antenna spacings caused by projection effects, diffuse emission produces edge--heavy {\it two--pronged fork}--like features even on wide antenna spacings, hitherto not predicted.
 Features in the inner regions of the wedge are predominantly due to compact foreground objects.
-This results in a ``pitchfork'' shaped delay spectrum.
-We also provide a practical diagnostic tool for data analysis that helps in identifying interferometers with high foreground contamination.
+This results in a {\it pitchfork}--shaped delay spectrum.
+The bright Galactic center severly contaminates the {\it EoR window} on certain antenna spacings in certain observations.
+For such cases, we present an effective practical tool for screening interferometers with high foreground contamination and mitigating their effects on EoR data analysis.
 
@@
 It captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument.
-At the same time, it is very closely related to the sought power $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ in $k$--modes.
+At the same time, it is very closely related to the sought power $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ in $k$--modes containing critical information about spatial scales.
 For our analysis, we use an equivalent quantity, $V_\eta(\overline{\mathbf{x}},\eta) \equiv V_\eta(\lambda\overline{\mathbf{u}},\eta)$.
@@
 
-Throughout this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east and north (along local meridian) directions at the MWA site.
+Throughout this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the MWA site.
 For the conventions of Fourier transform and its inverse we use in this paper, in accordance with equation~\ref{eqn:delay}, signals from the sky towards east and north are recorded with positive delays on eastward and northward oriented antenna spacings respectively.
@@
 \caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with MWA data for {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
-White lines mark the boundaries of {\it foreground wedge} determined by the horizon limit and antenna spacing.
+White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
 The features in the wedge appear to be brighter and the spillover contamination appears to be higher in the {\it off--zenith} pointing relative to the {\it zenith} pointing due to the presence of brighter foregrounds such as the Galactic center and the Galactic plane.
@@
 Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the data perfectly.
-This is predominant at short baselines.
+This is predominant at short antenna spacings.
 This is because the maximum delay envelope (boundary of the wedge) consists of mixed emission from large portions of the sky into a narrow range of delays and the deconvolution algorithm does not have sufficient support in delay space to fit the delay spectrum.
-However, we note that the purpose of deconvolution is to rid the delay spectra of instrumental artifacts to the best extent possible in order to see the underlying signatures of foreground emission.
+However, we wish to emphasize that the purpose of deconvolution is to rid the delay spectra of instrumental artifacts to the best extent possible in order to see the underlying signatures of foreground emission.
 It is not our intention in this paper to use the deconvolution as a foreground removal tool.
-Hence, our discussion and conclusions still hold despite imperfections in deconvolution.
+Hence, our discussion and conclusions are valid despite imperfections in deconvolution.
 
@@
 The models of power pattern used in data analysis take into account effects of mutual coupling of dipoles in the tile besides finite ground plane effects.
-Our models were found to be in good agreement with those used in the data analysis.
+Our models were found to be in reasonable agreement with those used in the data analysis.
 
@@
 
-We use a combination of NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz due to their similar flux sensitivity and angular resolution, and complimentary survey footprints covering the entire sky.
+We use a combination of NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz due to their matched flux sensitivity and angular resolution, and complimentary survey footprints covering the entire sky.
 The SUMSS catalog covers the sky with declination $\delta < -30$\arcdeg with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec.
@@
 
-From the SUMSS catalog, we select compact sources whose deconvolved major axes are equal to 0.
+From the SUMSS catalog, we select compact sources whose deconvolved major axes are equal to 0\arcsec.
 From the NVSS catalog, we excluded objects that overlap with those in the SUMSS survey footprint.
@@
 \caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
-White lines mark the boundaries of {\it foreground wedge} determined by the horizon limit and antenna spacing.
+White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
 The features resemble those obtained with MWA data shown in figure~\ref{fig:fhd_data}.
@@
 Our foreground models are derived from other higher frequency catalogs and sky maps.
-The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observing frequency.
-Using simple error propagation, the fractional error in the delay spectrum caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the observing frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index.
+The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observed frequency.
+Using simple error propagation, the fractional error in the delay spectrum caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the observed frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index.
 From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for compact sources from NVSS and SUMSS catalogs.
-We assume the same holds true for our diffuse sky model as well, which is predominantly derived from 408~MHz map of \citet{has82}.
-Besides, delay spectra from simulations and data each have uncertainties arising from thermal noise with rms, $\Delta V^\textrm{N}_\eta(|\overline{\mathbf{x}}|,\eta) \sim 1.4\times 10^7$~Jy~Hz, where the superscript N stands for thermal noise.
-We estimate the ratio of delay spectra from data and simulations as, $\rho = |V^\textrm{D}_\eta(|\overline{\mathbf{x}}|,\eta)|/|V^\textrm{S}_\eta(|\overline{\mathbf{x}}|,\eta)|$, where superscripts D and S denote data and simulation respectively.
-Using the aforementioned uncertainties added in quadrature, we estimate the approximate expected uncertainty in $\log_{10}\rho$ and denote it by $\Delta\,\log_{10}\rho$.
+Although the model of \citet{deo08} yields a spectral index per direction on the sky, we assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from 408~MHz map of \citet{has82}.
+Thus, fractional errors in delay spectrum from compact foreground objects and diffuse emission are $\sim$0.7 and $\sim$0.3 respectively.
+In addition, delay spectra from simulations and data each have uncertainties due to thermal noise with rms, $\Delta V^\textrm{N}_\eta(|\overline{\mathbf{x}}|,\eta) \sim$ 1.4$\times 10^7$~Jy~Hz, where the superscript N stands for thermal noise.
+We estimate the ratio of delay spectra from data and simulations as, $\rho = |V^\textrm{D}_\eta(|\overline{\mathbf{x}}|,\eta)|\,/\,|V^\textrm{S}_\eta(|\overline{\mathbf{x}}|,\eta)|$, where superscripts D and S denote data and simulation respectively.
+% Using the aforementioned uncertainties added in quadrature, we estimate the approximate expected uncertainty in $\log_{10}\rho$ and denote it by $\Delta\,\log_{10}\rho$.
 
-Figure~\ref{fig:data-sim-ratio-snr} shows the significance of $\log_{10}\rho$ relative to the expected error $\Delta\,\log_{10}\rho$ for all $|\overline{\mathbf{x}}|$ and $\eta$ inside the {\it foreground wedge}.
+% Figure~\ref{fig:data-sim-ratio} shows the significance of $\log_{10}\rho$ relative to the expected error $\Delta\,\log_{10}\rho$ for all $|\overline{\mathbf{x}}|$ and $\eta$ inside the {\it foreground wedge}.
 The probability densities and cumulative distributions are shown in black and gray respectively.
@@
 
+Figure~\ref{fig:data-sim-ratio} shows histograms of distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}.
+The median absolute deviation from these distributions correspond to $\sim$~90\% fractional difference between data and modeling on average in either case.
+This is in line with aforementioned uncertainties in foreground models and thermal noise in measurements.
+Currently we are significantly limited by unavailability of all--sky foreground models at the observed frequency of 185~MHz.
+Availability of an accurate foreground model will improve the agreement of simulations with observed data.
+
 \begin{figure}[htb]
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
-\caption{Significance of logarithm of ratio of observed to simulated visibilities relative to expected uncertainty restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
+\caption{Histogram of logarithm of ratio of observed to simulated visibilities restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+Median absolute deviations of $\log_{10}\rho$ are 0.29 and 0.27 respectively.
+A value of zero indicates when the ratio $\rho=1$.
+\label{fig:data-sim-ratio}}
+% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
+% \caption{Significance of logarithm of ratio of observed to simulated visibilities relative to expected uncertainty restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
 Probability densities and cumulative distributions are shown in black and gray respectively.
@@
 Accounting for such uncertainties will only improve our confidence level in this comparison.
-Our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground window}.
+Our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.
 
@@
 
-\item {\it Sky Model}: Our model of the sky consists of bright compact sources and diffuse emission.
-Both of these kinds of emission are anisotropic and so is the resulting composite model.
-In fact, the pointing for MWA observations is chosen from regions of low foreground emission.
-In our study, we have divided the transiting sky into four different ``bow-tie'' shaped regions with equal areas and zenith as the origin.
+\item {\it Sky Model}: Our model of the sky consists of bright compact sources and diffuse emission on diverse spatial scales as seen from figures~\ref{fig:DSM} and \ref{fig:CSM}.
+It also consists of localized regions of strong emission such as the Galactic plane and Galactic center.
+The resulting sky emission is anisotropic.
+In fact, the patch of sky for MWA observations is chosen from regions of low foreground emission.
+In our detailed study, we have divided the transiting sky into four different ``bow-tie'' shaped regions with equal areas around zenith as the origin.
 This results in the ability to attribute the features in the delay spectrum to eight different sectors of the sky.
 
-\item {\it Baseline Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our interferometer baselines by their orientation ($\theta_\textrm{b}$) measured counter-clockwise from East.
+\item {\it Interferometer Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our interferometer antenna spacings by their orientation ($\theta_\textrm{b}$) measured counter-clockwise from East.
 We use the following bins: $-67\fdg 5\le \theta_\textrm{b} < -22\fdg 5$, $-22\fdg 5\le \theta_\textrm{b} < 22\fdg 5$, $22\fdg 5\le \theta_\textrm{b} < 67\fdg 5$, and $67\fdg 5\le \theta_\textrm{b} < 112\fdg 5$.
 The bin centers are oriented towards South-East, East, North-East, and North respectively.
+Since the delays registered at an interferometer depend on the orientation of its antenna spacing vector, binning by interferometer orientation allows us to attribute the delay spectrum to different sky directions.
 
-\item {\it Tile Pointing and Power Pattern}: Since the MWA tiles are phased electronically, the power pattern of the tile changes in any observing mode that tracks the source.
+\item {\it Tile Pointing and Power Pattern}: Since the MWA tiles are steered electronically, the power pattern of the tile changes in any observing mode that tracks the source.
 Our observations consist of a combination of allowing the sky to drift and tracking the sky.
-In our study, we take into account the effect of changing power pattern of the antennas on the delay spectrum.
-The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours (see figure~\ref{fig:pointings}).
-For comparisons, we choose two snapshot observations at LST of 22.08~hours (off--zenith pointing) and 0.09~hours (zenith pointing).
+In our study, we take into account the effect of changing power pattern of the antennas on the delay spectrum (figure~\ref{fig:power_pattern}).
+The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose {\it off--zenith} and {\it zenith} pointings shown in figure~\ref{fig:pointings}.
 
-\item {\it Instrumental Bandpass}: The instrumental bandpass characteristics have a direct effect on the delay transform owing to a Fourier relationship between the two.
-Due to excision of noisy edge channels repeated in each coarse channel of our bandpass, we see repeated patterns of the {\it foreground wedge} in the delay spectra.
+\item {\it Instrumental Passband}: The instrumental passband characteristics have a direct effect on the delay transform owing to a Fourier relationship between the two.
+Due to excision of noisy and aliased edge channels repeated in each coarse channel of our passband, we saw repeated patterns of the {\it foreground wedge} in the delay spectra.
 As already mentioned earlier, we have used the deconvolution algorithm ({\it CLEAN}) to rid the delay spectra of band shape effects.
@@
 
-The signatures we henceforth identify in the delay spectrum are characterized as arising out of the aforementioned factors.
+The foreground signatures we henceforth identify in the delay spectrum are characterized as arising out of the aforementioned factors.
+Due to numerous combinations of such factors, we use unique annotations to characterize these signatures.
+
+% The signatures we henceforth identify in the delay spectrum are characterized as arising out of the aforementioned factors.
 Due to numerous combinations of such factors, we use unique codes for these combinations as described in table~\ref{tab:codes}.
@@
 
-\begin{table*}[htp]
-\caption[]{Codes used to describe signatures in delay spectrum.\label{tab:codes}}
-\centering
-\begin{tabular}{ccc}
-\tableline\tableline\tableline
-Order & Parameter Description & Code \\
-\tableline\tableline
-1 & Snapshot index & 1, 2, \ldots \\
-2 & Type of emission & C (compact), D (diffuse), GC (Galactic center), GP (Galactic plane) \\
-3 & Direction of emission & E (east), N (north), W (west), S (south), Z (zenith)\\
-& & NE (north-east), NW (north-west), SW (south-west), SE (south-east) \\
-4 & Baseline orientation & E (east), S (south), SE (south-east), NE (north-east), A (all) \\
-5 & Region in antenna power pattern & P (primary beam), Sn ($n$-th sidelobe, $n=1,2,\ldots$)\\
-& (in the direction of emission) & \\
-\tableline
-\end{tabular}
-\end{table*}
+% \begin{table*}[htp]
+% \caption[]{Codes used to describe signatures in delay spectrum.\label{tab:codes}}
+% \centering
+% \begin{tabular}{ccc}
+% \tableline\tableline\tableline
+% Order & Parameter Description & Code \\
+% \tableline\tableline
+% 1 & Snapshot index & 1, 2, \ldots \\
+% 2 & Type of emission & C (compact), D (diffuse), GC (Galactic center), GP (Galactic plane) \\
+% 3 & Direction of emission & E (east), N (north), W (west), S (south), Z (zenith)\\
+% & & NE (north-east), NW (north-west), SW (south-west), SE (south-east) \\
+% 4 & Baseline orientation & E (east), S (south), SE (south-east), NE (north-east), A (all) \\
+% 5 & Region in antenna power pattern & P (primary beam), Sn ($n$-th sidelobe, $n=1,2,\ldots$)\\
+% & (in the direction of emission) & \\
+% \tableline
+% \end{tabular}
+% \end{table*}
 
-We wish to note that numerous features overlap at varying levels of significance as a result of various combiantions of parameters.
+It may be noted that numerous features overlap at varying levels of significance as a result of various combinations of parameters.
 We assign the features to their predominant causes.
@@
 With the addition of noise in the visibilities which is subject to observing time, some of the weaker features may not be as prominently visible.
+Since the foreground signatures are far too numerous and they reside in a vast volume of parameter space such as interferometer orientation, power pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features observed in the delay spectrum of foreground emission.
 
@@
 
-Figure~\ref{fig:dsm} shows the noiseless delay spectra due to diffuse emission.
+Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the amplitude of delay spectrum, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$, simulated with the diffuse foreground model.
+The top and bottom panels correspond to {\it off--zenith} and {\it zenith} pointings respectively.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
+\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
+White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
+The logarithmic color scale is common to both panels.
+In the {\it zenith} pointing, diffuse emission delay spectrum has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings.
+\label{fig:noiseless-dsm-delay-spectrum}}
+\end{figure}
+
+Here we examine in detail some examples of interesting features observed with the diffuse foreground model.
+
+\subsubsection{Galactic Center on Eastward Antenna Spacings}\label{sec:GC-east}
+
+The most prominent signature seen in the {\it off--zenith} pointing (top panel of figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the far west on one of the sidelobes of the power pattern.
+It appears as a bright branch in the delay spectrum near the negative delay horizon delay limit.
+This feature is strongest at short antenna spacings and fades with increase in antenna spacing.
+This Galactic center signature is almost invisible in the {\it zenith} pointing because the Galactic center has already set over the horizon (see bottom panel in figure~\ref{fig:noiseless-dsm-delay-spectrum}).
+
+% Figure~\ref{fig:dsm} shows the noiseless delay spectra due to diffuse emission.
 Figure~\ref{fig:dsm_delay_spectrum_snap0} and \ref{fig:dsm_delay_spectrum_snap1} correspond to off-zenith and on-zenith snapshots respectively.
@@
 
-The most prominent feature, however, is 1-GC-W-E-S3 in figure~\ref{fig:dsm_delay_spectrum_snap0}.
+% The most prominent feature, however, is 1-GC-W-E-S3 in figure~\ref{fig:dsm_delay_spectrum_snap0}.
 This is noted in the off-zenith snapshot caused by the bright Galactic center which is on the west at an altitude of $\sim 33$\arcdeg with $\sim 100$~minutes remaining before it sets.
-This emission is observed as it is coincident with the third sidelobe of the power pattern in the westward direction with a neagtive delay of value almost close to the horizon limit on eastward baselines.
+This emission is observed as it is coincident with the third sidelobe of the power pattern in the westward direction with a neagtive delay of value almost near the horizon limit on eastward baselines.
 
-\begin{figure*}[htp]
-\centering
-\subfloat[][Off-zenith delay spectra from diffuse foreground model]{\label{fig:dsm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_dsm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_0.eps}}
-\subfloat[][On-zenith delay spectra from diffuse foreground model]{\label{fig:dsm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_dsm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_1.eps}}\\
-\caption{Delay spectrum visibility amplitudes (in units of Jy~Hz) from diffuse foreground emission obtained on baselines of length $\lesssim 314.7$~m.
+% \begin{figure*}[htp]
+% \centering
+% \subfloat[][Off-zenith delay spectra from diffuse foreground model]{\label{fig:dsm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_dsm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_0.eps}}
+% \subfloat[][On-zenith delay spectra from diffuse foreground model]{\label{fig:dsm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_dsm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_1.eps}}\\
+% \caption{Delay spectrum visibility amplitudes (in units of Jy~Hz) from diffuse foreground emission obtained on baselines of length $\lesssim 314.7$~m.
 The left and right panels correspond to off-zenith and on-zenith snapshots respectively.
@@
 A logarithmic color scale is used.
-Most prominent features are due to the bright Galactic center shining through the third sidelobe (1-GC-W-E-S3), diffuse emission of short baselines at zero delays (2-D-Z-A-P), and diffuse emission on long baselines close to the horizon limit leaking in through the second sidelobes (1-GP-NE-NE-S2, 2-GP-N-N-S2, and 2-GP-S-N-S2).
+Most prominent features are due to the bright Galactic center shining through the third sidelobe (1-GC-W-E-S3), diffuse emission of short baselines at zero delays (2-D-Z-A-P), and diffuse emission on long baselines near the horizon limit leaking in through the second sidelobes (1-GP-NE-NE-S2, 2-GP-N-N-S2, and 2-GP-S-N-S2).
 Features from diffuse emission are predominantly edge-heavy resembling a two-pronged fork.
 See text for a detailed description of these features.}
-\label{fig:dsm}
-\end{figure*}
+% \label{fig:dsm}
+% \end{figure*}
 
-\subsubsection{Diffuse Emission On Long Baselines }\label{sec:fork}
+\subsubsection{Diffuse Emission}\label{sec:diffuse-features}
 
-Interestingly, in contrast to feature code 2-D-Z-A-P, all the other features in figures~\ref{fig:dsm_delay_spectrum_snap0} and \ref{fig:dsm_delay_spectrum_snap1} extend from the shortest ($\sim 7.7$~m) to longest baselines ($\sim 314.7$~m).
-We argue this is because even longer baselines are shortened when projected against directions along the baseline vector towards the horizon while considering an all-sky foreground model.
-Such a shortening due to projection effects allows diffuse emission from close to the horizon to be observed at maximum delays that correspond to the horizon limit even on long baselines.
+Diffuse emission from the main lobe of the power patterns manifests as a branch at the top with positive delays and a zero--delay horizontal branch in the {\it off--zenith} and {\it zenith} pointings respectively, seen in figure~\ref{fig:noiseless-dsm-delay-spectrum}.
+The former is seen at positive delays because the main lobe of the power pattern is centered eastward of zenith, whereas in the latter it is centered at zenith.
+The horizontal arm in the {\it zenith} pointing fades away on antenna spacings wider than $\gtrsim 125$~m indicating that the diffuse emission is devoid of spatial structures on scales $\lesssim$~0\fdg 75.
+
+\subsubsection{Diffuse Emission on Wide Antenna Spacings}\label{sec:diffuse-long-baselines}
+
+A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim \overline{\mathbf{x}}\le$~200~m) contrary to expectations, which are not expected to be sensitive to diffuse emission.
+We claim this is because wide antenna spacings appear shortened due to projection effects when sensitivity towards off--axis emission, especially towards the horizon, is concerned.
+It naturally holds for short antenna spacings as well.
+Thus, diffuse emission from far off--axis directions manifests as an edge--heavy {\it two--pronged fork} across all antenna spacings.
+It decreases in strength with increasing antenna spacing as expected but is present in all orientations of the antenna spacing vector.
+This contributes to a {\it pitchfork}--shaped feature which will be discussed later in \S\ref{sec:pitchfork} in more detail.
+
+% Interestingly, in contrast to feature code 2-D-Z-A-P, all the other features in figures~\ref{fig:dsm_delay_spectrum_snap0} and \ref{fig:dsm_delay_spectrum_snap1} extend from the shortest ($\sim 7.7$~m) to longest baselines ($\sim 314.7$~m).
+We argue this is because even wider baselines are shortened when projected against directions along the baseline vector towards the horizon while considering an all-sky foreground model.
+Such a shortening due to projection effects allows diffuse emission from near the horizon to be observed at maximum delays that correspond to the horizon limit even on long baselines.
 Feature 1-GP-NE-NE-S2 in figure~\ref{fig:dsm_delay_spectrum_snap0}, and features 2-GP-N-N-S2 and 2-GP-S-N-S2 in figure~\ref{fig:dsm_delay_spectrum_snap0} demonstrate this effect of shortening of baselines.
@@
 
-Thus, the most notable features from diffuse emission are those from primary beam area on short baselines, bright Galactic center from the third sidelobe, and the two-pronged fork.
+% Thus, the most notable features from diffuse emission are those from primary beam area on short baselines, bright Galactic center from the third sidelobe, and the two-pronged fork.
 
@@
 
-Figure~\ref{fig:csm} shows the noiseless delay spectra obtained from compact foreground objects.
+Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the delay spectrum arising out of our compact objects sky model for {\it off--zenith} and {\it zenith} pointings (shown in figure~\ref{fig:CSM}) in the top and bottom panels respectively.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
+\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise.
+White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
+The logarithmic color scale is identical to that in figure~\ref{fig:noiseless-dsm-delay-spectrum}.
+Emission from compact objects gives rise to centrally concentrated features in the {\it foreground wedge}.\label{fig:noiseless-csm-delay-spectrum}}
+\end{figure}
+
+In contrast to delay spectrum of diffuse emission, compact emission manifests as a center--heavy structure in the delay spectrum with either pointing.
+This contributes to the {\it pitchfork} feature which will be discussed in detail in \S\ref{sec:pitchfork}.
+Below we summarize some specific notable features obtained with our compact objects foreground model.
+
+\subsubsection{Compact Emission in Power Pattern Main Lobe}\label{sec:csm-main-lobe}
+
+Since the main lobe of the power pattern in the {\it off--zenith} pointing is centered eastward of zenith, bulk of the compact foreground emission is seen in a branch with positive delay but short of the positive horizon delay limit.
+Compact emission from the same patch of sky is seen as a bright horizontal arm at zero delay in the {\it zenith} pointing since the main lobe of the power pattern is centered at zenith.
+
+\subsubsection{Compact Emission in Power Pattern Sidelobes}\label{sec:csm-side-lobe}
+
+Foreground emission in zero delay and negative delay regions in the {\it off--zenith} pointing is caused by compact objects co--located with sidelobes of the power pattern.
+On the other hand, compact objects co--located with sidelobes of power pattern in the {\it zenith} pointing are revealed as distinct but faint branches at positive and negative delays depending on the orientation of antenna spacing and direction of emission on the sky.
+
+% Figure~\ref{fig:csm} shows the noiseless delay spectra obtained from compact foreground objects.
 The off-zenith and on-zenith snapshot delay spectra are shown in figures~\ref{fig:csm_delay_spectrum_snap0} and \ref{fig:csm_delay_spectrum_snap1} respectively.
 
-In the off-zenith snapshot, the most prominent feature 1-C-E-E-P arises due to eastward baselines observing the eastward sky at positive delays centered on the primary beam.
+% In the off-zenith snapshot, the most prominent feature 1-C-E-E-P arises due to eastward baselines observing the eastward sky at positive delays centered on the primary beam.
 The next prominent feature 1-C-E-N-P at zero delays arises due to the compact emission from eastward sky centered on the primary beam observed by northward baselines.
@@
 
-The most prominent feature in the on-zenith snapshot is 2-C-Z-A-P appearing at zero delays caused by compact sources at zenith observed by all baselines along their primary beams.
+% The most prominent feature in the on-zenith snapshot is 2-C-Z-A-P appearing at zero delays caused by compact sources at zenith observed by all baselines along their primary beams.
 First and second sidelobes along north and south observed by northward baselines contribute to symmetric positive and negative delays as demonstrated by features 2-C-S-N-S1, 2-C-N-N-S1, 2-C-S-N-S2, and 2-C-N-N-S2.
 
-\begin{figure*}[htp]
-\centering
-\subfloat[][Off-zenith delay spectra from compact foreground objects]{\label{fig:csm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_csm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_0.eps}}
-\subfloat[][On-zenith delay spectra from compact foreground objects]{\label{fig:csm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_csm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_1.eps}}\\
-\caption{Same as figure~\ref{fig:dsm} but for compact foreground objects.
+% \begin{figure*}[htp]
+% \centering
+% \subfloat[][Off-zenith delay spectra from compact foreground objects]{\label{fig:csm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_csm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_0.eps}}
+% \subfloat[][On-zenith delay spectra from compact foreground objects]{\label{fig:csm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_csm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_1.eps}}\\
+% \caption{Same as figure~\ref{fig:dsm} but for compact foreground objects.
 Color scale is the same as in figure~\ref{fig:dsm}.
@@
 See text for a detailed description of the different features.}
-\label{fig:csm}
-\end{figure*}
+% \label{fig:csm}
+% \end{figure*}
 
-While compact foreground objects give rise to features throughout the foreground wedge, the primary distinction between delay spectra of compact and diffuse foregrounds is that the latter are predominantly center-heavy when compared to edge-heavy features of the former.
+% While compact foreground objects give rise to features throughout the foreground wedge, the primary distinction between delay spectra of compact and diffuse foregrounds is that the latter are predominantly center-heavy when compared to edge-heavy features of the former.
 
-% Figure~\ref{fig:obs-begin} shows the delay spectra obtained with the compact foreground model described in \S\ref{sec:CSM} at the beginning of the observation when the MWA tiles are phased to point at RA = 4\fdg 387, Dec = -29\fdg 86.
+% % Figure~\ref{fig:obs-begin} shows the delay spectra obtained with the compact foreground model described in \S\ref{sec:CSM} at the beginning of the observation when the MWA tiles are phased to point at RA = 4\fdg 387, Dec = -29\fdg 86.
 The contours of the antenna power pattern are shown in the central panel.
@@
 
-% \begin{figure*}[htb]
-% \centering
-% % \subfloat[][Delay spectra at the beginning of observation]{\label{fig:SNR1D_6hrs_167_single_e0}\includegraphics[width=\linewidth]{}}
-% \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_CLEAN_visibilities_custom_gaussian_FG_model_csm_all_sky_nside_128_30.7_MHz_bnw2.0_snapshot_0.eps}
-% \caption{Delay spectra, compact foreground sources, and delay maps for a snapshot at the beginning of the observation pointed at RA = 4\fdg 387, Dec = -29\fdg 86 at LST = 22.08 hours.
+% % \begin{figure*}[htb]
+% % \centering
+% % % \subfloat[][Delay spectra at the beginning of observation]{\label{fig:SNR1D_6hrs_167_single_e0}\includegraphics[width=\linewidth]{}}
+% % \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_CLEAN_visibilities_custom_gaussian_FG_model_csm_all_sky_nside_128_30.7_MHz_bnw2.0_snapshot_0.eps}
+% % \caption{Delay spectra, compact foreground sources, and delay maps for a snapshot at the beginning of the observation pointed at RA = 4\fdg 387, Dec = -29\fdg 86 at LST = 22.08 hours.
 The phase center is at RA = -28\fdg 8, Dec = -26\fdg 701.
@@
 The contour levels shown are: 0.0078125, 0.03125, 0.125, and 0.5.}
-% \label{fig:obs-begin}
-% \end{figure*}
+% % \label{fig:obs-begin}
+% % \end{figure*}
 
-% \begin{figure*}[htb]
-% \centering
-% % \subfloat[][Delay spectra at the beginning of observation]{\label{fig:SNR1D_6hrs_167_single_e0}\includegraphics[width=\linewidth]{}}
-% \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_CLEAN_visibilities_custom_gaussian_FG_model_csm_all_sky_nside_128_30.7_MHz_bnw2.0_snapshot_1.eps}
-% \caption{Same as figure~\ref{fig:obs-begin} but for a snapshot during the middle of the observation pointed at RA = +1\fdg 320, Dec = -26\fdg 70 at LST = 0.09 hours.
+% % \begin{figure*}[htb]
+% % \centering
+% % % \subfloat[][Delay spectra at the beginning of observation]{\label{fig:SNR1D_6hrs_167_single_e0}\includegraphics[width=\linewidth]{}}
+% % \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_CLEAN_visibilities_custom_gaussian_FG_model_csm_all_sky_nside_128_30.7_MHz_bnw2.0_snapshot_1.eps}
+% % \caption{Same as figure~\ref{fig:obs-begin} but for a snapshot during the middle of the observation pointed at RA = +1\fdg 320, Dec = -26\fdg 70 at LST = 0.09 hours.
 The phase center is at RA = 1\fdg 35, Dec = -26\fdg 701.
 }
-% \label{fig:obs-middle}
-% \end{figure*}
+% % \label{fig:obs-middle}
+% % \end{figure*}
 
-\subsection{All-Sky Composite Foreground Model}\label{sec:composite}
+\subsection{All--Sky Composite Foreground Model}\label{sec:composite}
 
 Delay spectra from the all-sky foreground model in our study display a composite feature set drawn from the features of compact and diffuse foreground models.
-Figures~\ref{fig:asm_delay_spectrum_snap0} and \ref{fig:asm_delay_spectrum_snap1} show the delay spectra for the composite all-sky foreground model.
+Here we compare the relative strengths of emission from different spatial scales in our all--sky composite foreground model.
+
+\subsubsection{{\it Pitchfork} Effect}\label{sec:pitchfork}
+
+Excluding the bright emission from the Galactic center which is heavily dependent on the sky location pointed to by the antenna, delay spectrum resulting from the all--sky composite foreground model consists of signatures already noted in the diffuse and compact foreground cases.
+This is illustrated by a microscopic examination of the {\it zenith} pointing in our study.
+
+Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three interferometers of different antenna spacings oriented northward during the {\it zenith} pointing.
+The diffuse, compact, and composite components are are shown as dotted, dashed, and solid lines respectively.
+The horizon delay limits are shown as vertical dot--dashed lines.
+The gray shaded area denotes the envelope of expected uncertainty in the delay spectrum.
+This uncertainty is dominated by the spread in spectral index of compact foreground objects inside the {\it foreground} wedge (between the horizon delay limits) proprotional to the delay spectrum from compact objects, and by thermal noise outside the horizon delay limits.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
+\caption{Absolute value of delay spectrum, $|V_\eta(\overline{\mathbf{x}},\eta)|$, for three chosen antenna spacings, $\sim$84~m ({\it top}), $\sim$105~m ({\it middle}), and $\sim$170~m ({\it bottom}).
+The antenna spacing vector is specified in each panel along East, North and Upward directions.
+The dotted, dashed and solid lines denote contributions from diffuse emission, compact objects and composite foreground model respectively.
+The dot--dashed lines mark the horizon delay limits.
+The shaded region denotes the envelope of uncertainty around expected values for composite model.
+This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits.
+Outside the horizon delay limits, it is predominantly due to thermal noise.
+Compact objects dominate the central regions of power spectrum while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a {\it pitchfork}--shaped structure.
+\label{fig:pitchfork-baselines}}
+\end{figure}
+
+The zero--delay peak (corresponding to the main lobe in the power pattern) with a value of $\sim 10^8$--$10^9$~Jy~Hz, independent of antenna spacing, is predominantly determined by compact foreground objects.
+The corresponding peak at zero--delay from diffuse foreground model is at least 10 times fainter and decreases drastically with increase in antenna spacing.
+This is the response expected from different antenna spacings towards compact and diffuse foreground objects.
+
+Near the horizon delay limits at shorter antenna spacing, diffuse emission is brighter than that from compact objects.
+With increasing antenna spacing, although diffuse and compact components of emission become comparable, diffuse emission does not decrease as rapidly as seen at zero--delay.
+In fact, even with widely spaced antennas, diffuse emission in the delay spectrum near the horizon delay limits exceeds that in the main lobe by more than two orders of magnitude.
+This is due to apparent shortening of antenna spacing owing to projection effects while receiving emission from wide off--axis directions.
+This demonstrates the edge--heavy {\it two--pronged fork} feature discussed in \S\ref{sec:diffuse-long-baselines}.
+
+The resulting delay spectrum from a composite foreground model that combines diffuse and compact object emission consists of center--heavy features on all antenna spacings from compact objects and edge--heavy features from both components especially diffuse emission even on wide antenna spacings.
+Henceforth, we call this as the {\it pitchfork} structure in the {\it foreground} wedge.
+
+All these observations are consistent with those from figures~\ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}.
+The observability of the predicted {\it pitchfork} signature depends on the relative levels of uncertainty in the foreground model and thermal noise.
+In our simulations, since thermal noise is $\gtrsim 10^7$~Jy~Hz, and features near the horizon delay limits are $\sim 10^7$~Jy~Hz, the {\it pitchfork} feature is not expected to be detected in a noisy scenario although this feature is marginally visible in the {\it zenith} pointing of the observed data (see figure~\ref{fig:fhd_data}).
+We attribute this to differences between our foreground model and the actual sky.
+However, with improvements to the foreground model and instrument configuration that can reduce thermal noise and other systematic uncertainties, we are confident the {\it pitchfork} signature in delay spectrum will be observed with high significance.
+
+We also note that increasing the antenna spacing results in progressively improving the resolution along delay axis by increasing the number of delay bins inside the {\it foreground wedge}.
+This improves the localization of foreground objects whose signatures are imprinted in the delay spectrum.
+For instance, there is an increase in the number of secondary peaks in the delay spectrum between zero--delay and horizon delay limits as antenna spacing increases from $\sim$84~m to $\sim$170~m.
+In this case, these correspond to sidelobes of the power pattern that lie between the main lobe and the horizon along the North--South direction.
+In interferometers with short spacings, due to relatively lower resolution along delay axis inside the {\it foreground wedge} and a consequent loss of localization of direction of foreground emission, these secondary peaks blend in with other major peaks and are not distinctly visible.
+These findings are consistent with dicussion in \S\ref{sec:csm-side-lobe}.
+
+In summary, the brightest signature ($\gtrsim 10^{10}$~Jy~Hz) is that of the Galactic center in the {\it off--zenith} pointing co--located at a westward sidelobe with a significantly high gain.
+The next brightest signature ($\sim 10^8$--$\sim 10^9$~Jy~Hz) is caused by compact emission appearing to be concentrated in the inner regions of the foreground wedge (near zero delays) rather than towards the edges.
+Diffuse emission co--located at the main lobe of the power pattern is fainter by approximately an order of magnitude relative to compact emission from the same region for a $\sim 84$~m antenna spacing but unlike the latter, diffuse emission decreases rapidly by over two orders of magnitude as antenna spacing is widened to $\sim 170$~m.
+However, diffuse emission is significantly high near horizon delay limits compared to main lobe of the power pattern as antenna spacing is widened.
+This leads to edge--heavy features in the delay spectrum.
+Complemented by the center-heavy compact foreground features, we have demosntrated the presence of a {\it three--pronged pitchfork}--shaped signature in the delay spectrum of foreground sky.
+
+% Figures~\ref{fig:asm_delay_spectrum_snap0} and \ref{fig:asm_delay_spectrum_snap1} show the delay spectra for the composite all-sky foreground model.
 While features from both its constituents occupy and overlap in different regions of the foreground wedge, their relatives magnitudes determine the resultant signatures seen in the delay spectra.
@@
 
-On the other hand, in central and inner regions of the wedge, clearly the center-heavy compact foreground model features dominate by $\gtrsim 2$ orders of magnitude relative to features arising out of diffuse emission.
+% On the other hand, in central and inner regions of the wedge, clearly the center-heavy compact foreground model features dominate by $\gtrsim 2$ orders of magnitude relative to features arising out of diffuse emission.
 This is quite evident since almost all the diffuse emission features are confined to the edges of the wedge.
@@
 
-Figures~\ref{fig:baseline_composite_asm_delay_spectrum_snap0} and \ref{fig:baseline_composite_asm_delay_spectrum_snap1} show the delay spectra from the diffuse and compact foreground models as well as from the composite model for the off-zenith and on-zenith snapshots respectively for a selected baseline labeled as ``155-154''.
+% Figures~\ref{fig:baseline_composite_asm_delay_spectrum_snap0} and \ref{fig:baseline_composite_asm_delay_spectrum_snap1} show the delay spectra from the diffuse and compact foreground models as well as from the composite model for the off-zenith and on-zenith snapshots respectively for a selected baseline labeled as ``155-154''.
 This baseline is of length $\simeq 201$~m and orientation $\theta_b\simeq 13\fdg 7$.
@@
 
-\begin{figure*}[htp]
-\centering
-\subfloat[][Off-zenith delay spectra from composite foreground model]{\label{fig:asm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_asm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_0.eps}}
-\subfloat[][Off-zenith delay spectra from composite foreground model]{\label{fig:asm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_asm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_1.eps}}\\
-\subfloat[][Off-zenith composite delay spectrum on eastward baseline]{\label{fig:baseline_composite_asm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/baseline_155-154_composite_noiseless_delay_spectrum_snapshot_0.edited.eps}}
-\subfloat[][On-zenith composite delay spectrum on eastward baseline]{\label{fig:baseline_composite_asm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/baseline_155-154_composite_noiseless_delay_spectrum_snapshot_1.edited.eps}}\\
-\caption{{\it Top}: Same as figure~\ref{fig:dsm} but for a composite all-sky foreground model consisting of both diffuse emission and compact objects.
+% \begin{figure*}[htp]
+% \centering
+% \subfloat[][Off-zenith delay spectra from composite foreground model]{\label{fig:asm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_asm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_0.eps}}
+% \subfloat[][Off-zenith delay spectra from composite foreground model]{\label{fig:asm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_asm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_1.eps}}\\
+% \subfloat[][Off-zenith composite delay spectrum on eastward baseline]{\label{fig:baseline_composite_asm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/baseline_155-154_composite_noiseless_delay_spectrum_snapshot_0.edited.eps}}
+% \subfloat[][On-zenith composite delay spectrum on eastward baseline]{\label{fig:baseline_composite_asm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/baseline_155-154_composite_noiseless_delay_spectrum_snapshot_1.edited.eps}}\\
+% \caption{{\it Top}: Same as figure~\ref{fig:dsm} but for a composite all-sky foreground model consisting of both diffuse emission and compact objects.
 The color scale used is the same.
@@
 Refer to text for detailed description of these features.}
-\label{fig:asm}
-\end{figure*}
+% \label{fig:asm}
+% \end{figure*}
 
-Thus the composite all-sky foreground model results in a delay spectrum structure that captures both the edge-heavy two-pronged fork structure from diffuse foreground emission and the center-heavy features from compact foreground objects.
+% Thus the composite all-sky foreground model results in a delay spectrum structure that captures both the edge-heavy two-pronged fork structure from diffuse foreground emission and the center-heavy features from compact foreground objects.
 Hereafter, we refer to this structure as a ``three-pronged pitchfork''.
 
-\section{Foreground Grading Diagnostic For Observations}\label{sec:fg-grading}
+\section{Screening of Severe Foreground Contaminated Interferometers}\label{sec:fg-grading}
+
+The delay spectrum, $V_\eta(\overline{\mathbf{x}},\eta)$, not only carries information on spatial scales of foreground emission but also offers a unique advantage of viewing the sky through a combination of antenna spacing vector and the delay axis.
+This allows us to programmatically screen data for antenna spacings that are severly contaminated by foregrounds which can be weighted appropriately during data analysis.
+While the {\it foreground wedge} is a region of high foreground contamination, of particular interest to the EoR community is foreground contamination near horizon delay limits.
+These regions are responsible for significant spillover into the relatively clean {\it EoR window} due to spectral response of the instrument and foregrounds.
+
+With a foreground model known {\it a priori} in which locations of very bright foreground objects such as the Galactic center or AGN are available, we can predict their response across antenna spacings as a function of observing parameters such as LST, power pattern, etc.
+Specifically, we can pick interferometers where these bright foreground objects will be visible at delays near horizon delay limits.
+One such example is already provided in our study.
+The Galactic center is the most dominant source of foreground contamination in the {\it off--zenith} pointing.
+Since it is located in the western sky, it is most distinctly seen near the negative horizon delay limit on eastward oriented antenna spacings in the delay spectrum (see figures~\ref{fig:fhd_data}, \ref{fig:sim_data} and \ref{fig:noiseless-dsm-delay-spectrum}, and \S\ref{sec:GC-east}).
+In fact these figures illustrate that deconvolution along delay axis is severly hampered by this extreme contamination, thereby leading to leakage beyond the {\it foreground wedge} into the {\it EoR window}.
+In practice, we could downweight these antenna spacings for the specific LST to mitigate the effects of spillover from such foreground contamination in EoR studies.
+With spatial scales of such objects known, besides downweighting based on antenna spacing orientation, we could also downweight based on antenna spacing distance.
+For instance, a bright compact object will excite interferometers uniformly across antenna spacings whereas an extended object will gradually fade with increase in antenna spacing.
+
+This provides us with a very simple and yet effective tool in adding a layer of sophistication to mitigate effects of foreground contamination in EoR data analysis.
+% The success behind clear identification of severely foreground--contaminated interferometers based on delay spectrum technique depends on the localization of the source of foreground contamination and its strength.
 
@@
 Our primary motivation in this work is to analyze the signatures of foreground components in the measured delay spectrum.
-Such an analysis will be helpful in extending our knowledge on the challenge posed by foregrounds in studying the EoR and in devising techniques to mitigate EoR signal contamination from foregrounds.
+Such an analysis will be helpful in extending our knowledge on the challenge posed by foregrounds in studying the EoR and in devising techniques to mitigate contamination of EoR H{\sc i} signal by foreground objects emitting in radio wavelengths.
 This is one of the first studies that uses an all-sky foreground model to study the resultant signatures on EoR delay spectrum.
 
-Using parameters that match the instrument and EoR observations using the MWA, and an all-sky foreground model that consists of diffuse emission on scales $\gtrsim 0\fdg 85$ and bright compact sources from the NVSS and SUMSS catalogs, we model delay spectra obtained with the MWA on each of its baselines of length $\lesssim 315$~m.
-We confirm that the modeled delay spectra match the data obtained with the MWA.
-A wedge shaped structure is clearly noted as predicted by previous studies.
+Using parameters that match the instrument and EoR observations using the MWA, and an all-sky foreground model that consists of diffuse emission from a ``global sky model'' and bright compact sources from the NVSS and SUMSS catalogs, we model delay spectra obtained with the MWA on antenna spacings of length $\le 200$~m.
+We confirm that the modeled delay spectra is in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise in measurements.
 
-We use our simulations in a noiseless scenario to understand many of the foreground signatures seen in the delay spectra.
-We find that the delay spectra depend on observing parameters such as antenna pointing and LST, instrument parameters such as antenna power pattern and bandpass shape, and foreground parameters such as the nature of emission, spectral index, etc.
-A {\it CLEAN} deconvolution in delay-space was applied to remove the contaminating harmonic effects of frequency flagging.
+We use our simulations in a noiseless scenario to identify numerous signatures of different compoenents of foreground emission seen in the delay spectra.
+We establish the relationship between these signatures and observing parameters such as antenna pointing and LST, instrument parameters such as antenna power pattern and bandpass shape, and foreground parameters such as the nature of emission, spectral index, etc.
 
-Notable features caused by diffuse emission include the bright Galactic center at the edge of the horizon to be observed through the far sidelobe of the antenna power pattern.
-As expected, diffuse emission in the primary beam of the antenna power pattern is present on short baselines of length $\lesssim 60$~m.
-Quite surprisingly, diffuse emission also leaves its footprint on long baselines on either side of the delay axis at the horizon delay limits.
-This clearly indicates that long baselines appear to be shortened in the direction of the horizon along the baseline vector and these short projected baselines are responsive to large scale diffuse emission.
-Thus, one of the most important findings of this study is the response of even long baselines to diffuse structures.
-The characteristic signature of diffuse emission appears to be a roughly symmetric edge-heavy feature taking the form of a {\it two-pronged fork}.
+The bright Galactic center at the edge of the western horizon co--located with one of the far sidelobes of MWA tile power pattern is the brightest source of foreground contamination in the {\it off--zenith} pointing.
+It manifests itself near negative horizon delay limit in the delay spectrum on antenna spacings oriented eastward.
 
-On the other hand, compact foreground objects predominantly map onto inner regions of the wedge.
-Features arising from compact sources coincident with primary beam and sidelobes of antenna power pattern are clearly noted.
-In general, compact objects produce delay spectrum signatures that are center-heavy, in stark contrast to those from diffuse foregrounds.
+On expected lines, diffuse emission in the primary beam of the antenna power pattern is prominent on shorter antenna spacings $\lesssim$~125~m corresponding to angular scales $\gtrsim$~0\fdg 75.
+However, the most interesting result is its footprint on wide antenna spacings near the horizon delay limits.
+This results in a edge--heavy {\it two--pronged fork}--shaped signature.
+This is due to apparent shortening of wide antenna spacings in the direction of foreground emission far off--axis thereby retaining their response to extended emission.
+On the other hand, compact foreground objects predominantly map onto central regions of the {\it foreground wedge}.
+Features arising from compact sources coincident with primary beam and sidelobes of antenna power pattern have been noted.
+In general, compact objects produce delay spectrum signatures that are center--heavy, in clear contrast to those from diffuse emission.
+A composite all--sky foreground model consisting of diffuse and compact foregrounds combines the characteristic individual shapes into a {\it pitchfork}--shaped structure in the {\it foreground wedge}.
+This will be distinctly visible when the thermal noise floor is sufficiently lowered.
 
-A composite all-sky foreground model consisting of diffuse and compact foregrounds gives rise to a combination of edge-heavy fork shaped features from diffuse foregrounds and center-heavy features from compact foreground objects in the delay spectrum.
-The compact foreground features are brighter by $\gtrsim 3$ orders of magnitude brighter than the diffuse foreground features in the central regions of delay spectrum.
-On the other hand, the edge-heavy diffuse foreground features in the delay spectrum are $\gtrsim 1$ order of magnitude brighter than that due to compact foregrounds at the edges of delay spectrum.
-This gives the appearance of a {\it pitchfork} shape to the resultant delay spectrum.
+We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination in EoR data analysis.
+It is possible to screen interferometers most affected by foreground contamination as a function of antenna spacing vector and time of observation, which can then be assigned optimal weights in the EoR data analysis chain.
 

### 2014-09-03 foreground_delay_spectrum_modeling.tex
more edits in intro

--- 2014-09-03 foreground_delay_spectrum_modeling.tex
@@
 
- Matching real observations from the Murchison Widefield Array Epoch of Reionization observing program with instrumental simulations of the entire known sky we are examine the signature of the foreground power spectrum on each baseline.
-We discover that the foregrounds most likely to obfuscate the Epoch of Reionization power are not caused by power emitted from the same sky location as the target HI signal, and that smooth galactic emission plays a larger role then previously thought.
+ Matching real observations from the Murchison Widefield Array Epoch of Reionization observing program with instrumental simulations of the entire known sky we examine the signature of the power spectrum of each baseline.
+We confirm that the foregrounds most likely to obfuscate the Epoch of Reionization power are not caused by emission central field of view but rather from near the horizon.
+We also find that the smooth galactic emission plays a larger role then previously thought.
 We quantify this contamination by breaking the simulations down by sky region and type of emission.

### 2014-09-03 foreground_delay_spectrum_modeling.tex
From v0.3.1: ===========
(*) Added the full author list

--- 2014-09-03 foreground_delay_spectrum_modeling.tex
@@
 \def\ASU{\altaffilmark{1}}
-\def\ASUtxt{\altaffiltext{1}{Arizona State University, Tempe, AZ 85287, USA; e-mail: t\_nithyanandan@asu.edu}}
+\def\ASUtxt{\altaffiltext{1}{School of Earth and Space Exploration, Arizona State University, Tempe, AZ 85287, USA; e-mail: t\_nithyanandan@asu.edu}}
 
-% \def\RRI{\altaffilmark{1}}
-% %% \def\RRItxt{\altaffiltext{1}{Raman Research Institute, Bangalore, India}}
-% \def\RRItxt{\altaffiltext{1}{Raman Research Institute, Bangalore, India; e-mail: nithya.rri@gmail.com}}
+\def\UW{\altaffilmark{2}}
+\def\UWtxt{\altaffiltext{2}{Department of Physics, University of Washington, Seattle, WA 98195, USA}}
 
-% \def\NRAONM{\altaffilmark{2}}
-% \def\NRAONMtxt{\altaffiltext{2}{National Radio Astronomy Observatory, Socorro, NM, USA}}
+\def\RRI{\altaffilmark{3}}
+\def\RRItxt{\altaffiltext{3}{Raman Research Institute, Bangalore 560080, India}}
 
-% \def\CAASTRO{\altaffilmark{3}}
-% \def\CAASTROtxt{\altaffiltext{3}{ARC Centre of Excellence for All-sky Astrophysics (CAASTRO), Sydney, Australia}}
+\def\SKASA{\altaffilmark{4}}
+\def\SKASAtxt{\altaffiltext{4}{Square Kilometre Array South Africa (SKA SA), Cape Town 7405, South Africa}}
 
-% \def\Curtin{\altaffilmark{4}}
-% \def\Curtintxt{\altaffiltext{4}{Curtin University, Perth, Australia}}
+\def\ANU{\altaffilmark{5}}
+\def\ANUtxt{\altaffiltext{5}{Research School of Astronomy and Astrophysics, Australian National University, Canberra, ACT 2611, Australia}}
 
-% %% \def\Swinburne{\altaffilmark{4}}
-% %% \def\Swinburnetxt{\altaffiltext{4}{Swinburne University of Technology, Melbourne, Australia}}
+\def\Haystack{\altaffilmark{6}}
+\def\Haystacktxt{\altaffiltext{6}{MIT Haystack Observatory, Westford, MA 01886, USA}}
 
-% \def\CfA{\altaffilmark{5}}
-% \def\CfAtxt{\altaffiltext{5}{Harvard-Smithsonian Center for Astrophysics, Cambridge, USA}}
+\def\Curtin{\altaffilmark{7}}
+\def\Curtintxt{\altaffiltext{7}{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}}
 
-% \def\ANU{\altaffilmark{7}}
-% \def\ANUtxt{\altaffiltext{7}{The Australian National University, Canberra, Australia}}
+\def\USydney{\altaffilmark{8}}
+\def\USydneytxt{\altaffiltext{8}{Sydney Institute for Astronomy, School of Physics, The University of Sydney, NSW 2006, Australia}}
 
-% \def\CSIRO{\altaffilmark{8}}
-% \def\CSIROtxt{\altaffiltext{8}{CSIRO Astronomy and Space Science, NSW, Australia}}
+\def\CAASTRO{\altaffilmark{9}}
+\def\CAASTROtxt{\altaffiltext{9}{ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}}
 
-% \def\Haystack{\altaffilmark{9}}
-% \def\Haystacktxt{\altaffiltext{9}{MIT Haystack Observatory, Westford, USA}}
+\def\MIT{\altaffilmark{10}}
+\def\MITtxt{\altaffiltext{10}{Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}}
 
-% \def\USydney{\altaffilmark{10}}
-% \def\USydneytxt{\altaffiltext{10}{Sydney Institute for Astronomy, School of Physics, The University of Sydney, Sydney, Australia}}
+\def\CfA{\altaffilmark{11}}
+\def\CfAtxt{\altaffiltext{11}{Harvard-Smithsonian Center for Astrophysics, Cambridge, MA 02138, USA}}
 
-% \def\MIT{\altaffilmark{11}}
-% \def\MITtxt{\altaffiltext{11}{MIT Kavli Institute for Astrophysics and Space Research, Cambridge, USA}}
+\def\Victoria{\altaffilmark{12}}
+\def\Victoriatxt{\altaffiltext{12}{School of Chemical \& Physical Sciences, Victoria University of Wellington, Wellington 6140, New Zealand}}
 
-% \def\UW{\altaffilmark{12}}
-% \def\UWtxt{\altaffiltext{12}{University of Washington, Seattle, USA}}
+\def\UWisc{\altaffilmark{13}}
+\def\UWisctxt{\altaffiltext{13}{Department of Physics, University of Wisconsin--Milwaukee, Milwaukee, WI 53201, USA}}
 
-% \def\Victoria{\altaffilmark{13}}
-% \def\Victoriatxt{\altaffiltext{13}{Victoria University of Wellington, Wellington, New Zealand}}
+\def\UMichigan{\altaffilmark{14}}
+\def\UMichigantxt{\altaffiltext{14}{Department of Atmospheric, Oceanic and Space Sciences, University of Michigan, Ann Arbor, MI 48109, USA}}
 
-% \def\UWisc{\altaffilmark{14}}
-% \def\UWisctxt{\altaffiltext{14}{University of Wisconsin--Milwaukee, Milwaukee, USA}}
+\def\CASS{\altaffilmark{15}}
+\def\CASStxt{\altaffiltext{15}{CSIRO Astronomy and Space Science (CASS), PO Box 76, Epping, NSW 1710, Australia}}
 
-% \def\UMelbourne{\altaffilmark{15}}
-% \def\UMelbournetxt{\altaffiltext{15}{The University of Melbourne, Melbourne, Australia}}
+\def\Tata{\altaffilmark{16}}
+\def\Tatatxt{\altaffiltext{16}{National Centre for Radio Astrophysics, Tata Institute for Fundamental Research, Pune 411007, India}}
 
-% \def\Tata{\altaffilmark{16}}
-% \def\Tatatxt{\altaffiltext{16}{Tata Institute of Fundamental Research, Pune, India}}
+\def\NRAO{\altaffilmark{17}}
+\def\NRAOtxt{\altaffiltext{17}{National Radio Astronomy Observatory, Charlottesville and Greenbank, USA}}
 
-% \def\NRAOGBT{\altaffilmark{17}}
-% \def\NRAOGBTtxt{\altaffiltext{17}{National Radio Astronomy Observatory, Charlottesville, USA}}
-
-% \def\UTasmania{\altaffilmark{18}}
-% \def\UTasmaniatxt{\altaffiltext{18}{University of Tasmania, Hobart, Australia}}
-
-% \def\UWA{\altaffilmark{19}}
-% \def\UWAtxt{\altaffiltext{19}{University of Western Australia, Perth, Australia}}
-
-% %% \def\ICRAR{\altaffilmark{9}}
-% %% \def\ICRARtxt{\altaffiltext{9}{International Centre for Radio Astronomy Research, Curtin University, Perth, Australia }}
-
-% %\def\VUW{\altaffilmark{13}}
-% %\def\VUWtxt{\altaffiltext{13}{Victoria University of Wellington, New Zealand}}
+\def\UMelbourne{\altaffilmark{18}}
+\def\UMelbournetxt{\altaffiltext{18}{School of Physics, The University of Melbourne, Parkville, VIC 3010, Australia}}
 
@@
 
-%% \author{NITHYANANDAN~THYAGARAJAN\altaffilmark{1}, 
-%%   N.~UDAYA~SHANKAR\altaffilmark{1}, 
-%%   \and RAVI~SUBRAHMANYAN\altaffilmark{1}}
-
 %% Author list
 \author{
-%% Lead authors
+%% Lead Authors
 Nithyanandan~Thyagarajan\ASU,
 Daniel~C.~Jacobs\ASU,
-Judd~D.~Bowman\ASU
+Judd~D.~Bowman\ASU,
+%% Supporting authors
+A.~P.~Beardsley\UW,
+B.~J.~Hazelton\UW, 
+M.~F.~Morales\UW, 
+J.~Pober\UW,
+I.~S.~Sullivan\UW,
+T.~Prabu\RRI, 
+%% Builders' list
+G.~Bernardi\SKASA,
+F.~Briggs\ANU,
+R.~J.~Cappallo\Haystack, 
+B.~E.~Corey\Haystack, 
+A.~A.~Deshpande\RRI, 
+D.~Emrich\Curtin,
+B.~M.~Gaensler\USydney$^,$\CAASTRO, 
+R.~Goeke\MIT,
+L.~J.~Greenhill\CfA,
+M.~Johnston-Hollitt\Victoria,
+D.~L.~Kaplan\UWisc, 
+J.~C.~Kasper\UMichigan$^,$\CfA, 
+E.~Kratzenberg\Haystack, 
+C.~J.~Lonsdale\Haystack, 
+M.~J.~Lynch\Curtin, 
+S.~R.~McWhirter\Haystack,
+D.~A.~Mitchell\CASS$^,$\CAASTRO, 
+E.~Morgan\MIT, 
+D.~Oberoi\Tata, 
+S.~M.~Ord\Curtin$^,$\CAASTRO,
+A.~E.~E.~Rogers\Haystack, 
+A.~Roshi\NRAO, 
+N.~Udaya~Shankar\RRI, 
+K.~S.~Srivani\RRI, 
+R.~Subrahmanyan\RRI$^,$\CAASTRO, 
+S.~J.~Tingay\Curtin$^,$\CAASTRO, 
+M.~Waterson\Curtin$^,$\ANU,
+R.~B.~Wayth\Curtin$^,$\CAASTRO, 
+R.~L.~Webster\UMelbourne$^,$\CAASTRO, 
+A.~R.~Whitney\Haystack, 
+A.~Williams\Curtin, 
+C.~L.~Williams\MIT
 }
@@
 \ASUtxt
-% \RRItxt
-% \NRAONMtxt
-% \CAASTROtxt
-% \Curtintxt
-% %% \Swinburnetxt
-% \CfAtxt
-% \ANUtxt
-% \CSIROtxt
-% \Haystacktxt
-% \USydneytxt
-% \MITtxt
-% \UWtxt
-% \Victoriatxt
-% \UWisctxt
-% \Tatatxt
-% \NRAOGBTtxt
-% \UMelbournetxt
-% \UTasmaniatxt
-% %% \UWAtxt
-% %% \ICRARtxt
-% %\VUWtxt
+\UWtxt
+\RRItxt
+\SKASAtxt
+\ANUtxt
+\Haystacktxt
+\Curtintxt
+\USydneytxt
+\CAASTROtxt
+\MITtxt
+\CfAtxt
+\Victoriatxt
+\UWisctxt
+\UMichigantxt
+\CASStxt
+\Tatatxt
+\NRAOtxt
+\UMelbournetxt
 

### 2014-09-03 foreground_delay_spectrum_modeling.tex

--- 2014-09-03 foreground_delay_spectrum_modeling.tex
@@
 
-We characterize signatures of foreground radio emission and instrument configuration on the observed power spectra of redshifted H{\sc i} 21~cm line emission from the epoch of reionization (EoR) using an antenna pair based ``delay spectrum'' analysis technique.
-We use the 128-tile Murchison Widefield Array (MWA) instrument configuration for our studies.
-% Our analysis is centered around 185~MHz with 30.72~MHz bandwidth at 80~kHz frequency resolution.
-We simulate the delay spectrum response of interferometers, for the first time, to an all-sky foreground model consisting of diffuse and compact Galactic and extragalactic emission.
-By tying model parameters to those of a subset of first season of observing with the MWA, our simulations agree with the observed data to within expected uncertainty.
-A wedge--shaped region of foreground contamination and a spillover from this contamination into relatively foreground--free regions ({\it EoR window}) in the delay spectrum due to spectral properties of foreground emission and the instrument is confirmed.
-Due to shortening of antenna spacings caused by projection effects, diffuse emission produces edge--heavy {\it two--pronged fork}--like features even on wide antenna spacings, hitherto not predicted.
-Features in the inner regions of the wedge are predominantly due to compact foreground objects.
-This results in a {\it pitchfork}--shaped delay spectrum.
-The bright Galactic center severly contaminates the {\it EoR window} on certain antenna spacings in certain observations.
-For such cases, we present an effective practical tool for screening interferometers with high foreground contamination and mitigating their effects on EoR data analysis.
+ Detection and measurement of 21cm HI emission from the Epoch of Reionization, at redshifts above 6, is limited primarily by foreground emission.
+With much of the information coming from the spectral line that maps to time and line of site distance, foregrounds affecting the spectral domain are deserving of close attention, particularly those that are due to the instrument because they are in theory removable.
+Here we investigate in detail the chromatic appearance of flat spectrum foregrounds imprinted by the instrumental transfer function.
+Using the delay spectrum technique to measure power spectrum contamination on a per-baseline basis we demonstrate that the foreground signatures which have the largest impact on the HI signal arise from power entering far from the center of the field of view.
+Comparing data from recent Murchison Widefield Array observations with simulations broken down by sky position and type of emission, we identify smooth emission near the horizon as the largest contributing factor, followed by bright sources in the same locations.
+The observation that these contaminants only effect a fraction of the baselines leads us to propose that by down-weighting certain baselines at certain times, a large fraction of foreground contamination could be mitigated.
 
@@
 
-One of the key challenges in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
-\citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which lack such symmetry.
-Foreground removal techniques rely on the spectral smoothness of foreground emission \citep{mor06,bow09,liu11,par12,dil13}.
+Perhaps the key challenge in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
+\citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness\citep{mor06,bow09,liu11,par12,dil13}.
+Considerable effort has been made towards understanding the $k$-space behavior foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,dat10,bow09}.
+A general model of how smooth spectrum foregrounds are scattered to higher $k$-modes by the wide-field response of the instrument, into the so-called `wedge',  has emerged that provides a high level explanation for the observed foreground power spectrum.
+The conservative foreground strategy that has developed alongside this work is to throw out $k$-modes which could be contaminated (Avoidance).
+The more aggressive option is to subtract a sky model and regain access to modes lost by Avoidance.
+In both cases one large unknown is which parts of the sky are most critical to either Avoid or Subtract.
+Here we will focus primarily on the Avoidance strategy, asking which foreground components are most likely to ``leak'' from foreground modes to Reionization modes.
 
-Precise understanding of these foreground signatures is necessary in order to remove foreground systematics from the measured power spectrum.
-Considerable effort has been made in mapping the residual foreground signatures in observed power spectrum \citep{thy13,pob13,mor12,tro12,dat10,bow09} after foreground subtraction.
+Foregrounds work their way into the power spectrum by either having intrinsic spectral variability, high instrumental chromaticity, or through approximations in the estimating the power spectrum.
+Here we are focusing on the instrumental effects.
+To de-couple from possible analysis effects, such as those pointed out by \citet{Haz13}, we compute power spectra using the per-baseline approach of \citet{pob13,arp14}.
+This approximates the power spectrum as the fourier transform of the spectra output by the correlator.
 
-In this paper, in one of the most detailed characterizations to date, using an interferometer (antenna pair) based delay spectrum technique, we demonstrate the signatures of unsubtracted diffuse and compact foreground models in delay spectrum, a quantity closely related to the sought power spectrum.
-The simplicity of this technique and usage of basic interferometric visibility data fourier transformed along frequency (delay spectrum) on individual interferometers offer very effective diagnostics for scheduling observations and data analysis, and an insightful view onto foreground signatures in the EoR H{\sc i} power spectrum.
-We explore the diversity in these signatures as a function of observation and instrument parameters and confirm our findings with data obtained from the MWA.
-This simple technique allows us to predict foreground signatures, hitherto not predicted.
+ Matching real observations from the Murchison Widefield Array Epoch of Reionization observing program with instrumental simulations of the entire known sky we examine the signature of the power spectrum of each baseline.
+We confirm that the foregrounds most likely to obfuscate the Epoch of Reionization power are not caused by emission central field of view but rather from near the horizon.
+We also find that the smooth galactic emission plays a larger role then previously thought.
+We quantify this contamination by breaking the simulations down by sky region and type of emission.
+Using these we arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to downweight badly contaminated baselines.
 
-We have organized this paper as follows.
-\S\ref{sec:delay-spectrum} defines the delay spectrum and its relation to the power spectrum.
-\S\ref{sec:instrument} and \S\ref{sec:obsparms} describe the MWA instrumental parameters and observation parameters used in this study respectively.
-Analysis of data obtained with the MWA is described in \S\ref{sec:data-analysis}.
-In \S\ref{sec:modeling}, we describe the building blocks of our delay spectrum modeling including the models used for the antenna power pattern and foreground.
-Here, we also establish that the results of modeling and data analysis agree with each other.
-In \S\ref{sec:delay-spectrum-analysis}, we present a detailed analysis of the signatures of foregrounds on the modeled delay spectrum and its dependence on instrumental and observation parameters.
-In \S\ref{sec:fg-grading}, we use our understanding to grade interferometers based on the amount of foreground contamination estimated by the delay spectrum technique.
-In \S\ref{sec:summary}, we present our summary and conclusions.
+In \S\ref{sec:delay-spectrum} we introduce the delay spectrum technique, then in  \S\ref{sec:instrument}, \S\ref{sec:obsparms}, and \S\ref{sec:data-analysis} we describe the MWA setup, summarize the observing parameters and present the resulting data.
+Simulations of these observing parameters are described in \S\ref{sec:modeling} and analyzed by sky location, and type of emission in \S\ref{sec:delay-spectrum-analysis}.
+\S\ref{sec:discussion} offers a discussion along with a basic comparison to other current and future instruments.
+\S\ref{sec:conclusion} offers conclusions and an initial description of a more precise foreground avoidance technique.
 

### 2014-09-04 foreground_delay_spectrum_modeling.tex
interferometer replaced with baseline to match vernacular usage

@@
 
-In summary, $V_\eta(\overline{\mathbf{u}},\eta)$, is closely related to visibilities, $V_f(\overline{\mathbf{u}},f)$, the basic data blocks measured by an interferometer, as seen from equation~\ref{eqn:delay-transform}.
+In summary, $V_\eta(\overline{\mathbf{u}},\eta)$, is closely related to visibilities, $V_f(\overline{\mathbf{u}},f)$, the basic data blocks measured by each baseline, as seen from equation~\ref{eqn:delay-transform}.
 It captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument.
@@
 
-\section{Instrument Parameters}\label{sec:instrument}
+\section{Murchison Widefield Array Observations}\label{sec:instrument}
+\subsection{The MWA}
+The MWA was completed in 2012, and after commissioning, began its Epoch of Reionization observing program  in 2013.
+In its final configuration the MWA is a 128 tile interferometer capable of observing a 30MHz band anywhere between 80 to 300 MHz.
+Each tile is a phased array of 16 bow-tie-shaped dipoles characterized by a grating response with a $\sim$ 25\arcdeg wide primary field of view and multiple secondary side lobes.
+The tile arrangement is a centrally condensed core of about 300m -there are many spacings in the range 5 to 50m- and a radial density that falls off as the inverse of the radius with the longest baselines at 3km \citep{bea12}.
+Here we focus on baselines shorter than 200m (spatial scales relevant to reionization) and a frequency range 170~MHz to 200~MHz.
 
-We use the 128-tile layout of the MWA \citep{bea12} and choose a center frequency of 185~MHz with a bandwidth of 30.72~MHz divided into 384 frequency channels each of width 80~kHz.
-The choice of center frequency translates to $z\approx 6.68$ for the 21~cm spin flip transition of H{\sc i}.
-The bandwidth was chosen to be close to the maximum instantaneous value of the MWA in order to maximize the resolution in delay coordinates and obtain detailed characteristics of delay spectra from foregrounds.
-We restrict our analysis to the innermost antenna spacings with $|\overline{\mathbf{x}}| \leq x_\textrm{max} = 200$~m, or $|\overline{\mathbf{u}}| \leq u_\textrm{max} = 123$ at a frequency of 185~MHz.
-These interferometers are responsive to transverse spatial structures on scales $\gtrsim 1/u_\textrm{max} \approx 28$\arcmin.
-Hereafter, we use the terms tile and antenna interchangeably to refer to the MWA tile.
-
-The bandpass weights in the frequency window of the MWA are obtained through a 8-tap Polyphase Filter Bank (PFB) using a Kaiser window with the $\beta$ parameter set to 5.
-The bandpass weights consist of coarse frequency channels of width 1.28~MHz each of which contains 16 finer frequency channels of width 80~kHz.
-Our analysis, therefore, contains 24 coarse channels.
-As the data are subsequently bandpass calibrated, the bandpass weights are normalized to unity.
-But the edges of the coarse channels are noisy and aliased from adjacent coarse channels.
-Thus, we remove one fine channel from each side of the coarse channel edges from our data.
-This constitutes $W_f(f)$ in equation~\ref{eqn:obsvis}.
+One important feature of the MWA's measured spectrum is the shape of its bandpass.
+The band is divided coarsely into 24 chunks with each chunk weighted by a digital filter.
+The channels at the edges of these chunks have been flagged because they are known to possibly be contaminated by aliasing at a low level.
+Any spectral flags, or other known weights constitutes $W_f(f)$ in equation~\ref{eqn:obsvis}.
 The bandpass properties used in this paper are illustrated in figure~\ref{fig:bandpass}.
@@
 
-Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=90$~K per polarization for all frequency channels for all interferometers throughout the course of observation.
+Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=90$~K per polarization for all frequency channels for all baselines throughout the course of observation.
 We take into account the increase in thermal uncertainty in frequency channels that results due to aforementioned bandpass correction.
+%XXX Since the bandpass correction is digital, shouldn't the noise stay flat after bandpass correction?
 
-\section{Observation Parameters}\label{sec:obsparms}
+\subsection{Observation Parameters}\label{sec:obsparms}
 
@@
 These beamformer delay settings can be changed only in discrete steps.
-Accordingly, observations were made in a ``drift and shift'' (D-S) mode.
-Figure~\ref{fig:pointings} shows the direction of pointing of MWA tiles during the course of the observation.
 The sky is allowed to drift for a certain period of time (usually $\sim 30$~mins) before the discrete shifts in the beamformer delay settings position the beam to be centered again on the same patch of sky.
 This process is repeated throughout the course of the observation $\approx 4.86$~hours.
-The LST of two snapshots we chose for our study are shown as vertical lines at -1.92~hours (wrapped by 360\arcdeg from 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings respectively.
+Figure~\ref{fig:pointings} illustrates how the apparent pointing oscillates around the desired pointing direction.
+
+The LST of two snapshots we chose for our study are shown as vertical lines at -1.92~hours (i.e.
+22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings respectively.
 These pointings (phased array power patterns) are centered at RA = 4\fdg 387, Dec = -29\fdg 86 and RA = -28\fdg 8, Dec = -26\fdg 701, whereas the visibilities themselves are phased to zenith corresponding to RA = -28\fdg 8, Dec = -26\fdg 701, and RA = 1\fdg 35, Dec = -26\fdg 701 respectively.
+
+The observations reported here were observed, near the beginning of the campaign, on Aug 23, 2013.
+This day has been chosen as a reference data set for comparison within the MWA EoR group.
+This day was selected on the basis of the low number of known system errors from among a long stretch of stable observing.
 
@@
 \end{figure}
-
+\subsection*{A note on baselines/delay spectra sign convention}
 Throughout this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the MWA site.
@@
 For instance, at the beginning of the observation, the Galactic center is in the westward sky just about to set.
-Signals from this direction are recorded with a negative delay on interferometers whose antenna spacings are oriented towards the east.
+Signals from this direction are recorded with a negative delay on baselines whose antenna spacings are oriented towards the east.
 Similarly, signals from directions eastward of the local meridian arrive with positive delays on eastward oriented antenna spacings.
@@
 For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all antenna spacings to lie in the range $-67\fdg 5 \leq \theta_b < 112\fdg 5$.
-Interferometers with antenna spacings oriented in the other half-plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
+baselines with antenna spacings oriented in the other half-plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
 
-\section{Data Analysis}\label{sec:data-analysis}
+\subsection{Data Analysis}\label{sec:data-analysis}
+The data are flagged for interference cite ao, removing 3\% of the data.  averaged in time and frequency to 2s, 40khz
 
@@
 The {\it off--zenith} pointing has emission notably higher than in the {\it zenith} pointing inside the wedge boundaries.
-This is shown later to be due to response of the interferometer array to the bright Galactic center and Galactic plane in the westward sky.
+This is shown later to be due to response of the array to the bright Galactic center and Galactic plane in the westward sky.
 Consequently, the contamination into the {\it EoR window} is also higher.
@@
 
-\item {\it Interferometer Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our interferometer antenna spacings by their orientation ($\theta_\textrm{b}$) measured counter-clockwise from East.
+\item {\it baseline Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our baselines by their orientation ($\theta_\textrm{b}$) measured counter-clockwise from East.
 We use the following bins: $-67\fdg 5\le \theta_\textrm{b} < -22\fdg 5$, $-22\fdg 5\le \theta_\textrm{b} < 22\fdg 5$, $22\fdg 5\le \theta_\textrm{b} < 67\fdg 5$, and $67\fdg 5\le \theta_\textrm{b} < 112\fdg 5$.
 The bin centers are oriented towards South-East, East, North-East, and North respectively.
-Since the delays registered at an interferometer depend on the orientation of its antenna spacing vector, binning by interferometer orientation allows us to attribute the delay spectrum to different sky directions.
+Because delays depend on baseline orientation, binning by baseline orientation allows us to match delay spectrum features to different sky directions.
 
@@
 With the addition of noise in the visibilities which is subject to observing time, some of the weaker features may not be as prominently visible.
-Since the foreground signatures are far too numerous and they reside in a vast volume of parameter space such as interferometer orientation, power pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features observed in the delay spectrum of foreground emission.
+Since the foreground signatures are far too numerous and they reside in a vast volume of parameter space such as baseline orientation, power pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features observed in the delay spectrum of foreground emission.
 
@@
 
-Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three interferometers of different antenna spacings oriented northward during the {\it zenith} pointing.
+Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three baselines of different antenna spacings oriented northward during the {\it zenith} pointing.
 The diffuse, compact, and composite components are are shown as dotted, dashed, and solid lines respectively.
@@
 In this case, these correspond to sidelobes of the power pattern that lie between the main lobe and the horizon along the North--South direction.
-In interferometers with short spacings, due to relatively lower resolution along delay axis inside the {\it foreground wedge} and a consequent loss of localization of direction of foreground emission, these secondary peaks blend in with other major peaks and are not distinctly visible.
+In baselines with short spacings, due to relatively lower resolution along delay axis inside the {\it foreground wedge} and a consequent loss of localization of direction of foreground emission, these secondary peaks blend in with other major peaks and are not distinctly visible.
 These findings are consistent with dicussion in \S\ref{sec:csm-side-lobe}.
@@
 
-\section{Screening of Severe Foreground Contaminated Interferometers}\label{sec:fg-grading}
+\section{Screening of Severe Foreground Contaminated baselines}\label{sec:fg-grading}
 
@@
 With a foreground model known {\it a priori} in which locations of very bright foreground objects such as the Galactic center or AGN are available, we can predict their response across antenna spacings as a function of observing parameters such as LST, power pattern, etc.
-Specifically, we can pick interferometers where these bright foreground objects will be visible at delays near horizon delay limits.
+Specifically, we can pick baselines where these bright foreground objects will be visible at delays near horizon delay limits.
 One such example is already provided in our study.
@@
 With spatial scales of such objects known, besides downweighting based on antenna spacing orientation, we could also downweight based on antenna spacing distance.
-For instance, a bright compact object will excite interferometers uniformly across antenna spacings whereas an extended object will gradually fade with increase in antenna spacing.
+For instance, a bright compact object will excite baselines uniformly across antenna spacings whereas an extended object will gradually fade with increase in antenna spacing.
 
 This provides us with a very simple and yet effective tool in adding a layer of sophistication to mitigate effects of foreground contamination in EoR data analysis.
-% The success behind clear identification of severely foreground--contaminated interferometers based on delay spectrum technique depends on the localization of the source of foreground contamination and its strength.
+% The success behind clear identification of severely foreground--contaminated baselines based on delay spectrum technique depends on the localization of the source of foreground contamination and its strength.
 
@@
 We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination in EoR data analysis.
-It is possible to screen interferometers most affected by foreground contamination as a function of antenna spacing vector and time of observation, which can then be assigned optimal weights in the EoR data analysis chain.
+It is possible to screen baselines most affected by foreground contamination as a function of antenna spacing vector and time of observation, which can then be assigned optimal weights in the EoR data analysis chain.
 

### 2014-09-05 foreground_delay_spectrum_modeling.tex
From danny1: ===========
(*) Fixed the citation labels included by Danny
(*) Added a bibtex entry to the eor.bib for an additional reference
(*) Added empty sections 'Discussion' and 'Conclusions' just to compile without
errors

---------------------------------------------------------------------------------
From v0.3.2:
===========

(*) Danny changed abstract and intro and also replaced interferometer with
baseline to match vernacular usage
(*) Danny added a few additional references
(*) Danny made a few other miscellaneous edits
(*) Nithya pulled Danny's version into a branch danny1

@@
 Here we are focusing on the instrumental effects.
-To de-couple from possible analysis effects, such as those pointed out by \citet{Haz13}, we compute power spectra using the per-baseline approach of \citet{pob13,arp14}.
+To de-couple from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the per-baseline approach of \citet{pob13,par14}.
 This approximates the power spectrum as the fourier transform of the spectra output by the correlator.
@@
 
-% \section{Discussion}\label{sec:discussion}
+\section{Discussion}\label{sec:discussion}
+
+\section{Conclusion}\label{sec:conclusion}
 

### 2014-09-05 foreground_delay_spectrum_modeling.tex
From v0.3.2: ===========
(*) Created branch v0.3.2b
(*) Removed some of the comments

--- 2014-09-03 foreground_delay_spectrum_modeling.tex
@@
 
- Detection and measurement of 21cm HI emission from the Epoch of Reionization, at redshifts above 6, is limited primarily by foreground emission.
-With much of the information coming from the spectral line that maps to time and line of site distance, foregrounds affecting the spectral domain are deserving of close attention, particularly those that are due to the instrument because they are in theory removable.
-Here we investigate in detail the chromatic appearance of flat spectrum foregrounds imprinted by the instrumental transfer function.
-Using the delay spectrum technique to measure power spectrum contamination on a per-baseline basis we demonstrate that the foreground signatures which have the largest impact on the HI signal arise from power entering far from the center of the field of view.
-Comparing data from recent Murchison Widefield Array observations with simulations broken down by sky position and type of emission, we identify smooth emission near the horizon as the largest contributing factor, followed by bright sources in the same locations.
-The observation that these contaminants only effect a fraction of the baselines leads us to propose that by down-weighting certain baselines at certain times, a large fraction of foreground contamination could be mitigated.
+We characterize signatures of foreground radio emission and instrument configuration on the observed power spectra of redshifted H{\sc i} 21~cm line emission from the epoch of reionization (EoR) using an antenna pair based ``delay spectrum'' analysis technique.
+We use the 128-tile Murchison Widefield Array (MWA) instrument configuration for our studies.
+% Our analysis is centered around 185~MHz with 30.72~MHz bandwidth at 80~kHz frequency resolution.
+We simulate the delay spectrum response of interferometers, for the first time, to an all-sky foreground model consisting of diffuse and compact Galactic and extragalactic emission.
+By tying model parameters to those of a subset of first season of observing with the MWA, our simulations agree with the observed data to within expected uncertainty.
+A wedge--shaped region of foreground contamination and a spillover from this contamination into relatively foreground--free regions ({\it EoR window}) in the delay spectrum due to spectral properties of foreground emission and the instrument is confirmed.
+Due to shortening of antenna spacings caused by projection effects, diffuse emission produces edge--heavy {\it two--pronged fork}--like features even on wide antenna spacings, hitherto not predicted.
+Features in the inner regions of the wedge are predominantly due to compact foreground objects.
+This results in a {\it pitchfork}--shaped delay spectrum.
+The bright Galactic center severly contaminates the {\it EoR window} on certain antenna spacings in certain observations.
+For such cases, we present an effective practical tool for screening interferometers with high foreground contamination and mitigating their effects on EoR data analysis.
 
@@
 
-Perhaps the key challenge in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
-\citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness\citep{mor06,bow09,liu11,par12,dil13}.
-Considerable effort has been made towards understanding the $k$-space behavior foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,dat10,bow09}.
-A general model of how smooth spectrum foregrounds are scattered to higher $k$-modes by the wide-field response of the instrument, into the so-called `wedge',  has emerged that provides a high level explanation for the observed foreground power spectrum.
-The conservative foreground strategy that has developed alongside this work is to throw out $k$-modes which could be contaminated (Avoidance).
-The more aggressive option is to subtract a sky model and regain access to modes lost by Avoidance.
-In both cases one large unknown is which parts of the sky are most critical to either Avoid or Subtract.
-Here we will focus primarily on the Avoidance strategy, asking which foreground components are most likely to ``leak'' from foreground modes to Reionization modes.
+One of the key challenges in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
+\citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which lack such symmetry.
+Foreground removal techniques rely on the spectral smoothness of foreground emission \citep{mor06,bow09,liu11,par12,dil13}.
 
-Foregrounds work their way into the power spectrum by either having intrinsic spectral variability, high instrumental chromaticity, or through approximations in the estimating the power spectrum.
-Here we are focusing on the instrumental effects.
-To de-couple from possible analysis effects, such as those pointed out by \citet{Haz13}, we compute power spectra using the per-baseline approach of \citet{pob13,arp14}.
-This approximates the power spectrum as the fourier transform of the spectra output by the correlator.
+Precise understanding of these foreground signatures is necessary in order to remove foreground systematics from the measured power spectrum.
+Considerable effort has been made in mapping the residual foreground signatures in observed power spectrum \citep{thy13,pob13,mor12,tro12,dat10,bow09} after foreground subtraction.
 
- Matching real observations from the Murchison Widefield Array Epoch of Reionization observing program with instrumental simulations of the entire known sky we examine the signature of the power spectrum of each baseline.
-We confirm that the foregrounds most likely to obfuscate the Epoch of Reionization power are not caused by emission central field of view but rather from near the horizon.
-We also find that the smooth galactic emission plays a larger role then previously thought.
-We quantify this contamination by breaking the simulations down by sky region and type of emission.
-Using these we arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to downweight badly contaminated baselines.
+In this paper, in one of the most detailed characterizations to date, using an interferometer (antenna pair) based delay spectrum technique, we demonstrate the signatures of unsubtracted diffuse and compact foreground models in delay spectrum, a quantity closely related to the sought power spectrum.
+The simplicity of this technique and usage of basic interferometric visibility data fourier transformed along frequency (delay spectrum) on individual interferometers offer very effective diagnostics for scheduling observations and data analysis, and an insightful view onto foreground signatures in the EoR H{\sc i} power spectrum.
+We explore the diversity in these signatures as a function of observation and instrument parameters and confirm our findings with data obtained from the MWA.
+This simple technique allows us to predict foreground signatures, hitherto not predicted.
 
-In \S\ref{sec:delay-spectrum} we introduce the delay spectrum technique, then in  \S\ref{sec:instrument}, \S\ref{sec:obsparms}, and \S\ref{sec:data-analysis} we describe the MWA setup, summarize the observing parameters and present the resulting data.
-Simulations of these observing parameters are described in \S\ref{sec:modeling} and analyzed by sky location, and type of emission in \S\ref{sec:delay-spectrum-analysis}.
-\S\ref{sec:discussion} offers a discussion along with a basic comparison to other current and future instruments.
-\S\ref{sec:conclusion} offers conclusions and an initial description of a more precise foreground avoidance technique.
+We have organized this paper as follows.
+\S\ref{sec:delay-spectrum} defines the delay spectrum and its relation to the power spectrum.
+\S\ref{sec:instrument} and \S\ref{sec:obsparms} describe the MWA instrumental parameters and observation parameters used in this study respectively.
+Analysis of data obtained with the MWA is described in \S\ref{sec:data-analysis}.
+In \S\ref{sec:modeling}, we describe the building blocks of our delay spectrum modeling including the models used for the antenna power pattern and foreground.
+Here, we also establish that the results of modeling and data analysis agree with each other.
+In \S\ref{sec:delay-spectrum-analysis}, we present a detailed analysis of the signatures of foregrounds on the modeled delay spectrum and its dependence on instrumental and observation parameters.
+In \S\ref{sec:fg-grading}, we use our understanding to grade interferometers based on the amount of foreground contamination estimated by the delay spectrum technique.
+In \S\ref{sec:summary}, we present our summary and conclusions.
 
@@
 We estimate the ratio of delay spectra from data and simulations as, $\rho = |V^\textrm{D}_\eta(|\overline{\mathbf{x}}|,\eta)|\,/\,|V^\textrm{S}_\eta(|\overline{\mathbf{x}}|,\eta)|$, where superscripts D and S denote data and simulation respectively.
-% Using the aforementioned uncertainties added in quadrature, we estimate the approximate expected uncertainty in $\log_{10}\rho$ and denote it by $\Delta\,\log_{10}\rho$.
-
-% Figure~\ref{fig:data-sim-ratio} shows the significance of $\log_{10}\rho$ relative to the expected error $\Delta\,\log_{10}\rho$ for all $|\overline{\mathbf{x}}|$ and $\eta$ inside the {\it foreground wedge}.
-The probability densities and cumulative distributions are shown in black and gray respectively.
-Also shown are dotted vertical lines which denote equality between $\Delta\,\log_{10}\rho$ and $\log_{10}\rho$.
-The fraction of the {\it foreground wedge} that lies in between these vertical lines is $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively.
-In other words, subject to expected uncertainties, the ratio of delay spectra obtained with MWA data and simulations are consistent with unity at $\sim$~90\% and $\sim$~95\% confidence levels in the {\it off--zenith} and {\it zenith} pointings respectively.
 
@@
 \label{fig:data-sim-ratio}}
-% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
-% \caption{Significance of logarithm of ratio of observed to simulated visibilities relative to expected uncertainty restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
-Probability densities and cumulative distributions are shown in black and gray respectively.
-Region inside the dotted vertical lines denotes when the ratio, $\rho$, is consistent with unity subject to expected uncertainty.
-Also indicated in percentages are confidence levels estimated conservatively for such a consistency.
-They are $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively.
-\label{fig:data-sim-ratio-snr}}
 \end{figure}
@@
 
-% The signatures we henceforth identify in the delay spectrum are characterized as arising out of the aforementioned factors.
-Due to numerous combinations of such factors, we use unique codes for these combinations as described in table~\ref{tab:codes}.
-The first column refers to the place holder of the specific parameter in the code sequence from left to right.
-The parameter codes are specified in this order.
-The second column gives the parameter used in describing features in delay spectrum.
-The third column denotes the code describing the parameter.
-Snapshot index uniquely specifies the snapshot and carries a numeric code (1 and 2 for off-zenith and on-zenith snapshots respectively) leading the code sequence.
-Following this, the type of emission is specified as compact, diffuse, Galactic center, or Galactic plane using codes C, D, GC, or GP respectively.
-Although in reality, a feature could be due to a combination of emission types, the code we use refers to the predominant type of emission that can be attributed to the feature.
-The next code specifies the direction of this predominant source of emission.
-The code values for the direction of emission are the usual abbreviations of the cardinal and ordinal directions besides Z (zenith).
-Following this is the code for the baseline orientation bin in the half-plane specified as S, SE, E, and NE along directions south, south-east, east, and north-east respectively.
-Code `A' is used to refer to all baselines.
-Finally, the parameter describing the region of antenna power pattern coincident with the predominant emission responsible for the feature is specified as P (primary lobe) and S (secondary lobe).
-The secondary lobe code carries a numeral suffix to further specify uniquely the sidelobe in the direction of emission.
-For example, a feature code 1-D-W-E-P indicates it is from the first snapshot (1) arising due to diffuse emission (D) westward (W) of zenith observed by predominantly eastward (E) baselines centered on the primary beam (P) of the antenna power pattern.
-Simiarly, feature code 2-GC-W-NE-S3 implies the feature is seen in the second snapshot (2) due to the Galactic Center (GC) that is westward (W) of zenith by north-eastward (NE) baselines and the region coincides with the third sidelobe (S3) in the direction of emission.
-
-% \begin{table*}[htp]
-% \caption[]{Codes used to describe signatures in delay spectrum.\label{tab:codes}}
-% \centering
-% \begin{tabular}{ccc}
-% \tableline\tableline\tableline
-% Order & Parameter Description & Code \\
-% \tableline\tableline
-% 1 & Snapshot index & 1, 2, \ldots \\
-% 2 & Type of emission & C (compact), D (diffuse), GC (Galactic center), GP (Galactic plane) \\
-% 3 & Direction of emission & E (east), N (north), W (west), S (south), Z (zenith)\\
-% & & NE (north-east), NW (north-west), SW (south-west), SE (south-east) \\
-% 4 & Baseline orientation & E (east), S (south), SE (south-east), NE (north-east), A (all) \\
-% 5 & Region in antenna power pattern & P (primary beam), Sn ($n$-th sidelobe, $n=1,2,\ldots$)\\
-% & (in the direction of emission) & \\
-% \tableline
-% \end{tabular}
-% \end{table*}
-
 It may be noted that numerous features overlap at varying levels of significance as a result of various combinations of parameters.
@@
 
-% Figure~\ref{fig:dsm} shows the noiseless delay spectra due to diffuse emission.
-Figure~\ref{fig:dsm_delay_spectrum_snap0} and \ref{fig:dsm_delay_spectrum_snap1} correspond to off-zenith and on-zenith snapshots respectively.
-In general, shorter baselines are expected to be receptive to large scales such as from diffuse emission as well as from small scales from compact objects.
-This effect is most clearly visible in the on-zenith snapshot noted by feature code 2-D-Z-A-P which stands for the signature observed due to diffuse emission from zenith direction on short baselines of all orientations restricted primarily to the primary beam.
-This is a notable feature on baselines of length $\lesssim 60$~m.
-For a frequency of 185~MHz, this length scale corresponds to angular scales of $\gtrsim 1\fdg 5$ on the sky.
-This feature is noted only for diffuse emission from inside the primary beam.
-
-% The most prominent feature, however, is 1-GC-W-E-S3 in figure~\ref{fig:dsm_delay_spectrum_snap0}.
-This is noted in the off-zenith snapshot caused by the bright Galactic center which is on the west at an altitude of $\sim 33$\arcdeg with $\sim 100$~minutes remaining before it sets.
-This emission is observed as it is coincident with the third sidelobe of the power pattern in the westward direction with a neagtive delay of value almost near the horizon limit on eastward baselines.
-
-% \begin{figure*}[htp]
-% \centering
-% \subfloat[][Off-zenith delay spectra from diffuse foreground model]{\label{fig:dsm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_dsm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_0.eps}}
-% \subfloat[][On-zenith delay spectra from diffuse foreground model]{\label{fig:dsm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_dsm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_1.eps}}\\
-% \caption{Delay spectrum visibility amplitudes (in units of Jy~Hz) from diffuse foreground emission obtained on baselines of length $\lesssim 314.7$~m.
-The left and right panels correspond to off-zenith and on-zenith snapshots respectively.
-These are noiseless delay spectra obtained after {\it CLEAN} deconvolution along delay axis.
-A logarithmic color scale is used.
-Most prominent features are due to the bright Galactic center shining through the third sidelobe (1-GC-W-E-S3), diffuse emission of short baselines at zero delays (2-D-Z-A-P), and diffuse emission on long baselines near the horizon limit leaking in through the second sidelobes (1-GP-NE-NE-S2, 2-GP-N-N-S2, and 2-GP-S-N-S2).
-Features from diffuse emission are predominantly edge-heavy resembling a two-pronged fork.
-See text for a detailed description of these features.}
-% \label{fig:dsm}
-% \end{figure*}
-
 \subsubsection{Diffuse Emission}\label{sec:diffuse-features}
@@
 This contributes to a {\it pitchfork}--shaped feature which will be discussed later in \S\ref{sec:pitchfork} in more detail.
-
-% Interestingly, in contrast to feature code 2-D-Z-A-P, all the other features in figures~\ref{fig:dsm_delay_spectrum_snap0} and \ref{fig:dsm_delay_spectrum_snap1} extend from the shortest ($\sim 7.7$~m) to longest baselines ($\sim 314.7$~m).
-We argue this is because even wider baselines are shortened when projected against directions along the baseline vector towards the horizon while considering an all-sky foreground model.
-Such a shortening due to projection effects allows diffuse emission from near the horizon to be observed at maximum delays that correspond to the horizon limit even on long baselines.
-Feature 1-GP-NE-NE-S2 in figure~\ref{fig:dsm_delay_spectrum_snap0}, and features 2-GP-N-N-S2 and 2-GP-S-N-S2 in figure~\ref{fig:dsm_delay_spectrum_snap0} demonstrate this effect of shortening of baselines.
-1-GP-NE-NE-S2 corresponds to emission observed from the Galactic plane from north-eastward directions by north-eastward baselines from the second sidelobe in their power pattern.
-Features 2-GP-N-N-S2 and 2-GP-S-N-S2 are also Galactic plane emission from southward and northward directions observed by second sidelobes of northward baselines with opposite delays.
-Owing to their apparent structure in delay spectra, we hereafter refer broadly to this edge-heavy structure as the ``two-pronged fork'' feature.
-
-% Thus, the most notable features from diffuse emission are those from primary beam area on short baselines, bright Galactic center from the third sidelobe, and the two-pronged fork.
 
@@
 On the other hand, compact objects co--located with sidelobes of power pattern in the {\it zenith} pointing are revealed as distinct but faint branches at positive and negative delays depending on the orientation of antenna spacing and direction of emission on the sky.
-
-% Figure~\ref{fig:csm} shows the noiseless delay spectra obtained from compact foreground objects.
-The off-zenith and on-zenith snapshot delay spectra are shown in figures~\ref{fig:csm_delay_spectrum_snap0} and \ref{fig:csm_delay_spectrum_snap1} respectively.
-
-% In the off-zenith snapshot, the most prominent feature 1-C-E-E-P arises due to eastward baselines observing the eastward sky at positive delays centered on the primary beam.
-The next prominent feature 1-C-E-N-P at zero delays arises due to the compact emission from eastward sky centered on the primary beam observed by northward baselines.
-Features 1-C-NE-NE-S2 and 1-C-W-E-S3 at positive and negative horizon limits are caused by north-eastward sky emission observed by the second sidelobe of north-eastward baselines and westward sky centered on the third sidelobe of eastward baselines respectively.
-
-% The most prominent feature in the on-zenith snapshot is 2-C-Z-A-P appearing at zero delays caused by compact sources at zenith observed by all baselines along their primary beams.
-First and second sidelobes along north and south observed by northward baselines contribute to symmetric positive and negative delays as demonstrated by features 2-C-S-N-S1, 2-C-N-N-S1, 2-C-S-N-S2, and 2-C-N-N-S2.
-
-% \begin{figure*}[htp]
-% \centering
-% \subfloat[][Off-zenith delay spectra from compact foreground objects]{\label{fig:csm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_csm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_0.eps}}
-% \subfloat[][On-zenith delay spectra from compact foreground objects]{\label{fig:csm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_csm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_1.eps}}\\
-% \caption{Same as figure~\ref{fig:dsm} but for compact foreground objects.
-Color scale is the same as in figure~\ref{fig:dsm}.
-The different features annotated are a result of emission observed by compact foreground objects through primary beam and sidelobes.
-Broadly, signatures in delay spectra of compact foreground objects are center-heavy in appearance.
-See text for a detailed description of the different features.}
-% \label{fig:csm}
-% \end{figure*}
-
-% While compact foreground objects give rise to features throughout the foreground wedge, the primary distinction between delay spectra of compact and diffuse foregrounds is that the latter are predominantly center-heavy when compared to edge-heavy features of the former.
-
-% % Figure~\ref{fig:obs-begin} shows the delay spectra obtained with the compact foreground model described in \S\ref{sec:CSM} at the beginning of the observation when the MWA tiles are phased to point at RA = 4\fdg 387, Dec = -29\fdg 86.
-The contours of the antenna power pattern are shown in the central panel.
-The flux densities of compact foreground objects multiplied by the antenna power pattern at the respective locations are shown on a logarithmic color scale right below the central panel.
-For reference, maps of delays between the received signals are shown for 100~m baselines with $\theta_b=0\arcdeg, 45\arcdeg, 90\arcdeg, \textrm{and}\,135\arcdeg$ to relate the strength of emission with the corresponding delay in the delay spectra.
-
-% % \begin{figure*}[htb]
-% % \centering
-% % % \subfloat[][Delay spectra at the beginning of observation]{\label{fig:SNR1D_6hrs_167_single_e0}\includegraphics[width=\linewidth]{}}
-% % \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_CLEAN_visibilities_custom_gaussian_FG_model_csm_all_sky_nside_128_30.7_MHz_bnw2.0_snapshot_0.eps}
-% % \caption{Delay spectra, compact foreground sources, and delay maps for a snapshot at the beginning of the observation pointed at RA = 4\fdg 387, Dec = -29\fdg 86 at LST = 22.08 hours.
-The phase center is at RA = -28\fdg 8, Dec = -26\fdg 701.
-The baselines have been binned by their orientation ($\theta_\textrm{b}$) anti-clockwise from east.
-The center-right panel shows delay spectra from baselines oriented between -22\fdg 5 $\le \theta_\textrm{b}<$+22\fdg 5.
-The upper-right, upper-center, and upper-left panels have baseline orientations lying between +22\fdg 5 $\le \theta_\textrm{b}<$+67\fdg 5, +67\fdg 5 $\le \theta_\textrm{b}<$+112\fdg 5, and +112\fdg 5 $\le \theta_\textrm{b}<$+157\fdg 5, respectively.
-The color scale in these delay spectra are given by the color bar on top.
-The panels diametrically opposite to aforementioned panels represent the delay map on the sky for a 100~m baseline oriented along east, north-east, north, and north-west respectively.
-The color scale for these delay maps is specified at the bottom.
-Delays are positive towards east and north sky positions for eastward and northward baseline orientations respectively.
-The central panel shows the compact foreground model multiplied by the antenna power pattern.
-The color scale is specified below the panel.
-Also shown are contours of the antenna power pattern.
-The contour levels shown are: 0.0078125, 0.03125, 0.125, and 0.5.}
-% % \label{fig:obs-begin}
-% % \end{figure*}
-
-% % \begin{figure*}[htb]
-% % \centering
-% % % \subfloat[][Delay spectra at the beginning of observation]{\label{fig:SNR1D_6hrs_167_single_e0}\includegraphics[width=\linewidth]{}}
-% % \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_CLEAN_visibilities_custom_gaussian_FG_model_csm_all_sky_nside_128_30.7_MHz_bnw2.0_snapshot_1.eps}
-% % \caption{Same as figure~\ref{fig:obs-begin} but for a snapshot during the middle of the observation pointed at RA = +1\fdg 320, Dec = -26\fdg 70 at LST = 0.09 hours.
-The phase center is at RA = 1\fdg 35, Dec = -26\fdg 701.
-}
-% % \label{fig:obs-middle}
-% % \end{figure*}
 
@@
 Complemented by the center-heavy compact foreground features, we have demosntrated the presence of a {\it three--pronged pitchfork}--shaped signature in the delay spectrum of foreground sky.
-
-% Figures~\ref{fig:asm_delay_spectrum_snap0} and \ref{fig:asm_delay_spectrum_snap1} show the delay spectra for the composite all-sky foreground model.
-While features from both its constituents occupy and overlap in different regions of the foreground wedge, their relatives magnitudes determine the resultant signatures seen in the delay spectra.
-For instance, the fork-shaped features of diffuse emission are brighter by a factor $\gtrsim 10$ relative to the compact foreground features at the edges of delay spectra in the respective snapshots.
-Hence, features 1-C-W-E-S3 and 1-C-NE-NE-S2 in figure~\ref{fig:csm_delay_spectrum_snap0} have been masked by features 1-GC-W-E-S3 and 1-GP-NE-NE-S2 respecively from diffuse emission in figure~\ref{fig:dsm_delay_spectrum_snap0}.
-Similarly, the edge features 2-C-S-N-S2 and 2-C-N-N-S2 from compact objects in figure~\ref{fig:csm_delay_spectrum_snap1} have been masked by diffuse emission features 2-GP-S-N-S2 and 2-GP-N-N-S2 respectively in figure~\ref{fig:dsm_delay_spectrum_snap1}.
-
-% On the other hand, in central and inner regions of the wedge, clearly the center-heavy compact foreground model features dominate by $\gtrsim 2$ orders of magnitude relative to features arising out of diffuse emission.
-This is quite evident since almost all the diffuse emission features are confined to the edges of the wedge.
-Further, 2-D-Z-A-P in figure~\ref{fig:dsm_delay_spectrum_snap1} is completely masked by 2-C-Z-A-P in figure~\ref{fig:csm_delay_spectrum_snap1}.
-Thus the inner regions of the wedge consist only of compact foreground features 1-C-E-N-P and 1-C-E-E-P in figure~\ref{fig:asm_delay_spectrum_snap0} and 2-C-Z-A-P, 2-C-S-N-S1 and 2-C-N-N-S1 in figure~\ref{fig:asm_delay_spectrum_snap1}.
-
-% Figures~\ref{fig:baseline_composite_asm_delay_spectrum_snap0} and \ref{fig:baseline_composite_asm_delay_spectrum_snap1} show the delay spectra from the diffuse and compact foreground models as well as from the composite model for the off-zenith and on-zenith snapshots respectively for a selected baseline labeled as ``155-154''.
-This baseline is of length $\simeq 201$~m and orientation $\theta_b\simeq 13\fdg 7$.
-These are effectively vertical slices of figures~\ref{fig:asm_delay_spectrum_snap0} and \ref{fig:asm_delay_spectrum_snap1} through the baseline location.
-The horizon delay limits are shown by vertical gray lines.
-Diffuse emission is negligible in the off-zenith snapshot except at the negative horizon delay limit.
-This is identified to be feature 1-GC-W-E-S3 noted in figure~\ref{fig:asm_delay_spectrum_snap0} due to the bright Galactic center on the westward sky observed by eastward baseline through its third sidelobe.
-All the other features are predominantly due to compact foreground objects.
-The peak at delay of $\sim 0.3\,\mu$~s is identified as feature 1-C-E-E-P in figure~\ref{fig:asm_delay_spectrum_snap0}.
-The peak at zero delay is from feature 1-C-Z-A-S1 while the other secondary peaks arise compact sources coincident with sidelobes of the power pattern.
-In the on-zenith snapshot, the feature 1-C-E-E-P has transformed to feature 2-C-Z-A-P due to change in pointing position.
-The diffuse emission is confined to the edges labeled by features 2-D-W-E-S2 and 2-D-E-E-S2.
-All intermediate peaks are due to compact foreground objects observed through the different sidelobes of the antenna power pattern.
-
-% \begin{figure*}[htp]
-% \centering
-% \subfloat[][Off-zenith delay spectra from composite foreground model]{\label{fig:asm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_asm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_0.eps}}
-% \subfloat[][Off-zenith delay spectra from composite foreground model]{\label{fig:asm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/annotated_combined_baseline_noiseless_CLEAN_visibilities_custom_FG_model_asm_all_sky_nside_128_185.0_MHz_30.7_MHz_bnw2.0_snap_1.eps}}\\
-% \subfloat[][Off-zenith composite delay spectrum on eastward baseline]{\label{fig:baseline_composite_asm_delay_spectrum_snap0}\includegraphics[width=0.49\linewidth]{figures/v1_0/baseline_155-154_composite_noiseless_delay_spectrum_snapshot_0.edited.eps}}
-% \subfloat[][On-zenith composite delay spectrum on eastward baseline]{\label{fig:baseline_composite_asm_delay_spectrum_snap1}\includegraphics[width=0.49\linewidth]{figures/v1_0/baseline_155-154_composite_noiseless_delay_spectrum_snapshot_1.edited.eps}}\\
-% \caption{{\it Top}: Same as figure~\ref{fig:dsm} but for a composite all-sky foreground model consisting of both diffuse emission and compact objects.
-The color scale used is the same.
-The different features annotated arise from those in figures~\ref{fig:dsm} and \ref{fig:csm}.
-Inner regions of the wedge are dominated by center-heavy features such as 1-C-E-N-P, 1-C-E-E-P, 2-C-Z-A-P, 2-C-N-N-S1 and 2-C-S-N-S1 characteristic of compact objects.
-Outer regions are dominated by edge-heavy fork-shaped features such as 1-GC-W-E-S3, 1-GP-NE-NE-S2, 2-GP-S-N-S2 and 2-GP-N-N-S2 characteristic of diffuse emission.
-The resultant structure in delay spectrum takes the form of a {\it pitchfork}.
-{\it Bottom}: Slices of delay spectrum visibility amplitudes from the top panel at the location of a selected eastward baseline.
-The baseline vector is $\sim 201$~m in length oriented $\sim 13\fdg 7$ measured anti-clockwise from east.
-The foreground components are: diffuse sky model (DSM, dashed lines), sky made of compact sources (CSM, dotted lines) and all-sky composite model (ASM, solid lines).
-The gray vertical lines signify the wedge boundaries set by the horizon delay limit.
-Compact sources dominate the delay spectra in the inner part of the wedge as marked by features 1-C-E-E-P and 2-C-Z-A-P.
-Edge-heavy fork-shaped features such as 1-GC-W-E-S3, 2-D-W-E-S2 and 2-D-E-E-S2 from diffuse emission are more significant at the edges of the wedge.
-These give rise to a ``pitchfork'' appearance.
-Other secondary peaks inside the wedge are due to compact objects observed through sidelobes of the antenna power pattern.
-The secondary bumps outside the wedge in all the model components is a result of imperfect deconvolution along delay axis.
-Refer to text for detailed description of these features.}
-% \label{fig:asm}
-% \end{figure*}
-
-% Thus the composite all-sky foreground model results in a delay spectrum structure that captures both the edge-heavy two-pronged fork structure from diffuse foreground emission and the center-heavy features from compact foreground objects.
-Hereafter, we refer to this structure as a ``three-pronged pitchfork''.
 

### 2014-09-05 foreground_delay_spectrum_modeling.tex
added calibration description, simplified delay CLEAN description

--- 2014-09-04 foreground_delay_spectrum_modeling.tex
@@
 
-\subsection{Data Analysis}\label{sec:data-analysis}
-The data are flagged for interference cite ao, removing 3\% of the data.  averaged in time and frequency to 2s, 40khz
-
-\subsection{Imaging and deconvolution}\label{sec:FHD}
-
-{\bf Describe the FHD data analysis up to the point of getting calibrated visibilities here.}
+\subsubsection{Initial Data processing}\label{sec:data-analysis}
+The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference, removing 3\% of the data and averaged in time  and frequency from the raw 1/2s,40kHz to 2s, 80khz.
+These data are then calibrated to a simulation of the sky containing 2420 point sources selected from the MWA Commissioning Survey (N.H Walker, in prep).
+MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning.
+It has a flux limit of 25~mJy and a the declination range of -12\arcdeg to -40\arcdeg evenly covering the field of view of the observations reported here.
+The calibration sources are selected to fall within the 10\% primary lobe of the tile beam.
+The calibration algorithm computes a per channel, per antenna, complex gain which flattens the bandpass and removes small phase slopes due to variation in cable delay.
+%possibly could add elaboration, if I knew exactly which version of the cal was done here.
+doesn't really matter
 
@@
 
-We obtain the delay spectrum of these calibrated visibilities using equation~\ref{eqn:delay-transform} while choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function.
-Due to periodic gaps in frequency bandpass occurring in intervals of 1.28~MHz (see figure~\ref{fig:bandpass}), the delay spectrum is expected to contain harmonics of actual foreground emission repeated at intervals of 0.78~$\mu$s.
-Hence, we have employed deconvolution by {\it CLEAN} algorithm \citep{tay99} along the delay axis \citep{par09,par12} to rid the delay spectra of such artifacts.
-The convolving kernel for the algorithm is given by the inverse Fourier transform of the instrumental bandpass shape.
+We obtain the delay spectrum of these calibrated visibilities taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function.
+The spectrum is multiplied by  the additional weight of the flagged channels which occur rarely for interference and at at regular spacing every 1.28~MHz where the edges of coarse channels are known to have a slight aliasing effect.
+These weights translate into a convolution in the delay space by a ringing harmonic point spread function.
+We deconvolve using a one dimensional CLEAN algorithm   \citep{tay99} as described for the delay axis by \citet{par09,par12}.
 

### 2014-09-05 foreground_delay_spectrum_modeling.tex

--- 2014-09-05 foreground_delay_spectrum_modeling.tex
@@
 
-We characterize signatures of foreground radio emission and instrument configuration on the observed power spectra of redshifted H{\sc i} 21~cm line emission from the epoch of reionization (EoR) using an antenna pair based ``delay spectrum'' analysis technique.
-We use the 128-tile Murchison Widefield Array (MWA) instrument configuration for our studies.
-% Our analysis is centered around 185~MHz with 30.72~MHz bandwidth at 80~kHz frequency resolution.
-We simulate the delay spectrum response of interferometers, for the first time, to an all-sky foreground model consisting of diffuse and compact Galactic and extragalactic emission.
-By tying model parameters to those of a subset of first season of observing with the MWA, our simulations agree with the observed data to within expected uncertainty.
-A wedge--shaped region of foreground contamination and a spillover from this contamination into relatively foreground--free regions ({\it EoR window}) in the delay spectrum due to spectral properties of foreground emission and the instrument is confirmed.
-Due to shortening of antenna spacings caused by projection effects, diffuse emission produces edge--heavy {\it two--pronged fork}--like features even on wide antenna spacings, hitherto not predicted.
-Features in the inner regions of the wedge are predominantly due to compact foreground objects.
-This results in a {\it pitchfork}--shaped delay spectrum.
-The bright Galactic center severly contaminates the {\it EoR window} on certain antenna spacings in certain observations.
-For such cases, we present an effective practical tool for screening interferometers with high foreground contamination and mitigating their effects on EoR data analysis.
+Detection of 21~cm H{\sc i} emission from the Epoch of Reionization (EoR), at redshifts $z>6$, is limited primarily by foreground emission.
+With much of the information coming from the spectral line that maps to distance along line of sight (cosmic time), foregrounds affecting the spectral domain deserve close attention.
+We investigate in detail the chromatic signatures of an all--sky foreground model imprinted by its spectral structure and instrumental transfer function.
+We use an antenna pair based delay spectrum technique.
+Besides confirming foreground contamination is predominantly restricted to a {\it wedge}--shaped region in Fourier space, we demonstrate that the foreground signatures which have the largest impact on the H{\sc i} signal via spillover outside the {\it wedge} arise from power received far away from the primary field of view.
+Comparing data from recent Murchison Widefield Array (MWA) observations with simulations broken down by sky position and type of emission, we identify diffuse emission near the horizon as the largest contributing factor even on wide antenna spacings relative to compact objects, hitherto unpredicted.
+These roles are switched in and around the primary field of view.
+This imprints a {\it pitchfork} on the {\it wedge} in Fourier space.
+The observation that the brightest contaminants are localized and only effect a fraction of the baselines leads us to propose that by down--weighting certain baselines at certain times, a large fraction of foreground contamination can be mitigated.
 
@@
 
-One of the key challenges in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
-\citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which lack such symmetry.
-Foreground removal techniques rely on the spectral smoothness of foreground emission \citep{mor06,bow09,liu11,par12,dil13}.
+Perhaps the key challenge in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
+\citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness\citep{mor06,bow09,liu11,par12,dil13}.
+Considerable effort has been made towards understanding the $k$-space behavior foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,dat10,bow09}.
+A general model of how smooth spectrum foregrounds are scattered to higher $k$-modes by the wide-field response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum.
+The conservative foreground strategy that has developed alongside this work is to discard $k$-modes which could be contaminated (Avoidance).
+The more aggressive alternative is to subtract a sky model and regain access to modes lost by Avoidance.
+In both cases, which parts of the sky are most critical to either Avoid or Subtract has remained largely uncertain.
+Here we will focus primarily on the Avoidance strategy, focusing on identifying foreground components most likely to ``leak'' from foreground modes to EoR modes.
 
-Precise understanding of these foreground signatures is necessary in order to remove foreground systematics from the measured power spectrum.
-Considerable effort has been made in mapping the residual foreground signatures in observed power spectrum \citep{thy13,pob13,mor12,tro12,dat10,bow09} after foreground subtraction.
+Foregrounds leak into the EoR H{\sc i} power spectrum by either having intrinsic spectral variability, high instrumental chromaticity, or through approximations in power spectrum analysis. 
+%Here we focus on the instrumental effects while. 
+To decouple from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the individual antenna pair based approach of \citet{pob13,par14}.
+This approximates the power spectrum as the inverse fourier transform of the spectra generated by the correlator.
 
-In this paper, in one of the most detailed characterizations to date, using an interferometer (antenna pair) based delay spectrum technique, we demonstrate the signatures of unsubtracted diffuse and compact foreground models in delay spectrum, a quantity closely related to the sought power spectrum.
-The simplicity of this technique and usage of basic interferometric visibility data fourier transformed along frequency (delay spectrum) on individual interferometers offer very effective diagnostics for scheduling observations and data analysis, and an insightful view onto foreground signatures in the EoR H{\sc i} power spectrum.
-We explore the diversity in these signatures as a function of observation and instrument parameters and confirm our findings with data obtained from the MWA.
-This simple technique allows us to predict foreground signatures, hitherto not predicted.
+ Matching real observations from the Murchison Widefield Array Epoch of Reionization observing program with instrumental simulations of the entire known sky we examine the signature of the power spectrum of each baseline.
+We confirm that the foregrounds most likely to obfuscate the Epoch of Reionization power are not caused by emission central field of view but rather from near the horizon.
+We also find that the smooth galactic emission plays a larger role then previously thought.
+We quantify this contamination by breaking the simulations down by sky region and type of emission.
+Using these we arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to downweight badly contaminated baselines.
 
-We have organized this paper as follows.
-\S\ref{sec:delay-spectrum} defines the delay spectrum and its relation to the power spectrum.
-\S\ref{sec:instrument} and \S\ref{sec:obsparms} describe the MWA instrumental parameters and observation parameters used in this study respectively.
-Analysis of data obtained with the MWA is described in \S\ref{sec:data-analysis}.
-In \S\ref{sec:modeling}, we describe the building blocks of our delay spectrum modeling including the models used for the antenna power pattern and foreground.
-Here, we also establish that the results of modeling and data analysis agree with each other.
-In \S\ref{sec:delay-spectrum-analysis}, we present a detailed analysis of the signatures of foregrounds on the modeled delay spectrum and its dependence on instrumental and observation parameters.
-In \S\ref{sec:fg-grading}, we use our understanding to grade interferometers based on the amount of foreground contamination estimated by the delay spectrum technique.
-In \S\ref{sec:summary}, we present our summary and conclusions.
+In \S\ref{sec:delay-spectrum} we introduce the delay spectrum technique, then in  \S\ref{sec:instrument}, \S\ref{sec:obsparms}, and \S\ref{sec:data-analysis} we describe the MWA setup, summarize the observing parameters and present the resulting data.
+Simulations of these observing parameters are described in \S\ref{sec:modeling} and analyzed by sky location, and type of emission in \S\ref{sec:delay-spectrum-analysis}.
+\S\ref{sec:discussion} offers a discussion along with a basic comparison to other current and future instruments.
+\S\ref{sec:conclusion} offers conclusions and an initial description of a more precise foreground avoidance technique.
 
@@
 
-In summary, $V_\eta(\overline{\mathbf{u}},\eta)$, is closely related to visibilities, $V_f(\overline{\mathbf{u}},f)$, the basic data blocks measured by an interferometer, as seen from equation~\ref{eqn:delay-transform}.
+In summary, $V_\eta(\overline{\mathbf{u}},\eta)$, is closely related to visibilities, $V_f(\overline{\mathbf{u}},f)$, the basic data blocks measured by each antenna pair, as seen from equation~\ref{eqn:delay-transform}.
 It captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument.
@@
 
-\section{Instrument Parameters}\label{sec:instrument}
+\section{Murchison Widefield Array Observations}\label{sec:instrument}
 
-We use the 128-tile layout of the MWA \citep{bea12} and choose a center frequency of 185~MHz with a bandwidth of 30.72~MHz divided into 384 frequency channels each of width 80~kHz.
-The choice of center frequency translates to $z\approx 6.68$ for the 21~cm spin flip transition of H{\sc i}.
-The bandwidth was chosen to be close to the maximum instantaneous value of the MWA in order to maximize the resolution in delay coordinates and obtain detailed characteristics of delay spectra from foregrounds.
-We restrict our analysis to the innermost antenna spacings with $|\overline{\mathbf{x}}| \leq x_\textrm{max} = 200$~m, or $|\overline{\mathbf{u}}| \leq u_\textrm{max} = 123$ at a frequency of 185~MHz.
-These interferometers are responsive to transverse spatial structures on scales $\gtrsim 1/u_\textrm{max} \approx 28$\arcmin.
-Hereafter, we use the terms tile and antenna interchangeably to refer to the MWA tile.
+\subsection{The MWA}
 
-The bandpass weights in the frequency window of the MWA are obtained through a 8-tap Polyphase Filter Bank (PFB) using a Kaiser window with the $\beta$ parameter set to 5.
-The bandpass weights consist of coarse frequency channels of width 1.28~MHz each of which contains 16 finer frequency channels of width 80~kHz.
-Our analysis, therefore, contains 24 coarse channels.
-As the data are subsequently bandpass calibrated, the bandpass weights are normalized to unity.
-But the edges of the coarse channels are noisy and aliased from adjacent coarse channels.
-Thus, we remove one fine channel from each side of the coarse channel edges from our data.
-This constitutes $W_f(f)$ in equation~\ref{eqn:obsvis}.
+The MWA was completed in 2012, and after commissioning, began its EoR observing program in 2013.
+In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30~MHz band anywhere in the range 80--300~MHz.
+Each tile is a phased array of 16 dipoles each in the shape of a bow--tie characterized by a grating response with a $\sim$~25\arcdeg wide primary field of view and multiple secondary side lobes.
+The tile arrangement is a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius with the longest baselines at 3~km \citep{bea12}.
+Here we focus on antenna spacings $|\overline{\mathbf{x}}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz.
+
+One important feature of the MWA's measured spectrum is the shape of its bandpass.
+The band is divided coarsely into 24 chunks with each chunk weighted by a digital filter.
+The channels at the edges of these chunks have been flagged because they are known to possibly be contaminated by aliasing at a low level.
+Any spectral flags, or other known weights constitutes $W_f(f)$ in equation~\ref{eqn:obsvis}.
 The bandpass properties used in this paper are illustrated in figure~\ref{fig:bandpass}.
@@
 \caption{Bandpass properties used in modeling and analysis of delay spectrum of MWA visibilities.
-The filled circles joined by dotted lines show the 8-tap PFB bandpass shapes obtained using a Kaiser window with parameter $\beta=5$.
+The filled circles joined by dotted lines show the 8--tap PFB bandpass shapes obtained using a Kaiser window with parameter $\beta=5$.
 The {\it plus} symbols joined by dashed lines are the bandpass correction factors.
@@
 
-Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=90$~K per polarization for all frequency channels for all interferometers throughout the course of observation.
+Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=90$~K per polarization for all frequency channels for all antenna pairs throughout the course of observation.
 We take into account the increase in thermal uncertainty in frequency channels that results due to aforementioned bandpass correction.
+({\bf Since the bandpass correction is digital, shouldn't the noise stay flat after bandpass correction?})
 
-\section{Observation Parameters}\label{sec:obsparms}
+\subsection{Observation Parameters}\label{sec:obsparms}
 
@@
 These beamformer delay settings can be changed only in discrete steps.
-Accordingly, observations were made in a ``drift and shift'' (D-S) mode.
-Figure~\ref{fig:pointings} shows the direction of pointing of MWA tiles during the course of the observation.
 The sky is allowed to drift for a certain period of time (usually $\sim 30$~mins) before the discrete shifts in the beamformer delay settings position the beam to be centered again on the same patch of sky.
 This process is repeated throughout the course of the observation $\approx 4.86$~hours.
-The LST of two snapshots we chose for our study are shown as vertical lines at -1.92~hours (wrapped by 360\arcdeg from 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings respectively.
+Figure~\ref{fig:pointings} illustrates how the apparent pointing oscillates around the desired pointing direction.
+
+The LST of two snapshots we chose for our study are shown as vertical lines at -1.92~hours (i.e., 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings respectively.
 These pointings (phased array power patterns) are centered at RA = 4\fdg 387, Dec = -29\fdg 86 and RA = -28\fdg 8, Dec = -26\fdg 701, whereas the visibilities themselves are phased to zenith corresponding to RA = -28\fdg 8, Dec = -26\fdg 701, and RA = 1\fdg 35, Dec = -26\fdg 701 respectively.
+
+The observations reported here were observed, near the beginning of the campaign, on Aug 23, 2013.
+This day has been chosen as a reference data set for comparison within the MWA EoR collaboration.
+This day was selected on the basis of low number of known system errors from among a long stretch of stable observing.
 
@@
 
+\subsection*{A note on baselines/delay spectra sign convention}
+
 Throughout this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the MWA site.
@@
 For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all antenna spacings to lie in the range $-67\fdg 5 \leq \theta_b < 112\fdg 5$.
-Interferometers with antenna spacings oriented in the other half-plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
+Antenna pairs with spacings oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
 
-\section{Data Analysis}\label{sec:data-analysis}
+\subsection{Data Analysis}\label{sec:data-analysis}
+
+The data are flagged for interference \citep{off12,off10}, removing $\sim 3$\% of the data. These are averaged in time and frequency to 2s and 40kHz respectively.
 
@@
 We estimate the ratio of delay spectra from data and simulations as, $\rho = |V^\textrm{D}_\eta(|\overline{\mathbf{x}}|,\eta)|\,/\,|V^\textrm{S}_\eta(|\overline{\mathbf{x}}|,\eta)|$, where superscripts D and S denote data and simulation respectively.
+% Using the aforementioned uncertainties added in quadrature, we estimate the approximate expected uncertainty in $\log_{10}\rho$ and denote it by $\Delta\,\log_{10}\rho$.
+
+% Figure~\ref{fig:data-sim-ratio} shows the significance of $\log_{10}\rho$ relative to the expected error $\Delta\,\log_{10}\rho$ for all $|\overline{\mathbf{x}}|$ and $\eta$ inside the {\it foreground wedge}.
+The probability densities and cumulative distributions are shown in black and gray respectively.
+Also shown are dotted vertical lines which denote equality between $\Delta\,\log_{10}\rho$ and $\log_{10}\rho$.
+The fraction of the {\it foreground wedge} that lies in between these vertical lines is $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively.
+In other words, subject to expected uncertainties, the ratio of delay spectra obtained with MWA data and simulations are consistent with unity at $\sim$~90\% and $\sim$~95\% confidence levels in the {\it off--zenith} and {\it zenith} pointings respectively.
 
@@
 \label{fig:data-sim-ratio}}
+% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
+% \caption{Significance of logarithm of ratio of observed to simulated visibilities relative to expected uncertainty restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+Probability densities and cumulative distributions are shown in black and gray respectively.
+Region inside the dotted vertical lines denotes when the ratio, $\rho$, is consistent with unity subject to expected uncertainty.
+Also indicated in percentages are confidence levels estimated conservatively for such a consistency.
+They are $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively.
+\label{fig:data-sim-ratio-snr}}
 \end{figure}
@@
 
-\item {\it Interferometer Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our interferometer antenna spacings by their orientation ($\theta_\textrm{b}$) measured counter-clockwise from East.
+\item {\it Antenna Pair Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our antenna pairs by their orientation ($\theta_\textrm{b}$) measured counter-clockwise from East.
 We use the following bins: $-67\fdg 5\le \theta_\textrm{b} < -22\fdg 5$, $-22\fdg 5\le \theta_\textrm{b} < 22\fdg 5$, $22\fdg 5\le \theta_\textrm{b} < 67\fdg 5$, and $67\fdg 5\le \theta_\textrm{b} < 112\fdg 5$.
 The bin centers are oriented towards South-East, East, North-East, and North respectively.
-Since the delays registered at an interferometer depend on the orientation of its antenna spacing vector, binning by interferometer orientation allows us to attribute the delay spectrum to different sky directions.
+Since delays depend on antenna pair orientation, binning by interferometer orientation allows us to match delay spectrum features to different sky directions.
 
@@
 With the addition of noise in the visibilities which is subject to observing time, some of the weaker features may not be as prominently visible.
-Since the foreground signatures are far too numerous and they reside in a vast volume of parameter space such as interferometer orientation, power pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features observed in the delay spectrum of foreground emission.
+Since the foreground signatures are far too numerous and they reside in a vast volume of parameter space such as antenna pair orientation, power pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features observed in the delay spectrum of foreground emission.
 
@@
 
-Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three interferometers of different antenna spacings oriented northward during the {\it zenith} pointing.
+Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing.
 The diffuse, compact, and composite components are are shown as dotted, dashed, and solid lines respectively.
@@
 In this case, these correspond to sidelobes of the power pattern that lie between the main lobe and the horizon along the North--South direction.
-In interferometers with short spacings, due to relatively lower resolution along delay axis inside the {\it foreground wedge} and a consequent loss of localization of direction of foreground emission, these secondary peaks blend in with other major peaks and are not distinctly visible.
+In short antenna spacings, due to relatively lower resolution along delay axis inside the {\it foreground wedge} and a consequent loss of localization of direction of foreground emission, these secondary peaks blend in with other major peaks and are not distinctly visible.
 These findings are consistent with dicussion in \S\ref{sec:csm-side-lobe}.
@@
 With a foreground model known {\it a priori} in which locations of very bright foreground objects such as the Galactic center or AGN are available, we can predict their response across antenna spacings as a function of observing parameters such as LST, power pattern, etc.
-Specifically, we can pick interferometers where these bright foreground objects will be visible at delays near horizon delay limits.
+Specifically, we can pick antenna pairs where these bright foreground objects will be visible at delays near horizon delay limits.
 One such example is already provided in our study.
@@
 With spatial scales of such objects known, besides downweighting based on antenna spacing orientation, we could also downweight based on antenna spacing distance.
-For instance, a bright compact object will excite interferometers uniformly across antenna spacings whereas an extended object will gradually fade with increase in antenna spacing.
+For instance, a bright compact object will yield uniform {\it visibility} amplitudes across antenna spacings whereas that of an extended object will gradually fade with increase in antenna spacing.
 
 This provides us with a very simple and yet effective tool in adding a layer of sophistication to mitigate effects of foreground contamination in EoR data analysis.
-% The success behind clear identification of severely foreground--contaminated interferometers based on delay spectrum technique depends on the localization of the source of foreground contamination and its strength.
+% The success behind clear identification of severely foreground--contaminated antenna pairs based on delay spectrum technique depends on the localization of the source of foreground contamination and its strength.
 
-% \section{Discussion}\label{sec:discussion}
+\section{Discussion}\label{sec:discussion}
+
+\section{Conclusion}\label{sec:conclusion}
 
@@
 We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination in EoR data analysis.
-It is possible to screen interferometers most affected by foreground contamination as a function of antenna spacing vector and time of observation, which can then be assigned optimal weights in the EoR data analysis chain.
+It is possible to screen antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation, which can then be assigned optimal weights in the EoR data analysis chain.
 

### 2014-09-07 foreground_delay_spectrum_modeling.tex
updated data processing, clean and leakage sections

--- 2014-09-05 foreground_delay_spectrum_modeling.tex
@@
 It has a flux limit of 25~mJy and a the declination range of -12\arcdeg to -40\arcdeg evenly covering the field of view of the observations reported here.
-The calibration sources are selected to fall within the 10\% primary lobe of the tile beam.
-The calibration algorithm computes a per channel, per antenna, complex gain which flattens the bandpass and removes small phase slopes due to variation in cable delay.
+The calibration sources are selected to fall within the 5\% primary lobe of the tile beam.
+The calibration algorithm computes a per channel, per antenna, complex gain solutions averaged to two minute intervals.
+The solutions are still fairly low signal to noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna to antenna variation.
+First we average the channel gains over all antenna to obtain a high signal to noise measurement of the bandpass.
+After applying this single passband, we do a second round of calibration and fit a 2nd order amplitude and first order phase for each antenna.
+This flattens any residual variation in bandpass and removes small phase slopes due to variation in cable delay.
+Finally we fit for an additional phase known to be caused by reflection in 150m cables.
 %possibly could add elaboration, if I knew exactly which version of the cal was done here.
@@
 The spectrum is multiplied by  the additional weight of the flagged channels which occur rarely for interference and at at regular spacing every 1.28~MHz where the edges of coarse channels are known to have a slight aliasing effect.
-These weights translate into a convolution in the delay space by a ringing harmonic point spread function.
-We deconvolve using a one dimensional CLEAN algorithm   \citep{tay99} as described for the delay axis by \citet{par09,par12}.
+In delay space, these weights translate into a convolution by a ringing harmonic point spread function (PSF).
+We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis by \citet{par09,par12}.
+The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights.
+We limit the selection of peaks to modes inside the delay horizon corresponding to smooth spectrum sources on the sky --ie beneath the black line in Figure \ref{fig:fhd_data}.
 
@@
 We show results of delay spectrum from MWA data after deconvolution along delay axis in figure~\ref{fig:fhd_data}.
-Boundaries of the {\it foreground wedge} are also shown.
+In this plot and others, we plot the amplitude of the delay spectrum sorted by baseline length.
+Note that no baselines have been averaged together in this plot.
+The principle features are the wedge-shaped feature the boundary of which increases approximately linearly with light travel time of the baseline length.
+Note that the wedge is not filled out uniformly.
+Each baseline has a different orientation and resulting response to different sources in the sky.
+It is this variation that we are investigating.
+
+In these power spectrum units [Mpc$^{-3}$], the reionization power spectrum is expected to be spherically symmetric and decrease quickly with $|k|$.
+The brightest signal will occur on the shortest baselines and smallest delays.
+Thus the region of interest is just above the black horizon line and below the harmonic due to the coarse channels.
+
  The {\it off--zenith} pointing has emission notably higher than in the {\it zenith} pointing inside the wedge boundaries.
@@
 Consequently, the contamination into the {\it EoR window} is also higher.
-Emission in {\it zenith} pointing is more centrally concentrated due to phasing of power pattern to the zenith, whereas the {\it off--zenith} pointing has a power pattern phased eastward of zenith and is also responsive to emission from angles far away from zenith.
+Emission in the {\it zenith} pointing is more centrally concentrated due to phasing of power pattern to the zenith, whereas the {\it off--zenith} pointing has a power pattern phased eastward of zenith and is also responsive to emission from angles far away from zenith.
 
@@
 Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the data perfectly.
-This is predominant at short antenna spacings.
+This is predominant at short antenna spacings where the number of delay modes within the horizon.
 This is because the maximum delay envelope (boundary of the wedge) consists of mixed emission from large portions of the sky into a narrow range of delays and the deconvolution algorithm does not have sufficient support in delay space to fit the delay spectrum.

### 2014-09-07 foreground_delay_spectrum_modeling.tex

--- 2014-09-05 foreground_delay_spectrum_modeling.tex
@@
 
-\subsection{Data Analysis}\label{sec:data-analysis}
-
-The data are flagged for interference \citep{off12,off10}, removing $\sim 3$\% of the data. These are averaged in time and frequency to 2s and 40kHz respectively.
-
-\subsection{Imaging and deconvolution}\label{sec:FHD}
-
-{\bf Describe the FHD data analysis up to the point of getting calibrated visibilities here.}
+\subsubsection{Initial Data processing}\label{sec:data-analysis}
+The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off12,off10}, removing 3\% of the data and averaged in time  and frequency from the raw 1/2s,40kHz to 2s, 80khz.
+These data are then calibrated to a simulation of the sky containing 2420 point sources selected from the MWA Commissioning Survey (N.H Walker, in prep).
+MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning.
+It has a flux limit of 25~mJy and a the declination range of -12\arcdeg to -40\arcdeg evenly covering the field of view of the observations reported here.
+The calibration sources are selected to fall within the 5\% primary lobe of the tile beam.
+The calibration algorithm computes a per channel, per antenna, complex gain solutions averaged to two minute intervals.
+The solutions are still fairly low signal to noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna to antenna variation.
+First we average the channel gains over all antenna to obtain a high signal to noise measurement of the bandpass.
+After applying this single passband, we do a second round of calibration and fit a 2nd order amplitude and first order phase for each antenna.
+This flattens any residual variation in bandpass and removes small phase slopes due to variation in cable delay.
+Finally we fit for an additional phase known to be caused by reflection in 150m cables.
+%possibly could add elaboration, if I knew exactly which version of the cal was done here.
+doesn't really matter
 
@@
 
-We obtain the delay spectrum of these calibrated visibilities using equation~\ref{eqn:delay-transform} while choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function.
-Due to periodic gaps in frequency bandpass occurring in intervals of 1.28~MHz (see figure~\ref{fig:bandpass}), the delay spectrum is expected to contain harmonics of actual foreground emission repeated at intervals of 0.78~$\mu$s.
-Hence, we have employed deconvolution by {\it CLEAN} algorithm \citep{tay99} along the delay axis \citep{par09,par12} to rid the delay spectra of such artifacts.
-The convolving kernel for the algorithm is given by the inverse Fourier transform of the instrumental bandpass shape.
+We obtain the delay spectrum of these calibrated visibilities taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function.
+The spectrum is multiplied by  the additional weight of the flagged channels which occur rarely for interference and at at regular spacing every 1.28~MHz where the edges of coarse channels are known to have a slight aliasing effect.
+In delay space, these weights translate into a convolution by a ringing harmonic point spread function (PSF).
+We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis by \citet{par09,par12}.
+The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights.
+We limit the selection of peaks to modes inside the delay horizon corresponding to smooth spectrum sources on the sky --ie beneath the black line in Figure \ref{fig:fhd_data}.
 

### 2014-09-07 foreground_delay_spectrum_modeling.tex
Editing Observation parameters, data figure caption, and primary beam modelling to include discussion about Curtin beam model

@@
 
-One of the primary targets for EoR observing with the MWA is a patch of sky centered around RA = 00h 00m 00.00s, Dec = -30\arcdeg$\,$00\arcmin$\,$00\farcs 0.
-This is one of the regions of low foreground emission within 27\arcdeg$\,$ from the zenith.
+The MWA Epoch of reionization targets two primary low-foreground fields.
+Here we choose to focus on RA = 0h, Dec = -30\arcdeg.
 The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of an array of dipoles in a MWA tile.
@@
 
-The LST of two snapshots we chose for our study are shown as vertical lines at -1.92~hours (i.e., 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings respectively.
-These pointings (phased array power patterns) are centered at RA = 4\fdg 387, Dec = -29\fdg 86 and RA = -28\fdg 8, Dec = -26\fdg 701, whereas the visibilities themselves are phased to zenith corresponding to RA = -28\fdg 8, Dec = -26\fdg 701, and RA = 1\fdg 35, Dec = -26\fdg 701 respectively.
+The observations reported here were observed, near the beginning of the campaign, on Aug 23, 2013.
+This day has been chosen as a reference data set for comparison within the MWA EoR collaboration and was selected on the basis of low number of known system errors from among a long stretch of stable observing.
 
-The observations reported here were observed, near the beginning of the campaign, on Aug 23, 2013.
-This day has been chosen as a reference data set for comparison within the MWA EoR collaboration.
-This day was selected on the basis of low number of known system errors from among a long stretch of stable observing.
+Here we have chosen two, 112 second long sections from this night for detailed study.
+These were chosen to provide a selection of possible foreground and instrumental conditions.
+As an example of ``nominal'' observing setup we choose a zenith pointing; as an example of poor foreground conditions we choose a pointing when the field is 2 hours from zenith.
+This pointing has a significantly higher secondary side lobe structure and is recorded when the bright galactic center is well above the horizon.
+
+The LSTs of these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at -1.92~hours (i.e., 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings respectively.
+%These pointings (phased array power patterns) are centered at RA = 4\fdg 387, Dec = -29\fdg 86 and RA = -28\fdg 8, Dec = -26\fdg 701, whereas the visibilities themselves are phased to zenith corresponding to RA = -28\fdg 8, Dec = -26\fdg 701, and RA = 1\fdg 35, Dec = -26\fdg 701 respectively.
 
@@
 \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
-\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with MWA data for {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
-White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
-The features in the wedge appear to be brighter and the spillover contamination appears to be higher in the {\it off--zenith} pointing relative to the {\it zenith} pointing due to the presence of brighter foregrounds such as the Galactic center and the Galactic plane.
-The logarithmic color scale (shown at the right) is common to both panels.\label{fig:fhd_data}}
+\caption{ Delay spectrum power spectra of two 112s  Murchison Widefield Array observations of the Epoch of Reionization field at RA=0, Dec=-30.
+At top an {\it off--zenith}  pointing when tiles have significant side lobe response in the direction of the galactic center.
+At bottom a {\it zenith} pointing where side lobes are minimal and the galaxy has completely set.
+To get the the power spectrum each baseline is Fourier transformed into delay space ($y$-axis) and the amplitude plotted as a logarithmic color in order of baseline length ($x$-axis).
+The logarithmic color scale (shown at the right) is common to both panels.
+White lines mark the boundaries of the {\it foreground wedge} determined by the light travel time across each baseline.
+Smooth spectrum sources are physically limited to the region below this line.
+The various features within the wedge depend on the \emph{orientation} of the baseline, the side lobe structure of the primary beam pointing and the sky at that time.
+\label{fig:fhd_data}}
 \end{figure}
@@
 
-\section{Delay Spectrum Modeling}\label{sec:modeling}
+\section{Instrument Modeling}\label{sec:modeling}
 
-We describe power pattern and foreground models we have used in modeling the measured delay spectra.
+We describe power pattern and foreground models we have used in modeling the instrument. %todo
 
@@
 
-The power pattern of a MWA tile, $A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}, over the entire hemisphere is analytically modeled based on a 4$\times$4 phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane.
-We also assume that the delay corrections during the phased addition of dipole voltages in the beamformer suffer from random fluctuations of rms 0.05~ns.
-The models of power pattern used in data analysis take into account effects of mutual coupling of dipoles in the tile besides finite ground plane effects.
-Our models were found to be in reasonable agreement with those used in the data analysis.
+The power pattern of a MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is a subject of active research.
+The most precise model currently available is that of Sutinjo et al (2014; submitted?) which models the 4$\times$4 array of bow-tie dipoles as a mutually-coupled array with the overall power-pattern of the dipole modeled via finite element electromagnetic simulation.
+To speed simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the detailed simulation.
+We also assume that the delay corrections during the phased addition of dipole voltages in the beamformer suffer from random fluctuations of rms 0.05~ns, a number in line with the known stability level of the analog signal chain.
 

### 2014-09-09 foreground_delay_spectrum_modeling.tex
updated beam description. small edits through section 4.2.1

@@
 
-The power pattern of a MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is a subject of active research.
+The precise power pattern of the MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is a subject of active investigation.
 The most precise model currently available is that of Sutinjo et al (2014; submitted?) which models the 4$\times$4 array of bow-tie dipoles as a mutually-coupled array with the overall power-pattern of the dipole modeled via finite element electromagnetic simulation.
-To speed simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the detailed simulation.
-We also assume that the delay corrections during the phased addition of dipole voltages in the beamformer suffer from random fluctuations of rms 0.05~ns, a number in line with the known stability level of the analog signal chain.
+To speed simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the FEKO simulation.
+We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known stability level of the analog signal chain.
 
@@
 \subsection{Foreground Model}\label{sec:foreground}
+%todo
+The MWA has a very wide field of view ($\gtrsim 20$\arcdeg at 185~MHz) and has non-zero response all the way to the horizon.
+As we have seen above, foreground sources contaminate the region inside the ``wedge'' with an amplitude that depends on the primary beam amplitude in that direction, with sources at the horizon being closest to the edge of the wedge.
 
-An instrument such as the MWA has a very wide field of view ($\gtrsim 20$\arcdeg at 185~MHz) for imaging purposes.
-In the context of EoR H{\sc i} power spectrum, it has been known that unsubtracted forground sources anywhere in the visible hemisphere up to the horizon directly contaminate the spatial frequency modes in the power spectrum.
-In addition, this contamination also spills over into the relatively cleaner regions, called {\it EoR window}, due to spectral properties of the instrument and the foregrounds \citep{thy13,pob13,ved12,par12}.
+The degree to which these sources leak power into spectral line of sight modes above the wedge, a region known as the ``EoR Window'', depends primarily on the brightness of foregrounds near the edge of the wedge \citep{thy13,pob13,ved12,par12}.
 Thus, it is important to consider an all-sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting to the primary field of view.
 
-\citet{bea13} estimated the sensitivity of the MWA to EoR H{\sc i} power spectrum detection taking into account theremal noise effects in the {\it Eor window}.
+\citet{bea13} estimated the sensitivity of the MWA to EoR H{\sc i} power spectrum detection taking into account thermal noise effects in the {\it Eor window}.
 \citet{thy13} estimated the sensitivity taking into account the spillover from foreground contamination from unsubtracted extragalactic point sources in the {\t EoR window} besides thermal noise.
 In this paper, we include the diffuse Galactic emission for enhancing our understanding of foreground signatures in the power spectrum.
-We use an all-sky foreground emission model that consists of diffuse and bright compact components.
+We use an all-sky foreground emission model that includes both diffuse and bright compact components.
 
@@
 However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals.
-Using the same source model, we also obtain the diffuse emission maps at 170~MHz and 200~MHz to estimate the spectral index map of the diffuse emission model.
+Using the same model, we use the diffuse emission maps at 170~MHz and 200~MHz to estimate the spectral index map of the diffuse emission model.
 
@@
 Contours of the MWA tile power pattern shown in figure~\ref{fig:power_pattern} are also overlaid.
-Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the visible hemisphere in the westward sky and the MWA tile's power receptivity is significantly high ($\gtrsim 0.125$) in that direction.
-In the {\it zenith} pointing, the Galactic plane has almost set and the contour level of power pattern in that direction is at least 16 times lesser.
+Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the visible hemisphere in the westward sky and the MWA tile's power receptivity is significantly high ($\gtrsim 12$\%) in that direction.
+In the {\it zenith} pointing, the Galactic plane has almost set and the contour level of power pattern in that direction is at least 16 times smaller.
 
@@
 Contours of power pattern shown in figure~\ref{fig:power_pattern} are overlaid.
-The contour levels shown are 0.001953125, 0.0078125, 0.03125, 0.125, and 0.5.
-The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off--zenith} pointing and the MWA tile power receptivity is significant ($\gtrsim 0.125$) in that direction.
+The contour levels shown are 0.00195, 0.00781, 0.0312, 0.125, and 0.5.
+The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off--zenith} pointing and the MWA tile power receptivity is significant ($\gtrsim 12$\%) in that direction.
 In contrast, emission from the Galactic plane in {\it zenith} pointing is significantly lesser.

### 2014-09-10 foreground_delay_spectrum_modeling.tex
edits in section 5, and conclusion. also a few more typos here and there

@@
 Foregrounds leak into the EoR H{\sc i} power spectrum by either having intrinsic spectral variability, high instrumental chromaticity, or through approximations in power spectrum analysis.
-%Here we focus on the instrumental effects while. 
+Here we use existing catalogs and a high fidelity instrumental model to focus on the possible sources of instrumental chromaticity.
 To decouple from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the individual antenna pair based approach of \citet{pob13,par14}.
@@
 \end{equation}
-The terms delay and lag are used interchangeably in this paper.
+Note that as the visibility is a fully complex quantity, its Fourier transform is fully Nyquist sampled over positive and negative delays.
+Negative and positive delay\footnote{The terms delay and lag are used interchangeably in this paper.} correspond to the two hemispheres of the sky as transected by the antenna pair's baseline vector.
+For example a perfectly East-West baseline will see a source in the Eastern sky at positive delay, and conversely for a source in the Western sky.
 
@@
 
-In order to make a quantitiative comparison of delay spectra obtained with the MWA data and our simulations, we consider the following uncertainties.
+In order to make a quantitative comparison of delay spectra obtained with the MWA data and our simulations, we consider the following uncertainties.
 Our foreground models are derived from other higher frequency catalogs and sky maps.
@@
 Although the model of \citet{deo08} yields a spectral index per direction on the sky, we assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from 408~MHz map of \citet{has82}.
-Thus, fractional errors in delay spectrum from compact foreground objects and diffuse emission are $\sim$0.7 and $\sim$0.3 respectively.
+Thus, fractional errors in delay spectrum from compact foreground objects and diffuse emission are $\sim$70\% and $\sim$30\% respectively.
 In addition, delay spectra from simulations and data each have uncertainties due to thermal noise with rms, $\Delta V^\textrm{N}_\eta(|\overline{\mathbf{x}}|,\eta) \sim$ 1.4$\times 10^7$~Jy~Hz, where the superscript N stands for thermal noise.
@@
 
-Figure~\ref{fig:data-sim-ratio} shows histograms of distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}.
+Figure~\ref{fig:data-sim-ratio} shows histograms of the distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}.
 The median absolute deviation from these distributions correspond to $\sim$~90\% fractional difference between data and modeling on average in either case.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
-\caption{Histogram of logarithm of ratio of observed to simulated visibilities restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+\caption{Histogram of logarithm of ratio of observed to simulated visibilities restricted to the {\it foreground wedge} --i.e.
+within the white lines of Figure \ref{fig:sim_data}-- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
 Median absolute deviations of $\log_{10}\rho$ are 0.29 and 0.27 respectively.
@@
 
-Having established that results from modeling match those from data, we proceed to examine in further detail the signatures seen in the modeled delay spectra.
+Having established that the model, by and large, matches the data, we proceed to examine in further detail the signatures seen in the modeled delay spectra.
 
@@
 This results in the ability to attribute the features in the delay spectrum to eight different sectors of the sky.
+%TODO possibly reference sections or figures here?
 
@@
 The bin centers are oriented towards South-East, East, North-East, and North respectively.
-Since delays depend on antenna pair orientation, binning by interferometer orientation allows us to match delay spectrum features to different sky directions.
+Since delays depend on antenna pair orientation, binning by orientation allows us to match delay spectrum features to different sky directions.
 
 \item {\it Tile Pointing and Power Pattern}: Since the MWA tiles are steered electronically, the power pattern of the tile changes in any observing mode that tracks the source.
-Our observations consist of a combination of allowing the sky to drift and tracking the sky.
-In our study, we take into account the effect of changing power pattern of the antennas on the delay spectrum (figure~\ref{fig:power_pattern}).
+The pointing system is capable of steering to points on a regular $\sim$7\arcdeg grid.
+During observations we allow the sky to drift across the nearest available pointing, shifting between grid points as necessary.
+In our study, we take into account the effect of changing power pattern of the antennas on the delay spectrum (Figure~\ref{fig:power_pattern}).
 The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose {\it off--zenith} and {\it zenith} pointings shown in figure~\ref{fig:pointings}.
 
-\item {\it Instrumental Passband}: The instrumental passband characteristics have a direct effect on the delay transform owing to a Fourier relationship between the two.
+\item {\it Instrumental Passband}: The instrumental passband characteristics have a direct effect on the delay transform owing to the Fourier relationship between the two.
 Due to excision of noisy and aliased edge channels repeated in each coarse channel of our passband, we saw repeated patterns of the {\it foreground wedge} in the delay spectra.
@@
 
-The foreground signatures we henceforth identify in the delay spectrum are characterized as arising out of the aforementioned factors.
+%The foreground signatures we henceforth identify in the delay spectrum are characterized as arising out of the aforementioned factors.
 Due to numerous combinations of such factors, we use unique annotations to characterize these signatures.
 
-It may be noted that numerous features overlap at varying levels of significance as a result of various combinations of parameters.
+It may be noted that numerous features overlap at varying levels of significance depending on the combinations of parameters.
 We assign the features to their predominant causes.
@@
 
-\subsection{Diffuse Foregrounds}\label{sec:diffuse}
+\subsection{Diffuse Foreground: Input model}\label{sec:diffuse}
 
@@
 
-\subsubsection{Diffuse Emission}\label{sec:diffuse-features}
+\subsubsection{Diffuse Emission Power Spectra}\label{sec:diffuse-features}
 
-Diffuse emission from the main lobe of the power patterns manifests as a branch at the top with positive delays and a zero--delay horizontal branch in the {\it off--zenith} and {\it zenith} pointings respectively, seen in figure~\ref{fig:noiseless-dsm-delay-spectrum}.
+As we also see in Figure~\ref{fig:noiseless-dsm-delay-spectrum}, diffuse emission from within the field of view manifests in the {\it off--zenith} pointing as a branch at the top of the wedge, while in the {\it zenith} pointing it shows up at near zero delay.
 The former is seen at positive delays because the main lobe of the power pattern is centered eastward of zenith, whereas in the latter it is centered at zenith.
-The horizontal arm in the {\it zenith} pointing fades away on antenna spacings wider than $\gtrsim 125$~m indicating that the diffuse emission is devoid of spatial structures on scales $\lesssim$~0\fdg 75.
+As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inverseley proportional to length of the baseline projected in the direction of emission.
+Thus, on the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at zero delay, fades away on antenna spacings wider than $\gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0\fdg 75.
+On the off-zenith pointing we see power out to much longer baselines.
 
@@
 
-Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the delay spectrum arising out of our compact objects sky model for {\it off--zenith} and {\it zenith} pointings (shown in figure~\ref{fig:CSM}) in the top and bottom panels respectively.
+Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the delay spectrum compact source sky model for {\it off--zenith} and {\it zenith} pointings (shown in Figure~\ref{fig:CSM}) in the top and bottom panels respectively.
 
@@
 White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
-The logarithmic color scale is identical to that in figure~\ref{fig:noiseless-dsm-delay-spectrum}.
+The logarithmic color scale is identical to that in Figure~\ref{fig:noiseless-dsm-delay-spectrum}.
 Emission from compact objects gives rise to centrally concentrated features in the {\it foreground wedge}.\label{fig:noiseless-csm-delay-spectrum}}
@@
 \subsubsection{Compact Emission in Power Pattern Main Lobe}\label{sec:csm-main-lobe}
-
-Since the main lobe of the power pattern in the {\it off--zenith} pointing is centered eastward of zenith, bulk of the compact foreground emission is seen in a branch with positive delay but short of the positive horizon delay limit.
-Compact emission from the same patch of sky is seen as a bright horizontal arm at zero delay in the {\it zenith} pointing since the main lobe of the power pattern is centered at zenith.
+The response of an interferometer to a compact source is, to first order, flat with respect to baseline length.
+Therefor the distribution of sources within the primary field of view depends mainly on how the beam pattern maps to delay space for each antenna pair.
+Since the main lobe of the power pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with positive delay corresponding to the position of the pointing center.
+In the {\it zenith} pointing compact emission from the same patch of sky is seen as a bright horizontal arm at zero delay  since the main lobe of the power pattern is centered at zenith.
 
@@
 
-\subsection{All--Sky Composite Foreground Model}\label{sec:composite}
+\subsection{Combined Foreground Model}\label{sec:composite}
 
@@
 
-Excluding the bright emission from the Galactic center which is heavily dependent on the sky location pointed to by the antenna, delay spectrum resulting from the all--sky composite foreground model consists of signatures already noted in the diffuse and compact foreground cases.
-This is illustrated by a microscopic examination of the {\it zenith} pointing in our study.
+When not dominated by the bright emission from the Galactic center, which happens only for certain pointings at certain times, the delay spectrum of the combined foreground model is dominated, in somewhat equal measure, by diffuse and compact emission.
+This is illustrated by a more detailed examination of the {\it zenith} pointing in our study.
 
-Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing.
+Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional power spectrum plots shown in Figures \ref{fig:fhd_data} et al.
 The diffuse, compact, and composite components are are shown as dotted, dashed, and solid lines respectively.
@@
 The gray shaded area denotes the envelope of expected uncertainty in the delay spectrum.
-This uncertainty is dominated by the spread in spectral index of compact foreground objects inside the {\it foreground} wedge (between the horizon delay limits) proprotional to the delay spectrum from compact objects, and by thermal noise outside the horizon delay limits.
+Sources inside the {\it foreground} wedge (between the horizon delay limits) are dominated by the $\sim$70\% error in predicting the spectral index of compact foreground objects, outside the wedge thermal noise dominates.
 
@@
 \includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
-\caption{Absolute value of delay spectrum, $|V_\eta(\overline{\mathbf{x}},\eta)|$, for three chosen antenna spacings, $\sim$84~m ({\it top}), $\sim$105~m ({\it middle}), and $\sim$170~m ({\it bottom}).
+\caption{Absolute value of delay spectrum, $|V_\eta(\overline{\mathbf{x}},\eta)|$, for three chosen antenna spacings of length: $\sim$84~m ({\it top}), $\sim$105~m ({\it middle}), and $\sim$170~m ({\it bottom}).
 The antenna spacing vector is specified in each panel along East, North and Upward directions.
@@
 Near the horizon delay limits at shorter antenna spacing, diffuse emission is brighter than that from compact objects.
-With increasing antenna spacing, although diffuse and compact components of emission become comparable, diffuse emission does not decrease as rapidly as seen at zero--delay.
-In fact, even with widely spaced antennas, diffuse emission in the delay spectrum near the horizon delay limits exceeds that in the main lobe by more than two orders of magnitude.
-This is due to apparent shortening of antenna spacing owing to projection effects while receiving emission from wide off--axis directions.
+Diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at zero--delay.
+In fact, even on widely spaced antennas, {\emph diffuse emission in the delay spectrum near the horizon delay limits exceeds that in the main lobe by more than two orders of magnitude}.
+This is due to apparent shortening of antenna spacings projected in the direction of smooth emission entering far from the primary field of view.
 This demonstrates the edge--heavy {\it two--pronged fork} feature discussed in \S\ref{sec:diffuse-long-baselines}.
 
-The resulting delay spectrum from a composite foreground model that combines diffuse and compact object emission consists of center--heavy features on all antenna spacings from compact objects and edge--heavy features from both components especially diffuse emission even on wide antenna spacings.
+The resulting baseline dependent power spectrum from a composite foreground model that combines diffuse and compact object emission consists of center--heavy features on all antenna spacings from compact objects and edge--heavy features from both components especially diffuse emission even on wide antenna spacings.
 Henceforth, we call this as the {\it pitchfork} structure in the {\it foreground} wedge.
 
-All these observations are consistent with those from figures~\ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}.
+All these observations are consistent with those from Figures~\ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}.
 The observability of the predicted {\it pitchfork} signature depends on the relative levels of uncertainty in the foreground model and thermal noise.
-In our simulations, since thermal noise is $\gtrsim 10^7$~Jy~Hz, and features near the horizon delay limits are $\sim 10^7$~Jy~Hz, the {\it pitchfork} feature is not expected to be detected in a noisy scenario although this feature is marginally visible in the {\it zenith} pointing of the observed data (see figure~\ref{fig:fhd_data}).
+In our simulations, since thermal noise in these very short snapshots is $\gtrsim 10^7$~Jy~Hz, and features near the horizon delay limits are $\sim 10^7$~Jy~Hz, the {\it pitchfork} feature is not expected to be detected in a noisy scenario although this feature is marginally visible in the {\it zenith} pointing of the observed data (see Figure~\ref{fig:fhd_data}).
 We attribute this to differences between our foreground model and the actual sky.
-However, with improvements to the foreground model and instrument configuration that can reduce thermal noise and other systematic uncertainties, we are confident the {\it pitchfork} signature in delay spectrum will be observed with high significance.
+However, with improvements to the foreground model and the addition of more data to reduce the thermal noise, we are confident the {\it pitchfork} signature in delay spectrum will be observed with high significance.
 
@@
 In short antenna spacings, due to relatively lower resolution along delay axis inside the {\it foreground wedge} and a consequent loss of localization of direction of foreground emission, these secondary peaks blend in with other major peaks and are not distinctly visible.
-These findings are consistent with dicussion in \S\ref{sec:csm-side-lobe}.
+These findings are consistent with the discussion in \S\ref{sec:csm-side-lobe}.
 
@@
 This leads to edge--heavy features in the delay spectrum.
-Complemented by the center-heavy compact foreground features, we have demosntrated the presence of a {\it three--pronged pitchfork}--shaped signature in the delay spectrum of foreground sky.
+Complemented by the center-heavy compact foreground features, we have demonstrated the presence of a {\it three--pronged pitchfork}--shaped signature in the delay spectrum of foreground sky.
 
@@
 The delay spectrum, $V_\eta(\overline{\mathbf{x}},\eta)$, not only carries information on spatial scales of foreground emission but also offers a unique advantage of viewing the sky through a combination of antenna spacing vector and the delay axis.
-This allows us to programmatically screen data for antenna spacings that are severly contaminated by foregrounds which can be weighted appropriately during data analysis.
+This allows us to programmatically screen data for antenna spacings that are severely contaminated by foregrounds which can be weighted appropriately during data analysis.
 While the {\it foreground wedge} is a region of high foreground contamination, of particular interest to the EoR community is foreground contamination near horizon delay limits.
@@
 The Galactic center is the most dominant source of foreground contamination in the {\it off--zenith} pointing.
-Since it is located in the western sky, it is most distinctly seen near the negative horizon delay limit on eastward oriented antenna spacings in the delay spectrum (see figures~\ref{fig:fhd_data}, \ref{fig:sim_data} and \ref{fig:noiseless-dsm-delay-spectrum}, and \S\ref{sec:GC-east}).
-In fact these figures illustrate that deconvolution along delay axis is severly hampered by this extreme contamination, thereby leading to leakage beyond the {\it foreground wedge} into the {\it EoR window}.
+Since it is located in the western sky, it is most distinctly seen near the negative horizon delay limit on eastward oriented antenna spacings in the delay spectrum (see Figures~\ref{fig:fhd_data}, \ref{fig:sim_data} and \ref{fig:noiseless-dsm-delay-spectrum}, and \S\ref{sec:GC-east}).
+In fact these figures illustrate quite well the leakage beyond the {\it foreground wedge} into the {\it EoR window} due to these extremely bright sources near the edge of the wedge.
 In practice, we could downweight these antenna spacings for the specific LST to mitigate the effects of spillover from such foreground contamination in EoR studies.
@@
 
-\section{Discussion}\label{sec:discussion}
+%\section{Discussion}\label{sec:discussion}
 
-\section{Conclusion}\label{sec:conclusion}
+%\section{Conclusion}\label{sec:conclusion}
 
@@
 
-Our primary motivation in this work is to analyze the signatures of foreground components in the measured delay spectrum.
-Such an analysis will be helpful in extending our knowledge on the challenge posed by foregrounds in studying the EoR and in devising techniques to mitigate contamination of EoR H{\sc i} signal by foreground objects emitting in radio wavelengths.
-This is one of the first studies that uses an all-sky foreground model to study the resultant signatures on EoR delay spectrum.
+Our primary motivation in this work is to understand how the various bright foregrounds will manifest in three dimensional power spectra of  Epoch of Reionization H{\sc i} observations.
+In the units presented here, the Epoch of Reionization is expected to be eight orders of magnitude below the foregrounds; a detailed understanding of how foregrounds can on the reionization power spectrum is therefor essential.
+This analysis extends previous work by simulating the entire sky rather then just the central field of view and by providing a comparison with the early observations from the MWA.
+Making use of the delay spectrum method to estimate the power spectrum we are able to observe the effects of foregrounds while avoiding entanglements with more complex power spectrum estimators.
 
-Using parameters that match the instrument and EoR observations using the MWA, and an all-sky foreground model that consists of diffuse emission from a ``global sky model'' and bright compact sources from the NVSS and SUMSS catalogs, we model delay spectra obtained with the MWA on antenna spacings of length $\le 200$~m.
-We confirm that the modeled delay spectra is in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise in measurements.
+Simulating in all important respects the response of the MWA to an all-sky foreground model that consists of diffuse emission from a ``global sky model'' and bright compact sources from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra is in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise in measurements.
 
-We use our simulations in a noiseless scenario to identify numerous signatures of different compoenents of foreground emission seen in the delay spectra.
+Looking deeper at our simulations, below the level of the instrumental noise,  we identify numerous signatures of different components of foreground emission seen in the delay spectra.
 We establish the relationship between these signatures and observing parameters such as antenna pointing and LST, instrument parameters such as antenna power pattern and bandpass shape, and foreground parameters such as the nature of emission, spectral index, etc.
@@
 
-On expected lines, diffuse emission in the primary beam of the antenna power pattern is prominent on shorter antenna spacings $\lesssim$~125~m corresponding to angular scales $\gtrsim$~0\fdg 75.
-However, the most interesting result is its footprint on wide antenna spacings near the horizon delay limits.
-This results in a edge--heavy {\it two--pronged fork}--shaped signature.
+As expected, diffuse emission in the primary beam of the antenna power pattern is prominent on shorter antenna spacings.
+However, the most interesting result is its footprint on wide antenna spacings near the horizon delay limits, an edge--heavy {\it two--pronged fork}--shaped signature.
 This is due to apparent shortening of wide antenna spacings in the direction of foreground emission far off--axis thereby retaining their response to extended emission.
@@
 
+In summary: We find that inclusion of sources, both diffuse and compact, all the way to the horizon is essential to explaining the observed power spectrum.  
 We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination in EoR data analysis.
-It is possible to screen antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation, which can then be assigned optimal weights in the EoR data analysis chain.
+It is possible to discard --or down-weight-- antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation.
 
@@
 We acknowledge the iVEC Petabyte Data Store, the Initiative in Innovative Computing and the CUDA Center for Excellence sponsored by NVIDIA at Harvard University, and the International Centre for Radio Astronomy Research (ICRAR), a Joint Venture of Curtin University and The University of Western Australia, funded by the Western Australian State government.
+DCJ acknowledges support by the National Science Foundation under award No.
+(AST-1401708).
 

### 2014-09-10 foreground_delay_spectrum_modeling.tex
From v0.4 (b4741f4) and 2e40b94 from Danny: ==========================================
(*) Created a branch danny2 and merged Danny's updated repository into this
branch
(*) Removed references to labels sec:discussion and sec:discussion to let the
file compile in LaTeX without warnings and modified those places accordingly.
(*) Minor edits throughout
(*) Changed structure slightly in section 3 to increase reading continuity

@@
 Perhaps the key challenge in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
-\citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness\citep{mor06,bow09,liu11,par12,dil13}.
+\citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}.
 Considerable effort has been made towards understanding the $k$-space behavior foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,dat10,bow09}.
-A general model of how smooth spectrum foregrounds are scattered to higher $k$-modes by the wide-field response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum.
-The conservative foreground strategy that has developed alongside this work is to discard $k$-modes which could be contaminated (Avoidance).
-The more aggressive alternative is to subtract a sky model and regain access to modes lost by Avoidance.
-In both cases, which parts of the sky are most critical to either Avoid or Subtract has remained largely uncertain.
-Here we will focus primarily on the Avoidance strategy, focusing on identifying foreground components most likely to ``leak'' from foreground modes to EoR modes.
+A general model of how smooth spectrum foregrounds are scattered to higher $k$--modes by the wide-field response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum.
+The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}).
+The more aggressive alternative is to subtract a sky model and regain access to modes lost by {\it avoidance}.
+In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain.
+Here we will focus primarily on the {\it avoidance} strategy, focusing on identifying foreground components most likely to ``leak'' from foreground modes to EoR modes.
 
-Foregrounds leak into the EoR H{\sc i} power spectrum by either having intrinsic spectral variability, high instrumental chromaticity, or through approximations in power spectrum analysis.
+Foregrounds leak into and contaminate the EoR H{\sc i} power spectrum by either having intrinsic spectral variability, high instrumental chromaticity, or through approximations in power spectrum analysis.
 Here we use existing catalogs and a high fidelity instrumental model to focus on the possible sources of instrumental chromaticity.
@@
 We quantify this contamination by breaking the simulations down by sky region and type of emission.
-Using these we arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to downweight badly contaminated baselines.
+Using these we arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.
 
-In \S\ref{sec:delay-spectrum} we introduce the delay spectrum technique, then in  \S\ref{sec:instrument}, \S\ref{sec:obsparms}, and \S\ref{sec:data-analysis} we describe the MWA setup, summarize the observing parameters and present the resulting data.
+In \S\ref{sec:delay-spectrum} we introduce the delay spectrum technique, then in  \S\ref{sec:instrument}, we describe the MWA setup, summarize the observing parameters and present the resulting data.
 Simulations of these observing parameters are described in \S\ref{sec:modeling} and analyzed by sky location, and type of emission in \S\ref{sec:delay-spectrum-analysis}.
-\S\ref{sec:discussion} offers a discussion along with a basic comparison to other current and future instruments.
-\S\ref{sec:conclusion} offers conclusions and an initial description of a more precise foreground avoidance technique.
+In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique.
+\S\ref{sec:summary} summarizes our work and findings.
 
@@
 $\overline{\mathbf{u}}$ is related to transverse spatial frequency modes ($\overline{\mathbf{k}}_\perp$) of brightness distribution of the sky as $\overline{\mathbf{k}}_\perp \equiv 2\pi\overline{\mathbf{u}}/D(z)$, where $D(z)$ is the transverse comoving distance at a redshift $z$.
-If $I(\overline{\mathbf{l}},f)$ is the emission at different frequencies on the sky as a function of direction unit vector ($\overline{\mathbf{l}}$) specified as direction cosines, $V_f(\overline{\mathbf{u}},f)$ represents the Fourier decomposition of $I(\overline{\mathbf{l}},f)$ attenuated by the antenna's angular power pattern $A(\overline{\mathbf{l}},f)$, into transverse spatial frequency modes.
+If $I(\overline{\mathbf{l}},f)$ is the emission at different frequencies on the sky as a function of direction unit vector ($\overline{\mathbf{l}}$) specified as direction cosines, $V_f(\overline{\mathbf{u}},f)$ represents the Fourier decomposition of $I(\overline{\mathbf{l}},f)$ attenuated by the antenna's angular power--pattern $A(\overline{\mathbf{l}},f)$, into transverse spatial frequency modes.
 Taking into account the instrumental bandpass weights $W_f(f)$, this relation can be written as:
@@
 Negative and positive delay\footnote{The terms delay and lag are used interchangeably in this paper.} correspond to the two hemispheres of the sky as transected by the antenna pair's baseline vector.
-For example a perfectly East-West baseline will see a source in the Eastern sky at positive delay, and conversely for a source in the Western sky.
+For example, a perfectly Eastward oriented baseline vector will obesrve an object in the Eastern sky at positive delay, and conversely for an object in the Western sky.
 
@@
 However, foreground objects emit in radio frequencies and contaminate the signals from the redshifted H{\sc i}.
-Due to their spatial and spectral properties, $V_\eta(\overline{\mathbf{u}},\eta)$ also contains contamination from foreground radio sources \citep{thy13,tro12,mor12,bow09}.
+Due to their spatial and spectral properties, $V_\eta(\overline{\mathbf{u}},\eta)$ also contains contamination from foreground radio objects \citep{thy13,tro12,mor12,bow09}.
 These contaminations have to be characterized precisely in order to reduce their impact on redshifted H{\sc i} power spectrum detection sensitivity.
@@
 In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30~MHz band anywhere in the range 80--300~MHz.
-Each tile is a phased array of 16 dipoles each in the shape of a bow--tie characterized by a grating response with a $\sim$~25\arcdeg wide primary field of view and multiple secondary side lobes.
+Each tile is a phased array of 16 dipoles each in the shape of a bow--tie characterized by a grating response with a $\gtrsim$~20\arcdeg wide primary field of view and multiple secondary side lobes.
 The tile arrangement is a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius with the longest baselines at 3~km \citep{bea12}.
@@
 The LSTs of these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at -1.92~hours (i.e., 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings respectively.
-%These pointings (phased array power patterns) are centered at RA = 4\fdg 387, Dec = -29\fdg 86 and RA = -28\fdg 8, Dec = -26\fdg 701, whereas the visibilities themselves are phased to zenith corresponding to RA = -28\fdg 8, Dec = -26\fdg 701, and RA = 1\fdg 35, Dec = -26\fdg 701 respectively.
+%These pointings (phased array power--patterns) are centered at RA = 4\fdg 387, Dec = -29\fdg 86 and RA = -28\fdg 8, Dec = -26\fdg 701, whereas the visibilities themselves are phased to zenith corresponding to RA = -28\fdg 8, Dec = -26\fdg 701, and RA = 1\fdg 35, Dec = -26\fdg 701 respectively.
 
@@
 \subsubsection{Initial Data processing}\label{sec:data-analysis}
-The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off12,off10}, removing 3\% of the data and averaged in time  and frequency from the raw 1/2s,40kHz to 2s, 80khz.
-These data are then calibrated to a simulation of the sky containing 2420 point sources selected from the MWA Commissioning Survey (N.H Walker, in prep).
+
+The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off12,off10}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz.
+These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey (MWACS; Hurley-Walker et al, in prep).
 MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning.
-It has a flux limit of 25~mJy and a the declination range of -12\arcdeg to -40\arcdeg evenly covering the field of view of the observations reported here.
-The calibration sources are selected to fall within the 5\% primary lobe of the tile beam.
-The calibration algorithm computes a per channel, per antenna, complex gain solutions averaged to two minute intervals.
-The solutions are still fairly low signal to noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna to antenna variation.
-First we average the channel gains over all antenna to obtain a high signal to noise measurement of the bandpass.
-After applying this single passband, we do a second round of calibration and fit a 2nd order amplitude and first order phase for each antenna.
+It has a flux density limit of 25~mJy and a declination range of -12\arcdeg to -40\arcdeg evenly covering the field of view of the observations reported here.
+The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern.
+The calibration algorithm computes complex gain solutions per channel per antenna averaged to two minute intervals.
+The solutions are still fairly low signal--to--noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation.
+First, we average the channel gains over all antenna to obtain a high signal--to--noise measurement of the bandpass.
+After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna.
 This flattens any residual variation in bandpass and removes small phase slopes due to variation in cable delay.
-Finally we fit for an additional phase known to be caused by reflection in 150m cables.
+Finally, we fit for an additional phase known to be caused by reflection in 150~m length cables.
 %possibly could add elaboration, if I knew exactly which version of the cal was done here.
@@
 We obtain the delay spectrum of these calibrated visibilities taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function.
-The spectrum is multiplied by  the additional weight of the flagged channels which occur rarely for interference and at at regular spacing every 1.28~MHz where the edges of coarse channels are known to have a slight aliasing effect.
-In delay space, these weights translate into a convolution by a ringing harmonic point spread function (PSF).
+The spectrum is multiplied by  the additional weight of the flagged channels which occur rarely for interference and at regular spacing every 1.28~MHz where the edges of coarse channels are known to have a slight aliasing effect.
+In delay--space, these weights translate into a convolution by a ringing harmonic point spread function (PSF).
 We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis by \citet{par09,par12}.
 The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights.
-We limit the selection of peaks to modes inside the delay horizon corresponding to smooth spectrum sources on the sky --ie beneath the black line in Figure \ref{fig:fhd_data}.
+We limit the selection of peaks to modes inside the horizon delay limits corresponding to smooth spectrum objects in the visible sky hemisphere.
 
@@
 
+In subsequent plots, we plot the amplitude of the delay spectrum, $|V(|\overline{\mathbf{x}}|,\eta)|$, sorted by baseline length, $|\overline{\mathbf{x}}|$.
+Note that visibilities from different baselines have not been averaged together in this analysis.
+
 We show results of delay spectrum from MWA data after deconvolution along delay axis in figure~\ref{fig:fhd_data}.
-In this plot and others, we plot the amplitude of the delay spectrum sorted by baseline length.
-Note that no baselines have been averaged together in this plot.
-The principle features are the wedge-shaped feature the boundary of which increases approximately linearly with light travel time of the baseline length.
+The principle feature is the wedge--shaped structure, the boundaries of which, marked by solid white lines, increases approximately ({\bf why exactly is it only approximate?}) linearly with light travel time corresponding to the baseline length.
 Note that the wedge is not filled out uniformly.
-Each baseline has a different orientation and resulting response to different sources in the sky.
+Each baseline has a different orientation and resulting response to different objects in the sky.
 It is this variation that we are investigating.
 
-In these power spectrum units [Mpc$^{-3}$], the reionization power spectrum is expected to be spherically symmetric and decrease quickly with $|k|$.
-The brightest signal will occur on the shortest baselines and smallest delays.
-Thus the region of interest is just above the black horizon line and below the harmonic due to the coarse channels.
+In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\overline{\mathbf{k}}|$.
+Hence, the brightest EoR signal will occur on the shortest baselines and smallest delays.
+Thus the region of interest for EoR studies is just beyond the horizon delay limits (solid white lines) and inside the first harmonic on either side ($|\eta| \lesssim 0.78\,\mu$s) due to the coarse channel structure.
+This forms a small region inside the so-called {\it EoR window}.
 
@@
 Consequently, the contamination into the {\it EoR window} is also higher.
-Emission in the {\it zenith} pointing is more centrally concentrated due to phasing of power pattern to the zenith, whereas the {\it off--zenith} pointing has a power pattern phased eastward of zenith and is also responsive to emission from angles far away from zenith.
+Emission in the {\it zenith} pointing is more centrally concentrated due to phasing of power--pattern to the zenith, whereas the {\it off--zenith} pointing has a power--pattern phased eastward of zenith and is also responsive to emission from angles far away from zenith.
 
@@
 White lines mark the boundaries of the {\it foreground wedge} determined by the light travel time across each baseline.
-Smooth spectrum sources are physically limited to the region below this line.
-The various features within the wedge depend on the \emph{orientation} of the baseline, the side lobe structure of the primary beam pointing and the sky at that time.
+Smooth spectrum objects are physically limited to the region below this line.
+The various features within the wedge depend on the \emph{orientation} of the baseline, the side lobe structure of the power--pattern and the sky at that instant.
 \label{fig:fhd_data}}
@@
 
-Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the data perfectly.
-This is predominant at short antenna spacings where the number of delay modes within the horizon.
+Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the measurements perfectly.
+This is predominant at short antenna spacings where the number of delay modes within the horizon is lesser.
 This is because the maximum delay envelope (boundary of the wedge) consists of mixed emission from large portions of the sky into a narrow range of delays and the deconvolution algorithm does not have sufficient support in delay space to fit the delay spectrum.
@@
 
-We describe power pattern and foreground models we have used in modeling the instrument. %todo
+We describe power--pattern and foreground models we have used in modeling the instrument. %todo
 
-\subsection{Power pattern}\label{sec:power_pattern}
+\subsection{power--pattern}\label{sec:power_pattern}
 
-The precise power pattern of the MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is a subject of active investigation.
-The most precise model currently available is that of Sutinjo et al (2014; submitted?) which models the 4$\times$4 array of bow-tie dipoles as a mutually-coupled array with the overall power-pattern of the dipole modeled via finite element electromagnetic simulation.
-To speed simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the FEKO simulation.
+The precise power--pattern of the MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is a subject of active investigation.
+The most precise model currently available is that of Sutinjo et al (2014; submitted?) which models the 4$\times$4 array of bow-tie dipoles as a mutually-coupled array with the overall power--pattern of the dipole modeled via finite element electromagnetic simulation.
+To speed simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the FEKO (or just aforementioned?) simulation.
 We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known stability level of the analog signal chain.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom.eps}
-\caption{Models of MWA tile power pattern in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates for the off--zenith (top) and zenith pointings (bottom) at 185~MHz.
+\caption{Models of MWA tile power--pattern in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates for the off--zenith (top) and zenith pointings (bottom) at 185~MHz.
 An MWA tile is modeled as a 4$\times$4 array of isotropic radiators placed 0.3~m above the ground plane.
@@
 The MWA has a very wide field of view ($\gtrsim 20$\arcdeg at 185~MHz) and has non-zero response all the way to the horizon.
-As we have seen above, foreground sources contaminate the region inside the ``wedge'' with an amplitude that depends on the primary beam amplitude in that direction, with sources at the horizon being closest to the edge of the wedge.
+As we have seen above, foreground objects contaminate the region inside the ``wedge'' with an amplitude that depends on the amplitude of power--pattern in that direction, with objects at the horizon being closest to the boundaries of the wedge.
 
-The degree to which these sources leak power into spectral line of sight modes above the wedge, a region known as the ``EoR Window'', depends primarily on the brightness of foregrounds near the edge of the wedge \citep{thy13,pob13,ved12,par12}.
-Thus, it is important to consider an all-sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting to the primary field of view.
+The degree to which these objects leak power into spectral line of sight modes above the wedge, a region known as the ``EoR Window'', depends primarily on the brightness of foregrounds near the boundaries of the wedge \citep{thy13,pob13,ved12,par12}.
+Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting to the primary field of view.
 
-\citet{bea13} estimated the sensitivity of the MWA to EoR H{\sc i} power spectrum detection taking into account thermal noise effects in the {\it Eor window}.
-\citet{thy13} estimated the sensitivity taking into account the spillover from foreground contamination from unsubtracted extragalactic point sources in the {\t EoR window} besides thermal noise.
+\citet{bea13} estimated the sensitivity of the MWA to EoR H{\sc i} power spectrum detection taking into account thermal noise effects in the {\it EoR window}.
+\citet{thy13} estimated the sensitivity taking into account the spillover from foreground contamination from unsubtracted extragalactic compact objects in the {\it EoR window} besides thermal noise.
 In this paper, we include the diffuse Galactic emission for enhancing our understanding of foreground signatures in the power spectrum.
-We use an all-sky foreground emission model that includes both diffuse and bright compact components.
+We use an all--sky foreground emission model that includes both diffuse and bright compact components.
 
@@
 
-For the diffuse component, we use an all-sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz.
-Since this map is predominantly based on the 408~MHz all-sky map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz diffuse emission map to the same resolution.
+For the diffuse component, we use an all--sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz.
+Since this map is predominantly based on the 408~MHz all--sky map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz diffuse emission map to the same resolution.
 However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals.
-Using the same model, we use the diffuse emission maps at 170~MHz and 200~MHz to estimate the spectral index map of the diffuse emission model.
+Using the same model, we use the diffuse emission maps at 170~MHz and 200~MHz to estimate the spectral index map of diffuse emission.
 
-A low resolution version of the diffuse foreground model used is show in figure~\ref{fig:DSM}.
-Contours of the MWA tile power pattern shown in figure~\ref{fig:power_pattern} are also overlaid.
+A low resolution version of the diffuse foreground model used is shown in figure~\ref{fig:DSM}.
+Contours of the MWA tile power--pattern shown in figure~\ref{fig:power_pattern} are also overlaid.
 Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the visible hemisphere in the westward sky and the MWA tile's power receptivity is significantly high ($\gtrsim 12$\%) in that direction.
-In the {\it zenith} pointing, the Galactic plane has almost set and the contour level of power pattern in that direction is at least 16 times smaller.
+In the {\it zenith} pointing, the Galactic plane has almost set and the contour level of power--pattern in that direction is at least 16 times smaller.
 
@@
 The color scale on the right is logarithmic and is common to both panels.
-Contours of power pattern shown in figure~\ref{fig:power_pattern} are overlaid.
+Contours of power--pattern shown in figure~\ref{fig:power_pattern} are overlaid.
 The contour levels shown are 0.00195, 0.00781, 0.0312, 0.125, and 0.5.
@@
 
-We use a combination of NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz due to their matched flux sensitivity and angular resolution, and complimentary survey footprints covering the entire sky.
+We use a combination of NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz due to their complimentary survey footprints covering the entire sky, and matched flux sensitivity and angular resolution.
 The SUMSS catalog covers the sky with declination $\delta < -30$\arcdeg with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec.
-The NVSS covers the sky with $\delta > -40$\arcdeg with a similar angular resolution and a limiting flux density of $\approx 2.5$~mJy for discrete sources.
+The NVSS covers the sky with $\delta > -40$\arcdeg with a similar angular resolution and a limiting flux density of $\approx 2.5$~mJy for compact objects.
 
-From the SUMSS catalog, we select compact sources whose deconvolved major axes are equal to 0\arcsec.
+From the SUMSS catalog, we select compact objects whose deconvolved major axes are equal to 0\arcsec.
 From the NVSS catalog, we excluded objects that overlap with those in the SUMSS survey footprint.
-Compact sources from NVSS were selected if the convolved major axes were not greater than $\approx 47$\arcsec, which is almost equal to the angular resolution of the survey.
+Compact objects from NVSS were selected if the convolved major axes were not greater than $\approx 47$\arcsec, which is almost equal to the angular resolution of the survey.
 Using a mean spectral index of $\langle\alpha_\textrm{sp}\rangle=-0.83$ (flux density, $S(f)\propto f^{\alpha_\textrm{sp}}$) obtained by \citet{mau03} for both NVSS and SUMSS catalog objcts, we calculate the corresponding flux densities at 185~MHz, $S_{185}$.
 From this subset, we choose compact objects with $S_{185}\geq 10$~Jy.
-The selection of such bright compact objects is not affected by minor differences in sensitivity of the two surveys.
-We verified that our selection criteria ensure a similar areal density of sources in the two surveys.
+The selection of such bright compact objects is not affected by minor differences in flux sensitivity of the two surveys.
+We verified that our selection criteria ensure a similar areal density of objects in the two surveys.
 
-Based on these criteria, we selected 133 sources from the SUMSS catalog and 336 sources from the NVSS catalog.
-Together with the diffuse foreground model, we obtain an all-sky foreground model consisting of both compact and diffuse emission.
+Based on these criteria, we selected 133 objects from the SUMSS catalog and 336 objects from the NVSS catalog.
+Together with the diffuse foreground model, we obtain an all--sky foreground model consisting of both compact and diffuse emission.
 
@@
 \includegraphics[width=\linewidth]{figures/v1_0/csm.eps}
-\caption{Flux densities (in Jy) of bright compact sources at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+\caption{Flux densities (in Jy) of bright compact objects at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
 The color scale on the right is logarithmic and is common to both panels and corresponds to the color--coded filled circles.
-Contours of power pattern shown in figure~\ref{fig:power_pattern} are overlaid.
+Contours of power--pattern shown in figure~\ref{fig:power_pattern} are overlaid.
 The contour levels shown are 0.001953125, 0.0078125, 0.03125, 0.125, and 0.5.
-The locations of bright compact sources appear to be random.\label{fig:CSM}}
+The locations of bright compact objects appear to be random.\label{fig:CSM}}
 \end{figure}
@@
 
-With the aforementioned all-sky foreground model, and instrumental and observational parameters, we simulate visibilities using equation~\ref{eqn:obsvis}.
+With the aforementioned all--sky foreground model, and instrumental and observational parameters, we simulate visibilities using equation~\ref{eqn:obsvis}.
 Figure~\ref{fig:sim_data} shows the amplitude of delay spectrum from {\it off--zenith} and {\it zenith} pointings.
@@
 Using simple error propagation, the fractional error in the delay spectrum caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the observed frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index.
-From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for compact sources from NVSS and SUMSS catalogs.
+From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for compact objects from NVSS and SUMSS catalogs.
 Although the model of \citet{deo08} yields a spectral index per direction on the sky, we assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from 408~MHz map of \citet{has82}.
@@
 These uncertainties are presented only to confirm the qualitative agreement already seen between figures~\ref{fig:fhd_data} and \ref{fig:sim_data} and are not intended to serve as a comprehensive estimate of all uncertainties involved.
-However, since we have not accounted for numerous other uncertainties, especially in the data, such as uncertainties in the antenna power pattern, calibration, radio frequency interference (RFI), and wide field imaging effects, our estimates are conservative.
+However, since we have not accounted for numerous other uncertainties, especially in the data, such as uncertainties in the antenna power--pattern, calibration, radio frequency interference (RFI), and wide field imaging effects, our estimates are conservative.
 Accounting for such uncertainties will only improve our confidence level in this comparison.
@@
 
-\item {\it Sky Model}: Our model of the sky consists of bright compact sources and diffuse emission on diverse spatial scales as seen from figures~\ref{fig:DSM} and \ref{fig:CSM}.
+\item {\it Sky Model}: Our model of the sky consists of bright compact objects and diffuse emission on diverse spatial scales as seen from figures~\ref{fig:DSM} and \ref{fig:CSM}.
 It also consists of localized regions of strong emission such as the Galactic plane and Galactic center.
@@
 
-\item {\it Tile Pointing and Power Pattern}: Since the MWA tiles are steered electronically, the power pattern of the tile changes in any observing mode that tracks the source.
+\item {\it Tile Pointing and Power--Pattern}: Since the MWA tiles are steered electronically, the power--pattern of the tile changes in any observing mode that tracks the object.
 The pointing system is capable of steering to points on a regular $\sim$7\arcdeg grid.
 During observations we allow the sky to drift across the nearest available pointing, shifting between grid points as necessary.
-In our study, we take into account the effect of changing power pattern of the antennas on the delay spectrum (Figure~\ref{fig:power_pattern}).
+In our study, we take into account the effect of changing power--pattern of the antennas on the delay spectrum (Figure~\ref{fig:power_pattern}).
 The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose {\it off--zenith} and {\it zenith} pointings shown in figure~\ref{fig:pointings}.
@@
 With the addition of noise in the visibilities which is subject to observing time, some of the weaker features may not be as prominently visible.
-Since the foreground signatures are far too numerous and they reside in a vast volume of parameter space such as antenna pair orientation, power pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features observed in the delay spectrum of foreground emission.
+Since the foreground signatures are far too numerous and they reside in a vast volume of parameter space such as antenna pair orientation, power--pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features observed in the delay spectrum of foreground emission.
 
-\subsection{Diffuse Foreground: Input model}\label{sec:diffuse}
+\subsection{Diffuse Foreground Signatures}\label{sec:diffuse}
 
@@
 
-The most prominent signature seen in the {\it off--zenith} pointing (top panel of figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the far west on one of the sidelobes of the power pattern.
+The most prominent signature seen in the {\it off--zenith} pointing (top panel of figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the far west on one of the sidelobes of the power--pattern.
 It appears as a bright branch in the delay spectrum near the negative delay horizon delay limit.
@@
 
-\subsubsection{Diffuse Emission Power Spectra}\label{sec:diffuse-features}
+\subsubsection{Diffuse Emission Delay Spectra}\label{sec:diffuse-features}
 
-As we also see in Figure~\ref{fig:noiseless-dsm-delay-spectrum}, diffuse emission from within the field of view manifests in the {\it off--zenith} pointing as a branch at the top of the wedge, while in the {\it zenith} pointing it shows up at near zero delay.
-The former is seen at positive delays because the main lobe of the power pattern is centered eastward of zenith, whereas in the latter it is centered at zenith.
+As we also see in Figure~\ref{fig:noiseless-dsm-delay-spectrum}, diffuse emission from within the primary field of view manifests in the {\it off--zenith} pointing as a branch at the top of the wedge, while in the {\it zenith} pointing it shows up at near zero delay.
+The former is seen at positive delays because the main lobe of the power--pattern is centered eastward of zenith, whereas in the latter it is centered at zenith.
 As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inverseley proportional to length of the baseline projected in the direction of emission.
-Thus, on the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at zero delay, fades away on antenna spacings wider than $\gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0\fdg 75.
-On the off-zenith pointing we see power out to much longer baselines.
+Thus, in the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at zero delay, fades away on antenna spacings $|\overline{\mathbf{x}}| \gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0\fdg 75.
+In the off-zenith pointing, we see power out to much longer baselines which is described below.
 
@@
 
-A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim \overline{\mathbf{x}}\le$~200~m) contrary to expectations, which are not expected to be sensitive to diffuse emission.
-We claim this is because wide antenna spacings appear shortened due to projection effects when sensitivity towards off--axis emission, especially towards the horizon, is concerned.
+A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\overline{\mathbf{x}}|\le$~200~m) contrary to expectations, which are not expected to be sensitive to diffuse emission.
+We claim this is because wide antenna spacings appear shortened due to projection effects when sensitivity towards off--axis emission, especially towards the horizon, is considered.
 It naturally holds for short antenna spacings as well.
@@
 
-\subsection{Compact Foregrounds}\label{sec:compact}
+\subsection{Compact Foreground Signatures}\label{sec:compact}
 
-Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the delay spectrum compact source sky model for {\it off--zenith} and {\it zenith} pointings (shown in Figure~\ref{fig:CSM}) in the top and bottom panels respectively.
+Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the delay spectrum of compact foreground sky model for {\it off--zenith} and {\it zenith} pointings (shown in Figure~\ref{fig:CSM}) in the top and bottom panels respectively.
 
@@
 
-\subsubsection{Compact Emission in Power Pattern Main Lobe}\label{sec:csm-main-lobe}
-The response of an interferometer to a compact source is, to first order, flat with respect to baseline length.
-Therefor the distribution of sources within the primary field of view depends mainly on how the beam pattern maps to delay space for each antenna pair.
-Since the main lobe of the power pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with positive delay corresponding to the position of the pointing center.
-In the {\it zenith} pointing compact emission from the same patch of sky is seen as a bright horizontal arm at zero delay  since the main lobe of the power pattern is centered at zenith.
+\subsubsection{Compact Emission in Power--Pattern Main Lobe}\label{sec:csm-main-lobe}
+The amplitude response of an interferometer to a compact object is, to first order, flat with respect to baseline length.
+Therefore, the distribution of objects within the primary field of view depends mainly on how the power--pattern maps to delay space for each antenna pair.
+Since the main lobe of the power--pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with positive delay corresponding to the position of the pointing center.
+In the {\it zenith} pointing compact emission from the same patch of sky is seen as a bright horizontal arm at zero delay  since the main lobe of the power--pattern is centered at zenith.
 
-\subsubsection{Compact Emission in Power Pattern Sidelobes}\label{sec:csm-side-lobe}
+\subsubsection{Compact Emission in Power--Pattern Sidelobes}\label{sec:csm-side-lobe}
 
-Foreground emission in zero delay and negative delay regions in the {\it off--zenith} pointing is caused by compact objects co--located with sidelobes of the power pattern.
-On the other hand, compact objects co--located with sidelobes of power pattern in the {\it zenith} pointing are revealed as distinct but faint branches at positive and negative delays depending on the orientation of antenna spacing and direction of emission on the sky.
+Foreground emission in zero delay and negative delay regions in the {\it off--zenith} pointing is caused by compact objects co--located with sidelobes of the power--pattern.
+On the other hand, compact objects co--located with sidelobes of power--pattern in the {\it zenith} pointing are revealed as faint but distinct branches at positive and negative delays depending on the orientation of antenna spacing and direction of emission on the sky.
 
-\subsection{Combined Foreground Model}\label{sec:composite}
+\subsection{Composite Foreground Signatures}\label{sec:composite}
 
-Delay spectra from the all-sky foreground model in our study display a composite feature set drawn from the features of compact and diffuse foreground models.
+Delay spectra from the all--sky foreground model in our study display a composite feature set drawn from the features of compact and diffuse foreground models.
 Here we compare the relative strengths of emission from different spatial scales in our all--sky composite foreground model.
@@
 
-Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional power spectrum plots shown in Figures \ref{fig:fhd_data} et al.
+Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional power spectrum plots shown in figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}.
 The diffuse, compact, and composite components are are shown as dotted, dashed, and solid lines respectively.
-The horizon delay limits are shown as vertical dot--dashed lines.
+The horizon delay limits are shown as a pair of vertical dot--dashed lines.
 The gray shaded area denotes the envelope of expected uncertainty in the delay spectrum.
-Sources inside the {\it foreground} wedge (between the horizon delay limits) are dominated by the $\sim$70\% error in predicting the spectral index of compact foreground objects, outside the wedge thermal noise dominates.
+Objects inside the {\it foreground} wedge (between the horizon delay limits) are dominated by the $\gtrsim$~70\% error owing to uncertainties in predicting the spectral index of compact foreground objects, outside the wedge thermal noise dominates.
 
@@
 
-The zero--delay peak (corresponding to the main lobe in the power pattern) with a value of $\sim 10^8$--$10^9$~Jy~Hz, independent of antenna spacing, is predominantly determined by compact foreground objects.
+The zero--delay peak (corresponding to the main lobe in the power--pattern) with a value of $\sim 10^8$--$10^9$~Jy~Hz, independent of antenna spacing, is predominantly determined by compact foreground objects.
 The corresponding peak at zero--delay from diffuse foreground model is at least 10 times fainter and decreases drastically with increase in antenna spacing.
@@
 The observability of the predicted {\it pitchfork} signature depends on the relative levels of uncertainty in the foreground model and thermal noise.
-In our simulations, since thermal noise in these very short snapshots is $\gtrsim 10^7$~Jy~Hz, and features near the horizon delay limits are $\sim 10^7$~Jy~Hz, the {\it pitchfork} feature is not expected to be detected in a noisy scenario although this feature is marginally visible in the {\it zenith} pointing of the observed data (see Figure~\ref{fig:fhd_data}).
+In our simulations, since thermal noise in these very short duration snapshots is $\gtrsim 10^7$~Jy~Hz, and features near the horizon delay limits are $\sim 10^7$~Jy~Hz, the {\it pitchfork} feature is not expected to be detected in a noisy scenario although this feature is marginally visible in the {\it zenith} pointing of the observed data (see Figure~\ref{fig:fhd_data}).
 We attribute this to differences between our foreground model and the actual sky.
-However, with improvements to the foreground model and the addition of more data to reduce the thermal noise, we are confident the {\it pitchfork} signature in delay spectrum will be observed with high significance.
+However, with improvements to the foreground model and the addition of more data to reduce thermal noise, we are confident the {\it pitchfork} signature in delay spectrum will be observed with high significance.
 
@@
 For instance, there is an increase in the number of secondary peaks in the delay spectrum between zero--delay and horizon delay limits as antenna spacing increases from $\sim$84~m to $\sim$170~m.
-In this case, these correspond to sidelobes of the power pattern that lie between the main lobe and the horizon along the North--South direction.
+In this case, these correspond to sidelobes of the power--pattern that lie between the main lobe and the horizon along the North--South direction.
 In short antenna spacings, due to relatively lower resolution along delay axis inside the {\it foreground wedge} and a consequent loss of localization of direction of foreground emission, these secondary peaks blend in with other major peaks and are not distinctly visible.
@@
 In summary, the brightest signature ($\gtrsim 10^{10}$~Jy~Hz) is that of the Galactic center in the {\it off--zenith} pointing co--located at a westward sidelobe with a significantly high gain.
-The next brightest signature ($\sim 10^8$--$\sim 10^9$~Jy~Hz) is caused by compact emission appearing to be concentrated in the inner regions of the foreground wedge (near zero delays) rather than towards the edges.
-Diffuse emission co--located at the main lobe of the power pattern is fainter by approximately an order of magnitude relative to compact emission from the same region for a $\sim 84$~m antenna spacing but unlike the latter, diffuse emission decreases rapidly by over two orders of magnitude as antenna spacing is widened to $\sim 170$~m.
-However, diffuse emission is significantly high near horizon delay limits compared to main lobe of the power pattern as antenna spacing is widened.
+The next brightest signature ($\sim 10^8$--$10^9$~Jy~Hz) is caused by compact emission appearing to be concentrated in the inner regions of the foreground wedge (near zero delays) rather than towards the edges.
+Diffuse emission co--located at the main lobe of the power--pattern is fainter by approximately an order of magnitude relative to compact emission from the same region for a $\sim 84$~m antenna spacing but unlike the latter, diffuse emission decreases rapidly by over two orders of magnitude as antenna spacing is widened to $\sim 170$~m.
+However, diffuse emission is significantly high near horizon delay limits compared to main lobe of the power--pattern as antenna spacing is widened.
 This leads to edge--heavy features in the delay spectrum.
@@
 
-With a foreground model known {\it a priori} in which locations of very bright foreground objects such as the Galactic center or AGN are available, we can predict their response across antenna spacings as a function of observing parameters such as LST, power pattern, etc.
+With a foreground model known {\it a priori} in which locations of very bright foreground objects such as the Galactic center or AGN are available, we can predict their response across antenna spacings as a function of observing parameters such as LST, power--pattern, etc.
 Specifically, we can pick antenna pairs where these bright foreground objects will be visible at delays near horizon delay limits.
@@
 Since it is located in the western sky, it is most distinctly seen near the negative horizon delay limit on eastward oriented antenna spacings in the delay spectrum (see Figures~\ref{fig:fhd_data}, \ref{fig:sim_data} and \ref{fig:noiseless-dsm-delay-spectrum}, and \S\ref{sec:GC-east}).
-In fact these figures illustrate quite well the leakage beyond the {\it foreground wedge} into the {\it EoR window} due to these extremely bright sources near the edge of the wedge.
+In fact these figures illustrate quite well the leakage beyond the {\it foreground wedge} into the {\it EoR window} due to these extremely bright objects near the boundaries of the wedge.
 In practice, we could downweight these antenna spacings for the specific LST to mitigate the effects of spillover from such foreground contamination in EoR studies.
@@
 
-Our primary motivation in this work is to understand how the various bright foregrounds will manifest in three dimensional power spectra of  Epoch of Reionization H{\sc i} observations.
-In the units presented here, the Epoch of Reionization is expected to be eight orders of magnitude below the foregrounds; a detailed understanding of how foregrounds can on the reionization power spectrum is therefor essential.
-This analysis extends previous work by simulating the entire sky rather then just the central field of view and by providing a comparison with the early observations from the MWA.
-Making use of the delay spectrum method to estimate the power spectrum we are able to observe the effects of foregrounds while avoiding entanglements with more complex power spectrum estimators.
+Our primary motivation in this work is to understand how the various bright foregrounds will manifest in three dimensional power spectrum of H{\sc i} from EoR observations.
+In the units presented here, the EoR H{\sc i} signal is expected to be eight orders of magnitude fainter than the foregrounds; a detailed understanding of how foregrounds can corrupt the EoR H{\sc i} power spectrum is therefore essential.
+This analysis extends previous work by simulating the entire sky rather then just the central field of view and by providing a comparison with early observations from the MWA.
+By making use of the delay spectrum technique to estimate power spectrum, we are able to observe the effects of foregrounds while avoiding entanglements with more complex power spectrum estimators.
 
-Simulating in all important respects the response of the MWA to an all-sky foreground model that consists of diffuse emission from a ``global sky model'' and bright compact sources from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra is in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise in measurements.
+Simulating in all important respects the response of the MWA to an all--sky foreground model that consists of diffuse emission from a ``global sky model'' and bright compact objects from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra is in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise in measurements.
 
 Looking deeper at our simulations, below the level of the instrumental noise,  we identify numerous signatures of different components of foreground emission seen in the delay spectra.
-We establish the relationship between these signatures and observing parameters such as antenna pointing and LST, instrument parameters such as antenna power pattern and bandpass shape, and foreground parameters such as the nature of emission, spectral index, etc.
+We establish the relationship between these signatures and observing parameters such as antenna pointing and LST, instrument parameters such as antenna power--pattern and bandpass shape, and foreground parameters such as the nature of emission, spectral index, etc.
 
-The bright Galactic center at the edge of the western horizon co--located with one of the far sidelobes of MWA tile power pattern is the brightest source of foreground contamination in the {\it off--zenith} pointing.
+The bright Galactic center at the edge of the western horizon co--located with one of the far sidelobes of MWA tile power--pattern is the brightest source of foreground contamination in the {\it off--zenith} pointing.
 It manifests itself near negative horizon delay limit in the delay spectrum on antenna spacings oriented eastward.
 
-As expected, diffuse emission in the primary beam of the antenna power pattern is prominent on shorter antenna spacings.
+As expected, diffuse emission in the primary beam of the antenna power--pattern is prominent on shorter antenna spacings.
 However, the most interesting result is its footprint on wide antenna spacings near the horizon delay limits, an edge--heavy {\it two--pronged fork}--shaped signature.
@@
 On the other hand, compact foreground objects predominantly map onto central regions of the {\it foreground wedge}.
-Features arising from compact sources coincident with primary beam and sidelobes of antenna power pattern have been noted.
-In general, compact objects produce delay spectrum signatures that are center--heavy, in clear contrast to those from diffuse emission.
+Features arising from compact objects coincident with primary beam and sidelobes of antenna power--pattern have been noted.
+In general, delay spectrum signatures of compact objects are center--heavy, in clear contrast to those from diffuse emission.
 A composite all--sky foreground model consisting of diffuse and compact foregrounds combines the characteristic individual shapes into a {\it pitchfork}--shaped structure in the {\it foreground wedge}.
@@
 
-In summary: We find that inclusion of sources, both diffuse and compact, all the way to the horizon is essential to explaining the observed power spectrum.  
+In conclusion, we find that inclusion of objects, both diffuse and compact, all the way to the horizon is essential to explaining the observed power spectrum.
 We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination in EoR data analysis.
-It is possible to discard --or down-weight-- antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation.
+It is possible to discard or down--weight antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation.
 

### 2014-09-10 foreground_delay_spectrum_modeling.tex
From f0b7176 on branch danny2: =============================
(*) Made few minor edits and removed some commented sentences

@@
 \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
-\caption{ Delay spectrum power spectra of two 112s  Murchison Widefield Array observations of the Epoch of Reionization field at RA=0, Dec=-30.
+\caption{ Delay spectra of two 112s  Murchison Widefield Array observations of the Epoch of Reionization field at RA=0, Dec=-30.
 At top an {\it off--zenith}  pointing when tiles have significant side lobe response in the direction of the galactic center.
 At bottom a {\it zenith} pointing where side lobes are minimal and the galaxy has completely set.
-To get the the power spectrum each baseline is Fourier transformed into delay space ($y$-axis) and the amplitude plotted as a logarithmic color in order of baseline length ($x$-axis).
+To get the the power spectrum each baseline is Fourier transformed into delay space ($y$--axis) and the amplitude plotted as a logarithmic color in order of baseline length ($x$--axis).
 The logarithmic color scale (shown at the right) is common to both panels.
@@
 \subsection{Foreground Model}\label{sec:foreground}
-%todo
+
 The MWA has a very wide field of view ($\gtrsim 20$\arcdeg at 185~MHz) and has non-zero response all the way to the horizon.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
-\caption{Histogram of logarithm of ratio of observed to simulated visibilities restricted to the {\it foreground wedge} --i.e.
-within the white lines of Figure \ref{fig:sim_data}-- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+\caption{Histogram of logarithm of ratio of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
 Median absolute deviations of $\log_{10}\rho$ are 0.29 and 0.27 respectively.
@@
 \end{itemize}
-
-%The foreground signatures we henceforth identify in the delay spectrum are characterized as arising out of the aforementioned factors.
-Due to numerous combinations of such factors, we use unique annotations to characterize these signatures.
 
@@
 \subsubsection{Compact Emission in Power--Pattern Main Lobe}\label{sec:csm-main-lobe}
+
 The amplitude response of an interferometer to a compact object is, to first order, flat with respect to baseline length.
@@
 
-%\section{Discussion}\label{sec:discussion}
-
-%\section{Conclusion}\label{sec:conclusion}
-
 \section{Summary}\label{sec:summary}

### 2014-09-11 foreground_delay_spectrum_modeling.tex
From v0.5: =========
(*) Added an entry to Sullivan et al.(2012) in the eor.bib file and cited this
reference in the tex file.
(*) Updated figures
(*) Removed empty appendix

@@
 The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern.
-The calibration algorithm computes complex gain solutions per channel per antenna averaged to two minute intervals.
+The calibration algorithm \citep{sul12} computes complex gain solutions per channel per antenna averaged to two minute intervals.
 The solutions are still fairly low signal--to--noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_visibilities_0.3m_ground_custom_gaussian_FG_model_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
 \caption{Absolute value of delay spectrum, $|V_\eta(\overline{\mathbf{x}},\eta)|$, for three chosen antenna spacings of length: $\sim$84~m ({\it top}), $\sim$105~m ({\it middle}), and $\sim$170~m ({\it bottom}).
@@
 
-\appendix
-
 \par\bigskip

### 2014-09-11 foreground_delay_spectrum_modeling.tex
From v0.5.1: ===========
(*) Mentioned cosmological parameters used in the paper.
(*) Wrote expression for relationship between k-parallel and eta
(*) Mentioned center frequency corresponds to redshift of 6.68
(*) Changed figure captions.

@@
 Since we are studying a redshifted H{\sc i} spectral line, $f$ is a measure of cosmological distance along the line of sight.
-Consequently, $\eta$ is a measure of line of sight spatial frequency mode, $k_\parallel$.
+Consequently, $\eta$ is a measure of line of sight spatial frequency mode, $k_\parallel$, given by $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
 Thus, $|V_\eta(\overline{\mathbf{u}},\eta)|^2$ is directly related to the spatial power spectrum of redshifted H{\sc i} distribution, $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$.
@@
 For purposes of illustration, we often use a simplified quantity $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$.
+
+Wherever applicable, we use $H_0=70$~km~s$^{-1}$~Mpc$^{-1}$ and assume a flat $\Lambda$CDM universe with $\Omega_\textrm{M}=0.27$ and $\Omega_\Lambda=0.73$.
 
@@
 Here we focus on antenna spacings $|\overline{\mathbf{x}}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz.
+The center frequency 185~MHz corresponds to a redshift, $z\simeq 6.68$.
 
@@
 To get the the power spectrum each baseline is Fourier transformed into delay space ($y$--axis) and the amplitude plotted as a logarithmic color in order of baseline length ($x$--axis).
+For reference, the $k$--space equivalent of $\eta$ and $|\overline{\mathbf{x}}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively.
 The logarithmic color scale (shown at the right) is common to both panels.
@@
 White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
+The axes and color scale are identical to those in figure~\ref{fig:fhd_data}.
 The features resemble those obtained with MWA data shown in figure~\ref{fig:fhd_data}.
 The Galactic center and Galactic plane prominently visible in the {\it off--zenith} pointing makes the emission in the {\it foreground wedge} brighter relative to that in the {\it zenith} pointing.
-The axes and color scale are identical to those in figure~\ref{fig:fhd_data}.
 \label{fig:sim_data}}
@@
 White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
-The logarithmic color scale is common to both panels.
-In the {\it zenith} pointing, diffuse emission delay spectrum has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings.
+The axes and color scale are identical to those in figures~\ref{fig:fhd_data} and \ref{fig:sim_data}.
+In the {\it zenith} pointing, delay spectrum from diffuse emission has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings.
 \label{fig:noiseless-dsm-delay-spectrum}}
@@
 White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
-The logarithmic color scale is identical to that in Figure~\ref{fig:noiseless-dsm-delay-spectrum}.
+The axes and logarithmic color scale are identical to those in figures~\ref{fig:fhd_data}, \ref{fig:sim_data} and \ref{fig:noiseless-dsm-delay-spectrum}.
 Emission from compact objects gives rise to centrally concentrated features in the {\it foreground wedge}.\label{fig:noiseless-csm-delay-spectrum}}

### 2014-09-12 foreground_delay_spectrum_modeling.tex
modified FHD citation sentence a little

--- 2014-09-11 foreground_delay_spectrum_modeling.tex
@@
 The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern.
-The calibration algorithm \citep{sul12} computes complex gain solutions per channel per antenna averaged to two minute intervals.
+The calibration algorithm --based on a forward modeling software by \citet{sul12}-- computes complex gain solutions per channel per antenna averaged to two minute intervals.
 The solutions are still fairly low signal--to--noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation.
@@
 We show results of delay spectrum from MWA data after deconvolution along delay axis in figure~\ref{fig:fhd_data}.
-The principle feature is the wedge--shaped structure, the boundaries of which, marked by solid white lines, increases approximately ({\bf why exactly is it only approximate?}) linearly with light travel time corresponding to the baseline length.
+The principle feature is the wedge--shaped structure, the boundaries of which, marked by solid white lines, increases linearly with light travel time corresponding to the baseline length.
 Note that the wedge is not filled out uniformly.

### 2014-09-12 foreground_delay_spectrum_modeling.tex
Figure should be capitalized

@@
 Any spectral flags, or other known weights constitutes $W_f(f)$ in equation~\ref{eqn:obsvis}.
-The bandpass properties used in this paper are illustrated in figure~\ref{fig:bandpass}.
+The bandpass properties used in this paper are illustrated in Figure~\ref{fig:bandpass}.
 
@@
 
-We show results of delay spectrum from MWA data after deconvolution along delay axis in figure~\ref{fig:fhd_data}.
+We show results of delay spectrum from MWA data after deconvolution along delay axis in Figure~\ref{fig:fhd_data}.
 The principle feature is the wedge--shaped structure, the boundaries of which, marked by solid white lines, increases linearly with light travel time corresponding to the baseline length.
@@
 
-A low resolution version of the diffuse foreground model used is shown in figure~\ref{fig:DSM}.
-Contours of the MWA tile power--pattern shown in figure~\ref{fig:power_pattern} are also overlaid.
+A low resolution version of the diffuse foreground model used is shown in Figure~\ref{fig:DSM}.
+Contours of the MWA tile power--pattern shown in Figure~\ref{fig:power_pattern} are also overlaid.
 Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the visible hemisphere in the westward sky and the MWA tile's power receptivity is significantly high ($\gtrsim 12$\%) in that direction.
@@
 The color scale on the right is logarithmic and is common to both panels.
-Contours of power--pattern shown in figure~\ref{fig:power_pattern} are overlaid.
+Contours of power--pattern shown in Figure~\ref{fig:power_pattern} are overlaid.
 The contour levels shown are 0.00195, 0.00781, 0.0312, 0.125, and 0.5.
@@
 The color scale on the right is logarithmic and is common to both panels and corresponds to the color--coded filled circles.
-Contours of power--pattern shown in figure~\ref{fig:power_pattern} are overlaid.
+Contours of power--pattern shown in Figure~\ref{fig:power_pattern} are overlaid.
 The contour levels shown are 0.001953125, 0.0078125, 0.03125, 0.125, and 0.5.
@@
 Figure~\ref{fig:sim_data} shows the amplitude of delay spectrum from {\it off--zenith} and {\it zenith} pointings.
-Notice the qualitative agreement of amplitude and structure with those obtained from data shown in figure~\ref{fig:fhd_data}.
+Notice the qualitative agreement of amplitude and structure with those obtained from data shown in Figure~\ref{fig:fhd_data}.
 The Galactic center and the Galactic plane visible in the {\it off--zenith} pointing make it appear brighter in the {\it foreground wedge}.
@@
 White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
-The features resemble those obtained with MWA data shown in figure~\ref{fig:fhd_data}.
+The features resemble those obtained with MWA data shown in Figure~\ref{fig:fhd_data}.
 The Galactic center and Galactic plane prominently visible in the {\it off--zenith} pointing makes the emission in the {\it foreground wedge} brighter relative to that in the {\it zenith} pointing.
-The axes and color scale are identical to those in figure~\ref{fig:fhd_data}.
+The axes and color scale are identical to those in Figure~\ref{fig:fhd_data}.
 \label{fig:sim_data}}
@@
 
-These uncertainties are presented only to confirm the qualitative agreement already seen between figures~\ref{fig:fhd_data} and \ref{fig:sim_data} and are not intended to serve as a comprehensive estimate of all uncertainties involved.
+These uncertainties are presented only to confirm the qualitative agreement already seen between Figures~\ref{fig:fhd_data} and \ref{fig:sim_data} and are not intended to serve as a comprehensive estimate of all uncertainties involved.
 However, since we have not accounted for numerous other uncertainties, especially in the data, such as uncertainties in the antenna power--pattern, calibration, radio frequency interference (RFI), and wide field imaging effects, our estimates are conservative.
@@
 
-\item {\it Sky Model}: Our model of the sky consists of bright compact objects and diffuse emission on diverse spatial scales as seen from figures~\ref{fig:DSM} and \ref{fig:CSM}.
+\item {\it Sky Model}: Our model of the sky consists of bright compact objects and diffuse emission on diverse spatial scales as seen from Figures~\ref{fig:DSM} and \ref{fig:CSM}.
 It also consists of localized regions of strong emission such as the Galactic plane and Galactic center.
@@
 In our study, we take into account the effect of changing power--pattern of the antennas on the delay spectrum (Figure~\ref{fig:power_pattern}).
-The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose {\it off--zenith} and {\it zenith} pointings shown in figure~\ref{fig:pointings}.
+The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose {\it off--zenith} and {\it zenith} pointings shown in Figure~\ref{fig:pointings}.
 
@@
 
-The most prominent signature seen in the {\it off--zenith} pointing (top panel of figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the far west on one of the sidelobes of the power--pattern.
+The most prominent signature seen in the {\it off--zenith} pointing (top panel of Figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the far west on one of the sidelobes of the power--pattern.
 It appears as a bright branch in the delay spectrum near the negative delay horizon delay limit.
 This feature is strongest at short antenna spacings and fades with increase in antenna spacing.
-This Galactic center signature is almost invisible in the {\it zenith} pointing because the Galactic center has already set over the horizon (see bottom panel in figure~\ref{fig:noiseless-dsm-delay-spectrum}).
+This Galactic center signature is almost invisible in the {\it zenith} pointing because the Galactic center has already set over the horizon (see bottom panel in Figure~\ref{fig:noiseless-dsm-delay-spectrum}).
 
@@
 
-Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional power spectrum plots shown in figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}.
+Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional power spectrum plots shown in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}.
 The diffuse, compact, and composite components are are shown as dotted, dashed, and solid lines respectively.
@@
 Since it is located in the western sky, it is most distinctly seen near the negative horizon delay limit on eastward oriented antenna spacings in the delay spectrum (see Figures~\ref{fig:fhd_data}, \ref{fig:sim_data} and \ref{fig:noiseless-dsm-delay-spectrum}, and \S\ref{sec:GC-east}).
-In fact these figures illustrate quite well the leakage beyond the {\it foreground wedge} into the {\it EoR window} due to these extremely bright objects near the boundaries of the wedge.
+In fact these Figures illustrate quite well the leakage beyond the {\it foreground wedge} into the {\it EoR window} due to these extremely bright objects near the boundaries of the wedge.
 In practice, we could downweight these antenna spacings for the specific LST to mitigate the effects of spillover from such foreground contamination in EoR studies.

### 2014-09-12 foreground_delay_spectrum_modeling.tex

@@
 The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern.
-The calibration algorithm --based on a forward modeling software by \citet{sul12}-- computes complex gain solutions per channel per antenna averaged to two minute intervals.
+The calibration algorithm -- based on a forward modeling software by \citet{sul12} -- computes complex gain solutions per channel per antenna averaged to two minute intervals.
 The solutions are still fairly low signal--to--noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation.
@@
 
-\subsection{power--pattern}\label{sec:power_pattern}
+\subsection{Power--pattern}\label{sec:power_pattern}
 
@@
 White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
+The axes and color scale are identical to those in Figure~\ref{fig:fhd_data}.
 The features resemble those obtained with MWA data shown in Figure~\ref{fig:fhd_data}.
 The Galactic center and Galactic plane prominently visible in the {\it off--zenith} pointing makes the emission in the {\it foreground wedge} brighter relative to that in the {\it zenith} pointing.
-The axes and color scale are identical to those in Figure~\ref{fig:fhd_data}.
 \label{fig:sim_data}}
@@
 This results in the ability to attribute the features in the delay spectrum to eight different sectors of the sky.
-%TODO possibly reference sections or figures here?
 
@@
 White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
-The logarithmic color scale is common to both panels.
-In the {\it zenith} pointing, diffuse emission delay spectrum has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings.
+The axes and color scale are identical to those in Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}.
+In the {\it zenith} pointing, delay spectrum from diffuse emission has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings.
 \label{fig:noiseless-dsm-delay-spectrum}}
@@
 White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
-The logarithmic color scale is identical to that in Figure~\ref{fig:noiseless-dsm-delay-spectrum}.
+The axes and logarithmic color scale are identical to those in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data} and \ref{fig:noiseless-dsm-delay-spectrum}.
 Emission from compact objects gives rise to centrally concentrated features in the {\it foreground wedge}.\label{fig:noiseless-csm-delay-spectrum}}
@@
 Since it is located in the western sky, it is most distinctly seen near the negative horizon delay limit on eastward oriented antenna spacings in the delay spectrum (see Figures~\ref{fig:fhd_data}, \ref{fig:sim_data} and \ref{fig:noiseless-dsm-delay-spectrum}, and \S\ref{sec:GC-east}).
-In fact these Figures illustrate quite well the leakage beyond the {\it foreground wedge} into the {\it EoR window} due to these extremely bright objects near the boundaries of the wedge.
+In fact these figures illustrate quite well the leakage beyond the {\it foreground wedge} into the {\it EoR window} due to these extremely bright objects near the boundaries of the wedge.
 In practice, we could downweight these antenna spacings for the specific LST to mitigate the effects of spillover from such foreground contamination in EoR studies.

### 2014-09-16 foreground_delay_spectrum_modeling.tex
Judd's initial edits

@@
 Detection of 21~cm H{\sc i} emission from the Epoch of Reionization (EoR), at redshifts $z>6$, is limited primarily by foreground emission.
-With much of the information coming from the spectral line that maps to distance along line of sight (cosmic time), foregrounds affecting the spectral domain deserve close attention.
-We investigate in detail the chromatic signatures of an all--sky foreground model imprinted by its spectral structure and instrumental transfer function.
-We use an antenna pair based delay spectrum technique.
-Besides confirming foreground contamination is predominantly restricted to a {\it wedge}--shaped region in Fourier space, we demonstrate that the foreground signatures which have the largest impact on the H{\sc i} signal via spillover outside the {\it wedge} arise from power received far away from the primary field of view.
-Comparing data from recent Murchison Widefield Array (MWA) observations with simulations broken down by sky position and type of emission, we identify diffuse emission near the horizon as the largest contributing factor even on wide antenna spacings relative to compact objects, hitherto unpredicted.
-These roles are switched in and around the primary field of view.
-This imprints a {\it pitchfork} on the {\it wedge} in Fourier space.
-The observation that the brightest contaminants are localized and only effect a fraction of the baselines leads us to propose that by down--weighting certain baselines at certain times, a large fraction of foreground contamination can be mitigated.
+We investigate the chromatic signatures of an all--sky foreground model imprinted by an instrumental transfer function.
+Using the delay spectrum technique, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal via spillover outside the {\it wedge} arise from power received far away from the primary field of view.
+Comparing data from recent Murchison Widefield Array (MWA) observations with simulations separated into components based on sky position and type of emission, we identify diffuse emission near the horizon as the largest contributing factor, even on wide antenna spacings.
+For signals entering through the primary field of view, compact objects dominate the foreground contamination.
+These two mechanisms imprint a characteristic {\it pitchfork} signature on the {\it wedge} in Fourier space.
+Based on these results, we propose that selective down--weighting of baselines based on length, direction, and time will mitigate a large fraction of foreground contamination in redshifted 21~cm power spectrum analysis.
 
@@
 Sensitive instruments such as the Square Kilometer Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}.
-Numerous precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum.
+Numerous precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13} {\bf tbd: and bowman et al.
+2013~}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum.
+{\bf  [JDB: add citations to Bowman et al.
+2006, Beardsley et al.
+2013, and Pober?]}
 
@@
 \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}.
-Considerable effort has been made towards understanding the $k$-space behavior foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,dat10,bow09}.
+Considerable effort has been made towards understanding the $k$-space behavior foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,dat10,bow09} {\bf [JDB: add citation to Liu et al.
+2009, 2010]}.
 A general model of how smooth spectrum foregrounds are scattered to higher $k$--modes by the wide-field response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum.
@@
 In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain.
-Here we will focus primarily on the {\it avoidance} strategy, focusing on identifying foreground components most likely to ``leak'' from foreground modes to EoR modes.
+Here we will focus primarily on extending the {\it avoidance} strategy by identifying foreground components most likely to ``leak'' from foreground modes to EoR modes and demonstrating a scheme for down--weighting these components.
 
-Foregrounds leak into and contaminate the EoR H{\sc i} power spectrum by either having intrinsic spectral variability, high instrumental chromaticity, or through approximations in power spectrum analysis.
-Here we use existing catalogs and a high fidelity instrumental model to focus on the possible sources of instrumental chromaticity.
-To decouple from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the individual antenna pair based approach of \citet{pob13,par14}.
-This approximates the power spectrum as the inverse fourier transform of the spectra generated by the correlator.
+Foregrounds contaminate the EoR H{\sc i} power spectrum by either having intrinsic spectral variability, high instrumental chromaticity, or through approximations in power spectrum analysis.
+Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity.
+To decouple our findings from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the individual baseline-based approach of \citet{pob13,par14}.
+This approximates the power spectrum as the inverse fourier transform of the spectra generated by the instrument's correlator.
 
- Matching real observations from the Murchison Widefield Array Epoch of Reionization observing program with instrumental simulations of the entire known sky we examine the signature of the power spectrum of each baseline.
-We confirm that the foregrounds most likely to obfuscate the Epoch of Reionization power are not caused by emission central field of view but rather from near the horizon.
-We also find that the smooth galactic emission plays a larger role then previously thought.
-We quantify this contamination by breaking the simulations down by sky region and type of emission.
+Matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the signature of the power spectrum of each baseline.
+We confirm {\bf [JDB: is confirm the right word, seems too weak}.
+Who else has published this?
+Are we the first to declare it?] that the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather from near the horizon.
+We find that diffuse Galactic emission plays a larger role then previously thought.
+We quantify this contamination by separating the simulations into componetes base on sky region and type of foreground emission.
 Using these we arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.
@@
 
+Throughout this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the telescope site.
+For the conventions of Fourier transform and its inverse we use in this paper, in accordance with equation~\ref{eqn:delay}, signals from the sky towards east and north are recorded with positive delays on eastward and northward oriented antenna spacings respectively.
+For instance, at the beginning of the observation, the Galactic center is in the westward sky just about to set.
+Signals from this direction are recorded with a negative delay.
+Similarly, signals from directions eastward of the local meridian arrive with positive delays.
+
+For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all antenna spacings to lie in the range $-67\fdg 5 \leq \theta_b < 112\fdg 5$.
+Antenna pairs with spacings oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
+
 Wherever applicable, we use $H_0=70$~km~s$^{-1}$~Mpc$^{-1}$ and assume a flat $\Lambda$CDM universe with $\Omega_\textrm{M}=0.27$ and $\Omega_\Lambda=0.73$.
 
-We now describe in detail an important ingredient, the MWA instrument.
+\section{Example Data}\label{sec:instrument}
 
-\section{Murchison Widefield Array Observations}\label{sec:instrument}
-
-\subsection{The MWA}
+\subsection{The Murchison Widefield Array (MWA)}
 
@@
 
-One important feature of the MWA's measured spectrum is the shape of its bandpass.
-The band is divided coarsely into 24 chunks with each chunk weighted by a digital filter.
+The MWA bandpass is divided coarsely into 24 chunks with each chunk weighted by a digital filter.
 The channels at the edges of these chunks have been flagged because they are known to possibly be contaminated by aliasing at a low level.
@@
 We take into account the increase in thermal uncertainty in frequency channels that results due to aforementioned bandpass correction.
-({\bf Since the bandpass correction is digital, shouldn't the noise stay flat after bandpass correction?})
+({\bf Since the bandpass correction is digital, shouldn't the noise stay flat after bandpass correction?
+-- JDB: yes})
 
@@
 
-The MWA Epoch of reionization targets two primary low-foreground fields.
-Here we choose to focus on RA = 0h, Dec = -30\arcdeg.
-The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of an array of dipoles in a MWA tile.
+The MWA  targets two primary low-foreground fields for reionization observations.
+Here, we focus on the field at RA = 0h, Dec = -30\arcdeg.
+The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of an array of dipoles in an MWA tile.
 These beamformer delay settings can be changed only in discrete steps.
@@
 
-The observations reported here were observed, near the beginning of the campaign, on Aug 23, 2013.
+The observations shown here were observed on Aug 23, 2013.
 This day has been chosen as a reference data set for comparison within the MWA EoR collaboration and was selected on the basis of low number of known system errors from among a long stretch of stable observing.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/custom_pointings.eps}
+\includegraphics[width=\linewidth]{figures/v1_0/custom_pointings}
 \caption{MWA tile beam pointing directions during the course of the observation.
@@
 
-\subsection*{A note on baselines/delay spectra sign convention}
-
-Throughout this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the MWA site.
-For the conventions of Fourier transform and its inverse we use in this paper, in accordance with equation~\ref{eqn:delay}, signals from the sky towards east and north are recorded with positive delays on eastward and northward oriented antenna spacings respectively.
-For instance, at the beginning of the observation, the Galactic center is in the westward sky just about to set.
-Signals from this direction are recorded with a negative delay on baselines whose antenna spacings are oriented towards the east.
-Similarly, signals from directions eastward of the local meridian arrive with positive delays on eastward oriented antenna spacings.
-
-For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all antenna spacings to lie in the range $-67\fdg 5 \leq \theta_b < 112\fdg 5$.
-Antenna pairs with spacings oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
-
 \subsubsection{Initial Data processing}\label{sec:data-analysis}
@@
 The solutions are still fairly low signal--to--noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation.
-First, we average the channel gains over all antenna to obtain a high signal--to--noise measurement of the bandpass.
+First, we average the channel gains over all antennas to obtain a high signal--to--noise measurement of the bandpass.
 After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna.
@@
 The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights.
-We limit the selection of peaks to modes inside the horizon delay limits corresponding to smooth spectrum objects in the visible sky hemisphere.
+We limit the selection of peaks to modes inside the horizon delay, corresponding to smooth spectrum objects in the visible sky hemisphere.
 
@@
 
-In subsequent plots, we plot the amplitude of the delay spectrum, $|V(|\overline{\mathbf{x}}|,\eta)|$, sorted by baseline length, $|\overline{\mathbf{x}}|$.
+In subsequent sections, we plot the amplitude of the delay spectrum, $|V(|\overline{\mathbf{x}}|,\eta)|$, sorted by baseline length, $|\overline{\mathbf{x}}|$.
 Note that visibilities from different baselines have not been averaged together in this analysis.
@@
 Each baseline has a different orientation and resulting response to different objects in the sky.
-It is this variation that we are investigating.
+It is this variation that we will explore below.
 
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
+\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots}
 \caption{ Delay spectra of two 112s  Murchison Widefield Array observations of the Epoch of Reionization field at RA=0, Dec=-30.
@@
 It is not our intention in this paper to use the deconvolution as a foreground removal tool.
-Hence, our discussion and conclusions are valid despite imperfections in deconvolution.
 
-\section{Instrument Modeling}\label{sec:modeling}
+\section{Simulations}\label{sec:modeling}
 
@@
 
-The precise power--pattern of the MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is a subject of active investigation.
-The most precise model currently available is that of Sutinjo et al (2014; submitted?) which models the 4$\times$4 array of bow-tie dipoles as a mutually-coupled array with the overall power--pattern of the dipole modeled via finite element electromagnetic simulation.
-To speed simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the FEKO (or just aforementioned?) simulation.
-We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known stability level of the analog signal chain.
+The power--pattern of the MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is modeled {\bf [TBD: cite Sutinjo et al (2014; submitted?)]} as a mutually-coupled 4-by-4 dipole array with the overall power--pattern of each inidivual dipole calculated via finite element electromagnetic simulation.
+To speed simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles.
+We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known repeatability and stability level of the analog signal chain {\bf [TBD: cite: bowman et al.
+2007 -- mwa prototype paper]}.
 
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom.eps}
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom}
 \caption{Models of MWA tile power--pattern in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates for the off--zenith (top) and zenith pointings (bottom) at 185~MHz.
@@
 
-The degree to which these objects leak power into spectral line of sight modes above the wedge, a region known as the ``EoR Window'', depends primarily on the brightness of foregrounds near the boundaries of the wedge \citep{thy13,pob13,ved12,par12}.
+The degree to which these objects leak power into spectral line of sight modes above the wedge, depends primarily on the brightness of foregrounds near the boundaries of the wedge \citep{thy13,pob13,ved12,par12}.
 Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting to the primary field of view.
 
-\citet{bea13} estimated the sensitivity of the MWA to EoR H{\sc i} power spectrum detection taking into account thermal noise effects in the {\it EoR window}.
+% JDB: I don't think we need this here.
+%% \citet{bea13} estimated the sensitivity of the MWA to EoR H{\sc i} power spectrum detection taking into account thermal noise effects in the {\it EoR window}.
 \citet{thy13} estimated the sensitivity taking into account the spillover from foreground contamination from unsubtracted extragalactic compact objects in the {\it EoR window} besides thermal noise.
 In this paper, we include the diffuse Galactic emission for enhancing our understanding of foreground signatures in the power spectrum.
+
 We use an all--sky foreground emission model that includes both diffuse and bright compact components.
@@
 However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals.
-Using the same model, we use the diffuse emission maps at 170~MHz and 200~MHz to estimate the spectral index map of diffuse emission.
+We model the spectral dependence of the diffuse foreground as a single spectral index at each pixel in our map, found by comparing model maps at 170~MHz and 200~MHz.
 
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/dsm.eps}
+\includegraphics[width=\linewidth]{figures/v1_0/dsm}
 \caption{Sky brightness temperature (in K) of the diffuse foreground model at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 
-We use a combination of NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz due to their complimentary survey footprints covering the entire sky, and matched flux sensitivity and angular resolution.
+We use a combination of NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz to provide our compact source catalog due to their complementary survey footprints covering the entire sky, and matched flux sensitivity and angular resolution.
 The SUMSS catalog covers the sky with declination $\delta < -30$\arcdeg with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/csm.eps}
+\includegraphics[width=\linewidth]{figures/v1_0/csm}
 \caption{Flux densities (in Jy) of bright compact objects at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
+\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots}
 \caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 
-In order to make a quantitative comparison of delay spectra obtained with the MWA data and our simulations, we consider the following uncertainties.
-Our foreground models are derived from other higher frequency catalogs and sky maps.
+In order to make a quantitative comparison of delay spectra obtained with the MWA data and our simulations, we consider the following uncertainties:
+
+{\bf Spectral index uncertainty.}  Our foreground models are derived from other higher frequency catalogs and sky maps.
 The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observed frequency.
@@
 Thus, fractional errors in delay spectrum from compact foreground objects and diffuse emission are $\sim$70\% and $\sim$30\% respectively.
-In addition, delay spectra from simulations and data each have uncertainties due to thermal noise with rms, $\Delta V^\textrm{N}_\eta(|\overline{\mathbf{x}}|,\eta) \sim$ 1.4$\times 10^7$~Jy~Hz, where the superscript N stands for thermal noise.
+
+{\bf Thermal uncertainty.} {\bf [JDB -- this is really a model component, not an uncertainy.
+We should probably move it to the earlier section.]}  In addition to intrinsic model uncertainy, delay spectra from simulations and data each have uncertainties due to thermal noise with rms, $\Delta V^\textrm{N}_\eta(|\overline{\mathbf{x}}|,\eta) \sim$ 1.4$\times 10^7$~Jy~Hz, where the superscript N stands for thermal noise.
 We estimate the ratio of delay spectra from data and simulations as, $\rho = |V^\textrm{D}_\eta(|\overline{\mathbf{x}}|,\eta)|\,/\,|V^\textrm{S}_\eta(|\overline{\mathbf{x}}|,\eta)|$, where superscripts D and S denote data and simulation respectively.
@@
 Currently we are significantly limited by unavailability of all--sky foreground models at the observed frequency of 185~MHz.
-Availability of an accurate foreground model will improve the agreement of simulations with observed data.
 
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0}
 \caption{Histogram of logarithm of ratio of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 \label{fig:data-sim-ratio}}
-% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
+% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0}
 % \caption{Significance of logarithm of ratio of observed to simulated visibilities relative to expected uncertainty restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 
-\section{Delay Spectrum Analysis}\label{sec:delay-spectrum-analysis}
+\section{Results}\label{sec:delay-spectrum-analysis}
 
-Having established that the model, by and large, matches the data, we proceed to examine in further detail the signatures seen in the modeled delay spectra.
+Having established that the mode matches the data to the level of expected uncertainties, we proceed to examine in further detail the key signatures seen in the modeled delay spectra.
+A number of factors are responsible for the characteristics noted in the delay spectra obtained from data and through modeling.
+We address these factors below: {\bf [JDB: These bullets below don't exactly match the analysis reported later -- we need to revise them so that they match one-to-one the results presented later.]}
 
-A number of factors are responsible for the characteristics noted in the delay spectra obtained from data and through modeling. We address these factors below:
 \begin{itemize}
@@
 In fact, the patch of sky for MWA observations is chosen from regions of low foreground emission.
-In our detailed study, we have divided the transiting sky into four different ``bow-tie'' shaped regions with equal areas around zenith as the origin.
-This results in the ability to attribute the features in the delay spectrum to eight different sectors of the sky.
+In our detailed study, we have divided the transiting sky into four different ``bow-tie'' shaped regions.
+Each region has equal area and the regions are aligned to cardinal directions around zenith as the origin.
+This results in the ability to isolate features in the delay spectrum to eight different sectors of the sky.
 
@@
 
-Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the amplitude of delay spectrum, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$, simulated with the diffuse foreground model.
+Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the amplitude of delay spectrum, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$, simulated with only the diffuse foreground model, omitting compact sources.
 The top and bottom panels correspond to {\it off--zenith} and {\it zenith} pointings respectively.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0}
 \caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
@@
 
-A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\overline{\mathbf{x}}|\le$~200~m) contrary to expectations, which are not expected to be sensitive to diffuse emission.
-We claim this is because wide antenna spacings appear shortened due to projection effects when sensitivity towards off--axis emission, especially towards the horizon, is considered.
-It naturally holds for short antenna spacings as well.
-Thus, diffuse emission from far off--axis directions manifests as an edge--heavy {\it two--pronged fork} across all antenna spacings.
-It decreases in strength with increasing antenna spacing as expected but is present in all orientations of the antenna spacing vector.
-This contributes to a {\it pitchfork}--shaped feature which will be discussed later in \S\ref{sec:pitchfork} in more detail.
+A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\overline{\mathbf{x}}|\le$~200~m)---contrary to expectations, since longer baselines are not expected to be sensitive to diffuse emission.
+In the widefield observations of the MA, long baselines appear shortened due to projection effects for off--axis emission, especially towards the horizon.
+This effect is evident at all LSTs in our simulations and it naturally holds for short baselines, as well.
+Thus, diffuse emission from far off--axis directions manifests as an edge--heavy {\it two--pronged fork} across all baselines.
+It decreases in strength with increasing baseline length as expected but is present in all orientations of the baseline vector.
+%% This contributes to a {\it pitchfork}--shaped feature which will be discussed later in \S\ref{sec:pitchfork} in more detail.
 
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0}
 \caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise.
@@
 
-\subsubsection{Compact Emission in Power--Pattern Main Lobe}\label{sec:csm-main-lobe}
+%\subsubsection{Compact Emission in Power--Pattern Main Lobe}\label{sec:csm-main-lobe}
 
@@
 
-\subsubsection{Compact Emission in Power--Pattern Sidelobes}\label{sec:csm-side-lobe}
+% \subsubsection{Compact Emission in Power--Pattern Sidelobes}\label{sec:csm-side-lobe}
 
@@
 
-\subsection{Composite Foreground Signatures}\label{sec:composite}
+\subsection{The "Pitchfork"}\label{sec:composite}
 
@@
 
-\subsubsection{{\it Pitchfork} Effect}\label{sec:pitchfork}
+% \subsubsection{{\it Pitchfork} Effect}\label{sec:pitchfork}
 
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_visibilities_0.3m_ground_custom_gaussian_FG_model_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_visibilities_0.3m_ground_custom_gaussian_FG_model_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0}
 \caption{Absolute value of delay spectrum, $|V_\eta(\overline{\mathbf{x}},\eta)|$, for three chosen antenna spacings of length: $\sim$84~m ({\it top}), $\sim$105~m ({\it middle}), and $\sim$170~m ({\it bottom}).
@@
 
-The resulting baseline dependent power spectrum from a composite foreground model that combines diffuse and compact object emission consists of center--heavy features on all antenna spacings from compact objects and edge--heavy features from both components especially diffuse emission even on wide antenna spacings.
-Henceforth, we call this as the {\it pitchfork} structure in the {\it foreground} wedge.
+Full simulations show the combination of center--heavy features on all antenna spacings from compact objects and edge--heavy features from both components especially diffuse emission even on wide antenna spacings.
+This results in a characteristic {\it pitchfork} structure in the {\it foreground} wedge that should be evident in observations.
 
@@
 We attribute this to differences between our foreground model and the actual sky.
-However, with improvements to the foreground model and the addition of more data to reduce thermal noise, we are confident the {\it pitchfork} signature in delay spectrum will be observed with high significance.
+Deeper observations should more clearly reveal the prominent feature.
 
@@
 This leads to edge--heavy features in the delay spectrum.
-Complemented by the center-heavy compact foreground features, we have demonstrated the presence of a {\it three--pronged pitchfork}--shaped signature in the delay spectrum of foreground sky.
+Complemented by the center-heavy compact foreground features, we predict the presence of a {\it three--pronged pitchfork}--shaped signature in the delay spectrum of foreground sky and identify evidence for the signature in actual MWA data, although deeper observations are needed to more clearly reveal the effect.
 
-\section{Screening of Severe Foreground Contaminated Interferometers}\label{sec:fg-grading}
+\section{Applications}\label{sec:fg-grading}
 
@@
 
-Our primary motivation in this work is to understand how the various bright foregrounds will manifest in three dimensional power spectrum of H{\sc i} from EoR observations.
-In the units presented here, the EoR H{\sc i} signal is expected to be eight orders of magnitude fainter than the foregrounds; a detailed understanding of how foregrounds can corrupt the EoR H{\sc i} power spectrum is therefore essential.
+Our primary motivation in this work is to understand how the various bright foregrounds will manifest in three dimensional power spectrum of H{\sc i} from 21~cm reionization observations.
+In the units presented here, the 21~cm signal is expected to be eight orders of magnitude fainter than the foregrounds; a detailed understanding of how foregrounds can corrupt the 21~cm power spectrum is therefore essential.
 This analysis extends previous work by simulating the entire sky rather then just the central field of view and by providing a comparison with early observations from the MWA.
@@
 
-Simulating in all important respects the response of the MWA to an all--sky foreground model that consists of diffuse emission from a ``global sky model'' and bright compact objects from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra is in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise in measurements.
+Simulating in all important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from [TBD: put actuall citation] and bright compact objects from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra is in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise in measurements.
 
@@
 A composite all--sky foreground model consisting of diffuse and compact foregrounds combines the characteristic individual shapes into a {\it pitchfork}--shaped structure in the {\it foreground wedge}.
-This will be distinctly visible when the thermal noise floor is sufficiently lowered.
+This will be distinctly visible when the thermal noise floor is sufficiently lowered, as longer observations are processed through the MWA analysis pipeline.
 
@@
 
-This scientific work makes use of the Murchison Radio-astronomy Observatory, operated by CSIRO.
+This work was supported by the U.
+S.
+National Science Foundation (NSF) through award AST--1109257.
+DCJ  is supported by an NSF Astronomy and Astrophysics Postdoctoral Fellowship under award AST--1401708.
+This work makes use of the Murchison Radio-astronomy Observatory, operated by CSIRO.
 We acknowledge the Wajarri Yamatji people as the traditional owners of the Observatory site.
-Support for the MWA comes from the U.S.
-National Science Foundation (grants AST-0457585, PHY-0835713, CAREER-0847753, and AST-0908884), the Australian Research Council (LIEF grants LE0775621 and LE0882938), the U.S.
+Support for the MWA comes from the NSF (awards: AST-0457585, PHY-0835713, CAREER-0847753, and AST-0908884), the Australian Research Council (LIEF grants LE0775621 and LE0882938), the U.S.
 Air Force Office of Scientific Research (grant FA9550-0510247), and the Centre for All-sky Astrophysics (an Australian Research Council Centre of Excellence funded by grant CE110001020).
@@
 We acknowledge the iVEC Petabyte Data Store, the Initiative in Innovative Computing and the CUDA Center for Excellence sponsored by NVIDIA at Harvard University, and the International Centre for Radio Astronomy Research (ICRAR), a Joint Venture of Curtin University and The University of Western Australia, funded by the Western Australian State government.
-DCJ acknowledges support by the National Science Foundation under award No.
-(AST-1401708).
 

### 2014-09-17 foreground_delay_spectrum_modeling.tex

@@
 Sensitive instruments such as the Square Kilometer Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}.
-Numerous precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13} {\bf tbd: and bowman et al.
-2013~}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum.
-{\bf  [JDB: add citations to Bowman et al.
-2006, Beardsley et al.
-2013, and Pober?]}
+Numerous precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{tin13,bow13,lon09}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{thy13,bea13,bow06,pob14}.
 
@@
 \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}.
-Considerable effort has been made towards understanding the $k$-space behavior foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,dat10,bow09} {\bf [JDB: add citation to Liu et al.
-2009, 2010]}.
+Considerable effort has been made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,liu11,dat10,bow09,liu09}.
 A general model of how smooth spectrum foregrounds are scattered to higher $k$--modes by the wide-field response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum.
@@
 To decouple our findings from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the individual baseline-based approach of \citet{pob13,par14}.
-This approximates the power spectrum as the inverse fourier transform of the spectra generated by the instrument's correlator.
+This approximates the power spectrum as the inverse Fourier transform of the spectra generated by the instrument's correlator.
 
@@
 We find that diffuse Galactic emission plays a larger role then previously thought.
-We quantify this contamination by separating the simulations into componetes base on sky region and type of foreground emission.
+We quantify this contamination by separating the simulations into components based on sky region and type of foreground emission.
 Using these we arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.
@@
 Negative and positive delay\footnote{The terms delay and lag are used interchangeably in this paper.} correspond to the two hemispheres of the sky as transected by the antenna pair's baseline vector.
-For example, a perfectly Eastward oriented baseline vector will obesrve an object in the Eastern sky at positive delay, and conversely for an object in the Western sky.
+For example, a perfectly Eastward oriented baseline vector will observe an object in the Eastern sky at positive delay, and conversely for an object in the Western sky.
 
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/bandpass_properties}
+\includegraphics[width=\linewidth]{figures/v1_0/bandpass_properties.eps}
 \caption{Bandpass properties used in modeling and analysis of delay spectrum of MWA visibilities.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/custom_pointings}
+\includegraphics[width=\linewidth]{figures/v1_0/custom_pointings.eps}
 \caption{MWA tile beam pointing directions during the course of the observation.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots}
+\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
 \caption{ Delay spectra of two 112s  Murchison Widefield Array observations of the Epoch of Reionization field at RA=0, Dec=-30.
@@
 
-The power--pattern of the MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is modeled {\bf [TBD: cite Sutinjo et al (2014; submitted?)]} as a mutually-coupled 4-by-4 dipole array with the overall power--pattern of each inidivual dipole calculated via finite element electromagnetic simulation.
+The power--pattern of the MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is modeled {\bf [TBD: cite Sutinjo et al (2014; submitted?)]} as a mutually-coupled 4-by-4 dipole array with the overall power--pattern of each individual dipole calculated via finite element electromagnetic simulation.
 To speed simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles.
-We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known repeatability and stability level of the analog signal chain {\bf [TBD: cite: bowman et al.
-2007 -- mwa prototype paper]}.
+We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known repeatability and stability level of the analog signal chain \citep{bow07b}.
 
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom}
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom.eps}
 \caption{Models of MWA tile power--pattern in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates for the off--zenith (top) and zenith pointings (bottom) at 185~MHz.
 An MWA tile is modeled as a 4$\times$4 array of isotropic radiators placed 0.3~m above the ground plane.
-Random fluctutations of rms 0.05~ns have been added to delay corrections during the phased addition of voltages from these isotropic radiators.
+Random fluctuations of rms 0.05~ns have been added to delay corrections during the phased addition of voltages from these isotropic radiators.
 The logarithmic color scale (shown at the right) is common to both panels.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/dsm}
+\includegraphics[width=\linewidth]{figures/v1_0/dsm.eps}
 \caption{Sky brightness temperature (in K) of the diffuse foreground model at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 Compact objects from NVSS were selected if the convolved major axes were not greater than $\approx 47$\arcsec, which is almost equal to the angular resolution of the survey.
-Using a mean spectral index of $\langle\alpha_\textrm{sp}\rangle=-0.83$ (flux density, $S(f)\propto f^{\alpha_\textrm{sp}}$) obtained by \citet{mau03} for both NVSS and SUMSS catalog objcts, we calculate the corresponding flux densities at 185~MHz, $S_{185}$.
+Using a mean spectral index of $\langle\alpha_\textrm{sp}\rangle=-0.83$ (flux density, $S(f)\propto f^{\alpha_\textrm{sp}}$) obtained by \citet{mau03} for both NVSS and SUMSS catalog objects, we calculate the corresponding flux densities at 185~MHz, $S_{185}$.
 From this subset, we choose compact objects with $S_{185}\geq 10$~Jy.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/csm}
+\includegraphics[width=\linewidth]{figures/v1_0/csm.eps}
 \caption{Flux densities (in Jy) of bright compact objects at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots}
+\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
 \caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 
-{\bf Thermal uncertainty.} {\bf [JDB -- this is really a model component, not an uncertainy.
-We should probably move it to the earlier section.]}  In addition to intrinsic model uncertainy, delay spectra from simulations and data each have uncertainties due to thermal noise with rms, $\Delta V^\textrm{N}_\eta(|\overline{\mathbf{x}}|,\eta) \sim$ 1.4$\times 10^7$~Jy~Hz, where the superscript N stands for thermal noise.
+{\bf Thermal uncertainty.} {\bf [JDB -- this is really a model component, not an uncertainty.
+We should probably move it to the earlier section.]}  In addition to intrinsic model uncertainty, delay spectra from simulations and data each have uncertainties due to thermal noise with rms, $\Delta V^\textrm{N}_\eta(|\overline{\mathbf{x}}|,\eta) \sim$ 1.4$\times 10^7$~Jy~Hz, where the superscript N stands for thermal noise.
 We estimate the ratio of delay spectra from data and simulations as, $\rho = |V^\textrm{D}_\eta(|\overline{\mathbf{x}}|,\eta)|\,/\,|V^\textrm{S}_\eta(|\overline{\mathbf{x}}|,\eta)|$, where superscripts D and S denote data and simulation respectively.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0}
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
 \caption{Histogram of logarithm of ratio of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 \label{fig:data-sim-ratio}}
-% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0}
+% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
 % \caption{Significance of logarithm of ratio of observed to simulated visibilities relative to expected uncertainty restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0}
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
 \caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
@@
 The former is seen at positive delays because the main lobe of the power--pattern is centered eastward of zenith, whereas in the latter it is centered at zenith.
-As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inverseley proportional to length of the baseline projected in the direction of emission.
+As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inversely proportional to length of the baseline projected in the direction of emission.
 Thus, in the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at zero delay, fades away on antenna spacings $|\overline{\mathbf{x}}| \gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0\fdg 75.
@@
 A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\overline{\mathbf{x}}|\le$~200~m)---contrary to expectations, since longer baselines are not expected to be sensitive to diffuse emission.
-In the widefield observations of the MA, long baselines appear shortened due to projection effects for off--axis emission, especially towards the horizon.
+In the Widefield observations of the MA, long baselines appear shortened due to projection effects for off--axis emission, especially towards the horizon.
 This effect is evident at all LSTs in our simulations and it naturally holds for short baselines, as well.
@@
 It decreases in strength with increasing baseline length as expected but is present in all orientations of the baseline vector.
-%% This contributes to a {\it pitchfork}--shaped feature which will be discussed later in \S\ref{sec:pitchfork} in more detail.
+%% This contributes to a {\it pitchfork}--shaped feature which will be discussed later in \S\ref{sec:composite} in more detail.
 
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0}
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
 \caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise.
@@
 In contrast to delay spectrum of diffuse emission, compact emission manifests as a center--heavy structure in the delay spectrum with either pointing.
-This contributes to the {\it pitchfork} feature which will be discussed in detail in \S\ref{sec:pitchfork}.
+This contributes to the {\it pitchfork} feature which will be discussed in detail in \S\ref{sec:composite}.
 Below we summarize some specific notable features obtained with our compact objects foreground model.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_visibilities_0.3m_ground_custom_gaussian_FG_model_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0}
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_visibilities_0.3m_ground_custom_gaussian_FG_model_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
 \caption{Absolute value of delay spectrum, $|V_\eta(\overline{\mathbf{x}},\eta)|$, for three chosen antenna spacings of length: $\sim$84~m ({\it top}), $\sim$105~m ({\it middle}), and $\sim$170~m ({\it bottom}).
@@
 In short antenna spacings, due to relatively lower resolution along delay axis inside the {\it foreground wedge} and a consequent loss of localization of direction of foreground emission, these secondary peaks blend in with other major peaks and are not distinctly visible.
-These findings are consistent with the discussion in \S\ref{sec:csm-side-lobe}.
+These findings are consistent with the discussion in \S\ref{sec:compact}.
 
@@
 
-Simulating in all important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from [TBD: put actuall citation] and bright compact objects from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra is in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise in measurements.
+Simulating in all important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from [TBD: put actual citation] and bright compact objects from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra is in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise in measurements.
 

### 2014-09-19 foreground_delay_spectrum_modeling.tex
From d78e0ca: ============
(*) Made minor change to title
(*) Added new figures
(*) Changed wording as Judd suggested
(*) Changed Tsys to 95 K and made corresponding changes to all figures
(*) Added reference(s) to eor.bib and cited them in text
(*) Introduction and nomenclature on 'baseline' yet to be reworded

@@
 %% \title{Limits on the detection of the Epoch of Reionization from MWA observations of the redshifted 21~cm line}
-\title{Signatures of Foreground Sky in Power Spectra of Redshifted Neutral Hydrogen from the Epoch of Reionization}
+\title{Signatures of Foreground Sky in Power Spectrum of Redshifted Neutral Hydrogen from the Epoch of Reionization}
 
@@
 Considerable effort has been made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,liu11,dat10,bow09,liu09}.
-A general model of how smooth spectrum foregrounds are scattered to higher $k$--modes by the wide-field response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum.
+A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide-field response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum.
 The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}).
@@
 Matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the signature of the power spectrum of each baseline.
-We confirm {\bf [JDB: is confirm the right word, seems too weak}.
-Who else has published this?
-Are we the first to declare it?] that the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather from near the horizon.
+We find that the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon.
 We find that diffuse Galactic emission plays a larger role then previously thought.
-We quantify this contamination by separating the simulations into components based on sky region and type of foreground emission.
+We quantify this contamination by separating the simulations into components based on type of foreground emission.
 Using these we arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.
@@
 Consequently, $\eta$ is a measure of line of sight spatial frequency mode, $k_\parallel$, given by $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
-Thus, $|V_\eta(\overline{\mathbf{u}},\eta)|^2$ is directly related to the spatial power spectrum of redshifted H{\sc i} distribution, $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$.
+We use $H_0=100\,h$~km~s$^{-1}$~Mpc$^{-1}$ and assume a flat $\Lambda$CDM universe with $\Omega_\textrm{M}=0.27$ and $\Omega_\Lambda=0.73$.
+
+Thus, $|V_\eta(\overline{\mathbf{u}},\eta)|^2$ is directly related to the spatial power spectrum of redshifted H{\sc i} distribution, $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ \citep{mor04}.
 However, foreground objects emit in radio frequencies and contaminate the signals from the redshifted H{\sc i}.
@@
 
+Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of H{\sc i} from the EoR are presented in this paper.
+$|\overline{\mathbf{x}}|$ and $k_\perp$, denoting spatial scales in the transverse direction, form the $x$--axis.
+$\eta$ and $k_\parallel$, denoting spatial scales along line of sight form the $y$--axis.
+Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge}, the boundaries of which are determined by antenna spacing and the light travel time across it.
+These boundaries are called horizon delay limits and are shown by solid lines.
+Spectral response of the instrument convolves the foreground wedge and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\eta$--axis.
+This effect has been quantified in \citet{thy13}.
+The width of this narrow strip is inversely proportional to operating bandwidth.
+Region of Fourier space excluding the {\it foreground wedge} and also the narrow strips is usually called the {\it EoR window}.
+These are shown in light and medium shades.
+However, the instrument used in our study, namely the MWA, has a passband constructed using coarse channels with a characteristic shape that results in harmonic repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.
+The underlying cause of this effect will be discussed later in this paper.
+Hence, we have identified a region of high significance (shown in medium shade) that forms a subset of the {\it EoR window} bounded by $|\eta|\le 0.78\,\mu$s where contamination from foregrounds is expected to be low and yet the EoR H{\sc i} power spectrum signal is reasonably strong.
+Hereafter, we refer to this region as the {\it EoR window} in the rest of this paper.
+The MWA EoR collaboration actively uses this as a test region in Fourier space for diagnostics on data quality.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\linewidth]{figures/v1_0/fourier_space_185.0_MHz_30.7_MHz.eps}
+\caption{Illustration of Fourier space in which delay (and power) spectra of H{\sc i} signals from the EoR are presented.
+The $x$--axis denoted by $|\overline{\mathbf{x}}|$ (antenna spacing in m) or $k_\perp$ (transverse wavenumber in units of $h$~Mpc$^{-1}$) are equivalent and represent transverse spatial scales.
+Similarly, the $y$--axis denoted by $\eta$ (delay in s or $\mu$s) or equivalently, $k_\parallel$ (line--of--sight wavenumber in units of $h$~Mpc$^{-1}$) represents spatial scales along the line of sight.
+The dark shaded region is often referred to as {\it foreground wedge} where smooth spectrum foregrounds reside.
+The boundaries of the {\it foreground wedge} (solid lines) are given by the light travel time for corresponding antenna spacings and is referred to as the horizon delay limits.
+In principle, region outside the wedge is expected to be relatively free of foreground contamination and is called the {\it EoR window} (light and medium shades).
+In practice, this contamination is also high in the narrow unshaded white regions (bound by solid and dashed lines) due to convolution by instrumental transfer function.
+The width of this narrow region is determined by inverse of operating bandwidth.
+Due to an undesirable spectral response of the instrument used in this study, namely the MWA, we identify a narrower region of high significance to EoR studies (medium shade) and hereafter refer to it as the {\it EoR window} in this paper.
+\label{fig:fourier-space}}
+\end{figure}
+
 Throughout this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the telescope site.
@@
 For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all antenna spacings to lie in the range $-67\fdg 5 \leq \theta_b < 112\fdg 5$.
-Antenna pairs with spacings oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
-
-Wherever applicable, we use $H_0=70$~km~s$^{-1}$~Mpc$^{-1}$ and assume a flat $\Lambda$CDM universe with $\Omega_\textrm{M}=0.27$ and $\Omega_\Lambda=0.73$.
+Antenna spacings oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
 
@@
 
-Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=90$~K per polarization for all frequency channels for all antenna pairs throughout the course of observation.
-We take into account the increase in thermal uncertainty in frequency channels that results due to aforementioned bandpass correction.
+Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=95$~K per polarization for all frequency channels for all antenna pairs throughout the course of the observation.
+% We take into account the increase in thermal uncertainty in frequency channels that results due to aforementioned bandpass correction.
 ({\bf Since the bandpass correction is digital, shouldn't the noise stay flat after bandpass correction?
@@
 These were chosen to provide a selection of possible foreground and instrumental conditions.
-As an example of ``nominal'' observing setup we choose a zenith pointing; as an example of poor foreground conditions we choose a pointing when the field is 2 hours from zenith.
+As an example of ``nominal'' observing setup we choose a zenith pointing; as an example of poor foreground conditions, we choose a pointing when the field is $\sim 2$~hours from zenith.
 This pointing has a significantly higher secondary side lobe structure and is recorded when the bright galactic center is well above the horizon.
@@
 \caption{MWA tile beam pointing directions during the course of the observation.
-The x-axis refers to the Local Sidereal Time (LST) in hours.
+The $x$--axis refers to the Local Sidereal Time (LST) in hours.
 The axis on the left refers to longitudes, namely, Right Ascension (RA) and Hour Angle (HA) in degrees.
@@
 The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off12,off10}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz.
-These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey (MWACS; Hurley-Walker et al, in prep).
+These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey (MWACS; Hurley-Walker et al, submitted).
 MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning.
@@
 The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights.
-We limit the selection of peaks to modes inside the horizon delay, corresponding to smooth spectrum objects in the visible sky hemisphere.
+We limit the selection of peaks to modes inside the horizon delay limit, corresponding to smooth spectrum objects in the visible sky hemisphere.
 
@@
 Hence, the brightest EoR signal will occur on the shortest baselines and smallest delays.
-Thus the region of interest for EoR studies is just beyond the horizon delay limits (solid white lines) and inside the first harmonic on either side ($|\eta| \lesssim 0.78\,\mu$s) due to the coarse channel structure.
-This forms a small region inside the so-called {\it EoR window}.
+Thus the region of interest for EoR studies is slightly beyond the horizon delay limits (solid white lines) and inside the first harmonic on either side ($|\eta| \lesssim 0.78\,\mu$s) due to the coarse channel structure.
+We call this region as the {\it EoR window} in this paper.
 
@@
 \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
-\caption{ Delay spectra of two 112s  Murchison Widefield Array observations of the Epoch of Reionization field at RA=0, Dec=-30.
-At top an {\it off--zenith}  pointing when tiles have significant side lobe response in the direction of the galactic center.
-At bottom a {\it zenith} pointing where side lobes are minimal and the galaxy has completely set.
-To get the the power spectrum each baseline is Fourier transformed into delay space ($y$--axis) and the amplitude plotted as a logarithmic color in order of baseline length ($x$--axis).
+\caption{Delay spectra of two 112~s Murchison Widefield Array observations of the Epoch of Reionization field at RA = 0h, Dec = -30\arcdeg at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has completely set.
+To get the delay spectrum, each baseline is Fourier transformed into delay space ($y$--axis) and the amplitude plotted in order of baseline length ($x$--axis).
 For reference, the $k$--space equivalent of $\eta$ and $|\overline{\mathbf{x}}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively.
@@
 Smooth spectrum objects are physically limited to the region below this line.
-The various features within the wedge depend on the \emph{orientation} of the baseline, the side lobe structure of the power--pattern and the sky at that instant.
+The various features within the {\it foreground wedge} depend on the \emph{orientation} of the antenna spacing, the secondary lobe structure of the power--pattern and the sky at that instant.
 \label{fig:fhd_data}}
@@
 
-The power--pattern of the MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is modeled {\bf [TBD: cite Sutinjo et al (2014; submitted?)]} as a mutually-coupled 4-by-4 dipole array with the overall power--pattern of each individual dipole calculated via finite element electromagnetic simulation.
-To speed simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles.
+The power--pattern of the MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is modeled as a mutually--coupled 4--by--4 dipole array with the overall power--pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14}.
+To speed up simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles.
 We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known repeatability and stability level of the analog signal chain \citep{bow07b}.
@@
 
-The MWA has a very wide field of view ($\gtrsim 20$\arcdeg at 185~MHz) and has non-zero response all the way to the horizon.
-As we have seen above, foreground objects contaminate the region inside the ``wedge'' with an amplitude that depends on the amplitude of power--pattern in that direction, with objects at the horizon being closest to the boundaries of the wedge.
+The MWA has a very wide field of view ($\gtrsim 20$\arcdeg at 185~MHz) and has non--zero response all the way to the horizon.
+As we have seen above, foreground objects contaminate the region inside the {\it wedge} with an amplitude that depends on the amplitude of power--pattern in that direction, with objects at the horizon being closest to the boundaries of the wedge.
 
-The degree to which these objects leak power into spectral line of sight modes above the wedge, depends primarily on the brightness of foregrounds near the boundaries of the wedge \citep{thy13,pob13,ved12,par12}.
-Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting to the primary field of view.
+The degree to which these objects leak power into spectral line of sight modes above the {\it foreground wedge}, depends primarily on the brightness of foregrounds near the boundaries of the {\it wedge} \citep{thy13,pob13,ved12,par12}.
+Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view.
 
@@
 Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the visible hemisphere in the westward sky and the MWA tile's power receptivity is significantly high ($\gtrsim 12$\%) in that direction.
-In the {\it zenith} pointing, the Galactic plane has almost set and the contour level of power--pattern in that direction is at least 16 times smaller.
+In the {\it zenith} pointing, the Galactic plane has set and the contour level of power--pattern in that direction is at least 16 times smaller.
 
@@
 
-We use a combination of NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz to provide our compact source catalog due to their complementary survey footprints covering the entire sky, and matched flux sensitivity and angular resolution.
+We use a combination of NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz to provide our compact object catalog due to their complementary survey footprints covering the entire sky, and matched flux sensitivity and angular resolution.
 The SUMSS catalog covers the sky with declination $\delta < -30$\arcdeg with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
-\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+\caption{Simulated delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), for {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
 White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
@@
 
-In order to make a quantitative comparison of delay spectra obtained with the MWA data and our simulations, we consider the following uncertainties:
-
-{\bf Spectral index uncertainty.}  Our foreground models are derived from other higher frequency catalogs and sky maps.
+In order to make a quantitative comparison of delay spectra obtained with the MWA data and our simulations, we consider the uncertainty in the assumed spectral index of our foreground model.
+Our foreground models are derived from other higher frequency catalogs and sky maps.
 The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observed frequency.
@@
 From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for compact objects from NVSS and SUMSS catalogs.
-Although the model of \citet{deo08} yields a spectral index per direction on the sky, we assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from 408~MHz map of \citet{has82}.
+Although the model of \citet{deo08} yields a spectral index per direction on the sky, we assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from the 408~MHz map of \citet{has82}.
 Thus, fractional errors in delay spectrum from compact foreground objects and diffuse emission are $\sim$70\% and $\sim$30\% respectively.
 
-{\bf Thermal uncertainty.} {\bf [JDB -- this is really a model component, not an uncertainty.
-We should probably move it to the earlier section.]}  In addition to intrinsic model uncertainty, delay spectra from simulations and data each have uncertainties due to thermal noise with rms, $\Delta V^\textrm{N}_\eta(|\overline{\mathbf{x}}|,\eta) \sim$ 1.4$\times 10^7$~Jy~Hz, where the superscript N stands for thermal noise.
+In addition to intrinsic model uncertainty, delay spectra from simulations and data each have fluctuations due to thermal noise in delay spectrum with rms $\sim$ 1.4$\times 10^7$~Jy~Hz.
 We estimate the ratio of delay spectra from data and simulations as, $\rho = |V^\textrm{D}_\eta(|\overline{\mathbf{x}}|,\eta)|\,/\,|V^\textrm{S}_\eta(|\overline{\mathbf{x}}|,\eta)|$, where superscripts D and S denote data and simulation respectively.
@@
 The median absolute deviation from these distributions correspond to $\sim$~90\% fractional difference between data and modeling on average in either case.
-This is in line with aforementioned uncertainties in foreground models and thermal noise in measurements.
+This is in line with aforementioned uncertainty in foreground models and thermal noise fluctuations in measurements.
 Currently we are significantly limited by unavailability of all--sky foreground models at the observed frequency of 185~MHz.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
 \caption{Histogram of logarithm of ratio of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
-Median absolute deviations of $\log_{10}\rho$ are 0.29 and 0.27 respectively.
+Median absolute deviation of $\log_{10}\rho$ is $\sim 0.28$ in both cases.
 A value of zero indicates when the ratio $\rho=1$.
 \label{fig:data-sim-ratio}}
-% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
+% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
 % \caption{Significance of logarithm of ratio of observed to simulated visibilities relative to expected uncertainty restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 
-Having established that the mode matches the data to the level of expected uncertainties, we proceed to examine in further detail the key signatures seen in the modeled delay spectra.
-A number of factors are responsible for the characteristics noted in the delay spectra obtained from data and through modeling.
-We address these factors below: {\bf [JDB: These bullets below don't exactly match the analysis reported later -- we need to revise them so that they match one-to-one the results presented later.]}
+Having established that the simulation matches the data to the level of expected uncertainties, we proceed to examine in further detail the key signatures seen in the simulated delay spectra.
+A number of factors are responsible for the characteristics noted in the delay spectra obtained from data and through simulations.
+We address these factors below:
 
@@
 In fact, the patch of sky for MWA observations is chosen from regions of low foreground emission.
-In our detailed study, we have divided the transiting sky into four different ``bow-tie'' shaped regions.
+% In our detailed study, we have divided the transiting sky into four different ``bow-tie'' shaped regions.
 Each region has equal area and the regions are aligned to cardinal directions around zenith as the origin.
@@
 
-\item {\it Antenna Pair Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our antenna pairs by their orientation ($\theta_\textrm{b}$) measured counter-clockwise from East.
+\item {\it Antenna Pair Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our antenna pairs by their orientation, $\theta_\textrm{b}$.
 We use the following bins: $-67\fdg 5\le \theta_\textrm{b} < -22\fdg 5$, $-22\fdg 5\le \theta_\textrm{b} < 22\fdg 5$, $22\fdg 5\le \theta_\textrm{b} < 67\fdg 5$, and $67\fdg 5\le \theta_\textrm{b} < 112\fdg 5$.
-The bin centers are oriented towards South-East, East, North-East, and North respectively.
-Since delays depend on antenna pair orientation, binning by orientation allows us to match delay spectrum features to different sky directions.
+The bin centers are oriented towards South--East, East, North--East, and North respectively.
+Since delays depend on $\theta_\textrm{b}$, binning by $\theta_\textrm{b}$ allows us to match delay spectrum features to different sky directions.
 
@@
 \item {\it Instrumental Passband}: The instrumental passband characteristics have a direct effect on the delay transform owing to the Fourier relationship between the two.
-Due to excision of noisy and aliased edge channels repeated in each coarse channel of our passband, we saw repeated patterns of the {\it foreground wedge} in the delay spectra.
+Due to excision of aliased edge channels repeated in each coarse channel of our passband, we observed repeated patterns of the {\it foreground wedge} in the delay spectra.
 As already mentioned earlier, we have used the deconvolution algorithm ({\it CLEAN}) to rid the delay spectra of band shape effects.
@@
 
-It may be noted that numerous features overlap at varying levels of significance depending on the combinations of parameters.
+It may be noted that numerous features overlap at varying levels of significance depending on combinations of parameters.
 We assign the features to their predominant causes.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
-\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
+\caption{Simulated delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
 White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
@@
 As we also see in Figure~\ref{fig:noiseless-dsm-delay-spectrum}, diffuse emission from within the primary field of view manifests in the {\it off--zenith} pointing as a branch at the top of the wedge, while in the {\it zenith} pointing it shows up at near zero delay.
-The former is seen at positive delays because the main lobe of the power--pattern is centered eastward of zenith, whereas in the latter it is centered at zenith.
+The former is seen at positive delays because the primary lobe of the power--pattern is centered eastward of zenith, whereas in the latter it is centered at zenith.
 As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inversely proportional to length of the baseline projected in the direction of emission.
@@
 A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\overline{\mathbf{x}}|\le$~200~m)---contrary to expectations, since longer baselines are not expected to be sensitive to diffuse emission.
-In the Widefield observations of the MA, long baselines appear shortened due to projection effects for off--axis emission, especially towards the horizon.
+In the Widefield observations of the MWA, long baselines appear shortened due to projection effects for off--axis emission, especially towards the horizon.
 This effect is evident at all LSTs in our simulations and it naturally holds for short baselines, as well.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
-\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise.
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
+\caption{Simulated delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise.
 White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
@@
 Therefore, the distribution of objects within the primary field of view depends mainly on how the power--pattern maps to delay space for each antenna pair.
-Since the main lobe of the power--pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with positive delay corresponding to the position of the pointing center.
-In the {\it zenith} pointing compact emission from the same patch of sky is seen as a bright horizontal arm at zero delay  since the main lobe of the power--pattern is centered at zenith.
+Since the main lobe of the power--pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with positive delay corresponding to the position of the primary lobe of the power--pattern.
+In the {\it zenith} pointing, compact emission from the same patch of sky is seen as a bright horizontal arm at zero delay  since the primary lobe of the power--pattern is centered at zenith.
 
@@
 Foreground emission in zero delay and negative delay regions in the {\it off--zenith} pointing is caused by compact objects co--located with sidelobes of the power--pattern.
-On the other hand, compact objects co--located with sidelobes of power--pattern in the {\it zenith} pointing are revealed as faint but distinct branches at positive and negative delays depending on the orientation of antenna spacing and direction of emission on the sky.
+On the other hand, compact objects co--located with secondary lobes of power--pattern in the {\it zenith} pointing are revealed as faint but distinct branches at positive and negative delays depending on the orientation of antenna spacing and direction of emission on the sky.
 
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_visibilities_0.3m_ground_custom_gaussian_FG_model_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
-\caption{Absolute value of delay spectrum, $|V_\eta(\overline{\mathbf{x}},\eta)|$, for three chosen antenna spacings of length: $\sim$84~m ({\it top}), $\sim$105~m ({\it middle}), and $\sim$170~m ({\it bottom}).
+\includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
+\caption{Simulated delay spectrum amplitudes, $|V_\eta(\overline{\mathbf{x}},\eta)|$, for three chosen antenna spacings of length: $\sim$84~m ({\it top}), $\sim$105~m ({\it middle}), and $\sim$170~m ({\it bottom}).
 The antenna spacing vector is specified in each panel along East, North and Upward directions.
@@
 This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits.
-Outside the horizon delay limits, it is predominantly due to thermal noise.
+Outside the horizon delay limits, it is predominantly due to thermal noise fluctuations.
 Compact objects dominate the central regions of power spectrum while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a {\it pitchfork}--shaped structure.
@@
 All these observations are consistent with those from Figures~\ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}.
-The observability of the predicted {\it pitchfork} signature depends on the relative levels of uncertainty in the foreground model and thermal noise.
+The observability of the predicted {\it pitchfork} signature depends on the relative levels of uncertainty in the foreground model and fluctuations from thermal noise.
 In our simulations, since thermal noise in these very short duration snapshots is $\gtrsim 10^7$~Jy~Hz, and features near the horizon delay limits are $\sim 10^7$~Jy~Hz, the {\it pitchfork} feature is not expected to be detected in a noisy scenario although this feature is marginally visible in the {\it zenith} pointing of the observed data (see Figure~\ref{fig:fhd_data}).
@@
 For instance, there is an increase in the number of secondary peaks in the delay spectrum between zero--delay and horizon delay limits as antenna spacing increases from $\sim$84~m to $\sim$170~m.
-In this case, these correspond to sidelobes of the power--pattern that lie between the main lobe and the horizon along the North--South direction.
+In this case, these correspond to secondary lobes of the power--pattern that lie between the main lobe and the horizon along the North--South direction.
 In short antenna spacings, due to relatively lower resolution along delay axis inside the {\it foreground wedge} and a consequent loss of localization of direction of foreground emission, these secondary peaks blend in with other major peaks and are not distinctly visible.
@@
 
-In summary, the brightest signature ($\gtrsim 10^{10}$~Jy~Hz) is that of the Galactic center in the {\it off--zenith} pointing co--located at a westward sidelobe with a significantly high gain.
-The next brightest signature ($\sim 10^8$--$10^9$~Jy~Hz) is caused by compact emission appearing to be concentrated in the inner regions of the foreground wedge (near zero delays) rather than towards the edges.
-Diffuse emission co--located at the main lobe of the power--pattern is fainter by approximately an order of magnitude relative to compact emission from the same region for a $\sim 84$~m antenna spacing but unlike the latter, diffuse emission decreases rapidly by over two orders of magnitude as antenna spacing is widened to $\sim 170$~m.
-However, diffuse emission is significantly high near horizon delay limits compared to main lobe of the power--pattern as antenna spacing is widened.
+In summary, the brightest signature ($\gtrsim 10^{10}$~Jy~Hz) is that of the Galactic center in the {\it off--zenith} pointing co--located at a westward secondary lobe with a significantly high gain.
+The next brightest signature ($\sim 10^8$--$10^9$~Jy~Hz) is caused by compact emission appearing to be concentrated in the inner regions of the {\it foreground wedge} (near zero delays) rather than towards the edges.
+Diffuse emission co--located at the primary lobe of the power--pattern is fainter by approximately an order of magnitude relative to compact emission from the same region for a $\sim 84$~m antenna spacing but unlike the latter, diffuse emission decreases rapidly by over two orders of magnitude as antenna spacing is widened to $\sim 170$~m.
+However, diffuse emission is significantly high near horizon delay limits compared to primary lobe of the power--pattern as antenna spacing is widened.
 This leads to edge--heavy features in the delay spectrum.
-Complemented by the center-heavy compact foreground features, we predict the presence of a {\it three--pronged pitchfork}--shaped signature in the delay spectrum of foreground sky and identify evidence for the signature in actual MWA data, although deeper observations are needed to more clearly reveal the effect.
+Complemented by the center-heavy compact foreground features, we predict the presence of a {\it three--pronged pitchfork}--shaped signature in the delay spectrum of foreground sky and identify evidence for presence of this signature in actual MWA data, although deeper observations are needed to more clearly reveal the effect.
 
@@
 
-The delay spectrum, $V_\eta(\overline{\mathbf{x}},\eta)$, not only carries information on spatial scales of foreground emission but also offers a unique advantage of viewing the sky through a combination of antenna spacing vector and the delay axis.
-This allows us to programmatically screen data for antenna spacings that are severely contaminated by foregrounds which can be weighted appropriately during data analysis.
 While the {\it foreground wedge} is a region of high foreground contamination, of particular interest to the EoR community is foreground contamination near horizon delay limits.
@@
 
-With a foreground model known {\it a priori} in which locations of very bright foreground objects such as the Galactic center or AGN are available, we can predict their response across antenna spacings as a function of observing parameters such as LST, power--pattern, etc.
-Specifically, we can pick antenna pairs where these bright foreground objects will be visible at delays near horizon delay limits.
-One such example is already provided in our study.
-The Galactic center is the most dominant source of foreground contamination in the {\it off--zenith} pointing.
-Since it is located in the western sky, it is most distinctly seen near the negative horizon delay limit on eastward oriented antenna spacings in the delay spectrum (see Figures~\ref{fig:fhd_data}, \ref{fig:sim_data} and \ref{fig:noiseless-dsm-delay-spectrum}, and \S\ref{sec:GC-east}).
-In fact these figures illustrate quite well the leakage beyond the {\it foreground wedge} into the {\it EoR window} due to these extremely bright objects near the boundaries of the wedge.
-In practice, we could downweight these antenna spacings for the specific LST to mitigate the effects of spillover from such foreground contamination in EoR studies.
-With spatial scales of such objects known, besides downweighting based on antenna spacing orientation, we could also downweight based on antenna spacing distance.
-For instance, a bright compact object will yield uniform {\it visibility} amplitudes across antenna spacings whereas that of an extended object will gradually fade with increase in antenna spacing.
+The Galactic center in the {\it off--zenith} observation is one such example already available in our study.
+Figure~\ref{fig:sky-model} shows the sky model ({\it top:} compact objects, {\it bottom:} diffuse sky model) in this pointing.
+The Galactic center is the most dominant source of foreground contamination from the diffuse sky model co--located with a bright secondary lobe of the power pattern.
+Figure~\ref{fig:delay-map} shows the sky mapped to delays registered by the baseline vectors, of length 100~m for instance, oriented northward (top panel) and eastward (bottom panel).
+Since the Galactic center is located in the western sky, this figure demonstrates that it is observed at negative delay on a baseline vector oriented eastward and at zero delay on a baseline vector oriented northward.
+Figure~\ref{fig:baseline-breakup} shows the delay spectra on baselines oriented northward (67\fdg 5~$\le \theta_\textrm{b} <$~112\fdg 5) at the top and eastward (-22\fdg 5~$\le \theta_\textrm{b} <$~22\fdg 5) at the bottom.
+The Galactic center manifests itself most distinctly near the negative horizon delay limit on short eastward baselines in the delay spectrum (bottom panel of Figure~\ref{fig:baseline-breakup}).
+Consequently, the spillover caused by the instrument's transfer function from the {\it foreground wedge} into the {\it EoR window} is most severe in eastward baselines relative to northward baselines evident by the bright vertical stripes of foreground contamination.
 
-This provides us with a very simple and yet effective tool in adding a layer of sophistication to mitigate effects of foreground contamination in EoR data analysis.
+\begin{figure*}[htb]
+\centering
+\subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{figures/v1_0/sky_model_snapshot_0.eps}}
+\subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{figures/v1_0/directional_delay_map_snapshot_0.eps}}
+\subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figures/v1_0/baseline_orientation_breakup_snapshot_0.eps}}
+\caption{(a): Sky model showing compact objects ({\it top}) and diffuse emission ({\it bottom}).
+Contours of power pattern are overlaid whose levels are identical to those in Figures~\ref{fig:DSM} and \ref{fig:CSM}.
+The Galactic center is very prominent in diffuse emission in the western sky and is co--located with a bright secondary lobe of the power pattern.
+(b): Sky hemisphere mapped to delays observed on antenna spacings with northward ({\it top}) and eastward ({\it bottom}) orientations.
+Color scale shown is for a 100~m antenna spacing.
+Delays vary linearly with antenna spacing length.
+The bright Galactic center in the western sky is recorded at zero delays in northward antenna spacings and close to negative horizon delay limit on eastward antenna spacings.
+(c): Simulated delay spectra $|V_\eta(\overline{\mathbf{x}},\eta)|$ on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}).
+The bright Galactic center is prominently visible close to negative horizon delay limit on eastward antenna spacings.
+These are also the most severely contaminated by foreground spillover.
+The northward antenna spacings, on the other hand, are the least contaminated.}
+\label{fig:breakup}
+\end{figure*}
+
+The delay spectrum, $V_\eta(\overline{\mathbf{x}},\eta)$, not only carries information on spatial scales of foreground emission but also offers a unique advantage of viewing the sky through a combination of antenna spacing vector and the delay axis.
+With a foreground model known {\it a priori} in which structures and locations of very bright foreground objects such as the Galactic center or AGN are available, we can predict their response across antenna spacings as a function of observing parameters such as LST, power--pattern, etc.
+This allows us to programmatically screen data for antenna spacings that are severely contaminated by foregrounds near the horizon delay limits in the delay spectrum.
+These can be weighted appropriately during data analysis.
+We demonstrate such a screening technique, whereby we use the bright object's location and structure to discard antenna spacings of certain lengths and orientations to mitigate foreground contamination in the {\it EoR window}.
+
+In our example, we discard eastward antenna spacings (-22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5) of lengths $|\overline{\mathbf{x}}| < 40$~m.
+Figure~\ref{fig:before-after} shows the delay spectra obtained with all antenna spacings (top panel) and after applying our screening technique on {\it off--zenith} observation.
+Notice the remarkable reduction -- by about an order of magnitude -- in foreground contamination in this region after application of our screening technique.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_before_after_nside_64_Tsys_95.0K.eps}
+\caption{Simulated delay spectra amplitude, $|V_\eta(\overline{\mathbf{x}},\eta)|$, for the {\it off--zenith} observation with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
+Discarded antenna spacings have lengths $|\overline{\mathbf{x}}| < 40$~m and orientations -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5.
+The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} prominent when all antenna spacings are used is reduced by an order of magnitude when short eastward antenna spacings are discarded.
+\label{fig:before-after}}
+\end{figure}
+
+We further demonstrate this screening technique can be customized when it is necessary to optimize between foreground mitigation and loss of sensitivity from discarding data.
+Figure~\ref{fig:screening} shows how the typical foreground contamination in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
+We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars).
+Among antenna spacings that satisfy these criteria, we discard data from those whose lengths are shorter than $|\overline{\mathbf{x}}|_\textrm{max}$ ($x$--axis) and show foreground contamination estimated in the {\it EoR window} from all remaining antenna spacings.
+Foreground contamination estimated from only northward (dashed line) and eastward (dot--dashed line) antenna spacings represent the lower and upper limits respectively, and are shown for reference.
+It is seen that foreground contamination can be mitigated by a factor between $\sim 1.5$ ($|\theta_\textrm{b}|\le$~7\fdg 5) and $\sim 10$ ($|\theta_\textrm{b}|\le$~22\fdg 5).
+It is notable that short antenna spacings play maximum role in determining foreground contamination.
+For instance, among antenna spacings with orientations $|\theta_\textrm{b}|\le$~22\fdg 5, discarding those with lengths $|\overline{\mathbf{x}}|>40$~m has no effect on foreground contamination and would only lead to loss of sensitivity.
+On the other hand, a bright compact object at the same location will give rise to foreground contamination even on longer antenna spacings.
+Such cases will necessitate discarding more or all of the eastward oriented antenna spacings.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_nside_64_Tsys_95.0K.eps}
+\caption{Foreground contamination in {\it EoR window} for the {\it off--zenith} pointing as a function of orientation and length of discarded antenna spacings.
+Antenna spacings oriented eastward with varying degrees of directedness are considered: -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars).
+Among antenna spacings that satisfy the above criteria, ones with lengths $\le |\overline{\mathbf{x}}|_\textrm{max}$ ($x$--axis) are discarded.
+Lower (dashed line) and upper (dot--dashed) limits on foreground contamination are derived from northward (67\fdg 5~$\le\theta_\textrm{b}<$~112\fdg 5) and eastward ($|\theta_\textrm{b}|\le$~22\fdg 5) antenna spacings respectively.
+Foreground contamination in the {\it EoR window} reduces by factor $\sim$1.5 ($|\theta_\textrm{b}|\le$~7\fdg 5) to $\sim$10 ($|\theta_\textrm{b}|\le$~22\fdg 5).
+Short baselines have maximum effect on foreground contamination.
+For instance, discarding antenna spacings oriented $|\theta_\textrm{b}|\le$~22\fdg 5 with lengths $|\overline{\mathbf{x}}|>40$~m has no effect in reducing foreground contamination.
+\label{fig:screening}}
+\end{figure}
+
+In principle, instead of discarding selected antenna spacings altogether, we could down--weight them based on an optimal scheme.
+This technique provides a very simple and yet effective tool in adding a layer of control to mitigate effects of foreground contamination in EoR data analysis.
 % The success behind clear identification of severely foreground--contaminated antenna pairs based on delay spectrum technique depends on the localization of the source of foreground contamination and its strength.
@@
 Our primary motivation in this work is to understand how the various bright foregrounds will manifest in three dimensional power spectrum of H{\sc i} from 21~cm reionization observations.
-In the units presented here, the 21~cm signal is expected to be eight orders of magnitude fainter than the foregrounds; a detailed understanding of how foregrounds can corrupt the 21~cm power spectrum is therefore essential.
+In the units presented here, the dynamic range between bright foregrounds and the 21~cm signal is expected to be $\sim 10^8$; a detailed understanding of how foregrounds can corrupt the 21~cm power spectrum is therefore essential.
 This analysis extends previous work by simulating the entire sky rather then just the central field of view and by providing a comparison with early observations from the MWA.
@@
 
-Simulating in all important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from [TBD: put actual citation] and bright compact objects from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra is in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise in measurements.
+Simulating in all important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright compact objects from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra is in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise fluctuations in measurements.
 
@@
 
-The bright Galactic center at the edge of the western horizon co--located with one of the far sidelobes of MWA tile power--pattern is the brightest source of foreground contamination in the {\it off--zenith} pointing.
-It manifests itself near negative horizon delay limit in the delay spectrum on antenna spacings oriented eastward.
+The bright Galactic center at the edge of the western horizon co--located with one of the far secondary lobes of MWA tile power--pattern is the brightest source of foreground contamination in the {\it off--zenith} pointing.
+It manifests itself near the negative horizon delay limit in the delay spectrum on antenna spacings oriented eastward.
 
@@
 In conclusion, we find that inclusion of objects, both diffuse and compact, all the way to the horizon is essential to explaining the observed power spectrum.
-We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination in EoR data analysis.
+We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination by an order of magnitude in EoR data analysis.
 It is possible to discard or down--weight antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation.

### 2014-09-20 foreground_delay_spectrum_modeling.tex
From v0.6: =========
(*) Made minor reshuffling of paragraphs up to section 4.
(*) Minor edits up to section 4.

@@
 
-\def\UW{\altaffilmark{2}}
-\def\UWtxt{\altaffiltext{2}{Department of Physics, University of Washington, Seattle, WA 98195, USA}}
+\def\SKASA{\altaffilmark{2}}
+\def\SKASAtxt{\altaffiltext{2}{Square Kilometre Array South Africa (SKA SA), Cape Town 7405, South Africa}}
 
-\def\RRI{\altaffilmark{3}}
-\def\RRItxt{\altaffiltext{3}{Raman Research Institute, Bangalore 560080, India}}
+\def\ANU{\altaffilmark{3}}
+\def\ANUtxt{\altaffiltext{3}{Research School of Astronomy and Astrophysics, Australian National University, Canberra, ACT 2611, Australia}}
 
-\def\SKASA{\altaffilmark{4}}
-\def\SKASAtxt{\altaffiltext{4}{Square Kilometre Array South Africa (SKA SA), Cape Town 7405, South Africa}}
+\def\Haystack{\altaffilmark{4}}
+\def\Haystacktxt{\altaffiltext{4}{MIT Haystack Observatory, Westford, MA 01886, USA}}
 
-\def\ANU{\altaffilmark{5}}
-\def\ANUtxt{\altaffiltext{5}{Research School of Astronomy and Astrophysics, Australian National University, Canberra, ACT 2611, Australia}}
+\def\RRI{\altaffilmark{5}}
+\def\RRItxt{\altaffiltext{5}{Raman Research Institute, Bangalore 560080, India}}
 
-\def\Haystack{\altaffilmark{6}}
-\def\Haystacktxt{\altaffiltext{6}{MIT Haystack Observatory, Westford, MA 01886, USA}}
+\def\Curtin{\altaffilmark{6}}
+\def\Curtintxt{\altaffiltext{6}{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}}
 
-\def\Curtin{\altaffilmark{7}}
-\def\Curtintxt{\altaffiltext{7}{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}}
+\def\USydney{\altaffilmark{7}}
+\def\USydneytxt{\altaffiltext{7}{Sydney Institute for Astronomy, School of Physics, The University of Sydney, NSW 2006, Australia}}
 
-\def\USydney{\altaffilmark{8}}
-\def\USydneytxt{\altaffiltext{8}{Sydney Institute for Astronomy, School of Physics, The University of Sydney, NSW 2006, Australia}}
+\def\CAASTRO{\altaffilmark{8}}
+\def\CAASTROtxt{\altaffiltext{8}{ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}}
 
-\def\CAASTRO{\altaffilmark{9}}
-\def\CAASTROtxt{\altaffiltext{9}{ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}}
+\def\MIT{\altaffilmark{9}}
+\def\MITtxt{\altaffiltext{9}{Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}}
 
-\def\MIT{\altaffilmark{10}}
-\def\MITtxt{\altaffiltext{10}{Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}}
+\def\CfA{\altaffilmark{10}}
+\def\CfAtxt{\altaffiltext{10}{Harvard-Smithsonian Center for Astrophysics, Cambridge, MA 02138, USA}}
 
-\def\CfA{\altaffilmark{11}}
-\def\CfAtxt{\altaffiltext{11}{Harvard-Smithsonian Center for Astrophysics, Cambridge, MA 02138, USA}}
+\def\UW{\altaffilmark{11}}
+\def\UWtxt{\altaffiltext{11}{Department of Physics, University of Washington, Seattle, WA 98195, USA}}
 
@@
 %% Supporting authors
-A.~P.~Beardsley\UW,
-B.~J.~Hazelton\UW, 
-M.~F.~Morales\UW, 
-J.~Pober\UW,
-I.~S.~Sullivan\UW,
-T.~Prabu\RRI, 
+% A.~P.~Beardsley\UW,
+% J.~Pober\UW,
+% I.~S.~Sullivan\UW,
 %% Builders' list
@@
 L.~J.~Greenhill\CfA,
+B.~J.~Hazelton\UW, 
 M.~Johnston-Hollitt\Victoria,
@@
 D.~A.~Mitchell\CASS$^,$\CAASTRO, 
+M.~F.~Morales\UW, 
 E.~Morgan\MIT, 
@@
 S.~M.~Ord\Curtin$^,$\CAASTRO,
+T.~Prabu\RRI, 
 A.~E.~E.~Rogers\Haystack, 
@@
 We investigate the chromatic signatures of an all--sky foreground model imprinted by an instrumental transfer function.
-Using the delay spectrum technique, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal via spillover outside the {\it wedge} arise from power received far away from the primary field of view.
-Comparing data from recent Murchison Widefield Array (MWA) observations with simulations separated into components based on sky position and type of emission, we identify diffuse emission near the horizon as the largest contributing factor, even on wide antenna spacings.
+Using the delay spectrum technique, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view.
+Comparing data from recent Murchison Widefield Array (MWA) observations with simulations separated into components based on type of emission, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings.
 For signals entering through the primary field of view, compact objects dominate the foreground contamination.
-These two mechanisms imprint a characteristic {\it pitchfork} signature on the {\it wedge} in Fourier space.
+These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space.
 Based on these results, we propose that selective down--weighting of baselines based on length, direction, and time will mitigate a large fraction of foreground contamination in redshifted 21~cm power spectrum analysis.
@@
 Studying the epoch of reionization holds the key to understanding this evolution.
-Recently, observing the redshifted 21~cm spin transition of neutral hydrogen has emerged as a very promising experiment to fill the gaps in our understanding of the universe's history.
+Direct observation of redshifted 21~cm spin flip transition of H{\sc i} has been identified as a useful probe of the EoR \citep{sun72,sco90,mad97,toz00,ili02} and has recently emerged as a very promising experiment to fill the gaps in our understanding of the universe's history.
 
@@
 
-Perhaps the key challenge in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
+A key challenge in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
 \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}.
@@
 In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain.
-Here we will focus primarily on extending the {\it avoidance} strategy by identifying foreground components most likely to ``leak'' from foreground modes to EoR modes and demonstrating a scheme for down--weighting these components.
+Here, we will focus primarily on extending the {\it avoidance} strategy by identifying foreground components most likely to ``leak'' from foreground modes to EoR modes and demonstrating a scheme for down--weighting these components.
 
@@
 Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity.
-To decouple our findings from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the individual baseline-based approach of \citet{pob13,par14}.
+To decouple our findings from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the individual baseline--based approach of \citet{pob13,par14}.
 This approximates the power spectrum as the inverse Fourier transform of the spectra generated by the instrument's correlator.
@@
 Matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the signature of the power spectrum of each baseline.
-We find that the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon.
+We report two important findings: the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a larger role than previously thought.
+% We find that the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon.
 We find that diffuse Galactic emission plays a larger role then previously thought.
@@
 \begin{align}\label{eqn:obsvis}
-  V_f(\overline{\mathbf{u}},f) &= \int A(\overline{\mathbf{l}},f)\,I(\overline{\mathbf{l}},f)\,W_f(f)\,e^{-i2\pi \overline{\mathbf{u}}\cdot\overline{\mathbf{l}}}\,\dif^2 \overline{\mathbf{l}}.
+  V_f(\overline{\mathbf{u}},f) &= \iint\limits_\textrm{sky} A(\overline{\mathbf{l}},f)\,I(\overline{\mathbf{l}},f)\,W_f(f)\,e^{-i2\pi \overline{\mathbf{u}}\cdot\overline{\mathbf{l}}}\,\dif^2 \overline{\mathbf{l}}.
 \end{align}
@@
 where, $W'_f(f)$ is a frequency window weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}.
-$\eta$ represents the signal delay between the antenna pairs in an interferometer given by:
+$\eta$ represents the signal delay between antenna pairs:
 \begin{equation}\label{eqn:delay}
@@
 Note that as the visibility is a fully complex quantity, its Fourier transform is fully Nyquist sampled over positive and negative delays.
-Negative and positive delay\footnote{The terms delay and lag are used interchangeably in this paper.} correspond to the two hemispheres of the sky as transected by the antenna pair's baseline vector.
-For example, a perfectly Eastward oriented baseline vector will observe an object in the Eastern sky at positive delay, and conversely for an object in the Western sky.
+Negative and positive delay\footnote{The terms delay and lag are used interchangeably in this paper.} correspond to the two hemispheres of the sky as transected by the antenna spacing vector.
+In this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the telescope site.
+Hence, a perfectly Eastward oriented antenna spacing will observe an object in the Eastern sky at positive delay, and conversely for an object in the Western sky.
+Similarly, an object in the Northern sky will appear at a positive delay at an antenna spacing oriented Northward.
 
 Since we are studying a redshifted H{\sc i} spectral line, $f$ is a measure of cosmological distance along the line of sight.
-Consequently, $\eta$ is a measure of line of sight spatial frequency mode, $k_\parallel$, given by $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
+Consequently, $\eta$ is a measure of line of sight spatial frequency mode given by $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
 We use $H_0=100\,h$~km~s$^{-1}$~Mpc$^{-1}$ and assume a flat $\Lambda$CDM universe with $\Omega_\textrm{M}=0.27$ and $\Omega_\Lambda=0.73$.
 
-Thus, $|V_\eta(\overline{\mathbf{u}},\eta)|^2$ is directly related to the spatial power spectrum of redshifted H{\sc i} distribution, $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ \citep{mor04}.
-However, foreground objects emit in radio frequencies and contaminate the signals from the redshifted H{\sc i}.
-Due to their spatial and spectral properties, $V_\eta(\overline{\mathbf{u}},\eta)$ also contains contamination from foreground radio objects \citep{thy13,tro12,mor12,bow09}.
-These contaminations have to be characterized precisely in order to reduce their impact on redshifted H{\sc i} power spectrum detection sensitivity.
+Thus, $|V_\eta(\overline{\mathbf{u}},\eta)|^2$ is directly related to the spatial power spectrum of EoR H{\sc i} distribution, $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ \citep{mor04}.
+However, foreground objects also emit in radio frequencies.
+Due to their spatial and spectral (chromatic) structure, the measured $V_\eta(\overline{\mathbf{u}},\eta)$ is contaminated by foreground radio objects \citep{thy13,tro12,mor12,bow09}.
+These contaminations have to be characterized precisely in order to reduce their impact on EoR H{\sc i} power spectrum detection sensitivity.
 
-In summary, $V_\eta(\overline{\mathbf{u}},\eta)$, is closely related to visibilities, $V_f(\overline{\mathbf{u}},f)$, the basic data blocks measured by each antenna pair, as seen from equation~\ref{eqn:delay-transform}.
+In summary, $V_\eta(\overline{\mathbf{u}},\eta)$ is closely related to visibilities, $V_f(\overline{\mathbf{u}},f)$, the basic data blocks measured by each antenna pair, as seen from equation~\ref{eqn:delay-transform}.
 It captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument.
-At the same time, it is very closely related to the sought power $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ in $k$--modes containing critical information about spatial scales.
+At the same time, it is directly related to the sought power $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ in $k$--modes containing critical information about spatial scales.
 For our analysis, we use an equivalent quantity, $V_\eta(\overline{\mathbf{x}},\eta) \equiv V_\eta(\lambda\overline{\mathbf{u}},\eta)$.
@@
 These boundaries are called horizon delay limits and are shown by solid lines.
-Spectral response of the instrument convolves the foreground wedge and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\eta$--axis.
+Spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\eta$--axis.
 This effect has been quantified in \citet{thy13}.
 The width of this narrow strip is inversely proportional to operating bandwidth.
-Region of Fourier space excluding the {\it foreground wedge} and also the narrow strips is usually called the {\it EoR window}.
+Region of Fourier space excluding the {\it foreground wedge} and the narrow strips is usually called the {\it EoR window}.
 These are shown in light and medium shades.
-However, the instrument used in our study, namely the MWA, has a passband constructed using coarse channels with a characteristic shape that results in harmonic repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.
+However, the instrument used in our study, namely the MWA, has a passband constructed using coarse channels with a characteristic shape giving rise to periodic gratings thereby resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.
 The underlying cause of this effect will be discussed later in this paper.
-Hence, we have identified a region of high significance (shown in medium shade) that forms a subset of the {\it EoR window} bounded by $|\eta|\le 0.78\,\mu$s where contamination from foregrounds is expected to be low and yet the EoR H{\sc i} power spectrum signal is reasonably strong.
+
+% Hence, we have identified a region of high significance (shown in medium shade) that forms a subset of the {\it EoR window} bounded by $|\eta|\le 0.78\,\mu$s where contamination from foregrounds is expected to be less and yet the EoR H{\sc i} power spectrum signal is reasonably strong.
 Hereafter, we refer to this region as the {\it EoR window} in the rest of this paper.
+
+In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\overline{\mathbf{k}}|$.
+Hence, the brightest EoR signal will be observed on the shortest baselines and smallest delays.
+Thus the region of interest for EoR studies is slightly beyond the horizon delay limits (dashed lines) and inside the first grating on either side ($|\eta| \lesssim 0.78\,\mu$s, dot--dashed lines) caused by the instrument's spectral transfer function.
+We call this region with medium shade as the {\it EoR window} in this paper.
 The MWA EoR collaboration actively uses this as a test region in Fourier space for diagnostics on data quality.
@@
 The dark shaded region is often referred to as {\it foreground wedge} where smooth spectrum foregrounds reside.
-The boundaries of the {\it foreground wedge} (solid lines) are given by the light travel time for corresponding antenna spacings and is referred to as the horizon delay limits.
+The boundaries of the {\it foreground wedge} (solid lines) are given by the light travel time for corresponding antenna spacings and are referred to as horizon delay limits.
 In principle, region outside the wedge is expected to be relatively free of foreground contamination and is called the {\it EoR window} (light and medium shades).
@@
 
-Throughout this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the telescope site.
-For the conventions of Fourier transform and its inverse we use in this paper, in accordance with equation~\ref{eqn:delay}, signals from the sky towards east and north are recorded with positive delays on eastward and northward oriented antenna spacings respectively.
-For instance, at the beginning of the observation, the Galactic center is in the westward sky just about to set.
-Signals from this direction are recorded with a negative delay.
-Similarly, signals from directions eastward of the local meridian arrive with positive delays.
+The terms antenna spacing and baseline are synonymous in radio interferometry.
+The latter term is also used broadly to refer to the antenna pair system in a radio interferometer array.
+We use these terms interchangeably in this paper.
 
-For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all antenna spacings to lie in the range $-67\fdg 5 \leq \theta_b < 112\fdg 5$.
-Antenna spacings oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
+For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67\fdg 5 \leq \theta_b < 112\fdg 5$.
+Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
 
@@
 In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30~MHz band anywhere in the range 80--300~MHz.
-Each tile is a phased array of 16 dipoles each in the shape of a bow--tie characterized by a grating response with a $\gtrsim$~20\arcdeg wide primary field of view and multiple secondary side lobes.
+Each tile is a phased array of 16 dipoles each in the shape of a bow--tie.
+This yields a primary field of view $\gtrsim$~20\arcdeg wide and multiple secondary lobes for the tile.
 The tile arrangement is a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius with the longest baselines at 3~km \citep{bea12}.
@@
 
-The MWA bandpass is divided coarsely into 24 chunks with each chunk weighted by a digital filter.
-The channels at the edges of these chunks have been flagged because they are known to possibly be contaminated by aliasing at a low level.
+The bandpass properties used in this paper are illustrated in Figure~\ref{fig:bandpass}.
+The MWA bandpass of width 30.72~MHz is divided coarsely into 24 1.28~MHz chunks (dashed lines connecting solid circles) with each chunk weighted by a digital filter.
+The coarse channel shape is obtained using a 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$.
+The bandpass correction ({\it plus} symbols connected by dotted lines) corrects for the shape of the coarse channels.
+After this correction, the channels at the edges of these chunks have been flagged (solid lines) because they are known to possibly be contaminated by aliasing at a low level.
 Any spectral flags, or other known weights constitutes $W_f(f)$ in equation~\ref{eqn:obsvis}.
-The bandpass properties used in this paper are illustrated in Figure~\ref{fig:bandpass}.
 
@@
 The filled circles joined by dotted lines show the 8--tap PFB bandpass shapes obtained using a Kaiser window with parameter $\beta=5$.
-The {\it plus} symbols joined by dashed lines are the bandpass correction factors.
-The solid line shows the flagged (gaps) and gain--corrected bandpass shape, $W_f(f)$.
+The bandpass correction factors applied are shown ({\it plus} symbols connected by dashed lines).
+The solid line shows the flagged (gaps) and gain--corrected bandpass shape.
 These shapes are repeated every 1.28~MHz for the entire bandwidth of 30.72~MHz centered around 185~MHz.
@@
 As an example of ``nominal'' observing setup we choose a zenith pointing; as an example of poor foreground conditions, we choose a pointing when the field is $\sim 2$~hours from zenith.
-This pointing has a significantly higher secondary side lobe structure and is recorded when the bright galactic center is well above the horizon.
+This pointing has a significantly higher secondary lobe structure and is observed when the bright galactic center is well above the horizon.
 
-The LSTs of these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at -1.92~hours (i.e., 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings respectively.
+The {\it local sidereal time} (LST) at these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at -1.92~hours (i.e., 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings respectively.
 %These pointings (phased array power--patterns) are centered at RA = 4\fdg 387, Dec = -29\fdg 86 and RA = -28\fdg 8, Dec = -26\fdg 701, whereas the visibilities themselves are phased to zenith corresponding to RA = -28\fdg 8, Dec = -26\fdg 701, and RA = 1\fdg 35, Dec = -26\fdg 701 respectively.
@@
 \caption{MWA tile beam pointing directions during the course of the observation.
-The $x$--axis refers to the Local Sidereal Time (LST) in hours.
+The $x$--axis refers to the {\it local sidereal time} (LST) in hours.
 The axis on the left refers to longitudes, namely, Right Ascension (RA) and Hour Angle (HA) in degrees.
@@
 The axis on the right refers to the declination of the pointing direction.
-The RA, HA, and declination are plotted with solid, dashed and dotted line styles respectively.
+The RA, HA, and declination are plotted with solid, dashed and dotted lines respectively.
 A ``drift and shift'' scheme is used.
-The sky is allowed to drift for $\sim 30$~mins before the beamformer delay settings change in a discrete step to center the beam around RA = 00h 00m 00.00s, Dec = -30\arcdeg$\,$00\arcmin$\,$00\farcs 0.
-Dot--dashed vertical lines show the LST of off-zenith and zenith pointings used in our study.
+The sky is allowed to drift for $\sim 30$~mins before the beamformer delay settings change in a discrete step to center the beam around RA = 0h, Dec = -30\arcdeg.
+Dot--dashed vertical lines show the LST of {\it off-zenith} and {\it zenith} pointings used in our study.
 \label{fig:pointings}}
@@
 
-\subsubsection{Initial Data processing}\label{sec:data-analysis}
+\subsection{Initial Data Processing}\label{sec:data-analysis}
 
 The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off12,off10}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz.
-These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey (MWACS; Hurley-Walker et al, submitted).
+These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey (MWACS; Hurley-Walker et al.
+2014, submitted).
 MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning.
@@
 
-\subsection{Deconvolution along Delay Axis}\label{sec:CLEAN}
+\subsection{Delay Spectrum after Delay Deconvolution}\label{sec:data-delay-spectrum}
 
 We obtain the delay spectrum of these calibrated visibilities taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function.
-The spectrum is multiplied by  the additional weight of the flagged channels which occur rarely for interference and at regular spacing every 1.28~MHz where the edges of coarse channels are known to have a slight aliasing effect.
-In delay--space, these weights translate into a convolution by a ringing harmonic point spread function (PSF).
+The spectrum is multiplied by the additional weight of the flagged channels which occur rarely for interference and regularly every 1.28~MHz where the edges of coarse channels are known to have a slight aliasing effect.
+In delay--space, these weights translate into a convolution by a grated point spread function (PSF) with gratings at multiples of $\eta=0.78\,\mu$s.
 We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis by \citet{par09,par12}.
@@
 We limit the selection of peaks to modes inside the horizon delay limit, corresponding to smooth spectrum objects in the visible sky hemisphere.
-
-\subsection{Delay Spectrum after Delay Deconvolution}\label{sec:data-delay-spectrum}
 
@@
 We show results of delay spectrum from MWA data after deconvolution along delay axis in Figure~\ref{fig:fhd_data}.
-The principle feature is the wedge--shaped structure, the boundaries of which, marked by solid white lines, increases linearly with light travel time corresponding to the baseline length.
+The principle feature is the {\it foreground wedge}, whose boundaries, marked by solid white lines, increase linearly with light travel time corresponding to baseline length.
 Note that the wedge is not filled out uniformly.
@@
 
-In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\overline{\mathbf{k}}|$.
-Hence, the brightest EoR signal will occur on the shortest baselines and smallest delays.
-Thus the region of interest for EoR studies is slightly beyond the horizon delay limits (solid white lines) and inside the first harmonic on either side ($|\eta| \lesssim 0.78\,\mu$s) due to the coarse channel structure.
-We call this region as the {\it EoR window} in this paper.
-
-The {\it off--zenith} pointing has emission notably higher than in the {\it zenith} pointing inside the wedge boundaries.
+The {\it off--zenith} pointing has emission notably higher than in the {\it zenith} pointing inside the {\it foreground wedge}.
 This is shown later to be due to response of the array to the bright Galactic center and Galactic plane in the westward sky.
 Consequently, the contamination into the {\it EoR window} is also higher.
-Emission in the {\it zenith} pointing is more centrally concentrated due to phasing of power--pattern to the zenith, whereas the {\it off--zenith} pointing has a power--pattern phased eastward of zenith and is also responsive to emission from angles far away from zenith.
+Emission in the {\it zenith} pointing is more centrally concentrated due to phasing of power--pattern to the zenith, whereas the {\it off--zenith} pointing has a power--pattern phased eastward of zenith but is also responsive to emission from angles far away from zenith.
 
@@
 \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
-\caption{Delay spectra of two 112~s Murchison Widefield Array observations of the Epoch of Reionization field at RA = 0h, Dec = -30\arcdeg at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has completely set.
+\caption{Delay spectra of two 112~s Murchison Widefield Array observations of the EoR field at RA = 0h, Dec = -30\arcdeg at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
 To get the delay spectrum, each baseline is Fourier transformed into delay space ($y$--axis) and the amplitude plotted in order of baseline length ($x$--axis).
@@
 The logarithmic color scale (shown at the right) is common to both panels.
-White lines mark the boundaries of the {\it foreground wedge} determined by the light travel time across each baseline.
-Smooth spectrum objects are physically limited to the region below this line.
-The various features within the {\it foreground wedge} depend on the \emph{orientation} of the antenna spacing, the secondary lobe structure of the power--pattern and the sky at that instant.
+White lines mark the boundaries of {\it foreground wedge} determined by light travel time across each baseline.
+Smooth spectrum objects predominantly reside in this {\it foreground wedge}.
+The various features in this region depend on antenna spacing orientation, secondary lobe structure of the power--pattern, and the sky at that instant.
 \label{fig:fhd_data}}

### 2014-09-21 foreground_delay_spectrum_modeling.tex
From 2dc2fc1: ============
(*) Minor edits from section 4 to the end
(*) Need to verify color scale on delay spectra

@@
 
-We describe power--pattern and foreground models we have used in modeling the instrument. %todo
+We describe power--pattern and foreground models we have used in our simulations. %todo
 
@@
 \includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom.eps}
-\caption{Models of MWA tile power--pattern in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates for the off--zenith (top) and zenith pointings (bottom) at 185~MHz.
-An MWA tile is modeled as a 4$\times$4 array of isotropic radiators placed 0.3~m above the ground plane.
+\caption{Models of MWA tile power--pattern in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates for the {\it off--zenith} (top) and {\it zenith} pointings (bottom) at 185~MHz.
+An MWA tile is modeled as a 4$\times$4 array of isotropic radiators placed 0.3~m above an infinite ground plane.
 Random fluctuations of rms 0.05~ns have been added to delay corrections during the phased addition of voltages from these isotropic radiators.
-The logarithmic color scale (shown at the right) is common to both panels.
+The logarithmic color scale (shown on the right) is common to both panels.
 \label{fig:power_pattern}}
@@
 The MWA has a very wide field of view ($\gtrsim 20$\arcdeg at 185~MHz) and has non--zero response all the way to the horizon.
-As we have seen above, foreground objects contaminate the region inside the {\it wedge} with an amplitude that depends on the amplitude of power--pattern in that direction, with objects at the horizon being closest to the boundaries of the wedge.
+As already noted, foreground objects fill the {\it foreground wedge} with an amplitude which depends on the tile power gain in that direction, with objects at the horizon being closest to the boundaries of the {\it wedge}.
 
-The degree to which these objects leak power into spectral line of sight modes above the {\it foreground wedge}, depends primarily on the brightness of foregrounds near the boundaries of the {\it wedge} \citep{thy13,pob13,ved12,par12}.
+The degree to which these objects leak power into spectral line of sight modes in the {\it EoR window} will be shown to depend primarily on the brightness of foregrounds near the boundaries of the {\it foreground wedge}.
 Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view.
@@
 For the diffuse component, we use an all--sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz.
-Since this map is predominantly based on the 408~MHz all--sky map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz diffuse emission map to the same resolution.
+Since this map is predominantly based on the 408~MHz all--sky map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz map to the same resolution.
 However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals.
-We model the spectral dependence of the diffuse foreground as a single spectral index at each pixel in our map, found by comparing model maps at 170~MHz and 200~MHz.
+We model the diffuse foreground spectra with a single spectral index at each pixel in the map, found by comparing model maps at 170~MHz and 200~MHz.
 
@@
 Contours of the MWA tile power--pattern shown in Figure~\ref{fig:power_pattern} are also overlaid.
-Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the visible hemisphere in the westward sky and the MWA tile's power receptivity is significantly high ($\gtrsim 12$\%) in that direction.
+Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the visible hemisphere in the westward sky and the MWA tile's power gain is significantly high ($\gtrsim 12$\%) in that direction.
 In the {\it zenith} pointing, the Galactic plane has set and the contour level of power--pattern in that direction is at least 16 times smaller.
@@
 The contour levels shown are 0.00195, 0.00781, 0.0312, 0.125, and 0.5.
-The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off--zenith} pointing and the MWA tile power receptivity is significant ($\gtrsim 12$\%) in that direction.
+The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off--zenith} pointing and the MWA tile power gain is significant ($\gtrsim 12$\%) in that direction.
 In contrast, emission from the Galactic plane in {\it zenith} pointing is significantly lesser.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
-\caption{Simulated delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), for {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
-White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
+\caption{Simulated delay spectrum amplitudes $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+White lines mark the boundaries of {\it foreground wedge} determined by horizon delay limit and antenna spacing.
 The axes and color scale are identical to those in Figure~\ref{fig:fhd_data}.
@@
 
-In order to make a quantitative comparison of delay spectra obtained with the MWA data and our simulations, we consider the uncertainty in the assumed spectral index of our foreground model.
+In order to make a quantitative comparison of delay spectra obtained with MWA data and our simulations, we consider the uncertainty in the assumed spectral index of our foreground model.
 Our foreground models are derived from other higher frequency catalogs and sky maps.
@@
 
-Figure~\ref{fig:data-sim-ratio} shows histograms of the distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}.
+Figure~\ref{fig:data-sim-ratio} shows normalized histograms of the distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}.
 The median absolute deviation from these distributions correspond to $\sim$~90\% fractional difference between data and modeling on average in either case.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-\caption{Histogram of logarithm of ratio of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+\caption{Normalized histograms of logarithm of ratios of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
 Median absolute deviation of $\log_{10}\rho$ is $\sim 0.28$ in both cases.
@@
 These uncertainties are presented only to confirm the qualitative agreement already seen between Figures~\ref{fig:fhd_data} and \ref{fig:sim_data} and are not intended to serve as a comprehensive estimate of all uncertainties involved.
-However, since we have not accounted for numerous other uncertainties, especially in the data, such as uncertainties in the antenna power--pattern, calibration, radio frequency interference (RFI), and wide field imaging effects, our estimates are conservative.
+However, since we have not accounted for numerous other uncertainties, especially in the data, such as uncertainties in the antenna power--pattern, calibration, interference, and wide field imaging effects, our estimates of uncertainty are conservative.
 Accounting for such uncertainties will only improve our confidence level in this comparison.
@@
 
-Having established that the simulation matches the data to the level of expected uncertainties, we proceed to examine in further detail the key signatures seen in the simulated delay spectra.
+Having established that the simulation matches the data to the level of expected uncertainties, we proceed to examine in further detail the key signatures seen in simulated delay spectra.
 A number of factors are responsible for the characteristics noted in the delay spectra obtained from data and through simulations.
@@
 
-\item {\it Sky Model}: Our model of the sky consists of bright compact objects and diffuse emission on diverse spatial scales as seen from Figures~\ref{fig:DSM} and \ref{fig:CSM}.
+\item {\it Sky Model}: Our model of the sky consists of bright compact objects and diffuse emission on diverse spatial scales as shown in Figures~\ref{fig:DSM} and \ref{fig:CSM}.
 It also consists of localized regions of strong emission such as the Galactic plane and Galactic center.
@@
 
-\item {\it Antenna Pair Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our antenna pairs by their orientation, $\theta_\textrm{b}$.
+\item {\it Baseline Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our antenna spacings by their orientation, $\theta_\textrm{b}$.
 We use the following bins: $-67\fdg 5\le \theta_\textrm{b} < -22\fdg 5$, $-22\fdg 5\le \theta_\textrm{b} < 22\fdg 5$, $22\fdg 5\le \theta_\textrm{b} < 67\fdg 5$, and $67\fdg 5\le \theta_\textrm{b} < 112\fdg 5$.
@@
 During observations we allow the sky to drift across the nearest available pointing, shifting between grid points as necessary.
-In our study, we take into account the effect of changing power--pattern of the antennas on the delay spectrum (Figure~\ref{fig:power_pattern}).
+In our study, we take into account the effect of changing tile power--pattern on delay spectrum (Figure~\ref{fig:power_pattern}).
 The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose {\it off--zenith} and {\it zenith} pointings shown in Figure~\ref{fig:pointings}.
@@
 Due to excision of aliased edge channels repeated in each coarse channel of our passband, we observed repeated patterns of the {\it foreground wedge} in the delay spectra.
-As already mentioned earlier, we have used the deconvolution algorithm ({\it CLEAN}) to rid the delay spectra of band shape effects.
+As already mentioned earlier, we have used a deconvolution algorithm ({\it CLEAN}) to rid the delay spectra of band shape effects.
 Since this paper focuses on studying the effects of foreground on the delay spectra, we do not pursue the characterization of artifacts from the band shape and imperfections in {\it CLEAN} deconvolution.
@@
 With the addition of noise in the visibilities which is subject to observing time, some of the weaker features may not be as prominently visible.
-Since the foreground signatures are far too numerous and they reside in a vast volume of parameter space such as antenna pair orientation, power--pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features observed in the delay spectrum of foreground emission.
+Since the foreground signatures are far too numerous and they reside in a vast volume of parameter space such as antenna spacing length and orientation, power--pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features observed in the delay spectra of foreground emission.
 
@@
 \includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-\caption{Simulated delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
+\caption{Simulated delay spectrum amplitudes $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
 White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
@@
 
-The most prominent signature seen in the {\it off--zenith} pointing (top panel of Figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the far west on one of the sidelobes of the power--pattern.
+The most prominent signature seen in the {\it off--zenith} pointing (top panel of Figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power--pattern.
 It appears as a bright branch in the delay spectrum near the negative delay horizon delay limit.
 This feature is strongest at short antenna spacings and fades with increase in antenna spacing.
-This Galactic center signature is almost invisible in the {\it zenith} pointing because the Galactic center has already set over the horizon (see bottom panel in Figure~\ref{fig:noiseless-dsm-delay-spectrum}).
+This Galactic center signature is absent in the {\it zenith} pointing (bottom panel) because the Galactic center has already set over the horizon.
 
@@
 
-As we also see in Figure~\ref{fig:noiseless-dsm-delay-spectrum}, diffuse emission from within the primary field of view manifests in the {\it off--zenith} pointing as a branch at the top of the wedge, while in the {\it zenith} pointing it shows up at near zero delay.
+Figure~\ref{fig:noiseless-dsm-delay-spectrum} also demonstrates diffuse emission from within the primary field of view manifests in the {\it off--zenith} pointing as a branch at the top of the wedge, and at zero delay in the {\it zenith} pointing.
 The former is seen at positive delays because the primary lobe of the power--pattern is centered eastward of zenith, whereas in the latter it is centered at zenith.
@@
 
-A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\overline{\mathbf{x}}|\le$~200~m)---contrary to expectations, since longer baselines are not expected to be sensitive to diffuse emission.
-In the Widefield observations of the MWA, long baselines appear shortened due to projection effects for off--axis emission, especially towards the horizon.
+A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\overline{\mathbf{x}}|\le$~200~m).
+This is contrary to expectations because longer baselines are not expected to be sensitive to diffuse emission.
+In the wide--field observations of the MWA, long baselines appear shortened due to projection effects for off--axis emission, especially towards the horizon.
 This effect is evident at all LSTs in our simulations and it naturally holds for short baselines, as well.
 Thus, diffuse emission from far off--axis directions manifests as an edge--heavy {\it two--pronged fork} across all baselines.
-It decreases in strength with increasing baseline length as expected but is present in all orientations of the baseline vector.
+It decreases in strength with increasing baseline length as expected but is present in all baseline orientations.
 %% This contributes to a {\it pitchfork}--shaped feature which will be discussed later in \S\ref{sec:composite} in more detail.
@@
 
-Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the delay spectrum of compact foreground sky model for {\it off--zenith} and {\it zenith} pointings (shown in Figure~\ref{fig:CSM}) in the top and bottom panels respectively.
+Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the delay spectrum of compact foreground sky model (see Figure~\ref{fig:CSM})for {\it off--zenith} and {\it zenith} pointings in the top and bottom panels respectively.
 
@@
 \includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-\caption{Simulated delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise.
+\caption{Simulated delay spectrum amplitudes $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise.
 White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
 The axes and logarithmic color scale are identical to those in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data} and \ref{fig:noiseless-dsm-delay-spectrum}.
-Emission from compact objects gives rise to centrally concentrated features in the {\it foreground wedge}.\label{fig:noiseless-csm-delay-spectrum}}
+Emission from compact objects gives rise to centrally concentrated features and stays flat across antenna spacings in the {\it foreground wedge}.\label{fig:noiseless-csm-delay-spectrum}}
 \end{figure}
@@
 The amplitude response of an interferometer to a compact object is, to first order, flat with respect to baseline length.
-Therefore, the distribution of objects within the primary field of view depends mainly on how the power--pattern maps to delay space for each antenna pair.
 Since the main lobe of the power--pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with positive delay corresponding to the position of the primary lobe of the power--pattern.
@@
 
-% \subsubsection{Compact Emission in Power--Pattern Sidelobes}\label{sec:csm-side-lobe}
-
-Foreground emission in zero delay and negative delay regions in the {\it off--zenith} pointing is caused by compact objects co--located with sidelobes of the power--pattern.
+Foreground emission in zero delay and negative delay regions in the {\it off--zenith} pointing is caused by compact objects co--located with secondary lobes of the power--pattern.
 On the other hand, compact objects co--located with secondary lobes of power--pattern in the {\it zenith} pointing are revealed as faint but distinct branches at positive and negative delays depending on the orientation of antenna spacing and direction of emission on the sky.
@@
 
-% \subsubsection{{\it Pitchfork} Effect}\label{sec:pitchfork}
-
 When not dominated by the bright emission from the Galactic center, which happens only for certain pointings at certain times, the delay spectrum of the combined foreground model is dominated, in somewhat equal measure, by diffuse and compact emission.
@@
 
-Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional power spectrum plots shown in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}.
+Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional delay spectrum plots shown in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}.
 The diffuse, compact, and composite components are are shown as dotted, dashed, and solid lines respectively.
@@
 The gray shaded area denotes the envelope of expected uncertainty in the delay spectrum.
-Objects inside the {\it foreground} wedge (between the horizon delay limits) are dominated by the $\gtrsim$~70\% error owing to uncertainties in predicting the spectral index of compact foreground objects, outside the wedge thermal noise dominates.
+Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the $\gtrsim$~70\% error owing to uncertainties in predicting the spectral index of compact foreground objects, while thermal noise fluctuations dominate outside.
 
@@
 \includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-\caption{Simulated delay spectrum amplitudes, $|V_\eta(\overline{\mathbf{x}},\eta)|$, for three chosen antenna spacings of length: $\sim$84~m ({\it top}), $\sim$105~m ({\it middle}), and $\sim$170~m ({\it bottom}).
-The antenna spacing vector is specified in each panel along East, North and Upward directions.
+\caption{Simulated delay spectrum amplitudes $|V_\eta(\overline{\mathbf{x}},\eta)|$ for three chosen antenna spacings of length: $\sim$84~m ({\it top}), $\sim$105~m ({\it middle}), and $\sim$170~m ({\it bottom}).
+The antenna spacing is specified in each panel along East, North and Upward directions.
 The dotted, dashed and solid lines denote contributions from diffuse emission, compact objects and composite foreground model respectively.
-The dot--dashed lines mark the horizon delay limits.
-The shaded region denotes the envelope of uncertainty around expected values for composite model.
+Vertical dot--dashed lines mark the horizon delay limits.
+The shaded region denotes the envelope of uncertainty around the composite model delay spectrum.
 This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits.
 Outside the horizon delay limits, it is predominantly due to thermal noise fluctuations.
-Compact objects dominate the central regions of power spectrum while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a {\it pitchfork}--shaped structure.
+Compact objects dominate the central regions of delay spectrum while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a {\it pitchfork}--shaped structure.
 \label{fig:pitchfork-baselines}}
@@
 Near the horizon delay limits at shorter antenna spacing, diffuse emission is brighter than that from compact objects.
-Diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at zero--delay.
+Here, diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at zero--delay.
 In fact, even on widely spaced antennas, {\emph diffuse emission in the delay spectrum near the horizon delay limits exceeds that in the main lobe by more than two orders of magnitude}.
-This is due to apparent shortening of antenna spacings projected in the direction of smooth emission entering far from the primary field of view.
+We attribute this to apparent shortening of antenna spacings projected in the direction of diffuse emission entering far from the primary field of view.
 This demonstrates the edge--heavy {\it two--pronged fork} feature discussed in \S\ref{sec:diffuse-long-baselines}.
@@
 Full simulations show the combination of center--heavy features on all antenna spacings from compact objects and edge--heavy features from both components especially diffuse emission even on wide antenna spacings.
-This results in a characteristic {\it pitchfork} structure in the {\it foreground} wedge that should be evident in observations.
+This results in a characteristic {\it pitchfork} structure imprinted in the {\it foreground} wedge and should be evident in observations.
 
 All these observations are consistent with those from Figures~\ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}.
-The observability of the predicted {\it pitchfork} signature depends on the relative levels of uncertainty in the foreground model and fluctuations from thermal noise.
-In our simulations, since thermal noise in these very short duration snapshots is $\gtrsim 10^7$~Jy~Hz, and features near the horizon delay limits are $\sim 10^7$~Jy~Hz, the {\it pitchfork} feature is not expected to be detected in a noisy scenario although this feature is marginally visible in the {\it zenith} pointing of the observed data (see Figure~\ref{fig:fhd_data}).
+The observability of the {\it pitchfork} signature predicted in this paper depends on relative levels of uncertainty in the foreground model and fluctuations from thermal noise.
+In our simulations, since thermal noise in these very short duration snapshots is $\gtrsim 10^7$~Jy~Hz, and features near the horizon delay limits are $\sim 10^7$~Jy~Hz, the {\it pitchfork} feature is not expected to be detected in a noisy scenario although this feature is marginally visible in the {\it zenith} pointing of observed data (see Figure~\ref{fig:fhd_data}).
 We attribute this to differences between our foreground model and the actual sky.
-Deeper observations should more clearly reveal the prominent feature.
+Deeper observations should reveal the feature clearly.
 
@@
 For instance, there is an increase in the number of secondary peaks in the delay spectrum between zero--delay and horizon delay limits as antenna spacing increases from $\sim$84~m to $\sim$170~m.
-In this case, these correspond to secondary lobes of the power--pattern that lie between the main lobe and the horizon along the North--South direction.
-In short antenna spacings, due to relatively lower resolution along delay axis inside the {\it foreground wedge} and a consequent loss of localization of direction of foreground emission, these secondary peaks blend in with other major peaks and are not distinctly visible.
+In this case, these correspond to secondary lobes of the power--pattern that lie between the primary lobe and the horizon along the local meridian.
+At short antenna spacings, due to relatively lower resolution along delay axis inside the {\it foreground wedge} and a consequent loss of localization of foreground emission, these secondary peaks blend in with other major peaks and are not distinctly visible.
 These findings are consistent with the discussion in \S\ref{sec:compact}.
@@
 In summary, the brightest signature ($\gtrsim 10^{10}$~Jy~Hz) is that of the Galactic center in the {\it off--zenith} pointing co--located at a westward secondary lobe with a significantly high gain.
-The next brightest signature ($\sim 10^8$--$10^9$~Jy~Hz) is caused by compact emission appearing to be concentrated in the inner regions of the {\it foreground wedge} (near zero delays) rather than towards the edges.
-Diffuse emission co--located at the primary lobe of the power--pattern is fainter by approximately an order of magnitude relative to compact emission from the same region for a $\sim 84$~m antenna spacing but unlike the latter, diffuse emission decreases rapidly by over two orders of magnitude as antenna spacing is widened to $\sim 170$~m.
-However, diffuse emission is significantly high near horizon delay limits compared to primary lobe of the power--pattern as antenna spacing is widened.
+The next brightest signature ($\sim 10^8$--$10^9$~Jy~Hz) is caused by compact emission appearing to be concentrated in the inner regions of the {\it foreground wedge} (near zero delays) rather than towards the boundaries.
+Diffuse emission co--located at the primary lobe of the power--pattern is fainter by approximately an order of magnitude relative to compact emission from the same region for a $\sim 84$~m antenna spacing.
+But unlike the latter, diffuse emission decreases rapidly by over two orders of magnitude as antenna spacing is widened to $\sim 170$~m.
+However, diffuse emission is significantly high near horizon delay limits compared to that in the primary lobe of the power--pattern as antenna spacing is widened.
 This leads to edge--heavy features in the delay spectrum.
-Complemented by the center-heavy compact foreground features, we predict the presence of a {\it three--pronged pitchfork}--shaped signature in the delay spectrum of foreground sky and identify evidence for presence of this signature in actual MWA data, although deeper observations are needed to more clearly reveal the effect.
+Complemented by the center-heavy compact foreground features, we predict the presence of a {\it three--pronged pitchfork}--shaped signature in the delay spectrum of foreground sky.
+We identify evidence for presence of this signature in actual MWA data, although deeper observations are needed to more clearly reveal the effect.
 
@@
 While the {\it foreground wedge} is a region of high foreground contamination, of particular interest to the EoR community is foreground contamination near horizon delay limits.
-These regions are responsible for significant spillover into the relatively clean {\it EoR window} due to spectral response of the instrument and foregrounds.
+These regions are believed to be responsible for significant spillover into the relatively clean {\it EoR window} due to spectral response of the instrument and foregrounds.
 
-The Galactic center in the {\it off--zenith} observation is one such example already available in our study.
+The Galactic center in the {\it off--zenith} pointing is one such example already available in our study.
 Figure~\ref{fig:sky-model} shows the sky model ({\it top:} compact objects, {\it bottom:} diffuse sky model) in this pointing.
-The Galactic center is the most dominant source of foreground contamination from the diffuse sky model co--located with a bright secondary lobe of the power pattern.
+The Galactic center is the most dominant source of foreground contamination from the diffuse sky model and is co--located with a bright secondary lobe of the power pattern.
 Figure~\ref{fig:delay-map} shows the sky mapped to delays registered by the baseline vectors, of length 100~m for instance, oriented northward (top panel) and eastward (bottom panel).
@@
 In our example, we discard eastward antenna spacings (-22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5) of lengths $|\overline{\mathbf{x}}| < 40$~m.
-Figure~\ref{fig:before-after} shows the delay spectra obtained with all antenna spacings (top panel) and after applying our screening technique on {\it off--zenith} observation.
+Figure~\ref{fig:before-after} shows the delay spectra obtained with all antenna spacings (top panel) and after applying our screening technique (bottom panel) on the {\it off--zenith} observation.
 Notice the remarkable reduction -- by about an order of magnitude -- in foreground contamination in this region after application of our screening technique.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_before_after_nside_64_Tsys_95.0K.eps}
-\caption{Simulated delay spectra amplitude, $|V_\eta(\overline{\mathbf{x}},\eta)|$, for the {\it off--zenith} observation with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
+\caption{Simulated delay spectra amplitudes $|V_\eta(\overline{\mathbf{x}},\eta)|$ for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
 Discarded antenna spacings have lengths $|\overline{\mathbf{x}}| < 40$~m and orientations -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5.
@@
 In the units presented here, the dynamic range between bright foregrounds and the 21~cm signal is expected to be $\sim 10^8$; a detailed understanding of how foregrounds can corrupt the 21~cm power spectrum is therefore essential.
-This analysis extends previous work by simulating the entire sky rather then just the central field of view and by providing a comparison with early observations from the MWA.
+This analysis extends previous work by simulating the entire sky rather than just the central field of view and by providing a comparison with early observations from the MWA.
 By making use of the delay spectrum technique to estimate power spectrum, we are able to observe the effects of foregrounds while avoiding entanglements with more complex power spectrum estimators.
@@
 On the other hand, compact foreground objects predominantly map onto central regions of the {\it foreground wedge}.
-Features arising from compact objects coincident with primary beam and sidelobes of antenna power--pattern have been noted.
+Features arising from compact objects co--located with primary and secondary lobes of antenna power--pattern have been identified.
 In general, delay spectrum signatures of compact objects are center--heavy, in clear contrast to those from diffuse emission.

### 2014-09-25 foreground_delay_spectrum_modeling.tex
From v0.6.2: ===========
(*) Made minor changes (danny's comments) in the first paragraph of Introduction
(*) Incorporated Judd's suggested changes.
(*) Modified the figures a little bit.
(*) Removed the bandpass figure
(*) Tweaked the title a little bit
(*) Submitted to MWA EoR collaboration

@@
 
-\shorttitle{Foreground signatures in EoR power spectra}
+\shorttitle{Foreground signatures and mitigation in EoR H{\sc i} power spectra}
 \shortauthors{Thyagarajan et~al.}
@@
 %% \title{Limits on the detection of the Epoch of Reionization from MWA observations of the redshifted 21~cm line}
-\title{Signatures of Foreground Sky in Power Spectrum of Redshifted Neutral Hydrogen from the Epoch of Reionization}
+\title{Foreground Signatures and Mitigation in Neutral Hydrogen Power Spectra From Reionization Epoch}
 
@@
 A.~Williams\Curtin, 
-C.~L.~Williams\MIT
+C.~L.~Williams\MIT,
+and~MWA~EoR~members
 }
@@
 
-Detection of 21~cm H{\sc i} emission from the Epoch of Reionization (EoR), at redshifts $z>6$, is limited primarily by foreground emission.
+Detection of 21~cm emission of neutral hydrogen from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission.
 We investigate the chromatic signatures of an all--sky foreground model imprinted by an instrumental transfer function.
 Using the delay spectrum technique, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view.
-Comparing data from recent Murchison Widefield Array (MWA) observations with simulations separated into components based on type of emission, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings.
+Comparing data from recent Murchison Widefield Array observations with simulations separated into components based on type of emission, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings.
 For signals entering through the primary field of view, compact objects dominate the foreground contamination.
@@
 
-After the CMB epoch, the period in the history of the universe often referred to as the {\it dark ages}, followed the epoch of reionization (EoR).
-This was a period of non-linear growth of density perturbations and astrophysical evolution.
+At the end of the recombination epoch, the universe turned completely neutral.
+This period, referred to as the {\it Dark Ages} in the universe's history, is characterized by collapse of matter under the influence of gravity.
+And it ended with the formation of the first stars and galaxies which started emitting ultra--violet and X--ray radiation, thereby reionizing the neutral medium in their surroundings.
+This commenced the epoch of reionization (EoR) --- a period of non--linear growth of matter density perturbations and astrophysical evolution.
 Studying the epoch of reionization holds the key to understanding this evolution.
-Direct observation of redshifted 21~cm spin flip transition of H{\sc i} has been identified as a useful probe of the EoR \citep{sun72,sco90,mad97,toz00,ili02} and has recently emerged as a very promising experiment to fill the gaps in our understanding of the universe's history.
+
+Observation of redshifted 21~cm spin flip transition of H{\sc i} has been identified as a direct probe of this epoch \citep{sun72,sco90,mad97,toz00,ili02}.
+Detecting this signal has recently emerged as a very promising experiment to fill the gaps in our understanding of the universe's history.
 
@@
 
-A key challenge in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
+A key challenge in the statistical detection of power spectrum of this redshifted H{\sc i} 21~cm signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
 \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}.
@@
 $\eta$ and $k_\parallel$, denoting spatial scales along line of sight form the $y$--axis.
-Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge}, the boundaries of which are determined by antenna spacing and the light travel time across it.
+Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge}, whose boundaries are determined by antenna spacing and the light travel time across it.
 These boundaries are called horizon delay limits and are shown by solid lines.
@@
 \caption{Illustration of Fourier space in which delay (and power) spectra of H{\sc i} signals from the EoR are presented.
-The $x$--axis denoted by $|\overline{\mathbf{x}}|$ (antenna spacing in m) or $k_\perp$ (transverse wavenumber in units of $h$~Mpc$^{-1}$) are equivalent and represent transverse spatial scales.
-Similarly, the $y$--axis denoted by $\eta$ (delay in s or $\mu$s) or equivalently, $k_\parallel$ (line--of--sight wavenumber in units of $h$~Mpc$^{-1}$) represents spatial scales along the line of sight.
-The dark shaded region is often referred to as {\it foreground wedge} where smooth spectrum foregrounds reside.
-The boundaries of the {\it foreground wedge} (solid lines) are given by the light travel time for corresponding antenna spacings and are referred to as horizon delay limits.
-In principle, region outside the wedge is expected to be relatively free of foreground contamination and is called the {\it EoR window} (light and medium shades).
-In practice, this contamination is also high in the narrow unshaded white regions (bound by solid and dashed lines) due to convolution by instrumental transfer function.
-The width of this narrow region is determined by inverse of operating bandwidth.
-Due to an undesirable spectral response of the instrument used in this study, namely the MWA, we identify a narrower region of high significance to EoR studies (medium shade) and hereafter refer to it as the {\it EoR window} in this paper.
+The $x$--axis is denoted by $|\overline{\mathbf{x}}|$ (antenna spacing) or $k_\perp$ (transverse wavenumber).
+The $y$--axis denoted by $\eta$ (delay) or $k_\parallel$ (line--of--sight wavenumber).
+The dark shaded region is referred to as {\it foreground wedge} where smooth spectrum foregrounds reside.
+Its boundaries (solid lines), given by light travel time for corresponding antenna spacings, are referred to as horizon delay limits.
+Regions outside the {\it wedge} and the narrow white unshaded strips (caused by convolution with the instrumental transfer function) are expected to be relatively free of foreground contamination.
+Due to an undesirable grating pattern (dot--dashed lines) of the instrument used in this study, namely the MWA, we identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the {\it EoR window} in this paper.
 \label{fig:fourier-space}}
@@
 
-The terms antenna spacing and baseline are synonymous in radio interferometry.
+% The terms antenna spacing and baseline are synonymous in radio interferometry.
 The latter term is also used broadly to refer to the antenna pair system in a radio interferometer array.
@@
 The MWA was completed in 2012, and after commissioning, began its EoR observing program in 2013.
-In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30~MHz band anywhere in the range 80--300~MHz.
+In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz.
 Each tile is a phased array of 16 dipoles each in the shape of a bow--tie.
@@
 
-The bandpass properties used in this paper are illustrated in Figure~\ref{fig:bandpass}.
+The MWA passband of width 30.72~MHz is divided coarsely into 24 1.28~MHz chunks with each chunk weighted by a digital filter.
+The coarse channel shape is obtained using a 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$.
+After correcting for the shape of these coarse channels, the channels at the edges of these chunks are flagged because they are known to be possibly contaminated by aliasing at a low level.
+Any spectral flags, or other known weights constitutes $W_f(f)$ in equation~\ref{eqn:obsvis}.
+
+% The bandpass properties used in this paper are illustrated in Figure~\ref{fig:bandpass}.
 The MWA bandpass of width 30.72~MHz is divided coarsely into 24 1.28~MHz chunks (dashed lines connecting solid circles) with each chunk weighted by a digital filter.
@@
 
-\begin{figure}[htb]
-\centering
-\includegraphics[width=\linewidth]{figures/v1_0/bandpass_properties.eps}
-\caption{Bandpass properties used in modeling and analysis of delay spectrum of MWA visibilities.
+% \begin{figure}[htb]
+% \centering
+% \includegraphics[width=\linewidth]{figures/v1_0/bandpass_properties.eps}
+% \caption{Bandpass properties used in modeling and analysis of delay spectrum of MWA visibilities.
 The filled circles joined by dotted lines show the 8--tap PFB bandpass shapes obtained using a Kaiser window with parameter $\beta=5$.
@@
 \label{fig:bandpass}}
-\end{figure}
+% \end{figure}
 
@@
 The MWA targets two primary low-foreground fields for reionization observations.
-Here, we focus on the field at RA = 0h, Dec = -30\arcdeg.
+Here, we focus on the field at RA = 0h, Dec = $-30$\arcdeg.
 The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of an array of dipoles in an MWA tile.
-These beamformer delay settings can be changed only in discrete steps.
-The sky is allowed to drift for a certain period of time (usually $\sim 30$~mins) before the discrete shifts in the beamformer delay settings position the beam to be centered again on the same patch of sky.
+The pointing system is capable of steering to points on a regular $\sim$7\arcdeg grid.
+During observations we allow the sky to drift across the nearest available pointing, shifting between grid points as necessary (once in $\sim 30$~minutes).
 This process is repeated throughout the course of the observation $\approx 4.86$~hours.
@@
 
-The {\it local sidereal time} (LST) at these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at -1.92~hours (i.e., 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings respectively.
+The {\it local sidereal time} (LST) at these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at $-1.92$~hours (i.e., 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings respectively.
 %These pointings (phased array power--patterns) are centered at RA = 4\fdg 387, Dec = -29\fdg 86 and RA = -28\fdg 8, Dec = -26\fdg 701, whereas the visibilities themselves are phased to zenith corresponding to RA = -28\fdg 8, Dec = -26\fdg 701, and RA = 1\fdg 35, Dec = -26\fdg 701 respectively.
@@
 \caption{MWA tile beam pointing directions during the course of the observation.
-The $x$--axis refers to the {\it local sidereal time} (LST) in hours.
-The axis on the left refers to longitudes, namely, Right Ascension (RA) and Hour Angle (HA) in degrees.
+The $x$--axis refers to the {\it local sidereal time} (LST).
+The axis on the left refers to longitudes, namely, Right Ascension (RA) and Hour Angle (HA).
 Negative values of RA, HA, and LST are to be interpreted as having been wrapped around by 360\arcdeg or 24 hours.
@@
 The RA, HA, and declination are plotted with solid, dashed and dotted lines respectively.
-A ``drift and shift'' scheme is used.
-The sky is allowed to drift for $\sim 30$~mins before the beamformer delay settings change in a discrete step to center the beam around RA = 0h, Dec = -30\arcdeg.
+The sky is allowed to drift for $\sim 30$~mins before the beamformer delay settings change in a discrete step to center the beam around RA = 0h, Dec = $-30$\arcdeg.
 Dot--dashed vertical lines show the LST of {\it off-zenith} and {\it zenith} pointings used in our study.
@@
 
-\subsection{Delay Spectrum after Delay Deconvolution}\label{sec:data-delay-spectrum}
+\subsection{Deconvolved Delay Spectrum}\label{sec:data-delay-spectrum}
 
@@
 \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
-\caption{Delay spectra of two 112~s Murchison Widefield Array observations of the EoR field at RA = 0h, Dec = -30\arcdeg at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
+\caption{Delay spectra of two observations of the EoR field (RA = 0h, Dec = $-30$\arcdeg) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
 To get the delay spectrum, each baseline is Fourier transformed into delay space ($y$--axis) and the amplitude plotted in order of baseline length ($x$--axis).
@@
 
-We describe power--pattern and foreground models we have used in our simulations. %todo
+We describe the instrumental model and foreground models we use in our simulations.
 
-\subsection{Power--pattern}\label{sec:power_pattern}
+\subsection{Instrument Model}\label{sec:instrument_model}
 
-The power--pattern of the MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is modeled as a mutually--coupled 4--by--4 dipole array with the overall power--pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14}.
+The instrument model consists primarily of the tile beam power--pattern, $A(\overline{\mathbf{l}},f)$ (see equation~\ref{eqn:obsvis}).
+It is modeled as a mutually--coupled 4--by--4 dipole array with the overall power--pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14}.
 To speed up simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom.eps}
-\caption{Models of MWA tile power--pattern in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates for the {\it off--zenith} (top) and {\it zenith} pointings (bottom) at 185~MHz.
+\caption{Models of MWA tile power--pattern for {\it off--zenith} (top) and {\it zenith} pointings (bottom) at 185~MHz.
 An MWA tile is modeled as a 4$\times$4 array of isotropic radiators placed 0.3~m above an infinite ground plane.
@@
 Contours of the MWA tile power--pattern shown in Figure~\ref{fig:power_pattern} are also overlaid.
-Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the visible hemisphere in the westward sky and the MWA tile's power gain is significantly high ($\gtrsim 12$\%) in that direction.
+Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the westward sky and the MWA tile's power gain is significant ($\gtrsim 12$\%) in that direction.
 In the {\it zenith} pointing, the Galactic plane has set and the contour level of power--pattern in that direction is at least 16 times smaller.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/dsm.eps}
-\caption{Sky brightness temperature (in K) of the diffuse foreground model at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+\caption{Sky brightness temperature of the diffuse foreground model at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
 The color scale on the right is logarithmic and is common to both panels.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/csm.eps}
-\caption{Flux densities (in Jy) of bright compact objects at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+\caption{Flux densities of bright compact objects at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
 The color scale on the right is logarithmic and is common to both panels and corresponds to the color--coded filled circles.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
-\caption{Simulated delay spectrum amplitudes $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
-White lines mark the boundaries of {\it foreground wedge} determined by horizon delay limit and antenna spacing.
-The axes and color scale are identical to those in Figure~\ref{fig:fhd_data}.
-The features resemble those obtained with MWA data shown in Figure~\ref{fig:fhd_data}.
-The Galactic center and Galactic plane prominently visible in the {\it off--zenith} pointing makes the emission in the {\it foreground wedge} brighter relative to that in the {\it zenith} pointing.
+\caption{Same as Figure~\ref{fig:fhd_data}, but generated from simulated visibilities using the model described in \S\ref{sec:modeling}.
+Note the similarity between the broad features to those in Figure~\ref{fig:fhd_data}.
 \label{fig:sim_data}}
@@
 
-% Figure~\ref{fig:data-sim-ratio} shows the significance of $\log_{10}\rho$ relative to the expected error $\Delta\,\log_{10}\rho$ for all $|\overline{\mathbf{x}}|$ and $\eta$ inside the {\it foreground wedge}.
-The probability densities and cumulative distributions are shown in black and gray respectively.
-Also shown are dotted vertical lines which denote equality between $\Delta\,\log_{10}\rho$ and $\log_{10}\rho$.
-The fraction of the {\it foreground wedge} that lies in between these vertical lines is $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively.
-In other words, subject to expected uncertainties, the ratio of delay spectra obtained with MWA data and simulations are consistent with unity at $\sim$~90\% and $\sim$~95\% confidence levels in the {\it off--zenith} and {\it zenith} pointings respectively.
-
 Figure~\ref{fig:data-sim-ratio} shows normalized histograms of the distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}.
@@
 Median absolute deviation of $\log_{10}\rho$ is $\sim 0.28$ in both cases.
-A value of zero indicates when the ratio $\rho=1$.
+A value of zero indicates $\rho=1$.
 \label{fig:data-sim-ratio}}
-% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-% \caption{Significance of logarithm of ratio of observed to simulated visibilities relative to expected uncertainty restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
-Probability densities and cumulative distributions are shown in black and gray respectively.
-Region inside the dotted vertical lines denotes when the ratio, $\rho$, is consistent with unity subject to expected uncertainty.
-Also indicated in percentages are confidence levels estimated conservatively for such a consistency.
-They are $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively.
-\label{fig:data-sim-ratio-snr}}
 \end{figure}
 
-These uncertainties are presented only to confirm the qualitative agreement already seen between Figures~\ref{fig:fhd_data} and \ref{fig:sim_data} and are not intended to serve as a comprehensive estimate of all uncertainties involved.
-However, since we have not accounted for numerous other uncertainties, especially in the data, such as uncertainties in the antenna power--pattern, calibration, interference, and wide field imaging effects, our estimates of uncertainty are conservative.
-Accounting for such uncertainties will only improve our confidence level in this comparison.
+These uncertainties are presented to confirm the qualitative agreement already seen between Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}.
+These uncertainty estimates are conservative.
+A full treatment of all uncertainties and deviations from ideal behavior such as uncertainties in tile power--pattern, calibration, data corruption due to interference, and wide--field imaging effects will bring the simulations much closer in agreement with observations.
 Our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.
@@
 In fact, the patch of sky for MWA observations is chosen from regions of low foreground emission.
-% In our detailed study, we have divided the transiting sky into four different ``bow-tie'' shaped regions.
-Each region has equal area and the regions are aligned to cardinal directions around zenith as the origin.
-This results in the ability to isolate features in the delay spectrum to eight different sectors of the sky.
+We will investigate the signatures of each component and their potential impact on {\it EoR window} contamination.
 
@@
 \item {\it Tile Pointing and Power--Pattern}: Since the MWA tiles are steered electronically, the power--pattern of the tile changes in any observing mode that tracks the object.
-The pointing system is capable of steering to points on a regular $\sim$7\arcdeg grid.
-During observations we allow the sky to drift across the nearest available pointing, shifting between grid points as necessary.
 In our study, we take into account the effect of changing tile power--pattern on delay spectrum (Figure~\ref{fig:power_pattern}).
 The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose {\it off--zenith} and {\it zenith} pointings shown in Figure~\ref{fig:pointings}.
+We will show the secondary lobe structures in the tile power--pattern are tied to the foreground signatures in Fourier space.
 
-\item {\it Instrumental Passband}: The instrumental passband characteristics have a direct effect on the delay transform owing to the Fourier relationship between the two.
+% \item {\it Instrumental Passband}: The instrumental passband characteristics have a direct effect on the delay transform owing to the Fourier relationship between the two.
 Due to excision of aliased edge channels repeated in each coarse channel of our passband, we observed repeated patterns of the {\it foreground wedge} in the delay spectra.
@@
 
-While the {\it foreground wedge} is a region of high foreground contamination, of particular interest to the EoR community is foreground contamination near horizon delay limits.
+% While the {\it foreground wedge} is a region of high foreground contamination, of particular interest to the EoR community is foreground contamination near horizon delay limits.
 These regions are believed to be responsible for significant spillover into the relatively clean {\it EoR window} due to spectral response of the instrument and foregrounds.
+
+Here, we will investigate the susceptibility of particular antenna spacings to foreground contamination arising out of bright foreground objects located near the horizon and present a technique to substantially mitigate such contamination.
 
@@
 Figure~\ref{fig:sky-model} shows the sky model ({\it top:} compact objects, {\it bottom:} diffuse sky model) in this pointing.
-The Galactic center is the most dominant source of foreground contamination from the diffuse sky model and is co--located with a bright secondary lobe of the power pattern.
+The Galactic center is the most dominant source of foreground contamination from the diffuse sky model and is co--located with a bright secondary lobe of the power pattern near the western horizon.
 Figure~\ref{fig:delay-map} shows the sky mapped to delays registered by the baseline vectors, of length 100~m for instance, oriented northward (top panel) and eastward (bottom panel).
@@
 Figure~\ref{fig:before-after} shows the delay spectra obtained with all antenna spacings (top panel) and after applying our screening technique (bottom panel) on the {\it off--zenith} observation.
-Notice the remarkable reduction -- by about an order of magnitude -- in foreground contamination in this region after application of our screening technique.
+Notice the remarkable reduction --- by about an order of magnitude --- in foreground contamination in this region after application of our screening technique.
 
@@
 \caption{Simulated delay spectra amplitudes $|V_\eta(\overline{\mathbf{x}},\eta)|$ for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
-Discarded antenna spacings have lengths $|\overline{\mathbf{x}}| < 40$~m and orientations -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5.
-The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} prominent when all antenna spacings are used is reduced by an order of magnitude when short eastward antenna spacings are discarded.
+Discarded antenna spacings (black vertical stripes) have lengths $|\overline{\mathbf{x}}| < 40$~m and orientations -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5.
+The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} prominent when all antenna spacings are used is lowered by an order of magnitude when short eastward antenna spacings are discarded.
+Color scale has been adjusted to emphasize mitigation achieved in foreground contamination.
 \label{fig:before-after}}
@@
 
-We further demonstrate this screening technique can be customized when it is necessary to optimize between foreground mitigation and loss of sensitivity from discarding data.
+This screening technique can be customized when it is necessary to optimize between foreground mitigation and loss of sensitivity from discarding data.
 Figure~\ref{fig:screening} shows how the typical foreground contamination in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
@@
 It is notable that short antenna spacings play maximum role in determining foreground contamination.
-For instance, among antenna spacings with orientations $|\theta_\textrm{b}|\le$~22\fdg 5, discarding those with lengths $|\overline{\mathbf{x}}|>40$~m has no effect on foreground contamination and would only lead to loss of sensitivity.
+For instance, among antenna spacings with orientations $|\theta_\textrm{b}|\le$~22\fdg 5, discarding those with lengths $|\overline{\mathbf{x}}|>40$~m has no effect on further foreground contamination and would only lead to loss of sensitivity.
 On the other hand, a bright compact object at the same location will give rise to foreground contamination even on longer antenna spacings.
@@
 Short baselines have maximum effect on foreground contamination.
-For instance, discarding antenna spacings oriented $|\theta_\textrm{b}|\le$~22\fdg 5 with lengths $|\overline{\mathbf{x}}|>40$~m has no effect in reducing foreground contamination.
+For instance, discarding antenna spacings oriented $|\theta_\textrm{b}|\le$~22\fdg 5 with lengths $|\overline{\mathbf{x}}|>40$~m has no effect in further reducing foreground contamination.
 \label{fig:screening}}

### 2014-10-10 foreground_delay_spectrum_modeling.tex
From v0.7-MWA-EoR: =================
(*) Incorporated most of Steven Tingay's comments on the first review of draft

@@
 %% \title{Limits on the detection of the Epoch of Reionization from MWA observations of the redshifted 21~cm line}
-\title{Foreground Signatures and Mitigation in Neutral Hydrogen Power Spectra From Reionization Epoch}
+\title{Foreground Signatures and Mitigation in Neutral Hydrogen Power Spectra From The Reionization Epoch}
 
@@
 
-At the end of the recombination epoch, the universe turned completely neutral.
-This period, referred to as the {\it Dark Ages} in the universe's history, is characterized by collapse of matter under the influence of gravity.
+At the end of the recombination epoch, the Universe was completely neutral.
+This period, referred to as the {\it Dark Ages} in the Universe's history, is characterized by the localized accumulation of matter under the influence of gravity.
 And it ended with the formation of the first stars and galaxies which started emitting ultra--violet and X--ray radiation, thereby reionizing the neutral medium in their surroundings.
 This commenced the epoch of reionization (EoR) --- a period of non--linear growth of matter density perturbations and astrophysical evolution.
-Studying the epoch of reionization holds the key to understanding this evolution.
+Studying the EoR holds the key to understanding this evolution.
 
-Observation of redshifted 21~cm spin flip transition of H{\sc i} has been identified as a direct probe of this epoch \citep{sun72,sco90,mad97,toz00,ili02}.
-Detecting this signal has recently emerged as a very promising experiment to fill the gaps in our understanding of the universe's history.
+Observations of redshifted 21~cm radiation generated by the spin flip transition of H{\sc i} has been identified as a direct probe of the EoR \citep{sun72,sco90,mad97,toz00,ili02}.
+Detecting this signal has recently emerged as a very promising experiment to fill the gaps in our understanding of the Universe's history.
 
-Sensitive instruments such as the Square Kilometer Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}.
+Sensitive instruments such as the Square Kilometre Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}.
 Numerous precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{tin13,bow13,lon09}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{thy13,bea13,bow06,pob14}.
 
-A key challenge in the statistical detection of power spectrum of this redshifted H{\sc i} 21~cm signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
+A key challenge in the statistical detection of the redshifted H{\sc i} 21~cm signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
 \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}.
+
 Considerable effort has been made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,liu11,dat10,bow09,liu09}.
@@
 
-Foregrounds contaminate the EoR H{\sc i} power spectrum by either having intrinsic spectral variability, high instrumental chromaticity, or through approximations in power spectrum analysis.
+Foregrounds contaminate the EoR H{\sc i} power spectrum by either having intrinsic spectral variability, high instrumental chromaticity, or through approximations in power spectrum analyses.
 Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity.
-To decouple our findings from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the individual baseline--based approach of \citet{pob13,par14}.
+To decouple these effects from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the individual baseline--based approach of \citet{pob13} and \citet{par14}.
 This approximates the power spectrum as the inverse Fourier transform of the spectra generated by the instrument's correlator.
@@
 We quantify this contamination by separating the simulations into components based on type of foreground emission.
-Using these we arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.
+We then arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.
 
-In \S\ref{sec:delay-spectrum} we introduce the delay spectrum technique, then in  \S\ref{sec:instrument}, we describe the MWA setup, summarize the observing parameters and present the resulting data.
-Simulations of these observing parameters are described in \S\ref{sec:modeling} and analyzed by sky location, and type of emission in \S\ref{sec:delay-spectrum-analysis}.
+In \S\ref{sec:delay-spectrum} we introduce the delay spectrum technique.
+In \S\ref{sec:instrument}, we describe the MWA setup, summarize the observing parameters, and present the resulting data.
+Simulations using these observing parameters are described in \S\ref{sec:modeling} and analyzed by sky location and type of emission in \S\ref{sec:delay-spectrum-analysis}.
 In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique.
-\S\ref{sec:summary} summarizes our work and findings.
+We present a summary of our work and findings in \S\ref{sec:summary}.
 
@@
 
-Interferometer array data known as {\it visibilities}, $V_f(\overline{\mathbf{u}},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\overline{\mathbf{x}}$ and then Fourier transformed along time axis to obtain a spectrum along frequency ($f$) axis.
-Here, $\overline{\mathbf{u}}\equiv \overline{\mathbf{x}}f/c \equiv \overline{\mathbf{x}}/\lambda$, $c$ is the speed of light and $\lambda$ is the wavelength.
-$\overline{\mathbf{u}}$ is related to transverse spatial frequency modes ($\overline{\mathbf{k}}_\perp$) of brightness distribution of the sky as $\overline{\mathbf{k}}_\perp \equiv 2\pi\overline{\mathbf{u}}/D(z)$, where $D(z)$ is the transverse comoving distance at a redshift $z$.
-If $I(\overline{\mathbf{l}},f)$ is the emission at different frequencies on the sky as a function of direction unit vector ($\overline{\mathbf{l}}$) specified as direction cosines, $V_f(\overline{\mathbf{u}},f)$ represents the Fourier decomposition of $I(\overline{\mathbf{l}},f)$ attenuated by the antenna's angular power--pattern $A(\overline{\mathbf{l}},f)$, into transverse spatial frequency modes.
+Interferometer array data known as {\it visibilities}, $V_f(\overline{\mathbf{u}},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\overline{\mathbf{x}}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis.
+Here, $\overline{\mathbf{u}}\equiv \overline{\mathbf{x}}f/c \equiv \overline{\mathbf{x}}/\lambda$, $c$ is the speed of light, and $\lambda$ is the wavelength of radio waves.
+$\overline{\mathbf{u}}$ is related to the transverse spatial frequency modes ($\overline{\mathbf{k}}_\perp$) of the brightness distribution of the sky as $\overline{\mathbf{k}}_\perp \equiv 2\pi\overline{\mathbf{u}}/D(z)$, where $D(z)$ is the transverse comoving distance at redshift $z$.
+If $I(\overline{\mathbf{l}},f)$ is the emission at different frequencies on the sky as a function of direction unit vector ($\overline{\mathbf{l}}$) specified as direction cosines, $V_f(\overline{\mathbf{u}},f)$ represents the Fourier decomposition of $I(\overline{\mathbf{l}},f)$ attenuated by the antenna's angular power--pattern, $A(\overline{\mathbf{l}},f)$, into transverse spatial frequency modes.
 Taking into account the instrumental bandpass weights $W_f(f)$, this relation can be written as:
@@
 
-We define the {\it delay spectrum} $V_\eta(\overline{\mathbf{u}},\eta)$ to be the inverse Fourier transform of $V_f(\overline{\mathbf{u}},f)$ along the frequency coordinate:
+We define the {\it delay spectrum}, $V_\eta(\overline{\mathbf{u}},\eta)$, to be the inverse Fourier transform of $V_f(\overline{\mathbf{u}},f)$ along the frequency coordinate:
 \begin{align}\label{eqn:delay-transform}
@@
 \end{align}
-where, $W'_f(f)$ is a frequency window weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}.
-$\eta$ represents the signal delay between antenna pairs:
+where, $W'_f(f)$ is a frequency window weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12} and $\eta$ represents the signal delay between antenna pairs:
 \begin{equation}\label{eqn:delay}
@@
 \end{equation}
-Note that as the visibility is a fully complex quantity, its Fourier transform is fully Nyquist sampled over positive and negative delays.
+Note that as the visibility is a complex quantity, its Fourier transform is Nyquist sampled over positive and negative delays.
 Negative and positive delay\footnote{The terms delay and lag are used interchangeably in this paper.} correspond to the two hemispheres of the sky as transected by the antenna spacing vector.
 In this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the telescope site.
-Hence, a perfectly Eastward oriented antenna spacing will observe an object in the Eastern sky at positive delay, and conversely for an object in the Western sky.
-Similarly, an object in the Northern sky will appear at a positive delay at an antenna spacing oriented Northward.
+Hence, a perfectly eastward oriented antenna spacing will observe an object in the eastern sky at positive delay, and conversely for an object in the western sky.
+Similarly, an object in the northern sky will appear at a positive delay at an antenna spacing oriented northward.
 
 Since we are studying a redshifted H{\sc i} spectral line, $f$ is a measure of cosmological distance along the line of sight.
-Consequently, $\eta$ is a measure of line of sight spatial frequency mode given by $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
+Consequently, $\eta$ is a measure of line--of--sight spatial frequency mode given by $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
 We use $H_0=100\,h$~km~s$^{-1}$~Mpc$^{-1}$ and assume a flat $\Lambda$CDM universe with $\Omega_\textrm{M}=0.27$ and $\Omega_\Lambda=0.73$.
@@
 Thus, $|V_\eta(\overline{\mathbf{u}},\eta)|^2$ is directly related to the spatial power spectrum of EoR H{\sc i} distribution, $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ \citep{mor04}.
-However, foreground objects also emit in radio frequencies.
+However, foreground objects also emit at radio frequencies.
 Due to their spatial and spectral (chromatic) structure, the measured $V_\eta(\overline{\mathbf{u}},\eta)$ is contaminated by foreground radio objects \citep{thy13,tro12,mor12,bow09}.
@@
 In summary, $V_\eta(\overline{\mathbf{u}},\eta)$ is closely related to visibilities, $V_f(\overline{\mathbf{u}},f)$, the basic data blocks measured by each antenna pair, as seen from equation~\ref{eqn:delay-transform}.
-It captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument.
+$V_\eta(\overline{\mathbf{u}},\eta)$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument.
 At the same time, it is directly related to the sought power $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ in $k$--modes containing critical information about spatial scales.
 For our analysis, we use an equivalent quantity, $V_\eta(\overline{\mathbf{x}},\eta) \equiv V_\eta(\lambda\overline{\mathbf{u}},\eta)$.
-For purposes of illustration, we often use a simplified quantity $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$.
+For purposes of illustration, we often use a simplified quantity, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$.
 
@@
 $|\overline{\mathbf{x}}|$ and $k_\perp$, denoting spatial scales in the transverse direction, form the $x$--axis.
-$\eta$ and $k_\parallel$, denoting spatial scales along line of sight form the $y$--axis.
+$\eta$ and $k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis.
 Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge}, whose boundaries are determined by antenna spacing and the light travel time across it.
 These boundaries are called horizon delay limits and are shown by solid lines.
-Spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\eta$--axis.
+The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\eta$--axis.
 This effect has been quantified in \citet{thy13}.
-The width of this narrow strip is inversely proportional to operating bandwidth.
-Region of Fourier space excluding the {\it foreground wedge} and the narrow strips is usually called the {\it EoR window}.
-These are shown in light and medium shades.
+The width of this narrow strip is inversely proportional to the operating bandwidth.
+The region of Fourier space excluding the {\it foreground wedge} and the narrow strips is the so--called {\it EoR window}, shown in light and medium shades.
 However, the instrument used in our study, namely the MWA, has a passband constructed using coarse channels with a characteristic shape giving rise to periodic gratings thereby resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.
@@
 The $y$--axis denoted by $\eta$ (delay) or $k_\parallel$ (line--of--sight wavenumber).
-The dark shaded region is referred to as {\it foreground wedge} where smooth spectrum foregrounds reside.
+The dark shaded region is referred to as the {\it foreground wedge} where smooth spectrum foregrounds reside.
 Its boundaries (solid lines), given by light travel time for corresponding antenna spacings, are referred to as horizon delay limits.
@@
 
-The MWA was completed in 2012, and after commissioning, began its EoR observing program in 2013.
+The MWA was completed in 2012 and, after commissioning, began its EoR observing program in 2013.
 In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz.
-Each tile is a phased array of 16 dipoles each in the shape of a bow--tie.
-This yields a primary field of view $\gtrsim$~20\arcdeg wide and multiple secondary lobes for the tile.
-The tile arrangement is a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius with the longest baselines at 3~km \citep{bea12}.
+Each tile is a phased array of 16 dipoles, each in the shape of a bow--tie.
+This yields a primary field of view $\gtrsim$~20\arcdeg wide and multiple secondary lobes.
+The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km \citep{bea12}.
 Here we focus on antenna spacings $|\overline{\mathbf{x}}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz.
-The center frequency 185~MHz corresponds to a redshift, $z\simeq 6.68$.
+The center frequency of 185~MHz corresponds to a redshift, $z\simeq 6.68$.
 
-The MWA passband of width 30.72~MHz is divided coarsely into 24 1.28~MHz chunks with each chunk weighted by a digital filter.
+The MWA passband of width 30.72~MHz is divided coarsely into 24$\times$1.28~MHz sub--bands with each sub--band weighted by a digital filter.
 The coarse channel shape is obtained using a 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$.
-After correcting for the shape of these coarse channels, the channels at the edges of these chunks are flagged because they are known to be possibly contaminated by aliasing at a low level.
+After correcting for the shape of these coarse channels, the fine channels (40~kHz width) at the edges of these sub--bands are flagged because they are known to be possibly contaminated by aliasing at a low level.
 Any spectral flags, or other known weights constitutes $W_f(f)$ in equation~\ref{eqn:obsvis}.
@@
 Here, we focus on the field at RA = 0h, Dec = $-30$\arcdeg.
-The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of an array of dipoles in an MWA tile.
+The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of the dipoles in an MWA tile.
 The pointing system is capable of steering to points on a regular $\sim$7\arcdeg grid.
@@
 
-The observations shown here were observed on Aug 23, 2013.
-This day has been chosen as a reference data set for comparison within the MWA EoR collaboration and was selected on the basis of low number of known system errors from among a long stretch of stable observing.
+The observations shown here were undertaken on Aug 23, 2013.
+This day has been chosen as a reference data set for comparison within the MWA EoR collaboration and was selected on the basis of a low number of known system errors from among a long stretch of stable observing.
 
@@
 These were chosen to provide a selection of possible foreground and instrumental conditions.
-As an example of ``nominal'' observing setup we choose a zenith pointing; as an example of poor foreground conditions, we choose a pointing when the field is $\sim 2$~hours from zenith.
+As an example of a nominal observing setup we choose a zenith pointing; as an example of poor foreground conditions, we choose a pointing when the field is $\sim 2$~hours from zenith.
 This pointing has a significantly higher secondary lobe structure and is observed when the bright galactic center is well above the horizon.
 
-The {\it local sidereal time} (LST) at these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at $-1.92$~hours (i.e., 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings respectively.
+The {\it local sidereal time} (LST) at these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at $-1.92$~hours (i.e., 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings, respectively.
 %These pointings (phased array power--patterns) are centered at RA = 4\fdg 387, Dec = -29\fdg 86 and RA = -28\fdg 8, Dec = -26\fdg 701, whereas the visibilities themselves are phased to zenith corresponding to RA = -28\fdg 8, Dec = -26\fdg 701, and RA = 1\fdg 35, Dec = -26\fdg 701 respectively.
@@
 2014, submitted).
-MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning.
-It has a flux density limit of 25~mJy and a declination range of -12\arcdeg to -40\arcdeg evenly covering the field of view of the observations reported here.
+MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning; it has a flux density limit of 25~mJy and a declination range of -12\arcdeg to -40\arcdeg evenly covering the field of view of the observations reported here.
 The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern.
-The calibration algorithm -- based on a forward modeling software by \citet{sul12} -- computes complex gain solutions per channel per antenna averaged to two minute intervals.
-The solutions are still fairly low signal--to--noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation.
+The calibration algorithm -- based on forward modeling software by \citet{sul12} -- computes complex gain solutions per channel per antenna averaged to two minute intervals.
+The solutions are fairly low signal--to--noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation.
 First, we average the channel gains over all antennas to obtain a high signal--to--noise measurement of the bandpass.
 After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna.
-This flattens any residual variation in bandpass and removes small phase slopes due to variation in cable delay.
-Finally, we fit for an additional phase known to be caused by reflection in 150~m length cables.
+This flattens any residual variation in bandpass and removes small phase slopes due to variations in cable delay.
+Finally, we fit for an additional phase known to be caused by reflections in 150~m length cables.
 %possibly could add elaboration, if I knew exactly which version of the cal was done here.
@@
 
-We show results of delay spectrum from MWA data after deconvolution along delay axis in Figure~\ref{fig:fhd_data}.
+We show results of delay spectrum from MWA data after deconvolution along the delay axis in Figure~\ref{fig:fhd_data}.
 The principle feature is the {\it foreground wedge}, whose boundaries, marked by solid white lines, increase linearly with light travel time corresponding to baseline length.
@@
 
-Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the measurements perfectly.
-This is predominant at short antenna spacings where the number of delay modes within the horizon is lesser.
+Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the measurements perfectly, dominant at short antenna spacings where the number of delay modes within the horizon is lesser.
 This is because the maximum delay envelope (boundary of the wedge) consists of mixed emission from large portions of the sky into a narrow range of delays and the deconvolution algorithm does not have sufficient support in delay space to fit the delay spectrum.
@@
 We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known repeatability and stability level of the analog signal chain \citep{bow07b}.
+The power--patterns are shown in Figure~\ref{fig:power_pattern} for the pointings chosen in our study.
 
@@
 Contours of the MWA tile power--pattern shown in Figure~\ref{fig:power_pattern} are also overlaid.
-Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the westward sky and the MWA tile's power gain is significant ($\gtrsim 12$\%) in that direction.
-In the {\it zenith} pointing, the Galactic plane has set and the contour level of power--pattern in that direction is at least 16 times smaller.
+Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the westward sky, where the MWA tile power gain is significant ($\gtrsim 12$\%).
+In the {\it zenith} pointing, the Galactic plane has set and the contour level of the power--pattern in that direction is at least 16 times smaller.
 
@@
 
-We use a combination of NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz to provide our compact object catalog due to their complementary survey footprints covering the entire sky, and matched flux sensitivity and angular resolution.
+We use a combination of the NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and the Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz to provide our compact object catalog due to their complementary survey footprints covering the entire sky, and matched flux density sensitivity and angular resolution.
 The SUMSS catalog covers the sky with declination $\delta < -30$\arcdeg with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec.
@@
 From this subset, we choose compact objects with $S_{185}\geq 10$~Jy.
-The selection of such bright compact objects is not affected by minor differences in flux sensitivity of the two surveys.
+The selection of such bright compact objects is not affected by minor differences in flux density sensitivity of the two surveys.
 We verified that our selection criteria ensure a similar areal density of objects in the two surveys.
@@
 With the aforementioned all--sky foreground model, and instrumental and observational parameters, we simulate visibilities using equation~\ref{eqn:obsvis}.
-Figure~\ref{fig:sim_data} shows the amplitude of delay spectrum from {\it off--zenith} and {\it zenith} pointings.
+Figure~\ref{fig:sim_data} shows the amplitude of the delay spectrum from {\it off--zenith} and {\it zenith} pointings.
 Notice the qualitative agreement of amplitude and structure with those obtained from data shown in Figure~\ref{fig:fhd_data}.
@@
 Our foreground models are derived from other higher frequency catalogs and sky maps.
-The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observed frequency.
-Using simple error propagation, the fractional error in the delay spectrum caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the observed frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index.
+The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observing frequency.
+Using simple error propagation, the fractional error in the delay spectrum caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the MWA observing frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index.
 From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for compact objects from NVSS and SUMSS catalogs.
@@
 
-In addition to intrinsic model uncertainty, delay spectra from simulations and data each have fluctuations due to thermal noise in delay spectrum with rms $\sim$ 1.4$\times 10^7$~Jy~Hz.
-We estimate the ratio of delay spectra from data and simulations as, $\rho = |V^\textrm{D}_\eta(|\overline{\mathbf{x}}|,\eta)|\,/\,|V^\textrm{S}_\eta(|\overline{\mathbf{x}}|,\eta)|$, where superscripts D and S denote data and simulation respectively.
+In addition to intrinsic model uncertainty, delay spectra from simulations and data each have fluctuations due to thermal noise in the delay spectrum with rms $\sim$ 1.4$\times 10^7$~Jy~Hz.
+We estimate the ratio of delay spectra from data and simulations as $\rho = |V^\textrm{D}_\eta(|\overline{\mathbf{x}}|,\eta)|\,/\,|V^\textrm{S}_\eta(|\overline{\mathbf{x}}|,\eta)|$, where superscripts D and S denote data and simulation, respectively.
 % Using the aforementioned uncertainties added in quadrature, we estimate the approximate expected uncertainty in $\log_{10}\rho$ and denote it by $\Delta\,\log_{10}\rho$.
@@
 Figure~\ref{fig:data-sim-ratio} shows normalized histograms of the distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}.
-The median absolute deviation from these distributions correspond to $\sim$~90\% fractional difference between data and modeling on average in either case.
-This is in line with aforementioned uncertainty in foreground models and thermal noise fluctuations in measurements.
-Currently we are significantly limited by unavailability of all--sky foreground models at the observed frequency of 185~MHz.
+The median absolute deviation from these distributions corresponds to $\sim$~90\% fractional difference between data and modeling on average in either case.
+This is in line with the aforementioned uncertainty in foreground models and thermal noise fluctuations in measurements.
+Currently we are significantly limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.
 
@@
 \item {\it Tile Pointing and Power--Pattern}: Since the MWA tiles are steered electronically, the power--pattern of the tile changes in any observing mode that tracks the object.
-In our study, we take into account the effect of changing tile power--pattern on delay spectrum (Figure~\ref{fig:power_pattern}).
-The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose {\it off--zenith} and {\it zenith} pointings shown in Figure~\ref{fig:pointings}.
+In our study, we take into account the effect of the changing tile power--pattern on the delay spectrum (Figure~\ref{fig:power_pattern}).
+The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose the {\it off--zenith} and {\it zenith} pointings shown in Figure~\ref{fig:pointings}.
 We will show the secondary lobe structures in the tile power--pattern are tied to the foreground signatures in Fourier space.
@@
 Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the amplitude of delay spectrum, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$, simulated with only the diffuse foreground model, omitting compact sources.
-The top and bottom panels correspond to {\it off--zenith} and {\it zenith} pointings respectively.
+The top and bottom panels correspond to {\it off--zenith} and {\it zenith} pointings, respectively.
 
@@
 \caption{Simulated delay spectrum amplitudes $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
-White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
+White lines mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
 The axes and color scale are identical to those in Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}.
@@
 It appears as a bright branch in the delay spectrum near the negative delay horizon delay limit.
-This feature is strongest at short antenna spacings and fades with increase in antenna spacing.
-This Galactic center signature is absent in the {\it zenith} pointing (bottom panel) because the Galactic center has already set over the horizon.
+This feature is strongest at short antenna spacings and fades with increasing antenna spacing.
+The signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-delay-spectrum}) because the Galactic center is below the horizon.
 
@@
 The former is seen at positive delays because the primary lobe of the power--pattern is centered eastward of zenith, whereas in the latter it is centered at zenith.
-As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inversely proportional to length of the baseline projected in the direction of emission.
+As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inversely proportional to the length of the baseline projected in the direction of the emission.
 Thus, in the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at zero delay, fades away on antenna spacings $|\overline{\mathbf{x}}| \gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0\fdg 75.
@@
 
-Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the delay spectrum of compact foreground sky model (see Figure~\ref{fig:CSM})for {\it off--zenith} and {\it zenith} pointings in the top and bottom panels respectively.
+Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the delay spectrum of the compact foreground sky model (see Figure~\ref{fig:CSM}) for {\it off--zenith} and {\it zenith} pointings in the top and bottom panels, respectively.
 
@@
 \caption{Simulated delay spectrum amplitudes $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise.
-White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
+White lines mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
 The axes and logarithmic color scale are identical to those in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data} and \ref{fig:noiseless-dsm-delay-spectrum}.
@@
 
-In contrast to delay spectrum of diffuse emission, compact emission manifests as a center--heavy structure in the delay spectrum with either pointing.
+In contrast to the delay spectrum of diffuse emission, compact emission manifests as a center--heavy structure in the delay spectrum with either pointing.
 This contributes to the {\it pitchfork} feature which will be discussed in detail in \S\ref{sec:composite}.
@@
 
-\subsection{The "Pitchfork"}\label{sec:composite}
+\subsection{The ``Pitchfork''}\label{sec:composite}
 
@@
 Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional delay spectrum plots shown in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}.
-The diffuse, compact, and composite components are are shown as dotted, dashed, and solid lines respectively.
+The diffuse, compact, and composite components are are shown as dotted, dashed, and solid lines, respectively.
 The horizon delay limits are shown as a pair of vertical dot--dashed lines.
@@
 The zero--delay peak (corresponding to the main lobe in the power--pattern) with a value of $\sim 10^8$--$10^9$~Jy~Hz, independent of antenna spacing, is predominantly determined by compact foreground objects.
-The corresponding peak at zero--delay from diffuse foreground model is at least 10 times fainter and decreases drastically with increase in antenna spacing.
+The corresponding peak at zero--delay from diffuse foreground model is at least 10 times fainter and decreases drastically with an increase in antenna spacing.
 This is the response expected from different antenna spacings towards compact and diffuse foreground objects.
@@
 All these observations are consistent with those from Figures~\ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}.
-The observability of the {\it pitchfork} signature predicted in this paper depends on relative levels of uncertainty in the foreground model and fluctuations from thermal noise.
+The observability of the {\it pitchfork} signature predicted in this paper depends on the relative levels of uncertainty in the foreground model and fluctuations from thermal noise.
 In our simulations, since thermal noise in these very short duration snapshots is $\gtrsim 10^7$~Jy~Hz, and features near the horizon delay limits are $\sim 10^7$~Jy~Hz, the {\it pitchfork} feature is not expected to be detected in a noisy scenario although this feature is marginally visible in the {\it zenith} pointing of observed data (see Figure~\ref{fig:fhd_data}).
@@
 
-We also note that increasing the antenna spacing results in progressively improving the resolution along delay axis by increasing the number of delay bins inside the {\it foreground wedge}.
+We also note that increasing the antenna spacing results in progressively improving the resolution along the delay axis by increasing the number of delay bins inside the {\it foreground wedge}.
 This improves the localization of foreground objects whose signatures are imprinted in the delay spectrum.
-For instance, there is an increase in the number of secondary peaks in the delay spectrum between zero--delay and horizon delay limits as antenna spacing increases from $\sim$84~m to $\sim$170~m.
+For instance, there is an increase in the number of secondary peaks in the delay spectrum between zero--delay and horizon delay limits as the antenna spacing increases from $\sim$84~m to $\sim$170~m.
 In this case, these correspond to secondary lobes of the power--pattern that lie between the primary lobe and the horizon along the local meridian.
-At short antenna spacings, due to relatively lower resolution along delay axis inside the {\it foreground wedge} and a consequent loss of localization of foreground emission, these secondary peaks blend in with other major peaks and are not distinctly visible.
+At short antenna spacings, due to relatively lower resolution along the delay axis inside the {\it foreground wedge} and a consequent loss of localization of foreground emission, these secondary peaks blend in with other major peaks and are not distinctly visible.
 These findings are consistent with the discussion in \S\ref{sec:compact}.
@@
 Diffuse emission co--located at the primary lobe of the power--pattern is fainter by approximately an order of magnitude relative to compact emission from the same region for a $\sim 84$~m antenna spacing.
-But unlike the latter, diffuse emission decreases rapidly by over two orders of magnitude as antenna spacing is widened to $\sim 170$~m.
-However, diffuse emission is significantly high near horizon delay limits compared to that in the primary lobe of the power--pattern as antenna spacing is widened.
+But unlike the latter, diffuse emission decreases rapidly by over two orders of magnitude as the antenna spacing is widened to $\sim 170$~m.
+However, diffuse emission is significantly higher near horizon delay limits compared to that in the primary lobe of the power--pattern as the antenna spacing is increased.
 This leads to edge--heavy features in the delay spectrum.
-Complemented by the center-heavy compact foreground features, we predict the presence of a {\it three--pronged pitchfork}--shaped signature in the delay spectrum of foreground sky.
+Complemented by the center-heavy compact foreground features, we predict the presence of a {\it three--pronged pitchfork}--shaped signature in the delay spectrum of the foreground sky.
 We identify evidence for presence of this signature in actual MWA data, although deeper observations are needed to more clearly reveal the effect.
@@
 
-The delay spectrum, $V_\eta(\overline{\mathbf{x}},\eta)$, not only carries information on spatial scales of foreground emission but also offers a unique advantage of viewing the sky through a combination of antenna spacing vector and the delay axis.
+The delay spectrum, $V_\eta(\overline{\mathbf{x}},\eta)$, not only carries information on spatial scales of foreground emission but also offers the unique advantage of viewing the sky through a combination of antenna spacing vectors and the delay axis.
 With a foreground model known {\it a priori} in which structures and locations of very bright foreground objects such as the Galactic center or AGN are available, we can predict their response across antenna spacings as a function of observing parameters such as LST, power--pattern, etc.
@@
 Among antenna spacings that satisfy these criteria, we discard data from those whose lengths are shorter than $|\overline{\mathbf{x}}|_\textrm{max}$ ($x$--axis) and show foreground contamination estimated in the {\it EoR window} from all remaining antenna spacings.
-Foreground contamination estimated from only northward (dashed line) and eastward (dot--dashed line) antenna spacings represent the lower and upper limits respectively, and are shown for reference.
+Foreground contamination estimated from only northward (dashed line) and eastward (dot--dashed line) antenna spacings represent the lower and upper limits, respectively, and are shown for reference.
 It is seen that foreground contamination can be mitigated by a factor between $\sim 1.5$ ($|\theta_\textrm{b}|\le$~7\fdg 5) and $\sim 10$ ($|\theta_\textrm{b}|\le$~22\fdg 5).
@@
 This analysis extends previous work by simulating the entire sky rather than just the central field of view and by providing a comparison with early observations from the MWA.
-By making use of the delay spectrum technique to estimate power spectrum, we are able to observe the effects of foregrounds while avoiding entanglements with more complex power spectrum estimators.
+By making use of the delay spectrum technique to estimate the power spectrum, we are able to observe the effects of foregrounds while avoiding entanglements with more complex power spectrum estimators.
 
-Simulating in all important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright compact objects from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra is in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise fluctuations in measurements.
+Simulating in all important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright compact objects from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise fluctuations in measurements.
 
@@
 However, the most interesting result is its footprint on wide antenna spacings near the horizon delay limits, an edge--heavy {\it two--pronged fork}--shaped signature.
-This is due to apparent shortening of wide antenna spacings in the direction of foreground emission far off--axis thereby retaining their response to extended emission.
+This is due to apparent shortening of wide antenna spacings in the direction of foreground emission far off--axis, thereby retaining their response to extended emission.
 On the other hand, compact foreground objects predominantly map onto central regions of the {\it foreground wedge}.
-Features arising from compact objects co--located with primary and secondary lobes of antenna power--pattern have been identified.
+Features arising from compact objects co--located with primary and secondary lobes of the antenna power--pattern have been identified.
 In general, delay spectrum signatures of compact objects are center--heavy, in clear contrast to those from diffuse emission.
 A composite all--sky foreground model consisting of diffuse and compact foregrounds combines the characteristic individual shapes into a {\it pitchfork}--shaped structure in the {\it foreground wedge}.
-This will be distinctly visible when the thermal noise floor is sufficiently lowered, as longer observations are processed through the MWA analysis pipeline.
+This will be distinctly visible when the thermal noise floor is sufficiently lowered, as longer observations are processed through the MWA analysis pipelines.
 

### 2014-10-10 foreground_delay_spectrum_modeling.tex
From 5b726ca: ============
(*) Added space after degree symbol.
(*) Merged branch 'Steven' with 'master'
(*) Revised author list (removed Desh and Bryan) and rearranged affiliation ordering
From 5b726ca: ============

(*) Added space after degree symbol.

@@
 
-\def\RRI{\altaffilmark{5}}
-\def\RRItxt{\altaffiltext{5}{Raman Research Institute, Bangalore 560080, India}}
+\def\Curtin{\altaffilmark{5}}
+\def\Curtintxt{\altaffiltext{5}{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}}
 
-\def\Curtin{\altaffilmark{6}}
-\def\Curtintxt{\altaffiltext{6}{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}}
+% \def\USydney{\altaffilmark{7}}
+% \def\USydneytxt{\altaffiltext{7}{Sydney Institute for Astronomy, School of Physics, The University of Sydney, NSW 2006, Australia}}
 
-\def\USydney{\altaffilmark{7}}
-\def\USydneytxt{\altaffiltext{7}{Sydney Institute for Astronomy, School of Physics, The University of Sydney, NSW 2006, Australia}}
+\def\MIT{\altaffilmark{6}}
+\def\MITtxt{\altaffiltext{6}{Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}}
 
-\def\CAASTRO{\altaffilmark{8}}
-\def\CAASTROtxt{\altaffiltext{8}{ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}}
+\def\CfA{\altaffilmark{7}}
+\def\CfAtxt{\altaffiltext{7}{Harvard-Smithsonian Center for Astrophysics, Cambridge, MA 02138, USA}}
 
-\def\MIT{\altaffilmark{9}}
-\def\MITtxt{\altaffiltext{9}{Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}}
+\def\UW{\altaffilmark{8}}
+\def\UWtxt{\altaffiltext{8}{Department of Physics, University of Washington, Seattle, WA 98195, USA}}
 
-\def\CfA{\altaffilmark{10}}
-\def\CfAtxt{\altaffiltext{10}{Harvard-Smithsonian Center for Astrophysics, Cambridge, MA 02138, USA}}
+\def\Victoria{\altaffilmark{9}}
+\def\Victoriatxt{\altaffiltext{9}{School of Chemical \& Physical Sciences, Victoria University of Wellington, Wellington 6140, New Zealand}}
 
-\def\UW{\altaffilmark{11}}
-\def\UWtxt{\altaffiltext{11}{Department of Physics, University of Washington, Seattle, WA 98195, USA}}
+\def\UWisc{\altaffilmark{10}}
+\def\UWisctxt{\altaffiltext{10}{Department of Physics, University of Wisconsin--Milwaukee, Milwaukee, WI 53201, USA}}
 
-\def\Victoria{\altaffilmark{12}}
-\def\Victoriatxt{\altaffiltext{12}{School of Chemical \& Physical Sciences, Victoria University of Wellington, Wellington 6140, New Zealand}}
+\def\UMichigan{\altaffilmark{11}}
+\def\UMichigantxt{\altaffiltext{11}{Department of Atmospheric, Oceanic and Space Sciences, University of Michigan, Ann Arbor, MI 48109, USA}}
 
-\def\UWisc{\altaffilmark{13}}
-\def\UWisctxt{\altaffiltext{13}{Department of Physics, University of Wisconsin--Milwaukee, Milwaukee, WI 53201, USA}}
+\def\CASS{\altaffilmark{12}}
+\def\CASStxt{\altaffiltext{12}{CSIRO Astronomy and Space Science (CASS), PO Box 76, Epping, NSW 1710, Australia}}
 
-\def\UMichigan{\altaffilmark{14}}
-\def\UMichigantxt{\altaffiltext{14}{Department of Atmospheric, Oceanic and Space Sciences, University of Michigan, Ann Arbor, MI 48109, USA}}
+\def\CAASTRO{\altaffilmark{13}}
+\def\CAASTROtxt{\altaffiltext{13}{ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}}
 
-\def\CASS{\altaffilmark{15}}
-\def\CASStxt{\altaffiltext{15}{CSIRO Astronomy and Space Science (CASS), PO Box 76, Epping, NSW 1710, Australia}}
+\def\Tata{\altaffilmark{14}}
+\def\Tatatxt{\altaffiltext{14}{National Centre for Radio Astrophysics, Tata Institute for Fundamental Research, Pune 411007, India}}
 
-\def\Tata{\altaffilmark{16}}
-\def\Tatatxt{\altaffiltext{16}{National Centre for Radio Astrophysics, Tata Institute for Fundamental Research, Pune 411007, India}}
+\def\RRI{\altaffilmark{15}}
+\def\RRItxt{\altaffiltext{15}{Raman Research Institute, Bangalore 560080, India}}
 
-\def\NRAO{\altaffilmark{17}}
-\def\NRAOtxt{\altaffiltext{17}{National Radio Astronomy Observatory, Charlottesville and Greenbank, USA}}
+\def\NRAO{\altaffilmark{16}}
+\def\NRAOtxt{\altaffiltext{16}{National Radio Astronomy Observatory, Charlottesville and Greenbank, USA}}
 
-\def\UMelbourne{\altaffilmark{18}}
-\def\UMelbournetxt{\altaffiltext{18}{School of Physics, The University of Melbourne, Parkville, VIC 3010, Australia}}
+\def\UMelbourne{\altaffilmark{17}}
+\def\UMelbournetxt{\altaffiltext{17}{School of Physics, The University of Melbourne, Parkville, VIC 3010, Australia}}
 
@@
 B.~E.~Corey\Haystack, 
-A.~A.~Deshpande\RRI, 
+% A.~A.~Deshpande\RRI, 
 D.~Emrich\Curtin,
-B.~M.~Gaensler\USydney$^,$\CAASTRO, 
+% B.~M.~Gaensler\USydney$^,$\CAASTRO, 
 R.~Goeke\MIT,
@@
 \ASUtxt
-\UWtxt
-\RRItxt
 \SKASAtxt
@@
 \Curtintxt
-\USydneytxt
-\CAASTROtxt
+% \USydneytxt
 \MITtxt
 \CfAtxt
+\UWtxt
 \Victoriatxt
@@
 \CASStxt
+\CAASTROtxt
 \Tatatxt
+\RRItxt
 \NRAOtxt
@@
 Each tile is a phased array of 16 dipoles, each in the shape of a bow--tie.
-This yields a primary field of view $\gtrsim$~20\arcdeg wide and multiple secondary lobes.
+This yields a primary field of view $\gtrsim$~20\arcdeg~ wide and multiple secondary lobes.
 The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km \citep{bea12}.
@@
 The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of the dipoles in an MWA tile.
-The pointing system is capable of steering to points on a regular $\sim$7\arcdeg grid.
+The pointing system is capable of steering to points on a regular $\sim$7\arcdeg~ grid.
 During observations we allow the sky to drift across the nearest available pointing, shifting between grid points as necessary (once in $\sim 30$~minutes).
@@
 The axis on the left refers to longitudes, namely, Right Ascension (RA) and Hour Angle (HA).
-Negative values of RA, HA, and LST are to be interpreted as having been wrapped around by 360\arcdeg or 24 hours.
+Negative values of RA, HA, and LST are to be interpreted as having been wrapped around by 360\arcdeg~ or 24 hours.
 The axis on the right refers to the declination of the pointing direction.
@@
 2014, submitted).
-MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning; it has a flux density limit of 25~mJy and a declination range of -12\arcdeg to -40\arcdeg evenly covering the field of view of the observations reported here.
+MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning; it has a flux density limit of 25~mJy and a declination range of -12\arcdeg~ to -40\arcdeg~ evenly covering the field of view of the observations reported here.
 The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern.
@@
 
-The MWA has a very wide field of view ($\gtrsim 20$\arcdeg at 185~MHz) and has non--zero response all the way to the horizon.
+The MWA has a very wide field of view ($\gtrsim 20$\arcdeg~ at 185~MHz) and has non--zero response all the way to the horizon.
 As already noted, foreground objects fill the {\it foreground wedge} with an amplitude which depends on the tile power gain in that direction, with objects at the horizon being closest to the boundaries of the {\it wedge}.
@@
 We use a combination of the NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and the Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz to provide our compact object catalog due to their complementary survey footprints covering the entire sky, and matched flux density sensitivity and angular resolution.
-The SUMSS catalog covers the sky with declination $\delta < -30$\arcdeg with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec.
-The NVSS covers the sky with $\delta > -40$\arcdeg with a similar angular resolution and a limiting flux density of $\approx 2.5$~mJy for compact objects.
+The SUMSS catalog covers the sky with declination $\delta < -30$\arcdeg~ with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec.
+The NVSS covers the sky with $\delta > -40$\arcdeg~ with a similar angular resolution and a limiting flux density of $\approx 2.5$~mJy for compact objects.
 
@@
 Figure~\ref{fig:screening} shows how the typical foreground contamination in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
-We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars).
+We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars).
 Among antenna spacings that satisfy these criteria, we discard data from those whose lengths are shorter than $|\overline{\mathbf{x}}|_\textrm{max}$ ($x$--axis) and show foreground contamination estimated in the {\it EoR window} from all remaining antenna spacings.
@@
 \caption{Foreground contamination in {\it EoR window} for the {\it off--zenith} pointing as a function of orientation and length of discarded antenna spacings.
-Antenna spacings oriented eastward with varying degrees of directedness are considered: -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars).
+Antenna spacings oriented eastward with varying degrees of directedness are considered: -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars).
 Among antenna spacings that satisfy the above criteria, ones with lengths $\le |\overline{\mathbf{x}}|_\textrm{max}$ ($x$--axis) are discarded.

### 2014-10-17 foreground_delay_spectrum_modeling.tex
From 1d1c070: ============
(*) Made minor changes based on Abraham's comments.

@@
 In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain.
-Here, we will focus primarily on extending the {\it avoidance} strategy by identifying foreground components most likely to ``leak'' from foreground modes to EoR modes and demonstrating a scheme for down--weighting these components.
+Here, we will focus primarily on extending the {\it avoidance} strategy by identifying foreground components at greatest risk to ``leak'' from foreground modes to EoR modes and demonstrating a scheme for down--weighting these components.
 
@@
 Matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the signature of the power spectrum of each baseline.
-We report two important findings: the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a larger role than previously thought.
+We report two important findings: the foregrounds that most severely obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a larger role than previously thought.
 % We find that the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon.
@@
 Regions outside the {\it wedge} and the narrow white unshaded strips (caused by convolution with the instrumental transfer function) are expected to be relatively free of foreground contamination.
-Due to an undesirable grating pattern (dot--dashed lines) of the instrument used in this study, namely the MWA, we identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the {\it EoR window} in this paper.
+Due to an undesirable grating pattern (dot--dashed lines) of the instrument used in this study, namely the MWA, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the {\it EoR window} in this paper.
 \label{fig:fourier-space}}
@@
 
-The MWA was completed in 2012 and, after commissioning, began its EoR observing program in 2013.
+The MWA construction was completed in 2012 and, after commissioning, began its EoR observing program in 2013.
 In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz.
@@
 The coarse channel shape is obtained using a 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$.
-After correcting for the shape of these coarse channels, the fine channels (40~kHz width) at the edges of these sub--bands are flagged because they are known to be possibly contaminated by aliasing at a low level.
+After correcting for the shape of these coarse channels, the fine channels (40~kHz width) at the edges of these sub--bands are flagged because they are known to be contaminated by aliasing at a low level.
 Any spectral flags, or other known weights constitutes $W_f(f)$ in equation~\ref{eqn:obsvis}.
@@
 Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=95$~K per polarization for all frequency channels for all antenna pairs throughout the course of the observation.
+Estimation of $T_\textrm{sys}$ from MWA data is a subject of active investigation.
+In our work, we have chosen a $T_\textrm{sys}$ that matches the thermal noise observed in the data.
 % We take into account the increase in thermal uncertainty in frequency channels that results due to aforementioned bandpass correction.
@@
 
-The observations shown here were undertaken on Aug 23, 2013.
-This day has been chosen as a reference data set for comparison within the MWA EoR collaboration and was selected on the basis of a low number of known system errors from among a long stretch of stable observing.
+The observations shown here were undertaken on Aug~23, 2013.
+Data set from this day has been chosen as a reference for comparison within the MWA EoR collaboration because this is one of the days featuring a low number of known system errors from among a long stretch of stable observing.
 
@@
 We obtain the delay spectrum of these calibrated visibilities taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function.
-The spectrum is multiplied by the additional weight of the flagged channels which occur rarely for interference and regularly every 1.28~MHz where the edges of coarse channels are known to have a slight aliasing effect.
+The spectrum is multiplied by the additional weight of the flagged channels which occur rarely for interference and regularly every 1.28~MHz where the edges of coarse channels are known to have a small aliasing effect.
 In delay--space, these weights translate into a convolution by a grated point spread function (PSF) with gratings at multiples of $\eta=0.78\,\mu$s.
@@
 
-The MWA has a very wide field of view ($\gtrsim 20$\arcdeg~ at 185~MHz) and has non--zero response all the way to the horizon.
+The MWA has a very wide field of view ($\gtrsim 20$\arcdeg~ at 185~MHz) and has non--negligible response all the way to the horizon.
 As already noted, foreground objects fill the {\it foreground wedge} with an amplitude which depends on the tile power gain in that direction, with objects at the horizon being closest to the boundaries of the {\it wedge}.
@@
 For the diffuse component, we use an all--sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz.
-Since this map is predominantly based on the 408~MHz all--sky map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz map to the same resolution.
+At this frequency, since this map is predominantly based on the 408~MHz all--sky map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz map to the same resolution.
 However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals.
@@
 
-It may be noted that numerous features overlap at varying levels of significance depending on combinations of parameters.
+Note that numerous features may overlap at different degrees of significance depending on combinations of parameters.
 We assign the features to their predominant causes.
@@
 A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\overline{\mathbf{x}}|\le$~200~m).
-This is contrary to expectations because longer baselines are not expected to be sensitive to diffuse emission.
+This is contrary to expectations because these longer baselines are not expected to be sensitive to diffuse emission (on scales $\gtrsim$~0\fdg 85) of our input foreground model.
 In the wide--field observations of the MWA, long baselines appear shortened due to projection effects for off--axis emission, especially towards the horizon.
@@
 
-When not dominated by the bright emission from the Galactic center, which happens only for certain pointings at certain times, the delay spectrum of the combined foreground model is dominated, in somewhat equal measure, by diffuse and compact emission.
+When not dominated by the bright emission from the Galactic center, which happens only for certain pointings at certain times, the delay spectrum of the combined foreground model is composed of diffuse and compact emission, in roughly equal measure.
 This is illustrated by a more detailed examination of the {\it zenith} pointing in our study.
@@
 Figure~\ref{fig:before-after} shows the delay spectra obtained with all antenna spacings (top panel) and after applying our screening technique (bottom panel) on the {\it off--zenith} observation.
-Notice the remarkable reduction --- by about an order of magnitude --- in foreground contamination in this region after application of our screening technique.
+Notice the remarkable reduction --- by about an order of magnitude --- in foreground contamination in the {\it EoR window} after application of our screening technique.
+The color scale range has been adjusted relative to the previous figures to show this significant reduction in foreground contamination.
 
@@
 
-This screening technique can be customized when it is necessary to optimize between foreground mitigation and loss of sensitivity from discarding data.
+This screening technique can be generalized to optimize between foreground mitigation and loss of sensitivity from discarding data.
 Figure~\ref{fig:screening} shows how the typical foreground contamination in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
@@
 Among antenna spacings that satisfy these criteria, we discard data from those whose lengths are shorter than $|\overline{\mathbf{x}}|_\textrm{max}$ ($x$--axis) and show foreground contamination estimated in the {\it EoR window} from all remaining antenna spacings.
+In other words, these plots demonstrate the progress in mitigation as orientation and maximum length of discarded antenna spacings are varied.
 Foreground contamination estimated from only northward (dashed line) and eastward (dot--dashed line) antenna spacings represent the lower and upper limits, respectively, and are shown for reference.
@@
 It is notable that short antenna spacings play maximum role in determining foreground contamination.
-For instance, among antenna spacings with orientations $|\theta_\textrm{b}|\le$~22\fdg 5, discarding those with lengths $|\overline{\mathbf{x}}|>40$~m has no effect on further foreground contamination and would only lead to loss of sensitivity.
+For instance, among antenna spacings with orientations $|\theta_\textrm{b}|\le$~22\fdg 5, discarding those with lengths $|\overline{\mathbf{x}}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity.
 On the other hand, a bright compact object at the same location will give rise to foreground contamination even on longer antenna spacings.
@@
 Short baselines have maximum effect on foreground contamination.
-For instance, discarding antenna spacings oriented $|\theta_\textrm{b}|\le$~22\fdg 5 with lengths $|\overline{\mathbf{x}}|>40$~m has no effect in further reducing foreground contamination.
+For instance, discarding antenna spacings oriented $|\theta_\textrm{b}|\le$~22\fdg 5 with lengths $|\overline{\mathbf{x}}|\gtrsim 30$~m has no effect in further reducing foreground contamination.
 \label{fig:screening}}

### 2014-10-17 foreground_delay_spectrum_modeling.tex
From 89b3f2f: ============
(*) Revised section on delay spectrum math and changed notations.
(*) Added a few references in eor.bib

@@
 
-Considerable effort has been made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,liu11,dat10,bow09,liu09}.
+Considerable effort has been and continues to be made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,liu14a,liu14b}.
 A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide-field response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum.
@@
 
-Interferometer array data known as {\it visibilities}, $V_f(\overline{\mathbf{u}},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\overline{\mathbf{x}}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis.
-Here, $\overline{\mathbf{u}}\equiv \overline{\mathbf{x}}f/c \equiv \overline{\mathbf{x}}/\lambda$, $c$ is the speed of light, and $\lambda$ is the wavelength of radio waves.
-$\overline{\mathbf{u}}$ is related to the transverse spatial frequency modes ($\overline{\mathbf{k}}_\perp$) of the brightness distribution of the sky as $\overline{\mathbf{k}}_\perp \equiv 2\pi\overline{\mathbf{u}}/D(z)$, where $D(z)$ is the transverse comoving distance at redshift $z$.
-If $I(\overline{\mathbf{l}},f)$ is the emission at different frequencies on the sky as a function of direction unit vector ($\overline{\mathbf{l}}$) specified as direction cosines, $V_f(\overline{\mathbf{u}},f)$ represents the Fourier decomposition of $I(\overline{\mathbf{l}},f)$ attenuated by the antenna's angular power--pattern, $A(\overline{\mathbf{l}},f)$, into transverse spatial frequency modes.
-Taking into account the instrumental bandpass weights $W_f(f)$, this relation can be written as:
+Interferometer array data known as {\it visibilities}, $V_{bf}(\vec{b},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\vec{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis.
+If $I(\hat{s},f)$ and $A(\hat{s},f)$ are the sky brightness and antenna's directional power--pattern, respectively, at different frequencies on the sky as a function of direction on the sky denoted by the unit vector ($\hat{s}$), and $W_f(f)$ denotes instrumental bandpass weights, then $V_{bf}(\vec{b},f)$ can be written as (with a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}):
 \begin{align}\label{eqn:obsvis}
-  V_f(\overline{\mathbf{u}},f) &= \iint\limits_\textrm{sky} A(\overline{\mathbf{l}},f)\,I(\overline{\mathbf{l}},f)\,W_f(f)\,e^{-i2\pi \overline{\mathbf{u}}\cdot\overline{\mathbf{l}}}\,\dif^2 \overline{\mathbf{l}}.
+  V_{bf}(\vec{b},f) &= \iint\limits_\textrm{sky} A(\hat{s},f)\,I(\hat{s},f)\,W_f(f)\,e^{-i2\pi f\frac{\vec{b}\cdot\hat{s}}{c}}\,\dif\Omega,
 \end{align}
+where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element along $\hat{s}$.
 
-We define the {\it delay spectrum}, $V_\eta(\overline{\mathbf{u}},\eta)$, to be the inverse Fourier transform of $V_f(\overline{\mathbf{u}},f)$ along the frequency coordinate:
+We define the {\it delay spectrum}, $V_{b\tau}(\vec{b},\tau)$, to be the inverse Fourier transform of $V_{bf}(\vec{b},f)$ along the frequency coordinate:
 \begin{align}\label{eqn:delay-transform}
-  V_\eta(\overline{\mathbf{u}},\eta) &= \int V_f(\overline{\mathbf{u}},f)\,W'_f(f)\,e^{i2\pi\eta f}\,\dif f,
+  V_{b\tau}(\vec{b},\tau) &= \int V_{bf}(\vec{b},f)\,W'_f(f)\,e^{i2\pi f\tau}\,\dif f,
 \end{align}
-where, $W'_f(f)$ is a frequency window weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12} and $\eta$ represents the signal delay between antenna pairs:
+where, $W'_f(f)$ is a spectral weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}, and $\tau$ represents the signal delay between antenna pairs:
 \begin{equation}\label{eqn:delay}
-  \eta = \frac{\overline{\mathbf{x}}\cdot\overline{\mathbf{l}}}{c} = \frac{\overline{\mathbf{u}}\cdot\overline{\mathbf{l}}}{f}.
+  \tau = \frac{\vec{b}\cdot\hat{s}}{c}.
 \end{equation}
-Note that as the visibility is a complex quantity, its Fourier transform is Nyquist sampled over positive and negative delays.
+
+Equation~\ref{eqn:obsvis} can be modified to express the Fourier decomposition of transverse sky brightness distribution as:
+\begin{align}\label{eqn:vis}
+  V_{uf}(\vec{u},f) &= \iint\limits_\textrm{sky} A(\hat{s},f)\,I(\hat{s},f)\,W_f(f)\,e^{-i2\pi\vec{u}\cdot\hat{s}}\,\dif\Omega,
+\end{align}
+where, $\vec{u}$ denotes the transverse spatial frequency vector.
+$\vec{u}$ is directly related to $\vec{b}$ as $\vec{u}\equiv \vec{b}/\lambda$, and to the transverse spatial wavenumber mode, $\vec{k}_\perp$, as $\vec{k}_\perp\equiv 2\pi\vec{u}/D(z)$, where, $D(z)$ is the transverse comoving distance at redshift $z$.
+Hence, $V_{uf}(\vec{u},f)\equiv V_{bf}(\lambda\vec{u},f)$.
+It is important to note that $\vec{b}$ is independent of frequency while $\vec{u}$ is not.
+
+Since we are concerned with a redshifted H{\sc i} spectral line from cosmological distances, $f$ is a measure of cosmological distance along the line of sight.
+$\eta$, which is the Fourier transform dual of $f$, is used to denote the spatial frequency along the line of sight and has units of time.
+It is directly related to the line--of--sight wavenumber, $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
+Thus,
+\begin{align}\label{eqn:los-transform}
+  V_{u\eta}(\vec{u},\eta) &= \int V_{uf}(\vec{u},f)\,W'_f(f)\,e^{i2\pi f\eta}\,\dif f
+\end{align}
+represents the true spatial Fourier representation of the three--dimensional sky brightness distribution.
+This representation has been discussed in detail in \citet{mor04}.
+The spatial power spectrum of EoR H{\sc i} distribution, $P(\vec{k}_\perp,k_\parallel)$, and $V_{u\eta}(\vec{u},\eta)$ are related by $P(\vec{k}_\perp,k_\parallel)\propto |V_{u\eta}(\vec{u},\eta)|^2$.
+
+While $V_{b\tau}(\vec{b},\tau)$ and $V_{u\eta}(\vec{u},\eta)$ in equations~\ref{eqn:obsvis} and \ref{eqn:vis} resemble each other very closely, they are not quite identical.
+The difference basically arises from the fact noted above that $\vec{b}$ is independent of frequency while $\vec{u}$ is not.
+\citet{par12} and \citet{liu14a} have discussed the mathematical correspondence between the two quantities.
+
+Radio emission from foreground objects, due to their spatial and spectral (chromatic) structure, contaminate the received signal \citep{bow09,mor12,tro12,thy13,liu14a,liu14b}.
+Since this contamination is expected to be several orders of magnitude stronger than the underlying EoR H{\sc i} signal, it is critical to characterize foregrounds precisely in order to reduce their impact on EoR H{\sc i} power spectrum detection sensitivity.
+Foregrounds can be described in both the delay spectrum and Fourier frameworks.
+For our study, which primarily concerns with foreground characterization, we find the former to be simple and yet extremely useful, while maintaining a close correspondence with the latter despite their subtle differences.
+
+In summary, the delay spectrum, $V_{b\tau}(\vec{b},\tau)$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}.
+$V_{b\tau}(\vec{b},\tau)$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument.
+At the same time, it is closely related to the sought power $|V_{u\eta}(\vec{u},\eta)|^2$ in Fourier modes containing critical information about spatial scales.
+For purposes of illustration later in this paper, we often use the quantity, $|V_{b\tau}(|\vec{b}|,\tau)|$.
+
+Note that since {\it visibility} is a complex quantity, its Fourier transform is Nyquist sampled over positive and negative delays.
 Negative and positive delay\footnote{The terms delay and lag are used interchangeably in this paper.} correspond to the two hemispheres of the sky as transected by the antenna spacing vector.
-In this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the telescope site.
+In this paper, $\vec{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the telescope site.
 Hence, a perfectly eastward oriented antenna spacing will observe an object in the eastern sky at positive delay, and conversely for an object in the western sky.
@@
 
-Since we are studying a redshifted H{\sc i} spectral line, $f$ is a measure of cosmological distance along the line of sight.
-Consequently, $\eta$ is a measure of line--of--sight spatial frequency mode given by $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
-We use $H_0=100\,h$~km~s$^{-1}$~Mpc$^{-1}$ and assume a flat $\Lambda$CDM universe with $\Omega_\textrm{M}=0.27$ and $\Omega_\Lambda=0.73$.
-
-Thus, $|V_\eta(\overline{\mathbf{u}},\eta)|^2$ is directly related to the spatial power spectrum of EoR H{\sc i} distribution, $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ \citep{mor04}.
-However, foreground objects also emit at radio frequencies.
-Due to their spatial and spectral (chromatic) structure, the measured $V_\eta(\overline{\mathbf{u}},\eta)$ is contaminated by foreground radio objects \citep{thy13,tro12,mor12,bow09}.
-These contaminations have to be characterized precisely in order to reduce their impact on EoR H{\sc i} power spectrum detection sensitivity.
-
-In summary, $V_\eta(\overline{\mathbf{u}},\eta)$ is closely related to visibilities, $V_f(\overline{\mathbf{u}},f)$, the basic data blocks measured by each antenna pair, as seen from equation~\ref{eqn:delay-transform}.
-$V_\eta(\overline{\mathbf{u}},\eta)$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument.
-At the same time, it is directly related to the sought power $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ in $k$--modes containing critical information about spatial scales.
-For our analysis, we use an equivalent quantity, $V_\eta(\overline{\mathbf{x}},\eta) \equiv V_\eta(\lambda\overline{\mathbf{u}},\eta)$.
-For purposes of illustration, we often use a simplified quantity, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$.
-
 Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of H{\sc i} from the EoR are presented in this paper.
-$|\overline{\mathbf{x}}|$ and $k_\perp$, denoting spatial scales in the transverse direction, form the $x$--axis.
-$\eta$ and $k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis.
+$|\vec{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction, form the $x$--axis.
+$\tau\approx\eta\propto k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis.
 Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge}, whose boundaries are determined by antenna spacing and the light travel time across it.
 These boundaries are called horizon delay limits and are shown by solid lines.
-The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\eta$--axis.
+The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\tau$--axis.
 This effect has been quantified in \citet{thy13}.
@@
 
-% Hence, we have identified a region of high significance (shown in medium shade) that forms a subset of the {\it EoR window} bounded by $|\eta|\le 0.78\,\mu$s where contamination from foregrounds is expected to be less and yet the EoR H{\sc i} power spectrum signal is reasonably strong.
+% Hence, we have identified a region of high significance (shown in medium shade) that forms a subset of the {\it EoR window} bounded by $|\tau|\le 0.78\,\mu$s where contamination from foregrounds is expected to be less and yet the EoR H{\sc i} power spectrum signal is reasonably strong.
 Hereafter, we refer to this region as the {\it EoR window} in the rest of this paper.
 
-In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\overline{\mathbf{k}}|$.
+In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\vec{k}|$.
 Hence, the brightest EoR signal will be observed on the shortest baselines and smallest delays.
-Thus the region of interest for EoR studies is slightly beyond the horizon delay limits (dashed lines) and inside the first grating on either side ($|\eta| \lesssim 0.78\,\mu$s, dot--dashed lines) caused by the instrument's spectral transfer function.
+Thus the region of interest for EoR studies is slightly beyond the horizon delay limits (dashed lines) and inside the first grating on either side ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) caused by the instrument's spectral transfer function.
 We call this region with medium shade as the {\it EoR window} in this paper.
@@
 \caption{Illustration of Fourier space in which delay (and power) spectra of H{\sc i} signals from the EoR are presented.
-The $x$--axis is denoted by $|\overline{\mathbf{x}}|$ (antenna spacing) or $k_\perp$ (transverse wavenumber).
-The $y$--axis denoted by $\eta$ (delay) or $k_\parallel$ (line--of--sight wavenumber).
+The $x$--axis is denoted by $|\vec{b}|$ (antenna spacing) or $k_\perp$ (transverse wavenumber).
+The $y$--axis denoted by $\tau$ (delay) or $k_\parallel$ (line--of--sight wavenumber).
 The dark shaded region is referred to as the {\it foreground wedge} where smooth spectrum foregrounds reside.
@@
 The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km \citep{bea12}.
-Here we focus on antenna spacings $|\overline{\mathbf{x}}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz.
+Here we focus on antenna spacings $|\vec{b}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz.
 The center frequency of 185~MHz corresponds to a redshift, $z\simeq 6.68$.
@@
 The spectrum is multiplied by the additional weight of the flagged channels which occur rarely for interference and regularly every 1.28~MHz where the edges of coarse channels are known to have a small aliasing effect.
-In delay--space, these weights translate into a convolution by a grated point spread function (PSF) with gratings at multiples of $\eta=0.78\,\mu$s.
+In delay--space, these weights translate into a convolution by a grated point spread function (PSF) with gratings at multiples of $\tau=0.78\,\mu$s.
 We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis by \citet{par09,par12}.
@@
 
-In subsequent sections, we plot the amplitude of the delay spectrum, $|V(|\overline{\mathbf{x}}|,\eta)|$, sorted by baseline length, $|\overline{\mathbf{x}}|$.
+In subsequent sections, we plot the amplitude of the delay spectrum, $|V_{b\tau}(|\vec{b}|,\tau)|$, sorted by baseline length, $|\vec{b}|$.
 Note that visibilities from different baselines have not been averaged together in this analysis.
@@
 To get the delay spectrum, each baseline is Fourier transformed into delay space ($y$--axis) and the amplitude plotted in order of baseline length ($x$--axis).
-For reference, the $k$--space equivalent of $\eta$ and $|\overline{\mathbf{x}}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively.
+For reference, the $k$--space equivalent of $\tau$ and $|\vec{b}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively.
 The logarithmic color scale (shown at the right) is common to both panels.
@@
 
-The instrument model consists primarily of the tile beam power--pattern, $A(\overline{\mathbf{l}},f)$ (see equation~\ref{eqn:obsvis}).
+The instrument model consists primarily of the tile beam power--pattern, $A(\hat{s},f)$ (see equation~\ref{eqn:obsvis}).
 It is modeled as a mutually--coupled 4--by--4 dipole array with the overall power--pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14}.
@@
 In addition to intrinsic model uncertainty, delay spectra from simulations and data each have fluctuations due to thermal noise in the delay spectrum with rms $\sim$ 1.4$\times 10^7$~Jy~Hz.
-We estimate the ratio of delay spectra from data and simulations as $\rho = |V^\textrm{D}_\eta(|\overline{\mathbf{x}}|,\eta)|\,/\,|V^\textrm{S}_\eta(|\overline{\mathbf{x}}|,\eta)|$, where superscripts D and S denote data and simulation, respectively.
+We estimate the ratio of delay spectra from data and simulations as $\rho = |V^\textrm{D}_\tau(|\vec{b}|,\tau)|\,/\,|V^\textrm{S}_\tau(|\vec{b}|,\tau)|$, where superscripts D and S denote data and simulation, respectively.
 % Using the aforementioned uncertainties added in quadrature, we estimate the approximate expected uncertainty in $\log_{10}\rho$ and denote it by $\Delta\,\log_{10}\rho$.
@@
 
-Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the amplitude of delay spectrum, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$, simulated with only the diffuse foreground model, omitting compact sources.
+Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the amplitude of delay spectrum, $|V_{b\tau}(|\vec{b}|,\tau)|$, simulated with only the diffuse foreground model, omitting compact sources.
 The top and bottom panels correspond to {\it off--zenith} and {\it zenith} pointings, respectively.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-\caption{Simulated delay spectrum amplitudes $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
+\caption{Simulated delay spectrum amplitudes $|V_{b\tau}(|\vec{b}|,\tau)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
 White lines mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
@@
 As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inversely proportional to the length of the baseline projected in the direction of the emission.
-Thus, in the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at zero delay, fades away on antenna spacings $|\overline{\mathbf{x}}| \gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0\fdg 75.
+Thus, in the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at zero delay, fades away on antenna spacings $|\vec{b}| \gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0\fdg 75.
 In the off-zenith pointing, we see power out to much longer baselines which is described below.
@@
 
-A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\overline{\mathbf{x}}|\le$~200~m).
+A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\vec{b}|\le$~200~m).
 This is contrary to expectations because these longer baselines are not expected to be sensitive to diffuse emission (on scales $\gtrsim$~0\fdg 85) of our input foreground model.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-\caption{Simulated delay spectrum amplitudes $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise.
+\caption{Simulated delay spectrum amplitudes $|V_{b\tau}(|\vec{b}|,\tau)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise.
 White lines mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-\caption{Simulated delay spectrum amplitudes $|V_\eta(\overline{\mathbf{x}},\eta)|$ for three chosen antenna spacings of length: $\sim$84~m ({\it top}), $\sim$105~m ({\it middle}), and $\sim$170~m ({\it bottom}).
+\caption{Simulated delay spectrum amplitudes $|V_{b\tau}(\vec{b},\tau)|$ for three chosen antenna spacings of length: $\sim$84~m ({\it top}), $\sim$105~m ({\it middle}), and $\sim$170~m ({\it bottom}).
 The antenna spacing is specified in each panel along East, North and Upward directions.
@@
 The bright Galactic center in the western sky is recorded at zero delays in northward antenna spacings and close to negative horizon delay limit on eastward antenna spacings.
-(c): Simulated delay spectra $|V_\eta(\overline{\mathbf{x}},\eta)|$ on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}).
+(c): Simulated delay spectra $|V_{b\tau}(\vec{b},\tau)|$ on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}).
 The bright Galactic center is prominently visible close to negative horizon delay limit on eastward antenna spacings.
@@
 
-The delay spectrum, $V_\eta(\overline{\mathbf{x}},\eta)$, not only carries information on spatial scales of foreground emission but also offers the unique advantage of viewing the sky through a combination of antenna spacing vectors and the delay axis.
+The delay spectrum, $V_{b\tau}(\vec{b},\tau)$, not only carries information on spatial scales of foreground emission but also offers the unique advantage of viewing the sky through a combination of antenna spacing vectors and the delay axis.
 With a foreground model known {\it a priori} in which structures and locations of very bright foreground objects such as the Galactic center or AGN are available, we can predict their response across antenna spacings as a function of observing parameters such as LST, power--pattern, etc.
@@
 
-In our example, we discard eastward antenna spacings (-22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5) of lengths $|\overline{\mathbf{x}}| < 40$~m.
+In our example, we discard eastward antenna spacings (-22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5) of lengths $|\vec{b}| < 40$~m.
 Figure~\ref{fig:before-after} shows the delay spectra obtained with all antenna spacings (top panel) and after applying our screening technique (bottom panel) on the {\it off--zenith} observation.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_before_after_nside_64_Tsys_95.0K.eps}
-\caption{Simulated delay spectra amplitudes $|V_\eta(\overline{\mathbf{x}},\eta)|$ for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
-Discarded antenna spacings (black vertical stripes) have lengths $|\overline{\mathbf{x}}| < 40$~m and orientations -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5.
+\caption{Simulated delay spectra amplitudes $|V_{b\tau}(\vec{b},\tau)|$ for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
+Discarded antenna spacings (black vertical stripes) have lengths $|\vec{b}| < 40$~m and orientations -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5.
 The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} prominent when all antenna spacings are used is lowered by an order of magnitude when short eastward antenna spacings are discarded.
@@
 We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars).
-Among antenna spacings that satisfy these criteria, we discard data from those whose lengths are shorter than $|\overline{\mathbf{x}}|_\textrm{max}$ ($x$--axis) and show foreground contamination estimated in the {\it EoR window} from all remaining antenna spacings.
+Among antenna spacings that satisfy these criteria, we discard data from those whose lengths are shorter than $|\vec{b}|_\textrm{max}$ ($x$--axis) and show foreground contamination estimated in the {\it EoR window} from all remaining antenna spacings.
 In other words, these plots demonstrate the progress in mitigation as orientation and maximum length of discarded antenna spacings are varied.
@@
 It is notable that short antenna spacings play maximum role in determining foreground contamination.
-For instance, among antenna spacings with orientations $|\theta_\textrm{b}|\le$~22\fdg 5, discarding those with lengths $|\overline{\mathbf{x}}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity.
+For instance, among antenna spacings with orientations $|\theta_\textrm{b}|\le$~22\fdg 5, discarding those with lengths $|\vec{b}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity.
 On the other hand, a bright compact object at the same location will give rise to foreground contamination even on longer antenna spacings.
@@
 Antenna spacings oriented eastward with varying degrees of directedness are considered: -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars).
-Among antenna spacings that satisfy the above criteria, ones with lengths $\le |\overline{\mathbf{x}}|_\textrm{max}$ ($x$--axis) are discarded.
+Among antenna spacings that satisfy the above criteria, ones with lengths $\le |\vec{b}|_\textrm{max}$ ($x$--axis) are discarded.
 Lower (dashed line) and upper (dot--dashed) limits on foreground contamination are derived from northward (67\fdg 5~$\le\theta_\textrm{b}<$~112\fdg 5) and eastward ($|\theta_\textrm{b}|\le$~22\fdg 5) antenna spacings respectively.
@@
 Short baselines have maximum effect on foreground contamination.
-For instance, discarding antenna spacings oriented $|\theta_\textrm{b}|\le$~22\fdg 5 with lengths $|\overline{\mathbf{x}}|\gtrsim 30$~m has no effect in further reducing foreground contamination.
+For instance, discarding antenna spacings oriented $|\theta_\textrm{b}|\le$~22\fdg 5 with lengths $|\vec{b}|\gtrsim 30$~m has no effect in further reducing foreground contamination.
 \label{fig:screening}}

### 2014-10-17 foreground_delay_spectrum_modeling.tex
From 164a6dd: ============
(*) Made changes to figure axes and few figures.
(*) Added a bit more description about how foreground contamination used in
figure 14 is defined

@@
 At the same time, it is closely related to the sought power $|V_{u\eta}(\vec{u},\eta)|^2$ in Fourier modes containing critical information about spatial scales.
-For purposes of illustration later in this paper, we often use the quantity, $|V_{b\tau}(|\vec{b}|,\tau)|$.
+For purposes of illustration later in this paper, we often use the quantity, $|V_{b\tau}(\vec{b},\tau)|$.
 
@@
 
-In subsequent sections, we plot the amplitude of the delay spectrum, $|V_{b\tau}(|\vec{b}|,\tau)|$, sorted by baseline length, $|\vec{b}|$.
+In subsequent sections, we plot the amplitude of the delay spectrum, $|V_{b\tau}(\vec{b},\tau)|$, sorted by baseline length, $|\vec{b}|$.
 Note that visibilities from different baselines have not been averaged together in this analysis.
@@
 In addition to intrinsic model uncertainty, delay spectra from simulations and data each have fluctuations due to thermal noise in the delay spectrum with rms $\sim$ 1.4$\times 10^7$~Jy~Hz.
-We estimate the ratio of delay spectra from data and simulations as $\rho = |V^\textrm{D}_\tau(|\vec{b}|,\tau)|\,/\,|V^\textrm{S}_\tau(|\vec{b}|,\tau)|$, where superscripts D and S denote data and simulation, respectively.
+We estimate the ratio of delay spectra from data and simulations as $\rho = |V^\textrm{D}_{b\tau}(\vec{b},\tau)|\,/\,|V^\textrm{S}_{b\tau}(\vec{b},\tau)|$, where superscripts D and S denote data and simulation, respectively.
 % Using the aforementioned uncertainties added in quadrature, we estimate the approximate expected uncertainty in $\log_{10}\rho$ and denote it by $\Delta\,\log_{10}\rho$.
@@
 
-Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the amplitude of delay spectrum, $|V_{b\tau}(|\vec{b}|,\tau)|$, simulated with only the diffuse foreground model, omitting compact sources.
+Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the amplitude of delay spectrum, $|V_{b\tau}(\vec{b},\tau)|$, simulated with only the diffuse foreground model, omitting compact sources.
 The top and bottom panels correspond to {\it off--zenith} and {\it zenith} pointings, respectively.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-\caption{Simulated delay spectrum amplitudes $|V_{b\tau}(|\vec{b}|,\tau)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
+\caption{Simulated delay spectrum amplitudes $|V_{b\tau}(\vec{b},\tau)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
 White lines mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-\caption{Simulated delay spectrum amplitudes $|V_{b\tau}(|\vec{b}|,\tau)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise.
+\caption{Simulated delay spectrum amplitudes $|V_{b\tau}(\vec{b},\tau)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise.
 White lines mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
@@
 This screening technique can be generalized to optimize between foreground mitigation and loss of sensitivity from discarding data.
-Figure~\ref{fig:screening} shows how the typical foreground contamination in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
+Figure~\ref{fig:screening} shows how the typical foreground contamination\footnote{We use a statistical measure of foreground contamination using the standard deviation of noiseless $V_{b\tau}(\vec{b},\tau)$ from foregrounds in the {\it EoR window}.} in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
 We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars).
@@
 \includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_nside_64_Tsys_95.0K.eps}
-\caption{Foreground contamination in {\it EoR window} for the {\it off--zenith} pointing as a function of orientation and length of discarded antenna spacings.
+\caption{Foreground contamination, estimated using standard deviation in the {\it EoR window}, for the {\it off--zenith} pointing as a function of orientation and length of discarded antenna spacings.
 Antenna spacings oriented eastward with varying degrees of directedness are considered: -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars).

### 2014-10-18 foreground_delay_spectrum_modeling.tex
From 071aa02: ============
(*) Changed color scale of figure 6 showing compact foreground model
(*) Made changes based on most of Jonnie's comments
(*) Major changes include justification of compact source flux density threshold
(*) Modified a couple of figures as per Jonnie's suggestion

@@
 These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space.
-Based on these results, we propose that selective down--weighting of baselines based on length, direction, and time will mitigate a large fraction of foreground contamination in redshifted 21~cm power spectrum analysis.
+Based on these results, we propose that selective down--weighting of baselines based on length, direction, and time will mitigate foreground contamination substantially (up to a factor $\sim 10$ with negligible loss of data) in redshifted 21~cm power spectrum analysis.
 
@@
 
-A key challenge in the statistical detection of the redshifted H{\sc i} 21~cm signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
+A key challenge in the statistical detection of the redshifted H{\sc i} 21~cm signal, via the spatial power spectrum of temperature fluctuations, arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
 \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}.
@@
 Considerable effort has been and continues to be made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,liu14a,liu14b}.
-A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide-field response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum.
+A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide--field (and chromatic) response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum.
 The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}).
@@
 
-Foregrounds contaminate the EoR H{\sc i} power spectrum by either having intrinsic spectral variability, high instrumental chromaticity, or through approximations in power spectrum analyses.
+Foregrounds that have intrinsic spectral variations, instruments with high chromaticity, or approximations in power spectrum analyses can contaminate the true EoR H{\sc i} power spectrum.
 Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity.
@@
 \end{align}
-where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element along $\hat{s}$.
+where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element along $\hat{s}$. We wish to emphasize that this equation is valid in general and does not involve any approximation.
 
@@
 $\vec{u}$ is directly related to $\vec{b}$ as $\vec{u}\equiv \vec{b}/\lambda$, and to the transverse spatial wavenumber mode, $\vec{k}_\perp$, as $\vec{k}_\perp\equiv 2\pi\vec{u}/D(z)$, where, $D(z)$ is the transverse comoving distance at redshift $z$.
-Hence, $V_{uf}(\vec{u},f)\equiv V_{bf}(\lambda\vec{u},f)$.
+Hence, $V_{uf}(\vec{u},f)\equiv V_{uf}(\vec{b}/\lambda,f)\equiv V_{bf}(\lambda\vec{u},f)$.
 It is important to note that $\vec{b}$ is independent of frequency while $\vec{u}$ is not.
@@
 The difference basically arises from the fact noted above that $\vec{b}$ is independent of frequency while $\vec{u}$ is not.
-\citet{par12} and \citet{liu14a} have discussed the mathematical correspondence between the two quantities.
+\citet{par12} and \citet{liu14a} have discussed the mathematical correspondence between the two.
 
@@
 In this paper, $\vec{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the telescope site.
-Hence, a perfectly eastward oriented antenna spacing will observe an object in the eastern sky at positive delay, and conversely for an object in the western sky.
+Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at positive and negative delays, respectively.
 Similarly, an object in the northern sky will appear at a positive delay at an antenna spacing oriented northward.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/fourier_space_185.0_MHz_30.7_MHz.eps}
-\caption{Illustration of Fourier space in which delay (and power) spectra of H{\sc i} signals from the EoR are presented.
+\caption{Fourier space in which delay (and power) spectra of EoR H{\sc i} signals are presented.
 The $x$--axis is denoted by $|\vec{b}|$ (antenna spacing) or $k_\perp$ (transverse wavenumber).
 The $y$--axis denoted by $\tau$ (delay) or $k_\parallel$ (line--of--sight wavenumber).
+Here, $k_\perp$ and $k_\parallel$ are obtained for a frequency of 185~MHz.
 The dark shaded region is referred to as the {\it foreground wedge} where smooth spectrum foregrounds reside.
 Its boundaries (solid lines), given by light travel time for corresponding antenna spacings, are referred to as horizon delay limits.
-Regions outside the {\it wedge} and the narrow white unshaded strips (caused by convolution with the instrumental transfer function) are expected to be relatively free of foreground contamination.
+Regions excluding the {\it wedge} and narrow extensions of the {\it wedge} (white unshaded strips) caused by convolution with the instrumental transfer function are expected to be relatively free of foreground contamination.
 Due to an undesirable grating pattern (dot--dashed lines) of the instrument used in this study, namely the MWA, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the {\it EoR window} in this paper.
@@
 This flattens any residual variation in bandpass and removes small phase slopes due to variations in cable delay.
-Finally, we fit for an additional phase known to be caused by reflections in 150~m length cables.
+Finally, we fit for an additional phase known to be caused by reflections in a few 150~m length cables.
 %possibly could add elaboration, if I knew exactly which version of the cal was done here.
@@
 We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known repeatability and stability level of the analog signal chain \citep{bow07b}.
-The power--patterns are shown in Figure~\ref{fig:power_pattern} for the pointings chosen in our study.
+Besides having the effect of adding a time--dependent uncertainty in the power pattern, these random delay fluctuations reduce the coherence in the phased addition of dipole signals resulting in deviations from predicted models of the power--pattern, most prominently at its nulls.
+The power--patterns so obtained are shown in Figure~\ref{fig:power_pattern} for the pointings chosen in our study.
 
@@
 
+The model described in the preceding section is primarily a model of the diffuse foreground sky.
+While it contains faint compact emission blended in with the diffuse emission, it is known to be devoid of bright compact sources of emission \citep{deo08}.
+In order to supplement it with missing bright compact emission, we use classical radio source confusion estimates to determine the nominal flux density threshold and use compact sources above this threshold.
+It may be noted that slightly different criteria are in common use in radio astronomy to estimate radio source confusion \citep[see Appendix of][and references therein]{thy13}.
+For an angular resolution of 0\fdg 85, using a conservative `$S_\textrm{c}=5\sigma_\textrm{c}$' criterion, we determine the flux density threshold to be $\approx 10$~Jy.
+Other liberal criteria that yield a lower threshold carry a greater risk of double--counting compact sources which might be already blended in the diffuse sky model.
+
 We use a combination of the NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and the Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz to provide our compact object catalog due to their complementary survey footprints covering the entire sky, and matched flux density sensitivity and angular resolution.
@@
 Together with the diffuse foreground model, we obtain an all--sky foreground model consisting of both compact and diffuse emission.
+Figure~\ref{fig:CSM} shows the compact foreground emission model used in our study whose with contours of the power pattern overlaid.
+The flux densities of the compact sources are indicated by the color scale.
 
@@
 \end{figure}
+
+It is worth emphasizing that the ideal foreground model is one with an angular resolution finer than that of the instrument that covers the whole sky at the observing frequency, a point we will revisit later for another reason.
+For this study, we note that our conclusions are not significantly affected despite minor differences in the thresholds from different radio source confusion estimates.
 
@@
 The Galactic center manifests itself most distinctly near the negative horizon delay limit on short eastward baselines in the delay spectrum (bottom panel of Figure~\ref{fig:baseline-breakup}).
-Consequently, the spillover caused by the instrument's transfer function from the {\it foreground wedge} into the {\it EoR window} is most severe in eastward baselines relative to northward baselines evident by the bright vertical stripes of foreground contamination.
+Consequently, the spillover caused by the instrument's transfer function from the {\it foreground wedge} into the {\it EoR window} is most severe in eastward baselines (particularly the short ones) relative to northward baselines evident by the bright vertical stripes of foreground contamination.
 
@@
 (c): Simulated delay spectra $|V_{b\tau}(\vec{b},\tau)|$ on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}).
-The bright Galactic center is prominently visible close to negative horizon delay limit on eastward antenna spacings.
+The bright Galactic center is prominently visible close to negative horizon delay limit, especially on short eastward antenna spacings.
 These are also the most severely contaminated by foreground spillover.
@@
 
-In our example, we discard eastward antenna spacings (-22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5) of lengths $|\vec{b}| < 40$~m.
+In our example, we discard eastward antenna spacings (-22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5) of lengths $|\vec{b}| < 30$~m.
 Figure~\ref{fig:before-after} shows the delay spectra obtained with all antenna spacings (top panel) and after applying our screening technique (bottom panel) on the {\it off--zenith} observation.
-Notice the remarkable reduction --- by about an order of magnitude --- in foreground contamination in the {\it EoR window} after application of our screening technique.
+Notice the remarkable reduction in foreground spillover into the {\it EoR window} via the removal of bright vertical stripes on short eastward antenna spacings resulting from the application of our screening technique.
 The color scale range has been adjusted relative to the previous figures to show this significant reduction in foreground contamination.
@@
 \caption{Simulated delay spectra amplitudes $|V_{b\tau}(\vec{b},\tau)|$ for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
-Discarded antenna spacings (black vertical stripes) have lengths $|\vec{b}| < 40$~m and orientations -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5.
+Discarded antenna spacings (black vertical stripes) have lengths $|\vec{b}| < 30$~m and orientations -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5.
 The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} prominent when all antenna spacings are used is lowered by an order of magnitude when short eastward antenna spacings are discarded.
@@
 This screening technique can be generalized to optimize between foreground mitigation and loss of sensitivity from discarding data.
-Figure~\ref{fig:screening} shows how the typical foreground contamination\footnote{We use a statistical measure of foreground contamination using the standard deviation of noiseless $V_{b\tau}(\vec{b},\tau)$ from foregrounds in the {\it EoR window}.} in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
+Figure~\ref{fig:screening} shows how the typical foreground contamination\footnote{We use standard deviation of noiseless $V_{b\tau}(\vec{b},\tau)$ from foregrounds in the {\it EoR window} as a measure of foreground contamination.} in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
 We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars).
@@
 Foreground contamination estimated from only northward (dashed line) and eastward (dot--dashed line) antenna spacings represent the lower and upper limits, respectively, and are shown for reference.
+The fraction of antenna spacings discarded relative to the total number of antenna spacings is shown in dotted lines for different ranges of $\theta_\textrm{b}$.
 It is seen that foreground contamination can be mitigated by a factor between $\sim 1.5$ ($|\theta_\textrm{b}|\le$~7\fdg 5) and $\sim 10$ ($|\theta_\textrm{b}|\le$~22\fdg 5).
 It is notable that short antenna spacings play maximum role in determining foreground contamination.
-For instance, among antenna spacings with orientations $|\theta_\textrm{b}|\le$~22\fdg 5, discarding those with lengths $|\vec{b}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity.
+For instance, among antenna spacings with orientations $|\theta_\textrm{b}|\le$~22\fdg 5, discarding those with lengths $|\vec{b}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity as the fraction of discarded baselines increases from $\sim 5$\% to $\sim 25$\%.
 On the other hand, a bright compact object at the same location will give rise to foreground contamination even on longer antenna spacings.
-Such cases will necessitate discarding more or all of the eastward oriented antenna spacings.
+Such cases will necessitate discarding more or, in the worst case, all of the eastward oriented antenna spacings accounting for $\sim 25$\% loss of data.
 
@@
 Antenna spacings oriented eastward with varying degrees of directedness are considered: -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars).
-Among antenna spacings that satisfy the above criteria, ones with lengths $\le |\vec{b}|_\textrm{max}$ ($x$--axis) are discarded.
+Among antenna spacings that satisfy the above criteria, ones with lengths $\le |\vec{b}|_\textrm{max}$ ($x$--axis) are discarded (dotted lines).
 Lower (dashed line) and upper (dot--dashed) limits on foreground contamination are derived from northward (67\fdg 5~$\le\theta_\textrm{b}<$~112\fdg 5) and eastward ($|\theta_\textrm{b}|\le$~22\fdg 5) antenna spacings respectively.
-Foreground contamination in the {\it EoR window} reduces by factor $\sim$1.5 ($|\theta_\textrm{b}|\le$~7\fdg 5) to $\sim$10 ($|\theta_\textrm{b}|\le$~22\fdg 5).
+Foreground contamination in the {\it EoR window} (solid lines) reduces by factor $\sim$1.5 ($|\theta_\textrm{b}|\le$~7\fdg 5) to $\sim$10 ($|\theta_\textrm{b}|\le$~22\fdg 5).
 Short baselines have maximum effect on foreground contamination.
@@
 In conclusion, we find that inclusion of objects, both diffuse and compact, all the way to the horizon is essential to explaining the observed power spectrum.
-We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination by an order of magnitude in EoR data analysis.
+We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination by an order of magnitude in EoR data analysis with a mere 5\% loss of data.
 It is possible to discard or down--weight antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation.

### 2014-10-19 foreground_delay_spectrum_modeling.tex
From 910390d: ============
(*) Made very minor change in abstract
(*) Addressed most of Gianni's comments
(*) Added few more references

@@
 \def\SKASA{\altaffilmark{2}}
-\def\SKASAtxt{\altaffiltext{2}{Square Kilometre Array South Africa (SKA SA), Cape Town 7405, South Africa}}
+\def\SKASAtxt{\altaffiltext{2}{Square Kilometre Array South Africa (SKA SA), Park Road, Pinelands 7405, South Africa}}
 
-\def\ANU{\altaffilmark{3}}
-\def\ANUtxt{\altaffiltext{3}{Research School of Astronomy and Astrophysics, Australian National University, Canberra, ACT 2611, Australia}}
+\def\RU{\altaffilmark{3}}
+\def\RUtxt{\altaffiltext{3}{Department of Physics and Electronics, Rhodes University, Grahamstown 6140, South Africa}}
 
-\def\Haystack{\altaffilmark{4}}
-\def\Haystacktxt{\altaffiltext{4}{MIT Haystack Observatory, Westford, MA 01886, USA}}
+\def\CfA{\altaffilmark{4}}
+\def\CfAtxt{\altaffiltext{4}{Harvard-Smithsonian Center for Astrophysics, Cambridge, MA 02138, USA}}
 
-\def\Curtin{\altaffilmark{5}}
-\def\Curtintxt{\altaffiltext{5}{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}}
+\def\ANU{\altaffilmark{5}}
+\def\ANUtxt{\altaffiltext{5}{Research School of Astronomy and Astrophysics, Australian National University, Canberra, ACT 2611, Australia}}
+
+\def\Haystack{\altaffilmark{6}}
+\def\Haystacktxt{\altaffiltext{6}{MIT Haystack Observatory, Westford, MA 01886, USA}}
+
+\def\Curtin{\altaffilmark{7}}
+\def\Curtintxt{\altaffiltext{7}{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}}
 
@@
 
-\def\MIT{\altaffilmark{6}}
-\def\MITtxt{\altaffiltext{6}{Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}}
+\def\MIT{\altaffilmark{8}}
+\def\MITtxt{\altaffiltext{8}{Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}}
 
-\def\CfA{\altaffilmark{7}}
-\def\CfAtxt{\altaffiltext{7}{Harvard-Smithsonian Center for Astrophysics, Cambridge, MA 02138, USA}}
+\def\UW{\altaffilmark{9}}
+\def\UWtxt{\altaffiltext{9}{Department of Physics, University of Washington, Seattle, WA 98195, USA}}
 
-\def\UW{\altaffilmark{8}}
-\def\UWtxt{\altaffiltext{8}{Department of Physics, University of Washington, Seattle, WA 98195, USA}}
+\def\Victoria{\altaffilmark{10}}
+\def\Victoriatxt{\altaffiltext{10}{School of Chemical \& Physical Sciences, Victoria University of Wellington, Wellington 6140, New Zealand}}
 
-\def\Victoria{\altaffilmark{9}}
-\def\Victoriatxt{\altaffiltext{9}{School of Chemical \& Physical Sciences, Victoria University of Wellington, Wellington 6140, New Zealand}}
+\def\UWisc{\altaffilmark{11}}
+\def\UWisctxt{\altaffiltext{11}{Department of Physics, University of Wisconsin--Milwaukee, Milwaukee, WI 53201, USA}}
 
-\def\UWisc{\altaffilmark{10}}
-\def\UWisctxt{\altaffiltext{10}{Department of Physics, University of Wisconsin--Milwaukee, Milwaukee, WI 53201, USA}}
+\def\UMichigan{\altaffilmark{12}}
+\def\UMichigantxt{\altaffiltext{12}{Department of Atmospheric, Oceanic and Space Sciences, University of Michigan, Ann Arbor, MI 48109, USA}}
 
-\def\UMichigan{\altaffilmark{11}}
-\def\UMichigantxt{\altaffiltext{11}{Department of Atmospheric, Oceanic and Space Sciences, University of Michigan, Ann Arbor, MI 48109, USA}}
+\def\CASS{\altaffilmark{13}}
+\def\CASStxt{\altaffiltext{13}{CSIRO Astronomy and Space Science (CASS), PO Box 76, Epping, NSW 1710, Australia}}
 
-\def\CASS{\altaffilmark{12}}
-\def\CASStxt{\altaffiltext{12}{CSIRO Astronomy and Space Science (CASS), PO Box 76, Epping, NSW 1710, Australia}}
+\def\CAASTRO{\altaffilmark{14}}
+\def\CAASTROtxt{\altaffiltext{14}{ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}}
 
-\def\CAASTRO{\altaffilmark{13}}
-\def\CAASTROtxt{\altaffiltext{13}{ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}}
+\def\Tata{\altaffilmark{15}}
+\def\Tatatxt{\altaffiltext{15}{National Centre for Radio Astrophysics, Tata Institute for Fundamental Research, Pune 411007, India}}
 
-\def\Tata{\altaffilmark{14}}
-\def\Tatatxt{\altaffiltext{14}{National Centre for Radio Astrophysics, Tata Institute for Fundamental Research, Pune 411007, India}}
+\def\RRI{\altaffilmark{16}}
+\def\RRItxt{\altaffiltext{16}{Raman Research Institute, Bangalore 560080, India}}
 
-\def\RRI{\altaffilmark{15}}
-\def\RRItxt{\altaffiltext{15}{Raman Research Institute, Bangalore 560080, India}}
+\def\NRAO{\altaffilmark{17}}
+\def\NRAOtxt{\altaffiltext{17}{National Radio Astronomy Observatory, Charlottesville and Greenbank, USA}}
 
-\def\NRAO{\altaffilmark{16}}
-\def\NRAOtxt{\altaffiltext{16}{National Radio Astronomy Observatory, Charlottesville and Greenbank, USA}}
-
-\def\UMelbourne{\altaffilmark{17}}
-\def\UMelbournetxt{\altaffiltext{17}{School of Physics, The University of Melbourne, Parkville, VIC 3010, Australia}}
+\def\UMelbourne{\altaffilmark{18}}
+\def\UMelbournetxt{\altaffiltext{18}{School of Physics, The University of Melbourne, Parkville, VIC 3010, Australia}}
 
@@
 %% Builders' list
-G.~Bernardi\SKASA,
+G.~Bernardi\SKASA$^,$\RU$^,$\CfA,
 F.~Briggs\ANU,
@@
 \SKASAtxt
+\RUtxt
+\CfAtxt
 \ANUtxt
@@
 \MITtxt
-\CfAtxt
 \UWtxt
@@
 These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space.
-Based on these results, we propose that selective down--weighting of baselines based on length, direction, and time will mitigate foreground contamination substantially (up to a factor $\sim 10$ with negligible loss of data) in redshifted 21~cm power spectrum analysis.
+Based on these results, we propose that selective down--weighting of baselines based on length, direction, and time will mitigate foreground contamination substantially (up to a factor $\sim 10$ with almost no loss of data) in redshifted 21~cm power spectrum analysis.
 
@@
 Sensitive instruments such as the Square Kilometre Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}.
-Numerous precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{tin13,bow13,lon09}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{thy13,bea13,bow06,pob14}.
+Numerous precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{tin13,bow13,lon09}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{bow06,bea13,dil13,thy13,pob14}.
 
-A key challenge in the statistical detection of the redshifted H{\sc i} 21~cm signal, via the spatial power spectrum of temperature fluctuations, arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}.
+A key challenge in the statistical detection of the redshifted H{\sc i} 21~cm signal, via the spatial power spectrum of temperature fluctuations, arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06,ali08,ber09,ber10,gho12}.
 \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}.
@@
 The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}).
-The more aggressive alternative is to subtract a sky model and regain access to modes lost by {\it avoidance}.
+The more aggressive alternative is to subtract a sky model and regain access to modes discarded by {\it avoidance}.
 In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain.
@@
 
-Foregrounds that have intrinsic spectral variations, instruments with high chromaticity, or approximations in power spectrum analyses can contaminate the true EoR H{\sc i} power spectrum.
+Foregrounds with intrinsic deviations from spectral smoothness, instruments with high chromaticity, polarization leakage, calibration errors, or approximations in power spectrum analyses can contaminate the true EoR H{\sc i} power spectrum.
 Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity.
@@
 
-Matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the signature of the power spectrum of each baseline.
-We report two important findings: the foregrounds that most severely obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a larger role than previously thought.
+After matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the signature of the power spectrum of each baseline.
+We report two important findings: the foregrounds that most severely obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted.
 % We find that the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon.
@@
 The observations shown here were undertaken on Aug~23, 2013.
-Data set from this day has been chosen as a reference for comparison within the MWA EoR collaboration because this is one of the days featuring a low number of known system errors from among a long stretch of stable observing.
-
-Here we have chosen two, 112 second long sections from this night for detailed study.
+% Data set from this day has been chosen as a reference for comparison within the MWA EoR collaboration because this is one of the days featuring a low number of known system errors from among a long stretch of stable observing.
+We have chosen two, 112 second long sections from this night for detailed study.
 These were chosen to provide a selection of possible foreground and instrumental conditions.
@@
 2014, submitted).
-MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning; it has a flux density limit of 25~mJy and a declination range of -12\arcdeg~ to -40\arcdeg~ evenly covering the field of view of the observations reported here.
+% MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning;
+It has a flux density limit of 25~mJy and a declination range of -12\arcdeg~ to -40\arcdeg~ evenly covering the field of view of the observations reported here.
 The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern.
@@
 This is shown later to be due to response of the array to the bright Galactic center and Galactic plane in the westward sky.
-Consequently, the contamination into the {\it EoR window} is also higher.
+Consequently, the contamination into the {\it EoR window} is also higher, especially on short antenna spacings.
 Emission in the {\it zenith} pointing is more centrally concentrated due to phasing of power--pattern to the zenith, whereas the {\it off--zenith} pointing has a power--pattern phased eastward of zenith but is also responsive to emission from angles far away from zenith.
@@
 
-Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the measurements perfectly, dominant at short antenna spacings where the number of delay modes within the horizon is lesser.
+% Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the measurements perfectly, dominant at short antenna spacings where the number of delay modes within the horizon is lesser.
 This is because the maximum delay envelope (boundary of the wedge) consists of mixed emission from large portions of the sky into a narrow range of delays and the deconvolution algorithm does not have sufficient support in delay space to fit the delay spectrum.
@@
 
-We describe the instrumental model and foreground models we use in our simulations.
+We describe the instrumental and foreground models we use in our simulations.
 
@@
 It is modeled as a mutually--coupled 4--by--4 dipole array with the overall power--pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14}.
-To speed up simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles.
+To speed up simulations, we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles.
 We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known repeatability and stability level of the analog signal chain \citep{bow07b}.
@@
 The median absolute deviation from these distributions corresponds to $\sim$~90\% fractional difference between data and modeling on average in either case.
-This is in line with the aforementioned uncertainty in foreground models and thermal noise fluctuations in measurements.
-Currently we are significantly limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.
+This is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account.
+Currently, we are significantly limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.
 
@@
 \caption{Normalized histograms of logarithm of ratios of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
-Median absolute deviation of $\log_{10}\rho$ is $\sim 0.28$ in both cases.
-A value of zero indicates $\rho=1$.
+Median absolute deviation of $\log_{10}\rho$ is $\sim 0.28$ in both cases indicating a fractional difference of $\sim 90$\%.
+A value of zero indicates perfect agreement ($\rho=1$).
 \label{fig:data-sim-ratio}}
@@
 These uncertainty estimates are conservative.
-A full treatment of all uncertainties and deviations from ideal behavior such as uncertainties in tile power--pattern, calibration, data corruption due to interference, and wide--field imaging effects will bring the simulations much closer in agreement with observations.
+A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power--pattern \citep{ber14}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide--field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations.
 Our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.
@@
 Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional delay spectrum plots shown in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}.
-The diffuse, compact, and composite components are are shown as dotted, dashed, and solid lines, respectively.
+The diffuse, compact, and composite components are shown as dotted, dashed, and solid lines, respectively.
 The horizon delay limits are shown as a pair of vertical dot--dashed lines.
@@
 In fact, even on widely spaced antennas, {\emph diffuse emission in the delay spectrum near the horizon delay limits exceeds that in the main lobe by more than two orders of magnitude}.
-We attribute this to apparent shortening of antenna spacings projected in the direction of diffuse emission entering far from the primary field of view.
-This demonstrates the edge--heavy {\it two--pronged fork} feature discussed in \S\ref{sec:diffuse-long-baselines}.
+We attribute this to apparent shortening of antenna spacings projected in the direction of diffuse emission entering far from the primary field of view, thereby giving rise to the edge--heavy {\it two--pronged fork} feature discussed in \S\ref{sec:diffuse-long-baselines}.
 
-Full simulations show the combination of center--heavy features on all antenna spacings from compact objects and edge--heavy features from both components especially diffuse emission even on wide antenna spacings.
+Simulations with the complete foreground model show the combination of center--heavy features on all antenna spacings from compact emission, and edge--heavy features from both compact and diffuse emission especially the latter even on wide antenna spacings.
 This results in a characteristic {\it pitchfork} structure imprinted in the {\it foreground} wedge and should be evident in observations.
@@
 Figure~\ref{fig:before-after} shows the delay spectra obtained with all antenna spacings (top panel) and after applying our screening technique (bottom panel) on the {\it off--zenith} observation.
-Notice the remarkable reduction in foreground spillover into the {\it EoR window} via the removal of bright vertical stripes on short eastward antenna spacings resulting from the application of our screening technique.
+Notice the remarkable reduction in foreground spillover into the {\it EoR window} via the removal of bright vertical stripes on short eastward antenna spacings.
 The color scale range has been adjusted relative to the previous figures to show this significant reduction in foreground contamination.
@@
 Our primary motivation in this work is to understand how the various bright foregrounds will manifest in three dimensional power spectrum of H{\sc i} from 21~cm reionization observations.
-In the units presented here, the dynamic range between bright foregrounds and the 21~cm signal is expected to be $\sim 10^8$; a detailed understanding of how foregrounds can corrupt the 21~cm power spectrum is therefore essential.
+When the signal is expressed in units of temperature variance, the dynamic range between bright foregrounds and the 21~cm signal is expected to be $\sim 10^8$; a detailed understanding of how foregrounds can corrupt the 21~cm power spectrum is therefore essential.
 This analysis extends previous work by simulating the entire sky rather than just the central field of view and by providing a comparison with early observations from the MWA.
@@
 
-Looking deeper at our simulations, below the level of the instrumental noise,  we identify numerous signatures of different components of foreground emission seen in the delay spectra.
-We establish the relationship between these signatures and observing parameters such as antenna pointing and LST, instrument parameters such as antenna power--pattern and bandpass shape, and foreground parameters such as the nature of emission, spectral index, etc.
+Our simulations enable us to identify numerous signatures of different components of foreground emission seen in the delay spectra.
+We establish the relationship between these signatures and observing parameters such as antenna pointing and LST, instrument parameters such as antenna power--pattern, and foreground parameters such as the nature of emission, spectral index, etc.
 
@@
 In conclusion, we find that inclusion of objects, both diffuse and compact, all the way to the horizon is essential to explaining the observed power spectrum.
-We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination by an order of magnitude in EoR data analysis with a mere 5\% loss of data.
-It is possible to discard or down--weight antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation.
+We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination by an order of magnitude in EoR data analysis by discarding or down--weighting antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation, without significant loss of data.
 

### 2014-10-23 foreground_delay_spectrum_modeling.tex
from 449b860: ============
(*) Incorporated Judd's comments
(*) Changed a few figures

@@
 
-Radio emission from foreground objects, due to their spatial and spectral (chromatic) structure, contaminate the received signal \citep{bow09,mor12,tro12,thy13,liu14a,liu14b}.
+Radio emission from foreground objects, due to their spatial and spectral (chromatic) structure, contaminate the received signal \citep{bow09,dat10,mor12,tro12,thy13,liu14a,liu14b}.
 Since this contamination is expected to be several orders of magnitude stronger than the underlying EoR H{\sc i} signal, it is critical to characterize foregrounds precisely in order to reduce their impact on EoR H{\sc i} power spectrum detection sensitivity.
@@
 $\tau\approx\eta\propto k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis.
-Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge}, whose boundaries are determined by antenna spacing and the light travel time across it.
+Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by antenna spacing and the light travel time across it.
 These boundaries are called horizon delay limits and are shown by solid lines.
@@
 However, the instrument used in our study, namely the MWA, has a passband constructed using coarse channels with a characteristic shape giving rise to periodic gratings thereby resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.
-The underlying cause of this effect will be discussed later in this paper.
 
@@
 We show results of delay spectrum from MWA data after deconvolution along the delay axis in Figure~\ref{fig:fhd_data}.
-The principle feature is the {\it foreground wedge}, whose boundaries, marked by solid white lines, increase linearly with light travel time corresponding to baseline length.
+The principle feature is the {\it foreground wedge}, whose boundaries, marked by white lines, increase linearly with light travel time corresponding to baseline length.
 Note that the wedge is not filled out uniformly.
@@
 \caption{Delay spectra of two observations of the EoR field (RA = 0h, Dec = $-30$\arcdeg) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
-To get the delay spectrum, each baseline is Fourier transformed into delay space ($y$--axis) and the amplitude plotted in order of baseline length ($x$--axis).
 For reference, the $k$--space equivalent of $\tau$ and $|\vec{b}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively.
-The logarithmic color scale (shown at the right) is common to both panels.
 White lines mark the boundaries of {\it foreground wedge} determined by light travel time across each baseline.
 Smooth spectrum objects predominantly reside in this {\it foreground wedge}.
-The various features in this region depend on antenna spacing orientation, secondary lobe structure of the power--pattern, and the sky at that instant.
+The logarithmic color scale (shown at the right) is common to both panels.
 \label{fig:fhd_data}}
@@
 Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the westward sky, where the MWA tile power gain is significant ($\gtrsim 12$\%).
-In the {\it zenith} pointing, the Galactic plane has set and the contour level of the power--pattern in that direction is at least 16 times smaller.
+In the {\it zenith} pointing, the Galactic plane has set and the power--pattern in that direction is at least 16 times smaller.
 
@@
 The model described in the preceding section is primarily a model of the diffuse foreground sky.
-While it contains faint compact emission blended in with the diffuse emission, it is known to be devoid of bright compact sources of emission \citep{deo08}.
+While it contains faint compact emission blended in with the diffuse emission, it is known to be devoid of bright compact sources \citep{deo08}.
 In order to supplement it with missing bright compact emission, we use classical radio source confusion estimates to determine the nominal flux density threshold and use compact sources above this threshold.
@@
 For an angular resolution of 0\fdg 85, using a conservative `$S_\textrm{c}=5\sigma_\textrm{c}$' criterion, we determine the flux density threshold to be $\approx 10$~Jy.
-Other liberal criteria that yield a lower threshold carry a greater risk of double--counting compact sources which might be already blended in the diffuse sky model.
+Other liberal criteria that yield a lower threshold carry a greater risk of double--counting compact sources which might be already blended in with the diffuse sky model.
 
@@
 
-It is worth emphasizing that the ideal foreground model is one with an angular resolution finer than that of the instrument that covers the whole sky at the observing frequency, a point we will revisit later for another reason.
+It is worth emphasizing that the ideal foreground model is one with an angular resolution finer than that of the instrument that covers the whole sky at the observing frequency, a point we will revisit later.
 For this study, we note that our conclusions are not significantly affected despite minor differences in the thresholds from different radio source confusion estimates.
@@
 Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the delay spectrum of the compact foreground sky model (see Figure~\ref{fig:CSM}) for {\it off--zenith} and {\it zenith} pointings in the top and bottom panels, respectively.
+In contrast to the delay spectrum of diffuse emission, compact emission manifests as a center--heavy structure in the delay spectrum with either pointing.
 
@@
 \includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-\caption{Simulated delay spectrum amplitudes $|V_{b\tau}(\vec{b},\tau)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise.
-White lines mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
-The axes and logarithmic color scale are identical to those in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data} and \ref{fig:noiseless-dsm-delay-spectrum}.
-Emission from compact objects gives rise to centrally concentrated features and stays flat across antenna spacings in the {\it foreground wedge}.\label{fig:noiseless-csm-delay-spectrum}}
+\caption{Same as Figure~\ref{fig:noiseless-dsm-delay-spectrum} but for a foreground model consisting of compact emission.
+Compact emission gives rise to centrally concentrated features and stays flat across antenna spacings in the {\it foreground wedge}.\label{fig:noiseless-csm-delay-spectrum}}
 \end{figure}
-
-In contrast to the delay spectrum of diffuse emission, compact emission manifests as a center--heavy structure in the delay spectrum with either pointing.
-This contributes to the {\it pitchfork} feature which will be discussed in detail in \S\ref{sec:composite}.
-Below we summarize some specific notable features obtained with our compact objects foreground model.
-
-%\subsubsection{Compact Emission in Power--Pattern Main Lobe}\label{sec:csm-main-lobe}
 
@@
 
-Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional delay spectrum plots shown in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}.
-The diffuse, compact, and composite components are shown as dotted, dashed, and solid lines, respectively.
-The horizon delay limits are shown as a pair of vertical dot--dashed lines.
+Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional delay spectrum plots shown in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum}, and \ref{fig:noiseless-csm-delay-spectrum}.
+The diffuse, compact, and composite components are shown as solid red, cyan, and black lines, respectively.
+The horizon delay limits are shown as a pair of vertical dotted lines.
 The gray shaded area denotes the envelope of expected uncertainty in the delay spectrum.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-\caption{Simulated delay spectrum amplitudes $|V_{b\tau}(\vec{b},\tau)|$ for three chosen antenna spacings of length: $\sim$84~m ({\it top}), $\sim$105~m ({\it middle}), and $\sim$170~m ({\it bottom}).
+\caption{Simulated delay spectrum amplitudes $|V_{b\tau}(\vec{b},\tau)|$ for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}).
 The antenna spacing is specified in each panel along East, North and Upward directions.
-The dotted, dashed and solid lines denote contributions from diffuse emission, compact objects and composite foreground model respectively.
-Vertical dot--dashed lines mark the horizon delay limits.
+The solid red, cyan, and black lines denote contributions from diffuse emission, compact objects and composite foreground model respectively.
+Vertical dotted lines mark the horizon delay limits.
 The shaded region denotes the envelope of uncertainty around the composite model delay spectrum.
@@
 The Galactic center manifests itself most distinctly near the negative horizon delay limit on short eastward baselines in the delay spectrum (bottom panel of Figure~\ref{fig:baseline-breakup}).
-Consequently, the spillover caused by the instrument's transfer function from the {\it foreground wedge} into the {\it EoR window} is most severe in eastward baselines (particularly the short ones) relative to northward baselines evident by the bright vertical stripes of foreground contamination.
+Consequently, the spillover caused by the instrument's transfer function from the {\it foreground wedge} into the {\it EoR window} affects the northward baselines the least and is most severe in eastward baselines (particularly the short ones) evident by the bright vertical stripes of foreground contamination.
 
@@
 (c): Simulated delay spectra $|V_{b\tau}(\vec{b},\tau)|$ on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}).
+White lines denote horizon delay limits.
 The bright Galactic center is prominently visible close to negative horizon delay limit, especially on short eastward antenna spacings.
@@
 \caption{Simulated delay spectra amplitudes $|V_{b\tau}(\vec{b},\tau)|$ for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
-Discarded antenna spacings (black vertical stripes) have lengths $|\vec{b}| < 30$~m and orientations -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5.
+Discarded antenna spacings (black vertical stripes) have lengths $|\vec{b}|<30$~m (leftward of vertical dashed line) and orientations $|\theta_\textrm{b}|<$~22\fdg 5.
 The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} prominent when all antenna spacings are used is lowered by an order of magnitude when short eastward antenna spacings are discarded.
@@
 In other words, these plots demonstrate the progress in mitigation as orientation and maximum length of discarded antenna spacings are varied.
-Foreground contamination estimated from only northward (dashed line) and eastward (dot--dashed line) antenna spacings represent the lower and upper limits, respectively, and are shown for reference.
-The fraction of antenna spacings discarded relative to the total number of antenna spacings is shown in dotted lines for different ranges of $\theta_\textrm{b}$.
+% Foreground contamination estimated from only northward (dashed line) and eastward (dot--dashed line) antenna spacings represent the lower and upper limits, respectively, and are shown for reference.
+The fraction of discarded antenna spacings discarded relative to the total number is shown in dotted lines for different ranges of $\theta_\textrm{b}$.
 It is seen that foreground contamination can be mitigated by a factor between $\sim 1.5$ ($|\theta_\textrm{b}|\le$~7\fdg 5) and $\sim 10$ ($|\theta_\textrm{b}|\le$~22\fdg 5).
@@
 Antenna spacings oriented eastward with varying degrees of directedness are considered: -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars).
-Among antenna spacings that satisfy the above criteria, ones with lengths $\le |\vec{b}|_\textrm{max}$ ($x$--axis) are discarded (dotted lines).
-Lower (dashed line) and upper (dot--dashed) limits on foreground contamination are derived from northward (67\fdg 5~$\le\theta_\textrm{b}<$~112\fdg 5) and eastward ($|\theta_\textrm{b}|\le$~22\fdg 5) antenna spacings respectively.
+Among antenna spacings that satisfy the above criteria, ones with lengths $\le |\vec{b}|_\textrm{max}$ ($x$--axis) are discarded.
+The fraction of discarded antenna spacings relative to the total number is denoted by dotted lines for the corresponding cases.
+% Lower (dashed line) and upper (dot--dashed) limits on foreground contamination are derived from northward (67\fdg 5~$\le\theta_\textrm{b}<$~112\fdg 5) and eastward ($|\theta_\textrm{b}|\le$~22\fdg 5) antenna spacings respectively.
   Foreground contamination in the {\it EoR window} (solid lines) reduces by factor $\sim$1.5 ($|\theta_\textrm{b}|\le$~7\fdg 5) to $\sim$10 ($|\theta_\textrm{b}|\le$~22\fdg 5).

### 2014-10-23 foreground_delay_spectrum_modeling.tex
from d1904dc: ============
(*) Incorporated most of Josh's comments
(*) Changed one figure a little
(*) Changed the usage of "compact objects" -- addressing comments from both Josh
and Abraham

@@
 Comparing data from recent Murchison Widefield Array observations with simulations separated into components based on type of emission, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings.
-For signals entering through the primary field of view, compact objects dominate the foreground contamination.
+For signals entering through the primary field of view, compact emission dominates the foreground contamination.
 These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space.
@@
 
-Considerable effort has been and continues to be made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,liu14a,liu14b}.
+Considerable effort has been and continues to be made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,dil14,liu14a,liu14b}.
 A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide--field (and chromatic) response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum.
@@
 The model described in the preceding section is primarily a model of the diffuse foreground sky.
-While it contains faint compact emission blended in with the diffuse emission, it is known to be devoid of bright compact sources \citep{deo08}.
-In order to supplement it with missing bright compact emission, we use classical radio source confusion estimates to determine the nominal flux density threshold and use compact sources above this threshold.
+While it contains faint compact emission blended in with the diffuse emission, it is known to be devoid of bright point sources \citep{deo08}.
+In order to supplement it with missing bright compact emission, we use classical radio source confusion estimates to determine the nominal flux density threshold and include point sources brighter than this threshold.
 It may be noted that slightly different criteria are in common use in radio astronomy to estimate radio source confusion \citep[see Appendix of][and references therein]{thy13}.
 For an angular resolution of 0\fdg 85, using a conservative `$S_\textrm{c}=5\sigma_\textrm{c}$' criterion, we determine the flux density threshold to be $\approx 10$~Jy.
-Other liberal criteria that yield a lower threshold carry a greater risk of double--counting compact sources which might be already blended in with the diffuse sky model.
+Other liberal criteria that yield a lower threshold carry a greater risk of double--counting point sources which might be already blended in with the diffuse sky model.
 
-We use a combination of the NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and the Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz to provide our compact object catalog due to their complementary survey footprints covering the entire sky, and matched flux density sensitivity and angular resolution.
+We use a combination of the NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and the Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz to provide our point source catalog due to their complementary survey footprints covering the entire sky, and matched flux density sensitivity and angular resolution.
 The SUMSS catalog covers the sky with declination $\delta < -30$\arcdeg~ with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec.
-The NVSS covers the sky with $\delta > -40$\arcdeg~ with a similar angular resolution and a limiting flux density of $\approx 2.5$~mJy for compact objects.
+The NVSS covers the sky with $\delta > -40$\arcdeg~ with a similar angular resolution and a limiting flux density of $\approx 2.5$~mJy for point sources.
 
-From the SUMSS catalog, we select compact objects whose deconvolved major axes are equal to 0\arcsec.
+From the SUMSS catalog, we select point sources whose deconvolved major axes are equal to 0\arcsec.
 From the NVSS catalog, we excluded objects that overlap with those in the SUMSS survey footprint.
-Compact objects from NVSS were selected if the convolved major axes were not greater than $\approx 47$\arcsec, which is almost equal to the angular resolution of the survey.
+Point sources from NVSS were selected if the convolved major axes were not greater than $\approx 47$\arcsec, which is almost equal to the angular resolution of the survey.
 Using a mean spectral index of $\langle\alpha_\textrm{sp}\rangle=-0.83$ (flux density, $S(f)\propto f^{\alpha_\textrm{sp}}$) obtained by \citet{mau03} for both NVSS and SUMSS catalog objects, we calculate the corresponding flux densities at 185~MHz, $S_{185}$.
-From this subset, we choose compact objects with $S_{185}\geq 10$~Jy.
-The selection of such bright compact objects is not affected by minor differences in flux density sensitivity of the two surveys.
+From this subset, we choose point sources with $S_{185}\geq 10$~Jy.
+The selection of such bright point sources is not affected by minor differences in flux density sensitivity of the two surveys.
 We verified that our selection criteria ensure a similar areal density of objects in the two surveys.
@@
 Figure~\ref{fig:CSM} shows the compact foreground emission model used in our study whose with contours of the power pattern overlaid.
-The flux densities of the compact sources are indicated by the color scale.
+The flux densities of the point sources are indicated by the color scale.
 
@@
 \includegraphics[width=\linewidth]{figures/v1_0/csm.eps}
-\caption{Flux densities of bright compact objects at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+\caption{Flux densities of bright point sources at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
 The color scale on the right is logarithmic and is common to both panels and corresponds to the color--coded filled circles.
@@
 The contour levels shown are 0.001953125, 0.0078125, 0.03125, 0.125, and 0.5.
-The locations of bright compact objects appear to be random.\label{fig:CSM}}
+\label{fig:CSM}}
 \end{figure}
@@
 Using simple error propagation, the fractional error in the delay spectrum caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the MWA observing frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index.
-From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for compact objects from NVSS and SUMSS catalogs.
-Although the model of \citet{deo08} yields a spectral index per direction on the sky, we assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from the 408~MHz map of \citet{has82}.
-Thus, fractional errors in delay spectrum from compact foreground objects and diffuse emission are $\sim$70\% and $\sim$30\% respectively.
+From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for point sources from NVSS and SUMSS catalogs.
+Although the model of \citet{deo08} yields a spectral index per direction on the sky, we could assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from the 408~MHz map of \citet{has82}.
+Thus, fractional errors in delay spectrum from compact and diffuse components are $\sim$70\% and $\sim$30\% respectively.
 
@@
 These uncertainty estimates are conservative.
-A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power--pattern \citep{ber14}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide--field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations.
+A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power--pattern \citep{ber14}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide--field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations, but is beyond the scope of this paper.
 Our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.
@@
 
-\item {\it Sky Model}: Our model of the sky consists of bright compact objects and diffuse emission on diverse spatial scales as shown in Figures~\ref{fig:DSM} and \ref{fig:CSM}.
+\item {\it Sky Model}: Our model of the sky consists of compact and diffuse emission on diverse spatial scales as shown in Figures~\ref{fig:DSM} and \ref{fig:CSM}.
 It also consists of localized regions of strong emission such as the Galactic plane and Galactic center.
@@
 
+In subsequent sections, we provide a detailed explanation of our results as a combination of factors noted above.
 Note that numerous features may overlap at different degrees of significance depending on combinations of parameters.
@@
 
-Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the amplitude of delay spectrum, $|V_{b\tau}(\vec{b},\tau)|$, simulated with only the diffuse foreground model, omitting compact sources.
+Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the amplitude of delay spectrum, $|V_{b\tau}(\vec{b},\tau)|$, simulated with only the diffuse foreground model.
 The top and bottom panels correspond to {\it off--zenith} and {\it zenith} pointings, respectively.
@@
 
-The amplitude response of an interferometer to a compact object is, to first order, flat with respect to baseline length.
+The amplitude response of an interferometer to a point source is, to first order, flat with respect to baseline length.
 Since the main lobe of the power--pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with positive delay corresponding to the position of the primary lobe of the power--pattern.
@@
 
-Foreground emission in zero delay and negative delay regions in the {\it off--zenith} pointing is caused by compact objects co--located with secondary lobes of the power--pattern.
-On the other hand, compact objects co--located with secondary lobes of power--pattern in the {\it zenith} pointing are revealed as faint but distinct branches at positive and negative delays depending on the orientation of antenna spacing and direction of emission on the sky.
+Foreground emission in zero delay and negative delay regions in the {\it off--zenith} pointing is caused by point sources co--located with secondary lobes of the power--pattern.
+On the other hand, point sources co--located with secondary lobes of power--pattern in the {\it zenith} pointing are revealed as faint but distinct branches at positive and negative delays depending on the orientation of antenna spacing and direction of emission on the sky.
 
@@
 The gray shaded area denotes the envelope of expected uncertainty in the delay spectrum.
-Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the $\gtrsim$~70\% error owing to uncertainties in predicting the spectral index of compact foreground objects, while thermal noise fluctuations dominate outside.
+Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the $\gtrsim$~70\% error owing to uncertainties in predicting the spectral index of compact foreground model, while thermal noise fluctuations dominate outside.
 
@@
 The antenna spacing is specified in each panel along East, North and Upward directions.
-The solid red, cyan, and black lines denote contributions from diffuse emission, compact objects and composite foreground model respectively.
+The solid red, cyan, and black lines denote contributions from diffuse, compact, and composite foreground models respectively.
 Vertical dotted lines mark the horizon delay limits.
@@
 Outside the horizon delay limits, it is predominantly due to thermal noise fluctuations.
-Compact objects dominate the central regions of delay spectrum while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a {\it pitchfork}--shaped structure.
+Compact emission dominates the central regions of delay spectrum while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a {\it pitchfork}--shaped structure.
 \label{fig:pitchfork-baselines}}
@@
 
-The zero--delay peak (corresponding to the main lobe in the power--pattern) with a value of $\sim 10^8$--$10^9$~Jy~Hz, independent of antenna spacing, is predominantly determined by compact foreground objects.
+The zero--delay peak (corresponding to the main lobe in the power--pattern) with a value of $\sim 10^8$--$10^9$~Jy~Hz, independent of antenna spacing, is predominantly determined by compact emission.
 The corresponding peak at zero--delay from diffuse foreground model is at least 10 times fainter and decreases drastically with an increase in antenna spacing.
-This is the response expected from different antenna spacings towards compact and diffuse foreground objects.
+This is the response expected from different antenna spacings towards compact and diffuse emission.
 
-Near the horizon delay limits at shorter antenna spacing, diffuse emission is brighter than that from compact objects.
+Near the horizon delay limits at shorter antenna spacing, the diffuse component is brighter relative to the compact component.
 Here, diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at zero--delay.
@@
 The Galactic center in the {\it off--zenith} pointing is one such example already available in our study.
-Figure~\ref{fig:sky-model} shows the sky model ({\it top:} compact objects, {\it bottom:} diffuse sky model) in this pointing.
+Figure~\ref{fig:sky-model} shows the sky model ({\it top:} compact component, {\it bottom:} diffuse component) in this pointing.
 The Galactic center is the most dominant source of foreground contamination from the diffuse sky model and is co--located with a bright secondary lobe of the power pattern near the western horizon.
@@
 \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figures/v1_0/baseline_orientation_breakup_snapshot_0.eps}}
-\caption{(a): Sky model showing compact objects ({\it top}) and diffuse emission ({\it bottom}).
+\caption{(a): Sky model showing compact ({\it top}) and diffuse ({\it bottom}) emission.
 Contours of power pattern are overlaid whose levels are identical to those in Figures~\ref{fig:DSM} and \ref{fig:CSM}.
@@
 For instance, among antenna spacings with orientations $|\theta_\textrm{b}|\le$~22\fdg 5, discarding those with lengths $|\vec{b}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity as the fraction of discarded baselines increases from $\sim 5$\% to $\sim 25$\%.
-On the other hand, a bright compact object at the same location will give rise to foreground contamination even on longer antenna spacings.
+On the other hand, a bright point source at the same location will give rise to foreground contamination even on longer antenna spacings.
 Such cases will necessitate discarding more or, in the worst case, all of the eastward oriented antenna spacings accounting for $\sim 25$\% loss of data.
@@
 
-Simulating in all important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright compact objects from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise fluctuations in measurements.
+Simulating in all important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright point sources from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise fluctuations in measurements.
 
@@
 This is due to apparent shortening of wide antenna spacings in the direction of foreground emission far off--axis, thereby retaining their response to extended emission.
-On the other hand, compact foreground objects predominantly map onto central regions of the {\it foreground wedge}.
-Features arising from compact objects co--located with primary and secondary lobes of the antenna power--pattern have been identified.
-In general, delay spectrum signatures of compact objects are center--heavy, in clear contrast to those from diffuse emission.
+On the other hand, compact emission predominantly maps onto central regions of the {\it foreground wedge}.
+Features arising from compact emission co--located with primary and secondary lobes of the antenna power--pattern have been identified.
+In general, delay spectrum signatures of compact emission are center--heavy, in clear contrast to those from diffuse emission.
 A composite all--sky foreground model consisting of diffuse and compact foregrounds combines the characteristic individual shapes into a {\it pitchfork}--shaped structure in the {\it foreground wedge}.
@@
 
-In conclusion, we find that inclusion of objects, both diffuse and compact, all the way to the horizon is essential to explaining the observed power spectrum.
+In conclusion, we find that inclusion of emission models, both diffuse and compact, all the way to the horizon is essential to explaining the observed power spectrum.
 We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination by an order of magnitude in EoR data analysis by discarding or down--weighting antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation, without significant loss of data.

### 2014-10-23 foreground_delay_spectrum_modeling.tex
from 4d54f99: ============
(*) Changed paragraph justifying discrepancy between data and simulations.
(*) Commented out the figure showing the histogram of log(rho) and instead
justified by text

@@
 We estimate the ratio of delay spectra from data and simulations as $\rho = |V^\textrm{D}_{b\tau}(\vec{b},\tau)|\,/\,|V^\textrm{S}_{b\tau}(\vec{b},\tau)|$, where superscripts D and S denote data and simulation, respectively.
-% Using the aforementioned uncertainties added in quadrature, we estimate the approximate expected uncertainty in $\log_{10}\rho$ and denote it by $\Delta\,\log_{10}\rho$.
 
-Figure~\ref{fig:data-sim-ratio} shows normalized histograms of the distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}.
+The median absolute deviation of $\log_{10}\rho$ inside the {\it foreground wedge} for both pointings is $\approx 0.28$.
+This corresponds to $\sim$~90\% fractional difference between data and modeling on average with either pointing.
+We also simulated delay spectra after assigning spectral indices drawn randomly from a {\it gaussian} distribution with a mean of $\langle\alpha_\textrm{sp}\rangle=-0.83$ and a HWHM of $\Delta\alpha_\textrm{sp}=0.35$ to the point sources in our compact foreground model.
+These simulations typically yielded a median absolute deviation of $\approx 0.29$ for $\log_{10}\rho$ indicating fractional differences of $\sim$~95\%.
+This demonstrates that a fractional deviation of $\sim$~90\% observed between data and simulations is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account.
+Currently, we are significantly limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.
+
+% Figure~\ref{fig:data-sim-ratio} shows normalized histograms of the distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}.
 The median absolute deviation from these distributions corresponds to $\sim$~90\% fractional difference between data and modeling on average in either case.
@@
 
-\begin{figure}[htb]
-\centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-\caption{Normalized histograms of logarithm of ratios of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+% \begin{figure}[htb]
+% \centering
+% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
+% \caption{Normalized histograms of logarithm of ratios of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
 Median absolute deviation of $\log_{10}\rho$ is $\sim 0.28$ in both cases indicating a fractional difference of $\sim 90$\%.
@@
 \label{fig:data-sim-ratio}}
-\end{figure}
+% \end{figure}
 
-These uncertainties are presented to confirm the qualitative agreement already seen between Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}.
-These uncertainty estimates are conservative.
+These uncertainties are presented only to confirm the qualitative agreement already seen between Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}.
+These estimates are conservative.
 A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power--pattern \citep{ber14}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide--field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations, but is beyond the scope of this paper.
@@
 Secondly, we have used noiseless cases to clearly illustrate the observed foreground signatures.
-With the addition of noise in the visibilities which is subject to observing time, some of the weaker features may not be as prominently visible.
+With the addition of noise in the visibilities, some of the weaker features may not be as prominently visible.
 Since the foreground signatures are far too numerous and they reside in a vast volume of parameter space such as antenna spacing length and orientation, power--pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features observed in the delay spectra of foreground emission.

### 2014-10-23 foreground_delay_spectrum_modeling.tex
from 6865031: ============
(*) Removed the histogram figure from git tracking
(*) Removed the corresponding commented portions relating to the histogram
figure
Minor edits to address collab comments. Mostly section 3

@@
 
-The MWA construction was completed in 2012 and, after commissioning, began its EoR observing program in 2013.
+MWA construction was completed in 2012 and, after commissioning, began its EoR observing program in 2013.
 In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz.
@@
 The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern.
-The calibration algorithm -- based on forward modeling software by \citet{sul12} -- computes complex gain solutions per channel per antenna averaged to two minute intervals.
-The solutions are fairly low signal--to--noise so we iteratively project along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation.
+The calibration algorithm -- based on forward modeling software by \citet{sul12} and the calibration method described by \citet{salvini2014fast} -- computes complex gain solutions per channel per antenna averaged to two minute intervals.
+The solutions are fairly low signal--to--noise so we iteratively average along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation.
 First, we average the channel gains over all antennas to obtain a high signal--to--noise measurement of the bandpass.
@@
 This flattens any residual variation in bandpass and removes small phase slopes due to variations in cable delay.
-Finally, we fit for an additional phase known to be caused by reflections in a few 150~m length cables.
+Finally, we fit for an additional phase known to be caused by small reflections in subset of cables.
 %possibly could add elaboration, if I knew exactly which version of the cal was done here.
@@
 
+%@article{salvini2014fast,
+%  title={Fast gain calibration in radio astronomy using alternating direction implicit methods: Analysis and applications},
+%  author={Salvini, Stefano and Wijnholds, Stefan J},
+%  journal={arXiv preprint arXiv:1410.2101},
+%  year={2014}
+%}
+
 \subsection{Deconvolved Delay Spectrum}\label{sec:data-delay-spectrum}
 
-We obtain the delay spectrum of these calibrated visibilities taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function.
+We obtain the delay spectrum of these calibrated visibilities by taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function.
 The spectrum is multiplied by the additional weight of the flagged channels which occur rarely for interference and regularly every 1.28~MHz where the edges of coarse channels are known to have a small aliasing effect.
@@
 Currently, we are significantly limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.
-
-% Figure~\ref{fig:data-sim-ratio} shows normalized histograms of the distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}.
-The median absolute deviation from these distributions corresponds to $\sim$~90\% fractional difference between data and modeling on average in either case.
-This is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account.
-Currently, we are significantly limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.
-
-% \begin{figure}[htb]
-% \centering
-% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-% \caption{Normalized histograms of logarithm of ratios of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
-Median absolute deviation of $\log_{10}\rho$ is $\sim 0.28$ in both cases indicating a fractional difference of $\sim 90$\%.
-A value of zero indicates perfect agreement ($\rho=1$).
-\label{fig:data-sim-ratio}}
-% \end{figure}
 

### 2014-10-24 foreground_delay_spectrum_modeling.tex
from 2845225: ============
(*) Added a reference to eor.bib

@@
 The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern.
-The calibration algorithm -- based on forward modeling software by \citet{sul12} and the calibration method described by \citet{salvini2014fast} -- computes complex gain solutions per channel per antenna averaged to two minute intervals.
+The calibration algorithm -- based on forward modeling software by \citet{sul12} and the calibration method described by \citet{sal14} -- computes complex gain solutions per channel per antenna averaged to two minute intervals.
 The solutions are fairly low signal--to--noise so we iteratively average along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation.
@@
 Finally, we fit for an additional phase known to be caused by small reflections in subset of cables.
-%possibly could add elaboration, if I knew exactly which version of the cal was done here.
-doesn't really matter
-
-%@article{salvini2014fast,
-%  title={Fast gain calibration in radio astronomy using alternating direction implicit methods: Analysis and applications},
-%  author={Salvini, Stefano and Wijnholds, Stefan J},
-%  journal={arXiv preprint arXiv:1410.2101},
-%  year={2014}
-%}
 

### 2014-10-24 foreground_delay_spectrum_modeling.tex
Changed title

@@
 
-\shorttitle{Foreground signatures and mitigation in EoR H{\sc i} power spectra}
+\shorttitle{Dissection of foregrounds in EoR H{\sc i} power spectra}
 \shortauthors{Thyagarajan et~al.}
@@
 %% \title{Limits on the detection of the Epoch of Reionization from MWA observations of the redshifted 21~cm line}
-\title{Foreground Signatures and Mitigation in Neutral Hydrogen Power Spectra From The Reionization Epoch}
+\title{A Dissection of Foregrounds in Epoch of Reionization Observations from the Murchison Widefield Array}
 

### 2014-10-24 foreground_delay_spectrum_modeling.tex
Minor change from 170 m to 171 m for a baseline length

@@
 This improves the localization of foreground objects whose signatures are imprinted in the delay spectrum.
-For instance, there is an increase in the number of secondary peaks in the delay spectrum between zero--delay and horizon delay limits as the antenna spacing increases from $\sim$84~m to $\sim$170~m.
+For instance, there is an increase in the number of secondary peaks in the delay spectrum between zero--delay and horizon delay limits as the antenna spacing increases from $\sim$84~m to $\sim$171~m.
 In this case, these correspond to secondary lobes of the power--pattern that lie between the primary lobe and the horizon along the local meridian.
@@
 Diffuse emission co--located at the primary lobe of the power--pattern is fainter by approximately an order of magnitude relative to compact emission from the same region for a $\sim 84$~m antenna spacing.
-But unlike the latter, diffuse emission decreases rapidly by over two orders of magnitude as the antenna spacing is widened to $\sim 170$~m.
+But unlike the latter, diffuse emission decreases rapidly by over two orders of magnitude as the antenna spacing is widened to $\sim 171$~m.
 However, diffuse emission is significantly higher near horizon delay limits compared to that in the primary lobe of the power--pattern as the antenna spacing is increased.

### 2014-10-24 foreground_delay_spectrum_modeling.tex
Explanation added for approximation between eta and k_parallel

@@
 It is directly related to the line--of--sight wavenumber, $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
+This approximation holds under the assumption that there is negligible cosmological evolution within the redshift range (or frequency band).
 Thus,
@@
 $\tau\approx\eta\propto k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis.
-Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by antenna spacing and the light travel time across it.
+Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by the antenna spacings and the light travel times across them.
 These boundaries are called horizon delay limits and are shown by solid lines.

### 2014-10-27 foreground_delay_spectrum_modeling.tex
Changed the number of SUMSS and NVSS sources in the compact foreground model
Rearranged author list and affiliations after including EoR authors and changed the acknowledgments

@@
 
-\def\SKASA{\altaffilmark{2}}
-\def\SKASAtxt{\altaffiltext{2}{Square Kilometre Array South Africa (SKA SA), Park Road, Pinelands 7405, South Africa}}
+\def\UW{\altaffilmark{2}}
+\def\UWtxt{\altaffiltext{2}{Department of Physics, University of Washington, Seattle, WA 98195, USA}}
 
-\def\RU{\altaffilmark{3}}
-\def\RUtxt{\altaffiltext{3}{Department of Physics and Electronics, Rhodes University, Grahamstown 6140, South Africa}}
+\def\SKASA{\altaffilmark{3}}
+\def\SKASAtxt{\altaffiltext{3}{Square Kilometre Array South Africa (SKA SA), Park Road, Pinelands 7405, South Africa}}
 
-\def\CfA{\altaffilmark{4}}
-\def\CfAtxt{\altaffiltext{4}{Harvard-Smithsonian Center for Astrophysics, Cambridge, MA 02138, USA}}
+\def\RU{\altaffilmark{4}}
+\def\RUtxt{\altaffiltext{4}{Department of Physics and Electronics, Rhodes University, Grahamstown 6140, South Africa}}
 
-\def\ANU{\altaffilmark{5}}
-\def\ANUtxt{\altaffiltext{5}{Research School of Astronomy and Astrophysics, Australian National University, Canberra, ACT 2611, Australia}}
+\def\CfA{\altaffilmark{5}}
+\def\CfAtxt{\altaffiltext{5}{Harvard-Smithsonian Center for Astrophysics, Cambridge, MA 02138, USA}}
 
-\def\Haystack{\altaffilmark{6}}
-\def\Haystacktxt{\altaffiltext{6}{MIT Haystack Observatory, Westford, MA 01886, USA}}
+\def\ANU{\altaffilmark{6}}
+\def\ANUtxt{\altaffiltext{6}{Research School of Astronomy and Astrophysics, Australian National University, Canberra, ACT 2611, Australia}}
 
-\def\Curtin{\altaffilmark{7}}
-\def\Curtintxt{\altaffiltext{7}{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}}
+\def\CAASTRO{\altaffilmark{7}}
+\def\CAASTROtxt{\altaffiltext{7}{ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}}
 
-% \def\USydney{\altaffilmark{7}}
-% \def\USydneytxt{\altaffiltext{7}{Sydney Institute for Astronomy, School of Physics, The University of Sydney, NSW 2006, Australia}}
+\def\Haystack{\altaffilmark{8}}
+\def\Haystacktxt{\altaffiltext{8}{MIT Haystack Observatory, Westford, MA 01886, USA}}
 
-\def\MIT{\altaffilmark{8}}
-\def\MITtxt{\altaffiltext{8}{Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}}
+\def\MIT{\altaffilmark{9}}
+\def\MITtxt{\altaffiltext{9}{Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}}
 
-\def\UW{\altaffilmark{9}}
-\def\UWtxt{\altaffiltext{9}{Department of Physics, University of Washington, Seattle, WA 98195, USA}}
+\def\Curtin{\altaffilmark{10}}
+\def\Curtintxt{\altaffiltext{10}{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}}
 
-\def\Victoria{\altaffilmark{10}}
-\def\Victoriatxt{\altaffiltext{10}{School of Chemical \& Physical Sciences, Victoria University of Wellington, Wellington 6140, New Zealand}}
+\def\Victoria{\altaffilmark{11}}
+\def\Victoriatxt{\altaffiltext{11}{School of Chemical \& Physical Sciences, Victoria University of Wellington, Wellington 6140, New Zealand}}
 
-\def\UWisc{\altaffilmark{11}}
-\def\UWisctxt{\altaffiltext{11}{Department of Physics, University of Wisconsin--Milwaukee, Milwaukee, WI 53201, USA}}
+\def\UWisc{\altaffilmark{12}}
+\def\UWisctxt{\altaffiltext{12}{Department of Physics, University of Wisconsin--Milwaukee, Milwaukee, WI 53201, USA}}
 
-\def\UMichigan{\altaffilmark{12}}
-\def\UMichigantxt{\altaffiltext{12}{Department of Atmospheric, Oceanic and Space Sciences, University of Michigan, Ann Arbor, MI 48109, USA}}
+\def\UMichigan{\altaffilmark{13}}
+\def\UMichigantxt{\altaffiltext{13}{Department of Atmospheric, Oceanic and Space Sciences, University of Michigan, Ann Arbor, MI 48109, USA}}
 
-\def\CASS{\altaffilmark{13}}
-\def\CASStxt{\altaffiltext{13}{CSIRO Astronomy and Space Science (CASS), PO Box 76, Epping, NSW 1710, Australia}}
+\def\UMelbourne{\altaffilmark{14}}
+\def\UMelbournetxt{\altaffiltext{14}{School of Physics, The University of Melbourne, Parkville, VIC 3010, Australia}}
 
-\def\CAASTRO{\altaffilmark{14}}
-\def\CAASTROtxt{\altaffiltext{14}{ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}}
+\def\USydney{\altaffilmark{15}}
+\def\USydneytxt{\altaffiltext{15}{Sydney Institute for Astronomy, School of Physics, The University of Sydney, NSW 2006, Australia}}
 
-\def\Tata{\altaffilmark{15}}
-\def\Tatatxt{\altaffiltext{15}{National Centre for Radio Astrophysics, Tata Institute for Fundamental Research, Pune 411007, India}}
+\def\CASS{\altaffilmark{16}}
+\def\CASStxt{\altaffiltext{16}{CSIRO Astronomy and Space Science (CASS), PO Box 76, Epping, NSW 1710, Australia}}
 
-\def\RRI{\altaffilmark{16}}
-\def\RRItxt{\altaffiltext{16}{Raman Research Institute, Bangalore 560080, India}}
+\def\Tata{\altaffilmark{17}}
+\def\Tatatxt{\altaffiltext{17}{National Centre for Radio Astrophysics, Tata Institute for Fundamental Research, Pune 411007, India}}
 
-\def\NRAO{\altaffilmark{17}}
-\def\NRAOtxt{\altaffiltext{17}{National Radio Astronomy Observatory, Charlottesville and Greenbank, USA}}
+\def\RRI{\altaffilmark{18}}
+\def\RRItxt{\altaffiltext{18}{Raman Research Institute, Bangalore 560080, India}}
 
-\def\UMelbourne{\altaffilmark{18}}
-\def\UMelbournetxt{\altaffiltext{18}{School of Physics, The University of Melbourne, Parkville, VIC 3010, Australia}}
+\def\NRAO{\altaffilmark{19}}
+\def\NRAOtxt{\altaffiltext{19}{National Radio Astronomy Observatory, Charlottesville and Greenbank, USA}}
 
@@
 Judd~D.~Bowman\ASU,
-%% Supporting authors
-% A.~P.~Beardsley\UW,
-% J.~Pober\UW,
-% I.~S.~Sullivan\UW,
-%% Builders' list
+N.~Barry\UW,
+A.~P.~Beardsley\UW,
 G.~Bernardi\SKASA$^,$\RU$^,$\CfA,
-F.~Briggs\ANU,
+F.~Briggs\ANU$^,$\CAASTRO,
 R.~J.~Cappallo\Haystack, 
+P.~Carroll\UW,
 B.~E.~Corey\Haystack, 
 % A.~A.~Deshpande\RRI, 
+A.~de~Oliveira--Costa\MIT,
+J.~S.~Dillon\MIT,
 D.~Emrich\Curtin,
 % B.~M.~Gaensler\USydney$^,$\CAASTRO, 
+A.~Ewall--Wice\MIT,
+L.~Feng\MIT,
 R.~Goeke\MIT,
@@
 B.~J.~Hazelton\UW, 
+J.~N.~Hewitt\MIT,
+N.~Hurley--Walker\Curtin,
 M.~Johnston-Hollitt\Victoria,
@@
 J.~C.~Kasper\UMichigan$^,$\CfA, 
+Han--Seek Kim\UMelbourne$^,$\CAASTRO,
+P.~Kittiwisit\ASU,
 E.~Kratzenberg\Haystack, 
+E.~Lenc\USydney$^,$\CAASTRO,
+J.~Line\UMelbourne$^,$\CAASTRO,
+A.~Loeb\CfA,
 C.~J.~Lonsdale\Haystack, 
 M.~J.~Lynch\Curtin, 
+B.~McKinley\UMelbourne$^,$\CAASTRO,
 S.~R.~McWhirter\Haystack,
@@
 E.~Morgan\MIT, 
+A.~R.~Neben\MIT,
 D.~Oberoi\Tata, 
+A.~R.~Offringa\ANU$^,$\CAASTRO, 
 S.~M.~Ord\Curtin$^,$\CAASTRO,
+Sourabh Paul\RRI,
+B.~Pindor\UMelbourne$^,$\CAASTRO,
+J.~C.~Pober\UW,
 T.~Prabu\RRI, 
+P.~Procopio\UMelbourne$^,$\CAASTRO,
+J.~Riding\UMelbourne$^,$\CAASTRO,
 A.~E.~E.~Rogers\Haystack, 
@@
 N.~Udaya~Shankar\RRI, 
+Shiv~K.~Sethi\RRI,
 K.~S.~Srivani\RRI, 
 R.~Subrahmanyan\RRI$^,$\CAASTRO, 
+I.~S.~Sullivan\UW,
+M.~Tegmark\MIT,
 S.~J.~Tingay\Curtin$^,$\CAASTRO, 
+C.~M.~Trott\Curtin$^,$\CAASTRO,
 M.~Waterson\Curtin$^,$\ANU,
@@
 C.~L.~Williams\MIT,
-and~MWA~EoR~members
+J.~S.~B.~Wyithe\UMelbourne$^,$\CAASTRO
 }
@@
 \ASUtxt
+\UWtxt
 \SKASAtxt
@@
 \ANUtxt
+\CAASTROtxt
 \Haystacktxt
+\MITtxt
 \Curtintxt
-% \USydneytxt
-\MITtxt
-\UWtxt
 \Victoriatxt
@@
 \UMichigantxt
+\UMelbournetxt
+\USydneytxt
 \CASStxt
-\CAASTROtxt
 \Tatatxt
@@
 \NRAOtxt
-\UMelbournetxt
 
@@
 
-Based on these criteria, we selected 133 objects from the SUMSS catalog and 336 objects from the NVSS catalog.
+Based on these criteria, we selected 100 objects from the SUMSS catalog and 250 objects from the NVSS catalog.
 Together with the diffuse foreground model, we obtain an all--sky foreground model consisting of both compact and diffuse emission.
@@
 DCJ is supported by an NSF Astronomy and Astrophysics Postdoctoral Fellowship under award AST--1401708.
+JCP is supported by an NSF Astronomy and Astrophysics Fellowship under award AST-1302774.
 This work makes use of the Murchison Radio-astronomy Observatory, operated by CSIRO.

### 2014-10-27 foreground_delay_spectrum_modeling.tex
Rearranged addresses in affiliations

@@
 \def\ASU{\altaffilmark{1}}
-\def\ASUtxt{\altaffiltext{1}{School of Earth and Space Exploration, Arizona State University, Tempe, AZ 85287, USA; e-mail: t\_nithyanandan@asu.edu}}
+\def\ASUtxt{\altaffiltext{1}{Arizona State University, School of Earth and Space Exploration, Tempe, AZ 85287, USA; e-mail: t\_nithyanandan@asu.edu}}
 
 \def\UW{\altaffilmark{2}}
-\def\UWtxt{\altaffiltext{2}{Department of Physics, University of Washington, Seattle, WA 98195, USA}}
+\def\UWtxt{\altaffiltext{2}{University of Washington, Department of Physics, Seattle, WA 98195, USA}}
 
@@
 \def\ANU{\altaffilmark{6}}
-\def\ANUtxt{\altaffiltext{6}{Research School of Astronomy and Astrophysics, Australian National University, Canberra, ACT 2611, Australia}}
+\def\ANUtxt{\altaffiltext{6}{Australian National University, Research School of Astronomy and Astrophysics, Canberra, ACT 2611, Australia}}
 
@@
 \def\MIT{\altaffilmark{9}}
-\def\MITtxt{\altaffiltext{9}{Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}}
+\def\MITtxt{\altaffiltext{9}{MIT Kavli Institute for Astrophysics and Space Research, Cambridge, MA 02139, USA}}
 
 \def\Curtin{\altaffilmark{10}}
-\def\Curtintxt{\altaffiltext{10}{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}}
+\def\Curtintxt{\altaffiltext{10}{International Centre for Radio Astronomy Research, Curtin University, Perth, WA 6845, Australia}}
 
 \def\Victoria{\altaffilmark{11}}
-\def\Victoriatxt{\altaffiltext{11}{School of Chemical \& Physical Sciences, Victoria University of Wellington, Wellington 6140, New Zealand}}
+\def\Victoriatxt{\altaffiltext{11}{Victoria University of Wellington, School of Chemical \& Physical Sciences, Wellington 6140, New Zealand}}
 
 \def\UWisc{\altaffilmark{12}}
-\def\UWisctxt{\altaffiltext{12}{Department of Physics, University of Wisconsin--Milwaukee, Milwaukee, WI 53201, USA}}
+\def\UWisctxt{\altaffiltext{12}{University of Wisconsin--Milwaukee, Department of Physics, Milwaukee, WI 53201, USA}}
 
 \def\UMichigan{\altaffilmark{13}}
-\def\UMichigantxt{\altaffiltext{13}{Department of Atmospheric, Oceanic and Space Sciences, University of Michigan, Ann Arbor, MI 48109, USA}}
+\def\UMichigantxt{\altaffiltext{13}{University of Michigan, Department of Atmospheric, Oceanic and Space Sciences, Ann Arbor, MI 48109, USA}}
 
 \def\UMelbourne{\altaffilmark{14}}
-\def\UMelbournetxt{\altaffiltext{14}{School of Physics, The University of Melbourne, Parkville, VIC 3010, Australia}}
+\def\UMelbournetxt{\altaffiltext{14}{The University of Melbourne, School of Physics, Parkville, VIC 3010, Australia}}
 
 \def\USydney{\altaffilmark{15}}
-\def\USydneytxt{\altaffiltext{15}{Sydney Institute for Astronomy, School of Physics, The University of Sydney, NSW 2006, Australia}}
+\def\USydneytxt{\altaffiltext{15}{The University of Sydney, Sydney Institute for Astronomy, School of Physics, NSW 2006, Australia}}
 

### 2014-10-28 foreground_delay_spectrum_modeling.tex
Modified caption of figure showing baseline screening and minor changes

@@
 \def\ASU{\altaffilmark{1}}
-\def\ASUtxt{\altaffiltext{1}{Arizona State University, School of Earth and Space Exploration, Tempe, AZ 85287, USA; e-mail: t\_nithyanandan@asu.edu}}
+\def\ASUtxt{\altaffiltext{1}{Arizona State University, School of Earth and Space Exploration, Tempe, AZ 85287, USA}}
+
+\def\myemail{\altaffilmark{*}}
+\def\myemailtxt{\altaffiltext{*}{e-mail: t\_nithyanandan@asu.edu}}
 
@@
 %% Lead Authors
-Nithyanandan~Thyagarajan\ASU,
+Nithyanandan~Thyagarajan\ASU\myemail,
 Daniel~C.~Jacobs\ASU,
@@
 \NRAOtxt
+\myemailtxt
 
@@
 These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space.
-Based on these results, we propose that selective down--weighting of baselines based on length, direction, and time will mitigate foreground contamination substantially (up to a factor $\sim 10$ with almost no loss of data) in redshifted 21~cm power spectrum analysis.
+Based on these results, we propose that selective down--weighting of baselines based on length, direction, and time will mitigate foreground contamination substantially (up to a factor $\sim 10$ with negligible loss of sensitivity) in redshifted 21~cm power spectrum analysis.
 
@@
 
-The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off12,off10}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz.
+The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off10,off12}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz.
 These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey (MWACS; Hurley-Walker et al.
@@
 The degree to which these objects leak power into spectral line of sight modes in the {\it EoR window} will be shown to depend primarily on the brightness of foregrounds near the boundaries of the {\it foreground wedge}.
-Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view.
+Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view (a point also supported by Pober~et~al.~(2014), in prep.).
 
@@
 For the diffuse component, we use an all--sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz.
-At this frequency, since this map is predominantly based on the 408~MHz all--sky map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz map to the same resolution.
+At this frequency, since this map is predominantly based on the 408~MHz map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz map to the same resolution.
 However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals.
@@
 
-With the aforementioned all--sky foreground model, and instrumental and observational parameters, we simulate visibilities using equation~\ref{eqn:obsvis}.
+With the aforementioned foreground model, and instrumental and observational parameters, we simulate visibilities using equation~\ref{eqn:obsvis}.
 Figure~\ref{fig:sim_data} shows the amplitude of the delay spectrum from {\it off--zenith} and {\it zenith} pointings.
@@
 
-Having established that the simulation matches the data to the level of expected uncertainties, we proceed to examine in further detail the key signatures seen in simulated delay spectra.
+Having shown that the simulation matches the data to the level of expected uncertainties, we proceed to examine in further detail the key signatures seen in simulated delay spectra.
 A number of factors are responsible for the characteristics noted in the delay spectra obtained from data and through simulations.
@@
 
-Delay spectra from the all--sky foreground model in our study display a composite feature set drawn from the features of compact and diffuse foreground models.
-Here we compare the relative strengths of emission from different spatial scales in our all--sky composite foreground model.
+Delay spectra from the foreground model in our study display a composite feature set drawn from the features of compact and diffuse foreground models.
+Here we compare the relative strengths of emission from different spatial scales in our composite foreground model.
 
@@
 \includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_nside_64_Tsys_95.0K.eps}
-\caption{Foreground contamination, estimated using standard deviation in the {\it EoR window}, for the {\it off--zenith} pointing as a function of orientation and length of discarded antenna spacings.
-Antenna spacings oriented eastward with varying degrees of directedness are considered: -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars).
-Among antenna spacings that satisfy the above criteria, ones with lengths $\le |\vec{b}|_\textrm{max}$ ($x$--axis) are discarded.
-The fraction of discarded antenna spacings relative to the total number is denoted by dotted lines for the corresponding cases.
-% Lower (dashed line) and upper (dot--dashed) limits on foreground contamination are derived from northward (67\fdg 5~$\le\theta_\textrm{b}<$~112\fdg 5) and eastward ($|\theta_\textrm{b}|\le$~22\fdg 5) antenna spacings respectively.
-  Foreground contamination in the {\it EoR window} (solid lines) reduces by factor $\sim$1.5 ($|\theta_\textrm{b}|\le$~7\fdg 5) to $\sim$10 ($|\theta_\textrm{b}|\le$~22\fdg 5).
-Short baselines have maximum effect on foreground contamination.
-For instance, discarding antenna spacings oriented $|\theta_\textrm{b}|\le$~22\fdg 5 with lengths $|\vec{b}|\gtrsim 30$~m has no effect in further reducing foreground contamination.
+\caption{Drop in foreground contamination in the {\it EoR window}, and loss of data for the {\it off--zenith} pointing as a function of discarded baselines.
+Eastward baselines with varying degrees of directedness -- $|\theta_\textrm{b}|<$~7\fdg 5 (solid circles), $|\theta_\textrm{b}|<$~15\arcdeg~ (solid squares), and $|\theta_\textrm{b}|<$~22\fdg 5 (solid stars) -- and lengths $|\vec{b}| \le |\vec{b}|_\textrm{max}$ ($x$--axis) are discarded.
+Loss of data (dotted lines) is measured by discarded baselines as a fraction of the total number for the corresponding cases.
+Foreground contamination in the {\it EoR window} (solid lines) drops by a factor $\sim$1.5 ($|\theta_\textrm{b}|\le$~7\fdg 5) to $\sim$10 ($|\theta_\textrm{b}|\le$~22\fdg 5).
+The latter limit can be achieved with a mere 5\% loss of data at $|\vec{b}|_\textrm{max}\simeq30$~m, and discarding longer baselines ($|\vec{b}|_\textrm{max}\gtrsim 30$~m) has no effect in further reducing foreground contamination.
 \label{fig:screening}}
@@
 In general, delay spectrum signatures of compact emission are center--heavy, in clear contrast to those from diffuse emission.
-A composite all--sky foreground model consisting of diffuse and compact foregrounds combines the characteristic individual shapes into a {\it pitchfork}--shaped structure in the {\it foreground wedge}.
+A composite foreground model consisting of diffuse and compact foregrounds combines the characteristic individual shapes into a {\it pitchfork}--shaped structure in the {\it foreground wedge}.
 This will be distinctly visible when the thermal noise floor is sufficiently lowered, as longer observations are processed through the MWA analysis pipelines.

### 2014-10-28 foreground_delay_spectrum_modeling.tex
Minor changes in wording of impact of baseline screening

@@
 This screening technique can be generalized to optimize between foreground mitigation and loss of sensitivity from discarding data.
-Figure~\ref{fig:screening} shows how the typical foreground contamination\footnote{We use standard deviation of noiseless $V_{b\tau}(\vec{b},\tau)$ from foregrounds in the {\it EoR window} as a measure of foreground contamination.} in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
+Figure~\ref{fig:screening} shows how the typical foreground contamination\footnote{Foreground contamination is measured by standard deviation of noiseless $V_{b\tau}(\vec{b},\tau)$ from foregrounds in the {\it EoR window}.} in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
 We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars).
@@
 It is seen that foreground contamination can be mitigated by a factor between $\sim 1.5$ ($|\theta_\textrm{b}|\le$~7\fdg 5) and $\sim 10$ ($|\theta_\textrm{b}|\le$~22\fdg 5).
-It is notable that short antenna spacings play maximum role in determining foreground contamination.
-For instance, among antenna spacings with orientations $|\theta_\textrm{b}|\le$~22\fdg 5, discarding those with lengths $|\vec{b}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity as the fraction of discarded baselines increases from $\sim 5$\% to $\sim 25$\%.
+The latter limit is achieved with a mere 5\% loss of data for $|\vec{b}|_\textrm{max}\simeq30$~m.
+And discarding antenna spacings with lengths $|\vec{b}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity as the fraction of discarded baselines increases from $\sim 5$\% to $\sim 25$\%.
 On the other hand, a bright point source at the same location will give rise to foreground contamination even on longer antenna spacings.
-Such cases will necessitate discarding more or, in the worst case, all of the eastward oriented antenna spacings accounting for $\sim 25$\% loss of data.
+Such cases will necessitate discarding more or, in the worst case, all of the eastward oriented antenna spacings.
 
@@
 Foreground contamination in the {\it EoR window} (solid lines) drops by a factor $\sim$1.5 ($|\theta_\textrm{b}|\le$~7\fdg 5) to $\sim$10 ($|\theta_\textrm{b}|\le$~22\fdg 5).
-The latter limit can be achieved with a mere 5\% loss of data at $|\vec{b}|_\textrm{max}\simeq30$~m, and discarding longer baselines ($|\vec{b}|_\textrm{max}\gtrsim 30$~m) has no effect in further reducing foreground contamination.
+The latter limit can be achieved with a mere 5\% loss of data at $|\vec{b}|_\textrm{max}\simeq30$~m, and discarding longer baselines ($|\vec{b}|\gtrsim 30$~m) has no effect in further reducing foreground contamination.
 \label{fig:screening}}
@@
 In conclusion, we find that inclusion of emission models, both diffuse and compact, all the way to the horizon is essential to explaining the observed power spectrum.
-We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination by an order of magnitude in EoR data analysis by discarding or down--weighting antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation, without significant loss of data.
+We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination by an order of magnitude in EoR data analysis by discarding or down--weighting antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation, with negligible loss of sensitivity.
 

### 2014-10-29 foreground_delay_spectrum_modeling.tex
Added Chen Wu as an author with UWA affiliation

@@
 
+\def\UWA{\altaffilmark{20}}
+\def\UWAtxt{\altaffiltext{20}{International Centre for Radio Astronomy Research, University of Western Australia, Crawley, WA 6009, Australia}}
+
 %% \definenote[thanks][conversion=set 2]
@@
 C.~L.~Williams\MIT,
+C.~Wu\UWA,
 J.~S.~B.~Wyithe\UMelbourne$^,$\CAASTRO
@@
 \NRAOtxt
+\UWAtxt
 \myemailtxt

### 2014-10-30 foreground_delay_spectrum_modeling.tex
Changed Josh's listed name
Added a citation to MWACS paper in eor.bib and changed the citation in text

@@
 % A.~A.~Deshpande\RRI, 
-A.~de~Oliveira--Costa\MIT,
-J.~S.~Dillon\MIT,
+A.~de~Oliveira-Costa\MIT,
+Joshua~S.~Dillon\MIT,
 D.~Emrich\Curtin,
 % B.~M.~Gaensler\USydney$^,$\CAASTRO, 
-A.~Ewall--Wice\MIT,
+A.~Ewall-Wice\MIT,
 L.~Feng\MIT,
@@
 J.~N.~Hewitt\MIT,
-N.~Hurley--Walker\Curtin,
+N.~Hurley-Walker\Curtin,
 M.~Johnston-Hollitt\Victoria,
@@
 J.~C.~Kasper\UMichigan$^,$\CfA, 
-Han--Seek Kim\UMelbourne$^,$\CAASTRO,
+Han-Seek Kim\UMelbourne$^,$\CAASTRO,
 P.~Kittiwisit\ASU,
@@
 The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off10,off12}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz.
-These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey (MWACS; Hurley-Walker et al.
-2014, submitted).
+These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey \citep[MWACS;][]{hur14}.
 % MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning;

### 2014-11-14 ms.tex
Renamed the tex, ps, and pdf files to ms.tex, ms.ps and ms.pdf respectively

--- 2014-10-30 foreground_delay_spectrum_modeling.tex
@@
 These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space.
-Based on these results, we propose that selective down--weighting of baselines based on length, direction, and time will mitigate foreground contamination substantially (up to a factor $\sim 10$ with negligible loss of sensitivity) in redshifted 21~cm power spectrum analysis.
+Based on these results, we propose that selective down--weighting of data based on antenna spacing vector and time will mitigate foreground contamination substantially (up to a factor $\sim 10$ with negligible loss of sensitivity) in redshifted 21~cm power spectrum analysis.
 
@@
 
-\keywords{large-scale structure of Universe --- methods: statistical --- radio continuum: galaxies --- radio lines: general --- reionization --- techniques: interferometric}
+\keywords{cosmology: observations --- large-scale structure of universe --- methods: statistical --- radio continuum: galaxies --- reionization --- techniques: interferometric}
 
@@
 We report two important findings: the foregrounds that most severely obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted.
-% We find that the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon.
-We find that diffuse Galactic emission plays a larger role then previously thought.
 We quantify this contamination by separating the simulations into components based on type of foreground emission.
@@
 
-% Hence, we have identified a region of high significance (shown in medium shade) that forms a subset of the {\it EoR window} bounded by $|\tau|\le 0.78\,\mu$s where contamination from foregrounds is expected to be less and yet the EoR H{\sc i} power spectrum signal is reasonably strong.
-Hereafter, we refer to this region as the {\it EoR window} in the rest of this paper.
-
 In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\vec{k}|$.
@@
 \end{figure}
-
-% The terms antenna spacing and baseline are synonymous in radio interferometry.
-The latter term is also used broadly to refer to the antenna pair system in a radio interferometer array.
-We use these terms interchangeably in this paper.
 
@@
 
-% The bandpass properties used in this paper are illustrated in Figure~\ref{fig:bandpass}.
-The MWA bandpass of width 30.72~MHz is divided coarsely into 24 1.28~MHz chunks (dashed lines connecting solid circles) with each chunk weighted by a digital filter.
-The coarse channel shape is obtained using a 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$.
-The bandpass correction ({\it plus} symbols connected by dotted lines) corrects for the shape of the coarse channels.
-After this correction, the channels at the edges of these chunks have been flagged (solid lines) because they are known to possibly be contaminated by aliasing at a low level.
-Any spectral flags, or other known weights constitutes $W_f(f)$ in equation~\ref{eqn:obsvis}.
-
-% \begin{figure}[htb]
-% \centering
-% \includegraphics[width=\linewidth]{figures/v1_0/bandpass_properties.eps}
-% \caption{Bandpass properties used in modeling and analysis of delay spectrum of MWA visibilities.
-The filled circles joined by dotted lines show the 8--tap PFB bandpass shapes obtained using a Kaiser window with parameter $\beta=5$.
-The bandpass correction factors applied are shown ({\it plus} symbols connected by dashed lines).
-The solid line shows the flagged (gaps) and gain--corrected bandpass shape.
-These shapes are repeated every 1.28~MHz for the entire bandwidth of 30.72~MHz centered around 185~MHz.
-\label{fig:bandpass}}
-% \end{figure}
-
 Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=95$~K per polarization for all frequency channels for all antenna pairs throughout the course of the observation.
@@
 In our work, we have chosen a $T_\textrm{sys}$ that matches the thermal noise observed in the data.
-% We take into account the increase in thermal uncertainty in frequency channels that results due to aforementioned bandpass correction.
-({\bf Since the bandpass correction is digital, shouldn't the noise stay flat after bandpass correction?
--- JDB: yes})
 
@@
 The MWA targets two primary low-foreground fields for reionization observations.
-Here, we focus on the field at RA = 0h, Dec = $-30$\arcdeg.
+Here, we focus on the field at RA = 0$^\textrm{h}$, Dec = $-30$\arcdeg.
 The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of the dipoles in an MWA tile.
@@
 
-The observations shown here were undertaken on Aug~23, 2013.
+The observations shown here were undertaken on 2013~August~23.
 % Data set from this day has been chosen as a reference for comparison within the MWA EoR collaboration because this is one of the days featuring a low number of known system errors from among a long stretch of stable observing.
@@
 The {\it local sidereal time} (LST) at these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at $-1.92$~hours (i.e., 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings, respectively.
-%These pointings (phased array power--patterns) are centered at RA = 4\fdg 387, Dec = -29\fdg 86 and RA = -28\fdg 8, Dec = -26\fdg 701, whereas the visibilities themselves are phased to zenith corresponding to RA = -28\fdg 8, Dec = -26\fdg 701, and RA = 1\fdg 35, Dec = -26\fdg 701 respectively.
 
@@
 The RA, HA, and declination are plotted with solid, dashed and dotted lines respectively.
-The sky is allowed to drift for $\sim 30$~mins before the beamformer delay settings change in a discrete step to center the beam around RA = 0h, Dec = $-30$\arcdeg.
+The sky is allowed to drift for $\sim 30$~mins before the beamformer delay settings change in a discrete step to center the beam around RA = 0$^\textrm{h}$, Dec = $-30$\arcdeg.
 Dot--dashed vertical lines show the LST of {\it off-zenith} and {\it zenith} pointings used in our study.
@@
 These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey \citep[MWACS;][]{hur14}.
-% MWACS was compiled using observations taken with subsets of the array during the final buildout and commissioning;
 It has a flux density limit of 25~mJy and a declination range of -12\arcdeg~ to -40\arcdeg~ evenly covering the field of view of the observations reported here.
@@
 \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
-\caption{Delay spectra of two observations of the EoR field (RA = 0h, Dec = $-30$\arcdeg) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
+\caption{Delay spectra of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30$\arcdeg) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
 For reference, the $k$--space equivalent of $\tau$ and $|\vec{b}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively.
@@
 \end{figure}
-
-% Despite delay deconvolution, we still see leakage beyond the cleaned regions indicating the inability of the deconvolution algorithm to fit the measurements perfectly, dominant at short antenna spacings where the number of delay modes within the horizon is lesser.
-This is because the maximum delay envelope (boundary of the wedge) consists of mixed emission from large portions of the sky into a narrow range of delays and the deconvolution algorithm does not have sufficient support in delay space to fit the delay spectrum.
-However, we wish to emphasize that the purpose of deconvolution is to rid the delay spectra of instrumental artifacts to the best extent possible in order to see the underlying signatures of foreground emission.
-It is not our intention in this paper to use the deconvolution as a foreground removal tool.
 
@@
 The degree to which these objects leak power into spectral line of sight modes in the {\it EoR window} will be shown to depend primarily on the brightness of foregrounds near the boundaries of the {\it foreground wedge}.
-Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view (a point also supported by Pober~et~al.~(2014), in prep.).
-
-% JDB: I don't think we need this here.
-%% \citet{bea13} estimated the sensitivity of the MWA to EoR H{\sc i} power spectrum detection taking into account thermal noise effects in the {\it EoR window}.
-\citet{thy13} estimated the sensitivity taking into account the spillover from foreground contamination from unsubtracted extragalactic compact objects in the {\it EoR window} besides thermal noise.
-In this paper, we include the diffuse Galactic emission for enhancing our understanding of foreground signatures in the power spectrum.
+Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view, a point also supported by Pober~et~al.~(2014, in preparation).
 
@@
 
-% \item {\it Instrumental Passband}: The instrumental passband characteristics have a direct effect on the delay transform owing to the Fourier relationship between the two.
-Due to excision of aliased edge channels repeated in each coarse channel of our passband, we observed repeated patterns of the {\it foreground wedge} in the delay spectra.
-As already mentioned earlier, we have used a deconvolution algorithm ({\it CLEAN}) to rid the delay spectra of band shape effects.
-Since this paper focuses on studying the effects of foreground on the delay spectra, we do not pursue the characterization of artifacts from the band shape and imperfections in {\it CLEAN} deconvolution.
-
 \end{itemize}
@@
 It decreases in strength with increasing baseline length as expected but is present in all baseline orientations.
-%% This contributes to a {\it pitchfork}--shaped feature which will be discussed later in \S\ref{sec:composite} in more detail.
 
@@
 
-% While the {\it foreground wedge} is a region of high foreground contamination, of particular interest to the EoR community is foreground contamination near horizon delay limits.
-These regions are believed to be responsible for significant spillover into the relatively clean {\it EoR window} due to spectral response of the instrument and foregrounds.
-
 Here, we will investigate the susceptibility of particular antenna spacings to foreground contamination arising out of bright foreground objects located near the horizon and present a technique to substantially mitigate such contamination.
@@
 This technique provides a very simple and yet effective tool in adding a layer of control to mitigate effects of foreground contamination in EoR data analysis.
-% The success behind clear identification of severely foreground--contaminated antenna pairs based on delay spectrum technique depends on the localization of the source of foreground contamination and its strength.
 
@@
 In conclusion, we find that inclusion of emission models, both diffuse and compact, all the way to the horizon is essential to explaining the observed power spectrum.
-We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination by an order of magnitude in EoR data analysis by discarding or down--weighting antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation, with negligible loss of sensitivity.
+We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination by an order of magnitude in EoR data analysis by discarding or down--weighting data from antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation, with negligible loss of sensitivity.
 

### 2014-11-14 ms.tex
Renamed figure file names, moved them to parent older and removed separate figures folder

@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/fourier_space_185.0_MHz_30.7_MHz.eps}
+% \includegraphics[width=\linewidth]{figures/v1_0/fourier_space_185.0_MHz_30.7_MHz.eps}
+\includegraphics[width=\linewidth]{f1.eps}
 \caption{Fourier space in which delay (and power) spectra of EoR H{\sc i} signals are presented.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/custom_pointings.eps}
+% \includegraphics[width=\linewidth]{figures/v1_0/custom_pointings.eps}
+\includegraphics[width=\linewidth]{f2.eps}
 \caption{MWA tile beam pointing directions during the course of the observation.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
+% \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
+\includegraphics[width=\linewidth]{f3.eps}
 \caption{Delay spectra of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30$\arcdeg) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom.eps}
+% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom.eps}
+\includegraphics[width=\linewidth]{f4.eps}
 \caption{Models of MWA tile power--pattern for {\it off--zenith} (top) and {\it zenith} pointings (bottom) at 185~MHz.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/dsm.eps}
+% \includegraphics[width=\linewidth]{figures/v1_0/dsm.eps}
+\includegraphics[width=\linewidth]{f5.eps}
 \caption{Sky brightness temperature of the diffuse foreground model at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/csm.eps}
+% \includegraphics[width=\linewidth]{figures/v1_0/csm.eps}
+\includegraphics[width=\linewidth]{f6.eps}
 \caption{Flux densities of bright point sources at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
+% \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
+\includegraphics[width=\linewidth]{f7.eps}
 \caption{Same as Figure~\ref{fig:fhd_data}, but generated from simulated visibilities using the model described in \S\ref{sec:modeling}.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
+% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
+\includegraphics[width=\linewidth]{f8.eps}
 \caption{Simulated delay spectrum amplitudes $|V_{b\tau}(\vec{b},\tau)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
+% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
+\includegraphics[width=\linewidth]{f9.eps}
 \caption{Same as Figure~\ref{fig:noiseless-dsm-delay-spectrum} but for a foreground model consisting of compact emission.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
+% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
+\includegraphics[width=\linewidth]{f10.eps}
 \caption{Simulated delay spectrum amplitudes $|V_{b\tau}(\vec{b},\tau)|$ for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}).
@@
 \centering
-\subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{figures/v1_0/sky_model_snapshot_0.eps}}
-\subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{figures/v1_0/directional_delay_map_snapshot_0.eps}}
-\subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figures/v1_0/baseline_orientation_breakup_snapshot_0.eps}}
+% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{figures/v1_0/sky_model_snapshot_0.eps}}
+% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{figures/v1_0/directional_delay_map_snapshot_0.eps}}
+% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figures/v1_0/baseline_orientation_breakup_snapshot_0.eps}}
+\subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{f11a.eps}}
+\subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{f11b.eps}}
+\subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{f11c.eps}}
 \caption{(a): Sky model showing compact ({\it top}) and diffuse ({\it bottom}) emission.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_before_after_nside_64_Tsys_95.0K.eps}
+% \includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_before_after_nside_64_Tsys_95.0K.eps}
+\includegraphics[width=\linewidth]{f12.eps}
 \caption{Simulated delay spectra amplitudes $|V_{b\tau}(\vec{b},\tau)|$ for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
@@
 \centering
-\includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_nside_64_Tsys_95.0K.eps}
+% \includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_nside_64_Tsys_95.0K.eps}
+\includegraphics[width=\linewidth]{f13.eps}
 \caption{Drop in foreground contamination in the {\it EoR window}, and loss of data for the {\it off--zenith} pointing as a function of discarded baselines.

### 2015-01-03 ms.tex
Minor changes to notations, citations, and wording. Added a file addressing referee comments

@@
 We investigate the chromatic signatures of an all--sky foreground model imprinted by an instrumental transfer function.
-Using the delay spectrum technique, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view.
-Comparing data from recent Murchison Widefield Array observations with simulations separated into components based on type of emission, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings.
+Using the delay spectrum technique {\bf that maps the Fourier transform of the measurements to locations of foreground objects on the celestial sphere through delays in signals received by antenna pairs}, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view.
+Comparing data from recent Murchison Widefield Array observations with simulations separated into components based on type of emission, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings {\bf which are usually not sensitive to diffuse emission}.
 For signals entering through the primary field of view, compact emission dominates the foreground contamination.
@@
 Sensitive instruments such as the Square Kilometre Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}.
-Numerous precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{tin13,bow13,lon09}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{bow06,bea13,dil13,thy13,pob14}.
+Numerous {\bf pathfinders and} precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{tin13,bow13,lon09}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{bow06,par12a,bea13,dil13,thy13,pob14}.
+{\bf The Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} (HERA) is currently under construction using new insights gained with the MWA and PAPER}.
 
 A key challenge in the statistical detection of the redshifted H{\sc i} 21~cm signal, via the spatial power spectrum of temperature fluctuations, arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06,ali08,ber09,ber10,gho12}.
-\citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}.
+\citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12b,dil13,pob13}.
 
 Considerable effort has been and continues to be made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,dil14,liu14a,liu14b}.
-A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide--field (and chromatic) response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum.
+{\bf A model that provides a generic explanation for the observed foreground power spectrum has emerged, whereby the wide--field (and chromatic) response of the instrument causes the power in smooth spectrum foregrounds to occupy higher $k$--modes into the so-called `wedge'}.
+% A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide--field (and chromatic) response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum.
 The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}).
@@
 Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity.
-To decouple these effects from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the individual baseline--based approach of \citet{pob13} and \citet{par14}.
+To decouple these effects from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using a {\it \textbf{Per-Baseline}} approach of \citet{par12b}.
 This approximates the power spectrum as the inverse Fourier transform of the spectra generated by the instrument's correlator.
 
-After matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the signature of the power spectrum of each baseline.
-We report two important findings: the foregrounds that most severely obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted.
+After matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the {\bf shape of the delay spectrum} of each baseline.
+We report two important findings: the foregrounds that most severely {\bf obscure} the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted.
 We quantify this contamination by separating the simulations into components based on type of foreground emission.
@@
 
-In \S\ref{sec:delay-spectrum} we introduce the delay spectrum technique.
+In \S\ref{sec:delay-spectrum} we {\bf provide a brief overview of the delay spectrum technique} \citep{par12a,par12b}.
 In \S\ref{sec:instrument}, we describe the MWA setup, summarize the observing parameters, and present the resulting data.
@@
 
-Interferometer array data known as {\it visibilities}, $V_{bf}(\vec{b},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\vec{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis.
-If $I(\hat{s},f)$ and $A(\hat{s},f)$ are the sky brightness and antenna's directional power--pattern, respectively, at different frequencies on the sky as a function of direction on the sky denoted by the unit vector ($\hat{s}$), and $W_f(f)$ denotes instrumental bandpass weights, then $V_{bf}(\vec{b},f)$ can be written as (with a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}):
+Interferometer array data known as {\it visibilities}, $V_{bf}\equiv V_{bf}(\boldsymbol{b},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis.
+If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power--pattern, respectively, at different frequencies on the sky as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_f\equiv W_f(f)$ denotes instrumental bandpass weights, then $V_{bf}$ can be written as (with a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}):
 \begin{align}\label{eqn:obsvis}
-  V_{bf}(\vec{b},f) &= \iint\limits_\textrm{sky} A(\hat{s},f)\,I(\hat{s},f)\,W_f(f)\,e^{-i2\pi f\frac{\vec{b}\cdot\hat{s}}{c}}\,\dif\Omega,
+  V_{bf} &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_f\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif\Omega,
 \end{align}
-where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element along $\hat{s}$. We wish to emphasize that this equation is valid in general and does not involve any approximation.
+where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector.
+We wish to {\bf note} that this equation is valid {\bf even for wide--field applications where the sky cannot be approximated to be flat}.
 
-We define the {\it delay spectrum}, $V_{b\tau}(\vec{b},\tau)$, to be the inverse Fourier transform of $V_{bf}(\vec{b},f)$ along the frequency coordinate:
+We define the {\it delay spectrum}, $V_{b\tau}\equiv V_{b\tau}(\boldsymbol{b},\tau)$, to be the inverse Fourier transform of $V_{bf}$ along the frequency coordinate:
 \begin{align}\label{eqn:delay-transform}
-  V_{b\tau}(\vec{b},\tau) &= \int V_{bf}(\vec{b},f)\,W'_f(f)\,e^{i2\pi f\tau}\,\dif f,
+  V_{b\tau} &\equiv \int V_{bf}\,W'_f\,e^{i2\pi f\tau}\,\dif f,
 \end{align}
-where, $W'_f(f)$ is a spectral weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}, and $\tau$ represents the signal delay between antenna pairs:
+where, $W'_f\equiv W'_f(f)$ is a spectral weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}, and $\tau$ represents the signal delay between antenna pairs:
 \begin{equation}\label{eqn:delay}
-  \tau = \frac{\vec{b}\cdot\hat{s}}{c}.
+  \tau = \frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}.
 \end{equation}
 
-Equation~\ref{eqn:obsvis} can be modified to express the Fourier decomposition of transverse sky brightness distribution as:
+Equation~\ref{eqn:obsvis} can be {\bf equivalently re-expressed} as:
 \begin{align}\label{eqn:vis}
-  V_{uf}(\vec{u},f) &= \iint\limits_\textrm{sky} A(\hat{s},f)\,I(\hat{s},f)\,W_f(f)\,e^{-i2\pi\vec{u}\cdot\hat{s}}\,\dif\Omega,
+  V_{uf} &\equiv V_{uf}(\boldsymbol{u},f) \nonumber\\
+  &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_f\,e^{-i2\pi\boldsymbol{u}\cdot\hat{\boldsymbol{s}}}\,\dif\Omega,
 \end{align}
-where, $\vec{u}$ denotes the transverse spatial frequency vector.
-$\vec{u}$ is directly related to $\vec{b}$ as $\vec{u}\equiv \vec{b}/\lambda$, and to the transverse spatial wavenumber mode, $\vec{k}_\perp$, as $\vec{k}_\perp\equiv 2\pi\vec{u}/D(z)$, where, $D(z)$ is the transverse comoving distance at redshift $z$.
-Hence, $V_{uf}(\vec{u},f)\equiv V_{uf}(\vec{b}/\lambda,f)\equiv V_{bf}(\lambda\vec{u},f)$.
-It is important to note that $\vec{b}$ is independent of frequency while $\vec{u}$ is not.
+where, $\boldsymbol{u}$ denotes the transverse spatial frequency vector.
+$\boldsymbol{u}$ is directly related to $\boldsymbol{b}$ as $\boldsymbol{u}\equiv \boldsymbol{b}/\lambda$, and to the transverse spatial wavenumber mode, $\boldsymbol{k}_\perp$, as $\boldsymbol{k}_\perp\equiv 2\pi\boldsymbol{u}/D(z)$, where, $D(z)$ is the transverse comoving distance at redshift $z$.
+Hence, $V_{uf}=V_{uf}(\boldsymbol{b}/\lambda,f)=V_{bf}(\lambda\boldsymbol{u},f)$.
+It is important to note that for a given antenna pair, $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not.
 
@@
 It is directly related to the line--of--sight wavenumber, $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
-This approximation holds under the assumption that there is negligible cosmological evolution within the redshift range (or frequency band).
+This approximation holds under the assumption that the redshift range (or frequency band) is small enough within which cosmological evolution is negligible.
 Thus,
 \begin{align}\label{eqn:los-transform}
-  V_{u\eta}(\vec{u},\eta) &= \int V_{uf}(\vec{u},f)\,W'_f(f)\,e^{i2\pi f\eta}\,\dif f
+  V_{u\eta} &\equiv V_{u\eta}(\boldsymbol{u},\eta) \nonumber\\
+  &\equiv \int V_{uf}(\boldsymbol{u},f)\,W'_f\,e^{i2\pi f\eta}\,\dif f
 \end{align}
@@
 This representation has been discussed in detail in \citet{mor04}.
-The spatial power spectrum of EoR H{\sc i} distribution, $P(\vec{k}_\perp,k_\parallel)$, and $V_{u\eta}(\vec{u},\eta)$ are related by $P(\vec{k}_\perp,k_\parallel)\propto |V_{u\eta}(\vec{u},\eta)|^2$.
+The spatial power spectrum of EoR H{\sc i} distribution, $P(\boldsymbol{k}_\perp,k_\parallel)$, and $V_{u\eta}$ are related by $P(\boldsymbol{k}_\perp,k_\parallel)\propto |V_{u\eta}|^2$.
 
-While $V_{b\tau}(\vec{b},\tau)$ and $V_{u\eta}(\vec{u},\eta)$ in equations~\ref{eqn:obsvis} and \ref{eqn:vis} resemble each other very closely, they are not quite identical.
-The difference basically arises from the fact noted above that $\vec{b}$ is independent of frequency while $\vec{u}$ is not.
-\citet{par12} and \citet{liu14a} have discussed the mathematical correspondence between the two.
+While $V_{b\tau}$ and $V_{u\eta}$ in equations~\ref{eqn:obsvis} and \ref{eqn:vis} resemble each other very closely, they are not quite identical.
+The difference basically arises from the fact noted above that $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not.
+\citet{par12b} and \citet{liu14a} have discussed the mathematical correspondence between the two.
 
@@
 
-In summary, the delay spectrum, $V_{b\tau}(\vec{b},\tau)$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}.
-$V_{b\tau}(\vec{b},\tau)$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument.
-At the same time, it is closely related to the sought power $|V_{u\eta}(\vec{u},\eta)|^2$ in Fourier modes containing critical information about spatial scales.
-For purposes of illustration later in this paper, we often use the quantity, $|V_{b\tau}(\vec{b},\tau)|$.
+In summary, the delay spectrum, $V_{b\tau}$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}.
+$V_{b\tau}$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument.
+At the same time, it is closely related to the sought power $|V_{u\eta}|^2$ in Fourier modes containing critical information about spatial scales.
+For purposes of illustration later in this paper, we often use the quantity, $|V_{b\tau}|$.
 
@@
 Negative and positive delay\footnote{The terms delay and lag are used interchangeably in this paper.} correspond to the two hemispheres of the sky as transected by the antenna spacing vector.
-In this paper, $\vec{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the telescope site.
+In this paper, $\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the telescope site.
 Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at positive and negative delays, respectively.
@@
 Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of H{\sc i} from the EoR are presented in this paper.
-$|\vec{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction, form the $x$--axis.
+$|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction, form the $x$--axis.
 $\tau\approx\eta\propto k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis.
@@
 
-In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\vec{k}|$.
+In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\boldsymbol{k}|$.
 Hence, the brightest EoR signal will be observed on the shortest baselines and smallest delays.
@@
 \caption{Fourier space in which delay (and power) spectra of EoR H{\sc i} signals are presented.
-The $x$--axis is denoted by $|\vec{b}|$ (antenna spacing) or $k_\perp$ (transverse wavenumber).
+The $x$--axis is denoted by $|\boldsymbol{b}|$ (antenna spacing) or $k_\perp$ (transverse wavenumber).
 The $y$--axis denoted by $\tau$ (delay) or $k_\parallel$ (line--of--sight wavenumber).
@@
 The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km \citep{bea12}.
-Here we focus on antenna spacings $|\vec{b}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz.
+Here we focus on antenna spacings $|\boldsymbol{b}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz.
 The center frequency of 185~MHz corresponds to a redshift, $z\simeq 6.68$.
@@
 After correcting for the shape of these coarse channels, the fine channels (40~kHz width) at the edges of these sub--bands are flagged because they are known to be contaminated by aliasing at a low level.
-Any spectral flags, or other known weights constitutes $W_f(f)$ in equation~\ref{eqn:obsvis}.
+Any spectral flags, or other known weights constitutes $W_f$ in equation~\ref{eqn:obsvis}.
 
@@
 
-We obtain the delay spectrum of these calibrated visibilities by taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f(f)$ to be a {\it Blackman--Harris} window function.
+We obtain the delay spectrum of these calibrated visibilities by taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f$ to be a {\it Blackman--Harris} window function.
 The spectrum is multiplied by the additional weight of the flagged channels which occur rarely for interference and regularly every 1.28~MHz where the edges of coarse channels are known to have a small aliasing effect.
 In delay--space, these weights translate into a convolution by a grated point spread function (PSF) with gratings at multiples of $\tau=0.78\,\mu$s.
-We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis by \citet{par09,par12}.
+We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis by \citet{par09,par12b}.
 The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights.
@@
 
-In subsequent sections, we plot the amplitude of the delay spectrum, $|V_{b\tau}(\vec{b},\tau)|$, sorted by baseline length, $|\vec{b}|$.
+In subsequent sections, we plot the amplitude of the delay spectrum, $|V_{b\tau}|$, sorted by baseline length, $|\boldsymbol{b}|$.
 Note that visibilities from different baselines have not been averaged together in this analysis.
@@
 \caption{Delay spectra of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30$\arcdeg) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
-For reference, the $k$--space equivalent of $\tau$ and $|\vec{b}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively.
+For reference, the $k$--space equivalent of $\tau$ and $|\boldsymbol{b}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively.
 White lines mark the boundaries of {\it foreground wedge} determined by light travel time across each baseline.
@@
 
-The instrument model consists primarily of the tile beam power--pattern, $A(\hat{s},f)$ (see equation~\ref{eqn:obsvis}).
+The instrument model consists primarily of the tile beam power--pattern, $A(\hat{\boldsymbol{s}},f)$ (see equation~\ref{eqn:obsvis}).
 It is modeled as a mutually--coupled 4--by--4 dipole array with the overall power--pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14}.
@@
 In addition to intrinsic model uncertainty, delay spectra from simulations and data each have fluctuations due to thermal noise in the delay spectrum with rms $\sim$ 1.4$\times 10^7$~Jy~Hz.
-We estimate the ratio of delay spectra from data and simulations as $\rho = |V^\textrm{D}_{b\tau}(\vec{b},\tau)|\,/\,|V^\textrm{S}_{b\tau}(\vec{b},\tau)|$, where superscripts D and S denote data and simulation, respectively.
+We estimate the ratio of delay spectra from data and simulations as $\rho = |V^\textrm{D}_{b\tau}(\boldsymbol{b},\tau)|\,/\,|V^\textrm{S}_{b\tau}(\boldsymbol{b},\tau)|$, where superscripts D and S denote data and simulation, respectively.
 
@@
 
-Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the amplitude of delay spectrum, $|V_{b\tau}(\vec{b},\tau)|$, simulated with only the diffuse foreground model.
+Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the amplitude of delay spectrum, $|V_{b\tau}|$, simulated with only the diffuse foreground model.
 The top and bottom panels correspond to {\it off--zenith} and {\it zenith} pointings, respectively.
@@
 \includegraphics[width=\linewidth]{f8.eps}
-\caption{Simulated delay spectrum amplitudes $|V_{b\tau}(\vec{b},\tau)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
+\caption{Simulated delay spectrum amplitudes $|V_{b\tau}|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
 White lines mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
@@
 As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inversely proportional to the length of the baseline projected in the direction of the emission.
-Thus, in the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at zero delay, fades away on antenna spacings $|\vec{b}| \gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0\fdg 75.
+Thus, in the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at zero delay, fades away on antenna spacings $|\boldsymbol{b}| \gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0\fdg 75.
 In the off-zenith pointing, we see power out to much longer baselines which is described below.
@@
 
-A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\vec{b}|\le$~200~m).
+A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\boldsymbol{b}|\le$~200~m).
 This is contrary to expectations because these longer baselines are not expected to be sensitive to diffuse emission (on scales $\gtrsim$~0\fdg 85) of our input foreground model.
@@
 \includegraphics[width=\linewidth]{f10.eps}
-\caption{Simulated delay spectrum amplitudes $|V_{b\tau}(\vec{b},\tau)|$ for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}).
+\caption{Simulated delay spectrum amplitudes $|V_{b\tau}|$ for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}).
 The antenna spacing is specified in each panel along East, North and Upward directions.
@@
 The bright Galactic center in the western sky is recorded at zero delays in northward antenna spacings and close to negative horizon delay limit on eastward antenna spacings.
-(c): Simulated delay spectra $|V_{b\tau}(\vec{b},\tau)|$ on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}).
+(c): Simulated delay spectra $|V_{b\tau}|$ on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}).
 White lines denote horizon delay limits.
@@
 
-The delay spectrum, $V_{b\tau}(\vec{b},\tau)$, not only carries information on spatial scales of foreground emission but also offers the unique advantage of viewing the sky through a combination of antenna spacing vectors and the delay axis.
+The delay spectrum, $V_{b\tau}$, not only carries information on spatial scales of foreground emission but also offers the unique advantage of viewing the sky through a combination of antenna spacing vectors and the delay axis.
 With a foreground model known {\it a priori} in which structures and locations of very bright foreground objects such as the Galactic center or AGN are available, we can predict their response across antenna spacings as a function of observing parameters such as LST, power--pattern, etc.
@@
 
-In our example, we discard eastward antenna spacings (-22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5) of lengths $|\vec{b}| < 30$~m.
+In our example, we discard eastward antenna spacings (-22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5) of lengths $|\boldsymbol{b}| < 30$~m.
 Figure~\ref{fig:before-after} shows the delay spectra obtained with all antenna spacings (top panel) and after applying our screening technique (bottom panel) on the {\it off--zenith} observation.
@@
 \includegraphics[width=\linewidth]{f12.eps}
-\caption{Simulated delay spectra amplitudes $|V_{b\tau}(\vec{b},\tau)|$ for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
-Discarded antenna spacings (black vertical stripes) have lengths $|\vec{b}|<30$~m (leftward of vertical dashed line) and orientations $|\theta_\textrm{b}|<$~22\fdg 5.
+\caption{Simulated delay spectra amplitudes $|V_{b\tau}|$ for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
+Discarded antenna spacings (black vertical stripes) have lengths $|\boldsymbol{b}|<30$~m (leftward of vertical dashed line) and orientations $|\theta_\textrm{b}|<$~22\fdg 5.
 The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} prominent when all antenna spacings are used is lowered by an order of magnitude when short eastward antenna spacings are discarded.
@@
 This screening technique can be generalized to optimize between foreground mitigation and loss of sensitivity from discarding data.
-Figure~\ref{fig:screening} shows how the typical foreground contamination\footnote{Foreground contamination is measured by standard deviation of noiseless $V_{b\tau}(\vec{b},\tau)$ from foregrounds in the {\it EoR window}.} in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
+Figure~\ref{fig:screening} shows how the typical foreground contamination\footnote{Foreground contamination is measured by standard deviation of noiseless $V_{b\tau}$ from foregrounds in the {\it EoR window}.} in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
 We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars).
-Among antenna spacings that satisfy these criteria, we discard data from those whose lengths are shorter than $|\vec{b}|_\textrm{max}$ ($x$--axis) and show foreground contamination estimated in the {\it EoR window} from all remaining antenna spacings.
+Among antenna spacings that satisfy these criteria, we discard data from those whose lengths are shorter than $|\boldsymbol{b}|_\textrm{max}$ ($x$--axis) and show foreground contamination estimated in the {\it EoR window} from all remaining antenna spacings.
 In other words, these plots demonstrate the progress in mitigation as orientation and maximum length of discarded antenna spacings are varied.
@@
 It is seen that foreground contamination can be mitigated by a factor between $\sim 1.5$ ($|\theta_\textrm{b}|\le$~7\fdg 5) and $\sim 10$ ($|\theta_\textrm{b}|\le$~22\fdg 5).
-The latter limit is achieved with a mere 5\% loss of data for $|\vec{b}|_\textrm{max}\simeq30$~m.
-And discarding antenna spacings with lengths $|\vec{b}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity as the fraction of discarded baselines increases from $\sim 5$\% to $\sim 25$\%.
+The latter limit is achieved with a mere 5\% loss of data for $|\boldsymbol{b}|_\textrm{max}\simeq30$~m.
+And discarding antenna spacings with lengths $|\boldsymbol{b}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity as the fraction of discarded baselines increases from $\sim 5$\% to $\sim 25$\%.
 On the other hand, a bright point source at the same location will give rise to foreground contamination even on longer antenna spacings.
@@
 \caption{Drop in foreground contamination in the {\it EoR window}, and loss of data for the {\it off--zenith} pointing as a function of discarded baselines.
-Eastward baselines with varying degrees of directedness -- $|\theta_\textrm{b}|<$~7\fdg 5 (solid circles), $|\theta_\textrm{b}|<$~15\arcdeg~ (solid squares), and $|\theta_\textrm{b}|<$~22\fdg 5 (solid stars) -- and lengths $|\vec{b}| \le |\vec{b}|_\textrm{max}$ ($x$--axis) are discarded.
+Eastward baselines with varying degrees of directedness -- $|\theta_\textrm{b}|<$~7\fdg 5 (solid circles), $|\theta_\textrm{b}|<$~15\arcdeg~ (solid squares), and $|\theta_\textrm{b}|<$~22\fdg 5 (solid stars) -- and lengths $|\boldsymbol{b}| \le |\boldsymbol{b}|_\textrm{max}$ ($x$--axis) are discarded.
 Loss of data (dotted lines) is measured by discarded baselines as a fraction of the total number for the corresponding cases.
 Foreground contamination in the {\it EoR window} (solid lines) drops by a factor $\sim$1.5 ($|\theta_\textrm{b}|\le$~7\fdg 5) to $\sim$10 ($|\theta_\textrm{b}|\le$~22\fdg 5).
-The latter limit can be achieved with a mere 5\% loss of data at $|\vec{b}|_\textrm{max}\simeq30$~m, and discarding longer baselines ($|\vec{b}|\gtrsim 30$~m) has no effect in further reducing foreground contamination.
+The latter limit can be achieved with a mere 5\% loss of data at $|\boldsymbol{b}|_\textrm{max}\simeq30$~m, and discarding longer baselines ($|\boldsymbol{b}|\gtrsim 30$~m) has no effect in further reducing foreground contamination.
 \label{fig:screening}}

### 2015-01-06 ms.tex
Restructured section 2 and minor changes addressing many referee comments

@@
 Sensitive instruments such as the Square Kilometre Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}.
-Numerous {\bf pathfinders and} precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{tin13,bow13,lon09}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{bow06,par12a,bea13,dil13,thy13,pob14}.
+Numerous {\bf pathfinders and} precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13,bow13}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{bow06,par12a,bea13,dil13,thy13,pob14}.
 {\bf The Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} (HERA) is currently under construction using new insights gained with the MWA and PAPER}.
@@
 \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12b,dil13,pob13}.
+{\bf Since this contamination is expected to be several orders of magnitude stronger than the underlying EoR H{\sc i} signal, it is critical to characterize foregrounds precisely in order to reduce their impact on EoR H{\sc i} power spectrum detection sensitivity}.
 
@@
 Interferometer array data known as {\it visibilities}, $V_{bf}\equiv V_{bf}(\boldsymbol{b},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis.
-If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power--pattern, respectively, at different frequencies on the sky as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_f\equiv W_f(f)$ denotes instrumental bandpass weights, then $V_{bf}$ can be written as (with a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}):
+If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power--pattern, respectively, at different frequencies as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_f\equiv W_f(f)$ denotes instrumental bandpass weights, then $V_{bf}$ can be written as (with a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}):
 \begin{align}\label{eqn:obsvis}
@@
 
-We define the {\it delay spectrum}, $V_{b\tau}\equiv V_{b\tau}(\boldsymbol{b},\tau)$, to be the inverse Fourier transform of $V_{bf}$ along the frequency coordinate:
+{\bf The {\it delay spectrum}, $V_{b\tau}\equiv V_{b\tau}(\boldsymbol{b},\tau)$, is defined as} the inverse Fourier transform of $V_{bf}$ along the frequency coordinate:
 \begin{align}\label{eqn:delay-transform}
@@
 \end{equation}
+$V_{b\tau}$ is expressed in observer's units as Jy~Hz.
+
+\subsection{Resemblance to Spatial Power Spectrum}\label{sec:power-spectrum}
 
@@
 \end{align}
-where, $\boldsymbol{u}$ denotes the transverse spatial frequency vector.
-$\boldsymbol{u}$ is directly related to $\boldsymbol{b}$ as $\boldsymbol{u}\equiv \boldsymbol{b}/\lambda$, and to the transverse spatial wavenumber mode, $\boldsymbol{k}_\perp$, as $\boldsymbol{k}_\perp\equiv 2\pi\boldsymbol{u}/D(z)$, where, $D(z)$ is the transverse comoving distance at redshift $z$.
-Hence, $V_{uf}=V_{uf}(\boldsymbol{b}/\lambda,f)=V_{bf}(\lambda\boldsymbol{u},f)$.
+where, $\hat{\boldsymbol{s}}$ is measured with reference to a location on the sky referred to as the {\it phase center}, and $\boldsymbol{u}\equiv (u,v,w)$ denotes the spatial frequency vector.
+$w$ is aligned parallel to the direction of the phase center, while $u$ and $v$ lie on the transverse plane perpendicular to it.
+For measurements that are constrained to lie on this plane, we can choose $w=0$ without loss of generality and $\boldsymbol{u}$ effectively reduces to $\boldsymbol{u}\equiv (u,v)$, a two--dimensional vector.
+Then, $\boldsymbol{u}$ is directly related to the transverse spatial wavenumber mode as:
+\begin{align}\label{eqn:k-perp}
+  \boldsymbol{k}_\perp &\equiv \frac{2\pi\boldsymbol{u}}{D}, 
+\end{align}
+where, $D\equiv D(z)$ is the transverse comoving distance at redshift $z$.
+% Hence, $V_{uf}=V_{uf}(\boldsymbol{b}/\lambda,f)=V_{bf}(\lambda\boldsymbol{u},f)$.
 It is important to note that for a given antenna pair, $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not.
@@
 $\eta$, which is the Fourier transform dual of $f$, is used to denote the spatial frequency along the line of sight and has units of time.
-It is directly related to the line--of--sight wavenumber, $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
+It is directly related to the line--of--sight wavenumber,
+\begin{align}\label{eqn:k-prll}
+  k_\parallel &\approx \frac{2\pi\eta\,f_{21}H_0\,E(z)}{c(1+z)^2}, 
+\end{align}
+where, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
 This approximation holds under the assumption that the redshift range (or frequency band) is small enough within which cosmological evolution is negligible.
@@
 represents the true spatial Fourier representation of the three--dimensional sky brightness distribution.
-This representation has been discussed in detail in \citet{mor04}.
-The spatial power spectrum of EoR H{\sc i} distribution, $P(\boldsymbol{k}_\perp,k_\parallel)$, and $V_{u\eta}$ are related by $P(\boldsymbol{k}_\perp,k_\parallel)\propto |V_{u\eta}|^2$.
+This approach has been discussed in detail in \citet{mor04}.
+The spatial power spectrum of EoR H{\sc i} distribution, $P(\boldsymbol{k}_\perp,k_\parallel)$, and $V_{u\eta}$ are related by \citep{mor04,mcq06,par12a}:
+\begin{align}\label{eqn:true-power_spectrum}
+  P(\boldsymbol{k}_\perp,k_\parallel) &\simeq |V_{u\eta}|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
+\end{align}
+where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, and $k_\textrm{B}$ is the Boltzmann constant.
+Thus, $V_{u\eta}$ inferred from observations, also in units of Jy~Hz, can be converted into an equivalent cosmological H{\sc i} power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$, in units of K$^2$~Mpc$^3$.
 
-While $V_{b\tau}$ and $V_{u\eta}$ in equations~\ref{eqn:obsvis} and \ref{eqn:vis} resemble each other very closely, they are not quite identical.
+Without loss of generality, the phase center can be assumed to be the zenith relative to the local tangent plane.
+Then $\boldsymbol{u}$ lies on this plane for measurements constrained to be on it.
+If the array of antennas are also coplanar lying on the local tangent plane, then $\boldsymbol{u}=\boldsymbol{b}/\lambda$.
+Under such circumstances, equations~\ref{eqn:obsvis} and \ref{eqn:delay} closely resemble equations~\ref{eqn:vis} and \ref{eqn:los-transform} respectively.
+However, they are not quite identical to each other.
 The difference basically arises from the fact noted above that $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not.
@@
 
-Radio emission from foreground objects, due to their spatial and spectral (chromatic) structure, contaminate the received signal \citep{bow09,dat10,mor12,tro12,thy13,liu14a,liu14b}.
-Since this contamination is expected to be several orders of magnitude stronger than the underlying EoR H{\sc i} signal, it is critical to characterize foregrounds precisely in order to reduce their impact on EoR H{\sc i} power spectrum detection sensitivity.
+% Radio emission from foreground objects, due to their spatial and spectral (chromatic) structure, contaminate the received signal \citep{bow09,dat10,mor12,tro12,thy13,liu14a,liu14b}. 
 Foregrounds can be described in both the delay spectrum and Fourier frameworks.
-For our study, which primarily concerns with foreground characterization, we find the former to be simple and yet extremely useful, while maintaining a close correspondence with the latter despite their subtle differences.
+For our study, which primarily concerns with foreground characterization, we find the former to be simple and yet extremely useful, while maintaining a close correspondence with the latter despite subtle differences.
+In order to express a quantity derived from $V_{u\eta}$ that has units identical to that of aforementioned cosmological H{\sc i} power spectrum derived, we define:
+\begin{align}\label{eqn:fake-power-spectrum}
+  P'(\boldsymbol{k}_\perp,k_\parallel) &\equiv |V_{b\tau}|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
+\end{align}
+where, $\boldsymbol{u}$ and $\eta$ in equations~\ref{eqn:k-perp} and \ref{eqn:k-prll} have been correspondingly replaced with $\boldsymbol{b}/\lambda$ and $\tau$ respectively.
 
@@
 $V_{b\tau}$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument.
-At the same time, it is closely related to the sought power $|V_{u\eta}|^2$ in Fourier modes containing critical information about spatial scales.
-For purposes of illustration later in this paper, we often use the quantity, $|V_{b\tau}|$.
+At the same time, it is closely related to the sought power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$ containing critical information about spatial scales.
+For purposes of illustration later in this paper, we often use the quantities $|V_{b\tau}|$ and $P'(\boldsymbol{k}_\perp,k_\parallel)$.
 
-Note that since {\it visibility} is a complex quantity, its Fourier transform is Nyquist sampled over positive and negative delays.
-Negative and positive delay\footnote{The terms delay and lag are used interchangeably in this paper.} correspond to the two hemispheres of the sky as transected by the antenna spacing vector.
-In this paper, $\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the telescope site.
+Note that since {\it visibility} is a complex quantity, its {\bf delay} transform {\bf contains} positive and negative delays.
+Negative and positive delays correspond to the two hemispheres of the sky as transected by the antenna spacing vector.
+In this paper, $\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward (zenith) directions at the telescope site.
 Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at positive and negative delays, respectively.
 Similarly, an object in the northern sky will appear at a positive delay at an antenna spacing oriented northward.
+For all observations used in this study, we use the zenith as the phase center.
 
-Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of H{\sc i} from the EoR are presented in this paper.
-$|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction, form the $x$--axis.
+{\bf We give a brief overview of some parameters of Fourier space which is generic to all experiments that use a similar approach}.
+Figure~\ref{fig:fourier-space} illustrates the Fourier space, {\bf specific to the MWA}, in which the delay (and power) spectra of H{\sc i} from the EoR are presented in this paper.
+$|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction {\bf (tangent plane to the celestial sphere)}, form the $x$--axis.
 $\tau\approx\eta\propto k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis.
 Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by the antenna spacings and the light travel times across them.
-These boundaries are called horizon delay limits and are shown by solid lines.
-The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\tau$--axis.
-This effect has been quantified in \citet{thy13}.
+These boundaries, called horizon delay limits \citep{ved12,par12b}, are shown by solid lines.
+The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\tau$--axis \citep{par12b,thy13}.
 The width of this narrow strip is inversely proportional to the operating bandwidth.
@@
 Its boundaries (solid lines), given by light travel time for corresponding antenna spacings, are referred to as horizon delay limits.
-Regions excluding the {\it wedge} and narrow extensions of the {\it wedge} (white unshaded strips) caused by convolution with the instrumental transfer function are expected to be relatively free of foreground contamination.
+Regions excluding the {\it wedge} and narrow extensions of the {\it wedge} (white unshaded strips) caused by convolution with the {\bf instrument's spectral transfer function} are expected to be relatively free of foreground contamination.
 Due to an undesirable grating pattern (dot--dashed lines) of the instrument used in this study, namely the MWA, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the {\it EoR window} in this paper.
@@
 
-For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67\fdg 5 \leq \theta_b < 112\fdg 5$.
+For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67.5\arcdeg \leq \theta_b < 112.5\arcdeg$.
 Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
@@
 The center frequency of 185~MHz corresponds to a redshift, $z\simeq 6.68$.
+{\bf At this frequency, we use $A_\textrm{e}=10.5$~m$^2$}.
 
-The MWA passband of width 30.72~MHz is divided coarsely into 24$\times$1.28~MHz sub--bands with each sub--band weighted by a digital filter.
+The MWA passband of width $\Delta B=30.72$~MHz is divided coarsely into 24$\times$1.28~MHz sub--bands with each sub--band weighted by a digital filter.
 The coarse channel shape is obtained using a 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$.
-After correcting for the shape of these coarse channels, the fine channels (40~kHz width) at the edges of these sub--bands are flagged because they are known to be contaminated by aliasing at a low level.
+{\bf Each of these coarse bands consists of 32 fine channels of width 40~kHz each}.
+After correcting for the shape of these coarse channels, the fine channels at the edges of these sub--bands are flagged because they are known to be contaminated by aliasing at a low level.
 Any spectral flags, or other known weights constitutes $W_f$ in equation~\ref{eqn:obsvis}.
 
-Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=95$~K per polarization for all frequency channels for all antenna pairs throughout the course of the observation.
+% Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=95$~K per polarization for all frequency channels for all antenna pairs throughout the course of the observation.
 Estimation of $T_\textrm{sys}$ from MWA data is a subject of active investigation.
@@
 The MWA targets two primary low-foreground fields for reionization observations.
-Here, we focus on the field at RA = 0$^\textrm{h}$, Dec = $-30$\arcdeg.
+Here, we focus on the field at RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$.
 The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of the dipoles in an MWA tile.
@@
 
-The {\it local sidereal time} (LST) at these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at $-1.92$~hours (i.e., 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings, respectively.
+The {\it local sidereal time} (LST) at these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at {\bf 22.08~hours ($-1.92$~hours in the figure)} and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings, respectively.
 
@@
 These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey \citep[MWACS;][]{hur14}.
-It has a flux density limit of 25~mJy and a declination range of -12\arcdeg~ to -40\arcdeg~ evenly covering the field of view of the observations reported here.
+It has a flux density limit of 25~mJy and a declination range of $-12\arcdeg$~ to $-40\arcdeg$~ evenly covering the field of view of the observations reported here.
 The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern.
@@
 
-The {\it off--zenith} pointing has emission notably higher than in the {\it zenith} pointing inside the {\it foreground wedge}.
-This is shown later to be due to response of the array to the bright Galactic center and Galactic plane in the westward sky.
+The {\it off--zenith} pointing {\bf has notably higher power in the {\it foreground wedge} than in the {\it zenith} pointing}.
+This is shown later in \S\ref{sec:diffuse} to be due to response of the array to the bright Galactic center and Galactic plane in the westward sky.
 Consequently, the contamination into the {\it EoR window} is also higher, especially on short antenna spacings.
-Emission in the {\it zenith} pointing is more centrally concentrated due to phasing of power--pattern to the zenith, whereas the {\it off--zenith} pointing has a power--pattern phased eastward of zenith but is also responsive to emission from angles far away from zenith.
+Emission in the {\it zenith} pointing is more centrally concentrated {\bf due to a zenith--centered power--pattern and faster suppression of response away from zenith, whereas the {\it off--zenith} pointing has a power--pattern phased eastward of zenith and is also more responsive to emission farther away from the primary lobe relative to that in the {\it zenith} pointing.
+The power pattern is discussed in detail in \S\ref{sec:instrument_model} and Figure~\ref{fig:power_pattern}}.
 
@@
 \includegraphics[width=\linewidth]{f3.eps}
-\caption{Delay spectra of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30$\arcdeg) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
+\caption{Delay spectra of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
+{\bf In both cases, the delay spectra are phased to the zenith and thus the {\it foreground wedge} boundaries are symmetric around $\tau=0$}.
 For reference, the $k$--space equivalent of $\tau$ and $|\boldsymbol{b}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively.
@@
 
+{\bf Estimation of $T_\textrm{sys}$ from MWA data is a subject of active investigation.
+For our work, we estimate thermal noise in the data using the {\it rms} of $V_{b\tau}$ obtained after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
+\begin{align}\label{eqn:Tsys}
+  \Delta V_{b\tau}^\textrm{rms} &= \sqrt{N_\textrm{ch}}\,\Delta V_{bf}^\textrm{rms} \Delta f, \textrm{and} \\
+  \Delta V_{bf}^\textrm{rms} &= \frac{2\,k_\textrm{B}\,T_\textrm{sys}}{A_\textrm{e}\sqrt{2\,\Delta f\,\Delta t}},
+\end{align}
+where, $\Delta f = 80$~kHz, $\Delta t = 112$~s, and $N_\textrm{ch}=\Delta B/\Delta f$ is the number of frequency channels.
+The choice of threshold for $\tau$ is well outside the {\it foreground window}, where foreground contamination is negligible and thus makes a robust estimation of $T_\textrm{sys}$ possible.
+We find the average system temperature to be $\sim 95$~K across all frequency channels and across all antenna pairs throughout the course of the observation.
+Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise observed in the data.}
+
 \subsection{Foreground Model}\label{sec:foreground}
 
-The MWA has a very wide field of view ($\gtrsim 20$\arcdeg~ at 185~MHz) and has non--negligible response all the way to the horizon.
+The MWA has a very wide field of view ($\gtrsim 20\arcdeg$~ at 185~MHz) and has non--negligible response all the way to the horizon.
 As already noted, foreground objects fill the {\it foreground wedge} with an amplitude which depends on the tile power gain in that direction, with objects at the horizon being closest to the boundaries of the {\it wedge}.
 
-The degree to which these objects leak power into spectral line of sight modes in the {\it EoR window} will be shown to depend primarily on the brightness of foregrounds near the boundaries of the {\it foreground wedge}.
-Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view, a point also supported by Pober~et~al.~(2014, in preparation).
+The degree to which these objects leak power into the spectral line--of--sight modes in the {\it EoR window} will be shown to depend primarily on the brightness of foregrounds near the boundaries of the {\it foreground wedge}.
+Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view, a point also supported by Pober~et~al.~(2015, in preparation).
 
@@
 For the diffuse component, we use an all--sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz.
-At this frequency, since this map is predominantly based on the 408~MHz map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz map to the same resolution.
+At this frequency, since this map is predominantly based on the 408~MHz map of \citet{has82} which has an angular resolution of 0.85\arcdeg, we smoothed the 185~MHz map to the same resolution.
 However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals.
@@
 The model described in the preceding section is primarily a model of the diffuse foreground sky.
-While it contains faint compact emission blended in with the diffuse emission, it is known to be devoid of bright point sources \citep{deo08}.
+While it contains faint compact emission blended in with the diffuse emission, {\bf bright point sources have been removed \citep{deo08}}.
 In order to supplement it with missing bright compact emission, we use classical radio source confusion estimates to determine the nominal flux density threshold and include point sources brighter than this threshold.
 It may be noted that slightly different criteria are in common use in radio astronomy to estimate radio source confusion \citep[see Appendix of][and references therein]{thy13}.
-For an angular resolution of 0\fdg 85, using a conservative `$S_\textrm{c}=5\sigma_\textrm{c}$' criterion, we determine the flux density threshold to be $\approx 10$~Jy.
+For an angular resolution of 0.85\arcdeg, using a conservative `$S_\textrm{c}=5\sigma_\textrm{c}$' criterion, we determine the flux density threshold to be $\approx 10$~Jy.
 Other liberal criteria that yield a lower threshold carry a greater risk of double--counting point sources which might be already blended in with the diffuse sky model.
@@
 We use a combination of the NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and the Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz to provide our point source catalog due to their complementary survey footprints covering the entire sky, and matched flux density sensitivity and angular resolution.
-The SUMSS catalog covers the sky with declination $\delta < -30$\arcdeg~ with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec.
-The NVSS covers the sky with $\delta > -40$\arcdeg~ with a similar angular resolution and a limiting flux density of $\approx 2.5$~mJy for point sources.
+The SUMSS catalog covers the sky with declination $\delta < -30\arcdeg$~ with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec.
+The NVSS covers the sky with $\delta > -40\arcdeg$~ with a similar angular resolution and a limiting flux density of $\approx 2.5$~mJy for point sources.
 
-From the SUMSS catalog, we select point sources whose deconvolved major axes are equal to 0\arcsec.
+From the SUMSS catalog, we select {\bf objects} whose deconvolved major axes are equal to 0\arcsec, {\bf thereby strictly selecting point sources}.
 From the NVSS catalog, we excluded objects that overlap with those in the SUMSS survey footprint.
-Point sources from NVSS were selected if the convolved major axes were not greater than $\approx 47$\arcsec, which is almost equal to the angular resolution of the survey.
+Point sources from NVSS were selected if the convolved major axes were not greater than $\approx 47$\arcsec, which matches the angular resolution of the survey.
 Using a mean spectral index of $\langle\alpha_\textrm{sp}\rangle=-0.83$ (flux density, $S(f)\propto f^{\alpha_\textrm{sp}}$) obtained by \citet{mau03} for both NVSS and SUMSS catalog objects, we calculate the corresponding flux densities at 185~MHz, $S_{185}$.
@@
 The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observing frequency.
-Using simple error propagation, the fractional error in the delay spectrum caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the MWA observing frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index.
+Using simple error propagation, the fractional error in the delay spectrum {\bf amplitude} caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the MWA observing frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index.
 From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for point sources from NVSS and SUMSS catalogs.
 Although the model of \citet{deo08} yields a spectral index per direction on the sky, we could assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from the 408~MHz map of \citet{has82}.
-Thus, fractional errors in delay spectrum from compact and diffuse components are $\sim$70\% and $\sim$30\% respectively.
+Thus, fractional errors in delay spectrum {\bf amplitudes} from compact and diffuse components are $\sim$70\% and $\sim$30\% respectively.
 
@@
 A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power--pattern \citep{ber14}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide--field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations, but is beyond the scope of this paper.
-Our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.
+{\bf Hereafter}, our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.
 
@@
 It also consists of localized regions of strong emission such as the Galactic plane and Galactic center.
-The resulting sky emission is anisotropic.
+% The resulting sky emission is anisotropic.
 In fact, the patch of sky for MWA observations is chosen from regions of low foreground emission.
@@
 \item {\it Baseline Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our antenna spacings by their orientation, $\theta_\textrm{b}$.
-We use the following bins: $-67\fdg 5\le \theta_\textrm{b} < -22\fdg 5$, $-22\fdg 5\le \theta_\textrm{b} < 22\fdg 5$, $22\fdg 5\le \theta_\textrm{b} < 67\fdg 5$, and $67\fdg 5\le \theta_\textrm{b} < 112\fdg 5$.
+We use the {\bf four following bins sweeping 45\arcdeg each}: $-67.5\arcdeg\le \theta_\textrm{b} < -22.5\arcdeg$, $-22.5\arcdeg\le \theta_\textrm{b} < 22.5\arcdeg$, $22.5\arcdeg\le \theta_\textrm{b} < 67.5\arcdeg$, and $67.5\arcdeg\le \theta_\textrm{b} < 112.5\arcdeg$.
 The bin centers are oriented towards South--East, East, North--East, and North respectively.
+{\bf While the bin boundaries are arbitrary, they were so chosen to allow bin centers to be aligned along the cardinal and ordinal directions in one half--plane}.
 Since delays depend on $\theta_\textrm{b}$, binning by $\theta_\textrm{b}$ allows us to match delay spectrum features to different sky directions.
 
-\item {\it Tile Pointing and Power--Pattern}: Since the MWA tiles are steered electronically, the power--pattern of the tile changes in any observing mode that tracks the object.
+\item {\it Tile Pointing and Power--Pattern}: Since the MWA tiles are steered electronically, the power--pattern of the tile changes in any observing mode that tracks the {\bf pointing center}.
 In our study, we take into account the effect of the changing tile power--pattern on the delay spectrum (Figure~\ref{fig:power_pattern}).
@@
 This feature is strongest at short antenna spacings and fades with increasing antenna spacing.
-The signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-delay-spectrum}) because the Galactic center is below the horizon.
+The {\bf bright} signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-delay-spectrum}) because the Galactic center is below the horizon.
 
@@
 As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inversely proportional to the length of the baseline projected in the direction of the emission.
-Thus, in the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at zero delay, fades away on antenna spacings $|\boldsymbol{b}| \gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0\fdg 75.
+Thus, in the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at zero delay, fades away on antenna spacings $|\boldsymbol{b}| \gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0.75\arcdeg.
 In the off-zenith pointing, we see power out to much longer baselines which is described below.
@@
 A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\boldsymbol{b}|\le$~200~m).
-This is contrary to expectations because these longer baselines are not expected to be sensitive to diffuse emission (on scales $\gtrsim$~0\fdg 85) of our input foreground model.
-In the wide--field observations of the MWA, long baselines appear shortened due to projection effects for off--axis emission, especially towards the horizon.
+This is contrary to expectations because these longer baselines are not expected to be sensitive to diffuse emission (on scales $\gtrsim$~0.85\arcdeg) of our input foreground model.
+{\bf In wide--field observations}, long baselines appear shortened due to projection effects for off--axis emission, especially towards the horizon.
+{\bf Consequently, they become sensitive to diffuse emission}.
 This effect is evident at all LSTs in our simulations and it naturally holds for short baselines, as well.
@@
 The gray shaded area denotes the envelope of expected uncertainty in the delay spectrum.
-Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the $\gtrsim$~70\% error owing to uncertainties in predicting the spectral index of compact foreground model, while thermal noise fluctuations dominate outside.
+Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the % $\gtrsim$~70\% error owing to
+uncertainty in predicting the spectral index of compact foreground model, while thermal noise fluctuations dominate outside.
 
@@
 Vertical dotted lines mark the horizon delay limits.
+{\bf Compact emission dominates the central regions of delay spectrum while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a {\it pitchfork}--shaped structure.
 The shaded region denotes the envelope of uncertainty around the composite model delay spectrum.
 This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits.
-Outside the horizon delay limits, it is predominantly due to thermal noise fluctuations.
-Compact emission dominates the central regions of delay spectrum while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a {\it pitchfork}--shaped structure.
+Outside the horizon delay limits, it is predominantly due to thermal noise fluctuations}.
 \label{fig:pitchfork-baselines}}
@@
 Since the Galactic center is located in the western sky, this figure demonstrates that it is observed at negative delay on a baseline vector oriented eastward and at zero delay on a baseline vector oriented northward.
-Figure~\ref{fig:baseline-breakup} shows the delay spectra on baselines oriented northward (67\fdg 5~$\le \theta_\textrm{b} <$~112\fdg 5) at the top and eastward (-22\fdg 5~$\le \theta_\textrm{b} <$~22\fdg 5) at the bottom.
+Figure~\ref{fig:baseline-breakup} shows the delay spectra on baselines oriented northward ($67.5\arcdeg\le \theta_\textrm{b} < 112.5\arcdeg$) at the top and eastward ($-22.5\arcdeg~\le \theta_\textrm{b} < 22.5\arcdeg$) at the bottom.
 The Galactic center manifests itself most distinctly near the negative horizon delay limit on short eastward baselines in the delay spectrum (bottom panel of Figure~\ref{fig:baseline-breakup}).
-Consequently, the spillover caused by the instrument's transfer function from the {\it foreground wedge} into the {\it EoR window} affects the northward baselines the least and is most severe in eastward baselines (particularly the short ones) evident by the bright vertical stripes of foreground contamination.
+Consequently, the spillover caused by the instrument's spectral transfer function from the {\it foreground wedge} into the {\it EoR window} affects the northward baselines the least and is most severe in eastward baselines (particularly the short ones) evident by the bright vertical stripes of foreground contamination.
 
@@
 
-In our example, we discard eastward antenna spacings (-22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5) of lengths $|\boldsymbol{b}| < 30$~m.
+In our example, we discard eastward antenna spacings ($-22.5\arcdeg\le\theta_\textrm{b}< 22.5\arcdeg$) of lengths $|\boldsymbol{b}| < 30$~m.
+{\bf This was determined as the limit beyond which foreground contamination levels are insensitive to any further removal of wider antenna spacings}.
 Figure~\ref{fig:before-after} shows the delay spectra obtained with all antenna spacings (top panel) and after applying our screening technique (bottom panel) on the {\it off--zenith} observation.
@@
 \caption{Simulated delay spectra amplitudes $|V_{b\tau}|$ for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
-Discarded antenna spacings (black vertical stripes) have lengths $|\boldsymbol{b}|<30$~m (leftward of vertical dashed line) and orientations $|\theta_\textrm{b}|<$~22\fdg 5.
+Discarded antenna spacings (black vertical stripes) have lengths $|\boldsymbol{b}|<30$~m (leftward of vertical dashed line) and orientations $|\theta_\textrm{b}|<$~22.5\arcdeg.
 The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} prominent when all antenna spacings are used is lowered by an order of magnitude when short eastward antenna spacings are discarded.
@@
 Figure~\ref{fig:screening} shows how the typical foreground contamination\footnote{Foreground contamination is measured by standard deviation of noiseless $V_{b\tau}$ from foregrounds in the {\it EoR window}.} in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
-We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars).
+We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., $-7.5\arcdeg~\le\theta_\textrm{b} < 7.5\arcdeg$ (solid circles), $-15\arcdeg\le\theta_\textrm{b}< 15\arcdeg$ (solid squares), and $-22.5\arcdeg\le\theta_\textrm{b}< 22.5\arcdeg$ (solid stars).
 Among antenna spacings that satisfy these criteria, we discard data from those whose lengths are shorter than $|\boldsymbol{b}|_\textrm{max}$ ($x$--axis) and show foreground contamination estimated in the {\it EoR window} from all remaining antenna spacings.
@@
 The fraction of discarded antenna spacings discarded relative to the total number is shown in dotted lines for different ranges of $\theta_\textrm{b}$.
-It is seen that foreground contamination can be mitigated by a factor between $\sim 1.5$ ($|\theta_\textrm{b}|\le$~7\fdg 5) and $\sim 10$ ($|\theta_\textrm{b}|\le$~22\fdg 5).
+It is seen that foreground contamination can be mitigated by a factor between $\sim 1.5$ ($|\theta_\textrm{b}|\le$~7.5\arcdeg) and $\sim 10$ ($|\theta_\textrm{b}|\le$~22.5\arcdeg).
 The latter limit is achieved with a mere 5\% loss of data for $|\boldsymbol{b}|_\textrm{max}\simeq30$~m.
@@
 Such cases will necessitate discarding more or, in the worst case, all of the eastward oriented antenna spacings.
+{\bf These results are specific to the MWA data.
+In general, the array configuration will be a significant factor in determining such thresholds}.
 
@@
 \caption{Drop in foreground contamination in the {\it EoR window}, and loss of data for the {\it off--zenith} pointing as a function of discarded baselines.
-Eastward baselines with varying degrees of directedness -- $|\theta_\textrm{b}|<$~7\fdg 5 (solid circles), $|\theta_\textrm{b}|<$~15\arcdeg~ (solid squares), and $|\theta_\textrm{b}|<$~22\fdg 5 (solid stars) -- and lengths $|\boldsymbol{b}| \le |\boldsymbol{b}|_\textrm{max}$ ($x$--axis) are discarded.
+Eastward baselines with varying degrees of directedness -- $|\theta_\textrm{b}|<$~7.5\arcdeg (solid circles), $|\theta_\textrm{b}|<$~15\arcdeg~ (solid squares), and $|\theta_\textrm{b}|<$~22.5\arcdeg (solid stars) -- and lengths $|\boldsymbol{b}| \le |\boldsymbol{b}|_\textrm{max}$ ($x$--axis) are discarded.
 Loss of data (dotted lines) is measured by discarded baselines as a fraction of the total number for the corresponding cases.
-Foreground contamination in the {\it EoR window} (solid lines) drops by a factor $\sim$1.5 ($|\theta_\textrm{b}|\le$~7\fdg 5) to $\sim$10 ($|\theta_\textrm{b}|\le$~22\fdg 5).
+Foreground contamination in the {\it EoR window} (solid lines) drops by a factor $\sim$1.5 ($|\theta_\textrm{b}|\le$~7.5\arcdeg) to $\sim$10 ($|\theta_\textrm{b}|\le$~22.5\arcdeg).
 The latter limit can be achieved with a mere 5\% loss of data at $|\boldsymbol{b}|_\textrm{max}\simeq30$~m, and discarding longer baselines ($|\boldsymbol{b}|\gtrsim 30$~m) has no effect in further reducing foreground contamination.
@@
 
-Simulating in all important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright point sources from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise fluctuations in measurements.
+Simulating in many important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright point sources from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise fluctuations in measurements.
 
@@
 In conclusion, we find that inclusion of emission models, both diffuse and compact, all the way to the horizon is essential to explaining the observed power spectrum.
-We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination by an order of magnitude in EoR data analysis by discarding or down--weighting data from antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation, with negligible loss of sensitivity.
+We also provide a simple and effective tool based on the delay spectrum technique {\bf that can potentially} mitigate foreground contamination by an order of magnitude in EoR data analysis by discarding or down--weighting data from antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation, with negligible loss of sensitivity.
 

### 2015-01-08 ms.tex
Added a discussion on possible pitchfork effect found in Pober et al 2013 and Thyagarajan et al 2013.

@@
 
+{\bf Interestingly, \citet{thy13} and \citet{pob13} have found a similar feature in their simulations and data respectively.
+In the latter, this feature is described as ``edge brightening'' and attributed to the Galactic plane near the horizon.
+From their discussion, it is unclear what fraction of power in that feature arises from ubiquitous diffuse emission unlike in our {\it zenith} pointing where the Galactic plane is not expected to play a significant role.
+On the other hand, \citet{thy13} statistically modeled point sources below the confusion limit and attributed the feature to an increase in solid angles subtended by delay bins closer to the horizon limits thereby, enhancing the resulting integrated emission.
+This must be true for diffuse emission as well.
+Thus we argue the feature arises primarily due to an increase in sensitivity of long baselines to diffuse emission from far off--axis angles caused by foreshortening, and partly due to a higher integrated emission in delay bins close to the horizon limits which map larger solid angles in these directions}.
+
 \subsection{Compact Foreground Signatures}\label{sec:compact}

### 2015-01-10 ms.tex
Addressed baseline screening technique in the context of covariance based power estimation techniques and updated reference to Sutinjo et al.(2014)

@@
 The instrument model consists primarily of the tile beam power--pattern, $A(\hat{\boldsymbol{s}},f)$ (see equation~\ref{eqn:obsvis}).
-It is modeled as a mutually--coupled 4--by--4 dipole array with the overall power--pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14}.
+It is modeled as a mutually--coupled 4--by--4 dipole array with the overall power--pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14_arXiv}.
 To speed up simulations, we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles.
@@
 In principle, instead of discarding selected antenna spacings altogether, we could down--weight them based on an optimal scheme.
+For instance, the estimates of covariance computed from the delay transform bins can be naturally fed into the covariance--weighted power spectrum estimation techniques \citep{liu14a,liu14b}.
+It could also be used to downweight or flag contaminated baselines in imaging applications.
 This technique provides a very simple and yet effective tool in adding a layer of control to mitigate effects of foreground contamination in EoR data analysis.

### 2015-01-17 ms.tex
Changed figures which now use units of power

@@
 \begin{align}\label{eqn:fake-power-spectrum}
-  P'(\boldsymbol{k}_\perp,k_\parallel) &\equiv |V_{b\tau}|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
+  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |V_{b\tau}|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
 \end{align}
@@
 At the same time, it is closely related to the sought power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$ containing critical information about spatial scales.
-For purposes of illustration later in this paper, we often use the quantities $|V_{b\tau}|$ and $P'(\boldsymbol{k}_\perp,k_\parallel)$.
+For purposes of illustration later in this paper, we often use the quantities $|V_{b\tau}|$ and $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$.
 
@@
 % \includegraphics[width=\linewidth]{figures/v1_0/fourier_space_185.0_MHz_30.7_MHz.eps}
-\includegraphics[width=\linewidth]{f1.eps}
+% \includegraphics[width=\linewidth]{f1.eps}
+\includegraphics[width=\linewidth]{fig1.eps}
 \caption{Fourier space in which delay (and power) spectra of EoR H{\sc i} signals are presented.
@@
 % \includegraphics[width=\linewidth]{figures/v1_0/custom_pointings.eps}
-\includegraphics[width=\linewidth]{f2.eps}
+% \includegraphics[width=\linewidth]{f2.eps}
+\includegraphics[width=\linewidth]{fig2.eps}
 \caption{MWA tile beam pointing directions during the course of the observation.
@@
 % \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
-\includegraphics[width=\linewidth]{f3.eps}
-\caption{Delay spectra of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
+% \includegraphics[width=\linewidth]{f3.eps}
+\includegraphics[width=\linewidth]{fig3.eps}
+\caption{Delay spectra power of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
 {\bf In both cases, the delay spectra are phased to the zenith and thus the {\it foreground wedge} boundaries are symmetric around $\tau=0$}.
@@
 % \includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom.eps}
-\includegraphics[width=\linewidth]{f4.eps}
+% \includegraphics[width=\linewidth]{f4.eps}
+\includegraphics[width=\linewidth]{fig4.eps}
 \caption{Models of MWA tile power--pattern for {\it off--zenith} (top) and {\it zenith} pointings (bottom) at 185~MHz.
@@
 % \includegraphics[width=\linewidth]{figures/v1_0/dsm.eps}
-\includegraphics[width=\linewidth]{f5.eps}
+% \includegraphics[width=\linewidth]{f5.eps}
+\includegraphics[width=\linewidth]{fig5.eps}
 \caption{Sky brightness temperature of the diffuse foreground model at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 % \includegraphics[width=\linewidth]{figures/v1_0/csm.eps}
-\includegraphics[width=\linewidth]{f6.eps}
+% \includegraphics[width=\linewidth]{f6.eps}
+\includegraphics[width=\linewidth]{fig6.eps}
 \caption{Flux densities of bright point sources at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 % \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
-\includegraphics[width=\linewidth]{f7.eps}
+% \includegraphics[width=\linewidth]{f7.eps}
+\includegraphics[width=\linewidth]{fig7.eps}
 \caption{Same as Figure~\ref{fig:fhd_data}, but generated from simulated visibilities using the model described in \S\ref{sec:modeling}.
@@
 % \includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-\includegraphics[width=\linewidth]{f8.eps}
-\caption{Simulated delay spectrum amplitudes $|V_{b\tau}|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
+% \includegraphics[width=\linewidth]{f8.eps}
+\includegraphics[width=\linewidth]{fig8.eps}
+\caption{Simulated delay spectrum {\bf power (in units of K$^2$~(Mpc/$h$)$^3$} for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
 White lines mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
@@
 % \includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-\includegraphics[width=\linewidth]{f9.eps}
+% \includegraphics[width=\linewidth]{f9.eps}
+\includegraphics[width=\linewidth]{fig9.eps}
 \caption{Same as Figure~\ref{fig:noiseless-dsm-delay-spectrum} but for a foreground model consisting of compact emission.
@@
 % \includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-\includegraphics[width=\linewidth]{f10.eps}
-\caption{Simulated delay spectrum amplitudes $|V_{b\tau}|$ for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}).
+% \includegraphics[width=\linewidth]{f10.eps}
+\includegraphics[width=\linewidth]{fig10.eps}
+\caption{Simulated delay spectrum power for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}).
 The antenna spacing is specified in each panel along East, North and Upward directions.
@@
 % \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figures/v1_0/baseline_orientation_breakup_snapshot_0.eps}}
-\subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{f11a.eps}}
-\subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{f11b.eps}}
-\subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{f11c.eps}}
+% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{f11a.eps}}
+% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{f11b.eps}}
+% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{f11c.eps}}
+\subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{fig11a.eps}}
+\subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{fig11b.eps}}
+\subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{fig11c.eps}}
 \caption{(a): Sky model showing compact ({\it top}) and diffuse ({\it bottom}) emission.
@@
 The bright Galactic center in the western sky is recorded at zero delays in northward antenna spacings and close to negative horizon delay limit on eastward antenna spacings.
-(c): Simulated delay spectra $|V_{b\tau}|$ on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}).
+(c): Simulated delay spectra {\bf power} on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}).
 White lines denote horizon delay limits.
@@
 % \includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_before_after_nside_64_Tsys_95.0K.eps}
-\includegraphics[width=\linewidth]{f12.eps}
-\caption{Simulated delay spectra amplitudes $|V_{b\tau}|$ for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
+% \includegraphics[width=\linewidth]{f12.eps}
+\includegraphics[width=\linewidth]{fig12.eps}
+\caption{Simulated delay spectra {\bf power} for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
 Discarded antenna spacings (black vertical stripes) have lengths $|\boldsymbol{b}|<30$~m (leftward of vertical dashed line) and orientations $|\theta_\textrm{b}|<$~22.5\arcdeg.
-The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} prominent when all antenna spacings are used is lowered by an order of magnitude when short eastward antenna spacings are discarded.
+The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} prominent when all antenna spacings are used is lowered by {\bf nearly two orders} of magnitude when short eastward antenna spacings are discarded.
 Color scale has been adjusted to emphasize mitigation achieved in foreground contamination.
@@
 % \includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_nside_64_Tsys_95.0K.eps}
-\includegraphics[width=\linewidth]{f13.eps}
+% \includegraphics[width=\linewidth]{f13.eps}
+\includegraphics[width=\linewidth]{fig13.eps}
 \caption{Drop in foreground contamination in the {\it EoR window}, and loss of data for the {\it off--zenith} pointing as a function of discarded baselines.
@@
 Loss of data (dotted lines) is measured by discarded baselines as a fraction of the total number for the corresponding cases.
-Foreground contamination in the {\it EoR window} (solid lines) drops by a factor $\sim$1.5 ($|\theta_\textrm{b}|\le$~7.5\arcdeg) to $\sim$10 ($|\theta_\textrm{b}|\le$~22.5\arcdeg).
+Foreground contamination in the {\it EoR window} (solid lines) drops by a factor $\sim$2 ($|\theta_\textrm{b}|\le$~7.5\arcdeg) to $\sim$100 ($|\theta_\textrm{b}|\le$~22.5\arcdeg).
 The latter limit can be achieved with a mere 5\% loss of data at $|\boldsymbol{b}|_\textrm{max}\simeq30$~m, and discarding longer baselines ($|\boldsymbol{b}|\gtrsim 30$~m) has no effect in further reducing foreground contamination.

### 2015-01-17 ms.tex
New section on role of antenna shapes partially complete

@@
 This must be true for diffuse emission as well.
-Thus we argue the feature arises primarily due to an increase in sensitivity of long baselines to diffuse emission from far off--axis angles caused by foreshortening, and partly due to a higher integrated emission in delay bins close to the horizon limits which map larger solid angles in these directions}.
+Thus we argue the feature arises primarily due to an increase in sensitivity of long baselines to diffuse emission caused by foreshortening towards angles far off--axis, where the increased size scales (or solid angles) containing higher levels of integrated emission and spatial scales given by the inverse of foreshortened baseline vectors become matched}.
 
@@
 This screening technique can be generalized to optimize between foreground mitigation and loss of sensitivity from discarding data.
-Figure~\ref{fig:screening} shows how the typical foreground contamination\footnote{Foreground contamination is measured by standard deviation of noiseless $V_{b\tau}$ from foregrounds in the {\it EoR window}.} in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
+Figure~\ref{fig:screening} shows how the typical foreground contamination\footnote{Foreground contamination is measured by standard deviation of noiseless $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$ from foregrounds in the {\it EoR window}.} in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
 We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., $-7.5\arcdeg~\le\theta_\textrm{b} < 7.5\arcdeg$ (solid circles), $-15\arcdeg\le\theta_\textrm{b}< 15\arcdeg$ (solid squares), and $-22.5\arcdeg\le\theta_\textrm{b}< 22.5\arcdeg$ (solid stars).
@@
 This technique provides a very simple and yet effective tool in adding a layer of control to mitigate effects of foreground contamination in EoR data analysis.
+
+\section{Role of Antenna Geometry}\label{sec:antenna-shape}
+
+Having examined the role the foregrounds play in determining the delay spectrum obtained with an interferometer array, we turn our attention to another crucial factor, namely, the power pattern of the antenna (refer to equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}).
+In the following discussion, we assume there are no random delay fluctuations in the MWA tile beamforming process.
+The model power pattern so obtained is different from the one we used earlier, but for the present discussion it still serves as a good reference (Figure~\ref{fig:power_pattern}).
+Since the previous model suffices, we do not plot the new power pattern model obtained with ideal beamforming.
+We also assume an ideal flat bandpass shape over the 30.72~MHz bandwidth unlike the flagging of the edges of the coarse sub--bands described earlier in this paper, while still applying a {\it Blackman--Harris} window function.
+
+\subsection{Uniformly Illuminated Sky}\label{sec:USM}
+
+The brightness distribution on the sky was fixed to be a constant equal to the mean over the whole sky at 185~MHz with no variation along frequency.
+This helps isolate the signatures of the antenna power pattern from those of the foregrounds.
+Figure~\ref{fig:noiseless-usm-delay-spectrum} shows the delay spectrum power of such a sky model with the model power pattern of an MWA tile.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\linewidth]{fig14.eps}
+\caption{Similar to Figure~\ref{fig:noiseless-dsm-delay-spectrum} but for a uniform sky brightness model.
+The features bear a close resemblance to those obtained with the diffuse emission model.
+The edge--heavy {\it two--pronged} features arise from foreshortening of baselines which increases their sensitivity to emission from larger size scales from near the horizon.
+The symmetries around $\tau=0$ or lack thereof correspond to symmetries in the power patterns of the respective pointings (see Figure~\ref{fig:power_pattern}).
+\label{fig:noiseless-usm-delay-spectrum}}
+\end{figure}
+
+Although not identical, these delay spectra power bear a close resemblance to those obtained with the diffuse sky model (see Figure~\ref{fig:noiseless-dsm-delay-spectrum}).
+This indicates that the delay spectra power seen from diffuse foregrounds is primarily explained by the delay spectrum signatures of the underlying power pattern, with the additional variations coming from anisotropic deviations of the diffuse foreground model from a uniformly illuminated sky.
+
+The edge--heavy {\it two--pronged} features are evident, which are very similar to those seen earlier in the case of diffuse emission.
+As reasoned in \S\ref{sec:diffuse-long-baselines}, the foreshortening of long baselines causes them to be sensitive to emission from near the horizon.
+
+Branches of foreground power seen at positive delays and close to the negative horizon delay limit in the {\it off--zenith} pointing correspond to radiation received from the direction of the primary lobe and the far eastward sidelobe, respectively, in the power pattern.
+On the other hand, the symmetry of the delay spectrum around $\tau=0$ seen in the {\it zenith} pointing is as expected from a symmetric power pattern.
+
+\subsection{Experiments with Antenna Geometry}\label{sec:antgeom}
+
+\begin{figure*}[htb]
+\centering
+\subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{fig15a.eps}} \\
+\subfloat[][Delay spectra power]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{fig15b.eps}}
+\caption{}
+\label{fig:multi-antenna}
+\end{figure*}
 

### 2015-01-18 ms.tex
Completed section on role of antenna shapes and changed abstract, intro and summary accordingly

@@
 For signals entering through the primary field of view, compact emission dominates the foreground contamination.
-These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space.
-Based on these results, we propose that selective down--weighting of data based on antenna spacing vector and time will mitigate foreground contamination substantially (up to a factor $\sim 10$ with negligible loss of sensitivity) in redshifted 21~cm power spectrum analysis.
+These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space --- {\bf a signature generic to all wide--field measurements}.
+Based on these results, we propose that selective down--weighting of data based on antenna spacing vector and time will mitigate foreground contamination substantially (up to a {\bf factor $\sim 100$} with negligible loss of sensitivity) in redshifted 21~cm power spectrum analysis.
+{\bf A dish--shaped antenna yields the most desirable features in terms of foreground contamination, relative to a dipole or a phased array}.
 
@@
 In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain.
-Here, we will focus primarily on extending the {\it avoidance} strategy by identifying foreground components at greatest risk to ``leak'' from foreground modes to EoR modes and demonstrating a scheme for down--weighting these components.
+Here, we focus primarily on extending the {\it avoidance} strategy by identifying foreground components at greatest risk to ``leak'' from foreground modes to EoR modes and proposing a scheme for down--weighting these components.
 
@@
 
-After matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the {\bf shape of the delay spectrum} of each baseline.
+After matching observations from the Murchison Widefield Array with instrumental simulations of the entire sky, we examine the {\bf shape of the delay spectrum} of each baseline.
 We report two important findings: the foregrounds that most severely {\bf obscure} the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted.
-We quantify this contamination by separating the simulations into components based on type of foreground emission.
+We quantify these by separating the simulations into components based on type of foreground emission.
 We then arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.
+{\bf We also examine the role of antenna shapes in determining foreground contamination and provide inputs on antenna design for future EoR experiments}.
 
@@
 In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique.
+{\bf The role of antenna shapes in determining foreground contamination is explored in \S\ref{sec:antenna-shape}}.
 We present a summary of our work and findings in \S\ref{sec:summary}.
@@
 where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, and $k_\textrm{B}$ is the Boltzmann constant.
-Thus, $V_{u\eta}$ inferred from observations, also in units of Jy~Hz, can be converted into an equivalent cosmological H{\sc i} power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$, in units of K$^2$~Mpc$^3$.
+Thus, $V_{u\eta}$ inferred from observations, also in units of Jy~Hz, can be converted into an equivalent cosmological H{\sc i} power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$, in units of K$^2$~Mpc$^3$ or, more generally, K$^2$(Mpc/$h$)$^3$, where $h$ is the Hubble constant factor.
 
@@
 For our study, which primarily concerns with foreground characterization, we find the former to be simple and yet extremely useful, while maintaining a close correspondence with the latter despite subtle differences.
-In order to express a quantity derived from $V_{u\eta}$ that has units identical to that of aforementioned cosmological H{\sc i} power spectrum derived, we define:
+In order to express a quantity derived from $V_{u\eta}$ that has units identical to that of aforementioned cosmological H{\sc i} power spectrum, we define:
 \begin{align}\label{eqn:fake-power-spectrum}
@@
 At the same time, it is closely related to the sought power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$ containing critical information about spatial scales.
-For purposes of illustration later in this paper, we often use the quantities $|V_{b\tau}|$ and $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$.
+In this paper, we often use the quantity $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$, and sometimes $V_{b\tau}$.
 
@@
 Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
+{\bf Note that visibilities from different baselines have not been averaged together in this analysis}.
 
@@
 
-The {\it local sidereal time} (LST) at these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at {\bf 22.08~hours ($-1.92$~hours in the figure)} and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings, respectively.
+The {\it local sidereal time} (LST) at these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at {\bf 22.08~hours (wrapped to $-1.92$~hours in the figure)} and 0.09~hours, which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings, respectively.
 
@@
 This flattens any residual variation in bandpass and removes small phase slopes due to variations in cable delay.
-Finally, we fit for an additional phase known to be caused by small reflections in subset of cables.
+Finally, we fit for an additional phase known to be caused by small reflections in a subset of cables.
 
@@
 
-In subsequent sections, we plot the amplitude of the delay spectrum, $|V_{b\tau}|$, sorted by baseline length, $|\boldsymbol{b}|$.
+% In subsequent sections, we plot the amplitude of the delay spectrum, $|V_{b\tau}|$, sorted by baseline length, $|\boldsymbol{b}|$.
 Note that visibilities from different baselines have not been averaged together in this analysis.
 
-We show results of delay spectrum from MWA data after deconvolution along the delay axis in Figure~\ref{fig:fhd_data}.
+We show {\bf delay power spectra} from MWA data after deconvolution along the delay axis in Figure~\ref{fig:fhd_data}.
 The principle feature is the {\it foreground wedge}, whose boundaries, marked by white lines, increase linearly with light travel time corresponding to baseline length.
@@
 \includegraphics[width=\linewidth]{fig3.eps}
-\caption{Delay spectra power of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
+\caption{{\bf Delay power spectra} of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
 {\bf In both cases, the delay spectra are phased to the zenith and thus the {\it foreground wedge} boundaries are symmetric around $\tau=0$}.
@@
 where, $\Delta f = 80$~kHz, $\Delta t = 112$~s, and $N_\textrm{ch}=\Delta B/\Delta f$ is the number of frequency channels.
-The choice of threshold for $\tau$ is well outside the {\it foreground window}, where foreground contamination is negligible and thus makes a robust estimation of $T_\textrm{sys}$ possible.
+The choice of threshold for $\tau$ is well outside the {\it foreground window}, where foreground contamination is negligible and thus yields a robust estimate of $T_\textrm{sys}$.
 We find the average system temperature to be $\sim 95$~K across all frequency channels and across all antenna pairs throughout the course of the observation.
@@
 With the aforementioned foreground model, and instrumental and observational parameters, we simulate visibilities using equation~\ref{eqn:obsvis}.
-Figure~\ref{fig:sim_data} shows the amplitude of the delay spectrum from {\it off--zenith} and {\it zenith} pointings.
+Figure~\ref{fig:sim_data} shows the {\bf delay power spectra} from {\it off--zenith} and {\it zenith} pointings.
 Notice the qualitative agreement of amplitude and structure with those obtained from data shown in Figure~\ref{fig:fhd_data}.
@@
 The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observing frequency.
-Using simple error propagation, the fractional error in the delay spectrum {\bf amplitude} caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the MWA observing frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index.
+Using simple error propagation, the fractional error in the delay spectrum {\bf amplitude}, $|V_{b\tau}|$, caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the MWA observing frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index.
 From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for point sources from NVSS and SUMSS catalogs.
@@
 {\bf While the bin boundaries are arbitrary, they were so chosen to allow bin centers to be aligned along the cardinal and ordinal directions in one half--plane}.
-Since delays depend on $\theta_\textrm{b}$, binning by $\theta_\textrm{b}$ allows us to match delay spectrum features to different sky directions.
+Since delays depend on $\theta_\textrm{b}$, binning by $\theta_\textrm{b}$ allows us to match {\bf features of the delay power spectra} to different sky directions.
 
 \item {\it Tile Pointing and Power--Pattern}: Since the MWA tiles are steered electronically, the power--pattern of the tile changes in any observing mode that tracks the {\bf pointing center}.
-In our study, we take into account the effect of the changing tile power--pattern on the delay spectrum (Figure~\ref{fig:power_pattern}).
+In our study, we take into account the effect of the changing tile power--pattern on the {\bf delay power spectra} (Figure~\ref{fig:power_pattern}).
 The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose the {\it off--zenith} and {\it zenith} pointings shown in Figure~\ref{fig:pointings}.
@@
 
-Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the amplitude of delay spectrum, $|V_{b\tau}|$, simulated with only the diffuse foreground model.
+Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the {\bf delay power spectra} simulated with only the diffuse foreground model.
 The top and bottom panels correspond to {\it off--zenith} and {\it zenith} pointings, respectively.
@@
 \includegraphics[width=\linewidth]{fig8.eps}
-\caption{Simulated delay spectrum {\bf power (in units of K$^2$~(Mpc/$h$)$^3$} for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
+\caption{Simulated {\bf delay power spectra (in units of K$^2$~(Mpc/$h$)$^3$} for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
 White lines mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
 The axes and color scale are identical to those in Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}.
-In the {\it zenith} pointing, delay spectrum from diffuse emission has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings.
+In the {\it zenith} pointing, {\bf delay power spectrum} from diffuse emission has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings.
 \label{fig:noiseless-dsm-delay-spectrum}}
@@
 The most prominent signature seen in the {\it off--zenith} pointing (top panel of Figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power--pattern.
-It appears as a bright branch in the delay spectrum near the negative delay horizon delay limit.
+It appears as a bright branch in the {\bf delay power spectrum} near the negative delay horizon delay limit.
 This feature is strongest at short antenna spacings and fades with increasing antenna spacing.
@@
 {\bf In wide--field observations}, long baselines appear shortened due to projection effects for off--axis emission, especially towards the horizon.
-{\bf Consequently, they become sensitive to diffuse emission}.
-This effect is evident at all LSTs in our simulations and it naturally holds for short baselines, as well.
+{\bf Consequently, they become sensitive to diffuse emission on large size scales.
+This holds for short baselines as well and is evident at all LSTs in our simulations}.
 Thus, diffuse emission from far off--axis directions manifests as an edge--heavy {\it two--pronged fork} across all baselines.
@@
 This must be true for diffuse emission as well.
-Thus we argue the feature arises primarily due to an increase in sensitivity of long baselines to diffuse emission caused by foreshortening towards angles far off--axis, where the increased size scales (or solid angles) containing higher levels of integrated emission and spatial scales given by the inverse of foreshortened baseline vectors become matched}.
+Thus we argue the feature arises primarily due to an increase in sensitivity of long baselines to diffuse emission caused by foreshortening towards angles far off--axis, where the increased size scales (or solid angles) containing higher levels of integrated emission, and spatial scales proportional to the inverse of foreshortened baseline lengths become matched}.
 
@@
 
-Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the delay spectrum of the compact foreground sky model (see Figure~\ref{fig:CSM}) for {\it off--zenith} and {\it zenith} pointings in the top and bottom panels, respectively.
-In contrast to the delay spectrum of diffuse emission, compact emission manifests as a center--heavy structure in the delay spectrum with either pointing.
+Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the {\bf delay power spectra} of the compact foreground sky model (see Figure~\ref{fig:CSM}) for {\it off--zenith} and {\it zenith} pointings in the top and bottom panels, respectively.
+In contrast to the {\bf delay power spectra} of diffuse emission, compact emission manifests as a center--heavy structure in either pointing.
 
@@
 
-When not dominated by the bright emission from the Galactic center, which happens only for certain pointings at certain times, the delay spectrum of the combined foreground model is composed of diffuse and compact emission, in roughly equal measure.
+When not dominated by the bright emission from the Galactic center, which happens only for certain pointings at certain times, the {\bf delay power spectrum} of the combined foreground model is composed of diffuse and compact emission, in roughly equal measure.
 This is illustrated by a more detailed examination of the {\it zenith} pointing in our study.
 
-Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional delay spectrum plots shown in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum}, and \ref{fig:noiseless-csm-delay-spectrum}.
+Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional {\bf delay power spectra} plots shown in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum}, and \ref{fig:noiseless-csm-delay-spectrum}.
 The diffuse, compact, and composite components are shown as solid red, cyan, and black lines, respectively.
 The horizon delay limits are shown as a pair of vertical dotted lines.
-The gray shaded area denotes the envelope of expected uncertainty in the delay spectrum.
+The gray shaded area denotes the envelope of expected uncertainty in the {\bf delay power spectrum}.
 Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the % $\gtrsim$~70\% error owing to
@@
 \includegraphics[width=\linewidth]{fig10.eps}
-\caption{Simulated delay spectrum power for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}).
+\caption{Simulated {\bf delay power spectra} for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}).
 The antenna spacing is specified in each panel along East, North and Upward directions.
@@
 Vertical dotted lines mark the horizon delay limits.
-{\bf Compact emission dominates the central regions of delay spectrum while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a {\it pitchfork}--shaped structure.
-The shaded region denotes the envelope of uncertainty around the composite model delay spectrum.
+{\bf Compact emission dominates the central regions of the delay power spectra while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a {\it pitchfork}--shaped structure.
+The shaded region denotes the envelope of uncertainty around the composite model delay power spectrum.
 This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits.
@@
 
-The zero--delay peak (corresponding to the main lobe in the power--pattern) with a value of $\sim 10^8$--$10^9$~Jy~Hz, independent of antenna spacing, is predominantly determined by compact emission.
-The corresponding peak at zero--delay from diffuse foreground model is at least 10 times fainter and decreases drastically with an increase in antenna spacing.
+The zero--delay peak (corresponding to the main lobe in the power--pattern) with a value of $\sim 10^7$--$10^8$~K$^2$(Mpc/$h$)$^3$, independent of antenna spacing, is predominantly determined by compact emission.
+The corresponding peak at zero--delay from diffuse foreground model is $\sim 10^3$ times fainter and decreases rapidly with increase in antenna spacing.
 This is the response expected from different antenna spacings towards compact and diffuse emission.
@@
 Here, diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at zero--delay.
-In fact, even on widely spaced antennas, {\emph diffuse emission in the delay spectrum near the horizon delay limits exceeds that in the main lobe by more than two orders of magnitude}.
+In fact, even on widely spaced antennas, diffuse emission in the {\bf delay power spectrum} near the horizon delay limits exceeds that in the main lobe by about three orders of magnitude.
 We attribute this to apparent shortening of antenna spacings projected in the direction of diffuse emission entering far from the primary field of view, thereby giving rise to the edge--heavy {\it two--pronged fork} feature discussed in \S\ref{sec:diffuse-long-baselines}.
@@
 The observability of the {\it pitchfork} signature predicted in this paper depends on the relative levels of uncertainty in the foreground model and fluctuations from thermal noise.
-In our simulations, since thermal noise in these very short duration snapshots is $\gtrsim 10^7$~Jy~Hz, and features near the horizon delay limits are $\sim 10^7$~Jy~Hz, the {\it pitchfork} feature is not expected to be detected in a noisy scenario although this feature is marginally visible in the {\it zenith} pointing of observed data (see Figure~\ref{fig:fhd_data}).
+In our simulations, since thermal noise in these very short duration snapshots is $\sim 10^4$~K$^2$(Mpc/$h$)$^3$ and features near the horizon delay limits are also of comparable amplitudes, the {\it pitchfork} feature is not expected to be detected in a noisy scenario although this feature is marginally visible in the {\it zenith} pointing of observed data (see Figure~\ref{fig:fhd_data}).
 We attribute this to differences between our foreground model and the actual sky.
@@
 We also note that increasing the antenna spacing results in progressively improving the resolution along the delay axis by increasing the number of delay bins inside the {\it foreground wedge}.
-This improves the localization of foreground objects whose signatures are imprinted in the delay spectrum.
-For instance, there is an increase in the number of secondary peaks in the delay spectrum between zero--delay and horizon delay limits as the antenna spacing increases from $\sim$84~m to $\sim$171~m.
+This improves the localization of foreground objects whose signatures are imprinted in the {\bf delay power spectrum}.
+For instance, there is an increase in the number of secondary peaks in the {\bf delay power spectrum} between zero--delay and horizon delay limits as the antenna spacing increases from $\sim$84~m to $\sim$171~m.
 In this case, these correspond to secondary lobes of the power--pattern that lie between the primary lobe and the horizon along the local meridian.
@@
 
-In summary, the brightest signature ($\gtrsim 10^{10}$~Jy~Hz) is that of the Galactic center in the {\it off--zenith} pointing co--located at a westward secondary lobe with a significantly high gain.
-The next brightest signature ($\sim 10^8$--$10^9$~Jy~Hz) is caused by compact emission appearing to be concentrated in the inner regions of the {\it foreground wedge} (near zero delays) rather than towards the boundaries.
-Diffuse emission co--located at the primary lobe of the power--pattern is fainter by approximately an order of magnitude relative to compact emission from the same region for a $\sim 84$~m antenna spacing.
-But unlike the latter, diffuse emission decreases rapidly by over two orders of magnitude as the antenna spacing is widened to $\sim 171$~m.
+In summary, the brightest feature $\sim 10^{12}$~K$^2$(Mpc/$h$)$^3$ is that of the Galactic center in the {\it off--zenith} pointing co--located at a westward secondary lobe with a significantly high gain.
+The next brightest signature $\gtrsim 10^8$~K$^2$(Mpc/$h$)$^3$ is caused by compact emission appearing to be concentrated in the inner regions of the {\it foreground wedge} (near zero delays) rather than towards the boundaries.
+Diffuse emission co--located at the primary lobe of the power--pattern is $\sim 10^3$ fainter relative to compact emission from the same region for a $\sim 84$~m antenna spacing.
+But unlike the latter, diffuse emission decreases rapidly by over four orders of magnitude as the antenna spacing is widened to $\sim 171$~m.
 However, diffuse emission is significantly higher near horizon delay limits compared to that in the primary lobe of the power--pattern as the antenna spacing is increased.
-This leads to edge--heavy features in the delay spectrum.
-Complemented by the center-heavy compact foreground features, we predict the presence of a {\it three--pronged pitchfork}--shaped signature in the delay spectrum of the foreground sky.
+This leads to edge--heavy features in the {\bf delay power spectra}.
+Complemented by the center-heavy compact foreground features, we predict the presence of a {\it three--pronged pitchfork}--shaped signature in the {\bf delay power spectrum} of the foreground sky.
 We identify evidence for presence of this signature in actual MWA data, although deeper observations are needed to more clearly reveal the effect.
@@
 Figure~\ref{fig:baseline-breakup} shows the delay spectra on baselines oriented northward ($67.5\arcdeg\le \theta_\textrm{b} < 112.5\arcdeg$) at the top and eastward ($-22.5\arcdeg~\le \theta_\textrm{b} < 22.5\arcdeg$) at the bottom.
-The Galactic center manifests itself most distinctly near the negative horizon delay limit on short eastward baselines in the delay spectrum (bottom panel of Figure~\ref{fig:baseline-breakup}).
+The Galactic center manifests itself most distinctly near the negative horizon delay limit on short eastward baselines in the {\bf delay power spectrum} (bottom panel of Figure~\ref{fig:baseline-breakup}).
 Consequently, the spillover caused by the instrument's spectral transfer function from the {\it foreground wedge} into the {\it EoR window} affects the northward baselines the least and is most severe in eastward baselines (particularly the short ones) evident by the bright vertical stripes of foreground contamination.
@@
 
-The delay spectrum, $V_{b\tau}$, not only carries information on spatial scales of foreground emission but also offers the unique advantage of viewing the sky through a combination of antenna spacing vectors and the delay axis.
+The {\bf delay power spectrum} not only carries information on spatial scales of foreground emission but also offers the unique advantage of viewing the sky through a combination of antenna spacing vectors and the delay axis.
 With a foreground model known {\it a priori} in which structures and locations of very bright foreground objects such as the Galactic center or AGN are available, we can predict their response across antenna spacings as a function of observing parameters such as LST, power--pattern, etc.
-This allows us to programmatically screen data for antenna spacings that are severely contaminated by foregrounds near the horizon delay limits in the delay spectrum.
+This allows us to programmatically screen data for antenna spacings that are severely contaminated by foregrounds near the horizon delay limits in the {\bf delay power spectrum}.
 These can be weighted appropriately during data analysis.
@@
 The fraction of discarded antenna spacings discarded relative to the total number is shown in dotted lines for different ranges of $\theta_\textrm{b}$.
-It is seen that foreground contamination can be mitigated by a factor between $\sim 1.5$ ($|\theta_\textrm{b}|\le$~7.5\arcdeg) and $\sim 10$ ($|\theta_\textrm{b}|\le$~22.5\arcdeg).
+It is seen that foreground contamination can be mitigated by a factor between $\sim 2$ ($|\theta_\textrm{b}|\le$~7.5\arcdeg) and $\sim 100$ ($|\theta_\textrm{b}|\le$~22.5\arcdeg).
 The latter limit is achieved with a mere 5\% loss of data for $|\boldsymbol{b}|_\textrm{max}\simeq30$~m.
@@
 
-Having examined the role the foregrounds play in determining the delay spectrum obtained with an interferometer array, we turn our attention to another crucial factor, namely, the power pattern of the antenna (refer to equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}).
+Having examined the role foregrounds play in determining the delay power spectrum obtained with an interferometer array, we turn our attention to another crucial factor, namely, the power pattern of the antenna (refer to equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}).
+Unlike the flagging of the edges of the coarse sub--bands described earlier in this paper, we assume an ideal flat passband over the entire 30.72~MHz bandwidth while still applying a {\it Blackman--Harris} window function.
+
+\subsection{Uniformly Illuminated Sky}\label{sec:USM}
+
 In the following discussion, we assume there are no random delay fluctuations in the MWA tile beamforming process.
-The model power pattern so obtained is different from the one we used earlier, but for the present discussion it still serves as a good reference (Figure~\ref{fig:power_pattern}).
-Since the previous model suffices, we do not plot the new power pattern model obtained with ideal beamforming.
-We also assume an ideal flat bandpass shape over the 30.72~MHz bandwidth unlike the flagging of the edges of the coarse sub--bands described earlier in this paper, while still applying a {\it Blackman--Harris} window function.
-
-\subsection{Uniformly Illuminated Sky}\label{sec:USM}
+The model power pattern so obtained is different from the one we used earlier, but the former still serves as a good reference (Figure~\ref{fig:power_pattern}) for the present discussion.
+Hence, we do not plot the new power pattern model obtained with ideal beamforming.
 
@@
 This helps isolate the signatures of the antenna power pattern from those of the foregrounds.
-Figure~\ref{fig:noiseless-usm-delay-spectrum} shows the delay spectrum power of such a sky model with the model power pattern of an MWA tile.
+
+Figure~\ref{fig:noiseless-usm-delay-spectrum} shows the delay power spectra of such a sky model with the model power pattern of an ideal MWA tile.
+Although not identical, these delay power spectra bear a close resemblance to those obtained with the diffuse sky model (see Figure~\ref{fig:noiseless-dsm-delay-spectrum}).
+This indicates that the delay power spectra seen from diffuse foregrounds is primarily explained by those of the underlying power pattern, with any additional variations arising out of anisotropic deviations of the diffuse foreground model from a uniformly illuminated sky.
 
@@
 
-Although not identical, these delay spectra power bear a close resemblance to those obtained with the diffuse sky model (see Figure~\ref{fig:noiseless-dsm-delay-spectrum}).
-This indicates that the delay spectra power seen from diffuse foregrounds is primarily explained by the delay spectrum signatures of the underlying power pattern, with the additional variations coming from anisotropic deviations of the diffuse foreground model from a uniformly illuminated sky.
-
 The edge--heavy {\it two--pronged} features are evident, which are very similar to those seen earlier in the case of diffuse emission.
@@
 Branches of foreground power seen at positive delays and close to the negative horizon delay limit in the {\it off--zenith} pointing correspond to radiation received from the direction of the primary lobe and the far eastward sidelobe, respectively, in the power pattern.
-On the other hand, the symmetry of the delay spectrum around $\tau=0$ seen in the {\it zenith} pointing is as expected from a symmetric power pattern.
-
-\subsection{Experiments with Antenna Geometry}\label{sec:antgeom}
+On the other hand, the symmetry of the delay power spectrum around $\tau=0$ seen in the {\it zenith} pointing is as expected from a symmetric power pattern.
+
+\subsection{Experiments with Aperture Shapes}\label{sec:apertures}
+
+The power pattern of an antenna is determined by the shape of its aperture.
+Through the use of a few examples, we examine the role the geometrical shape of the aperture plays in shaping the characteristics of delay power spectrum.
+We consider the following antenna elements placed at the MWA tile locations:
+\begin{itemize}
+\item {\it Dipole}: An East--West dipole of length 0.74~m at a height 0.3~m above a ground plane. $A_\textrm{e}=(\lambda/2)^2$.
+\item {\it Phased Array}: A 4$\times$4 array of isotropic radiators at a height 0.3~m above the ground plane placed in an arrangement similar to that in an MWA tile. $A_\textrm{e}=16\,(\lambda/2)^2$. 
+\item {\it Dish}: Diameter of 14~m similar to that proposed for HERA, with $A_\textrm{e}\approx 154$~m$^2$.
+The power pattern is simulated using an {\it Airy} pattern where its sensitivity beyond the horizon is forced to zero.
+\end{itemize}
+
+The power patterns of these antenna geometries at 185~MHz for the {\it zenith} pointing are shown in Figure~\ref{fig:power-patterns}.
+It must be noted that the presence of a ground plane below the dipole and the phased array introduces a smoother suppression of sensitivity near the horizon, whereas this transition is abrupt in the case of the dish.
 
@@
 \subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{fig15a.eps}} \\
-\subfloat[][Delay spectra power]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{fig15b.eps}}
-\caption{}
-\label{fig:multi-antenna}
+\subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{fig15b.eps}}
+\caption{Power patterns (top panels) and simulated delay power spectra (bottom panels) for different antenna shapes at 185~MHz centered on zenith.
+Antenna shapes used are: dipole (left), phased array (middle), and dish (right).
+Refer to text for details.
+The strength and occupancy of the power patterns are correlated with those of delay power spectra.
+The {\it foreground wedge} and the {\it EoR window} are most severely contaminated in the case of the dipole while it is the least for the dish.
+The phased array has intermediate levels of contamination.
+Foreground power close to the horizon delay limits in all three cases are significant even on long baselines.
+The foreshortening of baselines towards the horizon makes them sensitive to foreground emission on large size scales.
+The amplitude of this feature is highest for the dipole and least for the dish and hence, strongly depends on the shape of the antenna element.}
+\label{fig:aperture-shapes}
 \end{figure*}
+
+The delay power spectra for these antenna shapes are shown in Figure~\ref{fig:antennas-delay-spectra}.
+The occupancy of the power patterns in the sky hemisphere is clearly correlated with that in the delay spectra.
+Moreover, the following visual correlations are also apparent --- the strength of the primary lobe centered on the pointing center is correlated with the delay power spectrum centered on $\tau=0$; and, the overall rate of decrease in the power sensitivity away from the pointing center is correlated with the rate of drop in power away from $\tau=0$.
+
+The levels of foreground contamination in the {\it EoR window} varies substantially across the different antenna shapes: $\sim 10^4\,$K$^2\,$(Mpc/$h$)$^3$, $\lesssim 10^2\,$K$^2\,$(Mpc/$h$)$^3$, and $<1\,$K$^2\,$(Mpc/$h$)$^3$ while using a dipole, phased array, and a dish respectively.
+The severity of foregrounds inside the {\it foreground wedge} both in strength and occupancy also evidently decreases as the antenna element is changed from a dipole to a phased array to a dish.
+For instance, notice that the foreground contamination in $k$--modes between the $k_\parallel=0$ and the horizon limits decreases from $\sim 10^5$~K$^2\,$(Mpc/$h$)$^3$ in a phased array to $\sim 10$~K$^2\,$(Mpc/$h$)$^3$ in a dish.
+As a consequence, $k$--modes in the {\it foreground wedge} that may be deemed too contaminated for EoR studies in the case of a dipole or a phased array can potentially become accessible when using a dish.
+
+Finally, a distinct feature common to all these aperture shapes is that the foreground contamination near the horizon delay limits even on wide antenna spacings is significant ($\gtrsim 10^5$~K$^2$(Mpc/$h$)$^3$).
+This is contrary to expectation for two reasons --- the power pattern of the antenna near the horizon is negligible compared to the primary lobe region; and, a significant fraction of foreground emission near the horizon delay limits on wide antenna spacings arises from diffuse emission, to which usually only short antenna spacings are sensitive.
+We have argued that baselines are foreshortened towards the horizon and thus become sensitive to large size scales, thereby leading to significant integrated emission in delay bins close to the horizon limits.
+The prevalence of this feature across different antenna shapes demonstrates it is generic to all wide--field measurements, especially in EoR experiments.
+The amplitude of this effect, however, can be controlled via choice of antenna shape and through weighting of aperture illumination.
 
@@
 The bright Galactic center at the edge of the western horizon co--located with one of the far secondary lobes of MWA tile power--pattern is the brightest source of foreground contamination in the {\it off--zenith} pointing.
-It manifests itself near the negative horizon delay limit in the delay spectrum on antenna spacings oriented eastward.
+It manifests itself near the negative horizon delay limit in the {\bf delay power spectrum} on antenna spacings oriented eastward.
 
 As expected, diffuse emission in the primary beam of the antenna power--pattern is prominent on shorter antenna spacings.
-However, the most interesting result is its footprint on wide antenna spacings near the horizon delay limits, an edge--heavy {\it two--pronged fork}--shaped signature.
-This is due to apparent shortening of wide antenna spacings in the direction of foreground emission far off--axis, thereby retaining their response to extended emission.
+However, the most interesting result is its footprint on wide antenna spacings near the horizon delay limits --- an edge--heavy {\it two--pronged fork}--shaped signature.
+This is due to apparent shortening of wide antenna spacings in the direction of foreground emission far off--axis, thereby retaining their response to {\bf emission on large size scales}.
 On the other hand, compact emission predominantly maps onto central regions of the {\it foreground wedge}.
 Features arising from compact emission co--located with primary and secondary lobes of the antenna power--pattern have been identified.
-In general, delay spectrum signatures of compact emission are center--heavy, in clear contrast to those from diffuse emission.
+In general, {\bf delay power spectrum} signatures of compact emission are center--heavy, in clear contrast to those from diffuse emission.
 A composite foreground model consisting of diffuse and compact foregrounds combines the characteristic individual shapes into a {\it pitchfork}--shaped structure in the {\it foreground wedge}.
@@
 
+The role of antenna shape in determining the {\bf delay power spectrum} was also studied.
+We conclude the {\it pitchfork} feature to be generic to all wide--field measurements, especially EoR experiments.
+While comparing a dipole, a phased array such as an MWA tile, and a dish such as the one proposed for HERA, the dish is found to yield the least foreground contamination both in the {\it foreground wedge} and the {\it EoR window} and hence, the most desirable for EoR studies.
+
+We also provide a simple and effective tool based on the delay spectrum technique {\bf that can potentially} mitigate foreground contamination by {\bf nearly two orders} of magnitude in EoR data analysis by discarding or down--weighting data from antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation, with negligible loss of sensitivity.
 In conclusion, we find that inclusion of emission models, both diffuse and compact, all the way to the horizon is essential to explaining the observed power spectrum.
-We also provide a simple and effective tool based on the delay spectrum technique {\bf that can potentially} mitigate foreground contamination by an order of magnitude in EoR data analysis by discarding or down--weighting data from antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation, with negligible loss of sensitivity.
 

### 2015-01-19 ms.tex
Addressed most of the referee comments and made lots of minor changes throughout

@@
 
-Detection of 21~cm emission of neutral hydrogen from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission.
+Detection of 21~cm emission of H{\sc i} from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission.
 We investigate the chromatic signatures of an all--sky foreground model imprinted by an instrumental transfer function.
-Using the delay spectrum technique {\bf that maps the Fourier transform of the measurements to locations of foreground objects on the celestial sphere through delays in signals received by antenna pairs}, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view.
-Comparing data from recent Murchison Widefield Array observations with simulations separated into components based on type of emission, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings {\bf which are usually not sensitive to diffuse emission}.
+Using the delay spectrum technique {\bf that maps the measurements to foreground object locations through signal delays between antenna pairs}, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view.
+Comparing data from recent Murchison Widefield Array observations with simulations {\bf broken down by types of emission}, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings {\bf which are usually not sensitive to diffuse emission}.
 For signals entering through the primary field of view, compact emission dominates the foreground contamination.
-These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space --- {\bf a signature generic to all wide--field measurements}.
-Based on these results, we propose that selective down--weighting of data based on antenna spacing vector and time will mitigate foreground contamination substantially (up to a {\bf factor $\sim 100$} with negligible loss of sensitivity) in redshifted 21~cm power spectrum analysis.
-{\bf A dish--shaped antenna yields the most desirable features in terms of foreground contamination, relative to a dipole or a phased array}.
+These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space, {\bf which is found to be generic to wide--field measurements}.
+Based on these results, we propose that selective down--weighting of data based on antenna spacing and time {\bf can} mitigate foreground contamination substantially (up to a {\bf factor $\sim 100$} with negligible loss of sensitivity).
+{\bf A dish--shaped antenna yields the most desirable features from a foreground contamination viewpoint, relative to a dipole or a phased array}.
 
@@
 % A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide--field (and chromatic) response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum.
-The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}).
+The conservative foreground strategy, referred to as {\it avoidance}, that has developed alongside this work is to discard $k$--modes which could be contaminated \citep[for e.g.,][]{par14}.
 The more aggressive alternative is to subtract a sky model and regain access to modes discarded by {\it avoidance}.
@@
 Interferometer array data known as {\it visibilities}, $V_{bf}\equiv V_{bf}(\boldsymbol{b},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis.
-If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power--pattern, respectively, at different frequencies as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_f\equiv W_f(f)$ denotes instrumental bandpass weights, then $V_{bf}$ can be written as (with a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}):
+If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power pattern, respectively, at different frequencies as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_f\equiv W_f(f)$ denotes instrumental bandpass weights, then $V_{bf}$ can be written as (with a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}):
 \begin{align}\label{eqn:obsvis}
@@
 where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector.
-We wish to {\bf note} that this equation is valid {\bf even for wide--field applications where the sky cannot be approximated to be flat}.
+We wish to {\bf note} that this equation is valid {\bf even for wide--field measurements where the sky cannot be approximated to be flat}.
 
@@
 However, they are not quite identical to each other.
-The difference basically arises from the fact noted above that $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not.
+The difference is because $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not.
 \citet{par12b} and \citet{liu14a} have discussed the mathematical correspondence between the two.
@@
 For our study, which primarily concerns with foreground characterization, we find the former to be simple and yet extremely useful, while maintaining a close correspondence with the latter despite subtle differences.
-In order to express a quantity derived from $V_{u\eta}$ that has units identical to that of aforementioned cosmological H{\sc i} power spectrum, we define:
+In order to express a quantity derived from $V_{u\eta}$ that has units identical to that of the cosmological H{\sc i} power spectrum, we define delay power spectrum:
 \begin{align}\label{eqn:fake-power-spectrum}
@@
 At the same time, it is closely related to the sought power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$ containing critical information about spatial scales.
-In this paper, we often use the quantity $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$, and sometimes $V_{b\tau}$.
+% In this paper, we often use the quantity $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$, and sometimes $V_{b\tau}$.
 
@@
 
-\subsection{Observation Parameters}\label{sec:obsparms}
+\subsection{EoR Observations}\label{sec:obsparms}
+
+{\bf The MWA is expected to be sensitive to the power spectrum of the H{\sc i} signal over the redshift range $6<z<10$ \citep{thy13,bea13}.
+Over 600~hours have been currently observed using the MWA with the objective of characterizing the nature of the sources that are responsible for ionizing the IGM, charting the evolution of the global neutral fraction, and probing the nature of quasar emissions by constraining the properties of their ionized proximity zones.}
 
@@
 It has a flux density limit of 25~mJy and a declination range of $-12\arcdeg$~ to $-40\arcdeg$~ evenly covering the field of view of the observations reported here.
-The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern.
+The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power pattern.
 The calibration algorithm -- based on forward modeling software by \citet{sul12} and the calibration method described by \citet{sal14} -- computes complex gain solutions per channel per antenna averaged to two minute intervals.
@@
 
-We show {\bf delay power spectra} from MWA data after deconvolution along the delay axis in Figure~\ref{fig:fhd_data}.
-The principle feature is the {\it foreground wedge}, whose boundaries, marked by white lines, increase linearly with light travel time corresponding to baseline length.
+Figure~\ref{fig:fhd_data} shows {\bf delay power spectra} from MWA data after deconvolution along the delay axis.
+The principal feature is the {\it foreground wedge}, whose boundaries, marked by white lines, increase linearly with light travel time proportional to baseline length.
 Note that the wedge is not filled out uniformly.
-Each baseline has a different orientation and resulting response to different objects in the sky.
+Each baseline has a different orientation that determines the response to different objects in the sky.
 It is this variation that we will explore below.
@@
 Consequently, the contamination into the {\it EoR window} is also higher, especially on short antenna spacings.
-Emission in the {\it zenith} pointing is more centrally concentrated {\bf due to a zenith--centered power--pattern and faster suppression of response away from zenith, whereas the {\it off--zenith} pointing has a power--pattern phased eastward of zenith and is also more responsive to emission farther away from the primary lobe relative to that in the {\it zenith} pointing.
-The power pattern is discussed in detail in \S\ref{sec:instrument_model} and Figure~\ref{fig:power_pattern}}.
+Emission in the {\it zenith} pointing is more centrally concentrated {\bf due to a zenith--centered power pattern and faster suppression of response away from zenith}, whereas the {\it off--zenith} pointing has a power pattern phased eastward of zenith and is also more responsive to emission farther away from the primary lobe relative to that in the {\it zenith} pointing.
+{\bf The power pattern is discussed in detail in \S\ref{sec:instrument_model}}.
 
@@
 
-The instrument model consists primarily of the tile beam power--pattern, $A(\hat{\boldsymbol{s}},f)$ (see equation~\ref{eqn:obsvis}).
-It is modeled as a mutually--coupled 4--by--4 dipole array with the overall power--pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14_arXiv}.
+The instrument model consists primarily of the tile power pattern, $A(\hat{\boldsymbol{s}},f)$ (see equation~\ref{eqn:obsvis}).
+It is modeled as a mutually--coupled 4--by--4 dipole array with the overall power pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14_arXiv}.
 To speed up simulations, we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles.
 We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known repeatability and stability level of the analog signal chain \citep{bow07b}.
-Besides having the effect of adding a time--dependent uncertainty in the power pattern, these random delay fluctuations reduce the coherence in the phased addition of dipole signals resulting in deviations from predicted models of the power--pattern, most prominently at its nulls.
-The power--patterns so obtained are shown in Figure~\ref{fig:power_pattern} for the pointings chosen in our study.
+Besides having the effect of adding a time--dependent uncertainty in the power pattern, these random delay fluctuations reduce the coherence in the phased addition of dipole signals resulting in deviations from predicted models of the power pattern, most prominently at its nulls.
+The power patterns so obtained are shown in Figure~\ref{fig:power_pattern} for the pointings chosen in our study.
 
@@
 \includegraphics[width=\linewidth]{fig4.eps}
-\caption{Models of MWA tile power--pattern for {\it off--zenith} (top) and {\it zenith} pointings (bottom) at 185~MHz.
+\caption{Models of MWA tile power pattern for {\it off--zenith} (top) and {\it zenith} pointings (bottom) at 185~MHz.
 An MWA tile is modeled as a 4$\times$4 array of isotropic radiators placed 0.3~m above an infinite ground plane.
@@
 A low resolution version of the diffuse foreground model used is shown in Figure~\ref{fig:DSM}.
-Contours of the MWA tile power--pattern shown in Figure~\ref{fig:power_pattern} are also overlaid.
+Contours of the MWA tile power pattern shown in Figure~\ref{fig:power_pattern} are also overlaid.
 Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the westward sky, where the MWA tile power gain is significant ($\gtrsim 12$\%).
-In the {\it zenith} pointing, the Galactic plane has set and the power--pattern in that direction is at least 16 times smaller.
+In the {\it zenith} pointing, the Galactic plane has set and the power pattern in that direction is at least 16 times smaller.
 
@@
 The color scale on the right is logarithmic and is common to both panels.
-Contours of power--pattern shown in Figure~\ref{fig:power_pattern} are overlaid.
+Contours of power pattern shown in Figure~\ref{fig:power_pattern} are overlaid.
 The contour levels shown are 0.00195, 0.00781, 0.0312, 0.125, and 0.5.
@@
 The color scale on the right is logarithmic and is common to both panels and corresponds to the color--coded filled circles.
-Contours of power--pattern shown in Figure~\ref{fig:power_pattern} are overlaid.
+Contours of power pattern shown in Figure~\ref{fig:power_pattern} are overlaid.
 The contour levels shown are 0.001953125, 0.0078125, 0.03125, 0.125, and 0.5.
@@
 We estimate the ratio of delay spectra from data and simulations as $\rho = |V^\textrm{D}_{b\tau}(\boldsymbol{b},\tau)|\,/\,|V^\textrm{S}_{b\tau}(\boldsymbol{b},\tau)|$, where superscripts D and S denote data and simulation, respectively.
-
 The median absolute deviation of $\log_{10}\rho$ inside the {\it foreground wedge} for both pointings is $\approx 0.28$.
 This corresponds to $\sim$~90\% fractional difference between data and modeling on average with either pointing.
+
 We also simulated delay spectra after assigning spectral indices drawn randomly from a {\it gaussian} distribution with a mean of $\langle\alpha_\textrm{sp}\rangle=-0.83$ and a HWHM of $\Delta\alpha_\textrm{sp}=0.35$ to the point sources in our compact foreground model.
@@
 This demonstrates that a fractional deviation of $\sim$~90\% observed between data and simulations is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account.
-Currently, we are significantly limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.
+Currently, we are primarily limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.
 
@@
 These estimates are conservative.
-A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power--pattern \citep{ber14}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide--field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations, but is beyond the scope of this paper.
+A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power pattern \citep{ber14}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide--field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations, but is beyond the scope of this paper.
 {\bf Hereafter}, our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.
@@
 The bin centers are oriented towards South--East, East, North--East, and North respectively.
-{\bf While the bin boundaries are arbitrary, they were so chosen to allow bin centers to be aligned along the cardinal and ordinal directions in one half--plane}.
+{\bf While the bin boundaries are arbitrary, they were so chosen to allow bin centers to be aligned along the cardinal and ordinal directions in a half--plane}.
 Since delays depend on $\theta_\textrm{b}$, binning by $\theta_\textrm{b}$ allows us to match {\bf features of the delay power spectra} to different sky directions.
 
-\item {\it Tile Pointing and Power--Pattern}: Since the MWA tiles are steered electronically, the power--pattern of the tile changes in any observing mode that tracks the {\bf pointing center}.
-In our study, we take into account the effect of the changing tile power--pattern on the {\bf delay power spectra} (Figure~\ref{fig:power_pattern}).
+\item {\it Tile Pointing and Power Pattern}: Since the MWA tiles are steered electronically, the tile power pattern tile changes in any observing mode that tracks the {\bf pointing center}.
+In our study, we take into account the effect of the changing tile power pattern on the {\bf delay power spectra} (Figure~\ref{fig:power_pattern}).
 The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose the {\it off--zenith} and {\it zenith} pointings shown in Figure~\ref{fig:pointings}.
-We will show the secondary lobe structures in the tile power--pattern are tied to the foreground signatures in Fourier space.
+% We will show the secondary lobe structures in the tile power pattern are tied to the foreground signatures in Fourier space.
 
@@
 With the addition of noise in the visibilities, some of the weaker features may not be as prominently visible.
-Since the foreground signatures are far too numerous and they reside in a vast volume of parameter space such as antenna spacing length and orientation, power--pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features observed in the delay spectra of foreground emission.
+Since the foreground signatures are far too numerous {\bf and subject to a multitude of parameters like} baseline length and orientation, power pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features in the delay spectra of foreground emission.
 
@@
 
-The most prominent signature seen in the {\it off--zenith} pointing (top panel of Figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power--pattern.
-It appears as a bright branch in the {\bf delay power spectrum} near the negative delay horizon delay limit.
+The most prominent signature seen in the {\it off--zenith} pointing (top panel of Figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power pattern.
+It appears as a bright branch near the negative delay horizon delay limit.
 This feature is strongest at short antenna spacings and fades with increasing antenna spacing.
@@
 
-Figure~\ref{fig:noiseless-dsm-delay-spectrum} also demonstrates diffuse emission from within the primary field of view manifests in the {\it off--zenith} pointing as a branch at the top of the wedge, and at zero delay in the {\it zenith} pointing.
-The former is seen at positive delays because the primary lobe of the power--pattern is centered eastward of zenith, whereas in the latter it is centered at zenith.
+Figure~\ref{fig:noiseless-dsm-delay-spectrum} {\bf also shows diffuse emission in the primary field of view manifesting in the {\it off--zenith} pointing as a branch at $\tau>0$, and at $\tau=0$ in the {\it zenith} pointing.
+The former is seen at $\tau>0$ because the primary lobe of the power pattern is centered eastward of zenith, whereas in the latter it is centered at zenith}.
 As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inversely proportional to the length of the baseline projected in the direction of the emission.
-Thus, in the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at zero delay, fades away on antenna spacings $|\boldsymbol{b}| \gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0.75\arcdeg.
-In the off-zenith pointing, we see power out to much longer baselines which is described below.
+Thus, in the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at $\tau=0$, fades away on antenna spacings $|\boldsymbol{b}| \gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0.75\arcdeg.
 
@@
 
-A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\boldsymbol{b}|\le$~200~m).
-This is contrary to expectations because these longer baselines are not expected to be sensitive to diffuse emission (on scales $\gtrsim$~0.85\arcdeg) of our input foreground model.
-{\bf In wide--field observations}, long baselines appear shortened due to projection effects for off--axis emission, especially towards the horizon.
-{\bf Consequently, they become sensitive to diffuse emission on large size scales.
+{\bf In typical radio interferometric measurements, due to the Fourier relation between sky brightness and measured visibilities, only short antenna spacings are sensitive to emission from large size scales such as the diffuse component in our study (angular scales $\gtrsim$~0.85\arcdeg).
+On the contrary, we note a very interesting signature of diffuse emission revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing, and at both horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\boldsymbol{b}|\le$~200~m).
+This is a consequence of wide--field measurements, where long baselines are foreshortened due to projection effects towards off--axis emission, especially near the horizon.
+Consequently, they become sensitive to diffuse emission on large size scales.
 This holds for short baselines as well and is evident at all LSTs in our simulations}.
@@
 In the latter, this feature is described as ``edge brightening'' and attributed to the Galactic plane near the horizon.
-From their discussion, it is unclear what fraction of power in that feature arises from ubiquitous diffuse emission unlike in our {\it zenith} pointing where the Galactic plane is not expected to play a significant role.
+From their discussion, it is unclear what fraction of power in that feature arises from ubiquitous diffuse emission.
 On the other hand, \citet{thy13} statistically modeled point sources below the confusion limit and attributed the feature to an increase in solid angles subtended by delay bins closer to the horizon limits thereby, enhancing the resulting integrated emission.
 This must be true for diffuse emission as well.
-Thus we argue the feature arises primarily due to an increase in sensitivity of long baselines to diffuse emission caused by foreshortening towards angles far off--axis, where the increased size scales (or solid angles) containing higher levels of integrated emission, and spatial scales proportional to the inverse of foreshortened baseline lengths become matched}.
+Thus we argue the feature arises primarily due to an increase in sensitivity of long baselines to diffuse emission caused by foreshortening towards angles far off--axis.
+Consequently, the increased size scales (or solid angles) containing higher levels of integrated emission, and spatial scales inversely proportional to the foreshortened baseline lengths become matched}.
 
@@
 The amplitude response of an interferometer to a point source is, to first order, flat with respect to baseline length.
-Since the main lobe of the power--pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with positive delay corresponding to the position of the primary lobe of the power--pattern.
-In the {\it zenith} pointing, compact emission from the same patch of sky is seen as a bright horizontal arm at zero delay  since the primary lobe of the power--pattern is centered at zenith.
-
-Foreground emission in zero delay and negative delay regions in the {\it off--zenith} pointing is caused by point sources co--located with secondary lobes of the power--pattern.
-On the other hand, point sources co--located with secondary lobes of power--pattern in the {\it zenith} pointing are revealed as faint but distinct branches at positive and negative delays depending on the orientation of antenna spacing and direction of emission on the sky.
+Since the primary lobe of the power pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with $\tau>0$ corresponding to the position of the primary lobe of the power pattern.
+In the {\it zenith} pointing, compact emission from the same patch of sky is seen as a bright horizontal arm at $\tau=0$ since the primary lobe of the power pattern is centered at zenith.
+
+Foreground emission at $\tau=0$ and $\tau<0$ in the {\it off--zenith} pointing is caused by point sources co--located with secondary lobes of the power pattern.
+On the other hand, point sources co--located with secondary lobes of power pattern in the {\it zenith} pointing are revealed as faint but distinct branches at positive and negative delays depending on the orientation of antenna spacing and direction of emission on the sky.
 
@@
 
-When not dominated by the bright emission from the Galactic center, which happens only for certain pointings at certain times, the {\bf delay power spectrum} of the combined foreground model is composed of diffuse and compact emission, in roughly equal measure.
+When not dominated by the bright emission from the Galactic center, which happens only for certain pointings at certain times, the {\bf delay power spectrum} of the combined foreground model is composed of diffuse and compact emission, {\bf both of which are significant}.
 This is illustrated by a more detailed examination of the {\it zenith} pointing in our study.
@@
 The shaded region denotes the envelope of uncertainty around the composite model delay power spectrum.
-This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits.
-Outside the horizon delay limits, it is predominantly due to thermal noise fluctuations}.
+This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits and by thermal noise fluctuations elsewhere}.
 \label{fig:pitchfork-baselines}}
@@
 
-The zero--delay peak (corresponding to the main lobe in the power--pattern) with a value of $\sim 10^7$--$10^8$~K$^2$(Mpc/$h$)$^3$, independent of antenna spacing, is predominantly determined by compact emission.
+The peak at $\tau=0$ (corresponding to the primary lobe in the power pattern) with a value of $\sim 10^7$--$10^8$~K$^2$(Mpc/$h$)$^3$, independent of antenna spacing, is predominantly determined by compact emission.
 The corresponding peak at zero--delay from diffuse foreground model is $\sim 10^3$ times fainter and decreases rapidly with increase in antenna spacing.
@@
 Here, diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at zero--delay.
-In fact, even on widely spaced antennas, diffuse emission in the {\bf delay power spectrum} near the horizon delay limits exceeds that in the main lobe by about three orders of magnitude.
+In fact, even on widely spaced antennas, diffuse emission in the {\bf delay power spectrum} near the horizon delay limits exceeds that in the primary lobe by about three orders of magnitude.
 We attribute this to apparent shortening of antenna spacings projected in the direction of diffuse emission entering far from the primary field of view, thereby giving rise to the edge--heavy {\it two--pronged fork} feature discussed in \S\ref{sec:diffuse-long-baselines}.
@@
 For instance, there is an increase in the number of secondary peaks in the {\bf delay power spectrum} between zero--delay and horizon delay limits as the antenna spacing increases from $\sim$84~m to $\sim$171~m.
-In this case, these correspond to secondary lobes of the power--pattern that lie between the primary lobe and the horizon along the local meridian.
+In this case, these correspond to secondary lobes of the power pattern that lie between the primary lobe and the horizon along the local meridian.
 At short antenna spacings, due to relatively lower resolution along the delay axis inside the {\it foreground wedge} and a consequent loss of localization of foreground emission, these secondary peaks blend in with other major peaks and are not distinctly visible.
@@
 In summary, the brightest feature $\sim 10^{12}$~K$^2$(Mpc/$h$)$^3$ is that of the Galactic center in the {\it off--zenith} pointing co--located at a westward secondary lobe with a significantly high gain.
-The next brightest signature $\gtrsim 10^8$~K$^2$(Mpc/$h$)$^3$ is caused by compact emission appearing to be concentrated in the inner regions of the {\it foreground wedge} (near zero delays) rather than towards the boundaries.
-Diffuse emission co--located at the primary lobe of the power--pattern is $\sim 10^3$ fainter relative to compact emission from the same region for a $\sim 84$~m antenna spacing.
+The next brightest signature $\gtrsim 10^8$~K$^2$(Mpc/$h$)$^3$ is caused by compact emission appearing to be concentrated in the inner regions of the {\it foreground wedge} (around $\tau=0$) rather than towards the boundaries.
+Diffuse emission co--located at the primary lobe of the power pattern is $\sim 10^3$ fainter relative to compact emission from the same region for a $\sim 84$~m antenna spacing.
 But unlike the latter, diffuse emission decreases rapidly by over four orders of magnitude as the antenna spacing is widened to $\sim 171$~m.
-However, diffuse emission is significantly higher near horizon delay limits compared to that in the primary lobe of the power--pattern as the antenna spacing is increased.
+However, diffuse emission is significantly higher near horizon delay limits compared to that in the primary lobe of the power pattern as the antenna spacing is increased.
 This leads to edge--heavy features in the {\bf delay power spectra}.
@@
 Figure~\ref{fig:delay-map} shows the sky mapped to delays registered by the baseline vectors, of length 100~m for instance, oriented northward (top panel) and eastward (bottom panel).
-Since the Galactic center is located in the western sky, this figure demonstrates that it is observed at negative delay on a baseline vector oriented eastward and at zero delay on a baseline vector oriented northward.
+Since the Galactic center is located in the western sky, this figure demonstrates that it is observed at $\tau<0$ on a baseline vector oriented eastward and at $\tau=0$ on a baseline vector oriented northward.
 Figure~\ref{fig:baseline-breakup} shows the delay spectra on baselines oriented northward ($67.5\arcdeg\le \theta_\textrm{b} < 112.5\arcdeg$) at the top and eastward ($-22.5\arcdeg~\le \theta_\textrm{b} < 22.5\arcdeg$) at the bottom.
@@
 Delays vary linearly with antenna spacing length.
-The bright Galactic center in the western sky is recorded at zero delays in northward antenna spacings and close to negative horizon delay limit on eastward antenna spacings.
+The bright Galactic center in the western sky is recorded at $\tau=0$ in northward antenna spacings and close to negative horizon delay limit on eastward antenna spacings.
 (c): Simulated delay spectra {\bf power} on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}).
@@
 The {\bf delay power spectrum} not only carries information on spatial scales of foreground emission but also offers the unique advantage of viewing the sky through a combination of antenna spacing vectors and the delay axis.
-With a foreground model known {\it a priori} in which structures and locations of very bright foreground objects such as the Galactic center or AGN are available, we can predict their response across antenna spacings as a function of observing parameters such as LST, power--pattern, etc.
-This allows us to programmatically screen data for antenna spacings that are severely contaminated by foregrounds near the horizon delay limits in the {\bf delay power spectrum}.
+With a foreground model known {\it a priori} in which structures and locations of very bright foreground objects such as the Galactic center or AGN are available, we can predict their response across antenna spacings as a function of observing parameters such as LST, power pattern, etc.
+This allows us to programmatically screen data for antenna spacings that are severely contaminated by foregrounds near the horizon delay limits.
 These can be weighted appropriately during data analysis.
@@
 In our example, we discard eastward antenna spacings ($-22.5\arcdeg\le\theta_\textrm{b}< 22.5\arcdeg$) of lengths $|\boldsymbol{b}| < 30$~m.
-{\bf This was determined as the limit beyond which foreground contamination levels are insensitive to any further removal of wider antenna spacings}.
+{\bf This was found to be the limit beyond which foreground contamination levels are insensitive to any further removal of wider antenna spacings}.
 Figure~\ref{fig:before-after} shows the delay spectra obtained with all antenna spacings (top panel) and after applying our screening technique (bottom panel) on the {\it off--zenith} observation.
@@
 In principle, instead of discarding selected antenna spacings altogether, we could down--weight them based on an optimal scheme.
-For instance, the estimates of covariance computed from the delay transform bins can be naturally fed into the covariance--weighted power spectrum estimation techniques \citep{liu14a,liu14b}.
-It could also be used to downweight or flag contaminated baselines in imaging applications.
+{\bf For instance, the estimates of covariance computed from the delay transform bins can be naturally fed into the covariance--weighted power spectrum estimation techniques \citep{liu14a,liu14b}.
+It could also be used to downweight or flag contaminated baselines in imaging applications}.
 This technique provides a very simple and yet effective tool in adding a layer of control to mitigate effects of foreground contamination in EoR data analysis.
@@
 
-Having examined the role foregrounds play in determining the delay power spectrum obtained with an interferometer array, we turn our attention to another crucial factor, namely, the power pattern of the antenna (refer to equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}).
-Unlike the flagging of the edges of the coarse sub--bands described earlier in this paper, we assume an ideal flat passband over the entire 30.72~MHz bandwidth while still applying a {\it Blackman--Harris} window function.
+{\bf Having examined the role foregrounds play in determining the delay power spectrum obtained with an interferometer array, we turn our attention to another crucial factor, namely, the power pattern of the antenna (refer to equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}).
+Unlike the flagging of the edges of the coarse sub--bands described earlier in this paper, we assume an ideal flat passband over the entire 30.72~MHz bandwidth while still applying a {\it Blackman--Harris} window function}.
 
@@
 
-In the following discussion, we assume there are no random delay fluctuations in the MWA tile beamforming process.
+{\bf In the following discussion, we assume there are no random delay fluctuations in the MWA tile beamforming process.
 The model power pattern so obtained is different from the one we used earlier, but the former still serves as a good reference (Figure~\ref{fig:power_pattern}) for the present discussion.
@@
 
-The brightness distribution on the sky was fixed to be a constant equal to the mean over the whole sky at 185~MHz with no variation along frequency.
+The brightness distribution on the sky was fixed to be a constant equal to the mean over the whole sky at 185~MHz with no spectral variation.
 This helps isolate the signatures of the antenna power pattern from those of the foregrounds.
@@
 Figure~\ref{fig:noiseless-usm-delay-spectrum} shows the delay power spectra of such a sky model with the model power pattern of an ideal MWA tile.
-Although not identical, these delay power spectra bear a close resemblance to those obtained with the diffuse sky model (see Figure~\ref{fig:noiseless-dsm-delay-spectrum}).
-This indicates that the delay power spectra seen from diffuse foregrounds is primarily explained by those of the underlying power pattern, with any additional variations arising out of anisotropic deviations of the diffuse foreground model from a uniformly illuminated sky.
+Although not identical, these delay power spectra closely resemble those obtained with the diffuse sky model (see Figure~\ref{fig:noiseless-dsm-delay-spectrum}).
+This indicates that the delay power spectra seen from diffuse foregrounds is primarily explained by those of the underlying power pattern, with any additional variations arising out of anisotropic deviations of the diffuse foreground model from a uniformly illuminated sky}.
 
@@
 
-The edge--heavy {\it two--pronged} features are evident, which are very similar to those seen earlier in the case of diffuse emission.
+{\bf The edge--heavy {\it two--pronged} features are evident, which are very similar to those seen earlier in the case of diffuse emission.
 As reasoned in \S\ref{sec:diffuse-long-baselines}, the foreshortening of long baselines causes them to be sensitive to emission from near the horizon.
 
-Branches of foreground power seen at positive delays and close to the negative horizon delay limit in the {\it off--zenith} pointing correspond to radiation received from the direction of the primary lobe and the far eastward sidelobe, respectively, in the power pattern.
-On the other hand, the symmetry of the delay power spectrum around $\tau=0$ seen in the {\it zenith} pointing is as expected from a symmetric power pattern.
+Branches of foreground power seen at $\tau>0$ and close to the negative horizon delay limit in the {\it off--zenith} pointing correspond to radiation received from the direction of the primary lobe and the far eastward sidelobe, respectively, in the power pattern.
+On the other hand, the symmetry of the delay power spectrum around $\tau=0$ seen in the {\it zenith} pointing is as expected from a symmetric power pattern}.
 
@@
 
-The power pattern of an antenna is determined by the shape of its aperture.
+{\bf The power pattern of an antenna is determined by the shape of its aperture.
 Through the use of a few examples, we examine the role the geometrical shape of the aperture plays in shaping the characteristics of delay power spectrum.
@@
 The power patterns of these antenna geometries at 185~MHz for the {\it zenith} pointing are shown in Figure~\ref{fig:power-patterns}.
-It must be noted that the presence of a ground plane below the dipole and the phased array introduces a smoother suppression of sensitivity near the horizon, whereas this transition is abrupt in the case of the dish.
+It must be noted that the presence of a ground plane below the dipole and the phased array introduces a smoother suppression of sensitivity near the horizon, whereas this transition is abrupt in the case of the dish}.
 
@@
 
-The delay power spectra for these antenna shapes are shown in Figure~\ref{fig:antennas-delay-spectra}.
+{\bf The delay power spectra for these antenna shapes are shown in Figure~\ref{fig:antennas-delay-spectra}.
 The occupancy of the power patterns in the sky hemisphere is clearly correlated with that in the delay spectra.
@@
 
-The levels of foreground contamination in the {\it EoR window} varies substantially across the different antenna shapes: $\sim 10^4\,$K$^2\,$(Mpc/$h$)$^3$, $\lesssim 10^2\,$K$^2\,$(Mpc/$h$)$^3$, and $<1\,$K$^2\,$(Mpc/$h$)$^3$ while using a dipole, phased array, and a dish respectively.
-The severity of foregrounds inside the {\it foreground wedge} both in strength and occupancy also evidently decreases as the antenna element is changed from a dipole to a phased array to a dish.
+The levels of foreground contamination in the {\it EoR window} varies substantially across the different antenna shapes: $\sim 10^4\,$K$^2\,$(Mpc/$h$)$^3$, $\lesssim 10^2\,$K$^2\,$(Mpc/$h$)$^3$, and $<1\,$K$^2\,$(Mpc/$h$)$^3$ while using a dipole, phased array, and a dish, respectively.
+The severity of foreground contamination inside the {\it foreground wedge} both in strength and occupancy also evidently decreases as the antenna element is changed from a dipole to a phased array to a dish.
 For instance, notice that the foreground contamination in $k$--modes between the $k_\parallel=0$ and the horizon limits decreases from $\sim 10^5$~K$^2\,$(Mpc/$h$)$^3$ in a phased array to $\sim 10$~K$^2\,$(Mpc/$h$)$^3$ in a dish.
@@
 The prevalence of this feature across different antenna shapes demonstrates it is generic to all wide--field measurements, especially in EoR experiments.
-The amplitude of this effect, however, can be controlled via choice of antenna shape and through weighting of aperture illumination.
+The amplitude of this effect, however, can be controlled via choice of antenna shape and through weighting of aperture illumination}.
 
@@
 When the signal is expressed in units of temperature variance, the dynamic range between bright foregrounds and the 21~cm signal is expected to be $\sim 10^8$; a detailed understanding of how foregrounds can corrupt the 21~cm power spectrum is therefore essential.
-This analysis extends previous work by simulating the entire sky rather than just the central field of view and by providing a comparison with early observations from the MWA.
+This analysis extends previous work by simulating the entire sky rather than just the central field of view and by providing a comparison with early observations with the MWA.
 By making use of the delay spectrum technique to estimate the power spectrum, we are able to observe the effects of foregrounds while avoiding entanglements with more complex power spectrum estimators.
@@
 Our simulations enable us to identify numerous signatures of different components of foreground emission seen in the delay spectra.
-We establish the relationship between these signatures and observing parameters such as antenna pointing and LST, instrument parameters such as antenna power--pattern, and foreground parameters such as the nature of emission, spectral index, etc.
-
-The bright Galactic center at the edge of the western horizon co--located with one of the far secondary lobes of MWA tile power--pattern is the brightest source of foreground contamination in the {\it off--zenith} pointing.
+We establish the relationship between these signatures and observing parameters such as antenna pointing and LST, instrument parameters such as antenna power pattern, and foreground parameters such as the nature of emission, spectral index, etc.
+
+The bright Galactic center at the edge of the western horizon co--located with one of the far secondary lobes of MWA tile power pattern is the brightest source of foreground contamination in the {\it off--zenith} pointing.
 It manifests itself near the negative horizon delay limit in the {\bf delay power spectrum} on antenna spacings oriented eastward.
 
-As expected, diffuse emission in the primary beam of the antenna power--pattern is prominent on shorter antenna spacings.
+As expected, diffuse emission in the primary lobe of the antenna power pattern is prominent on shorter antenna spacings.
 However, the most interesting result is its footprint on wide antenna spacings near the horizon delay limits --- an edge--heavy {\it two--pronged fork}--shaped signature.
@@
 On the other hand, compact emission predominantly maps onto central regions of the {\it foreground wedge}.
-Features arising from compact emission co--located with primary and secondary lobes of the antenna power--pattern have been identified.
+Features arising from compact emission co--located with primary and secondary lobes of the antenna power pattern have been identified.
 In general, {\bf delay power spectrum} signatures of compact emission are center--heavy, in clear contrast to those from diffuse emission.
@@
 This will be distinctly visible when the thermal noise floor is sufficiently lowered, as longer observations are processed through the MWA analysis pipelines.
-
-The role of antenna shape in determining the {\bf delay power spectrum} was also studied.
-We conclude the {\it pitchfork} feature to be generic to all wide--field measurements, especially EoR experiments.
-While comparing a dipole, a phased array such as an MWA tile, and a dish such as the one proposed for HERA, the dish is found to yield the least foreground contamination both in the {\it foreground wedge} and the {\it EoR window} and hence, the most desirable for EoR studies.
+{\bf We conclude the {\it pitchfork} feature to be generic to all wide--field measurements, including EoR experiments}.
+
+{\bf The role of antenna shape in determining the {\bf delay power spectrum} was also studied.
+In contrast to a dipole and a phased array such as an MWA tile, a dish such as the one proposed for HERA is found to yield the least foreground contamination both in the {\it foreground wedge} and the {\it EoR window} and thus, preferable for EoR studies}.
 

### 2015-01-21 ms.tex
Modify motivation of uniformly illuminated sky
In ms.tex,

(*) Change first sentence in S7.1 motivating the simulation of uniformly
illuminated sky

In ApJ_referee_report_1_addressed.txt,

(*) Address flux scale question in the referee report

@@
 {\bf Having examined the role foregrounds play in determining the delay power spectrum obtained with an interferometer array, we turn our attention to another crucial factor, namely, the power pattern of the antenna (refer to equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}).
-Unlike the flagging of the edges of the coarse sub--bands described earlier in this paper, we assume an ideal flat passband over the entire 30.72~MHz bandwidth while still applying a {\it Blackman--Harris} window function}.
+In the following analysis, unlike the flagging of the edges of the coarse sub--bands described earlier in this paper, we assume an ideal flat passband over the entire 30.72~MHz bandwidth while still applying a {\it Blackman--Harris} window function.
+We also assume there are no random delay fluctuations in the MWA tile beamforming process}.
 
@@
 
-{\bf In the following discussion, we assume there are no random delay fluctuations in the MWA tile beamforming process.
+{\bf In order to confirm that the {\it pitchfork} effect is independent of the specific foreground model chosen in our study, we also simulated the response to a uniformly illuminated sky.
+The brightness distribution on the sky was fixed to be a constant equal to the mean over the whole sky at 185~MHz with no spectral variation.
+This helps isolate the signatures of the antenna power pattern from those of the foregrounds.
+
+ % In the following discussion, we assume there are no random delay fluctuations in the MWA tile beamforming process.
 The model power pattern so obtained is different from the one we used earlier, but the former still serves as a good reference (Figure~\ref{fig:power_pattern}) for the present discussion.
 Hence, we do not plot the new power pattern model obtained with ideal beamforming.
-
-The brightness distribution on the sky was fixed to be a constant equal to the mean over the whole sky at 185~MHz with no spectral variation.
-This helps isolate the signatures of the antenna power pattern from those of the foregrounds.
 

### 2015-01-21 ms.tex
Shorten sentence in summary on baseline screening
In ms.tex,

(*) Shorten a sentence in the summary on baseline screening

@@
 
-We also provide a simple and effective tool based on the delay spectrum technique {\bf that can potentially} mitigate foreground contamination by {\bf nearly two orders} of magnitude in EoR data analysis by discarding or down--weighting data from antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation, with negligible loss of sensitivity.
+We also provide a simple and effective tool based on the delay spectrum technique {\bf that can potentially} mitigate foreground contamination by {\bf nearly two orders} of magnitude in EoR data analysis by discarding or down--weighting data from antenna pairs most affected by foreground contamination, with negligible loss of sensitivity.
 In conclusion, we find that inclusion of emission models, both diffuse and compact, all the way to the horizon is essential to explaining the observed power spectrum.

### 2015-01-26 ms.tex
Add new figure to illustrate interferometer delays
from 837b478:
============

In ms.tex,

(*) Add a new figure (figure 1) and renamed all subsequent figure
file names. The new figure illustrates radio interferometer delays
and sign conventions used in the paper. Address one of the
referee's comments

In ApJ_referee_report_1_addressed.txt,

(*) Address comment to add a new figure illustrating radio
interferometer delays along with coordinate systems and sign
conventions used

@@
 Negative and positive delays correspond to the two hemispheres of the sky as transected by the antenna spacing vector.
-In this paper, $\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward (zenith) directions at the telescope site.
+{\bf Figure~\ref{fig:delay-cartoon} illustrates the radio interferometer delays and conventions used in our paper}.
+$\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward (zenith) directions at the telescope site.
 Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at positive and negative delays, respectively.
 Similarly, an object in the northern sky will appear at a positive delay at an antenna spacing oriented northward.
-For all observations used in this study, we use the zenith as the phase center.
+For all observations used in this study, we use the zenith as the phase center, {\bf for which $\tau\equiv 0$}.
+
+\begin{figure}[htb]
+\centering
+% \includegraphics[width=\linewidth]{delay_cartoon.eps}
+\includegraphics[width=\linewidth]{fig1.eps}
+\caption{{\bf Illustration of radio interferometer delays and conventions used in this paper.
+Two antennas labeled as A$_1$ and A$_2$ are separated by vector $\boldsymbol{b}$ on the local tangent plane.
+The local meridian, local north and local east are shown for reference.
+Points labeled as `z', `n', `s', `e' and `w' on the celestial sphere denote zenith, northward, southward, eastward and westward positions, respectively.
+$\tau_\textrm{z}$, $\tau_\textrm{n}$, $\tau_\textrm{s}$, $\tau_\textrm{e}$ and $\tau_\textrm{w}$ denote the respective delays measured between A$_1$ and A$_2$.
+Throughout this paper, zenith is chosen as the phase center.
+Hence, $\tau_\textrm{z}\equiv 0$.
+If $\boldsymbol{b}$ is oriented eastward as shown, then $\tau_\textrm{e} > 0$, $\tau_\textrm{w} < 0$, and $\tau_\textrm{n}=\tau_\textrm{s}=0$.
+Conversely, if $\boldsymbol{b}$ is oriented northward (not shown here), then $\tau_\textrm{n} > 0$, $\tau_\textrm{s} < 0$, and $\tau_\textrm{e}=\tau_\textrm{w}=0$}.
+\label{fig:delay-cartoon}}
+\end{figure}
 
@@
 % \includegraphics[width=\linewidth]{f1.eps}
-\includegraphics[width=\linewidth]{fig1.eps}
+\includegraphics[width=\linewidth]{fig2.eps}
 \caption{Fourier space in which delay (and power) spectra of EoR H{\sc i} signals are presented.
@@
 % \includegraphics[width=\linewidth]{f2.eps}
-\includegraphics[width=\linewidth]{fig2.eps}
+\includegraphics[width=\linewidth]{fig3.eps}
 \caption{MWA tile beam pointing directions during the course of the observation.
@@
 % \includegraphics[width=\linewidth]{f3.eps}
-\includegraphics[width=\linewidth]{fig3.eps}
+\includegraphics[width=\linewidth]{fig4.eps}
 \caption{{\bf Delay power spectra} of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
@@
 % \includegraphics[width=\linewidth]{f4.eps}
-\includegraphics[width=\linewidth]{fig4.eps}
+\includegraphics[width=\linewidth]{fig5.eps}
 \caption{Models of MWA tile power pattern for {\it off--zenith} (top) and {\it zenith} pointings (bottom) at 185~MHz.
@@
 % \includegraphics[width=\linewidth]{f5.eps}
-\includegraphics[width=\linewidth]{fig5.eps}
+\includegraphics[width=\linewidth]{fig6.eps}
 \caption{Sky brightness temperature of the diffuse foreground model at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 % \includegraphics[width=\linewidth]{f6.eps}
-\includegraphics[width=\linewidth]{fig6.eps}
+\includegraphics[width=\linewidth]{fig7.eps}
 \caption{Flux densities of bright point sources at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 % \includegraphics[width=\linewidth]{f7.eps}
-\includegraphics[width=\linewidth]{fig7.eps}
+\includegraphics[width=\linewidth]{fig8.eps}
 \caption{Same as Figure~\ref{fig:fhd_data}, but generated from simulated visibilities using the model described in \S\ref{sec:modeling}.
@@
 % \includegraphics[width=\linewidth]{f8.eps}
-\includegraphics[width=\linewidth]{fig8.eps}
+\includegraphics[width=\linewidth]{fig9.eps}
 \caption{Simulated {\bf delay power spectra (in units of K$^2$~(Mpc/$h$)$^3$} for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
@@
 % \includegraphics[width=\linewidth]{f9.eps}
-\includegraphics[width=\linewidth]{fig9.eps}
+\includegraphics[width=\linewidth]{fig10.eps}
 \caption{Same as Figure~\ref{fig:noiseless-dsm-delay-spectrum} but for a foreground model consisting of compact emission.
@@
 % \includegraphics[width=\linewidth]{f10.eps}
-\includegraphics[width=\linewidth]{fig10.eps}
+\includegraphics[width=\linewidth]{fig11.eps}
 \caption{Simulated {\bf delay power spectra} for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}).
@@
 % \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{f11c.eps}}
-\subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{fig11a.eps}}
-\subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{fig11b.eps}}
-\subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{fig11c.eps}}
+\subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{fig12a.eps}}
+\subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{fig12b.eps}}
+\subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{fig12c.eps}}
 \caption{(a): Sky model showing compact ({\it top}) and diffuse ({\it bottom}) emission.
@@
 % \includegraphics[width=\linewidth]{f12.eps}
-\includegraphics[width=\linewidth]{fig12.eps}
+\includegraphics[width=\linewidth]{fig13.eps}
 \caption{Simulated delay spectra {\bf power} for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
@@
 % \includegraphics[width=\linewidth]{f13.eps}
-\includegraphics[width=\linewidth]{fig13.eps}
+\includegraphics[width=\linewidth]{fig14.eps}
 \caption{Drop in foreground contamination in the {\it EoR window}, and loss of data for the {\it off--zenith} pointing as a function of discarded baselines.
@@
 \centering
-\includegraphics[width=\linewidth]{fig14.eps}
+\includegraphics[width=\linewidth]{fig15.eps}
 \caption{Similar to Figure~\ref{fig:noiseless-dsm-delay-spectrum} but for a uniform sky brightness model.
@@
 \centering
-\subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{fig15a.eps}} \\
-\subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{fig15b.eps}}
+\subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{fig16a.eps}} \\
+\subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{fig16b.eps}}
 \caption{Power patterns (top panels) and simulated delay power spectra (bottom panels) for different antenna shapes at 185~MHz centered on zenith.

### 2015-01-26 ms.tex
Update Bernardi et al 2014 to Bernardi et al 2015
from 572244e:
============

In eor.bib,

(*) Update Bernardi et al. 2014 to Bernardi et al. 2015

In ms.tex,

(*) Update Bernardi et al. 2014 to Bernardi et al. 2015

@@
 These estimates are conservative.
-A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power pattern \citep{ber14}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide--field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations, but is beyond the scope of this paper.
+A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power pattern \citep{ber15}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide--field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations, but is beyond the scope of this paper.
 {\bf Hereafter}, our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.

### 2015-01-27 ms.tex
Explain interferometer sensitivity to uniform sky
from 139fc7b:
============

In ms.tex,

(*) Add paragraph in S7.1 to explain wide-field interferometers are
sensitive to a uniformly sky

@@
 
+Na\"{i}vely, only single element antennas are sensitive to a uniform brightness distribution while interferometers are not.
+However, an examination of equation~\ref{eqn:obsvis} shows it resembles a Fourier transform when the sky brightness is expressed in a coordinate system of direction cosines \citep{tay99,tho01}.
+In such a framework, the solid angle subtended by a unit interval in direction cosine increases as the secant of the angle away from zenith, which is significant in wide--field measurements.
+A uniformly illuminated sky does not appear so in this frame.
+Therefore, a uniformly illuminated sky will excite wide--field interferometers.
+
 Figure~\ref{fig:noiseless-usm-delay-spectrum} shows the delay power spectra of such a sky model with the model power pattern of an ideal MWA tile.

### 2015-01-28 ms.tex
Re-word a few sentences in S5.1.3 and S7
from 0816602:
============

In ms.tex,

(*) Re-word a few sentences in S5.1.3 and S7

@@
 {\bf In typical radio interferometric measurements, due to the Fourier relation between sky brightness and measured visibilities, only short antenna spacings are sensitive to emission from large size scales such as the diffuse component in our study (angular scales $\gtrsim$~0.85\arcdeg).
-On the contrary, we note a very interesting signature of diffuse emission revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing, and at both horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\boldsymbol{b}|\le$~200~m).
+However, we note a very interesting signature of diffuse emission revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing, and at both horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\boldsymbol{b}|\le$~200~m).
 This is a consequence of wide--field measurements, where long baselines are foreshortened due to projection effects towards off--axis emission, especially near the horizon.
@@
 
-{\bf Interestingly, \citet{thy13} and \citet{pob13} have found a similar feature in their simulations and data respectively.
+{\bf \citet{thy13} and \citet{pob13} have found a similar feature in their simulations and data respectively.
 In the latter, this feature is described as ``edge brightening'' and attributed to the Galactic plane near the horizon.
@@
 
-{\bf Having examined the role foregrounds play in determining the delay power spectrum obtained with an interferometer array, we turn our attention to another crucial factor, namely, the power pattern of the antenna (refer to equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}).
-In the following analysis, unlike the flagging of the edges of the coarse sub--bands described earlier in this paper, we assume an ideal flat passband over the entire 30.72~MHz bandwidth while still applying a {\it Blackman--Harris} window function.
+{\bf We now examine the influence of the antenna power pattern on the delay power spectrum (refer to equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}).
+In the following analysis, we assume an ideal flat passband over the entire 30.72~MHz bandwidth while still applying a {\it Blackman--Harris} window function.
 We also assume there are no random delay fluctuations in the MWA tile beamforming process}.
@@
 
-{\bf In order to confirm that the {\it pitchfork} effect is independent of the specific foreground model chosen in our study, we also simulated the response to a uniformly illuminated sky.
+{\bf Simulations of the response to a uniformly illuminated sky confirm that the {\it pitchfork} effect is independent of the specific foreground model chosen in our study.
 The brightness distribution on the sky was fixed to be a constant equal to the mean over the whole sky at 185~MHz with no spectral variation.
@@
 The occupancy of the power patterns in the sky hemisphere is clearly correlated with that in the delay spectra.
-Moreover, the following visual correlations are also apparent --- the strength of the primary lobe centered on the pointing center is correlated with the delay power spectrum centered on $\tau=0$; and, the overall rate of decrease in the power sensitivity away from the pointing center is correlated with the rate of drop in power away from $\tau=0$.
+The following visual correlations are also apparent --- the strength of the primary lobe centered on the pointing center is correlated with the delay power spectrum centered on $\tau=0$; and, the overall rate of decrease in the power sensitivity away from the pointing center is correlated with the rate of drop in power away from $\tau=0$.
 

### 2015-02-03 ms.tex
Create branch alternate-version, change title
from 13ccbd7:
============

(*) Create a git branch alternate-version

In ms.tex,

(*) Change  title and short title
Modify abstract for alternate version
from ea7f9f8:
============

In ms.tex,

(*) Modify abstract for alternate version

@@
 
-\shorttitle{Dissection of foregrounds in EoR H{\sc i} power spectra}
+\shorttitle{EoR Foregrounds in Fourier Space}
 \shortauthors{Thyagarajan et~al.}
@@
 %% \title{Limits on the detection of the Epoch of Reionization from MWA observations of the redshifted 21~cm line}
-\title{A Dissection of Foregrounds in Epoch of Reionization Observations from the Murchison Widefield Array}
+\title{Foregrounds in Fourier Space for Redshifted 21~cm Observatories}
 
@@
 Detection of 21~cm emission of H{\sc i} from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission.
-We investigate the chromatic signatures of an all--sky foreground model imprinted by an instrumental transfer function.
-Using the delay spectrum technique {\bf that maps the measurements to foreground object locations through signal delays between antenna pairs}, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view.
-Comparing data from recent Murchison Widefield Array observations with simulations {\bf broken down by types of emission}, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings {\bf which are usually not sensitive to diffuse emission}.
+We investigate the signatures of wide--field measurements and an all--sky foreground model using the delay spectrum technique that maps the measurements to foreground object locations through signal delays between antenna pairs.
+We demonstrate interferometric measurements are inherently sensitive to all scales, including the largest angular scales, owing to the nature of wide--field measurements.
+These wide--field effects are generic to all observations but antenna shapes impact their amplitudes substantially.
+A dish--shaped antenna yields the most desirable features from a foreground contamination viewpoint, relative to a dipole or a phased array.
+Comparing data from recent Murchison Widefield Array observations, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view.
+We identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings.
 For signals entering through the primary field of view, compact emission dominates the foreground contamination.
-These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space, {\bf which is found to be generic to wide--field measurements}.
-Based on these results, we propose that selective down--weighting of data based on antenna spacing and time {\bf can} mitigate foreground contamination substantially (up to a {\bf factor $\sim 100$} with negligible loss of sensitivity).
-{\bf A dish--shaped antenna yields the most desirable features from a foreground contamination viewpoint, relative to a dipole or a phased array}.
+These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space.
+Based on these results, we propose that selective down--weighting of data based on antenna spacing and time can mitigate foreground contamination substantially by a factor $\sim 100$ with negligible loss of sensitivity.
 

### 2015-02-04 ms.tex
Add a new section "Wide-field measurements"
from cfc6578:
============

In ms.tex,

(*) Add a new section "Wide-field measurements"
(*) Need to add a figure showing wide-field effects

@@
 In \S\ref{sec:delay-spectrum} we {\bf provide a brief overview of the delay spectrum technique} \citep{par12a,par12b}.
-In \S\ref{sec:instrument}, we describe the MWA setup, summarize the observing parameters, and present the resulting data.
+In \S\ref{sec:data}, we describe the MWA setup, summarize the observing parameters, and present the resulting data.
 Simulations using these observing parameters are described in \S\ref{sec:modeling} and analyzed by sky location and type of emission in \S\ref{sec:delay-spectrum-analysis}.
@@
 
+In our present study, we use an antenna layout identical to that of the MWA \citep{bea12}.
+The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km.
+Here we focus on antenna spacings $|\boldsymbol{b}| \le 200$~m (spatial scales relevant to reionization).
 For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67.5\arcdeg \leq \theta_b < 112.5\arcdeg$.
@@
 
-\section{Example Data}\label{sec:instrument}
-
-\subsection{The Murchison Widefield Array (MWA)}
+\section{Wide--Field Measurements}\label{sec:wide-field}
+
+With $\hat{\boldsymbol{s}}\equiv (l,m,n)$, equation~\ref{eqn:obsvis} can be written as \citep{tay99,tho01}:
+\begin{align}\label{eqn:obsvis-lmn}
+  V_{bf} &= \iint\limits_\textrm{sky} \frac{A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)}{\sqrt{1-l^2-m^2}}\,W_f\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif l\,\dif m,
+\end{align}
+where, $l$, $m$, and $n$ denote the direction cosines toward east, north, and zenith respectively, with $n\equiv\sqrt{1-l^2-m^2}$, and:
+\begin{align}\label{eqn:solid-angle}
+  \dif\Omega = \frac{\dif l\,\dif m}{\sqrt{1-l^2-m^2}}.
+\end{align}
+
+When the synthesized field is small, where $A(\hat{\boldsymbol{s}},f)$ or $I(\hat{\boldsymbol{s}},f)$ is negligible for $|l| \ll 1$ and $|m| \ll 1$, equation~\ref{eqn:obsvis-lmn} reduces to a simple two--dimensional Fourier transform \citep{tay99,tho01} between the apparent sky brightness and measured visibilities.
+It is in this context that radio interferometers are understood to be sensitive only to fluctuations and not to a uniform sky brightness distribution.
+
+In a wide--field measurement, neither $A(\hat{\boldsymbol{s}},f)$ nor $I(\hat{\boldsymbol{s}},f)$, in general, is negligible anywhere in the visible hemisphere.
+The solid angle per pixel on the sky in direction cosine coordinates changes significantly with direction (equation~\ref{eqn:solid-angle}), increasing steeply towards the horizon.
+Hence, the approximations in the narrow--field scenario do not apply.
+For example, even if $A(\hat{\boldsymbol{s}},f)$ and $I(\hat{\boldsymbol{s}},f)$ are held constant across the visible hemisphere, the integrand in equation~\ref{eqn:obsvis-lmn} is still dependent on direction.
+Therefore, in a significant departure from a narrow--field measurement, the wide--field visibility from a uniform brightness distribution on a non--zero antenna spacing is not zero.
+
+Figure ???
+shows the wide--field delay power spectrum response of a uniformly illuminated sky with no spectral variation measured by antenna elements with a uniform power pattern across the sky.
+Notice the features arising due to the direction--dependent solid angle term and the steep rise in power towards the horizon limits.
+Wide--field effects are prevalent on all antenna spacings, including the longest ones used in this study.
+
+\citet{thy13} have found evidence of this feature in their statistical models.
+In line with our reasoning, they attribute it to an increase in solid angles subtended by delay bins closer to the horizon limits thereby, enhancing the resulting integrated emission.
+\citet{pob13} also find a similar ``edge brightening'' feature which they attribute to emission from the Galactic plane near the horizon.
+From their discussion, it is unclear what fraction of power in that feature arises from such wide--field effects.
+
+We conclude these are generic to all instruments making wide--field measurements.
+The nature of the specific instrument used for observing will control the amplitude of these effects, which we explore below.
+% Typically, the foreground sky consists of diffuse emission and point-like sources, and the antenna is more sensitive to power received from primary field of view than from the rest of the sky.
+
+\section{Simulations}\label{sec:sim}
+
+\section{Example Data}\label{sec:data}
+
+\subsection{The Murchison Widefield Array (MWA)}\label{sec:instrument}
 

### 2015-02-04 ms.tex
Bring figure3.eps under git version control
from 06e98c3:
============

(*) Bring figure3.eps under git version control as it was not done in
the previous commit
Add figure showing wide-field effects

from 28c0fe9:
============

(*) Bring figure3.eps under git version control

In ms.tex,

(*) Add figure showing wide-field effects under section "Wide-feild
measurements"

@@
 
-Figure ???
-shows the wide--field delay power spectrum response of a uniformly illuminated sky with no spectral variation measured by antenna elements with a uniform power pattern across the sky.
+Figure~\ref{fig:usm} shows the wide--field delay power spectrum response of a uniformly illuminated sky with no spectral variation measured by antenna elements with a uniform power pattern across the sky.
 Notice the features arising due to the direction--dependent solid angle term and the steep rise in power towards the horizon limits.
 Wide--field effects are prevalent on all antenna spacings, including the longest ones used in this study.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\linewidth]{figure3.eps}
+\caption{Delay power spectra of a uniform sky brightness distribution measured by antenna pairs with a uniform power pattern across the visible hemisphere.
+The non--zero response of the interferometer array to a uniform brightness distribution and the prominent {\it edge brightening} close to the horizon delay limits are due to wide--field effects.
+These effects are prevalent on all antenna spacings and are generic to all instruments used in wide--field measurements.
+\label{fig:usm}}
+\end{figure}
 

### 2015-02-04 ms.tex
Add interpretation to wide-field effects
from a805428:
============

In ms.tex,

(*) Add interpretation to wide-field effects under section "Wide-field
measurements"

@@
 
+We interpret this is due to equal--sized delay bins subtending larger solid angles near the horizon thereby containing larger integrated emission.
+An alternate, but equivalent, interpretation is that baseline vectors (including those with largest lengths) are foreshortened towards the horizon along their orientation.
+Thus, they become sensitive to larger angular scales that match the inverse of their foreshortened baseline lengths.
+
 \citet{thy13} have found evidence of this feature in their statistical models.
-In line with our reasoning, they attribute it to an increase in solid angles subtended by delay bins closer to the horizon limits thereby, enhancing the resulting integrated emission.
+In line with our reasoning, they attribute it to a steep rise in solid angles subtended by delay bins near the horizon limits.
 \citet{pob13} also find a similar ``edge brightening'' feature which they attribute to emission from the Galactic plane near the horizon.

### 2015-02-04 ms.tex
Add new section "Foreground Model"
from ed945db:
============

Add figure4.eps to git version control

In ms.tex,

(*) Add new section "Foreground Model"
(*) Minor modification to last paragraph of section 2.1

@@
 
-In our present study, we use an antenna layout identical to that of the MWA \citep{bea12}.
+In our present study, we use a latitude (-26.701\arcdeg) and an antenna layout identical to that of the MWA \citep{bea12}.
 The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km.
@@
 \section{Simulations}\label{sec:sim}
+
+We describe the instrumental and foreground models we use in our simulations.
+
+\subsection{Foreground Model}\label{sec:foreground}
+
+In wide--field measurements, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view.
+We use a foreground model that includes both diffuse and bright compact components.
+
+\subsubsection{Diffuse Foreground Model}\label{sec:DSM}
+
+For the diffuse component, we use an all--sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz.
+At this frequency, since this map is predominantly based on the 408~MHz map of \citet{has82} which has an angular resolution of 0.85\arcdeg, we smoothed the 185~MHz map to the same resolution.
+However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals.
+We model the diffuse foreground spectra with a spectral index at each pixel in the map, estimated from model maps at 170~MHz and 200~MHz.
+
+\subsubsection{Compact Foreground Model}\label{sec:CSM}
+
+The model described in the preceding section is primarily a model of the diffuse foreground sky.
+While it contains faint compact emission blended in with the diffuse emission, {\bf bright point sources have been removed \citep{deo08}}.
+In order to supplement it with missing bright compact emission, we use classical radio source confusion estimates to determine the nominal flux density threshold and include point sources brighter than this threshold.
+It may be noted that slightly different criteria are in common use in radio astronomy to estimate radio source confusion \citep[see Appendix of][and references therein]{thy13}.
+For an angular resolution of 0.85\arcdeg, using a conservative `$S_\textrm{c}=5\sigma_\textrm{c}$' criterion, we determine the flux density threshold to be $\approx 10$~Jy.
+Other liberal criteria that yield a lower threshold carry a greater risk of double--counting point sources which might be already blended in with the diffuse sky model.
+
+We use a combination of the NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and the Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz to provide our point source catalog due to their complementary survey footprints covering the entire sky, and matched flux density sensitivity and angular resolution.
+The SUMSS catalog covers the sky with declination $\delta < -30\arcdeg$~ with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec.
+The NVSS covers the sky with $\delta > -40\arcdeg$~ with a similar angular resolution and a limiting flux density of $\approx 2.5$~mJy for point sources.
+
+From the SUMSS catalog, we select objects whose deconvolved major axes are equal to 0\arcsec, thereby strictly selecting point sources.
+From the NVSS catalog, we excluded objects that overlap with those in the SUMSS survey footprint.
+Point sources from NVSS were selected if the convolved major axes were not greater than $\approx 47$\arcsec, which matches the angular resolution of the survey.
+Using a mean spectral index of $\langle\alpha_\textrm{sp}\rangle=-0.83$ (flux density, $S(f)\propto f^{\alpha_\textrm{sp}}$) obtained by \citet{mau03} for both NVSS and SUMSS catalog objects, we calculate the corresponding flux densities at 185~MHz, $S_{185}$.
+From this subset, we choose point sources with $S_{185}\geq 10$~Jy.
+The selection of such bright point sources is not affected by minor differences in flux density sensitivity of the two surveys.
+We verified that our selection criteria ensure a similar areal density of objects in the two surveys.
+
+Based on these criteria, we selected 100 objects from the SUMSS catalog and 250 objects from the NVSS catalog.
+Together with the diffuse foreground model, we obtain an all--sky foreground model consisting of both compact and diffuse emission.
+Figure~\ref{fig:sky-model-generic} shows the diffuse (top) and compact (bottom) foreground emission model used in our study.
+In this snapshot at 0.09~hours LST, the Galactic center in the diffuse model appears to have just set in the West.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\linewidth]{figure4.eps}
+\caption{Foreground model at 185~MHz consisting of diffuse emission (top) and bright point sources (bottom) visible during a snapshot at 0.09~hours LST.
+In the diffuse model, the Galactic center appears to have just set in the West.
+\label{fig:sky-model-generic}}
+\end{figure}
 

### 2015-02-04 ms.tex
Add subsection "Instrument Model"
from f851ac1:
============

Add figure1.eps, figure2.eps, figure5a.eps, figure5b.eps to git version
control

In ms.tex,

(*) Complete subsection "Foreground Model"
(*) Add subsection "Instrument Model" under section "Simulations"
(*) Minor change to section on "Delay Spectrum"

@@
 where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector.
-We wish to {\bf note} that this equation is valid {\bf even for wide--field measurements where the sky cannot be approximated to be flat}.
+We wish to {\bf note} that this equation is valid {\bf in general, including wide--field measurements}.
 
@@
 % \includegraphics[width=\linewidth]{delay_cartoon.eps}
-\includegraphics[width=\linewidth]{fig1.eps}
+% \includegraphics[width=\linewidth]{fig1.eps}
+\includegraphics[width=\linewidth]{figure1.eps}
 \caption{{\bf Illustration of radio interferometer delays and conventions used in this paper.
@@
 % \includegraphics[width=\linewidth]{f1.eps}
-\includegraphics[width=\linewidth]{fig2.eps}
+% \includegraphics[width=\linewidth]{fig2.eps}
+\includegraphics[width=\linewidth]{figure2.eps}
 \caption{Fourier space in which delay (and power) spectra of EoR H{\sc i} signals are presented.
@@
 
-In our present study, we use a latitude (-26.701\arcdeg) and an antenna layout identical to that of the MWA \citep{bea12}.
+In our present study, we use a latitude of -26.701\arcdeg and an antenna layout identical to that of the MWA \citep{bea12}.
 The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km.
@@
 
-We describe the instrumental and foreground models we use in our simulations.
+We choose an observing frequency of 185~MHz ($z\simeq 6.68$) and a flat passband of width $\Delta B = 30.72$~MHz to roughly match those of ongoing MWA EoR observations.
+Now we describe the foreground and instrumental models used in our simulations.
 
@@
 
-Based on these criteria, we selected 100 objects from the SUMSS catalog and 250 objects from the NVSS catalog.
+Based on these criteria, we select 100 objects from the SUMSS catalog and 250 objects from the NVSS catalog.
 Together with the diffuse foreground model, we obtain an all--sky foreground model consisting of both compact and diffuse emission.
@@
 \end{figure}
+
+\subsection{Instrument Model}\label{sec:instrument_model}
+
+The instrument model consists primarily of the antenna power pattern, $A(\hat{\boldsymbol{s}},f)$ (see equation~\ref{eqn:obsvis}).
+It is determined by the shape of its aperture.
+Through the use of a few examples, we examine the role the geometrical shape of the aperture plays in shaping the characteristics of delay power spectrum.
+We consider the following antenna elements placed at the MWA tile locations:
+\begin{itemize}
+\item {\it Dipole}: An East--West dipole of length 0.74~m at a height 0.3~m above a ground plane. $A_\textrm{e}=(\lambda/2)^2$.
+\item {\it Phased Array}: A 4$\times$4 array of isotropic radiators with a grid spacing of 1.1~m at a height 0.3~m above the ground plane placed in an arrangement similar to that in an MWA tile.
+$A_\textrm{e}=16\,(\lambda/2)^2$.
+\item {\it Dish}: Diameter of 14~m similar to that proposed for HERA, with $A_\textrm{e}\approx 154$~m$^2$.
+The power pattern is simulated using an {\it Airy} pattern where its sensitivity beyond the horizon is forced to zero.
+\end{itemize}
+
+The power patterns of these antenna geometries at 185~MHz for the {\it zenith} pointing are shown in Figure~\ref{fig:power-patterns}.
+It must be noted that the presence of a ground plane below the dipole and the phased array introduces a smoother suppression of sensitivity near the horizon, whereas this transition is abrupt in the case of the dish.
+
+\begin{figure*}[htb]
+\centering
+\subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{figure5a.eps}} \\
+\subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{figure5b.eps}}
+\caption{Power patterns (top panels) and simulated delay power spectra (bottom panels) for different antenna shapes at 185~MHz centered on zenith.
+Antenna shapes used are: dipole (left), phased array (middle), and dish (right).
+Refer to text for details.
+The strength and occupancy of the power patterns are correlated with those of delay power spectra.
+The {\it foreground wedge} and the {\it EoR window} are most severely contaminated in the case of the dipole while it is the least for the dish.
+The phased array has intermediate levels of contamination.
+Foreground power close to the horizon delay limits in all three cases are significant even on long baselines.
+The foreshortening of baselines towards the horizon makes them sensitive to foreground emission on large size scales.
+The amplitude of this feature is highest for the dipole and least for the dish and hence, strongly depends on the shape of the antenna element.}
+\label{fig:aperture-shapes}
+\end{figure*}
+
+The delay power spectra for these antenna shapes are shown in Figure~\ref{fig:antennas-delay-spectra}.
+The occupancy of the power patterns on the sky is clearly correlated with that in the delay spectra.
+The following visual correlations are also apparent --- the strength of the primary lobe centered on the pointing center is correlated with the delay power spectrum centered on $\tau=0$; and, the overall rate of decrease in the power sensitivity away from the pointing center is correlated with the rate of drop in power away from $\tau=0$.
+
+The levels of foreground contamination in the {\it EoR window} varies substantially across the different antenna shapes: $\sim 10^4\,$K$^2\,$(Mpc/$h$)$^3$, $\lesssim 10^2\,$K$^2\,$(Mpc/$h$)$^3$, and $<1\,$K$^2\,$(Mpc/$h$)$^3$ while using a dipole, phased array, and a dish, respectively.
+The severity of foreground contamination inside the {\it foreground wedge} both in strength and occupancy also evidently decreases as the antenna element is changed from a dipole to a phased array to a dish.
+For instance, notice that the foreground contamination in $k$--modes between the $k_\parallel=0$ and the horizon limits decreases from $\sim 10^5$~K$^2\,$(Mpc/$h$)$^3$ in a phased array to $\sim 10$~K$^2\,$(Mpc/$h$)$^3$ in a dish.
+As a consequence, $k$--modes in the {\it foreground wedge} that may be deemed too contaminated for EoR studies in the case of a dipole or a phased array can potentially become accessible when using a dish.
+
+Finally, a distinct feature common to all these aperture shapes is that the foreground contamination near the horizon delay limits even on wide antenna spacings is significant ($\gtrsim 10^5$~K$^2$(Mpc/$h$)$^3$).
+We have argued this arises due to wide--field effects.
+
+The prevalence of this feature across different antenna shapes demonstrates it is generic to all wide--field measurements, especially in EoR experiments.
+The amplitude of this effect, however, can be controlled via choice of antenna shape and through weighting of aperture illumination.
+A dish--shaped antenna appears to hold a significant advantage over a dipole or a phased array from the viewpoint of foreground contamination.
+
+Typically, the sensitivity of antennas to the primary field of view is high compared to the rest of regions on the sky.
+Combined with the wide--field effects seen earlier, it leads to a ``pitchfork''--shaped signature inside the {\it foreground wedge}, as exemplified in the case of a dish.
+Although the exact appearance of this signature depends on antenna the power pattern, we use the term ``pitchfork'', hereafter, to broadly refer to the combination of foreground power in the primary field of view and the enhancement of foreground power near the horizon limits due to the nature of wide--field measurements.
 
@@
 
-\begin{figure*}[htb]
-\centering
-\subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{fig16a.eps}} \\
-\subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{fig16b.eps}}
-\caption{Power patterns (top panels) and simulated delay power spectra (bottom panels) for different antenna shapes at 185~MHz centered on zenith.
+% \begin{figure*}[htb]
+% \centering
+% \subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{fig16a.eps}} \\
+% \subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{fig16b.eps}}
+% \caption{Power patterns (top panels) and simulated delay power spectra (bottom panels) for different antenna shapes at 185~MHz centered on zenith.
 Antenna shapes used are: dipole (left), phased array (middle), and dish (right).
@@
 The amplitude of this feature is highest for the dipole and least for the dish and hence, strongly depends on the shape of the antenna element.}
-\label{fig:aperture-shapes}
-\end{figure*}
+% \label{fig:aperture-shapes}
+% \end{figure*}
 

### 2015-02-04 ms.tex
Make sec. 2 compatible with alternate outline
from 379e27b:
============

In ms.tex,

(*) Make small changes to end of section 2
(*) Make small changes to caption of figure 2
(*) Remove reference to Fourier space figure as being specific to MWA
(*) Now sections 2, 3, and 4 are more or less compatible with the
alternate outline

@@
 
-Equation~\ref{eqn:obsvis} can be {\bf equivalently re-expressed} as:
+Equation~\ref{eqn:obsvis} can be {\bf equivalently expressed} as:
 \begin{align}\label{eqn:vis}
@@
 However, they are not quite identical to each other.
-The difference is because $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not.
+It is because $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not.
 \citet{par12b} and \citet{liu14a} have discussed the mathematical correspondence between the two.
@@
 {\bf We give a brief overview of some parameters of Fourier space which is generic to all experiments that use a similar approach}.
-Figure~\ref{fig:fourier-space} illustrates the Fourier space, {\bf specific to the MWA}, in which the delay (and power) spectra of H{\sc i} from the EoR are presented in this paper.
+Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of H{\sc i} from the EoR are presented in this paper.
 $|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction {\bf (tangent plane to the celestial sphere)}, form the $x$--axis.
@@
 The region of Fourier space excluding the {\it foreground wedge} and the narrow strips is the so--called {\it EoR window}, shown in light and medium shades.
-However, the instrument used in our study, namely the MWA, has a passband constructed using coarse channels with a characteristic shape giving rise to periodic gratings thereby resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.
-
-In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\boldsymbol{k}|$.
+{\bf In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\boldsymbol{k}|$.
 Hence, the brightest EoR signal will be observed on the shortest baselines and smallest delays.
-Thus the region of interest for EoR studies is slightly beyond the horizon delay limits (dashed lines) and inside the first grating on either side ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) caused by the instrument's spectral transfer function.
-We call this region with medium shade as the {\it EoR window} in this paper.
-The MWA EoR collaboration actively uses this as a test region in Fourier space for diagnostics on data quality.
+Thus the region of interest for EoR studies is just beyond the horizon delay limits (dashed lines) on short baselines, marked as the region of {\it maximal EoR sensitivity}}.
+
+{\bf In the specific case of the MWA which has a passband constructed using coarse channels, there are period grating responses resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.
+Thus with reference to the MWA, the {\it EoR window} is assumed to lie outside the dashed lines but inside the first grating response ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) and is shown in medium shade.
+The MWA EoR collaboration actively uses this as a test region in Fourier space for diagnostics on data quality}.
 
@@
 Its boundaries (solid lines), given by light travel time for corresponding antenna spacings, are referred to as horizon delay limits.
-Regions excluding the {\it wedge} and narrow extensions of the {\it wedge} (white unshaded strips) caused by convolution with the {\bf instrument's spectral transfer function} are expected to be relatively free of foreground contamination.
-Due to an undesirable grating pattern (dot--dashed lines) of the instrument used in this study, namely the MWA, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the {\it EoR window} in this paper.
+Regions excluding the {\it wedge} and narrow extensions of the {\it wedge} (white unshaded strips) caused by convolution with the {\bf instrument's spectral transfer function} are expected to be relatively free of foreground contamination {\bf and are generally referred to as the {\it EoR window}.
+There are undesirable grating responses (dot--dashed lines) specific to the MWA.
+Hence, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the {\it MWA EoR window}}.
 \label{fig:fourier-space}}
@@
 Here we focus on antenna spacings $|\boldsymbol{b}| \le 200$~m (spatial scales relevant to reionization).
+{\bf Their deviation from coplanarity is negligible}.
 For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67.5\arcdeg \leq \theta_b < 112.5\arcdeg$.
@@
 \includegraphics[width=\linewidth]{figure3.eps}
-\caption{Delay power spectra of a uniform sky brightness distribution measured by antenna pairs with a uniform power pattern across the visible hemisphere.
-The non--zero response of the interferometer array to a uniform brightness distribution and the prominent {\it edge brightening} close to the horizon delay limits are due to wide--field effects.
-These effects are prevalent on all antenna spacings and are generic to all instruments used in wide--field measurements.
+\caption{Wide--field effects on delay power spectra produced with a uniform sky brightness distribution measured by antenna pairs with a uniform power pattern across the visible hemisphere.
+The non--zero response of the interferometer array to a uniform brightness distribution and the prominent {\it edge brightening} close to the horizon delay limits are wide--field effects.
+These are prevalent on all antenna spacings and are generic to all instruments used in wide--field measurements.
 \label{fig:usm}}
@@
 An alternate, but equivalent, interpretation is that baseline vectors (including those with largest lengths) are foreshortened towards the horizon along their orientation.
-Thus, they become sensitive to larger angular scales that match the inverse of their foreshortened baseline lengths.
+Thus, they become sensitive to larger angular scales that match the inverse of their foreshortened baseline lengths along these directions.
 
@@
 
-The model described in the preceding section is primarily a model of the diffuse foreground sky.
+The model described above is primarily a model of the diffuse foreground sky.
 While it contains faint compact emission blended in with the diffuse emission, {\bf bright point sources have been removed \citep{deo08}}.
@@
 
-The delay power spectra for these antenna shapes are shown in Figure~\ref{fig:antennas-delay-spectra}.
+The delay power spectra without thermal noise component for these antenna shapes are shown in Figure~\ref{fig:antennas-delay-spectra}.
 The occupancy of the power patterns on the sky is clearly correlated with that in the delay spectra.
@@
 The severity of foreground contamination inside the {\it foreground wedge} both in strength and occupancy also evidently decreases as the antenna element is changed from a dipole to a phased array to a dish.
-For instance, notice that the foreground contamination in $k$--modes between the $k_\parallel=0$ and the horizon limits decreases from $\sim 10^5$~K$^2\,$(Mpc/$h$)$^3$ in a phased array to $\sim 10$~K$^2\,$(Mpc/$h$)$^3$ in a dish.
+For instance, notice that the foreground contamination in $k$--modes between $k_\parallel=0$ and the horizon limits decreases from $\sim 10^5$~K$^2\,$(Mpc/$h$)$^3$ in a phased array to $\sim 10$~K$^2\,$(Mpc/$h$)$^3$ in a dish.
 As a consequence, $k$--modes in the {\it foreground wedge} that may be deemed too contaminated for EoR studies in the case of a dipole or a phased array can potentially become accessible when using a dish.
@@
 We have argued this arises due to wide--field effects.
-
 The prevalence of this feature across different antenna shapes demonstrates it is generic to all wide--field measurements, especially in EoR experiments.
@@
 Combined with the wide--field effects seen earlier, it leads to a ``pitchfork''--shaped signature inside the {\it foreground wedge}, as exemplified in the case of a dish.
-Although the exact appearance of this signature depends on antenna the power pattern, we use the term ``pitchfork'', hereafter, to broadly refer to the combination of foreground power in the primary field of view and the enhancement of foreground power near the horizon limits due to the nature of wide--field measurements.
+Although the exact appearance of this signature depends on antenna the power pattern, we use the term {\it pitchfork} hereafter, to broadly refer to the combination of foreground power in the primary field of view and the enhancement of foreground power near the horizon limits due to the nature of wide--field measurements.
 

### 2015-02-05 ms.tex
Start rewriting section 5
from 48d0c33:
============

Add figure6a.eps, figure6b.eps to git version control

In ms.tex,

(*) Start rewriting section 5

@@
 
-\section{Example Data}\label{sec:data}
+\section{MWA Data Modeling}\label{sec:data}
+
+We now use our simulations to analyze features in the delay power spectrum in data obtained using the MWA.
 
@@
 This yields a primary field of view $\gtrsim$~20\arcdeg~ wide and multiple secondary lobes.
-The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km \citep{bea12}.
-Here we focus on antenna spacings $|\boldsymbol{b}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz.
-The center frequency of 185~MHz corresponds to a redshift, $z\simeq 6.68$.
-{\bf At this frequency, we use $A_\textrm{e}=10.5$~m$^2$}.
+The array layout is described in \citet{bea12}.
+%{\bf At the observing frequency of 185~MHz, we use $A_\textrm{e}=10.5$~m$^2$}.
 
@@
 This process is repeated throughout the course of the observation $\approx 4.86$~hours.
-Figure~\ref{fig:pointings} illustrates how the apparent pointing oscillates around the desired pointing direction.
-
-The observations shown here were undertaken on 2013~August~23.
-% Data set from this day has been chosen as a reference for comparison within the MWA EoR collaboration because this is one of the days featuring a low number of known system errors from among a long stretch of stable observing.
+% Figure~\ref{fig:pointings} illustrates how the apparent pointing oscillates around the desired pointing direction.
+
+The observations {\bf used} here were undertaken on 2013~August~23.
 We have chosen two, 112 second long sections from this night for detailed study.
@@
 
-The {\it local sidereal time} (LST) at these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at {\bf 22.08~hours (wrapped to $-1.92$~hours in the figure)} and 0.09~hours, which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings, respectively.
-
-\begin{figure}[htb]
-\centering
-% \includegraphics[width=\linewidth]{figures/v1_0/custom_pointings.eps}
-% \includegraphics[width=\linewidth]{f2.eps}
-\includegraphics[width=\linewidth]{fig3.eps}
-\caption{MWA tile beam pointing directions during the course of the observation.
+These two pointings are at LST 22.08~hours and 0.09~hours, which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings, respectively.
+
+% The {\it local sidereal time} (LST) at these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at {\bf 22.08~hours (wrapped to $-1.92$~hours in the figure)} and 0.09~hours, which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings, respectively.
+
+% \begin{figure}[htb]
+% \centering
+% % \includegraphics[width=\linewidth]{figures/v1_0/custom_pointings.eps}
+% % \includegraphics[width=\linewidth]{f2.eps}
+% \includegraphics[width=\linewidth]{fig3.eps}
+% \caption{MWA tile beam pointing directions during the course of the observation.
 The $x$--axis refers to the {\it local sidereal time} (LST).
@@
 \label{fig:pointings}}
-\end{figure}
+% \end{figure}
 
@@
 
-\section{Simulations}\label{sec:modeling}
-
-We describe the instrumental and foreground models we use in our simulations.
-
-\subsection{Instrument Model}\label{sec:instrument_model}
-
-The instrument model consists primarily of the tile power pattern, $A(\hat{\boldsymbol{s}},f)$ (see equation~\ref{eqn:obsvis}).
-It is modeled as a mutually--coupled 4--by--4 dipole array with the overall power pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14_arXiv}.
+\subsection{Modeling}\label{sec:modeling}
+
+The MWA tile power pattern is modeled as a mutually--coupled 4--by--4 dipole array with the overall power pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14_arXiv}.
 To speed up simulations, we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles.
@@
 Besides having the effect of adding a time--dependent uncertainty in the power pattern, these random delay fluctuations reduce the coherence in the phased addition of dipole signals resulting in deviations from predicted models of the power pattern, most prominently at its nulls.
-The power patterns so obtained are shown in Figure~\ref{fig:power_pattern} for the pointings chosen in our study.
-
-\begin{figure}[htb]
-\centering
-% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom.eps}
-% \includegraphics[width=\linewidth]{f4.eps}
-\includegraphics[width=\linewidth]{fig5.eps}
-\caption{Models of MWA tile power pattern for {\it off--zenith} (top) and {\it zenith} pointings (bottom) at 185~MHz.
-An MWA tile is modeled as a 4$\times$4 array of isotropic radiators placed 0.3~m above an infinite ground plane.
-Random fluctuations of rms 0.05~ns have been added to delay corrections during the phased addition of voltages from these isotropic radiators.
-The logarithmic color scale (shown on the right) is common to both panels.
-\label{fig:power_pattern}}
-\end{figure}
+% The power patterns so obtained are shown in Figure~\ref{fig:power_pattern} for the pointings chosen in our study.
+
+We use the model described in \S\ref{sec:foreground} for the foreground sky.
+Figure~\ref{fig:sky-model-MWA} shows the diffuse emission and bright point source foreground models for the two chosen pointings with the modeled MWA tile power pattern contours overlaid.
+Notice the presence of a portion of the Galactic plane and the bright Galactic center in the westward sky in the diffuse sky model, where the MWA tile power gain is significant ($\gtrsim 12$\%).
+In the {\it zenith} pointing, the Galactic plane has set and the power pattern in that direction is at least 16 times smaller.
+
+\begin{figure*}[htb]
+\centering
+\subfloat[][Diffuse Sky Model]{\label{fig:DSM}\includegraphics[width=0.45\linewidth]{figure6a.eps}}
+\subfloat[][Bright Point Source Model]{\label{fig:CSM}\includegraphics[width=0.45\linewidth]{figure6b.eps}}
+\caption{Sky brightness temperature of the diffuse foreground model (left) and flux densities of bright point sources (right) at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+The color scales are logarithmic.
+Power pattern contours are overlaid.
+The contour levels shown are 0.00195, 0.00781, 0.0312, 0.125, and 0.5.
+The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off--zenith} pointing in the diffuse sky model and the MWA tile power gain is significant ($\gtrsim 12$\%) in that direction.
+In contrast, emission from the Galactic plane in {\it zenith} pointing is significantly lesser.}
+\label{fig:sky-model-MWA}
+\end{figure*}
 
@@
 
-\subsection{Foreground Model}\label{sec:foreground}
-
-The MWA has a very wide field of view ($\gtrsim 20\arcdeg$~ at 185~MHz) and has non--negligible response all the way to the horizon.
-As already noted, foreground objects fill the {\it foreground wedge} with an amplitude which depends on the tile power gain in that direction, with objects at the horizon being closest to the boundaries of the {\it wedge}.
-
-The degree to which these objects leak power into the spectral line--of--sight modes in the {\it EoR window} will be shown to depend primarily on the brightness of foregrounds near the boundaries of the {\it foreground wedge}.
-Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view, a point also supported by Pober~et~al.~(2015, in preparation).
-
-We use an all--sky foreground emission model that includes both diffuse and bright compact components.
-
-\subsubsection{Diffuse Foreground Model}\label{sec:DSM}
-
-For the diffuse component, we use an all--sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz.
-At this frequency, since this map is predominantly based on the 408~MHz map of \citet{has82} which has an angular resolution of 0.85\arcdeg, we smoothed the 185~MHz map to the same resolution.
-However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals.
-We model the diffuse foreground spectra with a single spectral index at each pixel in the map, found by comparing model maps at 170~MHz and 200~MHz.
-
-A low resolution version of the diffuse foreground model used is shown in Figure~\ref{fig:DSM}.
-Contours of the MWA tile power pattern shown in Figure~\ref{fig:power_pattern} are also overlaid.
-Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the westward sky, where the MWA tile power gain is significant ($\gtrsim 12$\%).
-In the {\it zenith} pointing, the Galactic plane has set and the power pattern in that direction is at least 16 times smaller.
-
-\begin{figure}[htb]
-\centering
-% \includegraphics[width=\linewidth]{figures/v1_0/dsm.eps}
-% \includegraphics[width=\linewidth]{f5.eps}
-\includegraphics[width=\linewidth]{fig6.eps}
-\caption{Sky brightness temperature of the diffuse foreground model at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
-The color scale on the right is logarithmic and is common to both panels.
-Contours of power pattern shown in Figure~\ref{fig:power_pattern} are overlaid.
-The contour levels shown are 0.00195, 0.00781, 0.0312, 0.125, and 0.5.
-The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off--zenith} pointing and the MWA tile power gain is significant ($\gtrsim 12$\%) in that direction.
-In contrast, emission from the Galactic plane in {\it zenith} pointing is significantly lesser.
-\label{fig:DSM}}
-\end{figure}
-
-\subsubsection{Compact Foreground Model}\label{sec:CSM}
-
-The model described in the preceding section is primarily a model of the diffuse foreground sky.
-While it contains faint compact emission blended in with the diffuse emission, {\bf bright point sources have been removed \citep{deo08}}.
-In order to supplement it with missing bright compact emission, we use classical radio source confusion estimates to determine the nominal flux density threshold and include point sources brighter than this threshold.
-It may be noted that slightly different criteria are in common use in radio astronomy to estimate radio source confusion \citep[see Appendix of][and references therein]{thy13}.
-For an angular resolution of 0.85\arcdeg, using a conservative `$S_\textrm{c}=5\sigma_\textrm{c}$' criterion, we determine the flux density threshold to be $\approx 10$~Jy.
-Other liberal criteria that yield a lower threshold carry a greater risk of double--counting point sources which might be already blended in with the diffuse sky model.
-
-We use a combination of the NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and the Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz to provide our point source catalog due to their complementary survey footprints covering the entire sky, and matched flux density sensitivity and angular resolution.
-The SUMSS catalog covers the sky with declination $\delta < -30\arcdeg$~ with a limiting peak brightness of 6--10~mJy/beam and an angular resolution of $\sim 45$\arcsec.
-The NVSS covers the sky with $\delta > -40\arcdeg$~ with a similar angular resolution and a limiting flux density of $\approx 2.5$~mJy for point sources.
-
-From the SUMSS catalog, we select {\bf objects} whose deconvolved major axes are equal to 0\arcsec, {\bf thereby strictly selecting point sources}.
-From the NVSS catalog, we excluded objects that overlap with those in the SUMSS survey footprint.
-Point sources from NVSS were selected if the convolved major axes were not greater than $\approx 47$\arcsec, which matches the angular resolution of the survey.
-Using a mean spectral index of $\langle\alpha_\textrm{sp}\rangle=-0.83$ (flux density, $S(f)\propto f^{\alpha_\textrm{sp}}$) obtained by \citet{mau03} for both NVSS and SUMSS catalog objects, we calculate the corresponding flux densities at 185~MHz, $S_{185}$.
-From this subset, we choose point sources with $S_{185}\geq 10$~Jy.
-The selection of such bright point sources is not affected by minor differences in flux density sensitivity of the two surveys.
-We verified that our selection criteria ensure a similar areal density of objects in the two surveys.
-
-Based on these criteria, we selected 100 objects from the SUMSS catalog and 250 objects from the NVSS catalog.
-Together with the diffuse foreground model, we obtain an all--sky foreground model consisting of both compact and diffuse emission.
-Figure~\ref{fig:CSM} shows the compact foreground emission model used in our study whose with contours of the power pattern overlaid.
-The flux densities of the point sources are indicated by the color scale.
-
-\begin{figure}[htb]
-\centering
-% \includegraphics[width=\linewidth]{figures/v1_0/csm.eps}
-% \includegraphics[width=\linewidth]{f6.eps}
-\includegraphics[width=\linewidth]{fig7.eps}
-\caption{Flux densities of bright point sources at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
-The color scale on the right is logarithmic and is common to both panels and corresponds to the color--coded filled circles.
-Contours of power pattern shown in Figure~\ref{fig:power_pattern} are overlaid.
-The contour levels shown are 0.001953125, 0.0078125, 0.03125, 0.125, and 0.5.
-\label{fig:CSM}}
-\end{figure}
-
-It is worth emphasizing that the ideal foreground model is one with an angular resolution finer than that of the instrument that covers the whole sky at the observing frequency, a point we will revisit later.
+% It is worth emphasizing that the ideal foreground model is one with an angular resolution finer than that of the instrument that covers the whole sky at the observing frequency, a point we will revisit later.
 For this study, we note that our conclusions are not significantly affected despite minor differences in the thresholds from different radio source confusion estimates.

### 2015-02-05 ms.tex
Continue editing section 5
from 8d355ca:
============

(*) Update figure5a.eps, figure5b.eps
(*) Add figure7.eps to git version control

In ms.tex,

(*) Continue editing section 5

@@
 
-Figure~\ref{fig:fhd_data} shows {\bf delay power spectra} from MWA data after deconvolution along the delay axis.
+% Figure~\ref{fig:fhd_data} shows {\bf delay power spectra} from MWA data after deconvolution along the delay axis.
 The principal feature is the {\it foreground wedge}, whose boundaries, marked by white lines, increase linearly with light travel time proportional to baseline length.
@@
 
-\begin{figure}[htb]
-\centering
-% \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
-% \includegraphics[width=\linewidth]{f3.eps}
-\includegraphics[width=\linewidth]{fig4.eps}
-\caption{{\bf Delay power spectra} of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
+% \begin{figure}[htb]
+% \centering
+% % \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
+% % \includegraphics[width=\linewidth]{f3.eps}
+% \includegraphics[width=\linewidth]{fig4.eps}
+% \caption{{\bf Delay power spectra} of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
 {\bf In both cases, the delay spectra are phased to the zenith and thus the {\it foreground wedge} boundaries are symmetric around $\tau=0$}.
@@
 \label{fig:fhd_data}}
-\end{figure}
+% \end{figure}
 
@@
 With the aforementioned foreground model, and instrumental and observational parameters, we simulate visibilities using equation~\ref{eqn:obsvis}.
-Figure~\ref{fig:sim_data} shows the {\bf delay power spectra} from {\it off--zenith} and {\it zenith} pointings.
-Notice the qualitative agreement of amplitude and structure with those obtained from data shown in Figure~\ref{fig:fhd_data}.
-The Galactic center and the Galactic plane visible in the {\it off--zenith} pointing make it appear brighter in the {\it foreground wedge}.
-
-\begin{figure}[htb]
-\centering
-% \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
-% \includegraphics[width=\linewidth]{f7.eps}
-\includegraphics[width=\linewidth]{fig8.eps}
-\caption{Same as Figure~\ref{fig:fhd_data}, but generated from simulated visibilities using the model described in \S\ref{sec:modeling}.
-Note the similarity between the broad features to those in Figure~\ref{fig:fhd_data}.
-\label{fig:sim_data}}
-\end{figure}
+Figure~\ref{fig:fhd-sim-comparison} shows the {\bf delay power spectra} from {\it off--zenith} and {\it zenith} pointings obtained from MWA observations and modeling.
+Notice the qualitative agreement of amplitude and structure between the two.
+The Galactic center and the Galactic plane visible in the {\it off--zenith} pointing make it appear brighter in the {\it foreground wedge} as a branch with $\tau<0$.
+
+\begin{figure*}[htb]
+\centering
+\includegraphics[width=\linewidth]{figure7.eps}
+\caption{Delay power spectra from MWA data (left) and modeling (right) for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+The modeling matches the data to a level consistent with the uncertainties in foreground models and the antenna beam.
+\label{fig:fhd-sim-comparison}}
+\end{figure*}
 
@@
 
-These uncertainties are presented only to confirm the qualitative agreement already seen between Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}.
+These uncertainties are presented only to confirm the qualitative agreement already seen between data and modeling in Figure~\ref{fig:fhd-sim-comparison}.
 These estimates are conservative.
 A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power pattern \citep{ber15}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide--field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations, but is beyond the scope of this paper.
-{\bf Hereafter}, our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.
+{\bf Hereafter, our focus} is to explore in detail the foreground signatures embedded in the {\it foreground wedge} of the MWA instrument.
 

### 2015-02-05 ms.tex
Complete section 5 on MWA data, modeling analysis
from a4e4df6:
============

Add figure8.eps, figure9.eps to git version control

In ms.tex,

(*) Make minor changes to captions of figures 5 and 7
(*) Write subsection on foreground signature analysis under S5
(*) Complete section 5 now consisting of the entire MWA data, modeling,
comparison and foreground analysis including the "pitchfork" section
(*) Remove section "Results"

@@
 The strength and occupancy of the power patterns are correlated with those of delay power spectra.
+White dotted lines in the delay power spectra mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
 The {\it foreground wedge} and the {\it EoR window} are most severely contaminated in the case of the dipole while it is the least for the dish.
@@
 \caption{Delay power spectra from MWA data (left) and modeling (right) for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
-The modeling matches the data to a level consistent with the uncertainties in foreground models and the antenna beam.
+The {\it foreground wedge} is bounded by white dotted lines.
+Model matches the data to a level consistent with the uncertainties in foreground models and the antenna beam.
 \label{fig:fhd-sim-comparison}}
@@
 This demonstrates that a fractional deviation of $\sim$~90\% observed between data and simulations is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account.
-Currently, we are primarily limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.
+% Currently, we are primarily limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.
 
@@
 
-\section{Results}\label{sec:delay-spectrum-analysis}
+\subsection{Analysis of Foreground Signatures}\label{sec:FG-analysis}
 
@@
 
-\item {\it Sky Model}: Our model of the sky consists of compact and diffuse emission on diverse spatial scales as shown in Figures~\ref{fig:DSM} and \ref{fig:CSM}.
+\item {\it Sky Model}: Our model of the sky consists of diffuse and compact emission on diverse spatial scales as shown in Figures~\ref{fig:DSM} and \ref{fig:CSM}, respectively.
 It also consists of localized regions of strong emission such as the Galactic plane and Galactic center.
-% The resulting sky emission is anisotropic.
 In fact, the patch of sky for MWA observations is chosen from regions of low foreground emission.
@@
 \item {\it Tile Pointing and Power Pattern}: Since the MWA tiles are steered electronically, the tile power pattern tile changes in any observing mode that tracks the {\bf pointing center}.
-In our study, we take into account the effect of the changing tile power pattern on the {\bf delay power spectra} (Figure~\ref{fig:power_pattern}).
-The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose the {\it off--zenith} and {\it zenith} pointings shown in Figure~\ref{fig:pointings}.
-% We will show the secondary lobe structures in the tile power pattern are tied to the foreground signatures in Fourier space.
+In our study, we take into account the effect of the changing tile power pattern on the {\bf delay power spectra}.
 
@@
 
-\subsection{Diffuse Foreground Signatures}\label{sec:diffuse}
-
-Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the {\bf delay power spectra} simulated with only the diffuse foreground model.
-The top and bottom panels correspond to {\it off--zenith} and {\it zenith} pointings, respectively.
-
-\begin{figure}[htb]
-\centering
-% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-% \includegraphics[width=\linewidth]{f8.eps}
-\includegraphics[width=\linewidth]{fig9.eps}
-\caption{Simulated {\bf delay power spectra (in units of K$^2$~(Mpc/$h$)$^3$} for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise.
-White lines mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing.
-The axes and color scale are identical to those in Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}.
-In the {\it zenith} pointing, {\bf delay power spectrum} from diffuse emission has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings.
-\label{fig:noiseless-dsm-delay-spectrum}}
-\end{figure}
-
-Here we examine in detail some examples of interesting features observed with the diffuse foreground model.
+Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum} shows the delay power spectra obtained from the diffuse (left) and compact (right) foreground emission for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings without thermal noise component.
+Some of the notable signatures are discussed below.
+
+\begin{figure*}[htb]
+\centering
+\includegraphics[width=\linewidth]{figure8.eps}
+\caption{Simulated {\bf delay power spectra (in units of K$^2$~(Mpc/$h$)$^3$} for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse (left) and compact (right) foreground models without any thermal noise.
+The axes and color scale are identical to those in Figure~\ref{fig:fhd-sim-comparison}.
+In the {\it off-zenith} pointing, emission from the Galactic center is the most prominent feature seen as a branch at $\tau<0$.
+In the {\it zenith} pointing, {\bf delay power spectrum} from diffuse emission has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings due to wide--field effects.
+Compact emission is centrally concentrated.
+This combination gives rise to a {\it pitchfork} signature in the {\it foreground wedge}.
+\label{fig:noiseless-dsm-csm-delay-spectrum}}
+\end{figure*}
 
@@
 
-The most prominent signature seen in the {\it off--zenith} pointing (top panel of Figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power pattern.
+The most prominent signature seen in the {\it off--zenith} pointing (top left panel of Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power pattern.
 It appears as a bright branch near the negative delay horizon delay limit.
 This feature is strongest at short antenna spacings and fades with increasing antenna spacing.
-The {\bf bright} signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-delay-spectrum}) because the Galactic center is below the horizon.
-
-\subsubsection{Diffuse Emission Delay Spectra}\label{sec:diffuse-features}
-
-Figure~\ref{fig:noiseless-dsm-delay-spectrum} {\bf also shows diffuse emission in the primary field of view manifesting in the {\it off--zenith} pointing as a branch at $\tau>0$, and at $\tau=0$ in the {\it zenith} pointing.
-The former is seen at $\tau>0$ because the primary lobe of the power pattern is centered eastward of zenith, whereas in the latter it is centered at zenith}.
+The {\bf bright} signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) because the Galactic center is below the horizon.
+
+\subsubsection{Ubiquitous Diffuse Emission}\label{sec:diffuse-features}
+
+Diffuse emission outside the Galactic plane manifests in the primary field of view as a branch at $\tau>0$ and $\tau=0$ in the {\it off--zenith} and {\it zenith} pointings respectively.
+The former is seen at $\tau>0$ because the primary lobe of the power pattern is centered eastward of zenith, whereas in the latter it is centered at zenith.
 As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inversely proportional to the length of the baseline projected in the direction of the emission.
@@
 
-{\bf In typical radio interferometric measurements, due to the Fourier relation between sky brightness and measured visibilities, only short antenna spacings are sensitive to emission from large size scales such as the diffuse component in our study (angular scales $\gtrsim$~0.85\arcdeg).
-However, we note a very interesting signature of diffuse emission revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing, and at both horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\boldsymbol{b}|\le$~200~m).
-This is a consequence of wide--field measurements, where long baselines are foreshortened due to projection effects towards off--axis emission, especially near the horizon.
-Consequently, they become sensitive to diffuse emission on large size scales.
-This holds for short baselines as well and is evident at all LSTs in our simulations}.
+In both pointings the diffuse emission (left panels) is prominent near the horizon delay limits extending to the widest antenna spacings.
+This is a characteristic signature of the wide--field effects discussed in \S\ref{sec:wide-field}.
+It is evident at all LSTs in our simulations.
 Thus, diffuse emission from far off--axis directions manifests as an edge--heavy {\it two--pronged fork} across all baselines.
-It decreases in strength with increasing baseline length as expected but is present in all baseline orientations.
-
-{\bf \citet{thy13} and \citet{pob13} have found a similar feature in their simulations and data respectively.
-In the latter, this feature is described as ``edge brightening'' and attributed to the Galactic plane near the horizon.
-From their discussion, it is unclear what fraction of power in that feature arises from ubiquitous diffuse emission.
-On the other hand, \citet{thy13} statistically modeled point sources below the confusion limit and attributed the feature to an increase in solid angles subtended by delay bins closer to the horizon limits thereby, enhancing the resulting integrated emission.
-This must be true for diffuse emission as well.
-Thus we argue the feature arises primarily due to an increase in sensitivity of long baselines to diffuse emission caused by foreshortening towards angles far off--axis.
-Consequently, the increased size scales (or solid angles) containing higher levels of integrated emission, and spatial scales inversely proportional to the foreshortened baseline lengths become matched}.
-
-\subsection{Compact Foreground Signatures}\label{sec:compact}
-
-Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the {\bf delay power spectra} of the compact foreground sky model (see Figure~\ref{fig:CSM}) for {\it off--zenith} and {\it zenith} pointings in the top and bottom panels, respectively.
-In contrast to the {\bf delay power spectra} of diffuse emission, compact emission manifests as a center--heavy structure in either pointing.
-
-\begin{figure}[htb]
-\centering
-% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-% \includegraphics[width=\linewidth]{f9.eps}
-\includegraphics[width=\linewidth]{fig10.eps}
-\caption{Same as Figure~\ref{fig:noiseless-dsm-delay-spectrum} but for a foreground model consisting of compact emission.
-Compact emission gives rise to centrally concentrated features and stays flat across antenna spacings in the {\it foreground wedge}.\label{fig:noiseless-csm-delay-spectrum}}
-\end{figure}
-
-The amplitude response of an interferometer to a point source is, to first order, flat with respect to baseline length.
+It decreases in strength with increasing baseline length but is nevertheless present in all baseline orientations.
+
+\subsubsection{Compact Foreground Signatures}\label{sec:compact}
+
+In contrast to the {\bf delay power spectra} of diffuse emission, compact emission (right panels) manifests as a center--heavy structure in either pointing.
+
+The amplitude response of an interferometer to a point source is, to first order, flat cross baseline length.
 Since the primary lobe of the power pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with $\tau>0$ corresponding to the position of the primary lobe of the power pattern.
@@
 
-When not dominated by the bright emission from the Galactic center, which happens only for certain pointings at certain times, the {\bf delay power spectrum} of the combined foreground model is composed of diffuse and compact emission, {\bf both of which are significant}.
+When not dominated by the bright emission from the Galactic center, the {\bf delay power spectrum} of the combined foreground model is composed of diffuse and compact emission, {\bf both of which are significant}.
 This is illustrated by a more detailed examination of the {\it zenith} pointing in our study.
 
-Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional {\bf delay power spectra} plots shown in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum}, and \ref{fig:noiseless-csm-delay-spectrum}.
+Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional {\bf delay power spectra} plots shown in Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}.
 The diffuse, compact, and composite components are shown as solid red, cyan, and black lines, respectively.
@@
 \centering
-% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_3_baseline_comparison_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
-% \includegraphics[width=\linewidth]{f10.eps}
-\includegraphics[width=\linewidth]{fig11.eps}
+\includegraphics[width=\linewidth]{figure9.eps}
 \caption{Simulated {\bf delay power spectra} for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}).
-The antenna spacing is specified in each panel along East, North and Upward directions.
+The baseline length and orientation is specified in each panel.
 The solid red, cyan, and black lines denote contributions from diffuse, compact, and composite foreground models respectively.
@@
 The shaded region denotes the envelope of uncertainty around the composite model delay power spectrum.
-This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits and by thermal noise fluctuations elsewhere}.
+This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits and by thermal noise fluctuations outside}.
 \label{fig:pitchfork-baselines}}
@@
 
-Near the horizon delay limits at shorter antenna spacing, the diffuse component is brighter relative to the compact component.
+Near the horizon delay limits, the diffuse component is brighter relative to the compact component.
 Here, diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at zero--delay.
 In fact, even on widely spaced antennas, diffuse emission in the {\bf delay power spectrum} near the horizon delay limits exceeds that in the primary lobe by about three orders of magnitude.
-We attribute this to apparent shortening of antenna spacings projected in the direction of diffuse emission entering far from the primary field of view, thereby giving rise to the edge--heavy {\it two--pronged fork} feature discussed in \S\ref{sec:diffuse-long-baselines}.
-
-Simulations with the complete foreground model show the combination of center--heavy features on all antenna spacings from compact emission, and edge--heavy features from both compact and diffuse emission especially the latter even on wide antenna spacings.
+This feature is described in \S\ref{sec:diffuse-long-baselines}.
+We attribute this to wide--field measurement effects discussed in \S\ref{sec:wide-field}.
+
+Simulations with the complete foreground model show the combination of center--heavy features dominated by compact emission in primary field of view, and edge--heavy features from both types of emission especially the diffuse component near the horizon.
 This results in a characteristic {\it pitchfork} structure imprinted in the {\it foreground} wedge and should be evident in observations.
 
-All these observations are consistent with those from Figures~\ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}.
 The observability of the {\it pitchfork} signature predicted in this paper depends on the relative levels of uncertainty in the foreground model and fluctuations from thermal noise.
-In our simulations, since thermal noise in these very short duration snapshots is $\sim 10^4$~K$^2$(Mpc/$h$)$^3$ and features near the horizon delay limits are also of comparable amplitudes, the {\it pitchfork} feature is not expected to be detected in a noisy scenario although this feature is marginally visible in the {\it zenith} pointing of observed data (see Figure~\ref{fig:fhd_data}).
+In our simulations, since thermal noise in these very short duration snapshots is $\sim 10^4$~K$^2$(Mpc/$h$)$^3$ and features near the horizon delay limits are also of comparable amplitudes, the {\it pitchfork} feature is not expected to be detected in a noisy scenario, although this feature is marginally visible in the {\it zenith} pointing of observed data (see Figure~\ref{fig:fhd-sim-comparison}).
 We attribute this to differences between our foreground model and the actual sky.
@@
 
-We also note that increasing the antenna spacing results in progressively improving the resolution along the delay axis by increasing the number of delay bins inside the {\it foreground wedge}.
+We also note that increasing the antenna spacing progressively improves the resolution along the delay axis by increasing the number of delay bins inside the {\it foreground wedge}.
 This improves the localization of foreground objects whose signatures are imprinted in the {\bf delay power spectrum}.
-For instance, there is an increase in the number of secondary peaks in the {\bf delay power spectrum} between zero--delay and horizon delay limits as the antenna spacing increases from $\sim$84~m to $\sim$171~m.
+For instance, there is an increase in the number of secondary peaks in the {\bf delay power spectrum} between $\tau=0$ and horizon delay limits as the antenna spacing increases from $\sim$84~m to $\sim$171~m.
 In this case, these correspond to secondary lobes of the power pattern that lie between the primary lobe and the horizon along the local meridian.
 At short antenna spacings, due to relatively lower resolution along the delay axis inside the {\it foreground wedge} and a consequent loss of localization of foreground emission, these secondary peaks blend in with other major peaks and are not distinctly visible.
-These findings are consistent with the discussion in \S\ref{sec:compact}.
 
 In summary, the brightest feature $\sim 10^{12}$~K$^2$(Mpc/$h$)$^3$ is that of the Galactic center in the {\it off--zenith} pointing co--located at a westward secondary lobe with a significantly high gain.
-The next brightest signature $\gtrsim 10^8$~K$^2$(Mpc/$h$)$^3$ is caused by compact emission appearing to be concentrated in the inner regions of the {\it foreground wedge} (around $\tau=0$) rather than towards the boundaries.
-Diffuse emission co--located at the primary lobe of the power pattern is $\sim 10^3$ fainter relative to compact emission from the same region for a $\sim 84$~m antenna spacing.
+The next brightest signature $\sim 10^8$~K$^2$(Mpc/$h$)$^3$ is caused by compact emission {\bf from the primary field of view} appearing to be concentrated in the inner regions of the {\it foreground wedge} (around $\tau=0$).
+Diffuse emission from this region is $\sim 10^3$ fainter relative to that from compact emission for a $\sim 84$~m antenna spacing.
 But unlike the latter, diffuse emission decreases rapidly by over four orders of magnitude as the antenna spacing is widened to $\sim 171$~m.
-However, diffuse emission is significantly higher near horizon delay limits compared to that in the primary lobe of the power pattern as the antenna spacing is increased.
-This leads to edge--heavy features in the {\bf delay power spectra}.
-Complemented by the center-heavy compact foreground features, we predict the presence of a {\it three--pronged pitchfork}--shaped signature in the {\bf delay power spectrum} of the foreground sky.
-We identify evidence for presence of this signature in actual MWA data, although deeper observations are needed to more clearly reveal the effect.
+However, diffuse emission near the horizon is significantly boosted by wide--field effects compared to that in the primary field of view and is prominent across all antenna spacings.
+This results in a {\it three--pronged pitchfork}--shaped signature inside the {\it foreground wedge} (see also Figure~\ref{fig:antennas-delay-spectra}).
 
@@
 Here, we will investigate the susceptibility of particular antenna spacings to foreground contamination arising out of bright foreground objects located near the horizon and present a technique to substantially mitigate such contamination.
+We use the MWA as an example.
 

### 2015-02-05 ms.tex
Complete section 6 on baseline screening technique
from c202e24:
============

In ms.tex,

(*) Complete section 6 on baseline screening technique
(*) Remove earlier section 7 on role of antenna geometry

@@
 There are undesirable grating responses (dot--dashed lines) specific to the MWA.
-Hence, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the {\it MWA EoR window}}.
+Hence, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the MWA {\it EoR window}}.
 \label{fig:fourier-space}}
@@
 
-\section{Applications}\label{sec:fg-grading}
-
-Here, we will investigate the susceptibility of particular antenna spacings to foreground contamination arising out of bright foreground objects located near the horizon and present a technique to substantially mitigate such contamination.
+\section{Recipe: Baseline--Based Foreground Mitigation}\label{sec:fg-grading}
+
+Here, we investigate the susceptibility of particular antenna spacings to foreground contamination arising out of bright foreground objects located near the horizon and present a technique to substantially mitigate such contamination.
 We use the MWA as an example.
@@
 Figure~\ref{fig:delay-map} shows the sky mapped to delays registered by the baseline vectors, of length 100~m for instance, oriented northward (top panel) and eastward (bottom panel).
-Since the Galactic center is located in the western sky, this figure demonstrates that it is observed at $\tau<0$ on a baseline vector oriented eastward and at $\tau=0$ on a baseline vector oriented northward.
+Since the Galactic center is located in the western sky, this figure demonstrates that it is observed at $\tau<0$ on a baseline oriented eastward and at $\tau=0$ on a baseline oriented northward.
 Figure~\ref{fig:baseline-breakup} shows the delay spectra on baselines oriented northward ($67.5\arcdeg\le \theta_\textrm{b} < 112.5\arcdeg$) at the top and eastward ($-22.5\arcdeg~\le \theta_\textrm{b} < 22.5\arcdeg$) at the bottom.
@@
 \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{fig12c.eps}}
-\caption{(a): Sky model showing compact ({\it top}) and diffuse ({\it bottom}) emission.
-Contours of power pattern are overlaid whose levels are identical to those in Figures~\ref{fig:DSM} and \ref{fig:CSM}.
-The Galactic center is very prominent in diffuse emission in the western sky and is co--located with a bright secondary lobe of the power pattern.
+\caption{(a): Sky model showing compact ({\it top}) and diffuse ({\it bottom}) emission (adopted from Figure~\ref{fig:sky-model-MWA}).
+The Galactic center is very prominent in diffuse emission on the west co--located with a bright secondary lobe of the power pattern.
 (b): Sky hemisphere mapped to delays observed on antenna spacings with northward ({\it top}) and eastward ({\it bottom}) orientations.
+Delays vary linearly with antenna spacing length.
 Color scale shown is for a 100~m antenna spacing.
-Delays vary linearly with antenna spacing length.
-The bright Galactic center in the western sky is recorded at $\tau=0$ in northward antenna spacings and close to negative horizon delay limit on eastward antenna spacings.
+The bright Galactic center will appear at $\tau=0$ in northward antenna spacings and close to negative horizon delay limit on eastward antenna spacings.
 (c): Simulated delay spectra {\bf power} on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}).
@@
 
-The {\bf delay power spectrum} not only carries information on spatial scales of foreground emission but also offers the unique advantage of viewing the sky through a combination of antenna spacing vectors and the delay axis.
-With a foreground model known {\it a priori} in which structures and locations of very bright foreground objects such as the Galactic center or AGN are available, we can predict their response across antenna spacings as a function of observing parameters such as LST, power pattern, etc.
+% The {\bf delay power spectrum} not only carries information on spatial scales of foreground emission but also offers the unique advantage of viewing the sky through a combination of antenna spacing vectors and the delay axis.
+With a foreground model known {\it a priori} in which structures and locations of very bright foreground objects such as the Galactic center or AGN are available, the power of the delay spectrum technique lets us predict the response across antenna spacings as a function of observing parameters such as LST, power pattern, etc.
 This allows us to programmatically screen data for antenna spacings that are severely contaminated by foregrounds near the horizon delay limits.
@@
 In our example, we discard eastward antenna spacings ($-22.5\arcdeg\le\theta_\textrm{b}< 22.5\arcdeg$) of lengths $|\boldsymbol{b}| < 30$~m.
-{\bf This was found to be the limit beyond which foreground contamination levels are insensitive to any further removal of wider antenna spacings}.
+{\bf This was found to be the limit beyond which foreground contamination is insensitive to any further removal of wider antenna spacings}.
 Figure~\ref{fig:before-after} shows the delay spectra obtained with all antenna spacings (top panel) and after applying our screening technique (bottom panel) on the {\it off--zenith} observation.
 Notice the remarkable reduction in foreground spillover into the {\it EoR window} via the removal of bright vertical stripes on short eastward antenna spacings.
-The color scale range has been adjusted relative to the previous figures to show this significant reduction in foreground contamination.
+% The color scale range has been adjusted relative to the previous figures to show this significant reduction in foreground contamination.
 
@@
 Discarded antenna spacings (black vertical stripes) have lengths $|\boldsymbol{b}|<30$~m (leftward of vertical dashed line) and orientations $|\theta_\textrm{b}|<$~22.5\arcdeg.
-The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} prominent when all antenna spacings are used is lowered by {\bf nearly two orders} of magnitude when short eastward antenna spacings are discarded.
-Color scale has been adjusted to emphasize mitigation achieved in foreground contamination.
+The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} is lowered by {\bf nearly two orders} of magnitude when short eastward antenna spacings are discarded.
 \label{fig:before-after}}
@@
 This screening technique can be generalized to optimize between foreground mitigation and loss of sensitivity from discarding data.
-Figure~\ref{fig:screening} shows how the typical foreground contamination\footnote{Foreground contamination is measured by standard deviation of noiseless $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$ from foregrounds in the {\it EoR window}.} in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
+Figure~\ref{fig:screening} shows how the typical foreground contamination\footnote{Foreground contamination is measured by standard deviation of noiseless $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$ from foregrounds in the MWA {\it EoR window}.} in the MWA {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
 We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., $-7.5\arcdeg~\le\theta_\textrm{b} < 7.5\arcdeg$ (solid circles), $-15\arcdeg\le\theta_\textrm{b}< 15\arcdeg$ (solid squares), and $-22.5\arcdeg\le\theta_\textrm{b}< 22.5\arcdeg$ (solid stars).
 Among antenna spacings that satisfy these criteria, we discard data from those whose lengths are shorter than $|\boldsymbol{b}|_\textrm{max}$ ($x$--axis) and show foreground contamination estimated in the {\it EoR window} from all remaining antenna spacings.
-In other words, these plots demonstrate the progress in mitigation as orientation and maximum length of discarded antenna spacings are varied.
-% Foreground contamination estimated from only northward (dashed line) and eastward (dot--dashed line) antenna spacings represent the lower and upper limits, respectively, and are shown for reference.
-The fraction of discarded antenna spacings discarded relative to the total number is shown in dotted lines for different ranges of $\theta_\textrm{b}$.
-It is seen that foreground contamination can be mitigated by a factor between $\sim 2$ ($|\theta_\textrm{b}|\le$~7.5\arcdeg) and $\sim 100$ ($|\theta_\textrm{b}|\le$~22.5\arcdeg).
-The latter limit is achieved with a mere 5\% loss of data for $|\boldsymbol{b}|_\textrm{max}\simeq30$~m.
-And discarding antenna spacings with lengths $|\boldsymbol{b}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity as the fraction of discarded baselines increases from $\sim 5$\% to $\sim 25$\%.
-On the other hand, a bright point source at the same location will give rise to foreground contamination even on longer antenna spacings.
-Such cases will necessitate discarding more or, in the worst case, all of the eastward oriented antenna spacings.
-{\bf These results are specific to the MWA data.
-In general, the array configuration will be a significant factor in determining such thresholds}.
 
@@
 
+In other words, Figure~\ref{fig:screening} demonstrates the progress in foreground mitigation as orientation and maximum length of discarded antenna spacings are varied.
+The fraction of discarded antenna spacings discarded relative to the total number is shown in dotted lines for different ranges of $\theta_\textrm{b}$.
+It is seen that foreground contamination can be mitigated by a factor between $\sim 2$ ($|\theta_\textrm{b}|\le$~7.5\arcdeg) and $\sim 100$ ($|\theta_\textrm{b}|\le$~22.5\arcdeg).
+The latter limit is achieved with a mere 5\% loss of data for $|\boldsymbol{b}|_\textrm{max}\simeq30$~m.
+Discarding antenna spacings with lengths $|\boldsymbol{b}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity as the fraction of discarded baselines increases from $\sim 5$\% to $\sim 25$\%.
+On the other hand, a bright point source at the same location will give rise to foreground contamination even on longer antenna spacings.
+Such cases will necessitate discarding more or, in the worst case, all of the eastward oriented antenna spacings.
+{\bf These results use the MWA as an example.
+In general, the array layout will be a significant factor in determining such thresholds}.
+
 In principle, instead of discarding selected antenna spacings altogether, we could down--weight them based on an optimal scheme.
@@
 
-\section{Role of Antenna Geometry}\label{sec:antenna-shape}
-
-{\bf We now examine the influence of the antenna power pattern on the delay power spectrum (refer to equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}).
+% \section{Role of Antenna Geometry}\label{sec:antenna-shape}
+
+% {\bf We now examine the influence of the antenna power pattern on the delay power spectrum (refer to equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}).
 In the following analysis, we assume an ideal flat passband over the entire 30.72~MHz bandwidth while still applying a {\it Blackman--Harris} window function.
@@
 
-\subsection{Uniformly Illuminated Sky}\label{sec:USM}
-
-{\bf Simulations of the response to a uniformly illuminated sky confirm that the {\it pitchfork} effect is independent of the specific foreground model chosen in our study.
+% \subsection{Uniformly Illuminated Sky}\label{sec:USM}
+
+% {\bf Simulations of the response to a uniformly illuminated sky confirm that the {\it pitchfork} effect is independent of the specific foreground model chosen in our study.
 The brightness distribution on the sky was fixed to be a constant equal to the mean over the whole sky at 185~MHz with no spectral variation.
@@
 
- % In the following discussion, we assume there are no random delay fluctuations in the MWA tile beamforming process.
+%  % In the following discussion, we assume there are no random delay fluctuations in the MWA tile beamforming process.
 The model power pattern so obtained is different from the one we used earlier, but the former still serves as a good reference (Figure~\ref{fig:power_pattern}) for the present discussion.
@@
 
-Na\"{i}vely, only single element antennas are sensitive to a uniform brightness distribution while interferometers are not.
+% Na\"{i}vely, only single element antennas are sensitive to a uniform brightness distribution while interferometers are not.
 However, an examination of equation~\ref{eqn:obsvis} shows it resembles a Fourier transform when the sky brightness is expressed in a coordinate system of direction cosines \citep{tay99,tho01}.
@@
 
-Figure~\ref{fig:noiseless-usm-delay-spectrum} shows the delay power spectra of such a sky model with the model power pattern of an ideal MWA tile.
+% Figure~\ref{fig:noiseless-usm-delay-spectrum} shows the delay power spectra of such a sky model with the model power pattern of an ideal MWA tile.
 Although not identical, these delay power spectra closely resemble those obtained with the diffuse sky model (see Figure~\ref{fig:noiseless-dsm-delay-spectrum}).
@@
 
-\begin{figure}[htb]
-\centering
-\includegraphics[width=\linewidth]{fig15.eps}
-\caption{Similar to Figure~\ref{fig:noiseless-dsm-delay-spectrum} but for a uniform sky brightness model.
+% \begin{figure}[htb]
+% \centering
+% \includegraphics[width=\linewidth]{fig15.eps}
+% \caption{Similar to Figure~\ref{fig:noiseless-dsm-delay-spectrum} but for a uniform sky brightness model.
 The features bear a close resemblance to those obtained with the diffuse emission model.
@@
 \label{fig:noiseless-usm-delay-spectrum}}
-\end{figure}
-
-{\bf The edge--heavy {\it two--pronged} features are evident, which are very similar to those seen earlier in the case of diffuse emission.
+% \end{figure}
+
+% {\bf The edge--heavy {\it two--pronged} features are evident, which are very similar to those seen earlier in the case of diffuse emission.
 As reasoned in \S\ref{sec:diffuse-long-baselines}, the foreshortening of long baselines causes them to be sensitive to emission from near the horizon.
 
-Branches of foreground power seen at $\tau>0$ and close to the negative horizon delay limit in the {\it off--zenith} pointing correspond to radiation received from the direction of the primary lobe and the far eastward sidelobe, respectively, in the power pattern.
+% Branches of foreground power seen at $\tau>0$ and close to the negative horizon delay limit in the {\it off--zenith} pointing correspond to radiation received from the direction of the primary lobe and the far eastward sidelobe, respectively, in the power pattern.
 On the other hand, the symmetry of the delay power spectrum around $\tau=0$ seen in the {\it zenith} pointing is as expected from a symmetric power pattern}.
 
-\subsection{Experiments with Aperture Shapes}\label{sec:apertures}
-
-{\bf The power pattern of an antenna is determined by the shape of its aperture.
+% \subsection{Experiments with Aperture Shapes}\label{sec:apertures}
+
+% {\bf The power pattern of an antenna is determined by the shape of its aperture.
 Through the use of a few examples, we examine the role the geometrical shape of the aperture plays in shaping the characteristics of delay power spectrum.
 We consider the following antenna elements placed at the MWA tile locations:
-\begin{itemize}
-\item {\it Dipole}: An East--West dipole of length 0.74~m at a height 0.3~m above a ground plane. $A_\textrm{e}=(\lambda/2)^2$.
-\item {\it Phased Array}: A 4$\times$4 array of isotropic radiators at a height 0.3~m above the ground plane placed in an arrangement similar to that in an MWA tile. $A_\textrm{e}=16\,(\lambda/2)^2$. 
-\item {\it Dish}: Diameter of 14~m similar to that proposed for HERA, with $A_\textrm{e}\approx 154$~m$^2$.
+% \begin{itemize}
+% \item {\it Dipole}: An East--West dipole of length 0.74~m at a height 0.3~m above a ground plane. $A_\textrm{e}=(\lambda/2)^2$.
+% \item {\it Phased Array}: A 4$\times$4 array of isotropic radiators at a height 0.3~m above the ground plane placed in an arrangement similar to that in an MWA tile.
+$A_\textrm{e}=16\,(\lambda/2)^2$.
+% \item {\it Dish}: Diameter of 14~m similar to that proposed for HERA, with $A_\textrm{e}\approx 154$~m$^2$.
 The power pattern is simulated using an {\it Airy} pattern where its sensitivity beyond the horizon is forced to zero.
-\end{itemize}
-
-The power patterns of these antenna geometries at 185~MHz for the {\it zenith} pointing are shown in Figure~\ref{fig:power-patterns}.
+% \end{itemize}
+
+% The power patterns of these antenna geometries at 185~MHz for the {\it zenith} pointing are shown in Figure~\ref{fig:power-patterns}.
 It must be noted that the presence of a ground plane below the dipole and the phased array introduces a smoother suppression of sensitivity near the horizon, whereas this transition is abrupt in the case of the dish}.
 
-% \begin{figure*}[htb]
-% \centering
-% \subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{fig16a.eps}} \\
-% \subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{fig16b.eps}}
-% \caption{Power patterns (top panels) and simulated delay power spectra (bottom panels) for different antenna shapes at 185~MHz centered on zenith.
+% % \begin{figure*}[htb]
+% % \centering
+% % \subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{fig16a.eps}} \\
+% % \subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{fig16b.eps}}
+% % \caption{Power patterns (top panels) and simulated delay power spectra (bottom panels) for different antenna shapes at 185~MHz centered on zenith.
 Antenna shapes used are: dipole (left), phased array (middle), and dish (right).
@@
 The amplitude of this feature is highest for the dipole and least for the dish and hence, strongly depends on the shape of the antenna element.}
-% \label{fig:aperture-shapes}
-% \end{figure*}
-
-{\bf The delay power spectra for these antenna shapes are shown in Figure~\ref{fig:antennas-delay-spectra}.
+% % \label{fig:aperture-shapes}
+% % \end{figure*}
+
+% {\bf The delay power spectra for these antenna shapes are shown in Figure~\ref{fig:antennas-delay-spectra}.
 The occupancy of the power patterns in the sky hemisphere is clearly correlated with that in the delay spectra.
@@
 
-The levels of foreground contamination in the {\it EoR window} varies substantially across the different antenna shapes: $\sim 10^4\,$K$^2\,$(Mpc/$h$)$^3$, $\lesssim 10^2\,$K$^2\,$(Mpc/$h$)$^3$, and $<1\,$K$^2\,$(Mpc/$h$)$^3$ while using a dipole, phased array, and a dish, respectively.
+% The levels of foreground contamination in the {\it EoR window} varies substantially across the different antenna shapes: $\sim 10^4\,$K$^2\,$(Mpc/$h$)$^3$, $\lesssim 10^2\,$K$^2\,$(Mpc/$h$)$^3$, and $<1\,$K$^2\,$(Mpc/$h$)$^3$ while using a dipole, phased array, and a dish, respectively.
 The severity of foreground contamination inside the {\it foreground wedge} both in strength and occupancy also evidently decreases as the antenna element is changed from a dipole to a phased array to a dish.
@@
 
-Finally, a distinct feature common to all these aperture shapes is that the foreground contamination near the horizon delay limits even on wide antenna spacings is significant ($\gtrsim 10^5$~K$^2$(Mpc/$h$)$^3$).
+% Finally, a distinct feature common to all these aperture shapes is that the foreground contamination near the horizon delay limits even on wide antenna spacings is significant ($\gtrsim 10^5$~K$^2$(Mpc/$h$)$^3$).
 This is contrary to expectation for two reasons --- the power pattern of the antenna near the horizon is negligible compared to the primary lobe region; and, a significant fraction of foreground emission near the horizon delay limits on wide antenna spacings arises from diffuse emission, to which usually only short antenna spacings are sensitive.

### 2015-02-06 ms.tex
Complete summary section
from 3ed4ca5:
============

In ms.tex,

(*) Complete summary section
(*) Add emphasis on wide-field, pitchfork effects
(*) Add emphasis on effect of antenna shape on pitchfork signature

@@
 Our primary motivation in this work is to understand how the various bright foregrounds will manifest in three dimensional power spectrum of H{\sc i} from 21~cm reionization observations.
-When the signal is expressed in units of temperature variance, the dynamic range between bright foregrounds and the 21~cm signal is expected to be $\sim 10^8$; a detailed understanding of how foregrounds can corrupt the 21~cm power spectrum is therefore essential.
+{\bf In units of temperature variance}, the dynamic range between bright foregrounds and the 21~cm signal is expected to be $\sim 10^8$; a detailed understanding of how foregrounds can corrupt the 21~cm power spectrum is therefore essential.
 This analysis extends previous work by simulating the entire sky rather than just the central field of view and by providing a comparison with early observations with the MWA.
-By making use of the delay spectrum technique to estimate the power spectrum, we are able to observe the effects of foregrounds while avoiding entanglements with more complex power spectrum estimators.
-
-Simulating in many important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright point sources from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise fluctuations in measurements.
+By making use of the delay spectrum technique to estimate the power spectrum, we are able to observe the effects of foregrounds while avoiding entanglements with more {\bf sophisticated} power spectrum estimators.
+
+{\bf We find that all wide--field instruments, typical of modern EoR observatories, imprint a characteristic {\it two--pronged fork} signature in delay spectra.
+There are two equivalent interpretations: delay bins near the horizon subtend larger solid angles and therefore contain larger integrated emission; or, foreshortening of baselines towards the horizon makes them sensitive to emission on large angular scales which match the inverse of their foreshortened lengths.
+These effects combined with higher sensitivity of antennas in the primary field of view results in a characteristic {\it pitchfork} signature.
+The amplitude of these generic signatures can be controlled by careful design of antenna aperture.
+In contrast to a dipole and a phased array such as an MWA tile, a dish such as the one proposed for HERA is found to yield the least foreground contamination and thus preferable for EoR studies.}
+
+Simulating in many important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright point sources from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA {\bf to within expected uncertainties in foreground models}.
 
@@
 
-As expected, diffuse emission in the primary lobe of the antenna power pattern is prominent on shorter antenna spacings.
-However, the most interesting result is its footprint on wide antenna spacings near the horizon delay limits --- an edge--heavy {\it two--pronged fork}--shaped signature.
-This is due to apparent shortening of wide antenna spacings in the direction of foreground emission far off--axis, thereby retaining their response to {\bf emission on large size scales}.
+{\bf Diffuse emission in the primary field of view is prominent on shorter antenna spacings.
+However, it is also prominent near the horizon limits even on wide antenna spacings --- an effect of the wide--field nature of the measurement.
 On the other hand, compact emission predominantly maps onto central regions of the {\it foreground wedge}.
 Features arising from compact emission co--located with primary and secondary lobes of the antenna power pattern have been identified.
-In general, {\bf delay power spectrum} signatures of compact emission are center--heavy, in clear contrast to those from diffuse emission.
-A composite foreground model consisting of diffuse and compact foregrounds combines the characteristic individual shapes into a {\it pitchfork}--shaped structure in the {\it foreground wedge}.
-This will be distinctly visible when the thermal noise floor is sufficiently lowered, as longer observations are processed through the MWA analysis pipelines.
-{\bf We conclude the {\it pitchfork} feature to be generic to all wide--field measurements, including EoR experiments}.
-
-{\bf The role of antenna shape in determining the {\bf delay power spectrum} was also studied.
+In general, {\bf delay power spectrum} signatures of compact emission are center--heavy while those of diffuse emission are edge--heavy which results in the {\it pitchfork} signature.
+This will be distinctly visible when the thermal noise floor is sufficiently lowered, as longer observations are processed.}
+
+% {\bf The role of antenna shape in determining the {\bf delay power spectrum} was also studied.
 In contrast to a dipole and a phased array such as an MWA tile, a dish such as the one proposed for HERA is found to yield the least foreground contamination both in the {\it foreground wedge} and the {\it EoR window} and thus, preferable for EoR studies}.

### 2015-02-06 ms.tex
Complete Introduction section
from 23a1bee:
============

In ms.tex,

(*) Complete Introduction section

@@
 
+{\bf We investigate signatures generic to all wide--field measurements of EoR power spectrum and explore antenna shapes that are most desirable from the viewpoint of foreground contamination for future EoR experiments}.
 After matching observations from the Murchison Widefield Array with instrumental simulations of the entire sky, we examine the {\bf shape of the delay spectrum} of each baseline.
@@
 We then arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.
-{\bf We also examine the role of antenna shapes in determining foreground contamination and provide inputs on antenna design for future EoR experiments}.
 
 In \S\ref{sec:delay-spectrum} we {\bf provide a brief overview of the delay spectrum technique} \citep{par12a,par12b}.
-In \S\ref{sec:data}, we describe the MWA setup, summarize the observing parameters, and present the resulting data.
-Simulations using these observing parameters are described in \S\ref{sec:modeling} and analyzed by sky location and type of emission in \S\ref{sec:delay-spectrum-analysis}.
+We establish the signatures arising from the nature of wide--field measurements in \S\ref{sec:wide-field}.
+In \S\ref{sec:sim}, we present the foreground model and a variety of instrument models to rank antenna shapes based on the criterion of foreground contamination.
+In \S\ref{sec:MWA}, we describe the MWA setup, summarize the observing parameters, and present the resulting data.
+Simulations using these observing parameters are compared with the data and analyzed for foreground signatures.
 In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique.
-{\bf The role of antenna shapes in determining foreground contamination is explored in \S\ref{sec:antenna-shape}}.
 We present a summary of our work and findings in \S\ref{sec:summary}.
@@
 
-\section{MWA Data Modeling}\label{sec:data}
+\section{MWA Data Modeling}\label{sec:MWA}
 

### 2015-02-06 ms.tex
Move delay deconvolution to delay spectrum section
from c25e8ca:
============

In ms.tex,

(*) Move Delay spectrum deconvolution section to end of Delay Spectrum
section and simplify text to be more general

@@
 
+\subsection{Delay Spectrum Deconvolution}\label{sec:data-delay-spectrum}
+
+We obtain the delay spectrum of visibilities by taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f$ to be a {\it Blackman--Harris} window function.
+The spectrum is multiplied by the additional weights, $W_f$, consisting of instrumental passband and channels possibly flagged for interference.
+In delay--space, these weights translate into a convolution by a point spread function (PSF).
+We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis \citep{par09,par12b} to obtain the final delay spectra.
+The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights.
+We limit the selection of peaks to modes inside the horizon delay limit, corresponding to smooth spectrum objects in the visible sky hemisphere.
+
 \section{Wide--Field Measurements}\label{sec:wide-field}
@@
 Finally, we fit for an additional phase known to be caused by small reflections in a subset of cables.
-
-\subsection{Deconvolved Delay Spectrum}\label{sec:data-delay-spectrum}
-
-We obtain the delay spectrum of these calibrated visibilities by taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f$ to be a {\it Blackman--Harris} window function.
-The spectrum is multiplied by the additional weight of the flagged channels which occur rarely for interference and regularly every 1.28~MHz where the edges of coarse channels are known to have a small aliasing effect.
-In delay--space, these weights translate into a convolution by a grated point spread function (PSF) with gratings at multiples of $\tau=0.78\,\mu$s.
-We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis by \citet{par09,par12b}.
-The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights.
-We limit the selection of peaks to modes inside the horizon delay limit, corresponding to smooth spectrum objects in the visible sky hemisphere.
-
-% In subsequent sections, we plot the amplitude of the delay spectrum, $|V_{b\tau}|$, sorted by baseline length, $|\boldsymbol{b}|$.
-Note that visibilities from different baselines have not been averaged together in this analysis.
-
-% Figure~\ref{fig:fhd_data} shows {\bf delay power spectra} from MWA data after deconvolution along the delay axis.
-The principal feature is the {\it foreground wedge}, whose boundaries, marked by white lines, increase linearly with light travel time proportional to baseline length.
-Note that the wedge is not filled out uniformly.
-Each baseline has a different orientation that determines the response to different objects in the sky.
-It is this variation that we will explore below.
-
-The {\it off--zenith} pointing {\bf has notably higher power in the {\it foreground wedge} than in the {\it zenith} pointing}.
-This is shown later in \S\ref{sec:diffuse} to be due to response of the array to the bright Galactic center and Galactic plane in the westward sky.
-Consequently, the contamination into the {\it EoR window} is also higher, especially on short antenna spacings.
-Emission in the {\it zenith} pointing is more centrally concentrated {\bf due to a zenith--centered power pattern and faster suppression of response away from zenith}, whereas the {\it off--zenith} pointing has a power pattern phased eastward of zenith and is also more responsive to emission farther away from the primary lobe relative to that in the {\it zenith} pointing.
-{\bf The power pattern is discussed in detail in \S\ref{sec:instrument_model}}.
-
-% \begin{figure}[htb]
-% \centering
-% % \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
-% % \includegraphics[width=\linewidth]{f3.eps}
-% \includegraphics[width=\linewidth]{fig4.eps}
-% \caption{{\bf Delay power spectra} of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set.
-{\bf In both cases, the delay spectra are phased to the zenith and thus the {\it foreground wedge} boundaries are symmetric around $\tau=0$}.
-For reference, the $k$--space equivalent of $\tau$ and $|\boldsymbol{b}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively.
-White lines mark the boundaries of {\it foreground wedge} determined by light travel time across each baseline.
-Smooth spectrum objects predominantly reside in this {\it foreground wedge}.
-The logarithmic color scale (shown at the right) is common to both panels.
-\label{fig:fhd_data}}
-% \end{figure}
 

### 2015-02-08 ms.tex
Make changes suggested by Judd
from eb548c5:
============

In ms.tex,

(*) Change title to mention "Fourier delay space" and "Wide-field"
observatories
(*) Change abstract slightly to mention "Fourier delay space"
(*) Change introduction slightly
(*) Change Delay Spectrum section slightly
(*) Change Wide-field effects section slightly
(*) Re-number subsections in Simulations section and change slightly
(*) Trim "Analysis of Foreground Signatures" section by a few paragraphs
and change "MWA Data Modeling" section slightly
(*) Change title and wording of "Baseline based foreground mitigation"
se

@@
 %% \title{Limits on the detection of the Epoch of Reionization from MWA observations of the redshifted 21~cm line}
-\title{Foregrounds in Fourier Space for Redshifted 21~cm Observatories}
+\title{Foregrounds in Fourier Delay Space for Wide--Field Redshifted 21~cm Observatories}
 
@@
 For signals entering through the primary field of view, compact emission dominates the foreground contamination.
-These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space.
+These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier delay space.
 Based on these results, we propose that selective down--weighting of data based on antenna spacing and time can mitigate foreground contamination substantially by a factor $\sim 100$ with negligible loss of sensitivity.
@@
 This commenced the epoch of reionization (EoR) --- a period of non--linear growth of matter density perturbations and astrophysical evolution.
-Studying the EoR holds the key to understanding this evolution.
-
-Observations of redshifted 21~cm radiation generated by the spin flip transition of H{\sc i} has been identified as a direct probe of the EoR \citep{sun72,sco90,mad97,toz00,ili02}.
+% Studying the EoR holds the key to understanding this evolution.
+
+Observing redshifted 21~cm radiation generated by the spin flip transition of H{\sc i} has been identified as a direct probe of the EoR \citep{sun72,sco90,mad97,toz00,ili02}.
 Detecting this signal has recently emerged as a very promising experiment to fill the gaps in our understanding of the Universe's history.
@@
 
-Considerable effort has been and continues to be made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,dil14,liu14a,liu14b}.
+Considerable effort is being made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,dil14,liu14a,liu14b}.
 {\bf A model that provides a generic explanation for the observed foreground power spectrum has emerged, whereby the wide--field (and chromatic) response of the instrument causes the power in smooth spectrum foregrounds to occupy higher $k$--modes into the so-called `wedge'}.
-% A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide--field (and chromatic) response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum.
-The conservative foreground strategy, referred to as {\it avoidance}, that has developed alongside this work is to discard $k$--modes which could be contaminated \citep[for e.g.,][]{par14}.
-The more aggressive alternative is to subtract a sky model and regain access to modes discarded by {\it avoidance}.
+The conservative foreground strategy, referred to as {\it avoidance}, that has developed alongside this work is to discard $k$--modes which could be contaminated \citep[e.g.,][]{par14}.
+The more aggressive alternative is to subtract a sky model and regain access to modes that would be discarded by {\it avoidance}.
 In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain.
@@
 Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity.
-To decouple these effects from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using a {\it \textbf{Per-Baseline}} approach of \citet{par12b}.
+To decouple these effects from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using a {\it \textbf{per-baseline}} approach of \citet{par12b}.
 This approximates the power spectrum as the inverse Fourier transform of the spectra generated by the instrument's correlator.
 
-{\bf We investigate signatures generic to all wide--field measurements of EoR power spectrum and explore antenna shapes that are most desirable from the viewpoint of foreground contamination for future EoR experiments}.
+{\bf We investigate signatures generic to all wide--field measurements of EoR power spectra and explore antenna shapes that are most desirable from the viewpoint of foreground contamination for future EoR experiments}.
 After matching observations from the Murchison Widefield Array with instrumental simulations of the entire sky, we examine the {\bf shape of the delay spectrum} of each baseline.
-We report two important findings: the foregrounds that most severely {\bf obscure} the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted.
+We report two important findings: in the {\it avoidance} strategy, the foregrounds that most severely {\bf obscure} the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted.
 We quantify these by separating the simulations into components based on type of foreground emission.
@@
 Interferometer array data known as {\it visibilities}, $V_{bf}\equiv V_{bf}(\boldsymbol{b},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis.
-If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power pattern, respectively, at different frequencies as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_f\equiv W_f(f)$ denotes instrumental bandpass weights, then $V_{bf}$ can be written as (with a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}):
+If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power pattern, respectively, at different frequencies as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_f\equiv W_f(f)$ denotes instrumental bandpass weights, then $V_{bf}$ can be written as:
 \begin{align}\label{eqn:obsvis}
@@
 where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector.
-We wish to {\bf note} that this equation is valid {\bf in general, including wide--field measurements}.
+This equation is valid {\bf in general, including wide--field measurements and is a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}}.
 
@@
 where, $D\equiv D(z)$ is the transverse comoving distance at redshift $z$.
-% Hence, $V_{uf}=V_{uf}(\boldsymbol{b}/\lambda,f)=V_{bf}(\lambda\boldsymbol{u},f)$.
-It is important to note that for a given antenna pair, $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not.
 
@@
 
-% Radio emission from foreground objects, due to their spatial and spectral (chromatic) structure, contaminate the received signal \citep{bow09,dat10,mor12,tro12,thy13,liu14a,liu14b}. 
 Foregrounds can be described in both the delay spectrum and Fourier frameworks.
-For our study, which primarily concerns with foreground characterization, we find the former to be simple and yet extremely useful, while maintaining a close correspondence with the latter despite subtle differences.
-In order to express a quantity derived from $V_{u\eta}$ that has units identical to that of the cosmological H{\sc i} power spectrum, we define delay power spectrum:
+For our study, we find the former to be simple and yet extremely useful.
+In order to express a quantity derived from $V_{u\eta}$ that has units identical to that of the cosmological H{\sc i} power spectrum, we define the delay power spectrum:
 \begin{align}\label{eqn:fake-power-spectrum}
@@
 where, $\boldsymbol{u}$ and $\eta$ in equations~\ref{eqn:k-perp} and \ref{eqn:k-prll} have been correspondingly replaced with $\boldsymbol{b}/\lambda$ and $\tau$ respectively.
-
-In summary, the delay spectrum, $V_{b\tau}$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}.
-$V_{b\tau}$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument.
-At the same time, it is closely related to the sought power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$ containing critical information about spatial scales.
-% In this paper, we often use the quantity $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$, and sometimes $V_{b\tau}$.
 
@@
 Hence, the brightest EoR signal will be observed on the shortest baselines and smallest delays.
-Thus the region of interest for EoR studies is just beyond the horizon delay limits (dashed lines) on short baselines, marked as the region of {\it maximal EoR sensitivity}}.
-
-{\bf In the specific case of the MWA which has a passband constructed using coarse channels, there are period grating responses resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.
-Thus with reference to the MWA, the {\it EoR window} is assumed to lie outside the dashed lines but inside the first grating response ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) and is shown in medium shade.
-The MWA EoR collaboration actively uses this as a test region in Fourier space for diagnostics on data quality}.
+Thus the region of interest for EoR studies relying on {\it avoidance} strategy is just beyond the horizon delay limits (dashed lines) on short baselines, marked as the region of {\it maximal EoR sensitivity}}.
+
+{\bf In the specific case of the MWA, which has a passband constructed using coarse channels, there are period grating responses resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.
+Thus with reference to the MWA, the {\it EoR window} lies outside the dashed lines but inside the first grating response ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) and is shown in medium shade}.
 
@@
 We obtain the delay spectrum of visibilities by taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f$ to be a {\it Blackman--Harris} window function.
-The spectrum is multiplied by the additional weights, $W_f$, consisting of instrumental passband and channels possibly flagged for interference.
+The sky spectrum is multiplied by weights, $W_f$, due to the instrumental passband and channels possibly flagged for interference.
 In delay--space, these weights translate into a convolution by a point spread function (PSF).
@@
 
-When the synthesized field is small, where $A(\hat{\boldsymbol{s}},f)$ or $I(\hat{\boldsymbol{s}},f)$ is negligible for $|l| \ll 1$ and $|m| \ll 1$, equation~\ref{eqn:obsvis-lmn} reduces to a simple two--dimensional Fourier transform \citep{tay99,tho01} between the apparent sky brightness and measured visibilities.
+When the synthesized field is small, where $A(\hat{\boldsymbol{s}},f)$ or $I(\hat{\boldsymbol{s}},f)$ is significant only for $|l| \ll 1$ and $|m| \ll 1$, equation~\ref{eqn:obsvis-lmn} reduces to a simple two--dimensional Fourier transform \citep{tay99,tho01} between the apparent sky brightness and measured visibilities.
 It is in this context that radio interferometers are understood to be sensitive only to fluctuations and not to a uniform sky brightness distribution.
@@
 Hence, the approximations in the narrow--field scenario do not apply.
-For example, even if $A(\hat{\boldsymbol{s}},f)$ and $I(\hat{\boldsymbol{s}},f)$ are held constant across the visible hemisphere, the integrand in equation~\ref{eqn:obsvis-lmn} is still dependent on direction.
+For example, even if $A(\hat{\boldsymbol{s}},f)$ and $I(\hat{\boldsymbol{s}},f)$ are held constant across the visible hemisphere, the amplitude of the integrand in equation~\ref{eqn:obsvis-lmn} is still dependent on direction.
 Therefore, in a significant departure from a narrow--field measurement, the wide--field visibility from a uniform brightness distribution on a non--zero antenna spacing is not zero.
@@
 
-We interpret this is due to equal--sized delay bins subtending larger solid angles near the horizon thereby containing larger integrated emission.
+We interpret this as due to equal--sized delay bins subtending larger solid angles near the horizon thereby containing larger integrated emission.
 An alternate, but equivalent, interpretation is that baseline vectors (including those with largest lengths) are foreshortened towards the horizon along their orientation.
-Thus, they become sensitive to larger angular scales that match the inverse of their foreshortened baseline lengths along these directions.
-
-\citet{thy13} have found evidence of this feature in their statistical models.
+Thus, they become sensitive to larger angular scales that match the inverse of their foreshortened lengths along these directions.
+
+\citet{thy13} found evidence of this feature in their statistical models.
 In line with our reasoning, they attribute it to a steep rise in solid angles subtended by delay bins near the horizon limits.
-\citet{pob13} also find a similar ``edge brightening'' feature which they attribute to emission from the Galactic plane near the horizon.
+\citet{pob13} also find a similar ``edge brightening'' feature which they attribute to Galactic plane emission near the horizon.
 From their discussion, it is unclear what fraction of power in that feature arises from such wide--field effects.
@@
 The nature of the specific instrument used for observing will control the amplitude of these effects, which we explore below.
-% Typically, the foreground sky consists of diffuse emission and point-like sources, and the antenna is more sensitive to power received from primary field of view than from the rest of the sky.
 
@@
 
-\subsubsection{Diffuse Foreground Model}\label{sec:DSM}
+% \subsubsection{Diffuse Foreground Model}\label{sec:DSM}
 
@@
 However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals.
-We model the diffuse foreground spectra with a spectral index at each pixel in the map, estimated from model maps at 170~MHz and 200~MHz.
-
-\subsubsection{Compact Foreground Model}\label{sec:CSM}
+We model the diffuse foreground spectra with a unique spectral index at each pixel in the map, estimated from model maps at 170~MHz and 200~MHz.
+
+% \subsubsection{Compact Foreground Model}\label{sec:CSM}
 
@@
 In order to supplement it with missing bright compact emission, we use classical radio source confusion estimates to determine the nominal flux density threshold and include point sources brighter than this threshold.
-It may be noted that slightly different criteria are in common use in radio astronomy to estimate radio source confusion \citep[see Appendix of][and references therein]{thy13}.
+Slightly different criteria are in common use in radio astronomy to estimate radio source confusion \citep[see Appendix of][and references therein]{thy13}.
 For an angular resolution of 0.85\arcdeg, using a conservative `$S_\textrm{c}=5\sigma_\textrm{c}$' criterion, we determine the flux density threshold to be $\approx 10$~Jy.
@@
 
-Based on these criteria, we select 100 objects from the SUMSS catalog and 250 objects from the NVSS catalog.
+These criteria yield 100 objects from the SUMSS catalog and 250 objects from the NVSS catalog.
 Together with the diffuse foreground model, we obtain an all--sky foreground model consisting of both compact and diffuse emission.
 Figure~\ref{fig:sky-model-generic} shows the diffuse (top) and compact (bottom) foreground emission model used in our study.
-In this snapshot at 0.09~hours LST, the Galactic center in the diffuse model appears to have just set in the West.
+In this snapshot at 0.09~hours LST, the Galactic center in the diffuse model has just set in the West.
 
@@
 \item {\it Dipole}: An East--West dipole of length 0.74~m at a height 0.3~m above a ground plane. $A_\textrm{e}=(\lambda/2)^2$.
-\item {\it Phased Array}: A 4$\times$4 array of isotropic radiators with a grid spacing of 1.1~m at a height 0.3~m above the ground plane placed in an arrangement similar to that in an MWA tile.
+\item {\it Phased Array}: A 4$\times$4 array of isotropic radiators with a grid spacing of 1.1~m at a height 0.3~m above the ground plane placed in an arrangement similar to that of an MWA tile.
 $A_\textrm{e}=16\,(\lambda/2)^2$.
@@
 The power patterns of these antenna geometries at 185~MHz for the {\it zenith} pointing are shown in Figure~\ref{fig:power-patterns}.
-It must be noted that the presence of a ground plane below the dipole and the phased array introduces a smoother suppression of sensitivity near the horizon, whereas this transition is abrupt in the case of the dish.
+The presence of a ground plane below the dipole and the phased array introduces a smoother suppression of sensitivity near the horizon, whereas this transition is abrupt in the case of the dish.
 
@@
 The occupancy of the power patterns on the sky is clearly correlated with that in the delay spectra.
-The following visual correlations are also apparent --- the strength of the primary lobe centered on the pointing center is correlated with the delay power spectrum centered on $\tau=0$; and, the overall rate of decrease in the power sensitivity away from the pointing center is correlated with the rate of drop in power away from $\tau=0$.
-
-The levels of foreground contamination in the {\it EoR window} varies substantially across the different antenna shapes: $\sim 10^4\,$K$^2\,$(Mpc/$h$)$^3$, $\lesssim 10^2\,$K$^2\,$(Mpc/$h$)$^3$, and $<1\,$K$^2\,$(Mpc/$h$)$^3$ while using a dipole, phased array, and a dish, respectively.
+Further, the strength of the primary lobe centered on the pointing center is correlated with the delay power spectrum centered on $\tau=0$; and, the overall rate of decrease in the power sensitivity away from the pointing center is correlated with the rate of drop in power away from $\tau=0$.
+
+The levels of foreground contamination in the {\it EoR window} varies substantially across the different antenna shapes: $\sim 10^4\,$K$^2\,$(Mpc/$h$)$^3$, $\lesssim 10^2\,$K$^2\,$(Mpc/$h$)$^3$, and $<1\,$K$^2\,$(Mpc/$h$)$^3$ for the dipole, phased array, and dish, respectively.
 The severity of foreground contamination inside the {\it foreground wedge} both in strength and occupancy also evidently decreases as the antenna element is changed from a dipole to a phased array to a dish.
@@
 
-Finally, a distinct feature common to all these aperture shapes is that the foreground contamination near the horizon delay limits even on wide antenna spacings is significant ($\gtrsim 10^5$~K$^2$(Mpc/$h$)$^3$).
+Finally, a distinct feature common to all these aperture shapes is that the foreground contamination near the horizon delay edges is significant even on wide antenna spacings ($\gtrsim 10^5$~K$^2$(Mpc/$h$)$^3$).
 We have argued this arises due to wide--field effects.
-The prevalence of this feature across different antenna shapes demonstrates it is generic to all wide--field measurements, especially in EoR experiments.
+The prevalence of this feature across different antenna shapes demonstrates it is generic to all wide--field measurements.
 The amplitude of this effect, however, can be controlled via choice of antenna shape and through weighting of aperture illumination.
@@
 Combined with the wide--field effects seen earlier, it leads to a ``pitchfork''--shaped signature inside the {\it foreground wedge}, as exemplified in the case of a dish.
-Although the exact appearance of this signature depends on antenna the power pattern, we use the term {\it pitchfork} hereafter, to broadly refer to the combination of foreground power in the primary field of view and the enhancement of foreground power near the horizon limits due to the nature of wide--field measurements.
+Although the exact appearance of this signature depends on the antenna power pattern, we use the term {\it pitchfork} hereafter, to broadly refer to the combination of foreground power in the primary field of view and the enhancement of foreground power near the horizon limits due to the nature of wide--field measurements.
 
@@
 
-We now use our simulations to analyze features in the delay power spectrum in data obtained using the MWA.
+We now use our simulations to analyze features in the delay power spectrum obtained using the MWA.
 
@@
 MWA construction was completed in 2012 and, after commissioning, began its EoR observing program in 2013.
-In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz.
+In its final configuration, the MWA is a 128--tile interferometer capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz.
 Each tile is a phased array of 16 dipoles, each in the shape of a bow--tie.
@@
 The MWA passband of width $\Delta B=30.72$~MHz is divided coarsely into 24$\times$1.28~MHz sub--bands with each sub--band weighted by a digital filter.
-The coarse channel shape is obtained using a 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$.
+The coarse channel shape is obtained using an 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$.
 {\bf Each of these coarse bands consists of 32 fine channels of width 40~kHz each}.
 After correcting for the shape of these coarse channels, the fine channels at the edges of these sub--bands are flagged because they are known to be contaminated by aliasing at a low level.
-Any spectral flags, or other known weights constitutes $W_f$ in equation~\ref{eqn:obsvis}.
-
-% Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=95$~K per polarization for all frequency channels for all antenna pairs throughout the course of the observation.
-Estimation of $T_\textrm{sys}$ from MWA data is a subject of active investigation.
-In our work, we have chosen a $T_\textrm{sys}$ that matches the thermal noise observed in the data.
+% Any spectral flags, or other known weights constitutes $W_f$ in equation~\ref{eqn:obsvis}.
 
@@
 
-{\bf The MWA is expected to be sensitive to the power spectrum of the H{\sc i} signal over the redshift range $6<z<10$ \citep{thy13,bea13}.
+{\bf The MWA is expected to be sensitive to the power spectrum of the H{\sc i} signal over the redshift range $6<z<10$ \citep{bow06,thy13,bea13}.
 Over 600~hours have been currently observed using the MWA with the objective of characterizing the nature of the sources that are responsible for ionizing the IGM, charting the evolution of the global neutral fraction, and probing the nature of quasar emissions by constraining the properties of their ionized proximity zones.}
@@
 The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of the dipoles in an MWA tile.
-The pointing system is capable of steering to points on a regular $\sim$7\arcdeg~ grid.
-During observations we allow the sky to drift across the nearest available pointing, shifting between grid points as necessary (once in $\sim 30$~minutes).
+The pointing system is optimized to points on a regular $\sim$7\arcdeg~ grid.
+The sky drifts across to the nearest available pointing, shifting between grid points (once every $\sim 30$~minutes).
 This process is repeated throughout the course of the observation $\approx 4.86$~hours.
-% Figure~\ref{fig:pointings} illustrates how the apparent pointing oscillates around the desired pointing direction.
-
-The observations {\bf used} here were undertaken on 2013~August~23.
+
+The observations {\bf used} here were acquired on 2013~August~23.
 We have chosen two, 112 second long sections from this night for detailed study.
@@
 These two pointings are at LST 22.08~hours and 0.09~hours, which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings, respectively.
-
-% The {\it local sidereal time} (LST) at these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at {\bf 22.08~hours (wrapped to $-1.92$~hours in the figure)} and 0.09~hours, which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings, respectively.
-
-% \begin{figure}[htb]
-% \centering
-% % \includegraphics[width=\linewidth]{figures/v1_0/custom_pointings.eps}
-% % \includegraphics[width=\linewidth]{f2.eps}
-% \includegraphics[width=\linewidth]{fig3.eps}
-% \caption{MWA tile beam pointing directions during the course of the observation.
-The $x$--axis refers to the {\it local sidereal time} (LST).
-The axis on the left refers to longitudes, namely, Right Ascension (RA) and Hour Angle (HA).
-Negative values of RA, HA, and LST are to be interpreted as having been wrapped around by 360\arcdeg~ or 24 hours.
-The axis on the right refers to the declination of the pointing direction.
-The RA, HA, and declination are plotted with solid, dashed and dotted lines respectively.
-The sky is allowed to drift for $\sim 30$~mins before the beamformer delay settings change in a discrete step to center the beam around RA = 0$^\textrm{h}$, Dec = $-30$\arcdeg.
-Dot--dashed vertical lines show the LST of {\it off-zenith} and {\it zenith} pointings used in our study.
-\label{fig:pointings}}
-% \end{figure}
 
@@
 Besides having the effect of adding a time--dependent uncertainty in the power pattern, these random delay fluctuations reduce the coherence in the phased addition of dipole signals resulting in deviations from predicted models of the power pattern, most prominently at its nulls.
-% The power patterns so obtained are shown in Figure~\ref{fig:power_pattern} for the pointings chosen in our study.
 
@@
 
-{\bf Estimation of $T_\textrm{sys}$ from MWA data is a subject of active investigation.
-For our work, we estimate thermal noise in the data using the {\it rms} of $V_{b\tau}$ obtained after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
+{\bf We estimate thermal noise in the data using the {\it rms} of $V_{b\tau}$ obtained after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
 \begin{align}\label{eqn:Tsys}
@@
 Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise observed in the data.}
-
-% It is worth emphasizing that the ideal foreground model is one with an angular resolution finer than that of the instrument that covers the whole sky at the observing frequency, a point we will revisit later.
-For this study, we note that our conclusions are not significantly affected despite minor differences in the thresholds from different radio source confusion estimates.
 
@@
 We also simulated delay spectra after assigning spectral indices drawn randomly from a {\it gaussian} distribution with a mean of $\langle\alpha_\textrm{sp}\rangle=-0.83$ and a HWHM of $\Delta\alpha_\textrm{sp}=0.35$ to the point sources in our compact foreground model.
-These simulations typically yielded a median absolute deviation of $\approx 0.29$ for $\log_{10}\rho$ indicating fractional differences of $\sim$~95\%.
+These simulations typically yielded a median absolute deviation of $\approx 0.29$ for $\log_{10}\rho$ indicating fractional differences of $\sim$~95\% {\bf between different realizations}.
 This demonstrates that a fractional deviation of $\sim$~90\% observed between data and simulations is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account.
-% Currently, we are primarily limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.
 
@@
 A number of factors are responsible for the characteristics noted in the delay spectra obtained from data and through simulations.
-We address these factors below:
-
-\begin{itemize}
-
-\item {\it Sky Model}: Our model of the sky consists of diffuse and compact emission on diverse spatial scales as shown in Figures~\ref{fig:DSM} and \ref{fig:CSM}, respectively.
-It also consists of localized regions of strong emission such as the Galactic plane and Galactic center.
-In fact, the patch of sky for MWA observations is chosen from regions of low foreground emission.
-We will investigate the signatures of each component and their potential impact on {\it EoR window} contamination.
-
-\item {\it Baseline Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our antenna spacings by their orientation, $\theta_\textrm{b}$.
-We use the {\bf four following bins sweeping 45\arcdeg each}: $-67.5\arcdeg\le \theta_\textrm{b} < -22.5\arcdeg$, $-22.5\arcdeg\le \theta_\textrm{b} < 22.5\arcdeg$, $22.5\arcdeg\le \theta_\textrm{b} < 67.5\arcdeg$, and $67.5\arcdeg\le \theta_\textrm{b} < 112.5\arcdeg$.
-The bin centers are oriented towards South--East, East, North--East, and North respectively.
-{\bf While the bin boundaries are arbitrary, they were so chosen to allow bin centers to be aligned along the cardinal and ordinal directions in a half--plane}.
-Since delays depend on $\theta_\textrm{b}$, binning by $\theta_\textrm{b}$ allows us to match {\bf features of the delay power spectra} to different sky directions.
-
-\item {\it Tile Pointing and Power Pattern}: Since the MWA tiles are steered electronically, the tile power pattern tile changes in any observing mode that tracks the {\bf pointing center}.
-In our study, we take into account the effect of the changing tile power pattern on the {\bf delay power spectra}.
-
-\end{itemize}
-
-In subsequent sections, we provide a detailed explanation of our results as a combination of factors noted above.
+In subsequent sections, we provide a detailed explanation of our results as a combination of these factors.
 Note that numerous features may overlap at different degrees of significance depending on combinations of parameters.
@@
 The observability of the {\it pitchfork} signature predicted in this paper depends on the relative levels of uncertainty in the foreground model and fluctuations from thermal noise.
-In our simulations, since thermal noise in these very short duration snapshots is $\sim 10^4$~K$^2$(Mpc/$h$)$^3$ and features near the horizon delay limits are also of comparable amplitudes, the {\it pitchfork} feature is not expected to be detected in a noisy scenario, although this feature is marginally visible in the {\it zenith} pointing of observed data (see Figure~\ref{fig:fhd-sim-comparison}).
+In our simulations, since thermal noise in these very short duration snapshots is $\sim 10^4$~K$^2$(Mpc/$h$)$^3$ and features near the horizon delay limits are also of comparable amplitudes, the {\it pitchfork} feature is not expected to be detected, though this feature is marginally visible in the {\it zenith} pointing of observed data (see Figure~\ref{fig:fhd-sim-comparison}).
 We attribute this to differences between our foreground model and the actual sky.
@@
 
-In summary, the brightest feature $\sim 10^{12}$~K$^2$(Mpc/$h$)$^3$ is that of the Galactic center in the {\it off--zenith} pointing co--located at a westward secondary lobe with a significantly high gain.
-The next brightest signature $\sim 10^8$~K$^2$(Mpc/$h$)$^3$ is caused by compact emission {\bf from the primary field of view} appearing to be concentrated in the inner regions of the {\it foreground wedge} (around $\tau=0$).
-Diffuse emission from this region is $\sim 10^3$ fainter relative to that from compact emission for a $\sim 84$~m antenna spacing.
-But unlike the latter, diffuse emission decreases rapidly by over four orders of magnitude as the antenna spacing is widened to $\sim 171$~m.
-However, diffuse emission near the horizon is significantly boosted by wide--field effects compared to that in the primary field of view and is prominent across all antenna spacings.
-This results in a {\it three--pronged pitchfork}--shaped signature inside the {\it foreground wedge} (see also Figure~\ref{fig:antennas-delay-spectra}).
-
-\section{Recipe: Baseline--Based Foreground Mitigation}\label{sec:fg-grading}
+\section{Baseline--Based Foreground Mitigation}\label{sec:fg-grading}
 
@@
 Figure~\ref{fig:delay-map} shows the sky mapped to delays registered by the baseline vectors, of length 100~m for instance, oriented northward (top panel) and eastward (bottom panel).
-Since the Galactic center is located in the western sky, this figure demonstrates that it is observed at $\tau<0$ on a baseline oriented eastward and at $\tau=0$ on a baseline oriented northward.
 Figure~\ref{fig:baseline-breakup} shows the delay spectra on baselines oriented northward ($67.5\arcdeg\le \theta_\textrm{b} < 112.5\arcdeg$) at the top and eastward ($-22.5\arcdeg~\le \theta_\textrm{b} < 22.5\arcdeg$) at the bottom.
@@
 
-% The {\bf delay power spectrum} not only carries information on spatial scales of foreground emission but also offers the unique advantage of viewing the sky through a combination of antenna spacing vectors and the delay axis.
-With a foreground model known {\it a priori} in which structures and locations of very bright foreground objects such as the Galactic center or AGN are available, the power of the delay spectrum technique lets us predict the response across antenna spacings as a function of observing parameters such as LST, power pattern, etc.
+With a foreground model known {\it a priori} in which structures and locations of very bright foreground objects such as the Galactic center or AGN are available, we can predict the response across antenna spacings as a function of observing parameters such as LST, power pattern, etc.
 This allows us to programmatically screen data for antenna spacings that are severely contaminated by foregrounds near the horizon delay limits.
@@
 In our example, we discard eastward antenna spacings ($-22.5\arcdeg\le\theta_\textrm{b}< 22.5\arcdeg$) of lengths $|\boldsymbol{b}| < 30$~m.
-{\bf This was found to be the limit beyond which foreground contamination is insensitive to any further removal of wider antenna spacings}.
+{\bf Foreground contamination was found to be insensitive to removal of wider antenna spacings as discussed below}.
 Figure~\ref{fig:before-after} shows the delay spectra obtained with all antenna spacings (top panel) and after applying our screening technique (bottom panel) on the {\it off--zenith} observation.
-Notice the remarkable reduction in foreground spillover into the {\it EoR window} via the removal of bright vertical stripes on short eastward antenna spacings.
-% The color scale range has been adjusted relative to the previous figures to show this significant reduction in foreground contamination.
+Notice the significant reduction in foreground spillover into the {\it EoR window} via the removal of bright vertical stripes on short eastward antenna spacings.
 
@@
 {\bf These results use the MWA as an example.
-In general, the array layout will be a significant factor in determining such thresholds}.
+In general, the array layout will play a significant role in determining such thresholds}.
 
@@
 This technique provides a very simple and yet effective tool in adding a layer of control to mitigate effects of foreground contamination in EoR data analysis.
-
-% \section{Role of Antenna Geometry}\label{sec:antenna-shape}
-
-% {\bf We now examine the influence of the antenna power pattern on the delay power spectrum (refer to equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}).
-In the following analysis, we assume an ideal flat passband over the entire 30.72~MHz bandwidth while still applying a {\it Blackman--Harris} window function.
-We also assume there are no random delay fluctuations in the MWA tile beamforming process}.
-
-% \subsection{Uniformly Illuminated Sky}\label{sec:USM}
-
-% {\bf Simulations of the response to a uniformly illuminated sky confirm that the {\it pitchfork} effect is independent of the specific foreground model chosen in our study.
-The brightness distribution on the sky was fixed to be a constant equal to the mean over the whole sky at 185~MHz with no spectral variation.
-This helps isolate the signatures of the antenna power pattern from those of the foregrounds.
-
-%  % In the following discussion, we assume there are no random delay fluctuations in the MWA tile beamforming process.
-The model power pattern so obtained is different from the one we used earlier, but the former still serves as a good reference (Figure~\ref{fig:power_pattern}) for the present discussion.
-Hence, we do not plot the new power pattern model obtained with ideal beamforming.
-
-% Na\"{i}vely, only single element antennas are sensitive to a uniform brightness distribution while interferometers are not.
-However, an examination of equation~\ref{eqn:obsvis} shows it resembles a Fourier transform when the sky brightness is expressed in a coordinate system of direction cosines \citep{tay99,tho01}.
-In such a framework, the solid angle subtended by a unit interval in direction cosine increases as the secant of the angle away from zenith, which is significant in wide--field measurements.
-A uniformly illuminated sky does not appear so in this frame.
-Therefore, a uniformly illuminated sky will excite wide--field interferometers.
-
-% Figure~\ref{fig:noiseless-usm-delay-spectrum} shows the delay power spectra of such a sky model with the model power pattern of an ideal MWA tile.
-Although not identical, these delay power spectra closely resemble those obtained with the diffuse sky model (see Figure~\ref{fig:noiseless-dsm-delay-spectrum}).
-This indicates that the delay power spectra seen from diffuse foregrounds is primarily explained by those of the underlying power pattern, with any additional variations arising out of anisotropic deviations of the diffuse foreground model from a uniformly illuminated sky}.
-
-% \begin{figure}[htb]
-% \centering
-% \includegraphics[width=\linewidth]{fig15.eps}
-% \caption{Similar to Figure~\ref{fig:noiseless-dsm-delay-spectrum} but for a uniform sky brightness model.
-The features bear a close resemblance to those obtained with the diffuse emission model.
-The edge--heavy {\it two--pronged} features arise from foreshortening of baselines which increases their sensitivity to emission from larger size scales from near the horizon.
-The symmetries around $\tau=0$ or lack thereof correspond to symmetries in the power patterns of the respective pointings (see Figure~\ref{fig:power_pattern}).
-\label{fig:noiseless-usm-delay-spectrum}}
-% \end{figure}
-
-% {\bf The edge--heavy {\it two--pronged} features are evident, which are very similar to those seen earlier in the case of diffuse emission.
-As reasoned in \S\ref{sec:diffuse-long-baselines}, the foreshortening of long baselines causes them to be sensitive to emission from near the horizon.
-
-% Branches of foreground power seen at $\tau>0$ and close to the negative horizon delay limit in the {\it off--zenith} pointing correspond to radiation received from the direction of the primary lobe and the far eastward sidelobe, respectively, in the power pattern.
-On the other hand, the symmetry of the delay power spectrum around $\tau=0$ seen in the {\it zenith} pointing is as expected from a symmetric power pattern}.
-
-% \subsection{Experiments with Aperture Shapes}\label{sec:apertures}
-
-% {\bf The power pattern of an antenna is determined by the shape of its aperture.
-Through the use of a few examples, we examine the role the geometrical shape of the aperture plays in shaping the characteristics of delay power spectrum.
-We consider the following antenna elements placed at the MWA tile locations:
-% \begin{itemize}
-% \item {\it Dipole}: An East--West dipole of length 0.74~m at a height 0.3~m above a ground plane. $A_\textrm{e}=(\lambda/2)^2$.
-% \item {\it Phased Array}: A 4$\times$4 array of isotropic radiators at a height 0.3~m above the ground plane placed in an arrangement similar to that in an MWA tile.
-$A_\textrm{e}=16\,(\lambda/2)^2$.
-% \item {\it Dish}: Diameter of 14~m similar to that proposed for HERA, with $A_\textrm{e}\approx 154$~m$^2$.
-The power pattern is simulated using an {\it Airy} pattern where its sensitivity beyond the horizon is forced to zero.
-% \end{itemize}
-
-% The power patterns of these antenna geometries at 185~MHz for the {\it zenith} pointing are shown in Figure~\ref{fig:power-patterns}.
-It must be noted that the presence of a ground plane below the dipole and the phased array introduces a smoother suppression of sensitivity near the horizon, whereas this transition is abrupt in the case of the dish}.
-
-% % \begin{figure*}[htb]
-% % \centering
-% % \subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{fig16a.eps}} \\
-% % \subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{fig16b.eps}}
-% % \caption{Power patterns (top panels) and simulated delay power spectra (bottom panels) for different antenna shapes at 185~MHz centered on zenith.
-Antenna shapes used are: dipole (left), phased array (middle), and dish (right).
-Refer to text for details.
-The strength and occupancy of the power patterns are correlated with those of delay power spectra.
-The {\it foreground wedge} and the {\it EoR window} are most severely contaminated in the case of the dipole while it is the least for the dish.
-The phased array has intermediate levels of contamination.
-Foreground power close to the horizon delay limits in all three cases are significant even on long baselines.
-The foreshortening of baselines towards the horizon makes them sensitive to foreground emission on large size scales.
-The amplitude of this feature is highest for the dipole and least for the dish and hence, strongly depends on the shape of the antenna element.}
-% % \label{fig:aperture-shapes}
-% % \end{figure*}
-
-% {\bf The delay power spectra for these antenna shapes are shown in Figure~\ref{fig:antennas-delay-spectra}.
-The occupancy of the power patterns in the sky hemisphere is clearly correlated with that in the delay spectra.
-The following visual correlations are also apparent --- the strength of the primary lobe centered on the pointing center is correlated with the delay power spectrum centered on $\tau=0$; and, the overall rate of decrease in the power sensitivity away from the pointing center is correlated with the rate of drop in power away from $\tau=0$.
-
-% The levels of foreground contamination in the {\it EoR window} varies substantially across the different antenna shapes: $\sim 10^4\,$K$^2\,$(Mpc/$h$)$^3$, $\lesssim 10^2\,$K$^2\,$(Mpc/$h$)$^3$, and $<1\,$K$^2\,$(Mpc/$h$)$^3$ while using a dipole, phased array, and a dish, respectively.
-The severity of foreground contamination inside the {\it foreground wedge} both in strength and occupancy also evidently decreases as the antenna element is changed from a dipole to a phased array to a dish.
-For instance, notice that the foreground contamination in $k$--modes between the $k_\parallel=0$ and the horizon limits decreases from $\sim 10^5$~K$^2\,$(Mpc/$h$)$^3$ in a phased array to $\sim 10$~K$^2\,$(Mpc/$h$)$^3$ in a dish.
-As a consequence, $k$--modes in the {\it foreground wedge} that may be deemed too contaminated for EoR studies in the case of a dipole or a phased array can potentially become accessible when using a dish.
-
-% Finally, a distinct feature common to all these aperture shapes is that the foreground contamination near the horizon delay limits even on wide antenna spacings is significant ($\gtrsim 10^5$~K$^2$(Mpc/$h$)$^3$).
-This is contrary to expectation for two reasons --- the power pattern of the antenna near the horizon is negligible compared to the primary lobe region; and, a significant fraction of foreground emission near the horizon delay limits on wide antenna spacings arises from diffuse emission, to which usually only short antenna spacings are sensitive.
-We have argued that baselines are foreshortened towards the horizon and thus become sensitive to large size scales, thereby leading to significant integrated emission in delay bins close to the horizon limits.
-The prevalence of this feature across different antenna shapes demonstrates it is generic to all wide--field measurements, especially in EoR experiments.
-The amplitude of this effect, however, can be controlled via choice of antenna shape and through weighting of aperture illumination}.
 
@@
 This will be distinctly visible when the thermal noise floor is sufficiently lowered, as longer observations are processed.}
-
-% {\bf The role of antenna shape in determining the {\bf delay power spectrum} was also studied.
-In contrast to a dipole and a phased array such as an MWA tile, a dish such as the one proposed for HERA is found to yield the least foreground contamination both in the {\it foreground wedge} and the {\it EoR window} and thus, preferable for EoR studies}.
 
@@
 
-\par\bigskip
+% \par\bigskip
 \bibliographystyle{apj}

### 2015-02-08 ms.tex
Move delay conventions, spatial power spectrum to appendix
from 79fcddd:
============

In ms.tex,

(*) Create appendix to describe delay transform conventions
(*) Create appendix to describe similarities and differences between
delay spectra and true spatial power spectra
(*) Adjust Delay transform section accordingly

Rename figures under Git

@@
 %% preprint2 produces a double-column, single-spaced document:
-\documentclass[preprint2,iop,numberedappendix]{emulateapj}
+\documentclass[preprint2,iop,numberedappendix,twocolappendix,appendixfloats]{emulateapj}
 %% \documentclass[preprint2,iop]{aastex}
@@
 \end{equation}
-$V_{b\tau}$ is expressed in observer's units as Jy~Hz.
+The delay transform conventions used in this paper are described in appendix~\ref{sec:delay-conventions}. $V_{b\tau}$ is expressed in observer's units as Jy~Hz.
 
-\subsection{Resemblance to Spatial Power Spectrum}\label{sec:power-spectrum}
+The delay spectrum has a close resemblance to cosmological H{\sc i} spatial power spectrum.
+Appendix~\ref{sec:EoR-power-spectrum} gives an overview of the similarities and differences between the two.
+Foregrounds can be described in either framework.
+For our study, we find the delay spectrum approach to be simple and yet extremely useful.
 
-Equation~\ref{eqn:obsvis} can be {\bf equivalently expressed} as:
-\begin{align}\label{eqn:vis}
-  V_{uf} &\equiv V_{uf}(\boldsymbol{u},f) \nonumber\\
-  &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_f\,e^{-i2\pi\boldsymbol{u}\cdot\hat{\boldsymbol{s}}}\,\dif\Omega,
-\end{align}
-where, $\hat{\boldsymbol{s}}$ is measured with reference to a location on the sky referred to as the {\it phase center}, and $\boldsymbol{u}\equiv (u,v,w)$ denotes the spatial frequency vector.
-$w$ is aligned parallel to the direction of the phase center, while $u$ and $v$ lie on the transverse plane perpendicular to it.
-For measurements that are constrained to lie on this plane, we can choose $w=0$ without loss of generality and $\boldsymbol{u}$ effectively reduces to $\boldsymbol{u}\equiv (u,v)$, a two--dimensional vector.
-Then, $\boldsymbol{u}$ is directly related to the transverse spatial wavenumber mode as:
-\begin{align}\label{eqn:k-perp}
-  \boldsymbol{k}_\perp &\equiv \frac{2\pi\boldsymbol{u}}{D}, 
-\end{align}
-where, $D\equiv D(z)$ is the transverse comoving distance at redshift $z$.
-
-Since we are concerned with a redshifted H{\sc i} spectral line from cosmological distances, $f$ is a measure of cosmological distance along the line of sight.
-$\eta$, which is the Fourier transform dual of $f$, is used to denote the spatial frequency along the line of sight and has units of time.
-It is directly related to the line--of--sight wavenumber,
-\begin{align}\label{eqn:k-prll}
-  k_\parallel &\approx \frac{2\pi\eta\,f_{21}H_0\,E(z)}{c(1+z)^2}, 
-\end{align}
-where, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
-This approximation holds under the assumption that the redshift range (or frequency band) is small enough within which cosmological evolution is negligible.
-Thus,
-\begin{align}\label{eqn:los-transform}
-  V_{u\eta} &\equiv V_{u\eta}(\boldsymbol{u},\eta) \nonumber\\
-  &\equiv \int V_{uf}(\boldsymbol{u},f)\,W'_f\,e^{i2\pi f\eta}\,\dif f
-\end{align}
-represents the true spatial Fourier representation of the three--dimensional sky brightness distribution.
-This approach has been discussed in detail in \citet{mor04}.
-The spatial power spectrum of EoR H{\sc i} distribution, $P(\boldsymbol{k}_\perp,k_\parallel)$, and $V_{u\eta}$ are related by \citep{mor04,mcq06,par12a}:
-\begin{align}\label{eqn:true-power_spectrum}
-  P(\boldsymbol{k}_\perp,k_\parallel) &\simeq |V_{u\eta}|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
-\end{align}
-where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, and $k_\textrm{B}$ is the Boltzmann constant.
-Thus, $V_{u\eta}$ inferred from observations, also in units of Jy~Hz, can be converted into an equivalent cosmological H{\sc i} power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$, in units of K$^2$~Mpc$^3$ or, more generally, K$^2$(Mpc/$h$)$^3$, where $h$ is the Hubble constant factor.
-
-Without loss of generality, the phase center can be assumed to be the zenith relative to the local tangent plane.
-Then $\boldsymbol{u}$ lies on this plane for measurements constrained to be on it.
-If the array of antennas are also coplanar lying on the local tangent plane, then $\boldsymbol{u}=\boldsymbol{b}/\lambda$.
-Under such circumstances, equations~\ref{eqn:obsvis} and \ref{eqn:delay} closely resemble equations~\ref{eqn:vis} and \ref{eqn:los-transform} respectively.
-However, they are not quite identical to each other.
-It is because $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not.
-\citet{par12b} and \citet{liu14a} have discussed the mathematical correspondence between the two.
-
-Foregrounds can be described in both the delay spectrum and Fourier frameworks.
-For our study, we find the former to be simple and yet extremely useful.
-In order to express a quantity derived from $V_{u\eta}$ that has units identical to that of the cosmological H{\sc i} power spectrum, we define the delay power spectrum:
-\begin{align}\label{eqn:fake-power-spectrum}
+In order to express a quantity derived from $V_{b\tau}$ that has units K$^2$(Mpc/$h$)$^3$ which is identical to that of the cosmological H{\sc i} power spectrum, we define the delay power spectrum:
+\begin{align}\label{eqn:delay-power-spectrum}
   P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |V_{b\tau}|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
 \end{align}
-where, $\boldsymbol{u}$ and $\eta$ in equations~\ref{eqn:k-perp} and \ref{eqn:k-prll} have been correspondingly replaced with $\boldsymbol{b}/\lambda$ and $\tau$ respectively.
-
-Note that since {\it visibility} is a complex quantity, its {\bf delay} transform {\bf contains} positive and negative delays.
-Negative and positive delays correspond to the two hemispheres of the sky as transected by the antenna spacing vector.
-{\bf Figure~\ref{fig:delay-cartoon} illustrates the radio interferometer delays and conventions used in our paper}.
-$\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward (zenith) directions at the telescope site.
-Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at positive and negative delays, respectively.
-Similarly, an object in the northern sky will appear at a positive delay at an antenna spacing oriented northward.
-For all observations used in this study, we use the zenith as the phase center, {\bf for which $\tau\equiv 0$}.
-
-\begin{figure}[htb]
-\centering
-% \includegraphics[width=\linewidth]{delay_cartoon.eps}
-% \includegraphics[width=\linewidth]{fig1.eps}
-\includegraphics[width=\linewidth]{figure1.eps}
-\caption{{\bf Illustration of radio interferometer delays and conventions used in this paper.
-Two antennas labeled as A$_1$ and A$_2$ are separated by vector $\boldsymbol{b}$ on the local tangent plane.
-The local meridian, local north and local east are shown for reference.
-Points labeled as `z', `n', `s', `e' and `w' on the celestial sphere denote zenith, northward, southward, eastward and westward positions, respectively.
-$\tau_\textrm{z}$, $\tau_\textrm{n}$, $\tau_\textrm{s}$, $\tau_\textrm{e}$ and $\tau_\textrm{w}$ denote the respective delays measured between A$_1$ and A$_2$.
-Throughout this paper, zenith is chosen as the phase center.
-Hence, $\tau_\textrm{z}\equiv 0$.
-If $\boldsymbol{b}$ is oriented eastward as shown, then $\tau_\textrm{e} > 0$, $\tau_\textrm{w} < 0$, and $\tau_\textrm{n}=\tau_\textrm{s}=0$.
-Conversely, if $\boldsymbol{b}$ is oriented northward (not shown here), then $\tau_\textrm{n} > 0$, $\tau_\textrm{s} < 0$, and $\tau_\textrm{e}=\tau_\textrm{w}=0$}.
-\label{fig:delay-cartoon}}
-\end{figure}
+with
+\begin{align}
+  \boldsymbol{k}_\perp &\equiv \frac{2\pi(\frac{\boldsymbol{b}}{\lambda})}{D}, \\
+  k_\parallel &\equiv \frac{2\pi\tau\,f_{21}H_0\,E(z)}{c(1+z)^2}, 
+\end{align}
+where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
 
@@
 $|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction {\bf (tangent plane to the celestial sphere)}, form the $x$--axis.
-$\tau\approx\eta\propto k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis.
+$\tau$ and $k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis.
 Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by the antenna spacings and the light travel times across them.
 These boundaries, called horizon delay limits \citep{ved12,par12b}, are shown by solid lines.
-The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\tau$--axis \citep{par12b,thy13}.
-The width of this narrow strip is inversely proportional to the operating bandwidth.
-The region of Fourier space excluding the {\it foreground wedge} and the narrow strips is the so--called {\it EoR window}, shown in light and medium shades.
-{\bf In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\boldsymbol{k}|$.
-Hence, the brightest EoR signal will be observed on the shortest baselines and smallest delays.
-Thus the region of interest for EoR studies relying on {\it avoidance} strategy is just beyond the horizon delay limits (dashed lines) on short baselines, marked as the region of {\it maximal EoR sensitivity}}.
-
-{\bf In the specific case of the MWA, which has a passband constructed using coarse channels, there are period grating responses resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.
-Thus with reference to the MWA, the {\it EoR window} lies outside the dashed lines but inside the first grating response ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) and is shown in medium shade}.
 
@@
 % \includegraphics[width=\linewidth]{fig2.eps}
-\includegraphics[width=\linewidth]{figure2.eps}
+% \includegraphics[width=\linewidth]{figure2.eps}
+\includegraphics[width=\linewidth]{figure1.eps}
 \caption{Fourier space in which delay (and power) spectra of EoR H{\sc i} signals are presented.
@@
 \end{figure}
+
+The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\tau$--axis \citep{par12b,thy13}.
+The width of this narrow strip is inversely proportional to the operating bandwidth.
+The region of Fourier space excluding the {\it foreground wedge} and the narrow strips is the so--called {\it EoR window}, shown in light and medium shades.
+{\bf In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\boldsymbol{k}|$.
+Hence, the brightest EoR signal will be observed on the shortest baselines and smallest delays.
+Thus the region of interest for EoR studies relying on {\it avoidance} strategy is just beyond the horizon delay limits (dashed lines) on short baselines, marked as the region of {\it maximal EoR sensitivity}}.
+
+{\bf In the specific case of the MWA, which has a passband constructed using coarse channels, there are period grating responses resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.
+Thus with reference to the MWA, the {\it EoR window} lies outside the dashed lines but inside the first grating response ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) and is shown in medium shade}.
 
@@
 \centering
-\includegraphics[width=\linewidth]{figure3.eps}
+% \includegraphics[width=\linewidth]{figure3.eps}
+\includegraphics[width=\linewidth]{figure2.eps}
 \caption{Wide--field effects on delay power spectra produced with a uniform sky brightness distribution measured by antenna pairs with a uniform power pattern across the visible hemisphere.
@@
 \centering
-\includegraphics[width=\linewidth]{figure4.eps}
+% \includegraphics[width=\linewidth]{figure4.eps}
+\includegraphics[width=\linewidth]{figure3.eps}
 \caption{Foreground model at 185~MHz consisting of diffuse emission (top) and bright point sources (bottom) visible during a snapshot at 0.09~hours LST.
@@
 \centering
-\subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{figure5a.eps}} \\
-\subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{figure5b.eps}}
+% \subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{figure5a.eps}} \\
+% \subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{figure5b.eps}}
+\subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{figure4a.eps}} \\
+\subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{figure4b.eps}}
 \caption{Power patterns (top panels) and simulated delay power spectra (bottom panels) for different antenna shapes at 185~MHz centered on zenith.
@@
 \centering
-\subfloat[][Diffuse Sky Model]{\label{fig:DSM}\includegraphics[width=0.45\linewidth]{figure6a.eps}}
-\subfloat[][Bright Point Source Model]{\label{fig:CSM}\includegraphics[width=0.45\linewidth]{figure6b.eps}}
+% \subfloat[][Diffuse Sky Model]{\label{fig:DSM}\includegraphics[width=0.45\linewidth]{figure6a.eps}}
+% \subfloat[][Bright Point Source Model]{\label{fig:CSM}\includegraphics[width=0.45\linewidth]{figure6b.eps}}
+\subfloat[][Diffuse Sky Model]{\label{fig:DSM}\includegraphics[width=0.45\linewidth]{figure5a.eps}}
+\subfloat[][Bright Point Source Model]{\label{fig:CSM}\includegraphics[width=0.45\linewidth]{figure5b.eps}}
 \caption{Sky brightness temperature of the diffuse foreground model (left) and flux densities of bright point sources (right) at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 \centering
-\includegraphics[width=\linewidth]{figure7.eps}
+% \includegraphics[width=\linewidth]{figure7.eps}
+\includegraphics[width=\linewidth]{figure6.eps}
 \caption{Delay power spectra from MWA data (left) and modeling (right) for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
@@
 \centering
-\includegraphics[width=\linewidth]{figure8.eps}
+% \includegraphics[width=\linewidth]{figure8.eps}
+\includegraphics[width=\linewidth]{figure7.eps}
 \caption{Simulated {\bf delay power spectra (in units of K$^2$~(Mpc/$h$)$^3$} for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse (left) and compact (right) foreground models without any thermal noise.
@@
 \centering
-\includegraphics[width=\linewidth]{figure9.eps}
+% \includegraphics[width=\linewidth]{figure9.eps}
+\includegraphics[width=\linewidth]{figure8.eps}
 \caption{Simulated {\bf delay power spectra} for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}).
@@
 % \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{f11c.eps}}
-\subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{fig12a.eps}}
-\subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{fig12b.eps}}
-\subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{fig12c.eps}}
+% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{fig12a.eps}}
+% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{fig12b.eps}}
+% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{fig12c.eps}}
+\subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{figure9a.eps}}
+\subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{figure9b.eps}}
+\subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figure9c.eps}}
 \caption{(a): Sky model showing compact ({\it top}) and diffuse ({\it bottom}) emission (adopted from Figure~\ref{fig:sky-model-MWA}).
@@
 % \includegraphics[width=\linewidth]{f12.eps}
-\includegraphics[width=\linewidth]{fig13.eps}
+% \includegraphics[width=\linewidth]{fig13.eps}
+\includegraphics[width=\linewidth]{figure10.eps}
 \caption{Simulated delay spectra {\bf power} for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
@@
 % \includegraphics[width=\linewidth]{f13.eps}
-\includegraphics[width=\linewidth]{fig14.eps}
+% \includegraphics[width=\linewidth]{fig14.eps}
+\includegraphics[width=\linewidth]{figure11.eps}
 \caption{Drop in foreground contamination in the {\it EoR window}, and loss of data for the {\it off--zenith} pointing as a function of discarded baselines.
@@
 
+\appendix
+
+\section{Delay Transform Conventions}\label{sec:delay-conventions}
+
+Note that since {\it visibility} in equation~\ref{eqn:obsvis} is a complex quantity, its {\bf delay} transform {\bf contains} positive and negative delays.
+Negative and positive delays correspond to the two hemispheres of the sky as transected by the antenna spacing vector.
+{\bf Figure~\ref{fig:delay-cartoon} illustrates the radio interferometer delays and conventions used in our paper}.
+$\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward (zenith) directions at the telescope site.
+Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at positive and negative delays, respectively.
+Similarly, an object in the northern sky will appear at a positive delay at an antenna spacing oriented northward.
+For all observations used in this study, we use the zenith as the phase center, {\bf for which $\tau\equiv 0$}.
+
+\begin{figure}[htb]
+\centering
+% \includegraphics[width=\linewidth]{delay_cartoon.eps}
+% \includegraphics[width=\linewidth]{fig1.eps}
+% \includegraphics[width=\linewidth]{figure1.eps}
+\includegraphics[width=\linewidth]{figureA1.eps}
+\caption{{\bf Illustration of radio interferometer delays and conventions used in this paper.
+Two antennas labeled as A$_1$ and A$_2$ are separated by vector $\boldsymbol{b}$ on the local tangent plane.
+The local meridian, local north and local east are shown for reference.
+Points labeled as `z', `n', `s', `e' and `w' on the celestial sphere denote zenith, northward, southward, eastward and westward positions, respectively.
+$\tau_\textrm{z}$, $\tau_\textrm{n}$, $\tau_\textrm{s}$, $\tau_\textrm{e}$ and $\tau_\textrm{w}$ denote the respective delays measured between A$_1$ and A$_2$.
+Throughout this paper, zenith is chosen as the phase center.
+Hence, $\tau_\textrm{z}\equiv 0$.
+If $\boldsymbol{b}$ is oriented eastward as shown, then $\tau_\textrm{e} > 0$, $\tau_\textrm{w} < 0$, and $\tau_\textrm{n}=\tau_\textrm{s}=0$.
+Conversely, if $\boldsymbol{b}$ is oriented northward (not shown here), then $\tau_\textrm{n} > 0$, $\tau_\textrm{s} < 0$, and $\tau_\textrm{e}=\tau_\textrm{w}=0$}.
+\label{fig:delay-cartoon}}
+\end{figure}
+
+\section{Cosmological H{\sc i} Power Spectrum}\label{sec:EoR-power-spectrum}
+
+Equation~\ref{eqn:obsvis} can be {\bf equivalently expressed} as:
+\begin{align}\label{eqn:vis}
+  V_{uf} &\equiv V_{uf}(\boldsymbol{u},f) \nonumber\\
+  &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_f\,e^{-i2\pi\boldsymbol{u}\cdot\hat{\boldsymbol{s}}}\,\dif\Omega,
+\end{align}
+where, $\hat{\boldsymbol{s}}$ is measured with reference to a location on the sky referred to as the {\it phase center}, and $\boldsymbol{u}\equiv (u,v,w)$ denotes the spatial frequency vector.
+$w$ is aligned parallel to the direction of the phase center, while $u$ and $v$ lie on the transverse plane perpendicular to it.
+For measurements that are constrained to lie on this plane, we can choose $w=0$ without loss of generality and $\boldsymbol{u}$ effectively reduces to $\boldsymbol{u}\equiv (u,v)$, a two--dimensional vector.
+Then, $\boldsymbol{u}$ is directly related to the transverse spatial wavenumber mode as:
+\begin{align}\label{eqn:k-perp}
+  \boldsymbol{k}_\perp &\equiv \frac{2\pi\boldsymbol{u}}{D}, 
+\end{align}
+where, $D\equiv D(z)$ is the transverse comoving distance at redshift $z$.
+
+Since we are concerned with a redshifted H{\sc i} spectral line from cosmological distances, $f$ is a measure of cosmological distance along the line of sight.
+$\eta$, which is the Fourier transform dual of $f$, is used to denote the spatial frequency along the line of sight and has units of time.
+It is directly related to the line--of--sight wavenumber,
+\begin{align}\label{eqn:k-prll}
+  k_\parallel &\approx \frac{2\pi\eta\,f_{21}H_0\,E(z)}{c(1+z)^2}, 
+\end{align}
+where, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
+This approximation holds under the assumption that the redshift range (or frequency band) is small enough within which cosmological evolution is negligible.
+Thus,
+\begin{align}\label{eqn:los-transform}
+  V_{u\eta} &\equiv V_{u\eta}(\boldsymbol{u},\eta) \nonumber\\
+  &\equiv \int V_{uf}(\boldsymbol{u},f)\,W'_f\,e^{i2\pi f\eta}\,\dif f
+\end{align}
+represents the true spatial Fourier representation of the three--dimensional sky brightness distribution.
+This approach has been discussed in detail in \citet{mor04}.
+The spatial power spectrum of EoR H{\sc i} distribution, $P(\boldsymbol{k}_\perp,k_\parallel)$, and $V_{u\eta}$ are related by \citep{mor04,mcq06,par12a}:
+\begin{align}\label{eqn:true-power_spectrum}
+  P(\boldsymbol{k}_\perp,k_\parallel) &\simeq |V_{u\eta}|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
+\end{align}
+where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, $\lambda$ is the wavelength of the band center, and $k_\textrm{B}$ is the Boltzmann constant.
+Thus, $V_{u\eta}$ inferred from observations, in units of Jy~Hz, can be converted into an equivalent cosmological H{\sc i} power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$, in units of K$^2$~Mpc$^3$ or, more generally, K$^2$(Mpc/$h$)$^3$, where $h$ is the Hubble constant factor.
+
+Without loss of generality, the phase center can be assumed to be the zenith relative to the local tangent plane.
+Then $\boldsymbol{u}$ lies on this plane for measurements constrained to be on it.
+If the array of antennas are also coplanar lying on the local tangent plane, then $\boldsymbol{u}=\boldsymbol{b}/\lambda$.
+Under such circumstances, equations~\ref{eqn:obsvis} and \ref{eqn:delay} closely resemble equations~\ref{eqn:vis} and \ref{eqn:los-transform} respectively.
+However, they are not quite identical to each other.
+It is because $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not.
+\citet{par12b} and \citet{liu14a} have discussed the mathematical correspondence between the two.
+
 % \par\bigskip

### 2015-02-09 ms.tex
Create subsection for Fourier delay space
from edcebe0:
============

In ms.tex,

(*) Create new subsection Fourier delay space under Delay Spectrum
section
(*) Reshuffle a paragraph within th Delay Spectrum section
(*) Change wording in Delay Spectrum section slightly
(*) Change wording in subsection Delay Spectrum Deconvolution

@@
 
+In our present study, we use a latitude of -26.701\arcdeg~ and an antenna layout identical to that of the MWA \citep{bea12} for the observatory.
+The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km.
+Here we focus on antenna spacings $|\boldsymbol{b}| \le 200$~m (spatial scales relevant to reionization).
+{\bf Their deviation from coplanarity is negligible}.
+For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67.5\arcdeg \leq \theta_b < 112.5\arcdeg$.
+Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
+{\bf Note that visibilities from different baselines have not been averaged together in this analysis}.
+
+\subsection{Fourier Delay Space}\label{sec:delay-space}
+
 {\bf We give a brief overview of some parameters of Fourier space which is generic to all experiments that use a similar approach}.
@@
 Hence, the brightest EoR signal will be observed on the shortest baselines and smallest delays.
-Thus the region of interest for EoR studies relying on {\it avoidance} strategy is just beyond the horizon delay limits (dashed lines) on short baselines, marked as the region of {\it maximal EoR sensitivity}}.
+Thus the regions of interest for EoR studies relying on {\it avoidance} strategy are just beyond the horizon delay limits (dashed lines) on short baselines, marked as regions of {\it maximal EoR sensitivity}}.
 
@@
 
-In our present study, we use a latitude of -26.701\arcdeg and an antenna layout identical to that of the MWA \citep{bea12}.
-The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km.
-Here we focus on antenna spacings $|\boldsymbol{b}| \le 200$~m (spatial scales relevant to reionization).
-{\bf Their deviation from coplanarity is negligible}.
-For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67.5\arcdeg \leq \theta_b < 112.5\arcdeg$.
-Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
-{\bf Note that visibilities from different baselines have not been averaged together in this analysis}.
-
 \subsection{Delay Spectrum Deconvolution}\label{sec:data-delay-spectrum}
@@
 We obtain the delay spectrum of visibilities by taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f$ to be a {\it Blackman--Harris} window function.
-The sky spectrum is multiplied by weights, $W_f$, due to the instrumental passband and channels possibly flagged for interference.
+The sky spectrum is multiplied by weights, $W_f$, due to the instrumental passband and frequency channels possibly flagged for radio frequency interference.
 In delay--space, these weights translate into a convolution by a point spread function (PSF).

### 2015-02-09 ms.tex
Change title and Introduction slightly
from 3e35322:
============

In ms.tex,

(*) Change title and last paragraph of Introduction section
(*) Change wording in Pitchfork subsection slightly
(*) Change wording slightly in caption to figure on foreground
mitigation by baseline screening

@@
 
-\shorttitle{EoR Foregrounds in Fourier Space}
+\shorttitle{Foregrounds in EoR Power Spectra}
 \shortauthors{Thyagarajan et~al.}
@@
 %% \title{Limits on the detection of the Epoch of Reionization from MWA observations of the redshifted 21~cm line}
-\title{Foregrounds in Fourier Delay Space for Wide--Field Redshifted 21~cm Observatories}
+\title{Foregrounds in Wide--Field Redshifted 21~cm Power Spectra}
 
@@
 
-{\bf We investigate signatures generic to all wide--field measurements of EoR power spectra and explore antenna shapes that are most desirable from the viewpoint of foreground contamination for future EoR experiments}.
+% {\bf We investigate signatures generic to all wide--field measurements of EoR power spectra and explore antenna shapes that are most desirable from the viewpoint of foreground contamination for future EoR experiments}.
 After matching observations from the Murchison Widefield Array with instrumental simulations of the entire sky, we examine the {\bf shape of the delay spectrum} of each baseline.
@@
 
-In \S\ref{sec:delay-spectrum} we {\bf provide a brief overview of the delay spectrum technique} \citep{par12a,par12b}.
-We establish the signatures arising from the nature of wide--field measurements in \S\ref{sec:wide-field}.
-In \S\ref{sec:sim}, we present the foreground model and a variety of instrument models to rank antenna shapes based on the criterion of foreground contamination.
+{\bf In \S\ref{sec:delay-spectrum} we provide a brief overview of the delay spectrum technique \citep{par12a,par12b}.
+We investigate signatures generic to all wide--field measurements of EoR power spectra in \S\ref{sec:wide-field}.
+In \S\ref{sec:sim}, we present the foreground model and a variety of instrument models to rank antenna shapes based on foreground contamination.
 In \S\ref{sec:MWA}, we describe the MWA setup, summarize the observing parameters, and present the resulting data.
 Simulations using these observing parameters are compared with the data and analyzed for foreground signatures.
-In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique.
+We report two important findings: foregrounds that most severely {\bf obscure} the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted.
+In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique that minimizes foreground contamination using prior knowledge of the sky to down--weight adversely contaminated baselines}.
 We present a summary of our work and findings in \S\ref{sec:summary}.
@@
 The peak at $\tau=0$ (corresponding to the primary lobe in the power pattern) with a value of $\sim 10^7$--$10^8$~K$^2$(Mpc/$h$)$^3$, independent of antenna spacing, is predominantly determined by compact emission.
-The corresponding peak at zero--delay from diffuse foreground model is $\sim 10^3$ times fainter and decreases rapidly with increase in antenna spacing.
+The peak at $\tau=0$ from diffuse foreground model is $\sim 10^3$ times fainter and decreases rapidly with increase in antenna spacing.
 This is the response expected from different antenna spacings towards compact and diffuse emission.
@@
 Near the horizon delay limits, the diffuse component is brighter relative to the compact component.
-Here, diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at zero--delay.
+Here, diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at $\tau=0$.
 In fact, even on widely spaced antennas, diffuse emission in the {\bf delay power spectrum} near the horizon delay limits exceeds that in the primary lobe by about three orders of magnitude.
-This feature is described in \S\ref{sec:diffuse-long-baselines}.
+This feature was described in \S\ref{sec:diffuse-long-baselines}.
 We attribute this to wide--field measurement effects discussed in \S\ref{sec:wide-field}.
@@
 \includegraphics[width=\linewidth]{figure11.eps}
-\caption{Drop in foreground contamination in the {\it EoR window}, and loss of data for the {\it off--zenith} pointing as a function of discarded baselines.
+\caption{Drop in foreground contamination in the MWA {\it EoR window}, and loss of data for the {\it off--zenith} pointing as a function of discarded baselines.
 Eastward baselines with varying degrees of directedness -- $|\theta_\textrm{b}|<$~7.5\arcdeg (solid circles), $|\theta_\textrm{b}|<$~15\arcdeg~ (solid squares), and $|\theta_\textrm{b}|<$~22.5\arcdeg (solid stars) -- and lengths $|\boldsymbol{b}| \le |\boldsymbol{b}|_\textrm{max}$ ($x$--axis) are discarded.

### 2015-02-10 ms.tex
Reorganize Simulations section
from 89383e7:
============

In ms.tex,

(*) Reorganize Simulations section (S4)

@@
 
-In our present study, we use a latitude of -26.701\arcdeg~ and an antenna layout identical to that of the MWA \citep{bea12} for the observatory.
-The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km.
-Here we focus on antenna spacings $|\boldsymbol{b}| \le 200$~m (spatial scales relevant to reionization).
-{\bf Their deviation from coplanarity is negligible}.
-For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67.5\arcdeg \leq \theta_b < 112.5\arcdeg$.
-Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
-{\bf Note that visibilities from different baselines have not been averaged together in this analysis}.
+Note that visibilities from different baselines have not been averaged together in this analysis.
 
@@
 
-Figure~\ref{fig:usm} shows the wide--field delay power spectrum response of a uniformly illuminated sky with no spectral variation measured by antenna elements with a uniform power pattern across the sky.
+Figure~\ref{fig:usm} shows the wide--field delay power spectrum response of a uniformly illuminated sky with no spectral variation measured by antenna elements arranged identical to that in the MWA antenna array layout \citep{bea12} with a uniform power pattern across the sky and a bandwidth of 30.72~MHz centered around 185~MHz (refer to \S\ref{sec:instrument_model} for a detailed description of the instrument model).
 Notice the features arising due to the direction--dependent solid angle term and the steep rise in power towards the horizon limits.
@@
 
+We describe the instrument and foreground models used in our simulations.
+
+\subsection{Instrument Model}\label{sec:instrument_model}
+
+{\bf In our present study, we use a latitude of -26.701\arcdeg~ and an antenna layout identical to that of the MWA \citep{bea12} for the observatory.
+The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km.
+Here we focus on antenna spacings $|\boldsymbol{b}| \le 200$~m (spatial scales relevant to reionization).
+Their deviation from coplanarity is negligible.
+For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67.5\arcdeg \leq \theta_b < 112.5\arcdeg$.
+Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
 We choose an observing frequency of 185~MHz ($z\simeq 6.68$) and a flat passband of width $\Delta B = 30.72$~MHz to roughly match those of ongoing MWA EoR observations.
-Now we describe the foreground and instrumental models used in our simulations.
+See \citet{tin13} for a detailed description of the MWA instrument}.
+
+One of the principal components of the instrument model is the antenna power pattern, $A(\hat{\boldsymbol{s}},f)$ (see equation~\ref{eqn:obsvis}).
+It is determined by the shape of its aperture.
+Using a few examples, we examine the role the geometrical shape of the aperture plays in shaping the characteristics of delay power spectrum.
+We consider the following antenna elements placed at the MWA tile locations:
+\begin{itemize}
+\item {\it Dipole}: An East--West dipole of length 0.74~m at a height 0.3~m above a ground plane. $A_\textrm{e}=(\lambda/2)^2$.
+\item {\it Phased Array}: A 4$\times$4 array of isotropic radiators with a grid spacing of 1.1~m at a height 0.3~m above the ground plane placed in an arrangement similar to that of an MWA tile.
+$A_\textrm{e}=16\,(\lambda/2)^2$.
+\item {\it Dish}: Diameter of 14~m similar to that proposed for HERA, with $A_\textrm{e}\approx 154$~m$^2$.
+The power pattern is simulated using an {\it Airy} pattern where its sensitivity beyond the horizon is forced to zero.
+\end{itemize}
+
+% The presence of a ground plane below the dipole and the phased array introduces a smoother suppression of sensitivity near the horizon, whereas this transition is abrupt in the case of the dish.
 
@@
 
-\subsection{Instrument Model}\label{sec:instrument_model}
+\subsection{Role of Antenna Geometry}\label{sec:antenna-shape}
 
-The instrument model consists primarily of the antenna power pattern, $A(\hat{\boldsymbol{s}},f)$ (see equation~\ref{eqn:obsvis}).
-It is determined by the shape of its aperture.
-Through the use of a few examples, we examine the role the geometrical shape of the aperture plays in shaping the characteristics of delay power spectrum.
-We consider the following antenna elements placed at the MWA tile locations:
-\begin{itemize}
-\item {\it Dipole}: An East--West dipole of length 0.74~m at a height 0.3~m above a ground plane. $A_\textrm{e}=(\lambda/2)^2$.
-\item {\it Phased Array}: A 4$\times$4 array of isotropic radiators with a grid spacing of 1.1~m at a height 0.3~m above the ground plane placed in an arrangement similar to that of an MWA tile.
-$A_\textrm{e}=16\,(\lambda/2)^2$.
-\item {\it Dish}: Diameter of 14~m similar to that proposed for HERA, with $A_\textrm{e}\approx 154$~m$^2$.
-The power pattern is simulated using an {\it Airy} pattern where its sensitivity beyond the horizon is forced to zero.
-\end{itemize}
-
-The power patterns of these antenna geometries at 185~MHz for the {\it zenith} pointing are shown in Figure~\ref{fig:power-patterns}.
-The presence of a ground plane below the dipole and the phased array introduces a smoother suppression of sensitivity near the horizon, whereas this transition is abrupt in the case of the dish.
+The power patterns of the aforementioned antenna geometries at 185~MHz for the {\it zenith} pointing are shown in Figure~\ref{fig:power-patterns}.
 

### 2015-02-10 ms.tex
Add summary paragraph to Delay Spectrum section
from fb0bec1:
============

In ms.tex,

(*) Add summary paragraph to Delay Spectrum section
(*) Change wording in Wide-field measurements section slightly
(*) Remove citation of Tingay et al.(2013) in subsection Instrument
Model
(*) Rename subsection Fourier Delay Space to Delay Space
(*) Change caption to figure on wide-field effects slightly

@@
 
-Note that visibilities from different baselines have not been averaged together in this analysis.
+In summary, the delay spectrum, $V_{b\tau}$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}.
+$V_{b\tau}$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument.
+At the same time, it is closely related to the sought power spectrum containing critical information about spatial scales.
+In using the delay spectrum technique, visibilities from different baselines have not been averaged together.
 
-\subsection{Fourier Delay Space}\label{sec:delay-space}
+\subsection{Delay Space}\label{sec:delay-space}
 
@@
 
-Figure~\ref{fig:usm} shows the wide--field delay power spectrum response of a uniformly illuminated sky with no spectral variation measured by antenna elements arranged identical to that in the MWA antenna array layout \citep{bea12} with a uniform power pattern across the sky and a bandwidth of 30.72~MHz centered around 185~MHz (refer to \S\ref{sec:instrument_model} for a detailed description of the instrument model).
-Notice the features arising due to the direction--dependent solid angle term and the steep rise in power towards the horizon limits.
-Wide--field effects are prevalent on all antenna spacings, including the longest ones used in this study.
+Figure~\ref{fig:usm} shows the wide--field delay power spectrum of a uniformly illuminated sky with no spectral variation measured by antenna elements arranged identical to that in the MWA antenna array layout \citep{bea12} with a uniform power pattern across the sky and a bandwidth of 30.72~MHz centered around 185~MHz (refer to \S\ref{sec:instrument_model} for a detailed description of the instrument model).
+Notice the steep rise in power towards the horizon limits.
+These wide--field effects are prevalent on all antenna spacings, including the longest ones used in this study.
 
@@
 \caption{Wide--field effects on delay power spectra produced with a uniform sky brightness distribution measured by antenna pairs with a uniform power pattern across the visible hemisphere.
+Delay power spectra are obtained using equations~\ref{eqn:obsvis}, \ref{eqn:delay-transform} and \ref{eqn:delay-power-spectrum} for each baseline, which are then stacked by baseline length.
+The axes correspond to cosmological dimensions.
 The non--zero response of the interferometer array to a uniform brightness distribution and the prominent {\it edge brightening} close to the horizon delay limits are wide--field effects.
@@
 Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
-We choose an observing frequency of 185~MHz ($z\simeq 6.68$) and a flat passband of width $\Delta B = 30.72$~MHz to roughly match those of ongoing MWA EoR observations.
-See \citet{tin13} for a detailed description of the MWA instrument}.
+We choose an observing frequency of 185~MHz ($z\simeq 6.68$) and a flat passband of width $\Delta B = 30.72$~MHz to roughly match those of ongoing MWA EoR observations}.
 

### 2015-02-10 ms.tex
Rename and renumber S5 and its subsections
from 45a74e9:
============

In ms.tex,

(*) Chnage title of S5 to Murchison Widefield Array
(*) Rename and renumber subsections under S5
(*) Change wording in caption of figure on wide-field effects slightly
(*) Change wording in subsection Delay Space slightly
(*) Change wording in caption of Delay space figure slightly
(*) Change wording in subsection Instrument model slightly
(*) Change wording in caption of MWA data vs. model comparison figure
slightly
(*) Change wording in caption of simulated CSM and DSM delay spectra
figure slightly
(*) Change

@@
 {\bf We give a brief overview of some parameters of Fourier space which is generic to all experiments that use a similar approach}.
-Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of H{\sc i} from the EoR are presented in this paper.
+Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of redshifted H{\sc i} observations are calculated.
 $|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction {\bf (tangent plane to the celestial sphere)}, form the $x$--axis.
@@
 \includegraphics[width=\linewidth]{figure1.eps}
-\caption{Fourier space in which delay (and power) spectra of EoR H{\sc i} signals are presented.
+\caption{Fourier space in which delay (and power) spectra of EoR H{\sc i} signals are calculated.
 The $x$--axis is denoted by $|\boldsymbol{b}|$ (antenna spacing) or $k_\perp$ (transverse wavenumber).
@@
 Its boundaries (solid lines), given by light travel time for corresponding antenna spacings, are referred to as horizon delay limits.
-Regions excluding the {\it wedge} and narrow extensions of the {\it wedge} (white unshaded strips) caused by convolution with the {\bf instrument's spectral transfer function} are expected to be relatively free of foreground contamination {\bf and are generally referred to as the {\it EoR window}.
+Narrow extensions of the {\it wedge} (white unshaded strips) are caused by convolution with the {\bf instrument's spectral transfer function}.
+Regions excluding the {\it wedge} are expected to be relatively free of foreground contamination {\bf and are generally referred to as the {\it EoR window}.
 There are undesirable grating responses (dot--dashed lines) specific to the MWA.
@@
 We obtain the delay spectrum of visibilities by taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f$ to be a {\it Blackman--Harris} window function.
-The sky spectrum is multiplied by weights, $W_f$, due to the instrumental passband and frequency channels possibly flagged for radio frequency interference.
+The sky spectrum is multiplied in the instrument by weights, $W_f$, due to the instrumental passband and flagging of frequency channels possibly corrupted by radio frequency interference, which together are represented by the weights, $W_f$.
 In delay--space, these weights translate into a convolution by a point spread function (PSF).
@@
 
-Figure~\ref{fig:usm} shows the wide--field delay power spectrum of a uniformly illuminated sky with no spectral variation measured by antenna elements arranged identical to that in the MWA antenna array layout \citep{bea12} with a uniform power pattern across the sky and a bandwidth of 30.72~MHz centered around 185~MHz (refer to \S\ref{sec:instrument_model} for a detailed description of the instrument model).
+Figure~\ref{fig:usm} shows the wide--field delay power spectrum of a uniformly illuminated sky with no spectral variation as measured by antenna elements arranged identical to that in the MWA antenna array layout \citep{bea12} with a uniform power pattern across the sky and a bandwidth of 30.72~MHz centered around 185~MHz (refer to \S\ref{sec:instrument_model} for a detailed description of the instrument model).
 Notice the steep rise in power towards the horizon limits.
@@
 Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
-We choose an observing frequency of 185~MHz ($z\simeq 6.68$) and a flat passband of width $\Delta B = 30.72$~MHz to roughly match those of ongoing MWA EoR observations}.
+We choose an observing frequency of 185~MHz ($z\simeq 6.68$) and a flat passband of width $\Delta B = 30.72$~MHz to roughly match those of ongoing MWA EoR observations (discussed in detail in \S\ref{sec:MWA})}.
 
@@
 Antenna shapes used are: dipole (left), phased array (middle), and dish (right).
-Refer to text for details.
+Refer to \S\ref{sec:instrument_model} for details of the antenna models.
 The strength and occupancy of the power patterns are correlated with those of delay power spectra.
@@
 The foreshortening of baselines towards the horizon makes them sensitive to foreground emission on large size scales.
-The amplitude of this feature is highest for the dipole and least for the dish and hence, strongly depends on the shape of the antenna element.}
+The amplitude of this feature strongly depends on the shape of the antenna element.
+It is highest for a dipole (which has a strong response near the horizon) and least for a dish.}
 \label{fig:aperture-shapes}
@@
 
-\section{MWA Data Modeling}\label{sec:MWA}
+\section{The Murchison Widefield Array}\label{sec:MWA}
 
-We now use our simulations to analyze features in the delay power spectrum obtained using the MWA.
+We now use our simulations to analyze features in the observed delay power spectrum obtained using the MWA.
 
-\subsection{The Murchison Widefield Array (MWA)}\label{sec:instrument}
+% \subsection{The Murchison Widefield Array (MWA)}\label{sec:instrument}
 
 MWA construction was completed in 2012 and, after commissioning, began its EoR observing program in 2013.
-In its final configuration, the MWA is a 128--tile interferometer capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz.
+The MWA is a 128--tile interferometer capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz.
 Each tile is a phased array of 16 dipoles, each in the shape of a bow--tie.
@@
 
-\subsection{EoR Observations}\label{sec:obsparms}
+% \subsection{EoR Observations}\label{sec:obsparms}
 
@@
 The color scales are logarithmic.
-Power pattern contours are overlaid.
+MWA tile power pattern contours are overlaid.
 The contour levels shown are 0.00195, 0.00781, 0.0312, 0.125, and 0.5.
 The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off--zenith} pointing in the diffuse sky model and the MWA tile power gain is significant ($\gtrsim 12$\%) in that direction.
-In contrast, emission from the Galactic plane in {\it zenith} pointing is significantly lesser.}
+In contrast, emission from the Galactic plane in {\it zenith} pointing is significantly less.}
 \label{fig:sky-model-MWA}
@@
 
-\subsection{Comparison with Data}\label{sec:data-vs-model}
+\subsection{Comparison of Data and Model}\label{sec:data-vs-model}
 
@@
 The {\it foreground wedge} is bounded by white dotted lines.
-Model matches the data to a level consistent with the uncertainties in foreground models and the antenna beam.
+Model matches the data to a level consistent with the uncertainties in foreground models and the antenna beam (as discussed in \S\ref{sec:data-vs-model}).
 \label{fig:fhd-sim-comparison}}
@@
 \includegraphics[width=\linewidth]{figure7.eps}
-\caption{Simulated {\bf delay power spectra (in units of K$^2$~(Mpc/$h$)$^3$} for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse (left) and compact (right) foreground models without any thermal noise.
+\caption{Simulated {\bf delay power spectra (in units of K$^2$~(Mpc/$h$)$^3$)} for for the diffuse (left) and compact (right) foreground models in the {\it off--zenith} (top) and {\it zenith} (bottom) pointings without any thermal noise.
 The axes and color scale are identical to those in Figure~\ref{fig:fhd-sim-comparison}.
@@
 Compact emission is centrally concentrated.
-This combination gives rise to a {\it pitchfork} signature in the {\it foreground wedge}.
 \label{fig:noiseless-dsm-csm-delay-spectrum}}
@@
 Vertical dotted lines mark the horizon delay limits.
-{\bf Compact emission dominates the central regions of the delay power spectra while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a {\it pitchfork}--shaped structure.
+{\bf Compact emission dominates the central regions of the delay power spectra while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a characteristic {\it pitchfork}--shaped structure.
 The shaded region denotes the envelope of uncertainty around the composite model delay power spectrum.

### 2015-02-10 ms.tex
Change wording in appendix A
from cf27bde:
============

In ms.tex,

(*) Change appendix A slightly
Change subscript notation of quantities
from 2e67f14:
============

In ms.tex,

(*) Change notation of W_f to W_i(f)
(*) Change notation W'_f to W(f)
(*) Change notation V_{uf} to V_u(f)
(*) Change notation V_{bf} to V_b(f)
(*) Change notation V_{u\eta} to \tilde{V}_u(\eta)
(*) Change wording slightly in data-model comparison subsection
(*) Cite Bowman et al.(2013) for MWA science objectives

@@
 
-Interferometer array data known as {\it visibilities}, $V_{bf}\equiv V_{bf}(\boldsymbol{b},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis.
-If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power pattern, respectively, at different frequencies as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_f\equiv W_f(f)$ denotes instrumental bandpass weights, then $V_{bf}$ can be written as:
+Interferometer array data known as {\it visibilities}, $V_b(f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis.
+If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power pattern, respectively, at different frequencies as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_\textrm{i}(f)$ denotes instrumental bandpass weights, then $V_b(f)$ can be written as:
 \begin{align}\label{eqn:obsvis}
-  V_{bf} &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_f\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif\Omega,
+  V_b(f) &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_\textrm{i}(f)\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif\Omega,
 \end{align}
@@
 
-{\bf The {\it delay spectrum}, $V_{b\tau}\equiv V_{b\tau}(\boldsymbol{b},\tau)$, is defined as} the inverse Fourier transform of $V_{bf}$ along the frequency coordinate:
+{\bf The {\it delay spectrum}, $\tilde{V}_b(\tau)$, is defined as} the inverse Fourier transform of $V_b(f)$ along the frequency coordinate:
 \begin{align}\label{eqn:delay-transform}
-  V_{b\tau} &\equiv \int V_{bf}\,W'_f\,e^{i2\pi f\tau}\,\dif f,
+  \tilde{V}_b(\tau) &\equiv \int V_b(f)\,W(f)\,e^{i2\pi f\tau}\,\dif f,
 \end{align}
-where, $W'_f\equiv W'_f(f)$ is a spectral weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}, and $\tau$ represents the signal delay between antenna pairs:
+where, $W(f)$ is a spectral weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}, and $\tau$ represents the signal delay between antenna pairs:
 \begin{equation}\label{eqn:delay}
@@
 \end{equation}
-The delay transform conventions used in this paper are described in appendix~\ref{sec:delay-conventions}. $V_{b\tau}$ is expressed in observer's units as Jy~Hz.
+The delay transform conventions used in this paper are described in appendix~\ref{sec:delay-conventions}. $\tilde{V}_b(\tau)$ is expressed in observer's units as Jy~Hz.
 
@@
 
-In order to express a quantity derived from $V_{b\tau}$ that has units K$^2$(Mpc/$h$)$^3$ which is identical to that of the cosmological H{\sc i} power spectrum, we define the delay power spectrum:
+In order to express a quantity derived from $\tilde{V}_b(\tau)$ that has units K$^2$(Mpc/$h$)$^3$ which is identical to that of the cosmological H{\sc i} power spectrum, we define the delay power spectrum:
 \begin{align}\label{eqn:delay-power-spectrum}
-  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |V_{b\tau}|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
+  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |\tilde{V}_b(\tau)|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
 \end{align}
@@
 
-In summary, the delay spectrum, $V_{b\tau}$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}.
-$V_{b\tau}$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument.
+In summary, the delay spectrum, $\tilde{V}_b(\tau)$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}.
+$\tilde{V}_b(\tau)$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument.
 At the same time, it is closely related to the sought power spectrum containing critical information about spatial scales.
@@
 
-We obtain the delay spectrum of visibilities by taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W'_f$ to be a {\it Blackman--Harris} window function.
-The sky spectrum is multiplied in the instrument by weights, $W_f$, due to the instrumental passband and flagging of frequency channels possibly corrupted by radio frequency interference, which together are represented by the weights, $W_f$.
+We obtain the delay spectrum of visibilities by taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W(f)$ to be a {\it Blackman--Harris} window function.
+The sky spectrum is multiplied in the instrument by the instrumental passband and flagging of frequency channels possibly corrupted by radio frequency interference, which together are represented by the weights, $W_\textrm{i}(f)$.
 In delay--space, these weights translate into a convolution by a point spread function (PSF).
@@
 \begin{align}\label{eqn:obsvis-lmn}
-  V_{bf} &= \iint\limits_\textrm{sky} \frac{A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)}{\sqrt{1-l^2-m^2}}\,W_f\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif l\,\dif m,
+  V_b(f) &= \iint\limits_\textrm{sky} \frac{A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)}{\sqrt{1-l^2-m^2}}\,W_\textrm{i}(f)\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif l\,\dif m,
 \end{align}
@@
 
-We now use our simulations to analyze features in the observed delay power spectrum obtained using the MWA.
+We now use our simulations to analyze features in observed delay power spectrum obtained using the MWA.
 
@@
 After correcting for the shape of these coarse channels, the fine channels at the edges of these sub--bands are flagged because they are known to be contaminated by aliasing at a low level.
-% Any spectral flags, or other known weights constitutes $W_f$ in equation~\ref{eqn:obsvis}.
+% Any spectral flags, or other known weights constitutes $W_\textrm{i}(f)$ in equation~\ref{eqn:obsvis}.
 
@@
 
-{\bf The MWA is expected to be sensitive to the power spectrum of the H{\sc i} signal over the redshift range $6<z<10$ \citep{bow06,thy13,bea13}.
-Over 600~hours have been currently observed using the MWA with the objective of characterizing the nature of the sources that are responsible for ionizing the IGM, charting the evolution of the global neutral fraction, and probing the nature of quasar emissions by constraining the properties of their ionized proximity zones.}
+{\bf The MWA is expected to be sensitive to the power spectrum of the H{\sc i} signal over the redshift range $6<z<10$ \citep{bow06,bea13,thy13}.
+Over 600~hours have been currently observed using the MWA, targeting the science objectives outlined in \citet{bow13}.}
 
@@
 
-{\bf We estimate thermal noise in the data using the {\it rms} of $V_{b\tau}$ obtained after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
+% {\bf We estimate the system temperature, $T_\textrm{sys}$, using the {\it rms} of $\tilde{V}_b(\tau)$ obtained after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relation:
+% \begin{align}\label{eqn:Tsys}
+%   \Delta \tilde{V} &= \frac{2\,k_\textrm{B}\,T_\textrm{sys}\sqrt{\Delta B}}{A_\textrm{e}\sqrt{2\,\Delta t}} 
+% \end{align}
+% where, $\Delta f = 80$~kHz, and $\Delta t = 112$~s.
+The choice of threshold for $\tau$ is well outside the {\it foreground window}, where foreground contamination is negligible and thus yields a robust estimate of $T_\textrm{sys}$.
+We find the average system temperature to be $\sim 95$~K across all frequency channels and across all antenna pairs throughout the course of the observation.
+Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise observed in the data.}
+
+{\bf We estimate thermal noise, $\Delta V$, in the visibilities, $V_b(f)$, using the {\it rms} of $\tilde{V}_b(\tau)$ obtained after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
 \begin{align}\label{eqn:Tsys}
-  \Delta V_{b\tau}^\textrm{rms} &= \sqrt{N_\textrm{ch}}\,\Delta V_{bf}^\textrm{rms} \Delta f, \textrm{and} \\
-  \Delta V_{bf}^\textrm{rms} &= \frac{2\,k_\textrm{B}\,T_\textrm{sys}}{A_\textrm{e}\sqrt{2\,\Delta f\,\Delta t}},
+  \Delta \tilde{V} &= \sqrt{N_\textrm{ch}}\,\Delta V \Delta f, \quad\textrm{and} \\
+  \Delta V &= \frac{2\,k_\textrm{B}\,T_\textrm{sys}}{A_\textrm{e}\sqrt{2\,\Delta f\,\Delta t}},
 \end{align}
@@
 The choice of threshold for $\tau$ is well outside the {\it foreground window}, where foreground contamination is negligible and thus yields a robust estimate of $T_\textrm{sys}$.
-We find the average system temperature to be $\sim 95$~K across all frequency channels and across all antenna pairs throughout the course of the observation.
+We find the average system temperature to be $\sim 95$~K.
 Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise observed in the data.}
@@
 The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observing frequency.
-Using simple error propagation, the fractional error in the delay spectrum {\bf amplitude}, $|V_{b\tau}|$, caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the MWA observing frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index.
+Using simple error propagation, the fractional error in the delay spectrum {\bf amplitude} caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the MWA observing frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index.
 From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for point sources from NVSS and SUMSS catalogs.
@@
 In addition to intrinsic model uncertainty, delay spectra from simulations and data each have fluctuations due to thermal noise in the delay spectrum with rms $\sim$ 1.4$\times 10^7$~Jy~Hz.
-We estimate the ratio of delay spectra from data and simulations as $\rho = |V^\textrm{D}_{b\tau}(\boldsymbol{b},\tau)|\,/\,|V^\textrm{S}_{b\tau}(\boldsymbol{b},\tau)|$, where superscripts D and S denote data and simulation, respectively.
+We estimate the ratio of delay spectra from data and simulations as $\rho = |\tilde{V}^\textrm{D}_b(\tau)|\,/\,|\tilde{V}^\textrm{S}_b(\tau)|$, where superscripts D and S denote data and simulation, respectively.
 The median absolute deviation of $\log_{10}\rho$ inside the {\it foreground wedge} for both pointings is $\approx 0.28$.
@@
 
-Note that since {\it visibility} in equation~\ref{eqn:obsvis} is a complex quantity, its {\bf delay} transform {\bf contains} positive and negative delays.
+% Note that since {\it visibility} in equation~\ref{eqn:obsvis} is a complex quantity, its {\bf delay} transform {\bf contains} positive and negative delays.
 Negative and positive delays correspond to the two hemispheres of the sky as transected by the antenna spacing vector.
@@
 \begin{align}\label{eqn:vis}
-  V_{uf} &\equiv V_{uf}(\boldsymbol{u},f) \nonumber\\
-  &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_f\,e^{-i2\pi\boldsymbol{u}\cdot\hat{\boldsymbol{s}}}\,\dif\Omega,
+  V_u(f) &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_\textrm{i}(f)\,e^{-i2\pi\boldsymbol{u}\cdot\hat{\boldsymbol{s}}}\,\dif\Omega,
 \end{align}
@@
 \begin{align}\label{eqn:los-transform}
-  V_{u\eta} &\equiv V_{u\eta}(\boldsymbol{u},\eta) \nonumber\\
-  &\equiv \int V_{uf}(\boldsymbol{u},f)\,W'_f\,e^{i2\pi f\eta}\,\dif f
+  \tilde{V}_u(\eta) &\equiv \int V_u(f)\,W(f)\,e^{i2\pi f\eta}\,\dif f
 \end{align}
@@
 This approach has been discussed in detail in \citet{mor04}.
-The spatial power spectrum of EoR H{\sc i} distribution, $P(\boldsymbol{k}_\perp,k_\parallel)$, and $V_{u\eta}$ are related by \citep{mor04,mcq06,par12a}:
+The spatial power spectrum of EoR H{\sc i} distribution, $P(\boldsymbol{k}_\perp,k_\parallel)$, and $\tilde{V}_u(\eta)$ are related by \citep{mor04,mcq06,par12a}:
 \begin{align}\label{eqn:true-power_spectrum}
-  P(\boldsymbol{k}_\perp,k_\parallel) &\simeq |V_{u\eta}|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
+  P(\boldsymbol{k}_\perp,k_\parallel) &\simeq |\tilde{V}_u(\eta)|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
 \end{align}
 where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, $\lambda$ is the wavelength of the band center, and $k_\textrm{B}$ is the Boltzmann constant.
-Thus, $V_{u\eta}$ inferred from observations, in units of Jy~Hz, can be converted into an equivalent cosmological H{\sc i} power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$, in units of K$^2$~Mpc$^3$ or, more generally, K$^2$(Mpc/$h$)$^3$, where $h$ is the Hubble constant factor.
+Thus, $\tilde{V}_u(\eta)$ inferred from observations, in units of Jy~Hz, can be converted into an equivalent cosmological H{\sc i} power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$, in units of K$^2$~Mpc$^3$ or, more generally, K$^2$(Mpc/$h$)$^3$, where $h$ is the Hubble constant factor.
 

### 2015-02-11 ms.tex
Address referee comments for new draft
from d946469:
============

In ApJ_referee_report_1_addressed.txt,

(*) Address referee comments with reference to the alternate version of
the draft
(*) Draft contents of the cover letter

In ms.tex,

(*) Change wording in abstract slightly
(*) Change wording in S5 slightly and cite Lonsdale et al. 2009 and
Tingay et al. 2013 for the MWA instrument.
(*) Change wording on Tsys estimation slightly in S5.2 Modeling
subsection
(*) Change wording slightly and mention Pober et al (2015, in
preparation) Foreground model subsection to support use of wide-field
f

@@
 Detection of 21~cm emission of H{\sc i} from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission.
-We investigate the signatures of wide--field measurements and an all--sky foreground model using the delay spectrum technique that maps the measurements to foreground object locations through signal delays between antenna pairs.
+{\bf We investigate the signatures of wide--field measurements and an all--sky foreground model using the delay spectrum technique that maps the measurements to foreground object locations through signal delays between antenna pairs.
 We demonstrate interferometric measurements are inherently sensitive to all scales, including the largest angular scales, owing to the nature of wide--field measurements.
 These wide--field effects are generic to all observations but antenna shapes impact their amplitudes substantially.
-A dish--shaped antenna yields the most desirable features from a foreground contamination viewpoint, relative to a dipole or a phased array.
+A dish--shaped antenna yields the most desirable features from a foreground contamination viewpoint, relative to a dipole or a phased array}.
 Comparing data from recent Murchison Widefield Array observations, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view.
-We identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings.
+We identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings {\bf that usually represent structures on small scales}.
 For signals entering through the primary field of view, compact emission dominates the foreground contamination.
@@
 
-{\bf In \S\ref{sec:delay-spectrum} we provide a brief overview of the delay spectrum technique \citep{par12a,par12b}.
+{\bf In \S\ref{sec:delay-spectrum} we provide an overview of the delay spectrum technique.
 We investigate signatures generic to all wide--field measurements of EoR power spectra in \S\ref{sec:wide-field}.
@@
 
+We provide a short overview of the delay spectrum technique \citep{par12a,par12b}.
+
 Interferometer array data known as {\it visibilities}, $V_b(f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis.
@@
 where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector.
-This equation is valid {\bf in general, including wide--field measurements and is a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}}.
+This equation is valid {\bf in general, including wide--field measurements, and is a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}}.
 
@@
 
-In order to express a quantity derived from $\tilde{V}_b(\tau)$ that has units K$^2$(Mpc/$h$)$^3$ which is identical to that of the cosmological H{\sc i} power spectrum, we define the delay power spectrum:
+In order to express a quantity derived from $\tilde{V}_b(\tau)$ whose units are the same as that of the cosmological H{\sc i} power spectrum, we define the delay power spectrum:
 \begin{align}\label{eqn:delay-power-spectrum}
@@
 where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
+$P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$ is in units of K$^2$(Mpc/$h$)$^3$.
 
@@
 {\bf In the specific case of the MWA, which has a passband constructed using coarse channels, there are period grating responses resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.
-Thus with reference to the MWA, the {\it EoR window} lies outside the dashed lines but inside the first grating response ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) and is shown in medium shade}.
+Thus the MWA {\it EoR window} lies outside the dashed lines but inside the first grating response ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) and is shown in medium shade}.
 
@@
 
-In wide--field measurements, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view.
+In wide--field measurements, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view, a point also supported by Pober et al.
+(2015, in preparation).
 We use a foreground model that includes both diffuse and bright compact components.
@@
 Figure~\ref{fig:sky-model-generic} shows the diffuse (top) and compact (bottom) foreground emission model used in our study.
-In this snapshot at 0.09~hours LST, the Galactic center in the diffuse model has just set in the West.
+In this snapshot pointed towards zenith at 0.09~hours LST, the Galactic center in the diffuse model has just set in the West.
 
@@
 \includegraphics[width=\linewidth]{figure3.eps}
-\caption{Foreground model at 185~MHz consisting of diffuse emission (top) and bright point sources (bottom) visible during a snapshot at 0.09~hours LST.
-In the diffuse model, the Galactic center appears to have just set in the West.
+\caption{Foreground model at 185~MHz consisting of diffuse emission (top) in units of K and bright point sources (bottom) in units of Jy, visible during a snapshot at 0.09~hours LST.
+In the diffuse model, the Galactic center has just set in the West.
+Color scales are logarithmic.
 \label{fig:sky-model-generic}}
@@
 
-The power patterns of the aforementioned antenna geometries at 185~MHz for the {\it zenith} pointing are shown in Figure~\ref{fig:power-patterns}.
+The power patterns of the aforementioned antenna geometries at 185~MHz for this zenith pointing are shown in Figure~\ref{fig:power-patterns}.
 
@@
 
-We now use our simulations to analyze features in observed delay power spectrum obtained using the MWA.
+We now use our simulations to analyze features in observed delay power spectrum obtained using the MWA instrument \citep{lon09,tin13}.
 
@@
 This yields a primary field of view $\gtrsim$~20\arcdeg~ wide and multiple secondary lobes.
-The array layout is described in \citet{bea12}.
+See \citet{bea12} for the tile layout.
 %{\bf At the observing frequency of 185~MHz, we use $A_\textrm{e}=10.5$~m$^2$}.
@@
 {\bf The MWA is expected to be sensitive to the power spectrum of the H{\sc i} signal over the redshift range $6<z<10$ \citep{bow06,bea13,thy13}.
-Over 600~hours have been currently observed using the MWA, targeting the science objectives outlined in \citet{bow13}.}
+Over 600~hours have been currently observed using the MWA, targeting science objectives outlined in \citet{bow13}.}
 
-The MWA targets two primary low-foreground fields for reionization observations.
+The MWA targets two primary low--foreground fields for reionization observations.
 Here, we focus on the field at RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$.
@@
 The observations {\bf used} here were acquired on 2013~August~23.
-We have chosen two, 112 second long sections from this night for detailed study.
+We have chosen two sections of duration 112 seconds each from this night for detailed study.
 These were chosen to provide a selection of possible foreground and instrumental conditions.
@@
 
-{\bf We estimate thermal noise, $\Delta V$, in the visibilities, $V_b(f)$, using the {\it rms} of $\tilde{V}_b(\tau)$ obtained after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
+{\bf We estimate thermal noise, $\Delta V$, in the observed visibilities, $V_b(f)$, using the {\it rms} of $\tilde{V}_b(\tau)$ obtained from data after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
 \begin{align}\label{eqn:Tsys}
@@
 With the addition of noise in the visibilities, some of the weaker features may not be as prominently visible.
-Since the foreground signatures are far too numerous {\bf and subject to a multitude of parameters like} baseline length and orientation, power pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features in the delay spectra of foreground emission.
+Since the foreground signatures are far too numerous {\bf and subject to a multitude of parameters like} baseline length and orientation, power pattern, patch of sky under observation, and instrumental configuration, we highlight only the most notable features in the foreground delay power spectra.
 
@@
 This feature is strongest at short antenna spacings and fades with increasing antenna spacing.
-The {\bf bright} signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) because the Galactic center is below the horizon.
+The {\bf bright} signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) because the Galactic center has set below the horizon.
 
@@
 As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inversely proportional to the length of the baseline projected in the direction of the emission.
-Thus, in the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at $\tau=0$, fades away on antenna spacings $|\boldsymbol{b}| \gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0.75\arcdeg.
+Thus, in the {\it zenith} pointing, the horizontal line at $\tau=0$, fades away on antenna spacings $|\boldsymbol{b}| \gtrsim 125$~m because the diffuse sky model is devoid of spatial structures on scales $\lesssim$~0.75\arcdeg.
 
@@
 
-The amplitude response of an interferometer to a point source is, to first order, flat cross baseline length.
-Since the primary lobe of the power pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with $\tau>0$ corresponding to the position of the primary lobe of the power pattern.
-In the {\it zenith} pointing, compact emission from the same patch of sky is seen as a bright horizontal arm at $\tau=0$ since the primary lobe of the power pattern is centered at zenith.
+The amplitude response of an interferometer to a point source is, to first order, flat across baseline length.
+Since the primary field of view in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with $\tau>0$.
+In the {\it zenith} pointing, compact emission from the same patch of sky is seen as a bright horizontal arm at $\tau=0$ since the primary field of view is centered at zenith.
 
@@
 
-Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional {\bf delay power spectra} plots shown in Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}.
+Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the {\bf delay power spectra} plots shown in Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}.
 The diffuse, compact, and composite components are shown as solid red, cyan, and black lines, respectively.
@@
 The peak at $\tau=0$ (corresponding to the primary lobe in the power pattern) with a value of $\sim 10^7$--$10^8$~K$^2$(Mpc/$h$)$^3$, independent of antenna spacing, is predominantly determined by compact emission.
-The peak at $\tau=0$ from diffuse foreground model is $\sim 10^3$ times fainter and decreases rapidly with increase in antenna spacing.
+The peak at $\tau=0$ from diffuse emission is $\sim 10^3$ times fainter and decreases rapidly with increase in antenna spacing.
 This is the response expected from different antenna spacings towards compact and diffuse emission.
@@
 The bright Galactic center will appear at $\tau=0$ in northward antenna spacings and close to negative horizon delay limit on eastward antenna spacings.
-(c): Simulated delay spectra {\bf power} on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}).
+(c): Simulated delay {\bf power} spectra on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}).
 White lines denote horizon delay limits.
@@
 Discarded antenna spacings (black vertical stripes) have lengths $|\boldsymbol{b}|<30$~m (leftward of vertical dashed line) and orientations $|\theta_\textrm{b}|<$~22.5\arcdeg.
-The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} is lowered by {\bf nearly two orders} of magnitude when short eastward antenna spacings are discarded.
+The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} is lowered by a factor $\sim 100$ when short eastward antenna spacings are discarded.
 \label{fig:before-after}}
@@
 
-Simulating in many important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright point sources from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA {\bf to within expected uncertainties in foreground models}.
+Simulating in {\bf many} important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright point sources from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA {\bf to within expected uncertainties in foreground models}.
 
@@
 In general, {\bf delay power spectrum} signatures of compact emission are center--heavy while those of diffuse emission are edge--heavy which results in the {\it pitchfork} signature.
-This will be distinctly visible when the thermal noise floor is sufficiently lowered, as longer observations are processed.}
+This will be distinctly visible when the thermal noise floor is sufficiently lowered, as larger volumes of data are processed.}
 
@@
 $\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward (zenith) directions at the telescope site.
-Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at positive and negative delays, respectively.
-Similarly, an object in the northern sky will appear at a positive delay at an antenna spacing oriented northward.
-For all observations used in this study, we use the zenith as the phase center, {\bf for which $\tau\equiv 0$}.
+Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at $\tau>0$ and $\tau<0$, respectively.
+Similarly, an object in the northern sky will appear at $\tau>0$ for an antenna spacing oriented northward.
+For all observations used in this study, we use zenith as the phase center, {\bf for which $\tau\equiv 0$}.
 
@@
 If the array of antennas are also coplanar lying on the local tangent plane, then $\boldsymbol{u}=\boldsymbol{b}/\lambda$.
-Under such circumstances, equations~\ref{eqn:obsvis} and \ref{eqn:delay} closely resemble equations~\ref{eqn:vis} and \ref{eqn:los-transform} respectively.
+Under such circumstances, equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform} closely resemble equations~\ref{eqn:vis} and \ref{eqn:los-transform} respectively.
 However, they are not quite identical to each other.

### 2015-02-23 ms.tex
Specify cosmological parameters used in paper
from b87734d:
============

In ms.tex,

(*) Specify cosmological parameters used in S2
Address referee comments from second stage review
from 29407e6:
============

In ms.tex,

(*) Modify 3-baseline figure (figure 8) by removing gray shade showing
expected uncertainties. Modify caption accrodingly.
(*) Modify text describing figure 8 according to changes above
(*) Change "equivalent" to "related" in describing interpretations pf
the pitchfork effect in S3 and Summary

@@
 where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
-$P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$ is in units of K$^2$(Mpc/$h$)$^3$.
+Throughout the paper, we use $\Omega_\textrm{M}=0.27$, $\Omega_\Lambda=0.73$, $\Omega_\textrm{K}=1-\Omega_\textrm{M}-\Omega_\Lambda$, $H_0=100\,$km$\,$s$^{-1}\,$Mpc$^{-1}$, and $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$ is in units of K$^2$(Mpc/$h$)$^3$.
 
@@
 We interpret this as due to equal--sized delay bins subtending larger solid angles near the horizon thereby containing larger integrated emission.
-An alternate, but equivalent, interpretation is that baseline vectors (including those with largest lengths) are foreshortened towards the horizon along their orientation.
+An alternate, but related, interpretation is that baseline vectors (including those with largest lengths) are foreshortened towards the horizon along their orientation.
 Thus, they become sensitive to larger angular scales that match the inverse of their foreshortened lengths along these directions.
@@
 The horizon delay limits are shown as a pair of vertical dotted lines.
-The gray shaded area denotes the envelope of expected uncertainty in the {\bf delay power spectrum}.
+% The gray shaded area denotes the envelope of expected uncertainty in the {\bf delay power spectrum}.
 Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the % $\gtrsim$~70\% error owing to
-uncertainty in predicting the spectral index of compact foreground model, while thermal noise fluctuations dominate outside.
+% uncertainty in predicting the spectral index of compact foreground model, while thermal noise fluctuations dominate outside.
 
@@
 {\bf Compact emission dominates the central regions of the delay power spectra while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a characteristic {\it pitchfork}--shaped structure.
-The shaded region denotes the envelope of uncertainty around the composite model delay power spectrum.
-This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits and by thermal noise fluctuations outside}.
-\label{fig:pitchfork-baselines}}
+% The shaded region denotes the envelope of uncertainty around the composite model delay power spectrum.
+This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits and by thermal noise fluctuations outside
+} \label{fig:pitchfork-baselines}}
 \end{figure}
@@
 {\bf We find that all wide--field instruments, typical of modern EoR observatories, imprint a characteristic {\it two--pronged fork} signature in delay spectra.
-There are two equivalent interpretations: delay bins near the horizon subtend larger solid angles and therefore contain larger integrated emission; or, foreshortening of baselines towards the horizon makes them sensitive to emission on large angular scales which match the inverse of their foreshortened lengths.
+There are two related interpretations: delay bins near the horizon subtend larger solid angles and therefore contain larger integrated emission; and, foreshortening of baselines towards the horizon makes them sensitive to emission on large angular scales which match the inverse of their foreshortened lengths.
 These effects combined with higher sensitivity of antennas in the primary field of view results in a characteristic {\it pitchfork} signature.

### 2015-02-23 ms.tex
Remove all bold face fonts
from 37bab06:
============

In ms.tex,

(*) Remove all bold face fonts

@@
 Detection of 21~cm emission of H{\sc i} from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission.
-{\bf We investigate the signatures of wide--field measurements and an all--sky foreground model using the delay spectrum technique that maps the measurements to foreground object locations through signal delays between antenna pairs.
+We investigate the signatures of wide--field measurements and an all--sky foreground model using the delay spectrum technique that maps the measurements to foreground object locations through signal delays between antenna pairs.
 We demonstrate interferometric measurements are inherently sensitive to all scales, including the largest angular scales, owing to the nature of wide--field measurements.
 These wide--field effects are generic to all observations but antenna shapes impact their amplitudes substantially.
-A dish--shaped antenna yields the most desirable features from a foreground contamination viewpoint, relative to a dipole or a phased array}.
+A dish--shaped antenna yields the most desirable features from a foreground contamination viewpoint, relative to a dipole or a phased array.
 Comparing data from recent Murchison Widefield Array observations, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view.
-We identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings {\bf that usually represent structures on small scales}.
+We identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings that usually represent structures on small scales.
 For signals entering through the primary field of view, compact emission dominates the foreground contamination.
@@
 Sensitive instruments such as the Square Kilometre Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}.
-Numerous {\bf pathfinders and} precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13,bow13}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{bow06,par12a,bea13,dil13,thy13,pob14}.
-{\bf The Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} (HERA) is currently under construction using new insights gained with the MWA and PAPER}.
+Numerous pathfinders and precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13,bow13}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{bow06,par12a,bea13,dil13,thy13,pob14}.
+The Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} (HERA) is currently under construction using new insights gained with the MWA and PAPER.
 
@@
 \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12b,dil13,pob13}.
-{\bf Since this contamination is expected to be several orders of magnitude stronger than the underlying EoR H{\sc i} signal, it is critical to characterize foregrounds precisely in order to reduce their impact on EoR H{\sc i} power spectrum detection sensitivity}.
+Since this contamination is expected to be several orders of magnitude stronger than the underlying EoR H{\sc i} signal, it is critical to characterize foregrounds precisely in order to reduce their impact on EoR H{\sc i} power spectrum detection sensitivity.
 
 Considerable effort is being made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,dil14,liu14a,liu14b}.
-{\bf A model that provides a generic explanation for the observed foreground power spectrum has emerged, whereby the wide--field (and chromatic) response of the instrument causes the power in smooth spectrum foregrounds to occupy higher $k$--modes into the so-called `wedge'}.
+A model that provides a generic explanation for the observed foreground power spectrum has emerged, whereby the wide--field (and chromatic) response of the instrument causes the power in smooth spectrum foregrounds to occupy higher $k$--modes into the so-called `wedge'.
 The conservative foreground strategy, referred to as {\it avoidance}, that has developed alongside this work is to discard $k$--modes which could be contaminated \citep[e.g.,][]{par14}.
@@
 Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity.
-To decouple these effects from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using a {\it \textbf{per-baseline}} approach of \citet{par12b}.
+To decouple these effects from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using a {\it per-baseline} approach of \citet{par12b}.
 This approximates the power spectrum as the inverse Fourier transform of the spectra generated by the instrument's correlator.
@@
 
-{\bf In \S\ref{sec:delay-spectrum} we provide an overview of the delay spectrum technique.
+In \S\ref{sec:delay-spectrum} we provide an overview of the delay spectrum technique.
 We investigate signatures generic to all wide--field measurements of EoR power spectra in \S\ref{sec:wide-field}.
@@
 Simulations using these observing parameters are compared with the data and analyzed for foreground signatures.
-We report two important findings: foregrounds that most severely {\bf obscure} the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted.
-In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique that minimizes foreground contamination using prior knowledge of the sky to down--weight adversely contaminated baselines}.
+We report two important findings: foregrounds that most severely obscure the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted.
+In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique that minimizes foreground contamination using prior knowledge of the sky to down--weight adversely contaminated baselines.
 We present a summary of our work and findings in \S\ref{sec:summary}.
@@
 where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector.
-This equation is valid {\bf in general, including wide--field measurements, and is a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}}.
-
-{\bf The {\it delay spectrum}, $\tilde{V}_b(\tau)$, is defined as} the inverse Fourier transform of $V_b(f)$ along the frequency coordinate:
+This equation is valid in general, including wide--field measurements, and is a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}.
+
+The {\it delay spectrum}, $\tilde{V}_b(\tau)$, is defined as the inverse Fourier transform of $V_b(f)$ along the frequency coordinate:
 \begin{align}\label{eqn:delay-transform}
@@
 
-{\bf We give a brief overview of some parameters of Fourier space which is generic to all experiments that use a similar approach}.
+We give a brief overview of some parameters of Fourier space which is generic to all experiments that use a similar approach.
 Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of redshifted H{\sc i} observations are calculated.
-$|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction {\bf (tangent plane to the celestial sphere)}, form the $x$--axis.
+$|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction (tangent plane to the celestial sphere), form the $x$--axis.
 $\tau$ and $k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis.
@@
 Its boundaries (solid lines), given by light travel time for corresponding antenna spacings, are referred to as horizon delay limits.
-Narrow extensions of the {\it wedge} (white unshaded strips) are caused by convolution with the {\bf instrument's spectral transfer function}.
-Regions excluding the {\it wedge} are expected to be relatively free of foreground contamination {\bf and are generally referred to as the {\it EoR window}.
+Narrow extensions of the {\it wedge} (white unshaded strips) are caused by convolution with the instrument's spectral transfer function.
+Regions excluding the {\it wedge} are expected to be relatively free of foreground contamination and are generally referred to as the {\it EoR window}.
 There are undesirable grating responses (dot--dashed lines) specific to the MWA.
-Hence, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the MWA {\it EoR window}}.
+Hence, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the MWA {\it EoR window}.
 \label{fig:fourier-space}}
@@
 The region of Fourier space excluding the {\it foreground wedge} and the narrow strips is the so--called {\it EoR window}, shown in light and medium shades.
-{\bf In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\boldsymbol{k}|$.
+In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\boldsymbol{k}|$.
 Hence, the brightest EoR signal will be observed on the shortest baselines and smallest delays.
-Thus the regions of interest for EoR studies relying on {\it avoidance} strategy are just beyond the horizon delay limits (dashed lines) on short baselines, marked as regions of {\it maximal EoR sensitivity}}.
-
-{\bf In the specific case of the MWA, which has a passband constructed using coarse channels, there are period grating responses resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.
-Thus the MWA {\it EoR window} lies outside the dashed lines but inside the first grating response ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) and is shown in medium shade}.
+Thus the regions of interest for EoR studies relying on {\it avoidance} strategy are just beyond the horizon delay limits (dashed lines) on short baselines, marked as regions of {\it maximal EoR sensitivity}.
+
+In the specific case of the MWA, which has a passband constructed using coarse channels, there are period grating responses resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.
+Thus the MWA {\it EoR window} lies outside the dashed lines but inside the first grating response ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) and is shown in medium shade.
 
@@
 
-{\bf In our present study, we use a latitude of -26.701\arcdeg~ and an antenna layout identical to that of the MWA \citep{bea12} for the observatory.
+In our present study, we use a latitude of -26.701\arcdeg~ and an antenna layout identical to that of the MWA \citep{bea12} for the observatory.
 The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km.
@@
 Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
-We choose an observing frequency of 185~MHz ($z\simeq 6.68$) and a flat passband of width $\Delta B = 30.72$~MHz to roughly match those of ongoing MWA EoR observations (discussed in detail in \S\ref{sec:MWA})}.
+We choose an observing frequency of 185~MHz ($z\simeq 6.68$) and a flat passband of width $\Delta B = 30.72$~MHz to roughly match those of ongoing MWA EoR observations (discussed in detail in \S\ref{sec:MWA}).
 
@@
 The model described above is primarily a model of the diffuse foreground sky.
-While it contains faint compact emission blended in with the diffuse emission, {\bf bright point sources have been removed \citep{deo08}}.
+While it contains faint compact emission blended in with the diffuse emission, bright point sources have been removed \citep{deo08}.
 In order to supplement it with missing bright compact emission, we use classical radio source confusion estimates to determine the nominal flux density threshold and include point sources brighter than this threshold.
@@
 \subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{figure4a.eps}} \\
-\subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{figure4b.eps}}
+\subfloat[][Delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{figure4b.eps}}
 \caption{Power patterns (top panels) and simulated delay power spectra (bottom panels) for different antenna shapes at 185~MHz centered on zenith.
@@
 The coarse channel shape is obtained using an 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$.
-{\bf Each of these coarse bands consists of 32 fine channels of width 40~kHz each}.
+Each of these coarse bands consists of 32 fine channels of width 40~kHz each.
 After correcting for the shape of these coarse channels, the fine channels at the edges of these sub--bands are flagged because they are known to be contaminated by aliasing at a low level.
@@
 
-{\bf The MWA is expected to be sensitive to the power spectrum of the H{\sc i} signal over the redshift range $6<z<10$ \citep{bow06,bea13,thy13}.
-Over 600~hours have been currently observed using the MWA, targeting science objectives outlined in \citet{bow13}.}
+The MWA is expected to be sensitive to the power spectrum of the H{\sc i} signal over the redshift range $6<z<10$ \citep{bow06,bea13,thy13}.
+Over 600~hours have been currently observed using the MWA, targeting science objectives outlined in \citet{bow13}.
 
@@
 
-The observations {\bf used} here were acquired on 2013~August~23.
+The observations used here were acquired on 2013~August~23.
 We have chosen two sections of duration 112 seconds each from this night for detailed study.
@@
 
-{\bf We estimate thermal noise, $\Delta V$, in the observed visibilities, $V_b(f)$, using the {\it rms} of $\tilde{V}_b(\tau)$ obtained from data after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
+We estimate thermal noise, $\Delta V$, in the observed visibilities, $V_b(f)$, using the {\it rms} of $\tilde{V}_b(\tau)$ obtained from data after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
 \begin{align}\label{eqn:Tsys}
@@
 We find the average system temperature to be $\sim 95$~K.
-Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise observed in the data.}
+Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise observed in the data.
 
@@
 With the aforementioned foreground model, and instrumental and observational parameters, we simulate visibilities using equation~\ref{eqn:obsvis}.
-Figure~\ref{fig:fhd-sim-comparison} shows the {\bf delay power spectra} from {\it off--zenith} and {\it zenith} pointings obtained from MWA observations and modeling.
+Figure~\ref{fig:fhd-sim-comparison} shows the delay power spectra from {\it off--zenith} and {\it zenith} pointings obtained from MWA observations and modeling.
 Notice the qualitative agreement of amplitude and structure between the two.
@@
 The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observing frequency.
-Using simple error propagation, the fractional error in the delay spectrum {\bf amplitude} caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the MWA observing frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index.
+Using simple error propagation, the fractional error in the delay spectrum amplitude caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the MWA observing frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index.
 From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for point sources from NVSS and SUMSS catalogs.
 Although the model of \citet{deo08} yields a spectral index per direction on the sky, we could assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from the 408~MHz map of \citet{has82}.
-Thus, fractional errors in delay spectrum {\bf amplitudes} from compact and diffuse components are $\sim$70\% and $\sim$30\% respectively.
+Thus, fractional errors in delay spectrum amplitudes from compact and diffuse components are $\sim$70\% and $\sim$30\% respectively.
 
@@
 We also simulated delay spectra after assigning spectral indices drawn randomly from a {\it gaussian} distribution with a mean of $\langle\alpha_\textrm{sp}\rangle=-0.83$ and a HWHM of $\Delta\alpha_\textrm{sp}=0.35$ to the point sources in our compact foreground model.
-These simulations typically yielded a median absolute deviation of $\approx 0.29$ for $\log_{10}\rho$ indicating fractional differences of $\sim$~95\% {\bf between different realizations}.
+These simulations typically yielded a median absolute deviation of $\approx 0.29$ for $\log_{10}\rho$ indicating fractional differences of $\sim$~95\% between different realizations.
 This demonstrates that a fractional deviation of $\sim$~90\% observed between data and simulations is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account.
@@
 A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power pattern \citep{ber15}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide--field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations, but is beyond the scope of this paper.
-{\bf Hereafter, our focus} is to explore in detail the foreground signatures embedded in the {\it foreground wedge} of the MWA instrument.
+Hereafter, our focus is to explore in detail the foreground signatures embedded in the {\it foreground wedge} of the MWA instrument.
 
@@
 With the addition of noise in the visibilities, some of the weaker features may not be as prominently visible.
-Since the foreground signatures are far too numerous {\bf and subject to a multitude of parameters like} baseline length and orientation, power pattern, patch of sky under observation, and instrumental configuration, we highlight only the most notable features in the foreground delay power spectra.
+Since the foreground signatures are far too numerous and subject to a multitude of parameters like baseline length and orientation, power pattern, patch of sky under observation, and instrumental configuration, we highlight only the most notable features in the foreground delay power spectra.
 
@@
 \includegraphics[width=\linewidth]{figure7.eps}
-\caption{Simulated {\bf delay power spectra (in units of K$^2$~(Mpc/$h$)$^3$)} for for the diffuse (left) and compact (right) foreground models in the {\it off--zenith} (top) and {\it zenith} (bottom) pointings without any thermal noise.
+\caption{Simulated delay power spectra (in units of K$^2$~(Mpc/$h$)$^3$) for for the diffuse (left) and compact (right) foreground models in the {\it off--zenith} (top) and {\it zenith} (bottom) pointings without any thermal noise.
 The axes and color scale are identical to those in Figure~\ref{fig:fhd-sim-comparison}.
 In the {\it off-zenith} pointing, emission from the Galactic center is the most prominent feature seen as a branch at $\tau<0$.
-In the {\it zenith} pointing, {\bf delay power spectrum} from diffuse emission has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings due to wide--field effects.
+In the {\it zenith} pointing, delay power spectrum from diffuse emission has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings due to wide--field effects.
 Compact emission is centrally concentrated.
@@
 This feature is strongest at short antenna spacings and fades with increasing antenna spacing.
-The {\bf bright} signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) because the Galactic center has set below the horizon.
+The bright signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) because the Galactic center has set below the horizon.
 
@@
 
-In contrast to the {\bf delay power spectra} of diffuse emission, compact emission (right panels) manifests as a center--heavy structure in either pointing.
+In contrast to the delay power spectra of diffuse emission, compact emission (right panels) manifests as a center--heavy structure in either pointing.
 
@@
 
-When not dominated by the bright emission from the Galactic center, the {\bf delay power spectrum} of the combined foreground model is composed of diffuse and compact emission, {\bf both of which are significant}.
+When not dominated by the bright emission from the Galactic center, the delay power spectrum of the combined foreground model is composed of diffuse and compact emission, both of which are significant.
 This is illustrated by a more detailed examination of the {\it zenith} pointing in our study.
 
-Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the {\bf delay power spectra} plots shown in Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}.
+Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the delay power spectra plots shown in Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}.
 The diffuse, compact, and composite components are shown as solid red, cyan, and black lines, respectively.
@@
 \includegraphics[width=\linewidth]{figure8.eps}
-\caption{Simulated {\bf delay power spectra} for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}).
+\caption{Simulated delay power spectra for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}).
 The baseline length and orientation is specified in each panel.
@@
 Vertical dotted lines mark the horizon delay limits.
-{\bf Compact emission dominates the central regions of the delay power spectra while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a characteristic {\it pitchfork}--shaped structure.
+Compact emission dominates the central regions of the delay power spectra while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a characteristic {\it pitchfork}--shaped structure.
 % The shaded region denotes the envelope of uncertainty around the composite model delay power spectrum.
-This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits and by thermal noise fluctuations outside
-} \label{fig:pitchfork-baselines}}
+This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits and by thermal noise fluctuations outside.
+\label{fig:pitchfork-baselines}}
 \end{figure}
@@
 Here, diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at $\tau=0$.
-In fact, even on widely spaced antennas, diffuse emission in the {\bf delay power spectrum} near the horizon delay limits exceeds that in the primary lobe by about three orders of magnitude.
+In fact, even on widely spaced antennas, diffuse emission in the delay power spectrum near the horizon delay limits exceeds that in the primary lobe by about three orders of magnitude.
 This feature was described in \S\ref{sec:diffuse-long-baselines}.
@@
 We also note that increasing the antenna spacing progressively improves the resolution along the delay axis by increasing the number of delay bins inside the {\it foreground wedge}.
-This improves the localization of foreground objects whose signatures are imprinted in the {\bf delay power spectrum}.
-For instance, there is an increase in the number of secondary peaks in the {\bf delay power spectrum} between $\tau=0$ and horizon delay limits as the antenna spacing increases from $\sim$84~m to $\sim$171~m.
+This improves the localization of foreground objects whose signatures are imprinted in the delay power spectrum.
+For instance, there is an increase in the number of secondary peaks in the delay power spectrum between $\tau=0$ and horizon delay limits as the antenna spacing increases from $\sim$84~m to $\sim$171~m.
 In this case, these correspond to secondary lobes of the power pattern that lie between the primary lobe and the horizon along the local meridian.
@@
 Figure~\ref{fig:baseline-breakup} shows the delay spectra on baselines oriented northward ($67.5\arcdeg\le \theta_\textrm{b} < 112.5\arcdeg$) at the top and eastward ($-22.5\arcdeg~\le \theta_\textrm{b} < 22.5\arcdeg$) at the bottom.
-The Galactic center manifests itself most distinctly near the negative horizon delay limit on short eastward baselines in the {\bf delay power spectrum} (bottom panel of Figure~\ref{fig:baseline-breakup}).
+The Galactic center manifests itself most distinctly near the negative horizon delay limit on short eastward baselines in the delay power spectrum (bottom panel of Figure~\ref{fig:baseline-breakup}).
 Consequently, the spillover caused by the instrument's spectral transfer function from the {\it foreground wedge} into the {\it EoR window} affects the northward baselines the least and is most severe in eastward baselines (particularly the short ones) evident by the bright vertical stripes of foreground contamination.
@@
 The bright Galactic center will appear at $\tau=0$ in northward antenna spacings and close to negative horizon delay limit on eastward antenna spacings.
-(c): Simulated delay {\bf power} spectra on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}).
+(c): Simulated delay power spectra on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}).
 White lines denote horizon delay limits.
@@
 In our example, we discard eastward antenna spacings ($-22.5\arcdeg\le\theta_\textrm{b}< 22.5\arcdeg$) of lengths $|\boldsymbol{b}| < 30$~m.
-{\bf Foreground contamination was found to be insensitive to removal of wider antenna spacings as discussed below}.
+Foreground contamination was found to be insensitive to removal of wider antenna spacings as discussed below.
 Figure~\ref{fig:before-after} shows the delay spectra obtained with all antenna spacings (top panel) and after applying our screening technique (bottom panel) on the {\it off--zenith} observation.
@@
 \includegraphics[width=\linewidth]{figure10.eps}
-\caption{Simulated delay spectra {\bf power} for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
+\caption{Simulated delay spectra power for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
 Discarded antenna spacings (black vertical stripes) have lengths $|\boldsymbol{b}|<30$~m (leftward of vertical dashed line) and orientations $|\theta_\textrm{b}|<$~22.5\arcdeg.
@@
 Such cases will necessitate discarding more or, in the worst case, all of the eastward oriented antenna spacings.
-{\bf These results use the MWA as an example.
-In general, the array layout will play a significant role in determining such thresholds}.
+These results use the MWA as an example.
+In general, the array layout will play a significant role in determining such thresholds.
 
 In principle, instead of discarding selected antenna spacings altogether, we could down--weight them based on an optimal scheme.
-{\bf For instance, the estimates of covariance computed from the delay transform bins can be naturally fed into the covariance--weighted power spectrum estimation techniques \citep{liu14a,liu14b}.
-It could also be used to downweight or flag contaminated baselines in imaging applications}.
+For instance, the estimates of covariance computed from the delay transform bins can be naturally fed into the covariance--weighted power spectrum estimation techniques \citep{liu14a,liu14b}.
+It could also be used to downweight or flag contaminated baselines in imaging applications.
 This technique provides a very simple and yet effective tool in adding a layer of control to mitigate effects of foreground contamination in EoR data analysis.
@@
 Our primary motivation in this work is to understand how the various bright foregrounds will manifest in three dimensional power spectrum of H{\sc i} from 21~cm reionization observations.
-{\bf In units of temperature variance}, the dynamic range between bright foregrounds and the 21~cm signal is expected to be $\sim 10^8$; a detailed understanding of how foregrounds can corrupt the 21~cm power spectrum is therefore essential.
+In units of temperature variance, the dynamic range between bright foregrounds and the 21~cm signal is expected to be $\sim 10^8$; a detailed understanding of how foregrounds can corrupt the 21~cm power spectrum is therefore essential.
 This analysis extends previous work by simulating the entire sky rather than just the central field of view and by providing a comparison with early observations with the MWA.
-By making use of the delay spectrum technique to estimate the power spectrum, we are able to observe the effects of foregrounds while avoiding entanglements with more {\bf sophisticated} power spectrum estimators.
-
-{\bf We find that all wide--field instruments, typical of modern EoR observatories, imprint a characteristic {\it two--pronged fork} signature in delay spectra.
+By making use of the delay spectrum technique to estimate the power spectrum, we are able to observe the effects of foregrounds while avoiding entanglements with more sophisticated power spectrum estimators.
+
+We find that all wide--field instruments, typical of modern EoR observatories, imprint a characteristic {\it two--pronged fork} signature in delay spectra.
 There are two related interpretations: delay bins near the horizon subtend larger solid angles and therefore contain larger integrated emission; and, foreshortening of baselines towards the horizon makes them sensitive to emission on large angular scales which match the inverse of their foreshortened lengths.
@@
 The amplitude of these generic signatures can be controlled by careful design of antenna aperture.
-In contrast to a dipole and a phased array such as an MWA tile, a dish such as the one proposed for HERA is found to yield the least foreground contamination and thus preferable for EoR studies.}
-
-Simulating in {\bf many} important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright point sources from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA {\bf to within expected uncertainties in foreground models}.
+In contrast to a dipole and a phased array such as an MWA tile, a dish such as the one proposed for HERA is found to yield the least foreground contamination and thus preferable for EoR studies.
+
+Simulating in many important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright point sources from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA to within expected uncertainties in foreground models.
 
@@
 The bright Galactic center at the edge of the western horizon co--located with one of the far secondary lobes of MWA tile power pattern is the brightest source of foreground contamination in the {\it off--zenith} pointing.
-It manifests itself near the negative horizon delay limit in the {\bf delay power spectrum} on antenna spacings oriented eastward.
-
-{\bf Diffuse emission in the primary field of view is prominent on shorter antenna spacings.
+It manifests itself near the negative horizon delay limit in the delay power spectrum on antenna spacings oriented eastward.
+
+Diffuse emission in the primary field of view is prominent on shorter antenna spacings.
 However, it is also prominent near the horizon limits even on wide antenna spacings --- an effect of the wide--field nature of the measurement.
@@
 Features arising from compact emission co--located with primary and secondary lobes of the antenna power pattern have been identified.
-In general, {\bf delay power spectrum} signatures of compact emission are center--heavy while those of diffuse emission are edge--heavy which results in the {\it pitchfork} signature.
-This will be distinctly visible when the thermal noise floor is sufficiently lowered, as larger volumes of data are processed.}
-
-We also provide a simple and effective tool based on the delay spectrum technique {\bf that can potentially} mitigate foreground contamination by {\bf nearly two orders} of magnitude in EoR data analysis by discarding or down--weighting data from antenna pairs most affected by foreground contamination, with negligible loss of sensitivity.
+In general, delay power spectrum signatures of compact emission are center--heavy while those of diffuse emission are edge--heavy which results in the {\it pitchfork} signature.
+This will be distinctly visible when the thermal noise floor is sufficiently lowered, as larger volumes of data are processed.
+
+We also provide a simple and effective tool based on the delay spectrum technique that can potentially mitigate foreground contamination by nearly two orders of magnitude in EoR data analysis by discarding or down--weighting data from antenna pairs most affected by foreground contamination, with negligible loss of sensitivity.
 In conclusion, we find that inclusion of emission models, both diffuse and compact, all the way to the horizon is essential to explaining the observed power spectrum.
@@
 Negative and positive delays correspond to the two hemispheres of the sky as transected by the antenna spacing vector.
-{\bf Figure~\ref{fig:delay-cartoon} illustrates the radio interferometer delays and conventions used in our paper}.
+Figure~\ref{fig:delay-cartoon} illustrates the radio interferometer delays and conventions used in our paper.
 $\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward (zenith) directions at the telescope site.
@@
 Similarly, an object in the northern sky will appear at $\tau>0$ for an antenna spacing oriented northward.
-For all observations used in this study, we use zenith as the phase center, {\bf for which $\tau\equiv 0$}.
+For all observations used in this study, we use zenith as the phase center, for which $\tau\equiv 0$.
 
@@
 \includegraphics[width=\linewidth]{figureA1.eps}
-\caption{{\bf Illustration of radio interferometer delays and conventions used in this paper.
+\caption{Illustration of radio interferometer delays and conventions used in this paper.
 Two antennas labeled as A$_1$ and A$_2$ are separated by vector $\boldsymbol{b}$ on the local tangent plane.
@@
 Conversely, if $\boldsymbol{b}$ is oriented northward (not shown here), then $\tau_\textrm{n} > 0$, $\tau_\textrm{s} < 0$, and $\tau_\textrm{e}=\tau_\textrm{w}=0$}.
-\label{fig:delay-cartoon}}
+\label{fig:delay-cartoon}
 \end{figure}
@@
 
-Equation~\ref{eqn:obsvis} can be {\bf equivalently expressed} as:
+Equation~\ref{eqn:obsvis} can be equivalently expressed as:
 \begin{align}\label{eqn:vis}

### 2015-02-23 ms.tex
Cite COTTER pipeline by Offringa et al(2015)
from e109571:
============

In eor.bib,

(*) Add bibtex entry for Offring et al(2015) COTTER pipeline

In ms.tex,

(*) Add citation to COTTER pipeline by Offringa et al(2015)
(*) Remove citations to Offringa et al 2010, 2012
(*) Remove footnote for AOFLAGGER

@@
 
-The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off10,off12}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz.
+The data are flagged for interference \citep{off15}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz.
 These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey \citep[MWACS;][]{hur14}.

### 2015-02-23 ms.tex
Remove commented text
from 049772a:
============

In ms.tex,

(*) Remove commented text

@@
 
-%% \title{Detecting Epoch of Reionization in redshifted 21~cm: contamination in EoR window}
-%% \title{Limits on the detection of the Epoch of Reionization from MWA observations of the redshifted 21~cm line}
 \title{Foregrounds in Wide--Field Redshifted 21~cm Power Spectra}
@@
 This commenced the epoch of reionization (EoR) --- a period of non--linear growth of matter density perturbations and astrophysical evolution.
-% Studying the EoR holds the key to understanding this evolution.
 
@@
 This approximates the power spectrum as the inverse Fourier transform of the spectra generated by the instrument's correlator.
-
-% {\bf We investigate signatures generic to all wide--field measurements of EoR power spectra and explore antenna shapes that are most desirable from the viewpoint of foreground contamination for future EoR experiments}.
-After matching observations from the Murchison Widefield Array with instrumental simulations of the entire sky, we examine the {\bf shape of the delay spectrum} of each baseline.
-We report two important findings: in the {\it avoidance} strategy, the foregrounds that most severely {\bf obscure} the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted.
-We quantify these by separating the simulations into components based on type of foreground emission.
-We then arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.
 
@@
 
-% The presence of a ground plane below the dipole and the phased array introduces a smoother suppression of sensitivity near the horizon, whereas this transition is abrupt in the case of the dish.
-
 \subsection{Foreground Model}\label{sec:foreground}
@@
 
-% \subsubsection{Diffuse Foreground Model}\label{sec:DSM}
-
 For the diffuse component, we use an all--sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz.
@@
 We model the diffuse foreground spectra with a unique spectral index at each pixel in the map, estimated from model maps at 170~MHz and 200~MHz.
-
-% \subsubsection{Compact Foreground Model}\label{sec:CSM}
 
@@
 
-% \subsection{The Murchison Widefield Array (MWA)}\label{sec:instrument}
-
 MWA construction was completed in 2012 and, after commissioning, began its EoR observing program in 2013.
@@
 See \citet{bea12} for the tile layout.
-%{\bf At the observing frequency of 185~MHz, we use $A_\textrm{e}=10.5$~m$^2$}.
 
@@
 After correcting for the shape of these coarse channels, the fine channels at the edges of these sub--bands are flagged because they are known to be contaminated by aliasing at a low level.
-% Any spectral flags, or other known weights constitutes $W_\textrm{i}(f)$ in equation~\ref{eqn:obsvis}.
-
-% \subsection{EoR Observations}\label{sec:obsparms}
 
@@
 \end{figure*}
-
-% {\bf We estimate the system temperature, $T_\textrm{sys}$, using the {\it rms} of $\tilde{V}_b(\tau)$ obtained after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relation:
-% \begin{align}\label{eqn:Tsys}
-%   \Delta \tilde{V} &= \frac{2\,k_\textrm{B}\,T_\textrm{sys}\sqrt{\Delta B}}{A_\textrm{e}\sqrt{2\,\Delta t}} 
-% \end{align}
-% where, $\Delta f = 80$~kHz, and $\Delta t = 112$~s.
-The choice of threshold for $\tau$ is well outside the {\it foreground window}, where foreground contamination is negligible and thus yields a robust estimate of $T_\textrm{sys}$.
-We find the average system temperature to be $\sim 95$~K across all frequency channels and across all antenna pairs throughout the course of the observation.
-Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise observed in the data.}
 
@@
 The horizon delay limits are shown as a pair of vertical dotted lines.
-% The gray shaded area denotes the envelope of expected uncertainty in the {\bf delay power spectrum}.
-Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the % $\gtrsim$~70\% error owing to
-% uncertainty in predicting the spectral index of compact foreground model, while thermal noise fluctuations dominate outside.
 
@@
 Compact emission dominates the central regions of the delay power spectra while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a characteristic {\it pitchfork}--shaped structure.
-% The shaded region denotes the envelope of uncertainty around the composite model delay power spectrum.
-This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits and by thermal noise fluctuations outside.
 \label{fig:pitchfork-baselines}}
@@
 
-% Note that since {\it visibility} in equation~\ref{eqn:obsvis} is a complex quantity, its {\bf delay} transform {\bf contains} positive and negative delays.
-Negative and positive delays correspond to the two hemispheres of the sky as transected by the antenna spacing vector.
 Figure~\ref{fig:delay-cartoon} illustrates the radio interferometer delays and conventions used in our paper.

### 2015-02-23 ms.tex
Make minor edits throughout the paper
from a7b9a1f:
============

In ms.tex,

(*) Make minor edits throughout the paper

@@
 \end{equation}
-The delay transform conventions used in this paper are described in appendix~\ref{sec:delay-conventions}. $\tilde{V}_b(\tau)$ is expressed in observer's units as Jy~Hz.
+The delay transform conventions used in this paper are described in appendix~\ref{sec:delay-conventions}. $\tilde{V}_b(\tau)$ is expressed in observer's units of Jy~Hz.
 
@@
 
-We give a brief overview of some parameters of Fourier space which is generic to all experiments that use a similar approach.
+We give a brief overview of some parameters of Fourier space which are generic to all experiments that use a similar approach.
 Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of redshifted H{\sc i} observations are calculated.
@@
 \end{align}
-where, $l$, $m$, and $n$ denote the direction cosines toward east, north, and zenith respectively, with $n\equiv\sqrt{1-l^2-m^2}$, and:
+where, $l$, $m$, and $n$ denote the direction cosines toward East, North, and zenith respectively, with $n\equiv\sqrt{1-l^2-m^2}$, and:
 \begin{align}\label{eqn:solid-angle}
@@
 The phased array has intermediate levels of contamination.
-Foreground power close to the horizon delay limits in all three cases are significant even on long baselines.
+Foreground power close to the horizon delay limits in all three cases is significant even on long baselines.
 The foreshortening of baselines towards the horizon makes them sensitive to foreground emission on large size scales.
@@
 We find the average system temperature to be $\sim 95$~K.
-Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise observed in the data.
+Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise in data.
 
@@
 
-The most prominent signature seen in the {\it off--zenith} pointing (top left panel of Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power pattern.
+The most prominent signature seen in the {\it off--zenith} pointing (top left panel, Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power pattern.
 It appears as a bright branch near the negative delay horizon delay limit.
 This feature is strongest at short antenna spacings and fades with increasing antenna spacing.
-The bright signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) because the Galactic center has set below the horizon.
+The bright signature is absent in the {\it zenith} pointing (bottom left panel, Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) because the Galactic center has set below the horizon.
 
@@
 
-When not dominated by the bright emission from the Galactic center, the delay power spectrum of the combined foreground model is composed of diffuse and compact emission, both of which are significant.
-This is illustrated by a more detailed examination of the {\it zenith} pointing in our study.
+When not dominated by the bright emission from the Galactic center, the delay power spectrum of the combined foreground model is composed of diffuse and compact emission both of which are significant.
+This is illustrated by a detailed examination of the {\it zenith} pointing.
 
-Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the delay power spectra plots shown in Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}.
+Figure~\ref{fig:pitchfork-baselines} shows delay power spectra of three antenna pairs of different spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the delay power spectra plots shown in Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}.
 The diffuse, compact, and composite components are shown as solid red, cyan, and black lines, respectively.
@@
 In fact, even on widely spaced antennas, diffuse emission in the delay power spectrum near the horizon delay limits exceeds that in the primary lobe by about three orders of magnitude.
-This feature was described in \S\ref{sec:diffuse-long-baselines}.
-We attribute this to wide--field measurement effects discussed in \S\ref{sec:wide-field}.
+This feature is described in \S\ref{sec:diffuse-long-baselines} and attributed to wide--field measurement effects discussed in \S\ref{sec:wide-field}.
 
@@
 The Galactic center is the most dominant source of foreground contamination from the diffuse sky model and is co--located with a bright secondary lobe of the power pattern near the western horizon.
-Figure~\ref{fig:delay-map} shows the sky mapped to delays registered by the baseline vectors, of length 100~m for instance, oriented northward (top panel) and eastward (bottom panel).
+Figure~\ref{fig:delay-map} shows the sky mapped to delays registered by the baseline vectors, of length 100~m for instance, oriented towards North (top panel) and East (bottom panel).
 Figure~\ref{fig:baseline-breakup} shows the delay spectra on baselines oriented northward ($67.5\arcdeg\le \theta_\textrm{b} < 112.5\arcdeg$) at the top and eastward ($-22.5\arcdeg~\le \theta_\textrm{b} < 22.5\arcdeg$) at the bottom.
 The Galactic center manifests itself most distinctly near the negative horizon delay limit on short eastward baselines in the delay power spectrum (bottom panel of Figure~\ref{fig:baseline-breakup}).
-Consequently, the spillover caused by the instrument's spectral transfer function from the {\it foreground wedge} into the {\it EoR window} affects the northward baselines the least and is most severe in eastward baselines (particularly the short ones) evident by the bright vertical stripes of foreground contamination.
+Consequently, the spillover caused by the instrument's spectral transfer function from the {\it foreground wedge} into the {\it EoR window} affects the northward baselines the least and is most severe on eastward baselines (particularly the short ones) evident by the bright vertical stripes of foreground contamination.
 
@@
 \caption{(a): Sky model showing compact ({\it top}) and diffuse ({\it bottom}) emission (adopted from Figure~\ref{fig:sky-model-MWA}).
-The Galactic center is very prominent in diffuse emission on the west co--located with a bright secondary lobe of the power pattern.
-(b): Sky hemisphere mapped to delays observed on antenna spacings with northward ({\it top}) and eastward ({\it bottom}) orientations.
+The Galactic center is very prominent in diffuse emission on the West co--located with a bright secondary lobe of the power pattern.
+(b): Sky hemisphere mapped to delays observed on antenna spacings oriented North ({\it top}) and East ({\it bottom}).
 Delays vary linearly with antenna spacing length.
@@
 Discarding antenna spacings with lengths $|\boldsymbol{b}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity as the fraction of discarded baselines increases from $\sim 5$\% to $\sim 25$\%.
-On the other hand, a bright point source at the same location will give rise to foreground contamination even on longer antenna spacings.
-Such cases will necessitate discarding more or, in the worst case, all of the eastward oriented antenna spacings.
-These results use the MWA as an example.
-In general, the array layout will play a significant role in determining such thresholds.
+
+If there was a bright point source in place of the Galactic center, it will give rise to foreground contamination even on longer antenna spacings.
+Such cases will necessitate discarding more or all of the eastward antenna spacings.
+The MWA was used as an example.
+% These results use the MWA as an example.
+In general, the direction, strength and type of emission, and the array layout will determine such thresholds.
 
@@
 It could also be used to downweight or flag contaminated baselines in imaging applications.
-This technique provides a very simple and yet effective tool in adding a layer of control to mitigate effects of foreground contamination in EoR data analysis.
+This technique provides a very simple and yet effective tool to suppress the effects of foreground contamination in EoR data analysis.
 
@@
 
-Simulating in many important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright point sources from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA to within expected uncertainties in foreground models.
+Simulating in many important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission and bright point sources, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA to within expected uncertainties in foreground models.
 
 Our simulations enable us to identify numerous signatures of different components of foreground emission seen in the delay spectra.
-We establish the relationship between these signatures and observing parameters such as antenna pointing and LST, instrument parameters such as antenna power pattern, and foreground parameters such as the nature of emission, spectral index, etc.
+We establish the relationship between these signatures and observing parameters such as antenna pointing and LST, instrument parameters such as antenna power pattern, and foreground parameters such as type of emission, etc.
 
 The bright Galactic center at the edge of the western horizon co--located with one of the far secondary lobes of MWA tile power pattern is the brightest source of foreground contamination in the {\it off--zenith} pointing.
-It manifests itself near the negative horizon delay limit in the delay power spectrum on antenna spacings oriented eastward.
+It manifests itself near the negative horizon delay limit in the delay power spectrum on eastward antenna spacings.
 
@@
 Features arising from compact emission co--located with primary and secondary lobes of the antenna power pattern have been identified.
-In general, delay power spectrum signatures of compact emission are center--heavy while those of diffuse emission are edge--heavy which results in the {\it pitchfork} signature.
-This will be distinctly visible when the thermal noise floor is sufficiently lowered, as larger volumes of data are processed.
+In general, delay power spectrum signatures of compact emission are center--heavy while those of diffuse emission are edge--heavy which results in a characteristic {\it pitchfork} signature.
+This will be evident when thermal noise is sufficiently lowered, as larger volumes of data are processed.
 
@@
 Figure~\ref{fig:delay-cartoon} illustrates the radio interferometer delays and conventions used in our paper.
-$\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward (zenith) directions at the telescope site.
+$\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local East, North (along local meridian) and upward (zenith) directions at the telescope site.
 Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at $\tau>0$ and $\tau<0$, respectively.
@@
 \includegraphics[width=\linewidth]{figureA1.eps}
-\caption{Illustration of radio interferometer delays and conventions used in this paper.
+\caption{Radio interferometer delay conventions used in this paper.
 Two antennas labeled as A$_1$ and A$_2$ are separated by vector $\boldsymbol{b}$ on the local tangent plane.
-The local meridian, local north and local east are shown for reference.
+The local meridian, local North and local East are shown for reference.
 Points labeled as `z', `n', `s', `e' and `w' on the celestial sphere denote zenith, northward, southward, eastward and westward positions, respectively.
@@
 If $\boldsymbol{b}$ is oriented eastward as shown, then $\tau_\textrm{e} > 0$, $\tau_\textrm{w} < 0$, and $\tau_\textrm{n}=\tau_\textrm{s}=0$.
-Conversely, if $\boldsymbol{b}$ is oriented northward (not shown here), then $\tau_\textrm{n} > 0$, $\tau_\textrm{s} < 0$, and $\tau_\textrm{e}=\tau_\textrm{w}=0$}.
-\label{fig:delay-cartoon}
+Conversely, if $\boldsymbol{b}$ is oriented northward (not shown here), then $\tau_\textrm{n} > 0$, $\tau_\textrm{s} < 0$, and $\tau_\textrm{e}=\tau_\textrm{w}=0$.} \label{fig:delay-cartoon}
 \end{figure}
@@
 $w$ is aligned parallel to the direction of the phase center, while $u$ and $v$ lie on the transverse plane perpendicular to it.
-For measurements that are constrained to lie on this plane, we can choose $w=0$ without loss of generality and $\boldsymbol{u}$ effectively reduces to $\boldsymbol{u}\equiv (u,v)$, a two--dimensional vector.
+For measurements that lie on this plane, we can choose $w=0$ without loss of generality and $\boldsymbol{u}$ effectively reduces to $\boldsymbol{u}\equiv (u,v)$, a two--dimensional vector.
 Then, $\boldsymbol{u}$ is directly related to the transverse spatial wavenumber mode as:

### 2015-02-23 ms.tex
Change "alternate" interpretations of pitchfork to two related effects
from fa05a33:
============

In ms.tex,

(*) Change "alternate" interpretations of pitchfork effect to two
related effects contributing to it based on Judd's suggestion

@@
 We interpret this as due to equal--sized delay bins subtending larger solid angles near the horizon thereby containing larger integrated emission.
-An alternate, but related, interpretation is that baseline vectors (including those with largest lengths) are foreshortened towards the horizon along their orientation.
+Further, baseline vectors (including those with largest lengths) are foreshortened towards the horizon along their orientation.
 Thus, they become sensitive to larger angular scales that match the inverse of their foreshortened lengths along these directions.
@@
 We find that all wide--field instruments, typical of modern EoR observatories, imprint a characteristic {\it two--pronged fork} signature in delay spectra.
-There are two related interpretations: delay bins near the horizon subtend larger solid angles and therefore contain larger integrated emission; and, foreshortening of baselines towards the horizon makes them sensitive to emission on large angular scales which match the inverse of their foreshortened lengths.
+This arises from two related effects: delay bins near the horizon subtend larger solid angles and therefore contain larger integrated emission; and, foreshortening of baselines towards the horizon makes them sensitive to emission on large angular scales which match the inverse of their foreshortened lengths.
 These effects combined with higher sensitivity of antennas in the primary field of view results in a characteristic {\it pitchfork} signature.

### 2015-02-26 ms.tex
Remove commented parts of text
from a2e2364:
============

In ms.tex,

(*) Remove commented parts of text
Rearrange keywords
from d3b979d:
============

In ms.tex,

(*) Rearrange keywords

@@
 
-\keywords{cosmology: observations --- large-scale structure of universe --- methods: statistical --- radio continuum: galaxies --- reionization --- techniques: interferometric}
+\keywords{cosmology: observations --- dark ages, reionization, first stars --- large-scale structure of universe --- methods: statistical --- radio continuum: galaxies --- techniques: interferometric}
 
@@
 \centering
-% \includegraphics[width=\linewidth]{figures/v1_0/fourier_space_185.0_MHz_30.7_MHz.eps}
-% \includegraphics[width=\linewidth]{f1.eps}
-% \includegraphics[width=\linewidth]{fig2.eps}
-% \includegraphics[width=\linewidth]{figure2.eps}
 \includegraphics[width=\linewidth]{figure1.eps}
@@
 \centering
-% \includegraphics[width=\linewidth]{figure3.eps}
 \includegraphics[width=\linewidth]{figure2.eps}
@@
 \centering
-% \includegraphics[width=\linewidth]{figure4.eps}
 \includegraphics[width=\linewidth]{figure3.eps}
@@
 \centering
-% \subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{figure5a.eps}} \\
-% \subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{figure5b.eps}}
 \subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{figure4a.eps}} \\
@@
 \centering
-% \subfloat[][Diffuse Sky Model]{\label{fig:DSM}\includegraphics[width=0.45\linewidth]{figure6a.eps}}
-% \subfloat[][Bright Point Source Model]{\label{fig:CSM}\includegraphics[width=0.45\linewidth]{figure6b.eps}}
 \subfloat[][Diffuse Sky Model]{\label{fig:DSM}\includegraphics[width=0.45\linewidth]{figure5a.eps}}
@@
 \centering
-% \includegraphics[width=\linewidth]{figure7.eps}
 \includegraphics[width=\linewidth]{figure6.eps}
@@
 \centering
-% \includegraphics[width=\linewidth]{figure8.eps}
 \includegraphics[width=\linewidth]{figure7.eps}
@@
 \centering
-% \includegraphics[width=\linewidth]{figure9.eps}
 \includegraphics[width=\linewidth]{figure8.eps}
@@
 \centering
-% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{figures/v1_0/sky_model_snapshot_0.eps}}
-% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{figures/v1_0/directional_delay_map_snapshot_0.eps}}
-% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figures/v1_0/baseline_orientation_breakup_snapshot_0.eps}}
-% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{f11a.eps}}
-% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{f11b.eps}}
-% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{f11c.eps}}
-% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{fig12a.eps}}
-% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{fig12b.eps}}
-% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{fig12c.eps}}
 \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{figure9a.eps}}
@@
 \centering
-% \includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_before_after_nside_64_Tsys_95.0K.eps}
-% \includegraphics[width=\linewidth]{f12.eps}
-% \includegraphics[width=\linewidth]{fig13.eps}
 \includegraphics[width=\linewidth]{figure10.eps}
@@
 \centering
-% \includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_nside_64_Tsys_95.0K.eps}
-% \includegraphics[width=\linewidth]{f13.eps}
-% \includegraphics[width=\linewidth]{fig14.eps}
 \includegraphics[width=\linewidth]{figure11.eps}
@@
 The MWA was used as an example.
-% These results use the MWA as an example.
 In general, the direction, strength and type of emission, and the array layout will determine such thresholds.
@@
 \centering
-% \includegraphics[width=\linewidth]{delay_cartoon.eps}
-% \includegraphics[width=\linewidth]{fig1.eps}
-% \includegraphics[width=\linewidth]{figure1.eps}
 \includegraphics[width=\linewidth]{figureA1.eps}

### 2015-03-16 ms.tex
Make ApJ Proof corrections
from 95314e1:
============

In ms.tex,

Make ApJ Proof corrections.

(*) Change HI to H I
(*) Change double hyphens to single hyphens
(*) Change non-linear to nonlinear
(*) Minor changes
(*) Remove hyphens in line-of-sight
(*) Change "dot-dashed" to "dotted-dashed"
(*) Change notation of fractional degrees
(*) Change itemizd list to enumerated list
(*) Change hour/hours to hr
(*) Remove hyphens in signal-to-noise
(*) Change "towards" to "toward"

In eor.bib,

(*) Update bibtex reference to Taylor et al.(1999)

Add proof1.txt containing responses to proof corrections

@@
 
-\title{Foregrounds in Wide--Field Redshifted 21~cm Power Spectra}
+\title{Foregrounds in Wide-Field Redshifted 21~cm Power Spectra}
 
@@
 
-Detection of 21~cm emission of H{\sc i} from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission.
-We investigate the signatures of wide--field measurements and an all--sky foreground model using the delay spectrum technique that maps the measurements to foreground object locations through signal delays between antenna pairs.
-We demonstrate interferometric measurements are inherently sensitive to all scales, including the largest angular scales, owing to the nature of wide--field measurements.
-These wide--field effects are generic to all observations but antenna shapes impact their amplitudes substantially.
-A dish--shaped antenna yields the most desirable features from a foreground contamination viewpoint, relative to a dipole or a phased array.
-Comparing data from recent Murchison Widefield Array observations, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view.
+Detection of 21~cm emission of H~{\sc i} from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission.
+We investigate the signatures of wide-field measurements and an all-sky foreground model using the delay spectrum technique that maps the measurements to foreground object locations through signal delays between antenna pairs.
+We demonstrate interferometric measurements are inherently sensitive to all scales, including the largest angular scales, owing to the nature of wide-field measurements.
+These wide-field effects are generic to all observations but antenna shapes impact their amplitudes substantially.
+A dish-shaped antenna yields the most desirable features from a foreground contamination viewpoint, relative to a dipole or a phased array.
+Comparing data from recent Murchison Widefield Array observations, we demonstrate that the foreground signatures that have the largest impact on the H~{\sc i} signal arise from power received far away from the primary field of view.
 We identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings that usually represent structures on small scales.
@@
 These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier delay space.
-Based on these results, we propose that selective down--weighting of data based on antenna spacing and time can mitigate foreground contamination substantially by a factor $\sim 100$ with negligible loss of sensitivity.
+Based on these results, we propose that selective down-weighting of data based on antenna spacing and time can mitigate foreground contamination substantially by a factor $\sim 100$ with negligible loss of sensitivity.
 
@@
 This period, referred to as the {\it Dark Ages} in the Universe's history, is characterized by the localized accumulation of matter under the influence of gravity.
-And it ended with the formation of the first stars and galaxies which started emitting ultra--violet and X--ray radiation, thereby reionizing the neutral medium in their surroundings.
-This commenced the epoch of reionization (EoR) --- a period of non--linear growth of matter density perturbations and astrophysical evolution.
+And it ended with the formation of the first stars and galaxies which started emitting ultra-violet and X-ray radiation, thereby reionizing the neutral medium in their surroundings.
+This commenced the epoch of reionization (EoR) -- a period of nonlinear growth of matter density perturbations and astrophysical evolution.
 
-Observing redshifted 21~cm radiation generated by the spin flip transition of H{\sc i} has been identified as a direct probe of the EoR \citep{sun72,sco90,mad97,toz00,ili02}.
+Observing redshifted 21~cm radiation generated by the spin flip transition of H~{\sc i} has been identified as a direct probe of the EoR \citep{sun72,sco90,mad97,toz00,ili02}.
 Detecting this signal has recently emerged as a very promising experiment to fill the gaps in our understanding of the Universe's history.
 
-Sensitive instruments such as the Square Kilometre Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}.
-Numerous pathfinders and precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13,bow13}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{bow06,par12a,bea13,dil13,thy13,pob14}.
+Sensitive instruments such as the Square Kilometre Array (SKA) are required for direct observation and tomography of redshifted H~{\sc i}.
+Numerous pathfinders and precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13,bow13}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H~{\sc i} power spectrum \citep{bow06,par12a,bea13,dil13,thy13,pob14}.
 The Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} (HERA) is currently under construction using new insights gained with the MWA and PAPER.
 
-A key challenge in the statistical detection of the redshifted H{\sc i} 21~cm signal, via the spatial power spectrum of temperature fluctuations, arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06,ali08,ber09,ber10,gho12}.
+A key challenge in the statistical detection of the redshifted H~{\sc i} 21~cm signal, via the spatial power spectrum of temperature fluctuations, arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06,ali08,ber09,ber10,gho12}.
 \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12b,dil13,pob13}.
-Since this contamination is expected to be several orders of magnitude stronger than the underlying EoR H{\sc i} signal, it is critical to characterize foregrounds precisely in order to reduce their impact on EoR H{\sc i} power spectrum detection sensitivity.
+Since this contamination is expected to be several orders of magnitude stronger than the underlying EoR H~{\sc i} signal, it is critical to characterize foregrounds precisely in order to reduce their impact on EoR H~{\sc i} power spectrum detection sensitivity.
 
-Considerable effort is being made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,dil14,liu14a,liu14b}.
-A model that provides a generic explanation for the observed foreground power spectrum has emerged, whereby the wide--field (and chromatic) response of the instrument causes the power in smooth spectrum foregrounds to occupy higher $k$--modes into the so-called `wedge'.
-The conservative foreground strategy, referred to as {\it avoidance}, that has developed alongside this work is to discard $k$--modes which could be contaminated \citep[e.g.,][]{par14}.
+Considerable effort is being made toward understanding the $k$-space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,dil14,liu14a,liu14b}.
+A model that provides a generic explanation for the observed foreground power spectrum has emerged, whereby the wide-field (and chromatic) response of the instrument causes the power in smooth spectrum foregrounds to occupy higher $k$-modes into the so-called ``wedge''.
+The conservative foreground strategy, referred to as {\it avoidance}, that has developed alongside this work is to discard $k$-modes which could be contaminated \citep[e.g.,][]{par14}.
 The more aggressive alternative is to subtract a sky model and regain access to modes that would be discarded by {\it avoidance}.
 In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain.
-Here, we focus primarily on extending the {\it avoidance} strategy by identifying foreground components at greatest risk to ``leak'' from foreground modes to EoR modes and proposing a scheme for down--weighting these components.
+Here, we focus primarily on extending the {\it avoidance} strategy by identifying foreground components at greatest risk to ``leak'' from foreground modes to EoR modes and proposing a scheme for down-weighting these components.
 
-Foregrounds with intrinsic deviations from spectral smoothness, instruments with high chromaticity, polarization leakage, calibration errors, or approximations in power spectrum analyses can contaminate the true EoR H{\sc i} power spectrum.
+Foregrounds with intrinsic deviations from spectral smoothness, instruments with high chromaticity, polarization leakage, calibration errors, or approximations in power spectrum analyses can contaminate the true EoR H~{\sc i} power spectrum.
 Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity.
@@
 In \S\ref{sec:delay-spectrum} we provide an overview of the delay spectrum technique.
-We investigate signatures generic to all wide--field measurements of EoR power spectra in \S\ref{sec:wide-field}.
+We investigate signatures generic to all wide-field measurements of EoR power spectra in \S\ref{sec:wide-field}.
 In \S\ref{sec:sim}, we present the foreground model and a variety of instrument models to rank antenna shapes based on foreground contamination.
@@
 We report two important findings: foregrounds that most severely obscure the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted.
-In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique that minimizes foreground contamination using prior knowledge of the sky to down--weight adversely contaminated baselines.
+In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique that minimizes foreground contamination using prior knowledge of the sky to down-weight adversely contaminated baselines.
 We present a summary of our work and findings in \S\ref{sec:summary}.
@@
 
-Interferometer array data known as {\it visibilities}, $V_b(f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis.
+Interferometer array data known as {\it visibilities}, $V_b(f)$, represent correlations between time-series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis.
 If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power pattern, respectively, at different frequencies as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_\textrm{i}(f)$ denotes instrumental bandpass weights, then $V_b(f)$ can be written as:
@@
 where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector.
-This equation is valid in general, including wide--field measurements, and is a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}.
+This equation is valid in general, including wide-field measurements, and is a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}.
 
@@
 \end{align}
-where, $W(f)$ is a spectral weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}, and $\tau$ represents the signal delay between antenna pairs:
+where, $W(f)$ is a spectral weighting function which can be chosen to control the quality of the delay spectrum \citep{ved12,thy13}, and $\tau$ represents the signal delay between antenna pairs:
 \begin{equation}\label{eqn:delay}
@@
 
-The delay spectrum has a close resemblance to cosmological H{\sc i} spatial power spectrum.
+The delay spectrum has a close resemblance to cosmological H~{\sc i} spatial power spectrum.
 Appendix~\ref{sec:EoR-power-spectrum} gives an overview of the similarities and differences between the two.
@@
 
-In order to express a quantity derived from $\tilde{V}_b(\tau)$ whose units are the same as that of the cosmological H{\sc i} power spectrum, we define the delay power spectrum:
+In order to express a quantity derived from $\tilde{V}_b(\tau)$ whose units are the same as that of the cosmological H~{\sc i} power spectrum, we define the delay power spectrum:
 \begin{align}\label{eqn:delay-power-spectrum}
@@
 \end{align}
-where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
+where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H~{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
 Throughout the paper, we use $\Omega_\textrm{M}=0.27$, $\Omega_\Lambda=0.73$, $\Omega_\textrm{K}=1-\Omega_\textrm{M}-\Omega_\Lambda$, $H_0=100\,$km$\,$s$^{-1}\,$Mpc$^{-1}$, and $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$ is in units of K$^2$(Mpc/$h$)$^3$.
@@
 In summary, the delay spectrum, $\tilde{V}_b(\tau)$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}.
-$\tilde{V}_b(\tau)$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument.
+$\tilde{V}_b(\tau)$ captures all the effects of EoR H~{\sc i} signal corruption caused by foregrounds and the instrument.
 At the same time, it is closely related to the sought power spectrum containing critical information about spatial scales.
@@
 We give a brief overview of some parameters of Fourier space which are generic to all experiments that use a similar approach.
-Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of redshifted H{\sc i} observations are calculated.
-$|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction (tangent plane to the celestial sphere), form the $x$--axis.
-$\tau$ and $k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis.
-Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by the antenna spacings and the light travel times across them.
+Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of redshifted H~{\sc i} observations are calculated.
+$|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction (tangent plane to the celestial sphere), form the $x$-axis.
+$\tau$ and $k_\parallel$, denoting spatial scales along line of sight form the $y$-axis.
+Foreground emission maps to a wedge-shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by the antenna spacings and the light travel times across them.
 These boundaries, called horizon delay limits \citep{ved12,par12b}, are shown by solid lines.
@@
 \includegraphics[width=\linewidth]{figure1.eps}
-\caption{Fourier space in which delay (and power) spectra of EoR H{\sc i} signals are calculated.
-The $x$--axis is denoted by $|\boldsymbol{b}|$ (antenna spacing) or $k_\perp$ (transverse wavenumber).
-The $y$--axis denoted by $\tau$ (delay) or $k_\parallel$ (line--of--sight wavenumber).
+\caption{Fourier space in which delay (and power) spectra of EoR H~{\sc i} signals are calculated.
+The $x$-axis is denoted by $|\boldsymbol{b}|$ (antenna spacing) or $k_\perp$ (transverse wavenumber).
+The $y$-axis denoted by $\tau$ (delay) or $k_\parallel$ (line of sight wavenumber).
 Here, $k_\perp$ and $k_\parallel$ are obtained for a frequency of 185~MHz.
@@
 Regions excluding the {\it wedge} are expected to be relatively free of foreground contamination and are generally referred to as the {\it EoR window}.
-There are undesirable grating responses (dot--dashed lines) specific to the MWA.
+There are undesirable grating responses (dotted-dashed lines) specific to the MWA.
 Hence, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the MWA {\it EoR window}.
@@
 
-The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\tau$--axis \citep{par12b,thy13}.
+The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\tau$-axis \citep{par12b,thy13}.
 The width of this narrow strip is inversely proportional to the operating bandwidth.
-The region of Fourier space excluding the {\it foreground wedge} and the narrow strips is the so--called {\it EoR window}, shown in light and medium shades.
-In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\boldsymbol{k}|$.
+The region of Fourier space excluding the {\it foreground wedge} and the narrow strips is the so-called {\it EoR window}, shown in light and medium shades.
+In the context of EoR studies in Fourier space, the H~{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\boldsymbol{k}|$.
 Hence, the brightest EoR signal will be observed on the shortest baselines and smallest delays.
@@
 In the specific case of the MWA, which has a passband constructed using coarse channels, there are period grating responses resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.
-Thus the MWA {\it EoR window} lies outside the dashed lines but inside the first grating response ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) and is shown in medium shade.
+Thus the MWA {\it EoR window} lies outside the dashed lines but inside the first grating response ($|\tau| \lesssim 0.78\,\mu$s, dotted-dashed lines) and is shown in medium shade.
 
@@
 
-We obtain the delay spectrum of visibilities by taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W(f)$ to be a {\it Blackman--Harris} window function.
+We obtain the delay spectrum of visibilities by taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W(f)$ to be a {\it Blackman-Harris} window function.
 The sky spectrum is multiplied in the instrument by the instrumental passband and flagging of frequency channels possibly corrupted by radio frequency interference, which together are represented by the weights, $W_\textrm{i}(f)$.
-In delay--space, these weights translate into a convolution by a point spread function (PSF).
+In delay-space, these weights translate into a convolution by a point spread function (PSF).
 We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis \citep{par09,par12b} to obtain the final delay spectra.
@@
 
-\section{Wide--Field Measurements}\label{sec:wide-field}
+\section{Wide-Field Measurements}\label{sec:wide-field}
 
@@
 \end{align}
-where, $l$, $m$, and $n$ denote the direction cosines toward East, North, and zenith respectively, with $n\equiv\sqrt{1-l^2-m^2}$, and:
+where, $l$, $m$, and $n$ denote the direction cosines toward east, north, and zenith respectively, with $n\equiv\sqrt{1-l^2-m^2}$, and:
 \begin{align}\label{eqn:solid-angle}
@@
 
-When the synthesized field is small, where $A(\hat{\boldsymbol{s}},f)$ or $I(\hat{\boldsymbol{s}},f)$ is significant only for $|l| \ll 1$ and $|m| \ll 1$, equation~\ref{eqn:obsvis-lmn} reduces to a simple two--dimensional Fourier transform \citep{tay99,tho01} between the apparent sky brightness and measured visibilities.
+When the synthesized field is small, where $A(\hat{\boldsymbol{s}},f)$ or $I(\hat{\boldsymbol{s}},f)$ is significant only for $|l| \ll 1$ and $|m| \ll 1$, equation~\ref{eqn:obsvis-lmn} reduces to a simple two-dimensional Fourier transform \citep{tay99,tho01} between the apparent sky brightness and measured visibilities.
 It is in this context that radio interferometers are understood to be sensitive only to fluctuations and not to a uniform sky brightness distribution.
 
-In a wide--field measurement, neither $A(\hat{\boldsymbol{s}},f)$ nor $I(\hat{\boldsymbol{s}},f)$, in general, is negligible anywhere in the visible hemisphere.
-The solid angle per pixel on the sky in direction cosine coordinates changes significantly with direction (equation~\ref{eqn:solid-angle}), increasing steeply towards the horizon.
-Hence, the approximations in the narrow--field scenario do not apply.
+In a wide-field measurement, neither $A(\hat{\boldsymbol{s}},f)$ nor $I(\hat{\boldsymbol{s}},f)$, in general, is negligible anywhere in the visible hemisphere.
+The solid angle per pixel on the sky in direction cosine coordinates changes significantly with direction (equation~\ref{eqn:solid-angle}), increasing steeply toward the horizon.
+Hence, the approximations in the narrow-field scenario do not apply.
 For example, even if $A(\hat{\boldsymbol{s}},f)$ and $I(\hat{\boldsymbol{s}},f)$ are held constant across the visible hemisphere, the amplitude of the integrand in equation~\ref{eqn:obsvis-lmn} is still dependent on direction.
-Therefore, in a significant departure from a narrow--field measurement, the wide--field visibility from a uniform brightness distribution on a non--zero antenna spacing is not zero.
+Therefore, in a significant departure from a narrow-field measurement, the wide-field visibility from a uniform brightness distribution on a non-zero antenna spacing is not zero.
 
-Figure~\ref{fig:usm} shows the wide--field delay power spectrum of a uniformly illuminated sky with no spectral variation as measured by antenna elements arranged identical to that in the MWA antenna array layout \citep{bea12} with a uniform power pattern across the sky and a bandwidth of 30.72~MHz centered around 185~MHz (refer to \S\ref{sec:instrument_model} for a detailed description of the instrument model).
-Notice the steep rise in power towards the horizon limits.
-These wide--field effects are prevalent on all antenna spacings, including the longest ones used in this study.
+Figure~\ref{fig:usm} shows the wide-field delay power spectrum of a uniformly illuminated sky with no spectral variation as measured by antenna elements arranged identical to that in the MWA antenna array layout \citep{bea12} with a uniform power pattern across the sky and a bandwidth of 30.72~MHz centered around 185~MHz (refer to \S\ref{sec:instrument_model} for a detailed description of the instrument model).
+Notice the steep rise in power toward the horizon limits.
+These wide-field effects are prevalent on all antenna spacings, including the longest ones used in this study.
 
@@
 \includegraphics[width=\linewidth]{figure2.eps}
-\caption{Wide--field effects on delay power spectra produced with a uniform sky brightness distribution measured by antenna pairs with a uniform power pattern across the visible hemisphere.
+\caption{Wide-field effects on delay power spectra produced with a uniform sky brightness distribution measured by antenna pairs with a uniform power pattern across the visible hemisphere.
 Delay power spectra are obtained using equations~\ref{eqn:obsvis}, \ref{eqn:delay-transform} and \ref{eqn:delay-power-spectrum} for each baseline, which are then stacked by baseline length.
 The axes correspond to cosmological dimensions.
-The non--zero response of the interferometer array to a uniform brightness distribution and the prominent {\it edge brightening} close to the horizon delay limits are wide--field effects.
-These are prevalent on all antenna spacings and are generic to all instruments used in wide--field measurements.
+The non-zero response of the interferometer array to a uniform brightness distribution and the prominent {\it edge brightening} close to the horizon delay limits are wide-field effects.
+These are prevalent on all antenna spacings and are generic to all instruments used in wide-field measurements.
 \label{fig:usm}}
@@
 
-We interpret this as due to equal--sized delay bins subtending larger solid angles near the horizon thereby containing larger integrated emission.
-Further, baseline vectors (including those with largest lengths) are foreshortened towards the horizon along their orientation.
+We interpret this as due to equal-sized delay bins subtending larger solid angles near the horizon thereby containing larger integrated emission.
+Further, baseline vectors (including those with largest lengths) are foreshortened toward the horizon along their orientation.
 Thus, they become sensitive to larger angular scales that match the inverse of their foreshortened lengths along these directions.
@@
 \citet{pob13} also find a similar ``edge brightening'' feature which they attribute to Galactic plane emission near the horizon.
-From their discussion, it is unclear what fraction of power in that feature arises from such wide--field effects.
+From their discussion, it is unclear what fraction of power in that feature arises from such wide-field effects.
 
-We conclude these are generic to all instruments making wide--field measurements.
+We conclude these are generic to all instruments making wide-field measurements.
 The nature of the specific instrument used for observing will control the amplitude of these effects, which we explore below.
@@
 
-In our present study, we use a latitude of -26.701\arcdeg~ and an antenna layout identical to that of the MWA \citep{bea12} for the observatory.
+In our present study, we use a latitude of -26\fdg 701 and an antenna layout identical to that of the MWA \citep{bea12} for the observatory.
 The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km.
@@
 Their deviation from coplanarity is negligible.
-For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67.5\arcdeg \leq \theta_b < 112.5\arcdeg$.
-Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
+For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from east) of all baselines to lie in the range $-67\fdg 5 \leq \theta_b < 112\fdg 5$.
+Baselines oriented in the other half-plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.
 We choose an observing frequency of 185~MHz ($z\simeq 6.68$) and a flat passband of width $\Delta B = 30.72$~MHz to roughly match those of ongoing MWA EoR observations (discussed in detail in \S\ref{sec:MWA}).
@@
 We consider the following antenna elements placed at the MWA tile locations:
-\begin{itemize}
-\item {\it Dipole}: An East--West dipole of length 0.74~m at a height 0.3~m above a ground plane. $A_\textrm{e}=(\lambda/2)^2$.
-\item {\it Phased Array}: A 4$\times$4 array of isotropic radiators with a grid spacing of 1.1~m at a height 0.3~m above the ground plane placed in an arrangement similar to that of an MWA tile.
+\begin{enumerate}
+\item {\it Dipole}: an east-west dipole of length 0.74~m at a height 0.3~m above a ground plane. $A_\textrm{e}=(\lambda/2)^2$.
+\item {\it Phased Array}: a 4$\times$4 array of isotropic radiators with a grid spacing of 1.1~m at a height 0.3~m above the ground plane placed in an arrangement similar to that of an MWA tile.
 $A_\textrm{e}=16\,(\lambda/2)^2$.
-\item {\it Dish}: Diameter of 14~m similar to that proposed for HERA, with $A_\textrm{e}\approx 154$~m$^2$.
+\item {\it Dish}: diameter of 14~m similar to that proposed for HERA, with $A_\textrm{e}\approx 154$~m$^2$.
 The power pattern is simulated using an {\it Airy} pattern where its sensitivity beyond the horizon is forced to zero.
-\end{itemize}
+\end{enumerate}
 
@@
 
-In wide--field measurements, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view, a point also supported by Pober et al.
+In wide-field measurements, it is important to consider an all-sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view, a point also supported by Pober et al.
 (2015, in preparation).
@@
 
-For the diffuse component, we use an all--sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz.
-At this frequency, since this map is predominantly based on the 408~MHz map of \citet{has82} which has an angular resolution of 0.85\arcdeg, we smoothed the 185~MHz map to the same resolution.
+For the diffuse component, we use an all-sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz.
+At this frequency, since this map is predominantly based on the 408~MHz map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz map to the same resolution.
 However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals.
@@
 Slightly different criteria are in common use in radio astronomy to estimate radio source confusion \citep[see Appendix of][and references therein]{thy13}.
-For an angular resolution of 0.85\arcdeg, using a conservative `$S_\textrm{c}=5\sigma_\textrm{c}$' criterion, we determine the flux density threshold to be $\approx 10$~Jy.
-Other liberal criteria that yield a lower threshold carry a greater risk of double--counting point sources which might be already blended in with the diffuse sky model.
+For an angular resolution of 0\fdg 85, using a conservative `$S_\textrm{c}=5\sigma_\textrm{c}$' criterion, we determine the flux density threshold to be $\approx 10$~Jy.
+Other liberal criteria that yield a lower threshold carry a greater risk of double-counting point sources which might be already blended in with the diffuse sky model.
 
@@
 These criteria yield 100 objects from the SUMSS catalog and 250 objects from the NVSS catalog.
-Together with the diffuse foreground model, we obtain an all--sky foreground model consisting of both compact and diffuse emission.
+Together with the diffuse foreground model, we obtain an all-sky foreground model consisting of both compact and diffuse emission.
 Figure~\ref{fig:sky-model-generic} shows the diffuse (top) and compact (bottom) foreground emission model used in our study.
-In this snapshot pointed towards zenith at 0.09~hours LST, the Galactic center in the diffuse model has just set in the West.
+In this snapshot pointed toward zenith at 0.09~hr LST, the Galactic center in the diffuse model has just set in the west.
 
@@
 \includegraphics[width=\linewidth]{figure3.eps}
-\caption{Foreground model at 185~MHz consisting of diffuse emission (top) in units of K and bright point sources (bottom) in units of Jy, visible during a snapshot at 0.09~hours LST.
-In the diffuse model, the Galactic center has just set in the West.
+\caption{Foreground model at 185~MHz consisting of diffuse emission (top) in units of K and bright point sources (bottom) in units of Jy, visible during a snapshot at 0.09~hr LST.
+In the diffuse model, the Galactic center has just set in the west.
 Color scales are logarithmic.
@@
 Foreground power close to the horizon delay limits in all three cases is significant even on long baselines.
-The foreshortening of baselines towards the horizon makes them sensitive to foreground emission on large size scales.
+The foreshortening of baselines toward the horizon makes them sensitive to foreground emission on large size scales.
 The amplitude of this feature strongly depends on the shape of the antenna element.
@@
 The severity of foreground contamination inside the {\it foreground wedge} both in strength and occupancy also evidently decreases as the antenna element is changed from a dipole to a phased array to a dish.
-For instance, notice that the foreground contamination in $k$--modes between $k_\parallel=0$ and the horizon limits decreases from $\sim 10^5$~K$^2\,$(Mpc/$h$)$^3$ in a phased array to $\sim 10$~K$^2\,$(Mpc/$h$)$^3$ in a dish.
-As a consequence, $k$--modes in the {\it foreground wedge} that may be deemed too contaminated for EoR studies in the case of a dipole or a phased array can potentially become accessible when using a dish.
+For instance, notice that the foreground contamination in $k$-modes between $k_\parallel=0$ and the horizon limits decreases from $\sim 10^5$~K$^2\,$(Mpc/$h$)$^3$ in a phased array to $\sim 10$~K$^2\,$(Mpc/$h$)$^3$ in a dish.
+As a consequence, $k$-modes in the {\it foreground wedge} that may be deemed too contaminated for EoR studies in the case of a dipole or a phased array can potentially become accessible when using a dish.
 
 Finally, a distinct feature common to all these aperture shapes is that the foreground contamination near the horizon delay edges is significant even on wide antenna spacings ($\gtrsim 10^5$~K$^2$(Mpc/$h$)$^3$).
-We have argued this arises due to wide--field effects.
-The prevalence of this feature across different antenna shapes demonstrates it is generic to all wide--field measurements.
+We have argued this arises due to wide-field effects.
+The prevalence of this feature across different antenna shapes demonstrates it is generic to all wide-field measurements.
 The amplitude of this effect, however, can be controlled via choice of antenna shape and through weighting of aperture illumination.
-A dish--shaped antenna appears to hold a significant advantage over a dipole or a phased array from the viewpoint of foreground contamination.
+A dish-shaped antenna appears to hold a significant advantage over a dipole or a phased array from the viewpoint of foreground contamination.
 
 Typically, the sensitivity of antennas to the primary field of view is high compared to the rest of regions on the sky.
-Combined with the wide--field effects seen earlier, it leads to a ``pitchfork''--shaped signature inside the {\it foreground wedge}, as exemplified in the case of a dish.
-Although the exact appearance of this signature depends on the antenna power pattern, we use the term {\it pitchfork} hereafter, to broadly refer to the combination of foreground power in the primary field of view and the enhancement of foreground power near the horizon limits due to the nature of wide--field measurements.
+Combined with the wide-field effects seen earlier, it leads to a ``pitchfork''-shaped signature inside the {\it foreground wedge}, as exemplified in the case of a dish.
+Although the exact appearance of this signature depends on the antenna power pattern, we use the term {\it pitchfork} hereafter, to broadly refer to the combination of foreground power in the primary field of view and the enhancement of foreground power near the horizon limits due to the nature of wide-field measurements.
 
@@
 MWA construction was completed in 2012 and, after commissioning, began its EoR observing program in 2013.
-The MWA is a 128--tile interferometer capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz.
-Each tile is a phased array of 16 dipoles, each in the shape of a bow--tie.
+The MWA is a 128-tile interferometer capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz.
+Each tile is a phased array of 16 dipoles, each in the shape of a bow-tie.
 This yields a primary field of view $\gtrsim$~20\arcdeg~ wide and multiple secondary lobes.
@@
 
-The MWA passband of width $\Delta B=30.72$~MHz is divided coarsely into 24$\times$1.28~MHz sub--bands with each sub--band weighted by a digital filter.
-The coarse channel shape is obtained using an 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$.
+The MWA passband of width $\Delta B=30.72$~MHz is divided coarsely into 24$\times$1.28~MHz sub-bands with each sub-band weighted by a digital filter.
+The coarse channel shape is obtained using an eight-tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$.
 Each of these coarse bands consists of 32 fine channels of width 40~kHz each.
-After correcting for the shape of these coarse channels, the fine channels at the edges of these sub--bands are flagged because they are known to be contaminated by aliasing at a low level.
+After correcting for the shape of these coarse channels, the fine channels at the edges of these sub-bands are flagged because they are known to be contaminated by aliasing at a low level.
 
-The MWA is expected to be sensitive to the power spectrum of the H{\sc i} signal over the redshift range $6<z<10$ \citep{bow06,bea13,thy13}.
-Over 600~hours have been currently observed using the MWA, targeting science objectives outlined in \citet{bow13}.
+The MWA is expected to be sensitive to the power spectrum of the H~{\sc i} signal over the redshift range $6<z<10$ \citep{bow06,bea13,thy13}.
+Over 600~hr have been currently observed using the MWA, targeting science objectives outlined in \citet{bow13}.
 
-The MWA targets two primary low--foreground fields for reionization observations.
-Here, we focus on the field at RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$.
+The MWA targets two primary low-foreground fields for reionization observations.
+Here, we focus on the field at R.A.~$=0^\textrm{h}$, decl.~$=-30\arcdeg$.
 The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of the dipoles in an MWA tile.
@@
 The sky drifts across to the nearest available pointing, shifting between grid points (once every $\sim 30$~minutes).
-This process is repeated throughout the course of the observation $\approx 4.86$~hours.
+This process is repeated throughout the course of the observation $\approx 4.86$~hr.
 
 The observations used here were acquired on 2013~August~23.
-We have chosen two sections of duration 112 seconds each from this night for detailed study.
+We have chosen two sections of duration 112~s each from this night for detailed study.
 These were chosen to provide a selection of possible foreground and instrumental conditions.
-As an example of a nominal observing setup we choose a zenith pointing; as an example of poor foreground conditions, we choose a pointing when the field is $\sim 2$~hours from zenith.
+As an example of a nominal observing setup we choose a zenith pointing; as an example of poor foreground conditions, we choose a pointing when the field is $\sim 2$~hr from zenith.
 This pointing has a significantly higher secondary lobe structure and is observed when the bright galactic center is well above the horizon.
 
-These two pointings are at LST 22.08~hours and 0.09~hours, which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings, respectively.
+These two pointings are at LST 22.08~hr and 0.09~hr, which are hereafter denoted as {\it off-zenith} and {\it zenith} pointings, respectively.
 
@@
 The data are flagged for interference \citep{off15}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz.
-These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey \citep[MWACS;][]{hur14}.
+These data are then calibrated to a simulation of the sky containing 2420 point-like objects selected from the MWA Commissioning Survey \citep[MWACS;][]{hur14}.
 It has a flux density limit of 25~mJy and a declination range of $-12\arcdeg$~ to $-40\arcdeg$~ evenly covering the field of view of the observations reported here.
 The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power pattern.
-The calibration algorithm -- based on forward modeling software by \citet{sul12} and the calibration method described by \citet{sal14} -- computes complex gain solutions per channel per antenna averaged to two minute intervals.
-The solutions are fairly low signal--to--noise so we iteratively average along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation.
-First, we average the channel gains over all antennas to obtain a high signal--to--noise measurement of the bandpass.
+The calibration algorithm --- based on forward modeling software by \citet{sul12} and the calibration method described by \citet{sal14} --- computes complex gain solutions per channel per antenna averaged to two minute intervals.
+The solutions are fairly low signal to noise so we iteratively average along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation.
+First, we average the channel gains over all antennas to obtain a high signal to noise measurement of the bandpass.
 After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna.
@@
 
-The MWA tile power pattern is modeled as a mutually--coupled 4--by--4 dipole array with the overall power pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14_arXiv}.
+The MWA tile power pattern is modeled as a mutually-coupled 4-by-4 dipole array with the overall power pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14_arXiv}.
 To speed up simulations, we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles.
 We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known repeatability and stability level of the analog signal chain \citep{bow07b}.
-Besides having the effect of adding a time--dependent uncertainty in the power pattern, these random delay fluctuations reduce the coherence in the phased addition of dipole signals resulting in deviations from predicted models of the power pattern, most prominently at its nulls.
+Besides having the effect of adding a time-dependent uncertainty in the power pattern, these random delay fluctuations reduce the coherence in the phased addition of dipole signals resulting in deviations from predicted models of the power pattern, most prominently at its nulls.
 
@@
 \subfloat[][Bright Point Source Model]{\label{fig:CSM}\includegraphics[width=0.45\linewidth]{figure5b.eps}}
-\caption{Sky brightness temperature of the diffuse foreground model (left) and flux densities of bright point sources (right) at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+\caption{Sky brightness temperature of the diffuse foreground model (left) and flux densities of bright point sources (right) at 185~MHz visible during {\it off-zenith} (top) and {\it zenith} (bottom) pointings.
 The color scales are logarithmic.
@@
 The contour levels shown are 0.00195, 0.00781, 0.0312, 0.125, and 0.5.
-The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off--zenith} pointing in the diffuse sky model and the MWA tile power gain is significant ($\gtrsim 12$\%) in that direction.
+The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off-zenith} pointing in the diffuse sky model and the MWA tile power gain is significant ($\gtrsim 12$\%) in that direction.
 In contrast, emission from the Galactic plane in {\it zenith} pointing is significantly less.}
@@
 
-We estimate thermal noise, $\Delta V$, in the observed visibilities, $V_b(f)$, using the {\it rms} of $\tilde{V}_b(\tau)$ obtained from data after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
+We estimate thermal noise, $\Delta V$, in the observed visibilities, $V_b(f)$, using the rms of $\tilde{V}_b(\tau)$ obtained from data after {\it delay-deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
 \begin{align}\label{eqn:Tsys}
@@
 With the aforementioned foreground model, and instrumental and observational parameters, we simulate visibilities using equation~\ref{eqn:obsvis}.
-Figure~\ref{fig:fhd-sim-comparison} shows the delay power spectra from {\it off--zenith} and {\it zenith} pointings obtained from MWA observations and modeling.
+Figure~\ref{fig:fhd-sim-comparison} shows the delay power spectra from {\it off-zenith} and {\it zenith} pointings obtained from MWA observations and modeling.
 Notice the qualitative agreement of amplitude and structure between the two.
-The Galactic center and the Galactic plane visible in the {\it off--zenith} pointing make it appear brighter in the {\it foreground wedge} as a branch with $\tau<0$.
+The Galactic center and the Galactic plane visible in the {\it off-zenith} pointing make it appear brighter in the {\it foreground wedge} as a branch with $\tau<0$.
 
@@
 \includegraphics[width=\linewidth]{figure6.eps}
-\caption{Delay power spectra from MWA data (left) and modeling (right) for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings.
+\caption{Delay power spectra from MWA data (left) and modeling (right) for the {\it off-zenith} (top) and {\it zenith} (bottom) pointings.
 The {\it foreground wedge} is bounded by white dotted lines.
@@
 These estimates are conservative.
-A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power pattern \citep{ber15}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide--field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations, but is beyond the scope of this paper.
+A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power pattern \citep{ber15}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide-field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations, but is beyond the scope of this paper.
 Hereafter, our focus is to explore in detail the foreground signatures embedded in the {\it foreground wedge} of the MWA instrument.
@@
 
-Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum} shows the delay power spectra obtained from the diffuse (left) and compact (right) foreground emission for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings without thermal noise component.
+Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum} shows the delay power spectra obtained from the diffuse (left) and compact (right) foreground emission for the {\it off-zenith} (top) and {\it zenith} (bottom) pointings without thermal noise component.
 Some of the notable signatures are discussed below.
@@
 \includegraphics[width=\linewidth]{figure7.eps}
-\caption{Simulated delay power spectra (in units of K$^2$~(Mpc/$h$)$^3$) for for the diffuse (left) and compact (right) foreground models in the {\it off--zenith} (top) and {\it zenith} (bottom) pointings without any thermal noise.
+\caption{Simulated delay power spectra (in units of K$^2$~(Mpc/$h$)$^3$) for for the diffuse (left) and compact (right) foreground models in the {\it off-zenith} (top) and {\it zenith} (bottom) pointings without any thermal noise.
 The axes and color scale are identical to those in Figure~\ref{fig:fhd-sim-comparison}.
 In the {\it off-zenith} pointing, emission from the Galactic center is the most prominent feature seen as a branch at $\tau<0$.
-In the {\it zenith} pointing, delay power spectrum from diffuse emission has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings due to wide--field effects.
+In the {\it zenith} pointing, delay power spectrum from diffuse emission has a {\it two-pronged fork}-shaped structure and is present even at wide antenna spacings due to wide-field effects.
 Compact emission is centrally concentrated.
@@
 
-The most prominent signature seen in the {\it off--zenith} pointing (top left panel, Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power pattern.
+The most prominent signature seen in the {\it off-zenith} pointing (top left panel, Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co-located with one of the bright secondary lobes of the power pattern.
 It appears as a bright branch near the negative delay horizon delay limit.
@@
 
-Diffuse emission outside the Galactic plane manifests in the primary field of view as a branch at $\tau>0$ and $\tau=0$ in the {\it off--zenith} and {\it zenith} pointings respectively.
+Diffuse emission outside the Galactic plane manifests in the primary field of view as a branch at $\tau>0$ and $\tau=0$ in the {\it off-zenith} and {\it zenith} pointings respectively.
 The former is seen at $\tau>0$ because the primary lobe of the power pattern is centered eastward of zenith, whereas in the latter it is centered at zenith.
 As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inversely proportional to the length of the baseline projected in the direction of the emission.
-Thus, in the {\it zenith} pointing, the horizontal line at $\tau=0$, fades away on antenna spacings $|\boldsymbol{b}| \gtrsim 125$~m because the diffuse sky model is devoid of spatial structures on scales $\lesssim$~0.75\arcdeg.
+Thus, in the {\it zenith} pointing, the horizontal line at $\tau=0$, fades away on antenna spacings $|\boldsymbol{b}| \gtrsim 125$~m because the diffuse sky model is devoid of spatial structures on scales $\lesssim$~0\fdg 75.
 
@@
 In both pointings the diffuse emission (left panels) is prominent near the horizon delay limits extending to the widest antenna spacings.
-This is a characteristic signature of the wide--field effects discussed in \S\ref{sec:wide-field}.
+This is a characteristic signature of the wide-field effects discussed in \S\ref{sec:wide-field}.
 It is evident at all LSTs in our simulations.
-Thus, diffuse emission from far off--axis directions manifests as an edge--heavy {\it two--pronged fork} across all baselines.
+Thus, diffuse emission from far off-axis directions manifests as an edge-heavy {\it two-pronged fork} across all baselines.
 It decreases in strength with increasing baseline length but is nevertheless present in all baseline orientations.
@@
 
-In contrast to the delay power spectra of diffuse emission, compact emission (right panels) manifests as a center--heavy structure in either pointing.
+In contrast to the delay power spectra of diffuse emission, compact emission (right panels) manifests as a center-heavy structure in either pointing.
 
 The amplitude response of an interferometer to a point source is, to first order, flat across baseline length.
-Since the primary field of view in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with $\tau>0$.
+Since the primary field of view in the {\it off-zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with $\tau>0$.
 In the {\it zenith} pointing, compact emission from the same patch of sky is seen as a bright horizontal arm at $\tau=0$ since the primary field of view is centered at zenith.
 
-Foreground emission at $\tau=0$ and $\tau<0$ in the {\it off--zenith} pointing is caused by point sources co--located with secondary lobes of the power pattern.
-On the other hand, point sources co--located with secondary lobes of power pattern in the {\it zenith} pointing are revealed as faint but distinct branches at positive and negative delays depending on the orientation of antenna spacing and direction of emission on the sky.
+Foreground emission at $\tau=0$ and $\tau<0$ in the {\it off-zenith} pointing is caused by point sources co-located with secondary lobes of the power pattern.
+On the other hand, point sources co-located with secondary lobes of power pattern in the {\it zenith} pointing are revealed as faint but distinct branches at positive and negative delays depending on the orientation of antenna spacing and direction of emission on the sky.
 
@@
 Vertical dotted lines mark the horizon delay limits.
-Compact emission dominates the central regions of the delay power spectra while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a characteristic {\it pitchfork}--shaped structure.
+Compact emission dominates the central regions of the delay power spectra while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a characteristic {\it pitchfork}-shaped structure.
 \label{fig:pitchfork-baselines}}
@@
 The peak at $\tau=0$ from diffuse emission is $\sim 10^3$ times fainter and decreases rapidly with increase in antenna spacing.
-This is the response expected from different antenna spacings towards compact and diffuse emission.
+This is the response expected from different antenna spacings toward compact and diffuse emission.
 
@@
 In fact, even on widely spaced antennas, diffuse emission in the delay power spectrum near the horizon delay limits exceeds that in the primary lobe by about three orders of magnitude.
-This feature is described in \S\ref{sec:diffuse-long-baselines} and attributed to wide--field measurement effects discussed in \S\ref{sec:wide-field}.
+This feature is described in \S\ref{sec:diffuse-long-baselines} and attributed to wide-field measurement effects discussed in \S\ref{sec:wide-field}.
 
-Simulations with the complete foreground model show the combination of center--heavy features dominated by compact emission in primary field of view, and edge--heavy features from both types of emission especially the diffuse component near the horizon.
+Simulations with the complete foreground model show the combination of center-heavy features dominated by compact emission in primary field of view, and edge-heavy features from both types of emission especially the diffuse component near the horizon.
 This results in a characteristic {\it pitchfork} structure imprinted in the {\it foreground} wedge and should be evident in observations.
@@
 
-\section{Baseline--Based Foreground Mitigation}\label{sec:fg-grading}
+\section{Baseline-Based Foreground Mitigation}\label{sec:fg-grading}
 
@@
 
-The Galactic center in the {\it off--zenith} pointing is one such example already available in our study.
+The Galactic center in the {\it off-zenith} pointing is one such example already available in our study.
 Figure~\ref{fig:sky-model} shows the sky model ({\it top:} compact component, {\it bottom:} diffuse component) in this pointing.
-The Galactic center is the most dominant source of foreground contamination from the diffuse sky model and is co--located with a bright secondary lobe of the power pattern near the western horizon.
-Figure~\ref{fig:delay-map} shows the sky mapped to delays registered by the baseline vectors, of length 100~m for instance, oriented towards North (top panel) and East (bottom panel).
-Figure~\ref{fig:baseline-breakup} shows the delay spectra on baselines oriented northward ($67.5\arcdeg\le \theta_\textrm{b} < 112.5\arcdeg$) at the top and eastward ($-22.5\arcdeg~\le \theta_\textrm{b} < 22.5\arcdeg$) at the bottom.
+The Galactic center is the most dominant source of foreground contamination from the diffuse sky model and is co-located with a bright secondary lobe of the power pattern near the western horizon.
+Figure~\ref{fig:delay-map} shows the sky mapped to delays registered by the baseline vectors, of length 100~m for instance, oriented toward north (top panel) and east (bottom panel).
+Figure~\ref{fig:baseline-breakup} shows the delay spectra on baselines oriented northward ($67\fdg 5\le \theta_\textrm{b} < 112\fdg 5$) at the top and eastward ($-22\fdg 5~\le \theta_\textrm{b} < 22\fdg 5$) at the bottom.
 The Galactic center manifests itself most distinctly near the negative horizon delay limit on short eastward baselines in the delay power spectrum (bottom panel of Figure~\ref{fig:baseline-breakup}).
@@
 \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figure9c.eps}}
-\caption{(a): Sky model showing compact ({\it top}) and diffuse ({\it bottom}) emission (adopted from Figure~\ref{fig:sky-model-MWA}).
-The Galactic center is very prominent in diffuse emission on the West co--located with a bright secondary lobe of the power pattern.
-(b): Sky hemisphere mapped to delays observed on antenna spacings oriented North ({\it top}) and East ({\it bottom}).
+\caption{(a) Sky model showing compact ({\it top}) and diffuse ({\it bottom}) emission (adopted from Figure~\ref{fig:sky-model-MWA}).
+The Galactic center is very prominent in diffuse emission on the west co-located with a bright secondary lobe of the power pattern.
+(b) Sky hemisphere mapped to delays observed on antenna spacings oriented north ({\it top}) and east ({\it bottom}).
 Delays vary linearly with antenna spacing length.
@@
 The bright Galactic center will appear at $\tau=0$ in northward antenna spacings and close to negative horizon delay limit on eastward antenna spacings.
-(c): Simulated delay power spectra on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}).
+(c) Simulated delay power spectra on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}).
 White lines denote horizon delay limits.
@@
 
-In our example, we discard eastward antenna spacings ($-22.5\arcdeg\le\theta_\textrm{b}< 22.5\arcdeg$) of lengths $|\boldsymbol{b}| < 30$~m.
+In our example, we discard eastward antenna spacings ($-22\fdg 5\le\theta_\textrm{b}< 22\fdg 5$) of lengths $|\boldsymbol{b}| < 30$~m.
 Foreground contamination was found to be insensitive to removal of wider antenna spacings as discussed below.
-Figure~\ref{fig:before-after} shows the delay spectra obtained with all antenna spacings (top panel) and after applying our screening technique (bottom panel) on the {\it off--zenith} observation.
+Figure~\ref{fig:before-after} shows the delay spectra obtained with all antenna spacings (top panel) and after applying our screening technique (bottom panel) on the {\it off-zenith} observation.
 Notice the significant reduction in foreground spillover into the {\it EoR window} via the removal of bright vertical stripes on short eastward antenna spacings.
@@
 \includegraphics[width=\linewidth]{figure10.eps}
-\caption{Simulated delay spectra power for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
-Discarded antenna spacings (black vertical stripes) have lengths $|\boldsymbol{b}|<30$~m (leftward of vertical dashed line) and orientations $|\theta_\textrm{b}|<$~22.5\arcdeg.
+\caption{Simulated delay spectra power for the {\it off-zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}).
+Discarded antenna spacings (black vertical stripes) have lengths $|\boldsymbol{b}|<30$~m (leftward of vertical dashed line) and orientations $|\theta_\textrm{b}|<$~22\fdg 5.
 The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} is lowered by a factor $\sim 100$ when short eastward antenna spacings are discarded.
@@
 Figure~\ref{fig:screening} shows how the typical foreground contamination\footnote{Foreground contamination is measured by standard deviation of noiseless $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$ from foregrounds in the MWA {\it EoR window}.} in the MWA {\it EoR window} depends on the orientations and lengths of discarded antenna spacings.
-We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., $-7.5\arcdeg~\le\theta_\textrm{b} < 7.5\arcdeg$ (solid circles), $-15\arcdeg\le\theta_\textrm{b}< 15\arcdeg$ (solid squares), and $-22.5\arcdeg\le\theta_\textrm{b}< 22.5\arcdeg$ (solid stars).
-Among antenna spacings that satisfy these criteria, we discard data from those whose lengths are shorter than $|\boldsymbol{b}|_\textrm{max}$ ($x$--axis) and show foreground contamination estimated in the {\it EoR window} from all remaining antenna spacings.
+We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., $-7\fdg 5~\le\theta_\textrm{b} < 7\fdg 5$ (solid circles), $-15\arcdeg\le\theta_\textrm{b}< 15\arcdeg$ (solid squares), and $-22\fdg 5\le\theta_\textrm{b}< 22\fdg 5$ (solid stars).
+Among antenna spacings that satisfy these criteria, we discard data from those whose lengths are shorter than $|\boldsymbol{b}|_\textrm{max}$ ($x$-axis) and show foreground contamination estimated in the {\it EoR window} from all remaining antenna spacings.
 
@@
 \includegraphics[width=\linewidth]{figure11.eps}
-\caption{Drop in foreground contamination in the MWA {\it EoR window}, and loss of data for the {\it off--zenith} pointing as a function of discarded baselines.
-Eastward baselines with varying degrees of directedness -- $|\theta_\textrm{b}|<$~7.5\arcdeg (solid circles), $|\theta_\textrm{b}|<$~15\arcdeg~ (solid squares), and $|\theta_\textrm{b}|<$~22.5\arcdeg (solid stars) -- and lengths $|\boldsymbol{b}| \le |\boldsymbol{b}|_\textrm{max}$ ($x$--axis) are discarded.
+\caption{Drop in foreground contamination in the MWA {\it EoR window}, and loss of data for the {\it off-zenith} pointing as a function of discarded baselines.
+Eastward baselines with varying degrees of directedness --- $|\theta_\textrm{b}|<$~7\fdg 5 (solid circles), $|\theta_\textrm{b}|<$~15\arcdeg~ (solid squares), and $|\theta_\textrm{b}|<$~22\fdg 5 (solid stars) --- and lengths $|\boldsymbol{b}| \le |\boldsymbol{b}|_\textrm{max}$ ($x$-axis) are discarded.
 Loss of data (dotted lines) is measured by discarded baselines as a fraction of the total number for the corresponding cases.
-Foreground contamination in the {\it EoR window} (solid lines) drops by a factor $\sim$2 ($|\theta_\textrm{b}|\le$~7.5\arcdeg) to $\sim$100 ($|\theta_\textrm{b}|\le$~22.5\arcdeg).
+Foreground contamination in the {\it EoR window} (solid lines) drops by a factor $\sim$2 ($|\theta_\textrm{b}|\le$~7\fdg 5) to $\sim$100 ($|\theta_\textrm{b}|\le$~22\fdg 5).
 The latter limit can be achieved with a mere 5\% loss of data at $|\boldsymbol{b}|_\textrm{max}\simeq30$~m, and discarding longer baselines ($|\boldsymbol{b}|\gtrsim 30$~m) has no effect in further reducing foreground contamination.
@@
 The fraction of discarded antenna spacings discarded relative to the total number is shown in dotted lines for different ranges of $\theta_\textrm{b}$.
-It is seen that foreground contamination can be mitigated by a factor between $\sim 2$ ($|\theta_\textrm{b}|\le$~7.5\arcdeg) and $\sim 100$ ($|\theta_\textrm{b}|\le$~22.5\arcdeg).
+It is seen that foreground contamination can be mitigated by a factor between $\sim 2$ ($|\theta_\textrm{b}|\le$~7\fdg 5) and $\sim 100$ ($|\theta_\textrm{b}|\le$~22\fdg 5).
 The latter limit is achieved with a mere 5\% loss of data for $|\boldsymbol{b}|_\textrm{max}\simeq30$~m.
@@
 
-In principle, instead of discarding selected antenna spacings altogether, we could down--weight them based on an optimal scheme.
-For instance, the estimates of covariance computed from the delay transform bins can be naturally fed into the covariance--weighted power spectrum estimation techniques \citep{liu14a,liu14b}.
+In principle, instead of discarding selected antenna spacings altogether, we could down-weight them based on an optimal scheme.
+For instance, the estimates of covariance computed from the delay transform bins can be naturally fed into the covariance-weighted power spectrum estimation techniques \citep{liu14a,liu14b}.
 It could also be used to downweight or flag contaminated baselines in imaging applications.
@@
 
-Our primary motivation in this work is to understand how the various bright foregrounds will manifest in three dimensional power spectrum of H{\sc i} from 21~cm reionization observations.
+Our primary motivation in this work is to understand how the various bright foregrounds will manifest in three-dimensional power spectrum of H~{\sc i} from 21~cm reionization observations.
 In units of temperature variance, the dynamic range between bright foregrounds and the 21~cm signal is expected to be $\sim 10^8$; a detailed understanding of how foregrounds can corrupt the 21~cm power spectrum is therefore essential.
@@
 
-We find that all wide--field instruments, typical of modern EoR observatories, imprint a characteristic {\it two--pronged fork} signature in delay spectra.
-This arises from two related effects: delay bins near the horizon subtend larger solid angles and therefore contain larger integrated emission; and, foreshortening of baselines towards the horizon makes them sensitive to emission on large angular scales which match the inverse of their foreshortened lengths.
+We find that all wide-field instruments, typical of modern EoR observatories, imprint a characteristic {\it two-pronged fork} signature in delay spectra.
+This arises from two related effects: delay bins near the horizon subtend larger solid angles and therefore contain larger integrated emission; and, foreshortening of baselines toward the horizon makes them sensitive to emission on large angular scales which match the inverse of their foreshortened lengths.
 These effects combined with higher sensitivity of antennas in the primary field of view results in a characteristic {\it pitchfork} signature.
@@
 
-Simulating in many important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission and bright point sources, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA to within expected uncertainties in foreground models.
+Simulating in many important respects the response of the MWA to an all-sky foreground model that consists of diffuse Galactic emission and bright point sources, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA to within expected uncertainties in foreground models.
 
@@
 
-The bright Galactic center at the edge of the western horizon co--located with one of the far secondary lobes of MWA tile power pattern is the brightest source of foreground contamination in the {\it off--zenith} pointing.
+The bright Galactic center at the edge of the western horizon co-located with one of the far secondary lobes of MWA tile power pattern is the brightest source of foreground contamination in the {\it off-zenith} pointing.
 It manifests itself near the negative horizon delay limit in the delay power spectrum on eastward antenna spacings.
@@
 Diffuse emission in the primary field of view is prominent on shorter antenna spacings.
-However, it is also prominent near the horizon limits even on wide antenna spacings --- an effect of the wide--field nature of the measurement.
+However, it is also prominent near the horizon limits even on wide antenna spacings --- an effect of the wide-field nature of the measurement.
 On the other hand, compact emission predominantly maps onto central regions of the {\it foreground wedge}.
-Features arising from compact emission co--located with primary and secondary lobes of the antenna power pattern have been identified.
-In general, delay power spectrum signatures of compact emission are center--heavy while those of diffuse emission are edge--heavy which results in a characteristic {\it pitchfork} signature.
+Features arising from compact emission co-located with primary and secondary lobes of the antenna power pattern have been identified.
+In general, delay power spectrum signatures of compact emission are center-heavy while those of diffuse emission are edge-heavy which results in a characteristic {\it pitchfork} signature.
 This will be evident when thermal noise is sufficiently lowered, as larger volumes of data are processed.
 
-We also provide a simple and effective tool based on the delay spectrum technique that can potentially mitigate foreground contamination by nearly two orders of magnitude in EoR data analysis by discarding or down--weighting data from antenna pairs most affected by foreground contamination, with negligible loss of sensitivity.
+We also provide a simple and effective tool based on the delay spectrum technique that can potentially mitigate foreground contamination by nearly two orders of magnitude in EoR data analysis by discarding or down-weighting data from antenna pairs most affected by foreground contamination, with negligible loss of sensitivity.
 In conclusion, we find that inclusion of emission models, both diffuse and compact, all the way to the horizon is essential to explaining the observed power spectrum.
@@
 S.
-National Science Foundation (NSF) through award AST--1109257.
-DCJ is supported by an NSF Astronomy and Astrophysics Postdoctoral Fellowship under award AST--1401708.
+National Science Foundation (NSF) through award AST-1109257.
+DCJ is supported by an NSF Astronomy and Astrophysics Postdoctoral Fellowship under award AST-1401708.
 JCP is supported by an NSF Astronomy and Astrophysics Fellowship under award AST-1302774.
@@
 Figure~\ref{fig:delay-cartoon} illustrates the radio interferometer delays and conventions used in our paper.
-$\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local East, North (along local meridian) and upward (zenith) directions at the telescope site.
+$\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward (zenith) directions at the telescope site.
 Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at $\tau>0$ and $\tau<0$, respectively.
@@
 Two antennas labeled as A$_1$ and A$_2$ are separated by vector $\boldsymbol{b}$ on the local tangent plane.
-The local meridian, local North and local East are shown for reference.
+The local meridian, local north and local east are shown for reference.
 Points labeled as `z', `n', `s', `e' and `w' on the celestial sphere denote zenith, northward, southward, eastward and westward positions, respectively.
@@
 
-\section{Cosmological H{\sc i} Power Spectrum}\label{sec:EoR-power-spectrum}
+\section{Cosmological H~{\sc i} Power Spectrum}\label{sec:EoR-power-spectrum}
 
@@
 $w$ is aligned parallel to the direction of the phase center, while $u$ and $v$ lie on the transverse plane perpendicular to it.
-For measurements that lie on this plane, we can choose $w=0$ without loss of generality and $\boldsymbol{u}$ effectively reduces to $\boldsymbol{u}\equiv (u,v)$, a two--dimensional vector.
+For measurements that lie on this plane, we can choose $w=0$ without loss of generality and $\boldsymbol{u}$ effectively reduces to $\boldsymbol{u}\equiv (u,v)$, a two-dimensional vector.
 Then, $\boldsymbol{u}$ is directly related to the transverse spatial wavenumber mode as:
@@
 
-Since we are concerned with a redshifted H{\sc i} spectral line from cosmological distances, $f$ is a measure of cosmological distance along the line of sight.
+Since we are concerned with a redshifted H~{\sc i} spectral line from cosmological distances, $f$ is a measure of cosmological distance along the line of sight.
 $\eta$, which is the Fourier transform dual of $f$, is used to denote the spatial frequency along the line of sight and has units of time.
-It is directly related to the line--of--sight wavenumber,
+It is directly related to the line of sight wavenumber,
 \begin{align}\label{eqn:k-prll}
@@
 \end{align}
-where, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
+where, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H~{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.
 This approximation holds under the assumption that the redshift range (or frequency band) is small enough within which cosmological evolution is negligible.
@@
 \end{align}
-represents the true spatial Fourier representation of the three--dimensional sky brightness distribution.
+represents the true spatial Fourier representation of the three-dimensional sky brightness distribution.
 This approach has been discussed in detail in \citet{mor04}.
-The spatial power spectrum of EoR H{\sc i} distribution, $P(\boldsymbol{k}_\perp,k_\parallel)$, and $\tilde{V}_u(\eta)$ are related by \citep{mor04,mcq06,par12a}:
+The spatial power spectrum of EoR H~{\sc i} distribution, $P(\boldsymbol{k}_\perp,k_\parallel)$, and $\tilde{V}_u(\eta)$ are related by \citep{mor04,mcq06,par12a}:
 \begin{align}\label{eqn:true-power_spectrum}
@@
 where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, $\lambda$ is the wavelength of the band center, and $k_\textrm{B}$ is the Boltzmann constant.
-Thus, $\tilde{V}_u(\eta)$ inferred from observations, in units of Jy~Hz, can be converted into an equivalent cosmological H{\sc i} power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$, in units of K$^2$~Mpc$^3$ or, more generally, K$^2$(Mpc/$h$)$^3$, where $h$ is the Hubble constant factor.
+Thus, $\tilde{V}_u(\eta)$ inferred from observations, in units of Jy~Hz, can be converted into an equivalent cosmological H~{\sc i} power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$, in units of K$^2$~Mpc$^3$ or, more generally, K$^2$(Mpc/$h$)$^3$, where $h$ is the Hubble constant factor.
 
